### 2024-02-02 thesis/abstract.tex
added thesis, (the first commit is uploading the folder of the master thesis preparation latex)

\begin{abstract}
    The theory of kernel methods will be applied to the problem of probabilistic forecasting taking as input some data.
    Conceptually, the proposed analysis could be applied to any kind of data. In this study we considered the electricity market, because of the interesting implications  on risk management tasks.

  %This example thesis briefly shows the main features of our thesis
  %style, and how to use it for your purposes.
\end{abstract}

### 2024-02-02 thesis/appendix.tex
added thesis, (the first commit is uploading the folder of the master thesis preparation latex)

\chapter{Appendix}
\section{Feature Map Normalization}\label{appendix:new_feature}
\begin{proof}
\begin{align*}
    \|
    \phi^{new}(x)\|_{\mathcal{H}}^{2} &= \|\frac{\phi(x)}{\|\phi(x)\|_{\mathcal{H}}}
    \|_{\mathcal{H}}^{2}
    \\
    &=
    \|
    \frac{\phi(x)}
    {\sqrt{k(x,x)}}
    \|_{\mathcal{H}}^{2}
    \\
    &=
    \langle
    \frac{\phi(x)}
    {\sqrt{k(x,x)}}
    ,
    \frac{\phi(x)}
    {\sqrt{k(x,x)}}
    \rangle_{\mathcal{H}}
    \\
    &=
    \frac{1}{\sqrt{k(x,x)^{2}}}
    \langle
    \phi(x)
    ,
    \phi(x)
    \rangle_{\mathcal{H}}
    \\
    &=1
\end{align*}
\end{proof}

### 2024-02-02 thesis/input_files/literaturereview.tex
added thesis, (the first commit is uploading the folder of the master thesis preparation latex)

%INTRO KERNEL METHODS E ENERGY PRICE MODELLING

Kernel methods are a class of algorithms for patter analysis.
With kernel methods we are able to apply linear methods with predictors in a high dimensional space, without having to explicitly evaluate the involved dot products of the features.
In this thesis work I will address the performance of kernel methods in the context of probabilistic forecasting; the area of application will be the electricity market. 
Probabilistic forecasting may be useful to power producers, traders and consumers in order to improve their decision making process and managing risk(VaR). This is because probabilistic forecast enables them to simulate scenarios and carry out stress tests.

%SPIEGO UN PO' I PAPER CHE HO FATTO PASSARE VELOCEMENTE 
This problem has been already addressed in \cite{probablistic_electricity_forecast} and \cite{probablistic_electricity_forecast2}. 
These papers surveyed the performance of neural network architectures against simpler approaches like quantile regression and data fitting to Johnson distribution. Their conclusion is that distributional NN perform a little worse than quantile regression but the former has smaller computational cost; that is because the quantile regression is run for every quantile from 0.01 to 0.99.
Nevertheless kernel methods received very little attention in this specific setting.
\\
Kernel methods considered are kernel mean embedding \cite{pmlr}, \cite{Muandet_2017} and kernel herding \cite{supersamples}. Particularly extending the idea of \cite{2022nystrom}, where the Nyström approximation is employed in computing the kernel mean embedding, experiments with the Pivoted Cholesky decomposition will be performed.
\\
In section 2, three papers that lay the basis for this thesis’ work are summarized.
\\

### 2024-02-02 thesis/input_files/paper1.tex
added thesis, (the first commit is uploading the folder of the master thesis preparation latex)

% Some commands used in this file

\section{\href{https://arxiv.org/abs/1605.09522}{Kernel Mean Embedding of Distributions: A Review and Beyond}}
From this first paper \cite{Muandet_2017}, the notation and terms used in the theory of Reproducing Kernel Hilbert Spaces are summarized.
\\
Many algorithms use the inner product as similarity measure between data instances $x, x' \in \mathcal{X}$. However, this inner product spans only the class of linear similarity measures. 
\\
The idea behind kernel methods is to apply a non-linear transformation $\phi$ to the data $x$ 
in order to get a more powerful non linear similarity measure.
\begin{align*}
\phi(x):\mathcal{X} &\longrightarrow \mathcal{F}
    \\
    x&\mapsto \phi(x)
\end{align*}

Then we take the inner product in the high dimensional space $\mathcal{F}$ mapped by $\phi(x)$.
\\
\begin{align*}
k(x,x'):=&\langle \phi(x), \phi(x') \rangle_{\mathcal{F}}
\end{align*}
\\
$\phi(x)$ is referred as feature map while $k$ as kernel function.
\\
Therefore, we can kernelize any algorithm involving a dot product by substituting $\langle x, x' \rangle_{\mathcal{X}}$ with $\langle \phi(x), \phi(x') \rangle_{\mathcal{F}}$
\\
One would expect constructing the feature maps explicitly and then evaluate their inner product in $\mathcal{F}$ to be computationally expensive, and indeed it is. However, we do not have to explicitly perform such calculations. This is because of the existence of the kernel trick.
To illustrate the idea behind the kernel trick consider the following example. 
\\
Suppose $x \in \mathbb{R}^2$ and assume to select $\phi(x)=(x_{1}^{2},x_{2}^{2},\sqrt{2}x_{1}x_{2})$, then the inner product in the feature space is $x_{1}^{2}x_{1}^{'2},x_{2}^{2}x_{2}^{'2}+2x_{1}x_{2}x'_{1}x'_{2}$.
Notice that this is the same of $\langle \phi(x), \phi(x') \rangle$; thus the kernel trick consists of just using $k(x,x')=:(x^Tx')^2$.

\subsection{RKHS}
Following are the definitions that make up the basis for the theory of kernel methods.

\begin{definition}
    A sequence $\{v_n\}_{n=1}^{\infty}$ of elements of a normed space $\mathcal{V}$ is a Cauchy sequence if for every $\epsilon>0$, there exist $N=N(\epsilon) \in \mathbf{N}$ such that $\|v_n-v_m\|_{\mathcal{V}}<\epsilon \ \ \forall m,n\geq N$  
\end{definition}

\begin{definition}
    A complete metric space is a metric space in which every Cauchy sequence is convergent.
\end{definition}

\begin{definition}
    A Hilbert space is a vector space $\mathcal{H}$ with an inner product $\langle f, g \rangle$ such that the norm defined by $\|f\|=\sqrt{\langle f, f \rangle}$
turns $\mathcal{H}$ into a complete metric space.
\end{definition}

\begin{definition}
    RKHS.
    A Reproducing Kernel Hilbert Space is an Hilbert space with the evaluation functionals $\mathcal{F}_{x}(f):=f(x)$ bounded, i.e. $\forall x \in \mathcal{X}$ there exists some $C>0$ such that $\| \mathcal{F}_{x}(f)\|=\|f(x)\| \leq C \|f\|_{\mathcal{H}} \ \forall f \in \mathcal{H}$
\end{definition}

\begin{theorem}
    Riesz Representation. If $A : \mathcal{H} \rightarrow \mathbf{R}$ is a bounded linear operator in a Hilbert space $\mathcal{H}$ , there exists some $g_{A} \in \mathcal{H}$ such that $A(f) = \langle f,g_A\rangle_\mathcal{H}, \ \forall f \in \mathcal{H}$.
\end{theorem}

The Riesz representation theorem results in the following proposition for RKHS.
\begin{proposition}
For each $x \in \mathcal{X}$ there exists a function $k_{x} \in \mathcal{H}$ such that $\mathcal{F}_{x}(f)=\langle k_{x}, f\rangle_{\mathcal{H}}=f(x)$    
\end{proposition}

The function $k_{x}$ is the reproducing kernel for the point $x$.
Furthermore, note that $k_{x}$ is itself a function lying on $\mathcal{X}$

\begin{align*}
    k_{x}(y)=\mathcal{F}_{y}(k_{x})=\langle k_{x}, k_{y}\rangle_{\mathcal{H}=\langle \phi(x), \phi(x)\rangle_{\mathcal{H}}}
\end{align*}

### 2024-02-02 thesis/input_files/paper2.tex
added thesis, (the first commit is uploading the folder of the master thesis preparation latex)

\section{\href{http://proceedings.mlr.press/v33/kanagawa14.pdf}{Recovering Distributions from Gaussian RKHS Embeddings}}
This paper covers the RKHS embedding approach to nonparametric statistical inference \cite{pmlr}.
The idea here is computing an estimate of the kernel mean in order to obtain an approximation of the underlying distribution of the observed random variable.
The kernel mean embedding $\mu_{\mathbb{P}}$ of a probability $\mathbb{P}$ corresponds to the feauture map $\phi(x)$ integrated with respect to the $\mathbb{P}$ measure.
\\
That is $\mu_{\mathbb{P}}:=\int_{\mathcal{X}}k(x,\cdot)d\mathbb{P}(x)$.
\\
Kernel mean embedding serves as a unique representation of $\mathbb{P}$ in the RKHS $\mathcal{H}$. This holds provided  that $\mathcal{H}$ is characteristic. 
\begin{definition}
The RKHS $\mathcal{H}$ and the associated kernel k are said characteristic, when the mapping $\mu : \mathbb{P} \rightarrow \mathcal{H}$ is injective.
\end{definition}
%Hence, H is characteristic, if for any P,Q ∈ P, we have μP = μQ if and only if P = Q. 
When the mapping is injective, we have that $\mu_{\mathbb{P}}$ is uniquely associated with $\mathbb{P}$; thus, $\mu_{\mathbb{P}}$ is a unique representation of $\mathbb{P}$ in $\mathcal{H}$. 
%Characteristic kernels on X = Rd include {Gaussian, Matérn and Laplace}(Sriperumbudur et al., 2010).
%On the other hand, linear and polynomial kernels are not characteristic.
\\
Note that, by the reproducing property of RKHS $\langle f, k(x,\cdot) \rangle=f(x)$ we have:
\\
\begin{center}
    $E_{\mathbb{P}}[f(x)]=\langle f,\mu_{\mathbb{P}} \rangle_{\mathcal{H}}, \ \forall f \in \mathcal{H}$
\end{center}

\begin{proof}   
\begin{align*}
E_{\mathbb{P}}[f(x)]&=\int_{\mathcal{X}} f(x) d\mathbb{P}(x)\\
&=\int_{\mathcal{X}} \langle f, k(x,\cdot) \rangle_{\mathcal{H}} d\mathbb{P}(x)\\
&= \sum_{i=1}^{\infty} \langle f, k(x_{i}, \cdot) \rangle_{\mathcal{H}} \mathbb{P}(\mathcal{X}_{i}) \\
&=\langle f, \sum_{i=1}^{\infty} k(x_{i}, \cdot)\mathbb{P}(\mathcal{X}_{i}) \rangle_{\mathcal{H}}\\
&=\langle f, \mu_{\mathbb{P}} \rangle_{\mathcal{H}}
\end{align*}
\end{proof}

The kernel mean embedding can be employed to estimate the density $p$ at any fixed point $x_{0}$. Letting $\delta_{x_{0}}$ to be the dirac delta function we have:
\\
\begin{center}
$p(x_{0})=\int \delta_{x_{0}}p(x)dx=E_{\mathbb{P}}[\delta_{x_{0}}]$    
\end{center}

Therefore, the idea is to define an estimator for the expectation of $\delta_{x_{0}}$ through $\mu_{\mathbb{P}}$; this would result in an estimator of $p(x_{0})$.
\\
A kernel $k(x_{0},\cdot)$ is used to approximate the delta function, furthermore applying theorem 1 of \cite{pmlr} we have that a consistent estimator of $E_{\mathbb{P}}[k(x_{0}, \cdot)]$ is given by $\sum\limits_{i=1}^{n}w_{i}k(x_{0},x_{i})$
\\
When the weigths are all $1/n$ we end up with the standard kernel density estimation.
\\
Alternatively,  the optimal weights can be found by minimizing the following problem $ \| \hat{\mu}-\Phi w\|^2$ 
where $\Phi:\mathbb{R}^{n} \rightarrow \mathcal{H}$.
%, \alpha \rightarrow \sum\limits_{j=1}^{n}\alpha_{j} \phi(x_{j})$
\\

### 2024-02-02 thesis/input_files/paper3.tex
added thesis, (the first commit is uploading the folder of the master thesis preparation latex)

\section{\href{https://arxiv.org/pdf/1203.3472.pdf}{Super-Samples from Kernel Herding}}

Kernel herding is a deterministic sampling algorithm designed to draw "Super Samples" from probability distributions \cite{supersamples}. The idea of herding, is to generate pseudo-samples that greedily minimize the error betweeen the mean operator and the empirical mean operator resulting from the selected herding points.
% note the kernel mean embedding lies in a.
%\\

%Let $x \in \mathcal{X}$, where a general $x$ is a state over the index set $\mathcal{X}$. Note, $\mathcal{X}$ is typically the space of covariates.

%\\
%Let $\phi:\mathcal{X}\rightarrow \mathcal{H}$ denote a feature map into a Hilbert space $\mathcal{H}$ endowed with inner product $\langle \cdot, \cdot \rangle_{\mathcal{H}}$
Letting $p(x)$ be a probability distribution, kernel herding is recursively defined as follows:
\begin{align*}
x_{t+1} &=\argmax{x\forall \mathcal{X}}\langle w_{t},\phi(x)\rangle 
\\
w_{t+1} &= w_{t} + E_{\mathbb{P}}[\phi(x)]-\phi(x_{t+1})    
\end{align*}
\\
$w$ denotes a weight vector that lies in $\mathcal{H}$.
\\
Here, by assuming that the inner product between weights and the mean operator is equal to a general functional f evaluated at x, that is $\langle w, \phi(x) \rangle_{\mathcal{H}}=f(x)$.
We have:
\begin{align*}
\langle w, \mu_{\mathbb{P}} \rangle_{\mathcal{H}}
&= \langle w, \int k(x,\cdot)d\mathbb{P}(x) \rangle_{\mathcal{H}}
\\
&= \langle w, \sum\limits_{i=1}^{\infty} k(x_{i},\cdot)\mathbb{P}(\mathcal{X}_{i})\rangle_{\mathcal{H}}
\\
&= \sum \limits_{i=1}^{\infty} \langle w, k(x_{i}, \cdot) \rangle_{\mathcal{H}}\mathbb{P}(\mathcal{X}_{i})
\\
&= \sum \limits_{i=1}^{\infty}f(x_{i})\mathbb{P}(\mathcal{X}_{i})
\\
&=\int f(x) d\mathbb{P}(x)
\\
&= \mathbb{E}_{\mathbb{P}} [f(x)]
\end{align*}
\\
Moreover, second assumption of the model is that $\|\phi(x)\|_{\mathcal{H}}=R \quad \forall x \in X$.
That is the Hilbert space norm of the feature vector is equal to a constant R for all states in the set $\mathcal{X}$.
\\
\\
This can be achieved by taking the new feature vector as $\phi^{new} (x)=\frac{\phi(x)}{\|\phi(x)\|_{\mathcal{H}}}$.
See \ref{appendix:new_feature} for details.
\\
By rewriting the formula for the weights we end up with
\begin{align*}
w_{t+1} &= w_{t} + E_{\mathbb{P}}[\phi(x)]-\phi(x_{t+1})
\\
&= w_{t-1} + E_{\mathbb{P}}[\phi(x)]+E_{\mathbb{P}}[\phi(x)]-\phi(x_{t+1})-\phi(x_{t})
\\
&= w_{t-2} + E_{\mathbb{P}}[\phi(x)]+E_{\mathbb{P}}[\phi(x)]+E_{\mathbb{P}}[\phi(x)]-\phi(x_{t+1})-\phi(x_{t})-\phi(x_{t-1})
\\
&\text{Considering $w_{T}$, we have}
\\
&
w_{T}=w_{0}+TE_{\mathbb{P}}[\phi(x)]-\sum\limits_{t=1}^{T}\phi(x_{t})
\end{align*}
Note that
$E_{\mathbb{P}}[\phi(x)]=\int_{\mathcal{X}}\phi(x)d\mathbb{P}(x)$ which corresponds to the definition of $\mu_\mathbb{P}$. 
That is $\mu$ is the mean operator associated with the distribution $\mathbb{P}$; it lies in $\mathcal{H}$.
\\
Thus,
\begin{align*}
   w_{T}= w_{0}+T\mu_{\mathbb{P}}-\sum\limits_{t=1}^{T}\phi(x_{t})
\end{align*}
\\
Notice we do not have to compute $\mu_{\mathbb{P}}$  explicitly, the terms involving $\mu_{\mathbb  {P}}$ will be computed by applying the kernel trick.
\\
% In addition we can construct an estimate $\hat{\mu}=:\frac{1}{N}\sum \limits_{i=1}^{N} k(x_{i},\cdot)$.
\\
Now we have everything we need in order to reformulate the original problem in a way such that it depends just on the states x.
Plug the formula for the weights in the formula for the $x_{t}$ and use the kernel trick; we end up with
\\
\begin{align*}
    x_{T+1} &=\argmax{x\forall \mathcal{X}}\langle w_{0}+T\mu_{\mathbb{P}}-\sum\limits_{t=1}^{T}\phi(x_{t}),\phi(x)\rangle_{\mathcal{H}} \\
&=\argmax{x\forall \mathcal{X}}\langle w_{0},\phi(x)\rangle_{\mathcal{H}}
+\langle T\mu_{\mathbb{P}},\phi(x)\rangle_{\mathcal{H}}
-\langle \sum\limits_{t=1}^{T}\phi(x_{t}),\phi(x)\rangle_{\mathcal{H}}
\\
&=\argmax{x\forall \mathcal{X}}\langle w_{0},\phi(x)\rangle_{\mathcal{H}}
+T\langle \mu_{\mathbb{P}},\phi(x)\rangle_{\mathcal{H}}
- \sum\limits_{t=1}^{T}k(x_{t}, x)
\end{align*}
\\
Notice $\langle \mu_{\mathbb{P}},\phi(x)\rangle_{\mathcal{H}}$ can be rewritten in the following way
\\
\begin{align*}
\langle \mu_{\mathbb{P}},\phi(x)\rangle_{\mathcal{H}}&=
\langle \int_{\mathcal{X'}}\phi(x')d\mathbb{P}(x'),\phi(x)\rangle_{\mathcal{H}}
\\
&=
\langle  \sum\limits_{i=1}^{\infty}\phi(x'_{i})\mathbb{P}(\mathcal{X'}_{i}),\phi(x)\rangle_{\mathcal{H}}
\\
&=
\sum\limits_{i=1}^{\infty}\langle \phi(x'_{i}),\phi(x)\rangle_{\mathcal{H}} \mathbb{P}(\mathcal{X'}_{i})
\\
&=
\sum\limits_{i=1}^{\infty}k(x'_{i},x) \mathbb{P}(\mathcal{X'}_{i})
\\
&=
\int_{\mathcal{X'}}k(x',x)d\mathbb{P}(x')
\\
&=
E_{\mathbb{P}}[k(x',x)]
\end{align*}
\\
% Note the introduction of $x'$ is the same as as $x$ it is just notation.
\\
Furthermore, by initializing $w_{0}=\mu_{\mathbb{P}}$ we end up with the following function to be optimized, i.e.
\\
\begin{align*}
x_{T+1}&=\argmax{x\forall \mathcal{X}}\langle w_{0},\phi(x)\rangle +T\langle \mu_{\mathbb{P}},\phi(x)\rangle- \sum\limits_{t=1}^{T}k(x_{t}, x)    
\\
&=\argmax{x\forall \mathcal{X}} (T+1)E_{\mathbb{P}}[k(x',x)]- \sum\limits_{t=1}^{T}k(x_{t}, x)    
\end{align*}
\\ 
Now consider the error term between the mean kernel operator and its estimation through herding samples
\\
\begin{align*}
\varepsilon_{T+1}&=\| \mu_{\mathbb{P}} -\frac{1}{T+1}\sum\limits_{i=1}^{T+1} \phi(x_{t})\|_{\mathcal{H}}^{2}
\\
&=\mathbb{E}_{x,x' \sim \mathbb{P}} [k(x',x)] - \frac{2}{T+1}\sum \limits_{t=1}^{T+1} \mathbb{E}_{x \sim \mathbb{P}} [k(x,x_{t})] +\frac{1}{(T+1)^2}\sum \limits_{t,t'=1}^{T+1} k(x_{t}, x_{t'})
\\
&=
\mathbb{E}_{x,x' \sim \mathbb{P}} [k(x',x)] - \frac{2}{T+1}\sum \limits_{t=1}^{T+1} \mathbb{E}_{x \sim \mathbb{P}} [k(x,x_{t})] +\frac{1}{(T+1)^2}\sum \limits_{\substack{
t=1\\ t=t'}}^{T+1} k(x_{t}, x_{t'})+
\\
&\hspace{0.5cm}
+\frac{2}{(T+1)^2}\sum \limits_{\substack{
t=1\\ t\neq t'}}^{T+1} k(x_{t}, x_{t'})
\\
\end{align*}
So $\varepsilon_{T+1}$ depends on $x_{T+1}$ only through $-\frac{2}{T+1} \mathbb{E}_{x \sim \mathbb{P}} [k(x,x_{T+1})]+
\frac{2}{(T+1)^2} \sum \limits_{t=1}^{T}k(x_{t},x_{T+1})
$
The term $k(x_{T+1},x_{T+1})$ is not included, because by assumption it is equal to the constant R.
\\
Recognize that this term is the negative of the objective function maximized with respect to x. So the sample $x_{T+1}$ minimizes the error at time step $T+1$, i.e. $\varepsilon_{T+1}$ 
%Notice that for many kernels, explicit expressions for ${E}_{x' \sim \mathbb{P}} [k(x,x')]$ have been obtained, see \cite{jebara}.
\\
During the iterative step of kernel herding we maximize the negative of this quantity, thus we are minimizing the error greedily. In the sense that at each iteration we choose the x that minimizes our current error; however this does not guarantee that the samples states are jointly optimal.
\\
Intuitively, at each iteration, herding searches for a new sample to add to the pool; it is attracted to the regions where p is high and pushed away from regions where samples have already been selected.
\\
% At each iteration, this algorithm searches for a new datapoint to add to the set of supersamples. After that, kernel mean embedding using the supersamples can be computed.
% Additionally, note that the kernel herding algorithm can be used to generate samples from the learned distribution that are more informative then i.i.d. random samples.

### 2024-02-02 thesis/input_files/workplan.tex
added thesis, (the first commit is uploading the folder of the master thesis preparation latex)

Workplan is to start by considering the performance of kernel herding in generating an empirical distribution of the electricity prices and compare the performance to the methods in \cite{probablistic_electricity_forecast}.
To do so, with kernel herding we generate samples and as a consequence we also have an empirical distribution of electricity prices. Then we take the quantiles from 0.01 to 0.99 and finally we evaluate the distribution forecast through the Continous Ranked Probability Score CRPS.
\\
In this study, the data from the \href{https://www.epexspot.com/en/market-data?market_area=CH&trading_date=2023-12-11&delivery_date=2023-12-12&underlying_year=&modality=Auction&sub_modality=DayAhead&technology=&product=60&data_mode=table&period=&production_period=}{EPX} martket will be used; data is retrieved daily from the data provider through an automatic python script.
\\
Depending on the results of kernel herding, we may also consider how kernelized quantile regression behaves in the same setting. Next, additional kernel methods applied to problems concerning energy prices and related subjects could be taken into account.

### 2024-02-02 src/CDF.py
cleaned code for TestKernel and its dependencies

@@ -5,6 +5,9 @@ def cdf_funct(x):
 
 
 def inv_cdf_funct(y,x):
+    # x has to be in format np([])
+    x=x.reshape(x.size,)
+
     sorted_x_axis=np.sort(x)
     idx_pos= y*float(len(sorted_x_axis))
     # print(idx_pos)

### 2024-02-02 src/pinball_crps.py

@@ -1,7 +1,7 @@
 # function to compute pinball score
 
 from sklearn.metrics import mean_pinball_loss
-
+import numpy as np
 
 
 def pinball(tau, Yt_pred, Yt):
@@ -20,6 +20,11 @@ def CRPS(Y_pred_test, Y_test):
     Y_test= observed Y for t in test:=[T, T']
     Y_pred_test=predicted for Y_test
     """
+    # pinball requires input to be passed as np([[]])
+    Y_pred_test=np.array([Y_pred_test])
+    Y_test=np.array([Y_test])
+
+
     CRPS=0
     # for every t in test
     for t in range(len(Y_test)):
@@ -28,7 +33,7 @@ def CRPS(Y_pred_test, Y_test):
             # discrete approximation of CPRS
             CRPS+=pinball(tau/100, Y_pred_test[t], Y_test[t])
     
-    CRPS/=len(Y_test)
+    CRPS/=Y_test.size
 
     return CRPS
 

### 2024-02-24 thesis/input_files/analysis.tex
thesis chapters structure

Analysis of experiments and results

- Comments
\\
- Comparison
\\
- Table of loss scores
\\
Plots:
\\
- Plots for visualizing timeseries with quantiles bounds
\\
- Other plots that will come up to mind

### 2024-02-24 thesis/input_files/data_exploratory.tex
thesis chapters structure

- Explain how data has been retrieved.

- Data provider EPEX(entsoe retrieves its data)

- Explain the ETL(Extract Transform Load) pipeline
I set up.

- Explore data in order to get useful insights for
how to tune the models to get the most of them.

- Correlation and auto correlation plots

- Split train and test dataset
Explain carefully why it is important to carry
out an out of sample test and not in sample.
In sample test involves look ahead bias because we are
fitting the model on the data we want to predict,
thus it overfits on the data considered but it does
not generalize well.

### 2024-02-24 thesis/input_files/energymarket.tex
thesis chapters structure

Chapter explaining the energy market

- How auctions work

- Difference intraday dayhead all other stuff

- Prices can be negative

### 2024-02-24 thesis/input_files/implementation.tex
thesis chapters structure

Section documenting code

- Explain how quantile regression,
random forest, gradient boosting have been used
that is explain how their implementation has been
adapted to my specific setting.

- Explain in detail how to my src code has been implemented
its rationale and how to use it.

- As I explain code scripts go over the test, to 
explain better my ideas.

### 2024-02-24 thesis/input_files/literaturereview.tex
thesis chapters structure

@@
 %SPIEGO UN PO' I PAPER CHE HO FATTO PASSARE VELOCEMENTE 
+Every paper uses different datasets (heterogeneous)
+So it is not possible to compare directly results
+from one paper to another without implementing the
+paper specific algorithms and the applying them to
+your dataset.
+This is why sections after are destined to analyzing
+how these proposed methods so far work and their mathematical
+theory details
+\\
 This problem has been already addressed in \cite{probablistic_electricity_forecast} and \cite{probablistic_electricity_forecast2}. 

### 2024-02-24 thesis/input_files/problemdescription.tex
thesis chapters structure

- Here explain in detail what this thesis is about.

- what is the question? objectives of the thesis
the description of the problem tackled and the methodology used to solve
it.

- why is it relevant.

- an outline of the thesis with the main points presented in each chapter.

### 2024-02-24 thesis/input_files/symbols.tex
thesis chapters structure

\glsxtrnewsymbol[description={position}]{x}{\ensuremath{x}}
\glsxtrnewsymbol[description={velocity}]{v}{\ensuremath{v}}
\glsxtrnewsymbol[description={acceleration}]{a}{\ensuremath{a}}
\glsxtrnewsymbol[description={time}]{t}{\ensuremath{t}}
\glsxtrnewsymbol[description={force}]{F}{\ensuremath{F}}

### 2024-02-24 src/pinball_crps.py
.

@@ -3,7 +3,6 @@
 from sklearn.metrics import mean_pinball_loss
 import numpy as np
 
-
 def pinball(tau, Yt_pred, Yt):
     """
     - quantile tau
@@ -27,11 +26,13 @@ def CRPS(Y_pred_test, Y_test):
 
     CRPS=0
     # for every t in test
-    for t in range(len(Y_test)):
+    for t in range(Y_test):
         # for every tau in [0.01, ..., 0.99]
         for tau in range(1,100):
             # discrete approximation of CPRS
             CRPS+=pinball(tau/100, Y_pred_test[t], Y_test[t])
+            print("tau: ", tau, "t: ", t, "CRPS: ", CRPS)
+            print("------------------")
     
     CRPS/=Y_test.size
 

### 2024-03-06 thesis/input_files/literaturereview.tex
6 march

@@
+Energy Forecasting
+
+\subsection{Energy forecasting classification}
+
+\subsubsection{Types}
+In the context of energy forecasting, the quantities of most interest are price, load and renewables generation.
+
+\subsubsection{Forecasting horizons}
+\subsubsection{Size}
+- Forecasting horizons
+
+- Refer to section \ref*{metrics} for evaluating metrics
+
+- plots come quelli delle review con i dati di Scopus se é una cosa fattibile
+\label{sec:literaturereview}
+
+\begin{figure}[h!]
+    \includegraphics[width=\textwidth]{images/epf_evolution1.jpg}
+    \caption{EPF publications \cite{EPF_review}}
+    \label{fig:epf_evolution}
+  \end{figure}
+
+\begin{figure}[h!]
+  \includegraphics[width=\textwidth]{images/point_vs_prob.jpg}
+  \caption{Point vs probabilistic pubblications}
+  \label{fig:point_vs_prob}
+\end{figure}
+
+\begin{figure}[h!]
+  \includegraphics[width=\textwidth]{images/cs_stat_both.jpg}
+  \caption{Pubblications by method}
+  \label{fig:point_vs_prob}
+\end{figure}
+
+\begin{figure}[h!]
+  \includegraphics[width=\textwidth]{images/subject_area.jpg}
+  \caption{Pubblications by subject area}
+  \label{fig:subject_area}
+\end{figure}
+
+\begin{figure}[h!]
+  \includegraphics[width=\textwidth]{images/src_title.jpg}
+  \caption{Most popolar sources/outlets}
+  \label{fig:src_title}
+\end{figure}
+
 %INTRO KERNEL METHODS E ENERGY PRICE MODELLING
@@
 \\
+************************
+\\
+During the past 25 years %because paper of 2014 says 15, so now it is 25
+a wide range of new ideas have been proposed for point forecasting and for probabilistic forecasting.
+\\
+The field benefitted greatly from the increse of computing power, the greater availability of dataset and the interest in data science.
+As a consequence, the forecaster's toolbox has grown in size and complexity.
+\\
+Such variety of methods is characterized by heterogeneity in the fields from which they come from; methods come from statisitics, mathematics, econometrics, electrical engineering and the artificial intelligence communities.
+\\
+Before delving into the literature review, it is important to make clear that at this point in time there is no superior method.
+Different solutions may outperform or underperform compared to other techniques depending on the problem settings.
+Thus, understanding the complexity, strenghts and weaknesses of each method is crucial for fitting the right model to the right setting.
+\\
+Finally, within this research community, emerged the need of more homogeneity in the choice of the error valuation metrics and in the way of comparing model performances \cite{EPF_review}.
+\\
+************************
+\\
+Lately, the idea of combining forecasts has gained popularity in the forecasting communit \cite{forecasting_big}; in the literature, combined forecasts are called ensemble \cite{gneiting_weather_ensemble}.
+Experimental results have shown ensemble methods to outperform their component forecasts.
+\\
+Note that the more the errors of the combined models are not correlated the more we can benefit from ensembles.
+\\
+It is also worth noting that older and simpler methods are still valuable(in combination with other models or on their own); these being less subject to overfitting than complex models.

### 2024-03-06 thesis/input_files/metrics.tex
6 march

Proper evaluation methods guide researchers in choosing the model that best fits their needs; thus, this chapter is dedicated to the most common evaluation metrics adopted by academics in the field of EF.
\section{MAPE}\label{mape}

\section{CRPS}
Nevertheless, we can evaluate the integral in closed form. 
As pointed out in \cite{proper_scores}, we can use lemma 2.2 of \cite{new_multi_tes1} or identity 17 of \cite{new_multi_test2}.
\\
The lemma states that ...
\\
Note that in our case the distribution G of Y1 and Y2 is degenerate, with all probability mass on a single point (x notation will need to be cleaned here). Since $G(t)=:$
\\
It follows that the third addend in the summation is zero since the expectation of Y1 Y2 with distribution G means the difference of two equal constant numbers.
\\
Additionally, since x is constant in the first term we have Y=x.

### 2024-03-06 thesis/input_files/problemdescription.tex
6 march

@@
-- Here explain in detail what this thesis is about.
+Individuals and organizations constantly face situations of uncertainty.
+Thus, the need of robust forecasting methods.
+Such methods are crucial to the process of taking informed decision and to strategic planning.
+\\
+The basic idea of forecasting is that we can extract knowledge from the past in order to make educated guess about the future.
+Consequently, the range of fields where forecasting can be applied is very wide.
+In this thesis, our focus lies on applying forecasting to the energy sector. 
+\\
+\\
+The reason of our decision to focus on the energy market is mainly motivated by the rapid changes it has experienced.
+Over the last decades, electricity markets have gone through an unprecedented transformation; this shift was driven by the liberalization of such markets, the development and integration of renewable energy sources, the  increase of low carbon technologies and the adoption of smart meters.
+Events like the California electricity crisis help also motivating the choice of the electricity sector as subject of our studies, see \cite{california}.
+Interestingly, the process of deregulation lead to an increasing interest in the field of electricity price forecasting (EF) within the academic community \ref{fig:epf_evolution}.
+\\
+Moreover, the United Nations have identified the right to access affordable, reliable, sustainable and modern energy as one of their 17 SDGs \cite{un_sdgs}
+\\
+Finally, the electricity market has a set of features that make it unique: electricity cannot be stored in an efficient way and supply and demand have to be matched instantly.
+\\
+\section{Motivation}
+The are mutliple reasons why the energy sector needs robust forecasting techniques.
+For power market companies, being able to predict prices with a low MAPE \ref{mape} results in increased savings \cite{savings}.
+Furthermore,the adoption of smart meters provides power market companies with a ton of consumer data; this will enable them to better model consumer preferences.
+\\
+Transmission system operators main goal is to match supply and deman, generally TSO do so by increasing or decreasing the generation supplied.
+Thus, from their view point forecasting is critical for balancing the electricity netwrok.
+\\
+Other possible applications are: control of storage, demand side response, anomaly detection, network design and planning, simulating inputs and hangling missing data.
+\\
+\section{Point vs probabilistic forecast}
+A distinction has to be made between two types of forecasting approaches: point forecasts and probabilistic forecasts.
+Point prediction, also called deterministic forecasting in the literature, is all about predicting a particular value in time.
+On the other hand, with probabilistic forecasting we aim at predicting either a prediction interval, quantiles or a probability distribution for each point in time.
+For this reason, probabilistic forecasts are more informative than point forecasts; and this is why the interest of the research community is shifting towards them.
+Note that a probabilistic forecast can be turned into a point forecast by simply taking its expecatation.
+Alternatively, a probabilistic forecast can be derived from a point one by modeling the residuals of the point prediction.
+
+\section{Aims and objectives}
 
@@
 
-- why is it relevant.
+The scope of the thesis is analyzing state of the art forecasting methods in the energy market and to compare them with ideas coming from kernel theory. Applications of such methods to
 
-- an outline of the thesis with the main points presented in each chapter.
+\section{Outline}
+We start with a literature review and bibliometric analysis in section \ref{literature_review}.
+Then the theory underlying kernel methods is covered in section \ref{kernel_theory}.
+Following, section \ref{state_of_the_art} introduces the state of the art methods in the context of both point and probabilistic forecasting.
+Section \ref{energy_market} explains the core features and terminology of the energy market and of the electricity newtork.
+Evaluation metrics necessary to rank the forecasting techniques are presented in section \ref{metrics}. Section \ref{eda} goes on with the ETL pipeline and the exploratory analysis. 
+Implementation details are included in \ref{implementation}.
+Finally section \ref{experiments} presents the experiments, the results and discusses models' strenghts, weaknesses and possible improvements.

### 2024-03-06 thesis/input_files/symbols.tex
6 march

@@
-\glsxtrnewsymbol[description={position}]{x}{\ensuremath{x}}
-\glsxtrnewsymbol[description={velocity}]{v}{\ensuremath{v}}
 \glsxtrnewsymbol[description={acceleration}]{a}{\ensuremath{a}}
-\glsxtrnewsymbol[description={time}]{t}{\ensuremath{t}}
-\glsxtrnewsymbol[description={force}]{F}{\ensuremath{F}}
+
+%C
+\glsxtrnewsymbol[description={Continous ranked probability score}]{CRPS}{CRPS}
+
+%E
+\glsxtrnewsymbol[description={Exploratory data analysis}]{EDA}{EDA}
+\glsxtrnewsymbol[description={Electric forecasting}]{EF}{EF}
+\glsxtrnewsymbol[description={Electric vehicles}]{EVs}{EVs}
+\glsxtrnewsymbol[description={Extract transform load}]{ETL}{ETL}
+
+%L
+\glsxtrnewsymbol[description={Low carbon technologies}]{LCTs}{LCTs}
+
+\glsxtrnewsymbol[description={Low voltage}]{LV}{LV}
+
+%P
+\glsxtrnewsymbol[description={Prediction interval}]{PI}{PI}
+\glsxtrnewsymbol[description={Price Forecasting}]{PF}{PF}
+
+\glsxtrnewsymbol[description={Probabilistic Price Forecasting}]{PPF}{PPF}
+
+%S
+\glsxtrnewsymbol[description={Sustainable development goals}]{SDGs}{SDGs}
+
+%
+\glsxtrnewsymbol[description={Transmission system operator}]{TSO}{TSO}

### 2024-03-06 src/poly_fit.py

@@ -1,9 +1,10 @@
 import numpy as np
-import math
-from src.cdf import cdf_funct
-from src.miscellaneous import find
 from scipy.optimize import curve_fit
 
+import math
+
+from src.cdf import cdf_funct
+from src.miscellaneous import find
 
 def spl_polyfit(x,y):
     s=3

### 2024-03-09 thesis/input_files/literaturereview.tex
9 march 2024

--- 2024-02-02 thesis/input_files/literaturereview.tex
@@
+Electricity Forecasting 
+%intro
+During the past 25 years %because paper of 2014 says 15, so now it is 25
+a wide range of new ideas have been proposed for point forecasting and for probabilistic forecasting.
+\\
+The field benefitted greatly from the increse of computing power, the greater availability of dataset and the interest in data science.
+As a consequence, the forecaster's toolbox has grown in size and complexity.
+\\
+As already pointed out, such variety of methods is characterized by the heterogeneity in the fields from which they come from; methods come from statisitics, mathematics, econometrics, electrical engineering and the artificial intelligence communities.
+\\
+Before delving into the literature review, it is important to make clear that at this point in time there is no superior method.
+Different solutions may outperform or underperform compared to other techniques depending on the problem settings.
+Thus, understanding the complexity, strenghts and weaknesses of each method is crucial for fitting the right model to the right setting.
+\\
+Finally, within this research community, it emerged the need of more homogeneity in the choice of the error valuation metrics (section \ref{metrics}), data quality and in the way of comparing model performances \cite{EPF_review}; as a solution, \cite{review_2021} proposes a checklist to aid evaluating the meaningfulness of new research.
+Hence, throughout this thesis work, we will stick to the proposed principles and best practices peculiar of the EF field.
+%literature review
+
+\section{Electricity forecasting classification}
+Electricity forecasting is a vague term and it is used in the literature to refer to the whole field.
+Thus, in order to introduce some clarity it is useful to classify the range of EF articles in terms of their core attributes.
+\subsection{Types}
+In the context of energy forecasting, the quantities of most interest are prices (EPF), loads (ELF) and renewables generation (mostly wind and solar).
+
+\subsection{Forecasting horizons}\label{cls_time}
+In terms of forecasting horizons, we can group EF into four major categories: very short term forcasting (VSTF), short term load forecasting (STF), medum term forecasting (MTF) and long term forecasting (LTF).
+Consensus in the literature is to use as cut off horizons one day, two weeks and three years respectilvely \cite{hong_phd}; see \ref{fig:time} for a visualisation.
+\begin{figure}
+  \includegraphics[width=\textwidth]{images/time.jpg}
+  \caption{Time classsification \cite{prob_elf}}
+  \label{fig:time}
+\end{figure}
+
+\subsection{Size}
+Forecasts can either be for the whole target electricity network (system) or for a subset of it (zonal).
+
+\subsection{Point vs Probabilistic}\label{cls_pp}
+Finally, EF literature distinguishes between point and probabilistic forecasts.
+Each of the two has its advantages and disadvantages; point forecasts are easier to generate and less computationally intensive while probabilistic forecast are more informative.
+Industry and research efforts have focused primarily on point forcasting.
+Nevertheless, interest in probabilistic forecasting has risen considerably over the last years due to renewable integration requirements, introduction of smart grids and increased market competitivness.
+\\
+\\
+To conclude this section
+Ziel et Steinert carefully tables, are reported in the appendix for the interested reader.
+
+\section{Bibliographic analysis}
+This section presents the results of the bibliometric analysis we performed on 6th March 2024.
+This survey has been carried out by using the Scopus citation database.
+For details on the specific queries entered in Scopus, refer to \ref{src_code}.
+\\
+To get started, let us consider the evolution of the EF field over the years.
+This is visualized in figure \ref{fig:epf_evolution}, with results grouped by category; note, articles prior to the 2000 have been aggregated together due to their small number.
+Figure \ref{fig:epf_evolution} shows the trend of an increasing interest in EF.
+\begin{figure}
+    \includegraphics[width=\textwidth]{images/epf_evolution1.jpg}
+    \caption{EF publications}
+    \label{fig:epf_evolution}
+  \end{figure}
+\\
+The next question was to compare the state of point versus probabilistic forecasting, this is visualized in figure \ref{fig:point_vs_prob}.
+What it can be concluded is that probabilistic is less developed than point forecasting.
+To our mind this is due to the complexity of probabilistic forecast.
+Nevertheless, we can see a trend that suggests researchers are making an effort to fill this gap.
+\begin{figure}
+  \includegraphics[width=\textwidth]{images/point_vs_prob.jpg}
+  \caption{Point vs probabilistic pubblications}
+  \label{fig:point_vs_prob}
+\end{figure}
+\\
+The EF literature is dominated by statistical and computational intelligence methods as can be seen from figure \ref{fig:cs_stat_both}, with CI methods beign slightly preferred. 
+\begin{figure}
+  \includegraphics[width=\textwidth]{images/cs_stat_both.jpg}
+  \caption{Pubblications by method}
+  \label{fig:cs_stat_both}
+\end{figure}
+\\
+EF is an heterogenous field of research, its researchers come from a wide array of backgrounds, with electrical engineers and statisticians making up the top contributors; their different educational training may explain why the split between statistical and computational intelligence methods is so marked.
+Figure \ref{fig:subject_area} depitcs the EF publications by subject area.
+What can be concluded, is that the bulk of pubblications come from engineering, computer science, mathematics and econometrics.
+\begin{figure}
+  \includegraphics[width=\textwidth]{images/subject_area.jpg}
+  \caption{Pubblications by subject area}
+  \label{fig:subject_area}
+\end{figure}
+\\
+Finally, in order to refer to the most relevant source in the field, EF outlets have been ranked by popularity and plotted in \ref{fig:src_title}.
+\begin{figure}
+  \includegraphics[width=\textwidth]{images/src_title.jpg}
+  \caption{Most popolar sources/outlets}
+  \label{fig:src_title}
+\end{figure}
+\\
+\section{EF literature review}
+To get started a few review articles were collected in order to understand conventions, best practicies and terminology of the EF community.
+Weron \cite{EPF_review} reviews the state of the art for electricity price forecasting; beside analysing complexity of available solutions, strenghts and weaknesses it also stresses the need for objective comparative EPF studies.
+Specifically, it advocates for studies using similar datasets, using the same error evaluation metric and statistical testing model's outperformance.
+Hong et al.
+\cite{prob_elf} discusses the state of the art in probabilistic electric load forecasting; it differentiates between techniques and methodologies.
+With techniques they refer to a family of models, like multiple linear regression or artificial neural networks.
+On the other hand, methodologies consist of general frameworks that can be incorporated into any method, for example variable selection mechanisms.
+Also this paper stresses the need for some guidelines to standardize research in the field.
+Nowotarski et al. \cite{nowotarski} also carried out a thorough review of probabilistic forecasting.
+Weron et al. \cite{lago} offer a set of best practices when forecasting electricity prices in order to have a common framework to evaluate and compare future research.
+Zhang et al. \cite{zhang2014review} considers state of the art methods in wind power probabilistic forecasting and describes current challenges and possible future developments.
+Ziel et al. \cite{ziel2018probabilistic} provides detailed tables, grouping research papers by the time dimension \ref{cls_time} and objective \ref{cls_pp} and reporting dataset, model and accuracy 
+measures adopted.
+David et al.
+\cite{david2016probabilistic} adopt a combination of ARMA and GARCH in probabilistic forecasts of solar irradiance.
+Furthermore, they propose a recursive framework for parameter estimation.
+De Gooijer \cite{de200625} reviews 25 years of time series forecasting for the period 1990-2005, highlighting the most influential works.
+He et al.
+\cite{he2022cooperative} models  multistep wind speed probabilistic forecasting by mixing CEEDMAN, LASSO and QRNN.
+In this work, CEEMDAN is used to decompose the wind speed time series; LASSO compresses high dimensional features; QR is used for obtaining quantile forecasts; finally KDE converts quantile forecasts into density estimates.
+Wan et al. \cite{wan2016direct} proposes a combination of QR and ELM to generate non parametric probabilistic forecast of wind generation.
+Van der Meer et al. \cite{van2018review} provides another thorough anlysis of the probabilistic forecasting realm by covering recent advances and identifying research gaps.
+The IEEE Power and Energy Society provides also insightful lecture notes on probabilistic energy forecasting methodologies,
+implementations, and applications \cite{gm_22}.
+Marcjasz et al \cite{probablistic_electricity_forecast2} uses distributional neural networks to crete probabilistic forecasts for the day ahead electricity prices in the german market.
+Nowotarski et al.
+\cite{nowotarski2015computing} introduce a method for constructing PI and call it quantile regression averaging.
+Their idea is weighting a set of models' prediction such that the pinball loss of the weighted model is minimized.
+The observed results is that QRA performed better compared to twelve individual models.
+Arora et al.
+\cite{arora2016forecasting} focus on modelling electricity smart meter data by proposing a non parametric probabilistic technique based on kernel density estimation and conditional kernel density estimation \cite{rosenblatt1969conditional} \cite{hyndman1996estimating}.
+Their conclusion is that kernel density methods are competitive against exponential smoothing when forecasting residential data.
+Conversely, exponential smoothing has still an edge in predicting SME data.
+Zhang et al.
+\cite{zhang2020probability} introduce a framework based on quantile regression and kernel density estimation in the context of short term wind forecasting.
+The proposed methods behave well compared to a ARMA(1,1).
+Haben et al. \cite{haben2018short} analyses a variety of techniques in terms of both probabilistic and point forecasting, within this study they focus on load forecasting at the low voltage level.
+Zhang et al \cite{zhang2020two} proposes a two stage bootstrap sampling framework for probabilistic load forecasting.
+They test it for different regression models such as RF, GBRT, linear regression, and LSSVM regression.
+Jónsson et al. \cite{jonsson2014predictive}
+introduces a density model for the day ahead market extending the adaptive QR framework of \cite{moller2008time} by modelling the tails of the predicted density with an exponential distribution.
+Huurman et al.
+\cite{huurman2012power} surveys the predictive power of weather variables for electricity prices in danish market.
+Their empirical results suggest that weather is central for point forecasting day ahead prices.
+The opposite conclusion are drawn for density forecasting
+%GEFCom2012
+A major step forward in EF was the creation of the global energy forecasting competition (GEFCom) in 2012.
+Until then, no formal benchmarking process or data pool was established and new publications rarely reproduced the results from work done by others.
+Addressing these issues was the motivation behind the creation of GEFCom by the IEEE working group on energy forecasting.
+The EF field was positively affected by this competition; a number of ideas were tested on the same setting with only the best ones being published and it also contributed bridging the gap between industry practice and academic research.
+GEFCom 2012 had two tracks; the former about hierarchical load forecasting, the latter about wind power forecasting, see \cite{hong2014global} for a comprehensive review.
+\\
+%GEFCom 2014
+The focus of GEFCom 2014 was on probabilistic forecasting, Hong et al. \cite{hong2016probabilistic} discusses the problem tracks, the data and the winning methods.
+In this paragraph some of the winning entries of the 2014 GEFCom edition are discussed.
+Xie et al.
+\cite{xie2016gefcom2014} propose a two stage approach; in the first stage they use MLR to build a point forecast, then in the second stage they try different approaches for modelling the MLR residuals, among other they tried ESM, ANN and ARIMA.
+Maciejowska et al. \cite{maciejowska2016hybrid} proposes a new probabilistic model extending the idea of QRA \cite{nowotarski2015computing}.
+Haben et al. \cite{haben2016hybrid} mixes CKD and QR in their competition entry.
+Gaillard et al. \cite{gaillard2016additive} \cite{gaillardasemi} combines quantile regression with generalized additive models \cite{hastie2017generalized}.
+Ziel et al. \cite{ziel2016lasso} estimates an AR model through the LASSO \cite{tibshirani1996regression} instead of the standard OLS.
+\\
+%GEFCom 2017
+The last GEFCom was held in 2017, its focus was providing probabilistic load forecasts, see \cite{hong2019global}.
+The GEFCom competition has also inspired the organization of other competitions such as the RWEnpower competition in the UK, the RTE competition in France, the Tokyo electric power company competition in Japan and the BigDEAL forecasting competitions.
+\\
+\\
+Considerations that can be drawn from the above literature review.
+Every paper uses different datasets (heterogeneous)
+So it is not possible to compare directly results
+from one paper to another without implementing the
+paper specific algorithms and the applying them to
+your dataset.
+This is why sections after are destined to analyzing
+how these proposed methods so far work and their mathematical
+theory details
+\\
+************************
+\\
+************************
+\\
+Lately, the idea of combining forecasts has gained popularity in the forecasting communit \cite{forecasting_big}; in the literature, combined forecasts are called ensemble \cite{gneiting_weather_ensemble}.
+Experimental results have shown ensemble methods to outperform their component forecasts.
+\\
+Note that the more the errors of the combined models are not correlated the more we can benefit from ensembles.
+\\
+It is also worth noting that older and simpler methods are still valuable(in combination with other models or on their own); these being less subject to overfitting than complex models.
+\\
+\section{Kernel methods literature review}
 %INTRO KERNEL METHODS E ENERGY PRICE MODELLING
-
 Kernel methods are a class of algorithms for patter analysis.
 With kernel methods we are able to apply linear methods with predictors in a high dimensional space, without having to explicitly evaluate the involved dot products of the features.
-In this thesis work I will address the performance of kernel methods in the context of probabilistic forecasting; the area of application will be the electricity market. 
-Probabilistic forecasting may be useful to power producers, traders and consumers in order to improve their decision making process and managing risk(VaR).
-This is because probabilistic forecast enables them to simulate scenarios and carry out stress tests.
-
-%SPIEGO UN PO' I PAPER CHE HO FATTO PASSARE VELOCEMENTE 
-This problem has been already addressed in \cite{probablistic_electricity_forecast} and \cite{probablistic_electricity_forecast2}. 
-These papers surveyed the performance of neural network architectures against simpler approaches like quantile regression and data fitting to Johnson distribution.
-Their conclusion is that distributional NN perform a little worse than quantile regression but the former has smaller computational cost; that is because the quantile regression is run for every quantile from 0.01 to 0.99.
-Nevertheless kernel methods received very little attention in this specific setting.
+Throughout this thesis work, we will address the performance of kernel methods in the context of EF.
 \\
-Kernel methods considered are kernel mean embedding \cite{pmlr}, \cite{Muandet_2017} and kernel herding \cite{supersamples}.
-Particularly extending the idea of \cite{2022nystrom}, where the Nyström approximation is employed in computing the kernel mean embedding, experiments with the Pivoted Cholesky decomposition will be performed.
+- Kernel history evolution.
+Their name comes from the german word kern, which translates to core in english.
+Such term was first used by David Hilbert in his paper on integral equations \cite{hilbert} where he defined a definite kernel.
+Following, Hilbert's and Schmidt's \cite{schmidt} work lead to the introduction of a new space, the Hilbert space.
+In 1909, James Mercer improved Hilbert's work by proposing his theorem \cite{mercer}. This theorem underlies the power of kernel methods, that is the kernel trick.
+In 1938, Schoenberg \cite{schoenberg} developed the mathematical results that allow us to find the kernel associated to a specific feature space metric.
+In 1941, Kolmogorov \cite{kolmogorov} carried out stuides on representing kernel in linear spaces.
+In 1950, Aronszajn \cite{aronszajn} published the first work on RKHS; developing the general method for representing kernels in linear spaces.
+In 1964, Aizerman \cite{aizerman} further improved the theory of RKHS.
+It was in the nineties that theory of kernel methods got popular, particularly in the field of machine learning.
+Kernels have been used in various different tasks such as SVM \cite{vapnik1} \cite{vapnik2}, Gaussian process classifiers \cite{williams}, spline methods \cite{wahba}, neural networks \cite{poggio} and principal component analysis \cite{pca_scholkopf}.
+Nevertheless kernel methods received very little attention in the specific setting of EF literature.
 \\
-In section 2, three papers that lay the basis for this thesis’ work are summarized.
+The theory needed for this thesis work is covered in section \ref{kernel_theory}. 
+For an introduction to kernel methods, we referred to \cite{learning_with_kernels}.
+Kanagawa et Fukumizu introduces to the concept of kernel mean embedding \cite{pmlr}.
+Muandet et al.
+\cite{Muandet_2017} surveys established results and new advances in the theory of Hilbert space distribution embeddings.
+It has to be said that computing and storing such embeddings becomes prohibitive for large scale settings.
+Rudi et al.
+\cite{2022nystrom} proposes an efficient approximation procedure based on the Nyström method \cite{nystrom}, providing also an upper bound for the approximation error.
+Article \cite{supersamples} presents kernel herding; basically, Smola et al.
+used the kernel trick to extend the herding algorithm to continous spaces.
+The result is an infinite memory deterministic process that takes in a collection of samples and learns to approximate a pdf.
 \\
+%QUESTO DA VEDERE
+% Particularly extending the idea of \cite{2022nystrom}, where the Nyström approximation is employed in computing the kernel mean embedding, experiments with the Pivoted Cholesky decomposition will be performed.

### 2024-03-09 thesis/input_files/symbols.tex
9 march 2024

@@
-\glsxtrnewsymbol[description={acceleration}]{a}{\ensuremath{a}}
+%For mathematical terms \ensuremath
+% \glsxtrnewsymbol[description={Autoregressive exogenous model}]{ARX}{\ensuremath{ARX}}
+
+%A
+\glsxtrnewsymbol[description={Artificial neural network}]{ANN}{ANN}
+\glsxtrnewsymbol[description={Autoregressive model}]{AR}{AR}
+\glsxtrnewsymbol[description={Autoregressive moving average model}]{ARMA}{ARMA}
+\glsxtrnewsymbol[description={Autoregressive integrated moving average model}]{ARIMA}{ARIMA}
+\glsxtrnewsymbol[description={Autoregressive exogenous model}]{ARX}{ARX}
 
 %C
+\glsxtrnewsymbol[description={Computational intelligence}]{CI}{CI}
+\glsxtrnewsymbol[description={Conditional kernel density}]{CKD}{CKD}
 \glsxtrnewsymbol[description={Continous ranked probability score}]{CRPS}{CRPS}
+
+%D
+\glsxtrnewsymbol[description={Distributional neural network}]{DDNN}{DDNN}
 
@@
 \glsxtrnewsymbol[description={Exploratory data analysis}]{EDA}{EDA}
-\glsxtrnewsymbol[description={Electric forecasting}]{EF}{EF}
+\glsxtrnewsymbol[description={Electricity forecasting}]{EF}{EF}
+\glsxtrnewsymbol[description={Electricity load forecasting}]{ELF}{ELF}
+\glsxtrnewsymbol[description={Empirical mode decomposition}]{EMD}{EMD}
+\glsxtrnewsymbol[description={Electricity load forecasting}]{EPF}{EPF}
+\glsxtrnewsymbol[description={Exponential smoothing models}]{ESM}{ESM}
+\glsxtrnewsymbol[description={Extract transform load}]{ETL}{ETL}
 \glsxtrnewsymbol[description={Electric vehicles}]{EVs}{EVs}
-\glsxtrnewsymbol[description={Extract transform load}]{ETL}{ETL}
+
+%G
+\glsxtrnewsymbol[description={Gradient boosting regression tree}]{GBRT}{GBRT}
+\glsxtrnewsymbol[description={Global energy forecasting competition}]{GEFCom}{GEFCom}
+
+%H
+\glsxtrnewsymbol[description={Holt-Winters-Taylor exponential smoothing method}]{HWT}{HWT}
+
+%K
+\glsxtrnewsymbol[description={Kernel density estimation}]{KDE}{KDE}
 
@@
 \glsxtrnewsymbol[description={Low carbon technologies}]{LCTs}{LCTs}
+\glsxtrnewsymbol[description={Least squares support vector regression}]{LSSVR}{LSSVR}
 
 \glsxtrnewsymbol[description={Low voltage}]{LV}{LV}
+
+%M
+\glsxtrnewsymbol[description={Multiple linear regression}]{MLR}{MLR}
+
+%O
+\glsxtrnewsymbol[description={Ordinary least squares}]{OLS}{OLS}
+
+%R
+\glsxtrnewsymbol[description={Random forest}]{RF}{RF}
+\glsxtrnewsymbol[description={Reproducing kernel hilbert space}]{RKHS}{RKHS}
 
@@
 
+%Q
+\glsxtrnewsymbol[description={Quantile regression}]{QR}{QR}
+
+\glsxtrnewsymbol[description={Quantile regression averaging}]{QRA}{QRA}
+
 %S
+\glsxtrnewsymbol[description={Small and medium-sized enterprises}]{SME}{SME}
+
+\glsxtrnewsymbol[description={Smoothed nonparametric ARX}]{SNARX}{SNARX}
+
 \glsxtrnewsymbol[description={Sustainable development goals}]{SDGs}{SDGs}
+\glsxtrnewsymbol[description={Support vector machine}]{SVM}{SVM}
 
-%
+%T
+\glsxtrnewsymbol[description={Time of use tariffs}]{TOU}{TOU}
 \glsxtrnewsymbol[description={Transmission system operator}]{TSO}{TSO}

### 2024-03-10 thesis/abstract.tex
updated thesis project
6 march
March 10th 2024

@@
+% PRELIMINARY ABSTRACT, ONCE I AM DONE WRITE THE FINAL ABSTRACT
 \begin{abstract}
-    The theory of kernel methods will be applied to the problem of probabilistic forecasting taking as input some data.
-    Conceptually, the proposed analysis could be applied to any kind of data. In this study we considered the electricity market, because of the interesting implications  on risk management tasks.
+    The theory of kernel methods will be applied to the problem of point and probabilistic forecasting for the energy sector.
+    Such choice is motivated by its interesting implications.
+
+    % on risk management tasks.
+
+    % - Contribution: Kernel herding applied to the
+    % field of probabilistic forecasting
 

### 2024-03-10 thesis/input_files/literaturereview.tex
March 10th 2024

@@
-Electricity Forecasting 
+%Electricity Forecasting 
 %intro
 During the past 25 years %because paper of 2014 says 15, so now it is 25
-a wide range of new ideas have been proposed for point forecasting and for probabilistic forecasting.
+a wide range of new ideas have been proposed for electricity point forecasting and for probabilistic forecasting.
 \\
@@
 As a consequence, the forecaster's toolbox has grown in size and complexity.
-\\
-As already pointed out, such variety of methods is characterized by the heterogeneity in the fields from which they come from; methods come from statisitics, mathematics, econometrics, electrical engineering and the artificial intelligence communities.
 \\
@@
 \\
-Finally, within this research community, it emerged the need of more homogeneity in the choice of the error valuation metrics (section \ref{metrics}), data quality and in the way of comparing model performances \cite{EPF_review}; as a solution, \cite{review_2021} proposes a checklist to aid evaluating the meaningfulness of new research.
+Finally, within this research community, it emerged the need of more homogeneity in the choice of the error valuation metrics (section \ref{metrics}), data quality and in the way of comparing model performances \cite{EPF_review}; as a solution, \cite{lago} proposes a checklist to aid evaluating the meaningfulness of new research.
 Hence, throughout this thesis work, we will stick to the proposed principles and best practices peculiar of the EF field.
@@
 Nevertheless, interest in probabilistic forecasting has risen considerably over the last years due to renewable integration requirements, introduction of smart grids and increased market competitivness.
-\\
-\\
-To conclude this section
-Ziel et Steinert carefully tables, are reported in the appendix for the interested reader.
-
 \section{Bibliographic analysis}
-This section presents the results of the bibliometric analysis we performed on 6th March 2024.
+This section presents the results of the bibliometric analysis we performed on March 6th 2024.
 This survey has been carried out by using the Scopus citation database.
@@
 \\
-Finally, in order to refer to the most relevant source in the field, EF outlets have been ranked by popularity and plotted in \ref{fig:src_title}.
+Finally, in order to refer to the most relevant source in the field, EF outlets have been ranked by popularity and plotted in figure \ref{fig:src_title}.
 \begin{figure}
@@
 \cite{he2022cooperative} models  multistep wind speed probabilistic forecasting by mixing CEEDMAN, LASSO and QRNN.
-In this work, CEEMDAN is used to decompose the wind speed time series; LASSO compresses high dimensional features; QR is used for obtaining quantile forecasts; finally KDE converts quantile forecasts into density estimates.
-Wan et al. \cite{wan2016direct} proposes a combination of QR and ELM to generate non parametric probabilistic forecast of wind generation.
+In this work, CEEMDAN is used to decompose the wind speed time series, LASSO compresses high dimensional features, QR is used for obtaining quantile forecasts, finally KDE converts quantile forecasts into density estimates.
+Wan et al. \cite{wan2016direct} proposes a combination of QR and ELM to generate non parametric probabilistic forecasts of wind generation.
+Zhang et al. \cite{zhang2019wind} forecasts wind speed by adopting QRMGM, which is a combination of QR with a minimal gated memory network.
+Hyndman et al. \cite{hyndman1996estimating} explains kernel estimator of conditional density, analyses its asymptotic behaviour and covers an application for the daily temperatures of Melbourne.
+Kaur et al. \cite{kaur2022energy} carries out a comparative study of techniques spanning ideas from statistic and artificial intelligence.
 Van der Meer et al. \cite{van2018review} provides another thorough anlysis of the probabilistic forecasting realm by covering recent advances and identifying research gaps.
@@
 implementations, and applications \cite{gm_22}.
-Marcjasz et al \cite{probablistic_electricity_forecast2} uses distributional neural networks to crete probabilistic forecasts for the day ahead electricity prices in the german market.
+Marcjasz et al. \cite{probablistic_electricity_forecast2} uses distributional neural networks to create probabilistic forecasts for the day ahead electricity prices in the german market.
 Nowotarski et al.
@@
 \cite{zhang2020probability} introduce a framework based on quantile regression and kernel density estimation in the context of short term wind forecasting.
-The proposed methods behave well compared to a ARMA(1,1).
+The proposed methods behave well compared to an ARMA(1,1).
 Haben et al. \cite{haben2018short} analyses a variety of techniques in terms of both probabilistic and point forecasting, within this study they focus on load forecasting at the low voltage level.
+%Metrics Start
+Koochali et al. \cite{koochali2022random} reviews various existing methods for assessing probabilistic forecast models and discusses their advantages and disadvantages.
+Matheson et al. \cite{matheson1976scoring} develops classes of scoring rules for continuous probability distributions.
+Gneiting et al. \cite{gneiting2007strictly}
+provides a thorough overview of the theory of proper scoring rules for interval and density forecasts.
+Gneiting et al. \cite{gneiting2014probabilistic}
+covers theory and techniques state of the art in probabilistic forecasting.
+%Metrics End
 Zhang et al \cite{zhang2020two} proposes a two stage bootstrap sampling framework for probabilistic load forecasting.
@@
 introduces a density model for the day ahead market extending the adaptive QR framework of \cite{moller2008time} by modelling the tails of the predicted density with an exponential distribution.
+Fatema et al.
+\cite{fatema2023probabilistic} considers gaussian process regression for point forecasting and prediction intervals prediction.
+Then, it inputs PI to KDE in order to estimate a probability distribution.
+Dudek \cite{dudek2018probabilistic} proposes a probabilistic forecasting model based on the Nadaraya Watson estimator \cite{nadaraya1964estimating} \cite{watson1964smooth}.
 Huurman et al.
-\cite{huurman2012power} surveys the predictive power of weather variables for electricity prices in danish market.
+\cite{huurman2012power} surveys the predictive power of weather variables for electricity prices in the danish market.
 Their empirical results suggest that weather is central for point forecasting day ahead prices.
-The opposite conclusion are drawn for density forecasting
+The opposite conclusion are drawn for density forecasting.
+\\
+Lately, the idea of combining forecasts has gained popularity in the forecasting community \cite{forecasting_big}; in the literature, combined forecasts are called ensemble \cite{gneiting_weather_ensemble}.
+Experimental results have shown ensemble methods to outperform their component forecasts.
+Note that, the more the errors of the combined models are not correlated the more we can benefit from ensembles.
+It is also worth noting that older and simpler methods are still valuable (in combination with other models or on their own); these being less subject to overfitting than complex models.
+\\
 %GEFCom2012
@@
 \\
-Considerations that can be drawn from the above literature review.
-Every paper uses different datasets (heterogeneous)
-So it is not possible to compare directly results
-from one paper to another without implementing the
-paper specific algorithms and the applying them to
-your dataset.
-This is why sections after are destined to analyzing
-how these proposed methods so far work and their mathematical
-theory details
+A couple of considerations can be drawn from the above literature review.
+As already pointed out, the EF field is characterized by heterogeneity in its forecasting techniques; methods come from statisitics, mathematics, econometrics, electrical engineering and the artificial intelligence communities.
 \\
-************************
+Every paper uses different datasets.
+Therefore, it is not possible to compare directly results from one paper to another without implementing the paper specific algorithms and then applying them to your dataset.
+An additional hurdle is that some datasets are not freely accessible.
 \\
-************************
-\\
-Lately, the idea of combining forecasts has gained popularity in the forecasting communit \cite{forecasting_big}; in the literature, combined forecasts are called ensemble \cite{gneiting_weather_ensemble}.
-Experimental results have shown ensemble methods to outperform their component forecasts.
-\\
-Note that the more the errors of the combined models are not correlated the more we can benefit from ensembles.
-\\
-It is also worth noting that older and simpler methods are still valuable(in combination with other models or on their own); these being less subject to overfitting than complex models.
-\\
+Thus, a good understanding of state of the art methods in EF is required to carry out a rigorous comparison between methods.
+That is why the following chapters are devoted to summarising the mathematical theory underlying such techniques.
+
 \section{Kernel methods literature review}
@@
 \\
-- Kernel history evolution.
+%- Kernel history evolution.
 Their name comes from the german word kern, which translates to core in english.
-Such term was first used by David Hilbert in his paper on integral equations \cite{hilbert} where he defined a definite kernel.
+Such term was first used by David Hilbert in his paper on integral equations \cite{hilbert} where he introduced the term of definite kernels.
 Following, Hilbert's and Schmidt's \cite{schmidt} work lead to the introduction of a new space, the Hilbert space.
@@
 \\
-The theory needed for this thesis work is covered in section \ref{kernel_theory}. 
-For an introduction to kernel methods, we referred to \cite{learning_with_kernels}.
+The kernel theory needed for this thesis work is covered in section \ref{kernel_theory}. 
+For an introduction to kernel methods, we referred to \cite{learning_with_kernels} \cite{hofmann2006review}.
+\\
 Kanagawa et Fukumizu introduces to the concept of kernel mean embedding \cite{pmlr}.
@@
 \cite{Muandet_2017} surveys established results and new advances in the theory of Hilbert space distribution embeddings.
-It has to be said that computing and storing such embeddings becomes prohibitive for large scale settings.
+It has to be said that, computing and storing such embeddings becomes prohibitive for large scale settings.
 Rudi et al.

### 2024-03-10 thesis/input_files/problemdescription.tex
9 march 2024
March 10th 2024

@@
-Individuals and organizations constantly face situations of uncertainty.
-Thus, the need of robust forecasting methods.
+Individuals and organizations constantly face situations of uncertainty thus the need of robust forecasting methods.
 Such methods are crucial to the process of taking informed decision and to strategic planning.
 \\
-The basic idea of forecasting is that we can extract knowledge from the past in order to make educated guess about the future.
+The basic idea of forecasting is that we can extract knowledge from the past in order to make educated guesses about the future.
 Consequently, the range of fields where forecasting can be applied is very wide.
-In this thesis, our focus lies on applying forecasting to the energy sector. 
+In this thesis, our focus lies on applying forecasting techniques to the energy sector. 
 \\
-\\
-The reason of our decision to focus on the energy market is mainly motivated by the rapid changes it has experienced.
+The reason of our decision to focus on the energy market is mainly motivated by the rapid changes it has undergone. %/experienced 
 Over the last decades, electricity markets have gone through an unprecedented transformation; this shift was driven by the liberalization of such markets, the development and integration of renewable energy sources, the  increase of low carbon technologies and the adoption of smart meters.
 Events like the California electricity crisis help also motivating the choice of the electricity sector as subject of our studies, see \cite{california}.
-Interestingly, the process of deregulation lead to an increasing interest in the field of electricity price forecasting (EF) within the academic community \ref{fig:epf_evolution}.
+Furthermore, the process of deregulation lead to an increasing interest in the field of electricity price forecasting (EF) within the academic community \ref{fig:epf_evolution}.
 \\
-Moreover, the United Nations have identified the right to access affordable, reliable, sustainable and modern energy as one of their 17 SDGs \cite{un_sdgs}
+In addtion, the United Nations have identified the right to access affordable, reliable, sustainable and modern energy as one of their 17 SDGs \cite{un_sdgs}.
 \\
@@
 For power market companies, being able to predict prices with a low MAPE \ref{mape} results in increased savings \cite{savings}.
-Furthermore,the adoption of smart meters provides power market companies with a ton of consumer data; this will enable them to better model consumer preferences.
+Furthermore,the adoption of smart meters provides power market companies with a ton of consumer data; this can enable them to better model consumer preferences.
 \\
-Transmission system operators main goal is to match supply and deman, generally TSO do so by increasing or decreasing the generation supplied.
+Transmission system operators' main goal is to match supply and deman, generally TSO do so by increasing or decreasing the generation supplied.
 Thus, from their view point forecasting is critical for balancing the electricity netwrok.
 \\
-Other possible applications are: control of storage, demand side response, anomaly detection, network design and planning, simulating inputs and hangling missing data.
+Probabilistic forecasting may be useful to power producers, traders and consumers in order to improve their decision making process and managing risk(VaR).
+This holds in particular for traders, because probabilistic forecast enables them to simulate scenarios and carry out stress tests.
+\\
+Other possible applications are: control of storage, demand side response, anomaly detection, network design and planning, simulating inputs and handling missing data.
 \\
@@
 
-- what is the question? objectives of the thesis
-the description of the problem tackled and the methodology used to solve
-it.
+% - what is the question? objectives of the thesis
+% the description of the problem tackled and the methodology used to solve
+% it.
 
-The scope of the thesis is analyzing state of the art forecasting methods in the energy market and to compare them with ideas coming from kernel theory. Applications of such methods to
+The scope of the thesis is analyzing state of the art forecasting methods in the energy market and to compare and to integrate them with ideas coming from the theory of kernel methods.
 

### 2024-03-12 thesis/input_files/metrics.tex
March 11th 2024

@@
 Proper evaluation methods guide researchers in choosing the model that best fits their needs; thus, this chapter is dedicated to the most common evaluation metrics adopted by academics in the field of EF.
-\section{MAPE}\label{mape}
+Error metrics and measures vary depending on wethter we are concerned with point or probabilistic forecasts.
+Additionally, note that the latter can take different forms which therefore requires different measures.
+\section{MAE}\label{mae}
+Consider the time series with actual values given by $\mathrm{L}=(L_{n+1}, L_{n+2},\dots, L_{n+h})$
+and its h step ahead point forecast $\mathrm{\hat{L}}=(\hat{L}_{n+1}, \hat{L}_{n+2},\dots, \hat{L}_{n+h})$ the mean absolute error is defined as
+\begin{definition}
+    $\mathrm{MAE}(\mathrm{L},\mathrm{\hat{L}})=\frac{1}{h}\|\mathrm{L}-\mathrm{\hat{L}}\|_{1}=\frac{1}{h}\sum\limits_{k=1}^{h}|L_{n+k}-\hat{L}_{n+k}|$
+\end{definition}
 
+\section{RMSE}\label{rmse}
+\begin{definition}
+    $\mathrm{RMSE(\mathrm{L}, \mathrm{\hat{L}})}=\frac{1}{\sqrt{h}}\|\mathrm{L}-\mathrm{\hat{L}}\|_{2}=\sqrt{\frac{\sum\limits_{k=1}^{h}(L_{n+k}- \hat{L}_{n+k})^2}{h}}$
+\end{definition}
+
+MAE and RMSE posses the useful property of being expressed in the same units of the data thus enabling meaningful comparisons.
+However, a drawback of such measures is that we cannot use them to compare accuracy between time series wich have different magnitudes.
+For instance, a day ahead error of 1kWh is negligible when considering a daily demand of 100kWn while the same error is considerably big when daily demand is 2kWh.
+This consideration leads to relative accuracy scores, between those the MAPE is by far the most popular.
+
+\section{MAPE}\label{mape}. 
+\begin{definition}
+    $\mathrm{MAPE}(\mathrm{L},\mathrm{\hat{L}})=\frac{100}{h}\sum\limits_{k=1}^{h}\frac{|L_{n+k}-\hat{L}_{n+k}|}{|L_{n+k}|}$
+\end{definition}
+
+\section{RMSPE}\label{rmspe}
+\begin{definition}
+    $\mathrm{RMSPE}(\mathrm{L},\mathrm{\hat{L}})=100\cdot\sqrt{\frac{1}{h}\sum\limits_{k=1}^{h} \left(\frac{|L_{n+k}-\hat{L}_{n+k}|}{|L_{n+k}|}\right)^2}$
+\end{definition}
+Note, MAPE and RMSPE may not be appropriate for series which have zero or very small values, for example, electricity demand at the household level; the result is a large score regardless of the absolute errors.
+Scaled errors constitute a robust family of scores.
+\section{MASE}\label{mase}
+\begin{definition}
+    $\mathrm{MASE}(\mathrm{L},\mathrm{\hat{L}})=\frac{1}{h}\sum\limits_{k=1}^N\frac{|L_{n+k}-\hat{L}_{n+k}|}{\frac{1}{h-1}\sum\limits_{k=2}^{h}|L_{k}-L_{k-1}|}$
+\end{definition}
+In the denominator we have the error of the naïve/persistence model. 
+In this model, the current demand makes up the prediction for the next time step; that is $\mathrm{\hat{L}^{\mathrm{naive}}}_{n+1}=\mathrm{L}_{n}$.
+\section{RMSSE}\label{rmsse}
+\begin{definition}
+    $\mathrm{RMSSE}(\mathrm{L},\mathrm{\hat{L}})=\sqrt{\frac{1}{h}\sum\limits_{k=1}^N\left(\frac{|L_{n+k}-\hat{L}_{n+k}|}{\frac{1}{h-1}\sum\limits_{k=2}^{h}|L_{k}-L_{k-1}|}\right)^2}$
+\end{definition}
+
+\section{Pinball}\label{pinball}
+The pinball score or quantile score is used to measure the accuracy of a quantile forecast.
+\begin{definition}
+    $\mathrm{Pinball}(L_{t},\hat{L}_{t,q},q)=
+\begin{cases}
+(1-q)(\hat{L}_{t,q}-L_{t}) & L_t < \hat{L}_{t,q} \\
+q(L_t-\hat{L}_t,q) & L_t \geq \hat{L}_{t,q}
+\end{cases}$
+\end{definition}
+The pinball loss is an asymmetric function, it weights its score differently depending on the error sign and on the quantile considered, see \ref{fig:pinball}.
+\begin{figure}
+    \includegraphics[width=\textwidth]{images/pinball_loos.png}
+    \caption{Pinball loss with q=0.2 \cite{haben2023core}}
+    \label{fig:pinball}
+  \end{figure}
+By averaging all the pinball losses over all quantiles and over the whole forecast horizon, we obtain the pinball loss of the probabilistic forecast.
+\section{Winkler}
+\begin{definition}
+    $\mathrm{Winkler}_{\alpha}=\begin{cases}
+        \delta & Low_{t}\leq L_{t}\leq Upp{t}\\
+        \delta+2(Low_{t}-L_{t})/\alpha & L_{t}<Low_{t}\\
+        \delta+2(L{t}-Upp_{t})/\alpha & L_{t} \geq Upp_{t}
+    \end{cases}$
+\end{definition}
+Where delta is the PI width, that is $\delta=Upp_t-Low_t$. This score penalizes observations falling outside the PI and rewards narrow PI.
 \section{CRPS}
+The continous ranked probability score measures the differece between the estimated cumulative distribution $\hat{F}$ and the empirical Cdf.
+\begin{definition}\label{def_crps}
+    $\mathrm{CRPS}(L, \hat{F})=\int\limits_{-\infty}^{\infty}\left(\hat{F}(x)-\mathbb{1}(x-L) \right)^2 dx$
+\end{definition}
 Nevertheless, we can evaluate the integral in closed form. 
-As pointed out in \cite{proper_scores}, we can use lemma 2.2 of \cite{new_multi_tes1} or identity 17 of \cite{new_multi_test2}.
+Where the indicator function is defined as 
+    $\mathbb{1}(z)=
+\begin{cases}
+0 & z<0\\
+1 & z \geq 0
+\end{cases}$
 \\
-The lemma states that ...
+For a visualisation see, \ref{fig:crps}. The grey area is what contributes toward the CRPS score.
+The better the estimated cdf is the smaller the total CRPS score will be.
+\begin{figure}
+    \includegraphics[width=\textwidth]{images/crps.png}
+    \caption{CRPS integral \cite{haben2023core}}
+    \label{fig:crps}
+  \end{figure}
   \\
-Note that in our case the distribution G of Y1 and Y2 is degenerate, with all probability mass on a single point (x notation will need to be cleaned here). Since $G(t)=:$
+It is worth noting, that the CRPS integral can be rewritten in terms of expectations.
+This makes its evaluation easier, since we know that the sample mean converges to the expectation by the law of large numbers.
+This was first pointed out by \cite{proper_scores}, where authors take advantage of lemma 2.2 of \cite{new_multi_test2} or equivalently identity 17 of \cite{new_multi_tes1}.
 \\
-It follows that the third addend in the summation is zero since the expectation of Y1 Y2 with distribution G means the difference of two equal constant numbers.
+\begin{lemma}
+    Let $\mathrm{X}_1$, $\mathrm{X}_2$, $\mathrm{Y}_1$, $\mathrm{Y}_2$ be independent real random variables with finite expectations.
+Let $\mathrm{X}_1,\mathrm{X}_2$ be identically distributed with distribution function $\mathrm{F}$ and let $\mathrm{Y}_1,\mathrm{Y}_2$ be identically distributed with distribution function $\mathrm{G}$.
+Then
+   \begin{equation}
+    \mathbb{E}(|X_1-Y_1|)-\frac{1}{2}\mathbb{E}(|X_1-X_2|)-\frac{1}{2}\mathbb{E}(|Y_1-Y_2|)=\int\limits_{-\infty}^{\infty}\left(F(x)-G(x)\right)^2dx
+\end{equation}
+\end{lemma}
+Notice that, in our case  \ref{def_crps}, the distribution $G$ of $Y_1$ and $Y_2$ is degenerate, with all probability mass on a single point.
+%  (x notation will need to be cleaned here). Since $G(t)=:$
+It follows that, the third addend in the summation is zero.
+That is because the fact of $Y_1$ and $Y_2$ both following distribution $G$ implies that $\mathbb{E}(|Y_1-Y_2|)$ corresponds to the difference of two equal constant numbers.
 \\
-Additionally, since x is constant in the first term we have Y=x.
+Additionally, since $Y_1$ is just a constant, we have $Y_1=L$.
+\\
+Putting everything together we have obtained an alternative way of computing the CRPS score.
+\begin{equation}
+    \int\limits_{-\infty}^{\infty}\left(\hat{F}(x)-\mathbb{1}(x-L) \right)^2 dx=\mathbb{E}(|X_1-L|)-\frac{1}{2}\mathbb{E}(|X_1-X_2|)
+\end{equation}
+\\
+\section{PIT}
+The probability integral transform is a method to assess visually the quality of probabilistic forecasts.
+PIT is obtained by applying a cdf $F$ to your data; if applying such cdf to the data results in a uniform distributed PIT then $F$ is a valid prediction.
+If not, $F$ is not the suited cdf for the considered data.
+Figure \ref{fig:pit} provides an example, applying the true cdf results in a well calibrated PIT (left).
+Alternatively, applying a bad cdf results in either a overdispersed (middle) or underdispersed (right) PIT.
+\begin{figure}
+    \includegraphics[width=\textwidth]{images/pit.png}
+    \caption{PIT types \cite{haben2023core}}
+    \label{fig:pit}
+  \end{figure}
+\\
+%RELIABILITY PLOT
+%PICP,NMPIW,CWC,unconditional coverage(arora)-->Probability density forecasting of wind speed based on quantile regression and kernel density estimation Lei Zhang
+
+% Reliability and sharpness \section{Criteria}-->Recent advances in electricity price forecsating a review of probabilistic forecasting Weron
+% Two criteria are used in literature to evaluate probabilistic forecasting: sharpness and reliability. Reliability means

### 2024-03-12 thesis/input_files/point.tex
March 11th 2024

This chapter covers the theory of the most widely used method for point forecasting in the EF literature. Besides discussing the theory underlying these models, this chapter includes also a couple of worked examples in order to get acquainted with the practical applications of models.
\section{Exponential smoothing}

\section{Multiple linear regression}

\section{Autoregressive models}
explain their theory
and how the procedure how they are used
IMPORTANT: load series is not a stationary series so before applying AR we have to perform stationary tests or differencing steps
\subsection{ARIMA}
\subsection{ARIMAX}
\subsection{SARIMA}
\subsection{SARIMAX}

\section{Generalized additive model}

\section{K-nearest neighbors}

\section{Support vector regression}

\section{Artificial neural networks}
\subsection{DNN}
\subsection{LSTM}
\subsection{DeepAr}
\subsection{AR Net}

\section{Kernel methods}
\subsection{Kernel regression}
\subsection{Kernel support vector regression}

### 2024-03-12 thesis/input_files/probabilistic.tex
March 11th 2024

\section{Quantile regression}
\input{input_files/qr.tex}

\section{Kernel density estimation}
\input{input_files/kde.tex}

\section{Ensemble methods}
\input{input_files/ensemble.tex}

\section{Copula models}

\section{DMLP}

\section{DeepAR}

\section{Kernel methods}
\subsection{Kernel quantile regression}
\subsection{Kernel herding}

### 2024-03-12 thesis/input_files/symbols.tex
March 10th 2024
March 11th 2024

@@
 \glsxtrnewsymbol[description={Global energy forecasting competition}]{GEFCom}{GEFCom}
+\glsxtrnewsymbol[description={Gaussian process regression}]{GPR}{GPR}
 
@@
 %M
+\glsxtrnewsymbol[description={Mean absolute error}]{MAE}{MAE}
+\glsxtrnewsymbol[description={Mean absolute scaled error}]{MASE}{MASE}
+\glsxtrnewsymbol[description={Mean absolute percentage error}]{MAPE}{MAPE}
+\glsxtrnewsymbol[description={Minimal gated memory network}]{MGM}{MGM}
+
 \glsxtrnewsymbol[description={Multiple linear regression}]{MLR}{MLR}
@@
 \glsxtrnewsymbol[description={Reproducing kernel hilbert space}]{RKHS}{RKHS}
+\glsxtrnewsymbol[description={Root mean squared error}]{RMSE}{RMSE}
 
@@
 \glsxtrnewsymbol[description={Transmission system operator}]{TSO}{TSO}
+
+%MATHEMATICAL NOTATION
+\glsxtrnewsymbol[description={feature vector}]{phi(x)}{\ensuremath{\phi(x)}}
+
+\glsxtrnewsymbol[description={kernel function}]{k(x,x')}{\ensuremath{k(x,x')}}

### 2024-03-13 thesis/input_files/point.tex
March 13th 2024

@@
 This chapter covers the theory of the most widely used method for point forecasting in the EF literature.
-Besides discussing the theory underlying these models, this chapter includes also a couple of worked examples in order to get acquainted with the practical applications of models.
+Besides discussing the theory underlying these models, this and the following chapter include also a couple of worked examples in order to get acquainted with the practical applications of models.
 \section{Exponential smoothing}

### 2024-03-13 thesis/input_files/probabilistic.tex
March 13th 2024

@@
+When it comes to probabilistic forecasting, there are a couple of standard practical approaches; conceptually they can be grouped into two main categories.
+The former class of approaches tries modelling the distribution of the observed data directly.
+The latter family of approaches constructs first a point forecast and then learns the distribution of the model's errors.
 \section{Quantile regression}

### 2024-03-13 thesis/input_files/qr.tex
March 13th 2024

Quantile regression can be interpreted as an extension of standard regression. In this setting, you basically slice the dependent variables into quantiles and then fit a regression for each quantile. With standard regression we build a model for the conditional mean, conversely, with quantile regression we model the conditional quantile function for any desired quantile. 
Therefore, with quantile regression we are able to study the impact of covariates on quantiles directly.
\begin{definition}
    For any real valued random variable Y, we define its associated quantile function.
    \begin{equation}
        Q(q)=\inf{y:F(y)\geq q}
    \end{equation}
\end{definition}
Alternatively, in order  to ease the posing of the quantile regression problem, we can formulate quantiles as the solution to a simple optimization problem.
\\
For any $0<q<1$ consider the pinball loss function \ref{pinball} $\rho_q(u)=u(q-\mathbb{1}_{u<0})$. 
Such loss is minimized by the quantiles $Q(q)$.
We can estimate them by minimizing the expectation of $\rho_q(Y-g(x,\beta))$ with respect to the parameter $\beta$.
\\
Note, that in the special case $q=\frac{1}{2}$,  quantile regression corresponds exactly to standard regression with an absolute value loss function.
\\
It follows that the conditional linear quantile function $Q_Y(q|X=x)=x_i\beta(q)$ can be estimated by solving
\begin{equation}\label{}
    \hat{\beta(q)}=\argmin{\beta}\sum \rho_q   (y_i-x \beta)
\end{equation}
\\
Notice that, the cost function is not differentiable, therefore there is no analytical solution to the quantile regression problem. Nevertheless, we can easily solve it by employing linear programming.
\\
Furthermore, can extend to non linear quantile regression by choosing a non linear model in place of $x\beta$ in the above equation.
\\
In order to illustrate the approach, let us consider applying quantile regression to the maximum daily temperature dataset in Melbourne \cite{hyndman1996estimating}; models considered will be linear, gradient boosting machine, quantile forest\cite{} and kernel quantile regression.

- se si puo fare plot con stessi dati ma differenti f nel quantile regression

### 2024-03-13 dataset_applications/qr_melbourne.py

@@ -1,0 +1,169 @@
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas as pd
+from pprint import pprint
+
+from quantile_forest import RandomForestQuantileRegressor as rfr
+
+from sklearn.ensemble import GradientBoostingRegressor as gbr
+
+from sklearn.experimental import enable_halving_search_cv  # noqa
+from sklearn.model_selection import HalvingRandomSearchCV
+from sklearn.metrics import make_scorer
+from sklearn.metrics import mean_pinball_loss
+from sklearn.model_selection import train_test_split
+
+import statsmodels.regression.quantile_regression as qr 
+import sys
+
+from tqdm import tqdm
+
+
+
+if __name__=="__main__":
+    # load data
+    df=pd.read_csv("/Users/luca/Desktop/ThesisKernelMethods/dataset_applications/temperatures_melbourne.csv", sep=";", decimal=".")
+
+    quantiles = [0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9]
+    pinball_scores=pd.DataFrame(columns=["Linear qr", "Gbm qr", "Quantile forest"],index=quantiles)
+
+    # plot data
+    plt.plot(df["Yt-1"],df["Yt"],"o", alpha=0.2)
+    plt.show()
+
+    # train, test split
+    X_train, X_test, y_train, y_test = train_test_split(df["Yt-1"], df["Yt"], test_size=0.20, random_state=4)
+
+    eval_set=np.linspace(df["Yt"].min(), df["Yt"].max(), 100).T
+    eval_set=eval_set.reshape(-1,1)
+    X_train= X_train.values.reshape(-1, 1)
+    X_test = X_test.values.reshape(-1, 1)
+
+
+    # quantile regression: linear, random forest, gbm
+
+    # linear quantile regression
+    qr_models = [qr.QuantReg(y_train, X_train).fit(q=q) for q in quantiles]
+    
+    y_test_pred_qr=[qr_model.predict(X_test) for qr_model in qr_models]
+
+    for i, q in enumerate(quantiles):
+        pinball_scores.loc[q,"Linear qr"]=mean_pinball_loss(y_train,qr_models[i].predict(X_train), alpha=q)
+
+    # plot model fit
+    # linear quantile regression
+    plt.plot(df["Yt-1"],df["Yt"],"o", alpha=0.2)
+    for i,q in enumerate(quantiles):
[…]

### 2024-03-13 dataset_applications/temperatures_melbourne.R

@@ -1,0 +1,4 @@
+library(hdrcde)
+temperatures_melbourne=data.frame(Y=as.matrix(maxtemp))
+temperatures_melbourne
+write.csv(df, "temperatures_melbourne.csv", row.names=FALSE)

### 2024-03-15 thesis/input_files/probabilistic.tex
March 15th 2024

@@
 \subsection{Kernel quantile regression}
+The idea of quantile regression has been extended to kernel methods by Takeuchi et al. \cite{takeuchi2006nonparametric}.
+There they minimize a risk plus regularizer defined as follows.
+\begin{equation}\label{eq:kqr_min1}
+    R[f]:=\frac{1}{m}\sum\limits_i=1^{m}\rho_q(y_i-f(x_i))+\frac{\lambda}{2}\|g\|_H^2
+\end{equation}
+where $f=g+b$, $g \in H$ and $b \in R$.
+Using the link between RKHS and feature spaces, we can rewrite $f(x)=\langle w, \phi(x) \rangle+b$. Doing so we obtain a minimization problem equivalent to minimizing equation \ref{eq:kqr_min1}.
+\begin{equation}\label{eq:kqr_min2}
+    \begin{aligned}
+    \min_{w,b} \quad & C \sum \limits_{i=1}^{m}
+    (y_i-\langle w, \phi(x) \rangle-b)+ (1-q)(-y_i+\langle w, \phi(x) \rangle+b)+ \frac{1}{2}\|w\|^2\\
+    \end{aligned}
+    \end{equation}
+Note the division by $\lambda$ so that $C=\frac{1}{\lambda m}$
+We can next rephrase the optimisation in \ref{eq:kqr_min2} by introducing the slack variables $\xi_i$ and $\xi_i^*$.
+\begin{equation}\label{eq:kqr_min3}
+    \begin{aligned}
+        \min_{w,b\xi_i,\xi_i^*} \quad & C \sum \limits_{i=1}^{m}
+        q \xi_i+ (1-q)\xi_i^*+ \frac{1}{2}\|w\|^2\\
+    \textrm{s.t.} \quad & y_i-\langle w, \phi(x) \rangle+b \leq \xi_i\\
+    & -y_i+\langle w, \phi(x) \rangle+b \leq \xi_i^*\\
+      &\xi_i\geq0    \\
+      &\xi_i^*\geq0    \\
+    \end{aligned}
+    \end{equation}
+In order to make it more compact, we rewrite equation \ref{eq:kqr_min3} in matrix notation.
+\begin{equation}\label{eq:kqr_min4}
+    \begin{aligned}
+        \min_{w,b\xi,\xi^*} \quad & C q \xi^\intercal \mathbb{1}+ C q (\xi^*)^\intercal \mathbb{1}+ \frac{1}{2}w^\intercal w\\
+    \textrm{s.t.} \quad & y-\Phi^\intercal w -b \preceq \xi\\
+    & -y+\Phi^\intercal w +b \preceq \xi^*\\
+      &\xi\succeq0    \\
+      &\xi^*\succeq0    \\
+    \end{aligned}
+    \end{equation}
+Consider now the associated lagrangian $L$ to \ref{eq:kqr_min4}
+\begin{equation}\label{eq:kqr_min5}
+    \begin{aligned}
+    L(w,b,\xi,\xi^*)= & C q \xi^\intercal \mathbb{1}+ C q (\xi^*)^\intercal \mathbb{1}+ \frac{1}{2}w^\intercal w- \alpha^\intercal(\xi - y+\Phi^\intercal w +b)
+    - (\alpha^*)^\intercal(\xi^* -y+\Phi^\intercal w +b)
+    \\
+    & -\nu^\intercal \xi - (\nu^*)^\intercal \xi^*
+\end{aligned}
+\end{equation}
+The next step is deriving its dual formulation, since it is easier and more efficient to solve. This because the dual problem has the useful property of being always convex.
+
+- dual problem
+\begin{definition}
+    The dual function associated to the lagrangia $L(x,\lambda)$ is given by $g(v)=\inf_x L(x,\lambda)$
+\end{definition}
+where $\lambda$ is called the lagrange multiplier of the optimization problem. Such dual formulation has an useful property, that is \begin{equation}\label{weak_duality}
+    g(\lambda)\leq p^*
+\end{equation}
+where $p^*$ is the optimal value of your optimization problem.
+Consider a simple lagrangian $L(x,\lambda)=f(x)+\sum \lambda_i i(x) +\sum v_i h(x)$, where $i(x)$ are inequality constraints while $e(x)$ are equality constraints of the problem.
+What we can be noted is that the lower bound on $p^*$ is non trivial only when the lagrange multiplier $\lambda \succeq 0$.
+Therefore, the idea is that by maximizing the dual function subject to the contraint $\lambda \succeq 0$ we can obtain a approximate or perfect solution to the primal problem.
+\\
+To explain why we may or not be able to attain the best solution to the primal problem by maximizing its dual we have to introduce the concept of duality.
+\\
+We use $d^*$ to denote the optimal value of the lagrange dual problem; we can think of it as the best lower bound on $p^*$.
+The inequality \ref{weak_duality} is called weak duality.
+The difference $p^*-d^*$ is said the optimal duality gap; it is the gap between the optimal value of the primal problem and the best lower bound on it that can be obtained from the Lagrange dual function.
+Moreover, note that the optimal duality gap is always nonnegative.
+\\
+We say that strong duality holds, when the optimal duality gap is zero; in other words, the lagrange dual bound is tight.
+
+- Slater's condition
+- Apply Slater's theorem=>we have strong duality
+- calculations
 \subsection{Kernel herding}

### 2024-03-15 thesis/input_files/qr.tex
March 15th 2024

@@
 Notice that, the cost function is not differentiable, therefore there is no analytical solution to the quantile regression problem.
-Nevertheless, we can easily solve it by employing linear programming.
+Nevertheless, we can easily solve it by employing linear programming and convex optimisation \cite{boyd2004convex}.
 \\
@@
 \\
-In order to illustrate the approach, let us consider applying quantile regression to the maximum daily temperature dataset in Melbourne \cite{hyndman1996estimating}; models considered will be linear, gradient boosting machine, quantile forest\cite{} and kernel quantile regression.
-
+In order to illustrate the approach, let us consider applying quantile regression to the maximum daily temperature dataset in Melbourne \cite{hyndman1996estimating}; models considered will be linear, gradient boosting machine, quantile forest\cite{meinshausen2006quantile} and kernel quantile regression.
 - se si puo fare plot con stessi dati ma differenti f nel quantile regression

### 2024-03-18 thesis/input_files/probabilistic.tex
March 18th 2024

@@
 
-\section{Copula models}
+% \section{Copula models}
+\section{Quantile forest}
+\section{Quantile gradient boosting machine}
 
 \section{DMLP}
-
+%gluon nn
+% train a beta distribution to learn its parameters
 \section{DeepAR}
@@
 where $f=g+b$, $g \in H$ and $b \in R$.
-Using the link between RKHS and feature spaces, we can rewrite $f(x)=\langle w, \phi(x) \rangle+b$. Doing so we obtain a minimization problem equivalent to minimizing equation \ref{eq:kqr_min1}.
+Using the link between RKHS and feature spaces, we can rewrite $f(x)=\langle w, \phi(x) \rangle+b$.
+Moreover, note that the RKHS norm is defined as follows $\|f\|_{\varphi }=\inf\{\|w\|_{F}:w\in F,f(x)=\langle w,\varphi (x)\rangle _{F},\forall x\in X\}$, where $F$ is the feature space.
+ Doing so we obtain a minimization problem equivalent to minimizing equation \ref{eq:kqr_min1}.
 \begin{equation}\label{eq:kqr_min2}
@@
 where $p^*$ is the optimal value of your optimization problem.
-Consider a simple lagrangian $L(x,\lambda)=f(x)+\sum \lambda_i i(x) +\sum v_i h(x)$, where $i(x)$ are inequality constraints while $e(x)$ are equality constraints of the problem.
+Consider a simple lagrangian $L(x,\lambda)=f(x)+\sum \lambda_i r_i(x) +\sum v_i h_i(x)$, where $r_i(x)$ are inequality constraints while $h_i(x)$ are equality constraints of the problem.
 What we can be noted is that the lower bound on $p^*$ is non trivial only when the lagrange multiplier $\lambda \succeq 0$.
@@
 We say that strong duality holds, when the optimal duality gap is zero; in other words, the lagrange dual bound is tight.
-
-- Slater's condition
-- Apply Slater's theorem=>we have strong duality
+\\
+Constraint qualifications are condtions under which strong duality holds; one the most popular is Slater's condition.
+\begin{equation}\label{slater_condition}
+    \begin{aligned}
+        \exists x \in \textrm{relint} \ D \ \textrm{s.t.} & \ r_i(x)<0, \quad i=1, \dots, m \\
+        & h_i(x)=b
+    \end{aligned}
+\end{equation}
+Where relint $D$ is the relative interior of $ D:=\cap _{i=0}^{m}\operatorname {dom} (r_{i})$
+\\
+Slater's theorem naturally follows.
+\begin{theorem}
+    If Slater's condtion holds and the problem is convex then strong duality holds.
+\end{theorem}
+We can now check that our optimization problem posseses strong duality by checking Slater's condition.
+\\
+In our case we don't have equlity constraint, so we do not have the $h_i(x)=0$ \ref{slater_condition} to be checked.
+All we have to check is the convexity of our problem and that there exist a $x$ such that $r_i(x)<0$.
+For convexity, a sufficient condition is the positive definiteness of $Q$ in the quadratic programming problem 
+\begin{equation}
+    \begin{aligned}
+        \min & x^\intercal Q x+ c^\intercal x \\        
+        s.t & Ax\preceq b
+    \end{aligned}
+\end{equation}
+In our case \ref{eq:kqr_min5}, we have $w^\intercal w$, thus $Q$ is just the identity matrix which satisfies the positive definiteness requirement. Therefore our problem is convex.
+Next we check Slater's condition holds.
+Considering first the non negative constraints on $\xi$ and $\xi^*$, we conclude that $\xi$  and $\xi^*$ have to be greater or equal to zero for the existence of an $x$ satisfying Slater's condition.
+Thus, let us suppose that $0 \leq \xi \leq \alpha$ and $0 \leq \xi^* \leq \alpha$.
+\\
+Next, let us consider the other two inequalities and make the following ansatz.
+\\
+\begin{equation}
+    \begin{aligned}
+        w=& \Phi^\intercal(\Phi \Phi^\intercal)^{-1} y\\
+        b\leq& \alpha
+    \end{aligned}
+\end{equation}
+We then have
+\begin{equation}
+    \begin{aligned}
+        -\xi + y -\Phi\Phi^\intercal(\Phi \Phi^\intercal)^{-1}y-b<&0
+        \\
+        -\xi^* - y +\Phi\Phi^\intercal(\Phi \Phi^\intercal)^{-1}y+b<&0
+    \end{aligned}
+\end{equation}
+Hence, we conclude that our problem satisfies Slater's condition. Therefore the solution of the dual and primal problem are equivalent.
+\\
+We end this section with the derivation of the dual problem; that is the convex problem we will solve in order to get the qunatiles prdeiction of our quantile kernel algorithm.
+\\
+First, derive the dual function of \ref{eq:kqr_min5}.
+\begin{equation}
+    \begin{aligned}
+        g(\alpha, \alpha^*, \nu, \nu^*)=& \inf_x L(x,\lambda)\\
+    = & \inf_{\xi, \xi^*, w, b} L(w,b,\xi,\xi^*)
+\end{aligned}
+\end{equation}
+Setting the derivates to zero
+\begin{equation}\label{eq:lagrange_derivatives}
+    \begin{cases}
+        \frac{\partial L}{\partial w}=0 \implies w=\Phi^\intercal(\alpha-\alpha^*)
+        \\
+        \frac{\partial L}{\partial b}=0 \implies \alpha-\alpha^*=0
+        \\
+        \frac{\partial L}{\partial \xi}=0 \implies Cq \mathbb{1}-\alpha- \nu=0
+        \\
+        \frac{\partial L}{\partial \xi^*}=0 \implies C(1-q)\mathbb{1} -\alpha^* -\nu^*=0
+    \end{cases}
+\end{equation}
+As pointed out previously, the lower bound resulting from the dual formulation is non trivial only when the lagrange multipliers are $\succeq 0$.
+Looking at the last two equations of the system \ref{eq:lagrange_derivatives}, this implies the following two constraints $\alpha \in [0, Cq\mathbb{1}]$ and $\alpha \in [0, C(1-q)\mathbb{1}]$.
+\\
+Substitute the conditions for an optimum into \ref{eq:kqr_min5} we obtain the dual formulation.
+\begin{equation}
+    \begin{aligned}
+        g(\alpha, \alpha^*)=& \xi^\intercal(Cq\mathbb{1}-\alpha -\nu)+(\xi^*)^\intercal(C(1-q)\mathbb{1}-\alpha^*-\nu^*)-(\alpha-\alpha^*)^\intercal \Phi\Phi^\intercal(\alpha-\alpha^*)
+        \\
+        & +(\alpha-\alpha^*)^\intercal y-(\alpha-\alpha^*)^\intercal b+\frac{1}{2}(\alpha-\alpha^*)^\intercal \Phi\Phi^\intercal(\alpha-\alpha^*)
+        \\
+        \\
+        g(\alpha, \alpha^*)=& 0+0-\frac{1}{2}(\alpha-\alpha^*)^\intercal \Phi\Phi^\intercal(\alpha-\alpha^*)+(\alpha-\alpha^*)^\intercal y-0
+        \\
+        g(\alpha, \alpha^*)=& -\frac{1}{2}(\alpha-\alpha^*)^\intercal \Phi\Phi^\intercal(\alpha-\alpha^*)+(\alpha-\alpha^*)^\intercal y
+    \end{aligned}
+\end{equation}
+Defining $\alpha=(\alpha-\alpha^*)$ and letting $K$ the kernel matrix, we have that the dual optimisation problem reads as follows
+\begin{equation}\label{eq:kqr_min6}
+    \begin{aligned}
+        \max_{\alpha} \quad & -\frac{1}{2}\alpha^\intercal K\alpha+\alpha^\intercal y\\
+    \textrm{s.t.} \quad & 
+    C(q-1)\mathbb{1}\preceq \alpha \preceq Cq\mathbb{1}\\
+    &\alpha^\intercal\mathbb{1}
+    \end{aligned}
+    \end{equation}
+Switching sign we rephrase as a minisation problem, which is the common practice in convex optimisation.
+\begin{equation}\label{eq:kqr_min7}
+    \begin{aligned}
+        \min_{\alpha} \quad & +\frac{1}{2}\alpha^\intercal K\alpha-\alpha^\intercal y\\
+    \textrm{s.t.} \quad & 
+    C(q-1)\mathbb{1}\preceq \alpha \preceq Cq\mathbb{1}\\
+    &\alpha^\intercal\mathbb{1}
+    \end{aligned}
+    \end{equation}
+The kernel quantile regression estimator is then given by
+\begin{equation}
+    f(x)=\sum\limits_i \alpha_i k(x_i, x)+b
+\end{equation}
+Where an unbiased estimator for $b$ is given by
+\begin{equation}
+    \begin{aligned}
+    b=&\frac{1}{I_q}\sum\limits_{i\in I_q} b_i, \\ 
+    \textrm{with} \ b_i=& y_i-\sum\limits_i \alpha_i k(x_i, x_i) \\
+     \textrm{and} \ I_q=& \{i=1,\dots,n|0<\alpha_i<Cq, 0<\alpha_i<C(1-q)\}
+    \end{aligned}
+\end{equation}
+%contribution is implementation kernel quantile regression for pythos users since there exists only r version
 - calculations
 \subsection{Kernel herding}
+Select the best point forecasting method and create a probabilistic forecast by modelling its model errors with kernel herding (do something like the residual bootstrap ensembles if it is meaningful and possible to implement).

### 2024-03-20 thesis/appendix.tex
updated thesis project
6 march
9 march 2024
March 20th 2024

@@
 \end{proof}
+
+\newpage
+\section{Src code}\label{src_code}
+The whole code for the project is hosted on
+\url{https://github.com/luca-pernigo/ThesisKernelMethods}\label{github_repo}.
+\\
+\begin{itemize}
+    \item query: folder containing Scopus data and scripts to generate bibliometric survey plots in section \ref{literature_review}
+    \item kqr.py: file implementing our custom kernel quantile regression
+\end{itemize}

### 2024-03-20 thesis/input_files/analysis.tex
March 20th 2024

@@
 Analysis of experiments and results
+\\
+Building on the theory introduced in \ref{ch:point} and \ref{ch:prob}, this section covers the experiments carried out and the results obtained.
 
@@
 - Other plots that will come up to mind
+
+\section{Quantile forecasting}
+In this subsection we compare performance of quantile regressor considering the GEFCom2014 dataset.
+\\
+- experiment for quantile estimator on gefcom2014 and we are good

### 2024-03-20 thesis/input_files/implementation.tex
6 march
March 11th 2024
March 20th 2024

@@
+This section is inteneded to explain and aid for reproducibilty studies. Hereafter the specific libraries used and the custom implementatios are thoroughly documented.
+\\
+For the list of python packages needed see the requirement.txt file.
+
 Section documenting code
+- indicate computer specifics
 
-- Explain how quantile regression,
-random forest, gradient boosting have been used
-that is explain how their implementation has been
+% - see whether something can be parallelized
+
+- Explain how methods' implementation has been
 adapted to my specific setting.
@@
 explain better my ideas.
+\section{Linear quantile regression}
+The implementation of statsmodels.regression.quantile\_regression.QuantReg has been used.
+The model is fitted through iterative reweighted least squares.
+\section{Quantile gradient boosting machine}
+The implementation of sklearn.ensemble.GradientBoostingRegressor as been used.
+\section{Quantile forest}
+The implementation of quantile\_forest.RandomForestQuantileRegressor as been used. This estimator is compatible with scikit-learn.
+\section{Kernel quantile regression}
+Kernel quantile regression had no previously implemented open source library, thus the need of implementing our own version.
+\\
+The scikit-learn team provides a project template for the creation of estimators compatible with scikit-learn functionalities.
+Therefore, the KQR class is derived from the scikit-learn BaseEstimator and the mixin class RegressorMixin.
+Our KQR class is initialized by providing a quantile, the regularization term C and the rbf kernel bandwidth gamma; the latter are the two hyperparameters of our custom estimator.
+\\
+In the fit method, we set up and solve the convex optimisation problem by using the cvxopt library. 
+When using this library, it is important to keep two things in mind.
+First this library assumes the quadratic term of the optimisation problem to be multiplied by the 0.5 factor, thus we just have to provide the Q matrix with no 0.5 in front.
+Secondly, in order to specify multiple inequalities we have to stack them an provide as a unique matrix.
+\\
+Once a solution to the convex problem has been found, we create a mask for the support vectors of the estimator in order to estimate the constant term of our kernel quantile regressor.
+\\
+In the predict method, we pass a matrix X\_eval of independent variables, next we compute the kernel matrix between X\_train and X\_eval and obtain y\_eval with the formula $y=\alpha^\intercal K+b$.
+\\
+This estimator is compatible with useful scikit-learn in built methods like gridsearch, crossvalidation and scoring rules.
+\\
+Up to now, there are only two open source implementation of the quantile kernel regression.
+Nevertheless they are both in R, that is there exists no python, matlab or julia open source implementation.
+Following are reported the results of a comparative study between our own implementation and the one of the R library kernlab(i guess it is in C or C++ and then binded to R).

### 2024-03-20 thesis/input_files/probabilistic.tex
March 20th 2024

@@
 \section{Quantile forest}
+Meinshausen \cite{meinshausen2006quantile} extends the idea of random forest \cite{breiman2001random} generalising it, the result is the quantile forest algorithm.
+Quantile forest allows us to estimate conditional quantiles in a non parametric fashion.
+\\
+In order to understand this algorithm it is necessary to first cover the theory of decision trees and random forests.
+- decision trees
+- random forest
+- quantile forest
 \section{Quantile gradient boosting machine}
@@
 The idea of quantile regression has been extended to kernel methods by Takeuchi et al. \cite{takeuchi2006nonparametric}.
-There they minimize a risk plus regularizer defined as follows.
+There, they minimize a risk plus regularizer defined as follows.
 \begin{equation}\label{eq:kqr_min1}
-    R[f]:=\frac{1}{m}\sum\limits_i=1^{m}\rho_q(y_i-f(x_i))+\frac{\lambda}{2}\|g\|_H^2
+    R[f]:=\frac{1}{m}\sum\limits_{i=1}^{m}\rho_q(y_i-f(x_i))+\frac{\lambda}{2}\|g\|_H^2
 \end{equation}
@@
 Using the link between RKHS and feature spaces, we can rewrite $f(x)=\langle w, \phi(x) \rangle+b$.
-Moreover, note that the RKHS norm is defined as follows $\|f\|_{\varphi }=\inf\{\|w\|_{F}:w\in F,f(x)=\langle w,\varphi (x)\rangle _{F},\forall x\in X\}$, where $F$ is the feature space.
+Moreover, note that the RKHS norm is defined as follows $\|f\|_{H}=\inf\{\|w\|_{F}:w\in F,f(x)=\langle w,\varphi (x)\rangle _{F},\forall x\in X\}$, where $F$ is the feature space.
  Doing so we obtain a minimization problem equivalent to minimizing equation \ref{eq:kqr_min1}.
@@
     \min_{w,b} \quad & C \sum \limits_{i=1}^{m}
-    (y_i-\langle w, \phi(x) \rangle-b)+ (1-q)(-y_i+\langle w, \phi(x) \rangle+b)+ \frac{1}{2}\|w\|^2\\
+    q(y_i-\langle w, \phi(x) \rangle-b)+ (1-q)(-y_i+\langle w, \phi(x) \rangle+b)+ \frac{1}{2}\|w\|^2\\
     \end{aligned}
     \end{equation}
-Note the division by $\lambda$ so that $C=\frac{1}{\lambda m}$
+Note the division by $\lambda$ so that $C=\frac{1}{\lambda m}$.
+\\
 We can next rephrase the optimisation in \ref{eq:kqr_min2} by introducing the slack variables $\xi_i$ and $\xi_i^*$.
@@
     \begin{aligned}
-        \min_{w,b\xi_i,\xi_i^*} \quad & C \sum \limits_{i=1}^{m}
+        \min_{w,b,\xi_i,\xi_i^*} \quad & C \sum \limits_{i=1}^{m}
         q \xi_i+ (1-q)\xi_i^*+ \frac{1}{2}\|w\|^2\\
-    \textrm{s.t.} \quad & y_i-\langle w, \phi(x) \rangle+b \leq \xi_i\\
+    \textrm{s.t.} \quad & y_i-\langle w, \phi(x) \rangle-b \leq \xi_i\\
     & -y_i+\langle w, \phi(x) \rangle+b \leq \xi_i^*\\
@@
     \end{equation}
-Consider now the associated lagrangian $L$ to \ref{eq:kqr_min4}
+Consider now, the lagrangian $L$ associated to \ref{eq:kqr_min4}
 \begin{equation}\label{eq:kqr_min5}
@@
     L(w,b,\xi,\xi^*)= & C q \xi^\intercal \mathbb{1}+ C q (\xi^*)^\intercal \mathbb{1}+ \frac{1}{2}w^\intercal w- \alpha^\intercal(\xi - y+\Phi^\intercal w +b)
-    - (\alpha^*)^\intercal(\xi^* -y+\Phi^\intercal w +b)
+    - (\alpha^*)^\intercal(\xi^* +y-\Phi^\intercal w -b)
     \\
@@
 
-- dual problem
 \begin{definition}
@@
 Consider a simple lagrangian $L(x,\lambda)=f(x)+\sum \lambda_i r_i(x) +\sum v_i h_i(x)$, where $r_i(x)$ are inequality constraints while $h_i(x)$ are equality constraints of the problem.
-What we can be noted is that the lower bound on $p^*$ is non trivial only when the lagrange multiplier $\lambda \succeq 0$.
-Therefore, the idea is that by maximizing the dual function subject to the contraint $\lambda \succeq 0$ we can obtain a approximate or perfect solution to the primal problem.
+Then it can be noted that the lower bound on $p^*$ is non trivial only when the lagrange multiplier $\lambda \succeq 0$.
+Therefore, the idea is that by maximizing the dual function subject to the contraint $\lambda \succeq 0$, we can obtain an approximate or perfect solution to the primal problem.
 \\
@@
 We use $d^*$ to denote the optimal value of the lagrange dual problem; we can think of it as the best lower bound on $p^*$. 
+\\
 The inequality \ref{weak_duality} is called weak duality.
@@
 \\
-Constraint qualifications are condtions under which strong duality holds; one the most popular is Slater's condition.
+Constraint qualifications are condtions under which strong duality holds; one of the most popular is Slater's condition.
 \begin{equation}\label{slater_condition}
@@
 \\
-In our case we don't have equlity constraint, so we do not have the $h_i(x)=0$ \ref{slater_condition} to be checked.
+In our case we don't have any equality constraint, so we do not have to worry about the $h_i(x)=0$ term in \ref{slater_condition}.
 All we have to check is the convexity of our problem and that there exist a $x$ such that $r_i(x)<0$.
@@
     \begin{aligned}
-        \min & x^\intercal Q x+ c^\intercal x \\        
-        s.t & Ax\preceq b
+        \min \quad & x^\intercal Q x+ c^\intercal x \\        
+        s.t \quad& Ax\preceq b
     \end{aligned}
 \end{equation}
-In our case \ref{eq:kqr_min5}, we have $w^\intercal w$, thus $Q$ is just the identity matrix which satisfies the positive definiteness requirement. Therefore our problem is convex.
-Next we check Slater's condition holds.
-Considering first the non negative constraints on $\xi$ and $\xi^*$, we conclude that $\xi$  and $\xi^*$ have to be greater or equal to zero for the existence of an $x$ satisfying Slater's condition.
+In our case \ref{eq:kqr_min5}, we have $w^\intercal w$, thus $Q$ is just the identity matrix which satisfies the positive definiteness requirement. Therefore, our problem is convex.
+Next we check that Slater's condition holds.
+Considering first the two non negative constraints on $\xi$ and $\xi^*$, we conclude that $\xi$  and $\xi^*$ have to be greater or equal to zero for the existence of an $x$ satisfying Slater's condition.
 Thus, let us suppose that $0 \leq \xi \leq \alpha$ and $0 \leq \xi^* \leq \alpha$.
@@
         w=& \Phi^\intercal(\Phi \Phi^\intercal)^{-1} y\\
-        b\leq& \alpha
+        b<& \alpha
     \end{aligned}
@@
 \end{equation}
-Setting the derivates to zero
+Setting its derivates to zero
 \begin{equation}\label{eq:lagrange_derivatives}
@@
     C(q-1)\mathbb{1}\preceq \alpha \preceq Cq\mathbb{1}\\
-    &\alpha^\intercal\mathbb{1}
+    &\alpha^\intercal\mathbb{1}=1
     \end{aligned}
     \end{equation}
-Switching sign we rephrase as a minisation problem, which is the common practice in convex optimisation.
+Switching sign, we rephrase it as a minisation problem, which is the common practice in convex optimisation.
 \begin{equation}\label{eq:kqr_min7}
@@
     C(q-1)\mathbb{1}\preceq \alpha \preceq Cq\mathbb{1}\\
-    &\alpha^\intercal\mathbb{1}
+    &\alpha^\intercal\mathbb{1}=1
     \end{aligned}
@@
     b=&\frac{1}{I_q}\sum\limits_{i\in I_q} b_i, \\ 
-    \textrm{with} \ b_i=& y_i-\sum\limits_i \alpha_i k(x_i, x_i) \\
+    \textrm{with} \ b_i=& y_i-\sum\limits_k \alpha_k k(x_k, x_i) \\
      \textrm{and} \ I_q=& \{i=1,\dots,n|0<\alpha_i<Cq, 0<\alpha_i<C(1-q)\}
@@
 %contribution is implementation kernel quantile regression for pythos users since there exists only r version
-- calculations
 \subsection{Kernel herding}

### 2024-03-22 thesis/input_files/data_exploratory.tex
6 march
decision trees

@@
 
+- correlation between temperatures and load
+
 - Correlation and auto correlation plots
@@
 not generalize well.
+
+This section covers the datasets used in our experiments, attributes and features will be thoroughly described.
+Next, we will explain the ETL pipeline that we set up in order ease the workflow of our comparison studies.
+Finally, in order to to bettern understand patters within the data, detect outliers and discover interesting patterns we carry out an exploratory data analysis.
+
+\section{GEFCom2014}
+The chosen dataset for probabilistic load forecasting, was the data from the GEFCom 2014 competition. The data is freely hosted on Dr. Hong blog \cite{hong2016probabilistic}.
+Contestant were asked to provide one month ahead hourly probabilistic forecasts on a rolling basis for 15 consecutive round.
+To get started, in the first round, organizers provided 69 months of hourly load data and 117 months of hourly temperature data.
+As the competition task went on, the real observed data of the previous tasks was made available.
+The GEFCom2014 Data contains five zip files.
+We are interested in the probabilistic load forecasting data, which is contained in the GEFCom2014-L\_V2 zip file.
+Within this subfolder, we a txt with the competition instructions and 15 folders one for each of the consecutive tasks.
+For each of those we have the prediction of the competition benchmark model and the train file upon which we will fit our model.
+\\
+Each data from the various task folders are shifted by one months between others.
+\\
+In order to compare our model performance with the winning entries of the GEFCom2014 competition we will refer to the Provisional\_Leaderboard\_V2 file contained also in the GEFCom2014 Data directory.
+Notice a quick premise, within this xsls file the authors opted for various scores to compare the entries performance.
+This choice was motivated by the need
+to give a proportial weight to each task when averaging between them; that is in order to not penalise too much the tasks with higher point spreads.
+Furthermore, additional factors such as frequency of benchmark outperformance, and robustness and clarity in methodology explaination.
+\\
+Since we are only interested in the load forecasting track, we can directly compare between models by comparing their pinball scores.
+Such scores are stored inside the subtab L-score-0/L-score-2 of Provisional\_Leaderboard\_V2.

### 2024-03-22 thesis/input_files/probabilistic.tex
decision trees

@@
 \input{input_files/qr.tex}
-
-\section{Kernel density estimation}
-\input{input_files/kde.tex}
-
-\section{Ensemble methods}
-\input{input_files/ensemble.tex}
 
@@
 In order to understand this algorithm it is necessary to first cover the theory of decision trees and random forests.
-- decision trees
-- random forest
+\subsection{Decision trees}
+Decision trees methods partition recursively the feature space in a set of binary rectangles and then fit a simple model in each of those partitions (the most straightforward is fitting just a constant).
+To get started, we first split the space into two disjoint regions, then we model the response variable by the mean of the observed predicted variables with associated features falling in that specific region.
+Our goal is selecting the best features and best split point such that to achieve the best generalizing fit.
+For an example/visualisation consider the pictures \ref{fig:elements_statistical_learning1} and \ref{fig:elements_statistical_learning2} taken from \cite{hastie2009elements}, where the decision tree algorithm is visualised for a regression problem with two independent variables X2 and X1.
+\begin{figure}
+    \includegraphics[width=0.5\textwidth]{images/elsii1.png}
+    \caption{two-dimensional feature space partitioned by recursive binary splittin}
+    \label{fig:elements_statistical_learning1}
+  \end{figure}
+\\
+\begin{figure}
+\includegraphics[width=1\textwidth]{images/elsii2.png}
+\caption{partition tree and regression model}
+\label{fig:elements_statistical_learning2}
+\end{figure}
+\\
+Suppose to partition the feature space into M regions $R_1,\dots,R_M$, then the model reads as follows.
+\begin{equation}
+    f(x)=\sum\limits_{m=1}^{M}c_m\mathbb{1}(x \in R_m)
+\end{equation}
+\\
+It follows that the function minimising the sum of squares is the one with $\hat{c}_m=mean(y_i|x_i \in R_m)$
+Finding the best split in terms of minimum sum of squares is computationally infeasible in practice. Therefore we approximate a solution by approaching the problem in a greedy fashion.
+Let $j$ denote the splitting variable, $s$ be the the split point we define the two half planes
+\begin{equation}
+    R_1(j,s)=\{X|X_j\leq s\} \quad R_2(j,s)=\{X|X_j\>s\}
+\end{equation}
+Then we search for the $s$ and $j$ that solve
+\begin{equation}
+    \min_{j,s}\left[\min_{c_1} \sum\limits_{x_i \in R_1(j,s)}(y_i-c_1)^2+\min_{c_2} \sum\limits_{x_i \in R_2(j,s)}(y_i-c_2)^2\right]
+\end{equation}
+The inner problem is easy, as already pointed out we will have 
+\begin{equation}
+    \hat{c}_1=mean(y_i|x_i \in R_1) \quad \hat{c}_2=mean(y_i|x_i \in R_2)
+\end{equation}
+For the outer problem, we scan through all the (j,s) tuples and pick the best pair. 
+Next, one or both of these regions from the previous step are split into two more regions; we recurse this process until some stopping condition is triggered (max number of branches, max deth of tree, threshold on the MSE, minimum number of observation in each leaf node).
+Notice, this being a greedy algorithm implies that our final solution is guaranteed to be just a local optimum not a global one.
+Even though their simplicity, these models have proved themselves to be really powerful, see figure \ref{fig:decision_tree} for a visualisaiton .
+The most popular among these model is the CART \cite{breiman2017classification} tree, its name comes from the fact that it can handle both classification and regression problems.
+\begin{figure}
+    \includegraphics[width=\textwidth]{images/decision_tree.png}
+    \caption{decision tree regression}
+    \label{fig:decision_tree}
+    % image generated on coolab to be quick
+\end{figure}
+
+\subsection{Random forest}
 - quantile forest
+
 \section{Quantile gradient boosting machine}
-
-\section{DMLP}
-%gluon nn
-% train a beta distribution to learn its parameters
-\section{DeepAR}
 
@@
 %contribution is implementation kernel quantile regression for pythos users since there exists only r version
+\subsection{Weather quantiles}
+In order to get acquainted with the inner working of the presented methods, this section covers an application explaining practical details and comparing results.
+\\
+The dataset used is the Melbourne Daily Maximum Temperatures \cite{hdrcde}. It contains the daily maximum temperatures in Melbourne, Australia, from 1981-1990, excluding leap days.
+Due to the bimodality of the data, such dataset is commonly used to give a difficult quantile regression problem \cite{hyndman1996estimating}, thus why we chose it.
+The observed bimodality is that a hot day is likely to be followed by either an equally hot day or one much cooler.
+\\
+
+- metrics pinball, picw plot, mae error
+
+\section{Kernel density estimation}
+\input{input_files/kde.tex}
+
+\section{Ensemble methods}
+\input{input_files/ensemble.tex}
+
+\section{DMLP}
+%gluon nn
+% train a beta distribution to learn its parameters
+\section{DeepAR}
+
 \subsection{Kernel herding}

### 2024-03-22 dataset_applications/kqr.py

@@ -106,7 +106,6 @@ class KQR(RegressorMixin, BaseEstimator):
         sol = qp(P=K,q=-r,G=G,h=h,A=A,b=b)
         # alpha solution
         self.a=np.array(sol["x"]).flatten()
-        
         # see coefficients
         # print("coefficients a: ",self.a)
 
@@ -114,19 +113,14 @@ class KQR(RegressorMixin, BaseEstimator):
         # print("coefficients sum up to 1:", np.sum(self.a))
         
         # condition, index set of support vectors
-        squared_diff = (np.round(self.a, 3) - (self.C * self.alpha))**2 + (np.round(self.a, 3) - (self.C * (self.alpha - 1)))**2
+        squared_diff = (self.a - (self.C * self.alpha))**2 + (self.a - (self.C * (self.alpha - 1)))**2
 
-        # Find the index corresponding to the smallest squared difference
+        # get the smallest squared difference
         offshift = int(np.argmin(squared_diff))
-
-        # Compute the bias term b
-        print(offshift)
+        # print(offshift)
+        
+        # calculate bias term b
         self.b = y[offshift] - self.a.T@K[:,offshift]
-        # idx_arr=[i for i, ai in enumerate(self.a) if (ai>self.C*(self.alpha-1)) and (ai<self.C*(self.alpha))]
-        # # beta of the model
-        # self.b_i=[y[i]- self.a.T@K[:,idx_arr] for i in idx_arr]
-        # print("beta i: ", y[idx_arr]-self.a.T@K[:,idx_arr])
-        # self.b=np.mean(self.b_i)
         # print("beta mean: ", self.b)
         
         # Return the regressor

### 2024-03-22 dataset_applications/qr_melbourne.py

@@ -34,7 +34,7 @@ if __name__=="__main__":
     # train, test split
     X_train, X_test, y_train, y_test = train_test_split(df["Yt-1"], df["Yt"], test_size=0.20, random_state=4)
 
-    eval_set=np.linspace(df["Yt"].min(), df["Yt"].max(), 100).T
+    eval_set=np.linspace(df["Yt-1"].min(), df["Yt-1"].max(), 100).T
     eval_set=eval_set.reshape(-1,1)
     X_train= X_train.values.reshape(-1, 1)
     X_test = X_test.values.reshape(-1, 1)

### 2024-03-22 thesis/input_files/probabilistic.tex
March 22nd 2024

@@
 
+\subsection{Bootstrap}
+The idea behind boostrapping is to randomly sample from the training set with replacement B times and then fitting B models to each of the "artificial" datasets.
+Boostrapping can serve different tasks; we can use it to assess the accuracy of a parameter estimate or of a prediction but also to improve their estimates.
+
+\subsection{Bagging}
+Bagging stands for boostrap aggregation. 
+The bagging estimate is dafined by $\mathbb{E}_P\hat{f}^*$ where $P$ is the empirical distribution putting equal probability on each data point of the training set.
+Basically, for each bootstrap fitted model $\hat{f}^{*b}(x)$, we compute the bagging estimate by
+\begin{equation}
+    \hat{f}_{bag}(x)=\frac{1}{B}\sum\limits_{b=1}^{B}\hat{f}^{*b}(x)
+\end{equation}
+Bagging is particularly useful in reducing the variance of decision trees, resulting in an improved prediction(tradeoff bias variance).
+Note, this improvement comes from the fact that averaging reduces variance and leaves biases unchanged.
+
 \subsection{Random forest}
-- quantile forest
+Decision trees are characterised by high variance and low bias, thus, they can benefit extremely from bagging.
+Furthermore, every decision tree generated through bagging will be identically distributed (i.d.), thus the expectation of an average of B trees is probabilistically equivalent to the expectation of any such trees.
+As a consequence, the bias will stay fixed since the bias of the bagged estimator is the same as that of each individual tree.
+\\
+Consider positively correlated i.d. random variables, then the variance of their average is 
+\begin{equation}
+    \rho \sigma^2+\frac{1-\rho}{B}\sigma^2
+\end{equation}
+The second term disappears as B increases, while the first term depends heavily on the correlation between bagged trees.
+Random forest consists in is reducing the correlation between the trees by randomly selecting m of the p features as candidates for splitting; m tipically takes value of the order $\sqrt{p}$ or even 1 with default value $m=\left\lfloor \frac{p}{3} \right\rfloor$ while a good minimum node size is around five.
+\\
+Letting $T(x;\Theta_b)$ be the bth bagged tree where $\Theta_b$ denotes the randomness characterizing its splits, cutpoints and terminal node values, we have that the random forest regressor is given by
+\begin{equation}
+    \hat{f}_{rf}^{B}=\frac{1}{B}\sum\limits_{b=1}^{B}T(x;\Theta_b)
+\end{equation}
+For a simple visualisation compare figure \ref{fig:random_forest} with \ref{fig:decision_tree}
+\begin{figure}
+    \includegraphics[width=\textwidth]{images/random_forest.png}
+    \caption{random forest regression}
+    \label{fig:random_forest}
+    % image generated on coolab to be quick
+\end{figure}
+
+\subsection{Quantile forest}
+The key observation here is noting that random forests approximates the conditional mean $\mathbb{E}(Y|X=x)$ by a weighted average over the observed $Y$.
+Hence, we can extend this idea to the full conditional distribution by
+\begin{equation}
+    F(y|X=x)=P(Y\leq y|X=x)=\mathbb{E}(\mathbb{1}_{\{Y\leq y\}}|X=x)
+\end{equation}
+All we have to do is approximating $\mathbb{E}(\mathbb{1}_{\{Y\leq y\}}|X=x)$ by a weighted mean over the random variable $\mathbb{1}_{Y\leq y}$
+\begin{equation}
+    \hat{F}(y|X=x)=\sum\limits_{i=1}^{n}\omega_i(x)\mathbb{1}_{\{Y_i\leq y\}}
+\end{equation}
 

### 2024-03-23 thesis/input_files/probabilistic.tex
gradient boosting

@@
 Notice, this being a greedy algorithm implies that our final solution is guaranteed to be just a local optimum not a global one.
-Even though their simplicity, these models have proved themselves to be really powerful, see figure \ref{fig:decision_tree} for a visualisaiton .
+Even though their simplicity, these models have proved themselves to be really powerful.
+See figure \ref{fig:decision_tree} for a example where decision tree with different hyperparameters are fitted to a sine wave plus noise.
 The most popular among these model is the CART \cite{breiman2017classification} tree, its name comes from the fact that it can handle both classification and regression problems.
@@
 \end{equation}
-All we have to do is approximating $\mathbb{E}(\mathbb{1}_{\{Y\leq y\}}|X=x)$ by a weighted mean over the random variable $\mathbb{1}_{Y\leq y}$
+All we have to do is approximating $\mathbb{E}(\mathbb{1}_{\{Y\leq y\}}|X=x)$ by a weighted mean over the random variable $\mathbb{1}_{\{Y\leq y\}}$
 \begin{equation}
@@
 \end{equation}
+By swapping $F(y|X=x)$ with $\hat{F}(y|X=x)$ in the defintion of conditional quantiles we obtain their respective random forest estimator
+\begin{equation}
+    \hat{Q}_\alpha=\inf\{y:\hat{F}(y|X=x)\geq \alpha\}
+\end{equation}
 
 \section{Quantile gradient boosting machine}
+\subsection{Boosting}
+With boosting we fit an additive expansion of elementary basis functions; with M basis functions we have 
+\begin{equation}
+    f(x)=\sum\limits_{m=1}^{M}\beta_m b(x;\gamma_m)
+\end{equation}
+where $b(x;\gamma)$ are the basis functions while $\beta_m$ are the coefficients of the expansion.
+Boosted models are fitted by minimizing a loss function over the training data
+\begin{equation}
+    \min_{\beta_m, \gamma_m}\sum\limits_{i=1}^{N}L\left(y_i, \sum\limits_{m=1}^M \beta_m b(x_i;\gamma_m)\right)
+\end{equation}
+However, such problem is highly intensive in terms of computation.
+Therefore, what is done in the literature is approximating its solution by iteratively adding new basis function to the current expansion.
+That is, we construct $f_m$ by solving for the optimal basis function and coefficients to add to $f_{m-1}$.
+Considering the square loss, we would have
+\begin{equation}
+    L(y_i, f_{m-1}(x_i)+\beta_m b(x_i;\gamma))=(e_{im}-\beta_m b(x_i;\gamma))^2
+\end{equation}
+
+\subsection{Boosted trees}
+We combine several trees obtaining the boosted tree model
+\begin{equation}
+    f_M(x)=\sum\limits_{m=1}^M T(x;\Theta_m)
+\end{equation}
+Thus, at each step of the iterative optimisation procedure we have to solve
+\begin{equation}\label{boosting_minimisation_problem}
+    \hat{\Theta}_m=\argmin{\Theta_m}\sum\limits_{i=1}^{N}L(y_i, f_{m-1}(x_i)+T(x_i;\Theta_m))
+\end{equation}
+Remember, $\Theta_m$ referes to the parameters of the mth tree, $\Theta_m=\{R_{jm}, \gamma_{jm}\}_1^{J_m}$
+\subsection{Gradient boosting}
+In order to robustly solve \ref{boosting_minimisation_problem}, gradient boosting considers the following problem
+\begin{equation}
+    \tilde{\Theta_m}=\argmin{\Theta}\sum\limits_{i=1}^{N}\left(-g_{im}-T(x_i;\Theta)\right)^2
+\end{equation}
+where $g_{im}$ is the gradient of $L(f)=\sum\limits_{i=1}^N L(y_i, f(x_i))$ evaluated at $f_{m-1}$.
+Put simply, we are fitting the mth tree to the negative of the gradient values of f through least squares.
+In order to solve our quantile regression tasks through gradient boosting, all we need to do is specifying the pinball loss as our criterion to guide the minimisation algorithm.
+See figure \ref{fig:gradient_boosting} for a visualisation
+\begin{figure}
+    \includegraphics[width=\textwidth]{images/gradient_boosting.png}
+    \caption{gradient boosting regression}
+    \label{fig:gradient_boosting}
+    % image generated on coolab to be quick
+\end{figure}
 
@@
 \\
-We end this section with the derivation of the dual problem; that is the convex problem we will solve in order to get the qunatiles prdeiction of our quantile kernel algorithm.
+We end this section with the derivation of the dual problem; that is the convex problem we will solve in order to get the quantiles prdeiction of our quantile kernel algorithm.
 \\
@@
     \end{equation}
-Switching sign, we rephrase it as a minisation problem, which is the common practice in convex optimisation.
+Switching sign, we rephrase it as a minimisation problem, which is the common practice in convex optimisation.
 \begin{equation}\label{eq:kqr_min7}
@@
 \subsection{Weather quantiles}
-In order to get acquainted with the inner working of the presented methods, this section covers an application explaining practical details and comparing results.
+In order to get acquainted with the inner workings of the presented methods, this section covers an application explaining practical details and comparing results.
 \\
@@
 
-- metrics pinball, picw plot, mae error
+- metrics pinball, picw plot, mae error on test data
 

### 2024-03-23 thesis/appendix.tex
March 23rd 2024

@@
 
+\section{Cross validation}\label{appendix:cross_validation}
+
+\section{Quantile regressor extensive comparison}\label{appendix:quantile_regressor_extensive_comparison}
+\subsection{Boston housing dataset}
+The Boston housing dataset \href{https://www.kaggle.com/datasets/altavish/boston-housing-dataset}{https://www.kaggle.com/datasets/altavish/boston-housing-dataset} contains information about various attributes for suburbs in Boston.
+There are 13 indipendent variables:
+\begin{itemize}
+\item CRIM per capita crime rate by town
+\item ZN proportion of residential land zoned for lots over 25,000 sq.ft.
+\item INDUS proportion of non-retail business acres per town.
+\item CHAS Charles River dummy variable (1 if tract bounds river; 0 otherwise)
+\item NOX nitric oxides concentration (parts per 10 million)
+\item RM average number of rooms per dwelling
+\item AGE proportion of owner-occupied units built prior to 1940
+\item DIS weighted distances to five Boston employment centres
+\item RAD index of accessibility to radial highways
+\item TAX full-value property-tax rate per 10,000
+\item PTRATIO pupil-teacher ratio by town
+\item B $1000(Bk - 0.63)^2$ where Bk is the proportion of afroamericans by town
+\item LSTAT lower status of the population
+\end{itemize}
+The dependent variable is MEDV, that is the median value of owner occupied homes in \$1000's
+
+\begin{table}
+\caption{Pinball loss Boston housing data}
+\begin{tabular}{lllll}
+\toprule
+    & Linear qr & Gbm qr & Quantile forest & Kernel qr \\
+\midrule
+0 & 13.785678 & 11.418540 & 10.587686 & 10.297572 \\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{table}
+    \caption{Pinball loss quantile-wise Boston data}
+\begin{tabular}{lllll}
+\toprule
+    & Linear qr & Gbm qr & Quantile forest & Kernel qr \\
+\midrule
+0.100000 & 0.729749 & 0.771714 & 0.588441 & 0.578898 \\
+0.200000 & 1.122582 & 1.033442 & 0.932824 & 0.869145 \\
+0.300000 & 1.479486 & 1.170642 & 1.153765 & 1.142783 \\
+0.400000 & 1.712577 & 1.436263 & 1.352667 & 1.331955 \\
+0.500000 & 1.911385 & 1.344361 & 1.408333 & 1.396300 \\
+0.600000 & 1.989514 & 1.448885 & 1.464902 & 1.431705 \\
+0.700000 & 1.938362 & 1.508741 & 1.427912 & 1.382772 \\
+0.800000 & 1.658058 & 1.497901 & 1.275059 & 1.245288 \\
+0.900000 & 1.243965 & 1.206591 & 0.983784 & 0.918725 \\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{table}
+\caption{MAE Boston data}    
+\begin{tabular}{lllll}
+\toprule
+    & Linear qr & Gbm qr & Quantile forest & Kernel qr \\
+\midrule
+0 & 3.826326 & 2.845989 & 2.965490 & 2.810494 \\
+\bottomrule
+\end{tabular}
+
+\end{table}
+
+\subsection{Abalone dataset}
+The abalone data \href{https://archive.ics.uci.edu/dataset/1/abalone}{https://archive.ics.uci.edu/dataset/1/abalone} consist of measurements of abalone molluscs, the goal is predicting their age by building a model for estimating its number of rings; age is the number of rings plus 1.5
+The data has 8 attributes:
+\begin{itemize}
+    \item Sex Categorical variable either male, female or infant
+    \item Length
+    \item Diameter
+    \item Height
+    \item Whole height
+    \item Shucked height
+    \item Viscera weight
+    \item Shell weight
+\end{itemize}
+
+\begin{table}
+    \caption{Pinball loss Abalone data}
+\begin{tabular}{lllll}
+    \toprule
+     & Linear qr & Gbm qr & Quantile forest & Kernel qr \\
+    \midrule
+    0 & 5.613975 & 5.531938 & 5.212990 & 5.252491 \\
+    \bottomrule
+    \end{tabular}
+\end{table}
+
+\begin{table}
+    \caption{Pinball loss quantile-wise Abalone data}
+    \begin{tabular}{lllll}
+    \toprule
+     & Linear qr & Gbm qr & Quantile forest & Kernel qr \\
+    \midrule
+    0.100000 & 0.277903 & 0.290531 & 0.274079 & 0.269287 \\
+    0.200000 & 0.469361 & 0.488079 & 0.453110 & 0.457286 \\
+    0.300000 & 0.621961 & 0.625633 & 0.580766 & 0.596791 \\
+    0.400000 & 0.729757 & 0.715875 & 0.689904 & 0.691310 \\
+    0.500000 & 0.794695 & 0.766185 & 0.735945 & 0.740834 \\
+    0.600000 & 0.810691 & 0.785769 & 0.744928 & 0.746636 \\
+    0.700000 & 0.769587 & 0.730318 & 0.700287 & 0.710392 \\
+    0.800000 & 0.667776 & 0.656913 & 0.609378 & 0.608334 \\
+    0.900000 & 0.472244 & 0.472635 & 0.424593 & 0.431621 \\
+    \bottomrule
+    \end{tabular}
+\end{table}
+
+\begin{table}
+    \caption{MAE Abalone data}
+    \begin{tabular}{lllll}
+    \toprule
+     & Linear qr & Gbm qr & Quantile forest & Kernel qr \\
+    \midrule
+    0 & 1.627627 & 1.574179 & 1.499522 & 1.498583 \\
+    \bottomrule
+    \end{tabular}
+\end{table}
+
+\subsection{Vehicle dataset}
+This data contains info about used cars \href{https://www.kaggle.com/datasets/nehalbirla/vehicle-dataset-from-cardekho}{https://www.kaggle.com/datasets/nehalbirla/vehicle-dataset-from-cardekho}, the predictors are:
+\begin{itemize}
+    \item Year
+    \item Present\_price ex showroom price
+    \item Kms Driven
+    \item Fuel type
+    \item Seller type
+    \item Transmission
+    \item Owner number of previous owners
+\end{itemize}
+The dependent variable is the selling price.
+
+\begin{table}
+\caption{Pinball loss Vehicle data}
+\begin{tabular}{lllll}
+    \toprule
+     & Linear qr & Gbm qr & Quantile forest & Kernel qr \\
+    \midrule
+    0 & 4.054449 & 2.289554 & 2.844410 & 2.204343 \\
+    \bottomrule
+    \end{tabular}
+\end{table}
+
+\begin{table}
+    \caption{Pinball loss quantile-wise Vehicle data}
+    \begin{tabular}{lllll}
+    \toprule
+     & Linear qr & Gbm qr & Quantile forest & Kernel qr \\
+    \midrule
+    0.100000 & 0.254649 & 0.139849 & 0.170489 & 0.182490 \\
+    0.200000 & 0.403772 & 0.236339 & 0.285875 & 0.223165 \\
+    0.300000 & 0.548820 & 0.244086 & 0.357375 & 0.242963 \\
+    0.400000 & 0.576918 & 0.263169 & 0.389305 & 0.262835 \\
+    0.500000 & 0.554367 & 0.306123 & 0.410738 & 0.295878 \\
+    0.600000 & 0.563046 & 0.335363 & 0.398125 & 0.306062 \\
+    0.700000 & 0.516019 & 0.287490 & 0.326572 & 0.283716 \\
+    0.800000 & 0.407742 & 0.261882 & 0.313698 & 0.235793 \\
+    0.900000 & 0.229115 & 0.215253 & 0.192233 & 0.171440 \\
+    \bottomrule
+    \end{tabular}
+\end{table}
+
+\begin{table}
+    \caption{MAE Vehicle data}
+    \begin{tabular}{lllll}
+    \toprule
+     & Linear qr & Gbm qr & Quantile forest & Kernel qr \\
+    \midrule
+    0 & 1.117714 & 0.606971 & 0.752197 & 0.594292 \\
+    \bottomrule
+    \end{tabular}
+\end{table}
+
 \newpage

### 2024-03-23 thesis/input_files/probabilistic.tex
March 23rd 2024

@@
 \\
-The dataset used is the Melbourne Daily Maximum Temperatures \cite{hdrcde}. It contains the daily maximum temperatures in Melbourne, Australia, from 1981-1990, excluding leap days.
+The dataset used is the Melbourne Daily Maximum Temperatures \cite{hdrcde}.
+It contains the daily maximum temperatures in Melbourne, Australia, from 1981-1990, excluding leap days, see figure \ref{fig:melbourne_temperature_data}.
 Due to the bimodality of the data, such dataset is commonly used to give a difficult quantile regression problem \cite{hyndman1996estimating}, thus why we chose it.
 The observed bimodality is that a hot day is likely to be followed by either an equally hot day or one much cooler.
-\\
+\begin{figure}
+    \includegraphics[width=\textwidth]{images/melbourne_temperature.png}
+    \caption{Melbourne temperatures dataset}
+    \label{fig:melbourne_temperature_data}
+\end{figure}
+Hereafter, the results of the four presented methods on the Melbourne dataset are reported, see figure \ref{fig:melbourne_quantiles_comparison} for a visualisation.
+Hyperparameter tuning have been optimised through cross validation, see \ref{appendix:cross_validation}.
+\begin{figure}
+    \begin{subfigure}[b]{0.5\linewidth}
+      \centering
+      \includegraphics[width=1.1\textwidth]{images/melbourne_linear_quantile_regression.png} 
+      \caption{Linear quantile regression} 
+      \label{fig:melbourne_linear_quantile_regression} 
+      \vspace{4ex}
+    \end{subfigure}%% 
+    \begin{subfigure}[b]{0.5\linewidth}
+      \centering
+      \includegraphics[width=1.1\textwidth]{images/melbourne_gradient_boosting_quantile_regression.png} 
+      \caption{Gradient boosting quantile regression} 
+      \label{fig:melbourne_gradient_boosting_quantile_regression} 
+      \vspace{4ex}
+    \end{subfigure} 
+    \begin{subfigure}[b]{0.5\linewidth}
+      \centering
+      \includegraphics[width=1.1\textwidth]{images/melbourne_quantile_forest.png} 
+      \caption{Quantile forest} 
+      \label{fig:melbourne_quantile_forest} 
+    \end{subfigure}%%
+    \begin{subfigure}[b]{0.5\linewidth}
+      \centering
+      \includegraphics[width=1.1\textwidth]{images/melborune_kernel_quantile_regression.png} 
+      \caption{Kernel quantile regression} 
+      \label{fig:melborune_kernel_quantile_regression} 
+    \end{subfigure} 
+    \caption{Quantile regressors}
+    \label{fig:melbourne_quantiles_comparison} 
+  \end{figure}
 
-- metrics pinball, picw plot, mae error on test data
+% tables
+\begin{table}
+\caption{Pinball loss Melbourne data}
+\begin{tabular}{lllll}
+    \toprule
+     & Linear qr & Gbm qr & Quantile forest & Kernel qr \\
+    \midrule
+    Pinball loss & 11.278895 & 10.317612 & 10.340842 & 10.031708 \\
+    \bottomrule
+    \end{tabular}
+\end{table}
+
+\begin{table}
+    \caption{Pinball loss quantile-wise Melbourne data}
+    \begin{tabular}{lllll}
+    \toprule
+    Quantiles & Linear qr & Gbm qr & Quantile forest & Kernel qr \\
+    \midrule
+    0.100000 & 0.710644 & 0.549232 & 0.562888 & 0.540235 \\
+    0.200000 & 1.155014 & 0.938561 & 0.946712 & 0.903213 \\
+    0.300000 & 1.417805 & 1.212671 & 1.222407 & 1.173803 \\
+    0.400000 & 1.540108 & 1.399925 & 1.409293 & 1.367943 \\
+    0.500000 & 1.574957 & 1.517281 & 1.484589 & 1.455849 \\
+    0.600000 & 1.525114 & 1.498608 & 1.495474 & 1.447470 \\
+    0.700000 & 1.397918 & 1.372183 & 1.362173 & 1.331503 \\
+    0.800000 & 1.170140 & 1.115077 & 1.123195 & 1.096606 \\
+    0.900000 & 0.787195 & 0.714075 & 0.734112 & 0.715085 \\
+    \bottomrule
+    \end{tabular}
+\end{table}
+
+As already pointed out, the quantile regression with q=0.5 corresponds to the standard regression problem, hence we can compare the proposed methods also in terms of the mean absolute error.
+
+\begin{table}
+\caption{MAE Melbourne data}
+\begin{tabular}{lllll}
+    \toprule
+     & Linear qr & Gbm qr & Quantile forest & Kernel qr \\
+    \midrule
+    MAE & 3.253882 & 3.134805 & 3.095041 & 3.024336 \\
+    \bottomrule
+    \end{tabular}
+\end{table}  
+From these tables, we can see that kernel quantile regression outperformed the simple quantile regressor as well as the more complex models like quantile forest and gradient boosting quantile regression for the Melbourne temperatures dataset.
+Not only kernel quantile regression was the best in terms of total pinball loss but also the best in terms of each quantile pinball loss and mean absolute error.
+Comparison has been carried out on further dataset, yielding similar conclusion as the one of above, to know more see \ref{appendix:quantile_regressor_extensive_comparison}.
 

### 2024-03-23 dataset_applications/kqr.py

@@ -6,6 +6,7 @@ from cvxopt import blas
  
 import matplotlib.pyplot as plt
 import numpy as np
+from numpy import asarray
 import pandas as pd
 from pprint import pprint
 from operator import itemgetter
@@ -14,15 +15,15 @@ from scipy.stats import uniform
 from sklearn.base import BaseEstimator, RegressorMixin
 
 from sklearn.experimental import enable_halving_search_cv
-from sklearn.model_selection import HalvingRandomSearchCV
-from sklearn.model_selection import GridSearchCV
 
-from sklearn.metrics.pairwise import rbf_kernel
 from sklearn.metrics import make_scorer
 from sklearn.metrics import mean_pinball_loss
+from sklearn.metrics.pairwise import rbf_kernel
 
+from sklearn.model_selection import HalvingRandomSearchCV
+from sklearn.model_selection import GridSearchCV
 from sklearn.model_selection import train_test_split
-from numpy import asarray
+
 from sklearn.preprocessing import StandardScaler
 from sklearn.preprocessing import MinMaxScaler
 from sklearn.preprocessing import RobustScaler
@@ -65,7 +66,7 @@ class KQR(RegressorMixin, BaseEstimator):
         self.alpha=alpha
 
     def fit(self, X, y):
-        """Implementation of a fitting function.
+        """Implementation of fitting function.
 
         Parameters
         ----------
@@ -87,9 +88,10 @@ class KQR(RegressorMixin, BaseEstimator):
         self.y_ = y
         # build convex optimisation problem
         K=rbf_kernel(self.X_, gamma=self.gamma)
-        # the 0.5 in front is taken into account by cvxopt library
+        # the 0.5 in front in the optimisation probelm is taken into account by cvxopt library
         K = matrix(K)
-        r=matrix(y)
+        # multiply by one to convert matrix items to float https://stackoverflow.com/questions/36510859/cvxopt-qp-solver-typeerror-a-must-be-a-d-matrix-with-1000-columns
+        r=matrix(y)* 1.0
         # equality constraint
         A = matrix(np.ones(y.size)).T
         b = matrix(1.0)
@@ -101,18 +103,18 @@ class KQR(RegressorMixin, BaseEstimator):
         # concatenate
         G = matrix([G1,G2])
         h = matrix([h1,-h2])
-
         # Solve
[…]

### 2024-03-23 dataset_applications/qr.py

@@ -15,41 +15,26 @@ from sklearn.metrics import make_scorer
 from sklearn.metrics import mean_absolute_error
 from sklearn.metrics import mean_pinball_loss
 from sklearn.model_selection import train_test_split
+from sklearn.preprocessing import StandardScaler
 
 import statsmodels.regression.quantile_regression as qr 
 import sys
 
 from tqdm import tqdm
 
-
-
-if __name__=="__main__":
-    # load data
-    df=pd.read_csv("/Users/luca/Desktop/ThesisKernelMethods/dataset_applications/temperatures_melbourne.csv", sep=";", decimal=".")
-
+from cvxopt import matrix, spmatrix, sparse
+def qr_tests(X,y,write_to_path):
     # quantiles
     quantiles = [0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9]
     # table for output results
     pinball_scores=pd.DataFrame(columns=["Linear qr", "Gbm qr", "Quantile forest", "Kernel qr"],index=quantiles)
     mae_scores=pd.DataFrame(columns=["Linear qr", "Gbm qr", "Quantile forest", "Kernel qr"])
 
-    # plot data
-    plt.plot(df["Yt-1"],df["Yt"],"o", alpha=0.2)
-    plt.show()
-
     # train, test split
-    X_train, X_test, y_train, y_test = train_test_split(df["Yt-1"], df["Yt"], test_size=0.20, random_state=0)
-
-    # equally space dataset for plotting
-    eval_set=np.linspace(df["Yt-1"].min(), df["Yt-1"].max(), 100).T
-
-    # reshape data to be predicted
-    X_train= X_train.values.reshape(-1, 1)
-    X_test = X_test.values.reshape(-1, 1)
-    eval_set=eval_set.reshape(-1,1)
+    X_train, X_test, y_train, y_test = train_test_split(X, y, test_size=0.20, random_state=0)
 
 
-    # quantile regression: linear, random forest, gbm
+    # # quantile regression: linear, random forest, gbm
 
     # linear quantile regression
     qr_models = [qr.QuantReg(y_train, X_train).fit(q=q) for q in quantiles]
@@ -60,14 +45,6 @@ if __name__=="__main__":
     for i, q in enumerate(quantiles):
[…]

### 2024-03-23 dataset_applications/qr_abalone.py

@@ -1,0 +1,17 @@
+import numpy as np
+import pandas as pd
+
+from qr import qr_tests
+
+if __name__=="__main__":
+    # load data
+    df=pd.read_csv("/Users/luca/Desktop/ThesisKernelMethods/dataset_applications/abalone.csv", sep=",", decimal=".")
+
+    # map categorical regressor to integer form
+    key = {"I":0, "M":1, "F":2}
+    df['Sex'] = df['Sex'].map(key)
+
+    X=df.iloc[:,0:-1]
+    y=df.iloc[:,-1]
+    
+    qr_tests(X,y,"qr_abalone")

### 2024-03-23 dataset_applications/qr_boston.py

@@ -1,0 +1,12 @@
+import numpy as np
+import pandas as pd
+
+from qr import qr_tests
+
+if __name__=="__main__":
+    # load data
+    df=pd.read_csv("/Users/luca/Desktop/ThesisKernelMethods/dataset_applications/BostonHousing.csv", sep=",", decimal=".")
+
+    X=df.iloc[:,0:-1]
+    y=df.iloc[:,-1]
+    qr_tests(X,y, "qr_boston")

### 2024-03-23 dataset_applications/qr_cars.py

@@ -1,0 +1,24 @@
+import numpy as np
+import pandas as pd
+
+from qr import qr_tests
+
+if __name__=="__main__":
+    # load data
+    df=pd.read_csv("/Users/luca/Desktop/ThesisKernelMethods/dataset_applications/car_data.csv", sep=",", decimal=".")
+
+    # map categorical regressor to integer form
+    key6 = {"Petrol":0, "Diesel":1, "CNG":2}
+    df['Fuel_Type'] = df['Fuel_Type'].map(key6)
+
+    key7 = {"Dealer":0, "Individual":1}
+    df['Seller_Type'] = df['Seller_Type'].map(key7)
+
+    key8 = {"Manual":0, "Automatic":1}
+    df['Transmission'] = df['Transmission'].map(key8)
+
+    X=df.iloc[:,[1,3,4,5,6,7,8]]
+    y=df.iloc[:,2]
+    print(X)
+    
+    qr_tests(X,y,"qr_cars")

### 2024-03-24 thesis/appendix.tex
March 24th 2024

@@
 
+\section{Cross validation}
+Cross validation directly estimates the expected test error
+\begin{equation}
+    Err=\mathbb{E}\left[L(Y,\hat{f}(X))\right]=\mathbb{E}\left[Err_{\tau}\right]
+\end{equation}
+$Err_{\tau}$ is the prediction error over an independent test sample
+\begin{equation}
+    \mathbb{E}_{\tau}=\mathbb{E}\left[L(Y,\hat{f}(X))| \tau \right]
+\end{equation}
+That is $Err$ is the average over everything that is random: X,Y and the training set $\tau$ used to generate $\hat{f}$.
+Hence, our interest lies in estimating the $Err$ quantity in order to guide model selection.
+\subsection{K-fold cross validation}
+K-fold cross validation splits the data into K roughly equal sized parts, then K models are trained.
+For k from 1 to K we train the kth model on the whole data except the kth partition.
+Next,  the prediction error of the kth fitted model when predicting the kth partition is computed.
+Finally averaging all the k prediction erros we obtain an estimate for its true value.
+We select the model which performs best in terms of expected prediction error.
+\\
+Usually, K is set equal to 5 or 10. Leave-one-out cross validation is the case when K is set equal to the size of the data.
+% REMEMBER: Parameters allow the model to learn the rules from the data while hyperparameters control how the model is training.
+Parameters learn their own values from data.
+In contrast, hyperparameters do not learn their values from data.
+We need to manually specify them before training the model
+\subsection{Gridsearch}
+Model performance depends higly on the choice of hyperparameters.
+Notice, there is no way to get to know them in advance, therefore, all we can do is trying a lot of combinations until we fit a good enough set of hyperparameters.
+Essentially, gridsearch carries out hyperparameter tuning by performing a search over a predefined hyperparemeters grid.
+During its search, it tries all possible combination and evaluates the different models %(different hyperparameters define different models)
+using cross validation. For example, supposing the grid contains 100 possible candidates and we are doing 5-fold cross-validation then we will be doing 500 iterations.
+Therefore, we get an estimate of prediction error for each considered model and this guides us in hyperparameters(model) selection.
+
+\subsection{Randomized search}
+Randomizedsearch controls the number of step by choosing smartly the hyperparameter to try in each iterations.
+Let us say there are 100 candidates and we set the number of iterations to 20 then the search will stop after the 20th iteration and return the best set among the hyperparameters observed.
+The advantage is that it is much quicker than gridsearch, but on the other hand its performance is worse than gridsearch.
+
+\subsection{Halving gridsearch}
+Gridsearch has been the to go choice for hyperparameter tuning for the past years.
+However, such method is brute forcing all possible combinations, hence it is highly computationally intensive, especially when it comes to large datasets.
+\\
+The scikit-learn team addresses this disadvantage of gridsearch by introducing the halving gridsearch method \cite{scikithalvinggridsearch} (2020).
+Such technique has proved itself to greatly speed up hyperparameter tuning.
+The ground concept underlying this method is successive halving.
+During the first iteration of halving gridsearch, all candidates are trained on a small subset of the training set.
+Next, we keep only the candidates which performed best and compare  them again on a bigger subset of the training set.
+As the iterations pass, the surviving candidates will be given more and more training samples until when we are left with only the best set of hyperparameters.
+
+\subsection{Cross validation for time series data}
+When data points are dependent on preceding values we cannot use standard K-fold cross validation.
+The rationale is that k-fold will randomize the order of the data, thus i might happen to use future data to predict the past; we want to avoid such behaviour in a time series setting.
+What we to have to keep in mind is that we must test our models during selection in the same way we will run them on the test dataset.
+Hence, for time series crossvalidation we have the TimeSeriesSplit procedure, see figure \ref{fig:crossvalidationtimeseries} for a visualisation.
+\begin{figure}
+    \includegraphics[width=\textwidth]{images/crossvalidationtimeseries.png}
+    \caption{K-fold cross validation for timeseries data}
+    \label{fig:crossvalidationtimeseries}
+\end{figure}
+
+\section{Kernel methods best practices}
+Following, are reported a couple of considerations important to keep in mind when working with kernel methods.
+\subsection{Data scaling}
+\subsection{Data compression}
+Nystrom, Cholesky
+
 \newpage

### 2024-03-24 dataset_applications/bash1.sh

@@ -1,0 +1,11 @@
+#!/bin/bash
+# code for moving test data in respective folder
+for((i=2;i<=15;i++))
+do
+    i_minus_1=$((i - 1))
+    i_minus_1=$((i-1))
+    # get test data
+    cp "/Users/luca/Desktop/GEFCom2014 Data/Load/Task ${i}/L${i}-train.csv" "/Users/luca/Desktop/GEFCom2014 Data/Load/Task ${i_minus_1}/L${i_minus_1}-test.csv"
+done
+
+cp "/Users/luca/Desktop/GEFCom2014 Data/Load/Solution to Task 15/solution15_L_temperature.csv" "/Users/luca/Desktop/GEFCom2014 Data/Load/Task 15/L15-test.csv"

### 2024-03-24 dataset_applications/bash2.sh

@@ -1,0 +1,8 @@
+#!/bin/bash
+
+# for each task clean dataset
+for ((i=1;i<=15;i++))
+do
+    python /Users/luca/Desktop/ThesisKernelMethods/dataset_applications/gefcom_clean_test.py $i
+
+done

### 2024-03-24 dataset_applications/gefcom_clean_test.py

@@ -1,0 +1,90 @@
+import pandas as pd
+import sys
+
+import holidays
+
+def clean_test(ith):
+    if ith==15:
+        df=pd.read_csv(f"/Users/luca/Desktop/GEFCom2014 Data/Load/Task {ith}/L{ith}-test.csv")
+        # dataset 15 does not stick to convention used so far in the competition, thus we have an if conditional to handle it
+        df['TIMESTAMP'] = pd.to_datetime(df['date'], format='%m/%d/%Y')
+
+        # to be coherent hours go from 0 to 23
+        df['HOUR'] =df["hour"]-1
+
+        df['DAY'] = df['TIMESTAMP'].dt.day
+        df['MONTH'] = df['TIMESTAMP'].dt.month
+        df['HOUR'] = df['TIMESTAMP'].dt.hour
+        df['YEAR'] = df['TIMESTAMP'].dt.year
+
+        # add info weekday; saturday, sunday
+        df['DAY_OF_WEEK'] = df['TIMESTAMP'].dt.dayofweek
+
+        # holidays
+        hol=holidays.US(years=range(2001, 2011))
+        df["IS_HOLIDAY"]=df['TIMESTAMP'].isin(hol)
+        # convert it to categorical variable
+        df["IS_HOLIDAY"]=pd.Categorical(df["IS_HOLIDAY"].astype(int))
+
+
+        # drop not useful columns
+        df.drop(columns=['TIMESTAMP', 'date', 'hour'], inplace=True)
+
+
+        # average temperatures
+        weather_stat=df.filter(regex=("w.*")).columns.to_list()
+        # average them 
+        df['w_avg'] = df[weather_stat].mean(axis=1)
+
+        # drop columns
+        df.drop(columns=weather_stat, inplace=True)
+        df.drop(columns=["YEAR"], inplace=True)
+
+        df=df[["LOAD","DAY","MONTH","HOUR","DAY_OF_WEEK","IS_HOLIDAY","w_avg"]]
+        
+        # save to csv
+        df.to_csv(f"/Users/luca/Desktop/GEFCom2014 Data/Load/Task {ith}/L{ith}-test_clean.csv", index=False)
+
+        # print(df)
+
+    else:    
+        df=pd.read_csv(f"/Users/luca/Desktop/GEFCom2014 Data/Load/Task {ith}/L{ith}-test.csv")
+
[…]

### 2024-03-24 dataset_applications/gefcom_clean_train.py

@@ -1,0 +1,60 @@
+import numpy as np
+import pandas as pd
+from sklearn.metrics import mean_pinball_loss
+
+from qr import qr_tests
+
+import holidays
+
+
+if __name__=="__main__":
+    # load data
+    df=pd.read_csv("/Users/luca/Desktop/GEFCom2014 Data/Load/Task 1/L1-train.csv", sep=",", decimal=".")
+    
+    # clean data dates
+    dates = pd.date_range(start='2001-01-01 01:00:00', periods=len(df), freq='H')
+    df['DAY'] = dates.day
+    df['MONTH'] = dates.month
+    df['HOUR'] = dates.hour
+    df['YEAR'] = dates.year
+    
+    # add info weekday; saturday, sunday
+    df['DAY_OF_WEEK'] = dates.dayofweek
+
+    # holidays
+    hol=holidays.US(years=range(2001, 2011))
+    df["IS_HOLIDAY"]=pd.Series(dates.date).isin(hol)
+    # convert it to categorical variable
+    df["IS_HOLIDAY"]=pd.Categorical(df["IS_HOLIDAY"].astype(int))
+    
+
+    # drop not useful columns
+    # zoneid is dropped because are all 1, so it is not a meaningful independent variable
+    # the information in timestamp has been decomposed in four columns so we can drop it
+    df.drop(columns=['ZONEID', 'TIMESTAMP'], inplace=True)
+
+
+    # slice for training, from when we have load informations 2005
+    # slice from 2009 so it is less computations heavy
+    idx_start = df.loc[(df['DAY'] == 1) & (df["MONTH"]==1) & (df["YEAR"]==2009) & (df["HOUR"])==1].index[0]
+    
+    
+    idx_end = df.loc[(df['DAY'] == 1) & (df["MONTH"]==1) & (df["YEAR"]==2010) & (df["HOUR"])==1].index[0]
+
+
+    df = df.loc[idx_start-1:idx_end-1]
+
+    # average temperatures
+    # select weather stations data using regex
+    weather_stat=df.filter(regex=("w.*")).columns.to_list()
+    # average them 
+    df['w_avg'] = df[weather_stat].mean(axis=1)
+
+
+    df.drop(columns=weather_stat, inplace=True)
+    df.drop(columns=["YEAR"], inplace=True)
+
+    df.to_csv("/Users/luca/Desktop/GEFCom2014 Data/Load/L2009-train.csv",index=False)
+
+    
+    

### 2024-03-24 dataset_applications/qr.py

@@ -23,7 +23,7 @@ import sys
 from tqdm import tqdm
 
 from cvxopt import matrix, spmatrix, sparse
-def qr_tests(X,y,write_to_path):
+def qr_tests(X,y,write_to_path=None):
     # quantiles
     quantiles = [0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9]
     # table for output results
@@ -158,25 +158,26 @@ def qr_tests(X,y,write_to_path):
     avg_pinball_scores=pinball_scores.sum(axis=0).to_frame().T
     print(avg_pinball_scores)
     # mae score
-    mae_scores.loc[0,"Linear qr"]=mean_absolute_error(y_test, y_test_pred_qr[5])
-    mae_scores.loc[0,"Gbm qr"]=mean_absolute_error(y_test, y_test_pred_qr_gbr[5])
-    mae_scores.loc[0,"Quantile forest"]=mean_absolute_error(y_test, y_test_pred_qr_rfr[5])
-    mae_scores.loc[0,"Kernel qr"]=mean_absolute_error(y_test, y_test_pred_qr_krn[5])
+    mae_scores.loc[0,"Linear qr"]=mean_absolute_error(y_test, y_test_pred_qr[4])
+    mae_scores.loc[0,"Gbm qr"]=mean_absolute_error(y_test, y_test_pred_qr_gbr[4])
+    mae_scores.loc[0,"Quantile forest"]=mean_absolute_error(y_test, y_test_pred_qr_rfr[4])
+    mae_scores.loc[0,"Kernel qr"]=mean_absolute_error(y_test, y_test_pred_qr_krn[4])
 
     
-    original_stdout=sys.stdout
-    with open(f"/Users/luca/Desktop/ThesisKernelMethods/thesis/tables/{write_to_path}.txt", "w") as f:
-        sys.stdout=f
-        print(avg_pinball_scores.to_latex(index=True,
-                  formatters={"name": str.upper},
-                  float_format="{:.4f}".format))
-    
-        print(pinball_scores.to_latex(index=True,
-                  formatters={"name": str.upper},
-                  float_format="{:.4f}".format))
+    if write_to_path!=None:
+        original_stdout=sys.stdout
+        with open(f"/Users/luca/Desktop/ThesisKernelMethods/thesis/tables/{write_to_path}.txt", "w") as f:
+            sys.stdout=f
+            print(avg_pinball_scores.to_latex(index=True,
+                    formatters={"name": str.upper},
[…]

### 2024-03-24 dataset_applications/qr_gefcom2014.py

@@ -1,0 +1,136 @@
+import numpy as np
+import pandas as pd
+import pickle
+from pprint import pprint
+from sklearn.experimental import enable_halving_search_cv  # noqa
+from sklearn.model_selection import HalvingRandomSearchCV
+from sklearn.metrics import make_scorer
+from sklearn.metrics import mean_absolute_error
+from sklearn.metrics import mean_pinball_loss
+from sklearn.model_selection import train_test_split
+from sklearn.preprocessing import StandardScaler
+import sys
+from tqdm import tqdm
+
+from kqr import KQR
+
+from cvxopt import matrix, spmatrix, sparse
+
+
+
+
+if __name__=="__main__":
+    # load data 2009
+    df=pd.read_csv("/Users/luca/Desktop/GEFCom2014 Data/Load/L2009-train.csv")
+    
+    X_train=df[["DAY",  "MONTH",  "HOUR",  "DAY_OF_WEEK",  "IS_HOLIDAY",  "w_avg"]]
+    y_train=df["LOAD"]
+
+    # 99 quantiles
+    quantiles = [i/100 for i in range(1,100)]
+
+    # define loss to tune
+    neg_mean_pinball_loss_scorer_05 = make_scorer(
+    mean_pinball_loss,
+    alpha=0.5,
+    greater_is_better=False,  # maximize the negative of the loss
+    )
+
+    # kernel quantile regression
+    scaler = StandardScaler()
+
+    X_train_scaled = scaler.fit_transform(X_train)
+
+    qr_krn_models=[]
+    y_test_pred_qr_krn=[]
+
+    param_grid_krn = dict(
+    C=[1e-1,1e-2,1, 5, 10,1e2,1e4],
+    gamma=[1e-1,1e-2,0.5,1, 5, 10, 20]   
+    )
+    
+    krn_blueprint=KQR(alpha=0.5)
+    best_hyperparameters_krn=HalvingRandomSearchCV(
+            krn_blueprint,
+            param_grid_krn,
+            scoring=neg_mean_pinball_loss_scorer_05,
+            n_jobs=2,
+            random_state=0,
+        ).fit(X_train_scaled, y_train).best_params_
+    
+    for i,q in enumerate(tqdm(quantiles)):
+
+        # fit data for specific quantile
+        qr_krn_models+=[KQR(alpha=q, **best_hyperparameters_krn).fit(X_train_scaled, y_train)]
+
+        # save models to pickle
[…]

### 2024-03-24 dataset_applications/qr_gefcom2014_test.py

@@ -1,0 +1,39 @@
+
+import pandas as pd
+import pickle
+
+from sklearn.metrics import mean_absolute_error
+from sklearn.metrics import mean_pinball_loss
+from sklearn.preprocessing import StandardScaler
+import sys
+
+from tqdm import tqdm
+
+
+def test(ith):
+    df=pd.read_csv(f"/Users/luca/Desktop/GEFCom2014 Data/Load/Task {ith}/L{ith}-test_clean.csv")
+
+    X_test=df[["DAY",  "MONTH",  "HOUR",  "DAY_OF_WEEK",  "IS_HOLIDAY",  "w_avg"]]
+    y_test=df["LOAD"]
+
+    quantiles = [i/100 for i in range(1,100)]
+
+    # we need to scale X_test
+    df_train=pd.read_csv("/Users/luca/Desktop/GEFCom2014 Data/Load/L2009-train.csv")
+    X_train=df_train[["DAY",  "MONTH",  "HOUR",  "DAY_OF_WEEK",  "IS_HOLIDAY",  "w_avg"]]
+    scaler = StandardScaler()
+    X_train_scaled = scaler.fit_transform(X_train)
+    X_test_scaled = scaler.transform(X_test)
+
+    pinball_tot=0
+
+    for i,q in enumerate(quantiles):
+        krn_q=pickle.load(open(f"/Users/luca/Desktop/ThesisKernelMethods/dataset_applications/models/krn_qr_{i}.pkl", 'rb'))
+        pinball_q=mean_pinball_loss(y_test,krn_q.predict(X_test_scaled), alpha=q)
+        print(f"pinball loss quantile {q}: ", pinball_q)
+        pinball_tot+=pinball_q
+    
+    print("total quantile: ", pinball_tot/len(quantiles))
+
+if __name__=="__main__":
+    test(int(sys.argv[1]))

### 2024-03-25 thesis/appendix.tex
cv theory done

--- 2024-03-23 thesis/appendix.tex
@@
 
+\section{Cross validation}
+Cross validation directly estimates the expected test error
+\begin{equation}
+    Err=\mathbb{E}\left[L\left(Y,\hat{f}(X)\right)\right]=\mathbb{E}\left[Err_{\Tau}\right]
+\end{equation}
+$Err_{\Tau}$ is the prediction error over an independent test sample
+\begin{equation}
+    \mathbb{E}_{\Tau}=\mathbb{E}\left[L\left(Y,\hat{f}(X)\right)| \Tau \right]
+\end{equation}
+That is $Err$ is the average over everything that is random: X,Y and the training set $\Tau$ used to learn $\hat{f}$.
+Hence, our interest lies in estimating the $Err$ quantity in order to guide model selection.
+\subsection{K-fold cross validation}
+K-fold cross validation splits the data into K roughly equally sized parts, then K models are trained.
+For k from 1 to K we train the kth model on the whole dataset except the kth partition.
+Next,  the prediction error of the kth fitted model is computed.
+Finally averaging all the k prediction erros we obtain an estimate for the expected test error.
+We select the model which performs best in terms of expected prediction error.
+\\
+Usually, K is set equal to 5 or 10. Leave-one-out cross validation is the case when K is set equal to the size of the data.
+% REMEMBER: Parameters allow the model to learn the rules from the data while hyperparameters control how the model is training.
+Parameters learn their own values from data.
+In contrast, hyperparameters do not learn their values from data.
+We need to manually specify them before training the model
+\subsection{Gridsearch}
+Model performance depends higly on the choice of hyperparameters.
+Notice, there is no way to get to know them in advance, therefore, all we can do is trying a lot of combinations until we fit a good enough set of hyperparameters.
+Essentially, gridsearch carries out hyperparameter tuning by performing a search over a predefined hyperparemeters grid.
+During its search, it tries all possible combination and evaluates the different models %(different hyperparameters define different models)
+using cross validation. For example, suppose that the grid contains 100 possible candidates and that we are doing 5-fold cross-validation, then the gridsearch algorithm will carry out 500 iterations.
+Therefore, we get an estimate of prediction error for each considered model and this guides us in hyperparameters (model) selection.
+
+\subsection{Randomized search}
+Randomized search controls the number of steps by choosing smartly the hyperparameters to try in each iterations.
+Let us say, there are 100 candidates and we set the number of iterations to 20 then the search will stop after the 20th iteration and return the best set among the hyperparameters observed.
+The advantage is that it is much more quicker than gridsearch, but on the other hand its performance is worse than gridsearch.
+
+\subsection{Halving gridsearch}
+Gridsearch has been the to go choice for hyperparameters tuning for the past years.
+However, such method is brute forcing all possible combinations, hence it is highly computationally intensive, especially when it comes to large datasets.
+\\
+The scikit-learn team addressed this disadvantage of gridsearch by introducing the halving gridsearch method \cite{scikithalvinggridsearch} (2020).
+Such technique has proved itself to greatly speed up hyperparameter tuning.
+The ground concept underlying this method is successive halving.
+During the first iteration of halving gridsearch, all candidates are trained on a small subset of the training set.
+Next, we keep only the candidates which performed best and compare  them again on a bigger subset of the training set.
+As the iterations pass, the surviving candidates will be given more and more training samples.
+The algorithm stops when we are left with only the best set of hyperparameters.
+
+\subsection{Cross validation for time series data}
+When data points are dependent on preceding values, we cannot use standard K-fold cross validation.
+The rationale is that K-fold will randomize the order of the data, thus it might happen to use future data to predict the past; we want to avoid such behaviour in a time series setting.
+When carryout out any kind of cross validation, we must keep consistency in the way we evaluate our predictors during model selection and in the way we perform evaluation of the test data.
+Hence, for time series crossvalidation we have the timeseries split procedure, see figure \ref{fig:crossvalidationtimeseries} for a visualisation; the blue observations make up the training sets while the orange observations form the test sets.
+\begin{figure}
+    \includegraphics[width=\textwidth]{images/crossvalidationtimeseries.png}
+    \caption{Cross validation for one step ahead timeseries data \cite{hyndman2018forecasting}}
+    \label{fig:crossvalidationtimeseries}
+\end{figure}
+We then average all the observed losses in order to get an estimate for $Err$.
+Notice, in the literature this procedure is sometimes also referred to evaluation on a rolling forecasting origin.
+This comes from the fact that at each iteration we push forward the origin of our forecast.
+The same concept applies for multi step ahead forecasting, see picture \ref{fig:crossvalidationtimeseries2}.
+That is in predicting $\hat{L}_{N+m}$ we use as inputs $L_1, L_2, \dots, \hat{L}_{N},\hat{L}_{N+1},\dots, \hat{L}_{N+m-1}$; where $\hat{L}_{N},\hat{L}_{N+1},\dots, \hat{L}_{N+m-1}$ are one step ahead forecasts.
+\begin{figure}
+    \includegraphics[width=\textwidth]{images/crossvalidationtimeseries2.png}
+    \caption{Cross validation for m step ahead timeseries data \cite{hyndman2018forecasting}}
+    \label{fig:crossvalidationtimeseries2}
+\end{figure}
+
+\section{Kernel methods best practices}
+Following, are reported a couple of considerations important to keep in mind when working with kernel methods.
+\subsection{Data scaling}
+\subsection{Data compression}
+Nystrom, Cholesky
+
 \newpage

### 2024-03-26 thesis/input_files/probabilistic.tex
March 26th 2024

@@
         \\
-        \frac{\partial L}{\partial b}=0 \implies \alpha-\alpha^*=0
+        \frac{\partial L}{\partial b}=0 \implies (\alpha-\alpha^*)\mathbb{1}=0
         \\
@@
 \end{equation}
-Where an unbiased estimator for $b$ is given by
+Since our optimisation problem possesses strong duality and it is differentiable in both the objective and the constraint, we have that it must satisfy the Karush Kuhn Tucker conditions, see section 5.5.3 \cite{boyd2004convex}.
+Thanks the KKT conditions on the primal optimisation problem we have that $f(x_i)=y_i \ \mathrm{for} \ \alpha_i \not \in \{C(q-1), Cq\}$. 
+To see this, we have to consider the KKT conditions.
 \begin{equation}
     \begin{aligned}
-    b=&\frac{1}{I_q}\sum\limits_{i\in I_q} b_i, \\ 
-    \textrm{with} \ b_i=& y_i-\sum\limits_k \alpha_k k(x_k, x_i) \\
-     \textrm{and} \ I_q=& \{i=1,\dots,n|0<\alpha_i<Cq, 0<\alpha_i<C(1-q)\}
+    \lambda^* r_i(x^*)=&0, \quad i=1,\dots,m
+    \\
+    \nabla L(x^*)=&0
 \end{aligned}
 \end{equation}
+In our setting we have
+\begin{equation}
+    \begin{aligned}
+        \alpha_i(\xi_i-y_i+r_i)=&0
+        \\
+        \alpha_i^*(\xi_i^*+y_i-r_i)=&0
+        \\
+        \nu_i \xi_i=&0
+        \\
+        \nu_i^* \xi_i^*=&0
+        \\
+        \nabla L =&0
+    \end{aligned}
+\end{equation}
+Using the gradient of the lagrangian of equation \ref{eq:lagrange_derivatives} we end up with
+\begin{equation}
+    \begin{aligned}
+        \alpha_i(\xi_i-y_i+r_i)=&0
+        \\
+        \alpha_i^*(\xi_i^*+y_i-r_i)=&0
+        \\
+        (Cq-\alpha_i) \xi_i=&0
+        \\
+        (C(1-q)-\alpha_i) \xi_i^*=&0
+    \end{aligned}
+\end{equation}
+
+Now, let us break into cases
+
+\begin{equation}
+\begin{cases}
+    \alpha_i=Cq, \ \alpha_i^*=0 & \implies \alpha_i-\alpha_i^*=Cq, \xi_i\leq 0, \xi^*=0\implies \xi_i-y_i+f_i=0 \\
+    \alpha_i=0, \ \alpha_i^*=C(1-q) & \implies \alpha_i-\alpha_i^*=C(q-1), \xi_i= 0, \xi^*\leq 0 \implies \xi_i^*+y_i-f_i=0 \\
+    0\leq \alpha_i\leq Cq, \ 0\leq \alpha_i^*\leq C(1-q) & \implies \xi_i=0,  \xi^*_i= 0 \implies -y_i+f_i=0 \ y_i-f_i=0 \\
+\end{cases}
+\end{equation}
+
+Therefore in order to retrieve b we simply have to choose an index $i$ such that bold $\alpha_i \not \in \{C(q-1), Cq\}$
+and let
+\begin{equation}
+    \begin{aligned}
+    b=&y_i-\sum_i \alpha_i k(x_i,x)
+    \end{aligned}
+\end{equation}
+
+% In our implementatiton we take the point that minsimizes the distance between the boundary of domain of alpha, in this way we are guaranteed we do not take ith index of alpha_i for retrieving b, because it that case it would not work.
 %contribution is implementation kernel quantile regression for pythos users since there exists only r version

### 2024-03-26 dataset_applications/bash3.sh

@@ -1,0 +1,10 @@
+#!/bin/bash
+
+# clean
+rm /Users/luca/Desktop/ThesisKernelMethods/thesis/tables/qr_gefcom2014.txt
+
+for((i=1;i<=15;++i))
+do
+    python /Users/luca/Desktop/ThesisKernelMethods/dataset_applications/qr_gefcom2014_test.py $i
+
+done

### 2024-03-26 dataset_applications/qr_gefcom2014_test.py

@@ -34,6 +34,19 @@ def test(ith):
         pinball_tot+=pinball_q
     
     print("total quantile: ", pinball_tot/len(quantiles))
+    ans=pinball_tot/len(quantiles)
+    return ans
 
 if __name__=="__main__":
-    test(int(sys.argv[1]))
+    i=int(sys.argv[1])
+    ans=test(i)
+
+
+    original_stdout=sys.stdout
+    with open(f"/Users/luca/Desktop/ThesisKernelMethods/thesis/tables/qr_gefcom2014.txt", "a") as f:
+        sys.stdout=f
+        # print(f"Average pinball loss, task n. {i}: ")
+        print(f"{ans:.4f}")
+        print("&")
+
+        sys.stdout=original_stdout

### 2024-03-27 thesis/appendix.tex
March 27th 2024

@@
 \end{table}
+What can be concluded from these numerical examples is that, on average kernel quantile regression yields better results than quantile forest \cite{meinshausen2006quantile} and gradient boosting machine quantile regression \cite {friedman2001greedy} in terms of the pinball loss as well as in terms of the mean absolute error.
 
@@
 Following, are reported a couple of considerations important to keep in mind when working with kernel methods.
-\subsection{Data scaling}
+
+\subsection{Data normalization}
+With data normalization we transform the range of the feautures to a standard scale.
+Such preprocessing step is essential when employing distance based algorithms like svm or k-nearest neighbors.
+The rationale behind it is that by normalizing data we give an uniform weight to each feature in the learning process; in this way we do not favour larger scale features.
+Examples of the most popular features scaling are:
+\begin{itemize}
+    \item Standard scaler= it computes the standard score z of a sample x
+    \\
+    z=$\frac{x-\mu}{\sigma}$.
+    \item MinMax scaler= it maps every data sample to the range 0, 1.
+    \\
+    $\frac{x-\min(X)}{\max(X)-\min(X)}$
+    \item Robust scaler= it scales features using statistics that are robust to outliers.
+    Essentially, it subtracts the median and then scales the data according to the interquantile range.
+\end{itemize}
+Many other data scaling algorithms exists, we refer the reader to a thorough comparison \cite{scikitscalers}.
+\\
+We conclude this subsection with a custom example that motivates the need of feature scaling \cite{scikitscale_example}.
+The idea is to compare the results of modelling the data with k nearest neighbors on the unscaled data against the scaled data.
+The considered data is the wine recognition dataset \href{https://archive.ics.uci.edu/dataset/109/wine}{https://archive.ics.uci.edu/dataset/109/wine}.
+The goal for this dataset is recognising from whose cultivator the wine comes based on two features with a completely different scale.
+The first feature has values in the [0,1000] range while the second feature has values contained in [1,10].
+The unscaled and scaled version are compared in figure \ref{fig:feature_scaler_example}
+\begin{figure}[!h]
+    \includegraphics[width=\textwidth]{images/feature_scaler_example.png}
+    \caption{importance of feature scaling}
+    \label{fig:feature_scaler_example}
+\end{figure}
+What we can conclude from the image is that, the model trained on scaled data is greatly better than the other.
+On the left, we can see that distances between categories are impacted solely by the larger scale feature.
+Conversely, on the right, we have that the two features contributing equally in determining the neighbors.
+
 \subsection{Data compression}
-Nystrom, Cholesky
+When the number of points n is large, kernel methods suffer from high computational costs. 
+Using kernel methods, we have that the storage cost is of the order of $O(n^2)$ while the computational cost for finding the solution is of the order $O(n^3)$.
+A significant speed up can be obtained thanks to low rank approximation.
+
+\subsubsection{Nystrom decomposition}
+The Nyström approximation involves storing a submatrix of the whole kernel matrix. Thus, storage and computational cost are reduced to $O(nm)$ and $O(nm^2)$ respectively.
+Nyström works by selecting $m<n$ points, callled representative points; it approximates K as
+\begin{equation}
+    \tilde {K}=K_{n,m} K_{m,m}^{-1}K_{n,m}^\intercal
+\end{equation}
+
+\subsubsection{Pivoted Cholesky decomposition}
+Pivoted Cholesky approximate the Cholesky decomposition of a matrix.
+Since kernel matrices are positive definite, they can decomposed in terms of the Cholesky decomposition.
+Hence, we have that pivoted Cholesky can be used to approximate the full kernel matrix.
 

### 2024-03-27 thesis/input_files/paper1.tex
March 27th 2024

@@
 \end{align*}
+
+\subsection{Kernel families}
+Table \ref{tab:kernel types} contains popular kernel families in literature and applications.
+\begin{table}
+    \caption{Kernel types}
+    \begin{tabular}{lll}
+        \toprule
+       Kernel function & Equation & Hyperparameters \\
+       \midrule
+       Linear &  $k(x_1,x_2)=x_1x_2$ &   \\
+       Polynomial &  $k(x_1,x_2)=(x_1^\intercal x_2+c)^d$ &   c, d\\
+       Gaussian RBF &  $k(x_1,x_2)=e^{-\frac{\|x_1-x_2|\|^2}{2\sigma^2}}$ &   $\sigma$\\
+       Exponential RBF/Laplacian& $k(x_1,x_2)=e^{-\frac{|x_1-x_2|}{\gamma}}$ &   $\gamma$\\
+       Hyperbolic/Sigmoid Kernel &  $k(x_1,x_2)=\tanh(\gamma x_1^\intercal x_2+r)$ &  $\gamma$, r \\
+       Periodic &  $k(x_1,x_2)=e^{\frac{-2 sin^2\left(\frac{\pi}{p}|x_1-x_2| \right)}{l^2}} $ &   p, l\\
+        \bottomrule
+    \end{tabular}
+    \label{tab:kernel types}
+\end{table}

### 2024-03-27 thesis/input_files/point.tex
March 18th 2024
March 26th 2024
March 27th 2024

@@
 \section{Generalized additive model}
-
+%prophet meta
 \section{K-nearest neighbors}
@@
 \section{Support vector regression}
+Developed at the AT\&T Bell Laboratories by Vapnik et al.
+\cite{cortes1995support}, \cite{vapnik1997support}, support vector machines SVMs are one of the most popular techniques within the field of statistical learning.
+\\
+The goal of support vector regression is finding a function f(x) with at most $\epsilon$ deviation from the actual observed data $y_i$ $\forall i$ and as flat as possible. 
+Put differently, we would like a model to keep error less than the $\epsilon$ threshold,  
+%https://stats.stackexchange.com/questions/5945/understanding-svm-regression-objective-function-and-flatness
+In standard SVR we have
+\begin{equation}
+    f(x)=\langle w,x \rangle +b \ \textrm{with} \ w \in X, b \in R
+\end{equation}
+Where $X$ denotes the space of the input patters $x_i$.
+We can translate the flatness requirement into minimising the squared norm of w. Doing so we can formulate our problem as a convex optimisation problem.
+\begin{equation}
+    \begin{aligned}
+        \min \quad& \frac{1}{2}\|w\|^2
+        \\
+        s.t. \quad& y_i-\langle w, x_i\rangle-b\leq \epsilon
+        \\
+        \quad& \langle w, x_i\rangle +b-y_i\leq \epsilon
+    \end{aligned}
+\end{equation}
+Next, we need to introduce the slack variables $\xi$ and $\xi^*$, in order to handle the above formulate optimisation problem.
+We obtain the following formulation
+\begin{equation}
+    \begin{aligned}
+        \min \quad& \frac{1}{2}\|w\|^2+C\sum\limits_{i=1}^l(\xi_i+\xi_i^*)
+        \\
+        s.t. \quad& y_i-\langle w, x_i\rangle-b\leq \epsilon+\xi
+        \\
+        \quad& \langle w, x_i\rangle +b-y_i\leq \epsilon+\xi^*
+        \\
+        \quad& \xi_i\geq0
+        \\
+        \quad& \xi_i^*\geq0
+    \end{aligned}
+\end{equation}
+The C constant trades off between $\epsilon$ deviation tolerance and flatness of the function f.
+We seek to minimise the epsilon insensitive loss function, defined as follows
+\begin{equation}
+    \|\xi\|_\epsilon:=\begin{cases}
+        0 \quad& \textrm{if} \ \|\xi\|\leq \epsilon
+        \\
+        \|\xi\|-\epsilon \quad& \textrm{if} \ \|\xi\|\leq \epsilon
+    \end{cases}
+\end{equation}
+The model is depicted in figure \ref{fig:svm_simple}. The gry band is called the epsilon insensitive tube, only the points outside it are accounted by the loss function.
+\begin{figure}
+    \includegraphics[width=\textwidth]{images/svm_simple.png}    
+    \caption{Support vector regression,\cite{learning_with_kernels}}
+    \label{fig:svm_simple}
+\end{figure}
+Smola et al. \cite{smola2004tutorial} point out that considering the dual formulation makes our optimisation problem easier to solve. Doing so we have
+\begin{equation}
+    \begin{aligned}
+        \max \quad& -\frac{1}{2}\sum\limits_{i,j=1}^l(\alpha_i-\alpha_i^*)(\alpha_j-\alpha_j^*)\langle x_i,x_j\rangle +\sum\limits_{i=1}^l(\alpha_i-\alpha_i^*)(y_i-\epsilon)
+        \\
+        s.t. \quad& \sum\limits_i=1^l(\alpha_i-\alpha_i^*)=0
+        \\
+        \quad& \alpha_i, \alpha_i^* \in [0, C]
+    \end{aligned}
+\end{equation}
+Rearranging the gradient of the lagrangian with respect to w, we obtain the so called support vector expansion.
+\begin{equation}\label{eq:support vector expansion}
+    w=\sum\limits_{i=1}^l(\alpha_i-\alpha_i^*)x_i
+\end{equation}
+That means that w can be completely described as a linear combination of the training features $x_i$.
+Notice, the complexity of the function representation is independent of the feature space dimensionaly but depends only on the number of support vectors; that is those point $i$ for which $(\alpha_i-\alpha_i^*)\neq 0$
+Notice that, we do not needd to compute w explicitly in order to evaluate f(x).
+Furthermore, equation \ref{eq:support vector expansion} implies
+\begin{equation}
+    f(x)=\sum\limits_{i=1}^l(\alpha_i-\alpha_i^*)\langle x_i, x\rangle +b
+\end{equation}
+Employ the Karush-Kuhn-Tucker conditions, b can be retrieved easily.
+Particularly, use the condition that the product between the constraints and the dual variable has to vanish.
+These implies that the lagrange multipliers $\alpha_i, \alpha_i^*$ may be nonzero only for the samples inside the epsilon insensitive tube.
+Consequently for any of these data points, the equality $b=y_i-\langle w, x_i-\epsilon\rangle$ holds.
+To get an idea, see figure \ref{fig:svr1} for a how support vector regression handles a sinusoidal function with noise.
+\\
+\begin{figure}
+    \includegraphics[width=\textwidth]{images/svr1.png}
+    \caption{support vector regression}
+    \label{fig:svr1}
+\end{figure}
 
@@
 \subsection{Kernel support vector regression}
+Support vector regression can be kernelized by swapping the R2 euclidean dot product of data x with the dot product in the higher feature space F. Doing so, the optimisation problem can be restated as
+\begin{equation}
+    \begin{aligned}
+        \max \quad& -\frac{1}{2}\sum\limits_{i,j=1}^l(\alpha_i-\alpha_i^*)(\alpha_i-\alpha_i^*)k(x_i, x_j)+\sum\limits_{i=1}^l(y_i-\epsilon)(\alpha_i-\alpha_i^*)
+        \\
+        s.t. \quad& \sum\limits_{i=1}^l(\alpha_i-\alpha_i^*)=0
+        \\
+        \quad& \alpha_i, \alpha_i^* \in [0,C]
+    \end{aligned}
+\end{equation}
+Our regressor f is then given by
+\begin{equation}
+    \begin{aligned}
+        \sum\limits_{i=1}^l(\alpha_i-\alpha_i^*)k(x_i, x)+b
+    \end{aligned}
+\end{equation}
+Notice, in this setting w is no longer given explicitly and we seek for the flattest function in the feature space not the input space.
+See figure \ref{fig:svr2} and figure \ref{fig:svr3} for two examples. In the former we have used a polynomial kernel while in the latter the standard radial basis function kernel.
+\begin{figure}
+    \includegraphics[width=\textwidth]{images/svr2.png}
+    \caption{polynomial support vector regression}
+    \label{fig:svr2}
+\end{figure}
+
+\begin{figure}
+    \includegraphics[width=\textwidth]{images/svr3.png}
+    \caption{rbf support vector regression}
+    \label{fig:svr3}
+\end{figure}
+Comparing theese pictures with figure \ref{fig:svr1}, it can be concluded that introducing kernels allows support vector regression to handle non linearities in the data.

### 2024-03-27 thesis/input_files/probabilistic.tex
March 27th 2024

@@
 \\
-In order to understand this algorithm it is necessary to first cover the theory of decision trees and random forests.
+In order to understand this algorithm, it is first necessary to first cover the theory of decision trees and random forests.
 \subsection{Decision trees}
@@
 Our goal is selecting the best features and best split point such that to achieve the best generalizing fit.
-For an example/visualisation consider the pictures \ref{fig:elements_statistical_learning1} and \ref{fig:elements_statistical_learning2} taken from \cite{hastie2009elements}, where the decision tree algorithm is visualised for a regression problem with two independent variables X2 and X1.
+For an example/visualisation consider the pictures \ref{fig:elements_statistical_learning1} and \ref{fig:elements_statistical_learning2} taken from \cite{hastie2009elements}, where the decision tree algorithm is visualised for a regression problem with two independent variables $X_2$ and $X_1$.
 \begin{figure}
     \includegraphics[width=0.5\textwidth]{images/elsii1.png}
-    \caption{two-dimensional feature space partitioned by recursive binary splittin}
+    \caption{two-dimensional feature space partitioned by recursive binary splitting}
     \label{fig:elements_statistical_learning1}
@@
 \begin{equation}
-    f(x)=\sum\limits_{m=1}^{M}c_m\mathbb{1}(x \in R_m)
+    f(x)=\sum\limits_{m=1}^{M}c_m\mathbb{I}_{\{x \in R_m\}}
 \end{equation}
 \\
-It follows that the function minimising the sum of squares is the one with $\hat{c}_m=mean(y_i|x_i \in R_m)$
+It follows that the function minimising the sum of squares is the one with $\hat{c}_m=mean(y_i|x_i \in R_m)$.
 Finding the best split in terms of minimum sum of squares is computationally infeasible in practice. Therefore we approximate a solution by approaching the problem in a greedy fashion.
-Let $j$ denote the splitting variable, $s$ be the the split point we define the two half planes
+Let $j$ denote the splitting variable and $s$ be the the split point we define the two half planes
 \begin{equation}
@@
 Bagging stands for boostrap aggregation. 
-The bagging estimate is dafined by $\mathbb{E}_P\hat{f}^*$ where $P$ is the empirical distribution putting equal probability on each data point of the training set.
+The bagging estimate is dafined by $\mathbb{E}_P[\hat{f}^*]$ where $P$ is the empirical distribution putting equal probability on each data point of the training set.
 Basically, for each bootstrap fitted model $\hat{f}^{*b}(x)$, we compute the bagging estimate by
@@
 The second term disappears as B increases, while the first term depends heavily on the correlation between bagged trees.
-Random forest consists in is reducing the correlation between the trees by randomly selecting m of the p features as candidates for splitting; m tipically takes value of the order $\sqrt{p}$ or even 1 with default value $m=\left\lfloor \frac{p}{3} \right\rfloor$ while a good minimum node size is around five.
+Random forest consists in is reducing the correlation between the trees by randomly selecting m of the p features as candidates for splitting; m tipically takes a value in the order of $\sqrt{p}$ or even 1 with default value $m=\left\lfloor \frac{p}{3} \right\rfloor$, while a good minimum node size is around five.
 \\
@@
 \begin{equation}
-    F(y|X=x)=P(Y\leq y|X=x)=\mathbb{E}(\mathbb{1}_{\{Y\leq y\}}|X=x)
+    F(y|X=x)=P(Y\leq y|X=x)=\mathbb{E}(\mathbb{I}_{\{Y\leq y\}}|X=x)
 \end{equation}
-All we have to do is approximating $\mathbb{E}(\mathbb{1}_{\{Y\leq y\}}|X=x)$ by a weighted mean over the random variable $\mathbb{1}_{\{Y\leq y\}}$
+All we have to do is approximating $\mathbb{E}(\mathbb{I}_{\{Y\leq y\}}|X=x)$ by a weighted mean over the random variable $\mathbb{I}_{\{Y\leq y\}}$
 \begin{equation}
-    \hat{F}(y|X=x)=\sum\limits_{i=1}^{n}\omega_i(x)\mathbb{1}_{\{Y_i\leq y\}}
+    \hat{F}(y|X=x)=\sum\limits_{i=1}^{n}\omega_i(x)\mathbb{I}_{\{Y_i\leq y\}}
 \end{equation}
@@
 \begin{equation}
-    \hat{Q}_\alpha=\inf\{y:\hat{F}(y|X=x)\geq \alpha\}
+    \hat{Q}_q=\inf\{y:\hat{F}(y|X=x)\geq q\}
 \end{equation}
@@
 \end{equation}
-Thus, at each step of the iterative optimisation procedure we have to solve
+Thus, at each step of the iterative optimisation procedure, we have to solve
 \begin{equation}\label{boosting_minimisation_problem}
@@
 \begin{equation}
-    \tilde{\Theta_m}=\argmin{\Theta}\sum\limits_{i=1}^{N}\left(-g_{im}-T(x_i;\Theta)\right)^2
+    \tilde{\Theta}_m=\argmin{\Theta}\sum\limits_{i=1}^{N}\left(-g_{im}-T(x_i;\Theta)\right)^2
 \end{equation}
 where $g_{im}$ is the gradient of $L(f)=\sum\limits_{i=1}^N L(y_i, f(x_i))$ evaluated at $f_{m-1}$.
-Put simply, we are fitting the mth tree to the negative of the gradient values of f through least squares.
+Put simply, we are fitting the m-th tree to the negative of the gradient values of f through least squares.
 In order to solve our quantile regression tasks through gradient boosting, all we need to do is specifying the pinball loss as our criterion to guide the minimisation algorithm.
@@
 \begin{equation}\label{eq:kqr_min1}
-    R[f]:=\frac{1}{m}\sum\limits_{i=1}^{m}\rho_q(y_i-f(x_i))+\frac{\lambda}{2}\|g\|_H^2
+    R[f]:=\frac{1}{m}\sum\limits_{i=1}^{m}\rho_q(y_i-f(x_i))+\frac{\lambda}{2}\|g\|_\mathcal{H}^2
 \end{equation}
-where $f=g+b$, $g \in H$ and $b \in R$.
+where $f=g+b$, $g \in \mathcal{H}$ and $b \in R$.
 Using the link between RKHS and feature spaces, we can rewrite $f(x)=\langle w, \phi(x) \rangle+b$.
-Moreover, note that the RKHS norm is defined as follows $\|f\|_{H}=\inf\{\|w\|_{F}:w\in F,f(x)=\langle w,\varphi (x)\rangle _{F},\forall x\in X\}$, where $F$ is the feature space.
+Moreover, note that the RKHS norm is defined as follows $\|f\|_{\mathcal{H}}=\inf\{\|w\|_{F}:w\in F,f(x)=\langle w,\varphi (x)\rangle _{F},\forall x\in X\}$, where $F$ is the feature space.
  Doing so we obtain a minimization problem equivalent to minimizing equation \ref{eq:kqr_min1}.
@@
 \begin{definition}
-    The dual function associated to the lagrangia $L(x,\lambda)$ is given by $g(v)=\inf_x L(x,\lambda)$
+    The dual function associated to the lagrangian $L(x,\lambda)$ is given by $g(v)=\inf_x L(x,\lambda)$
 \end{definition}
@@
 Consider a simple lagrangian $L(x,\lambda)=f(x)+\sum \lambda_i r_i(x) +\sum v_i h_i(x)$, where $r_i(x)$ are inequality constraints while $h_i(x)$ are equality constraints of the problem.
-Then it can be noted that the lower bound on $p^*$ is non trivial only when the lagrange multiplier $\lambda \succeq 0$.
+Then it can be noted that, the lower bound on $p^*$ is non trivial only when the lagrange multiplier $\lambda \succeq 0$.
 Therefore, the idea is that by maximizing the dual function subject to the contraint $\lambda \succeq 0$, we can obtain an approximate or perfect solution to the primal problem.
 \\
-To explain why we may or not be able to attain the best solution to the primal problem by maximizing its dual we have to introduce the concept of duality.
+To explain why we may or not be able to attain the best solution to the primal problem by maximizing its dual, we have to introduce the concept of duality.
 \\
@@
         \exists x \in \textrm{relint} \ D \ \textrm{s.t.} & \ r_i(x)<0, \quad i=1, \dots, m \\
-        & h_i(x)=b
+        & h_i(x)=0
     \end{aligned}
@@
 \end{equation}
-In our case \ref{eq:kqr_min5}, we have $w^\intercal w$, thus $Q$ is just the identity matrix which satisfies the positive definiteness requirement. Therefore, our problem is convex.
+In our case, equation \ref{eq:kqr_min5}, we have $w^\intercal w$, thus $Q$ is just the identity matrix which satisfies the positive definiteness requirement. Therefore, our problem is convex.
 Next we check that Slater's condition holds.
@@
 \\
-We end this section with the derivation of the dual problem; that is the convex problem we will solve in order to get the quantiles prdeiction of our quantile kernel algorithm.
+We end this section with the derivation of the dual problem; that is the convex problem, we will solve in order to get the quantiles prdeiction of our quantile kernel algorithm.
 \\
@@
         \\
-        \frac{\partial L}{\partial b}=0 \implies (\alpha-\alpha^*)\mathbb{1}=0
+        \frac{\partial L}{\partial b}=0 \implies (\alpha-\alpha^*)^\intercal\mathbb{1}=0
         \\
@@
 As pointed out previously, the lower bound resulting from the dual formulation is non trivial only when the lagrange multipliers are $\succeq 0$.
-Looking at the last two equations of the system \ref{eq:lagrange_derivatives}, this implies the following two constraints $\alpha \in [0, Cq\mathbb{1}]$ and $\alpha \in [0, C(1-q)\mathbb{1}]$.
+Looking at the last two equations of the system \ref{eq:lagrange_derivatives}, this implies the following two constraints $\alpha \in [0, Cq\mathbb{1}]$ and $\alpha^*\in [0, C(1-q)\mathbb{1}]$.
 \\
-Substitute the conditions for an optimum into \ref{eq:kqr_min5} we obtain the dual formulation.
+Substitute the conditions for an optimum into \ref{eq:kqr_min5}, we obtain the dual formulation.
 \begin{equation}
@@
 \end{equation}
-Defining $\alpha=(\alpha-\alpha^*)$ and letting $K$ the kernel matrix, we have that the dual optimisation problem reads as follows
+Defining $\textrm{a}=(\alpha-\alpha^*)$ and letting $K$ the kernel matrix, we have that the dual optimisation problem reads as follows
 \begin{equation}\label{eq:kqr_min6}
     \begin{aligned}
-        \max_{\alpha} \quad & -\frac{1}{2}\alpha^\intercal K\alpha+\alpha^\intercal y\\
+        \max_{\textrm{a}} \quad & -\frac{1}{2}\textrm{a}^\intercal K\textrm{a}+\textrm{a}^\intercal y\\
     \textrm{s.t.} \quad & 
-    C(q-1)\mathbb{1}\preceq \alpha \preceq Cq\mathbb{1}\\
-    &\alpha^\intercal\mathbb{1}=1
+    C(q-1)\mathbb{1}\preceq \textrm{a}\preceq Cq\mathbb{1}\\
+    &\textrm{a}^\intercal\mathbb{1}=1
     \end{aligned}
@@
     \begin{aligned}
-        \min_{\alpha} \quad & +\frac{1}{2}\alpha^\intercal K\alpha-\alpha^\intercal y\\
+        \min_{\textrm{a}} \quad & +\frac{1}{2}\textrm{a}^\intercal K\textrm{a}-\textrm{a}^\intercal y\\
     \textrm{s.t.} \quad & 
-    C(q-1)\mathbb{1}\preceq \alpha \preceq Cq\mathbb{1}\\
-    &\alpha^\intercal\mathbb{1}=1
+    C(q-1)\mathbb{1}\preceq \textrm{a} \preceq Cq\mathbb{1}\\
+    &\textrm{a}^\intercal\mathbb{1}=1
     \end{aligned}
@@
 \begin{equation}
-    f(x)=\sum\limits_i \alpha_i k(x_i, x)+b
+    f(x)=\sum\limits_i \textrm{a}_i k(x_i, x)+b
 \end{equation}
 Since our optimisation problem possesses strong duality and it is differentiable in both the objective and the constraint, we have that it must satisfy the Karush Kuhn Tucker conditions, see section 5.5.3 \cite{boyd2004convex}.
-Thanks the KKT conditions on the primal optimisation problem we have that $f(x_i)=y_i \ \mathrm{for} \ \alpha_i \not \in \{C(q-1), Cq\}$. 
+Thanks to the KKT conditions on the primal optimisation problem we have that $f(x_i)=y_i \ \mathrm{for} \ \textrm{a}_i \not \in \{C(q-1), Cq\}$. 
 To see this, we have to consider the KKT conditions.
@@
     \alpha_i=0, \ \alpha_i^*=C(1-q) & \implies \alpha_i-\alpha_i^*=C(q-1), \xi_i= 0, \xi^*\leq 0 \implies \xi_i^*+y_i-f_i=0 \\
-    0\leq \alpha_i\leq Cq, \ 0\leq \alpha_i^*\leq C(1-q) & \implies \xi_i=0,  \xi^*_i= 0 \implies -y_i+f_i=0 \ y_i-f_i=0 \\
+    0\leq \alpha_i\leq Cq, \ 0\leq \alpha_i^*\leq C(1-q) & \implies \xi_i=0,  \xi^*_i= 0 \implies -y_i+f_i=0, \ y_i-f_i=0 \\
 \end{cases}
@@
 
-Therefore in order to retrieve b we simply have to choose an index $i$ such that bold $\alpha_i \not \in \{C(q-1), Cq\}$
+Therefore in order to retrieve b we simply have to choose an index $i$ such that  $\textrm{a}_i \not \in \{C(q-1), Cq\}$
 and let
@@
     \begin{aligned}
-    b=&y_i-\sum_i \alpha_i k(x_i,x)
+    b=&y_i-\sum_i \textrm{a}_i k(x_i,x)
     \end{aligned}
@@
 \\
-The dataset used is the Melbourne Daily Maximum Temperatures \cite{hdrcde}.
+The dataset used is the Melbourne daily maximum temperatures \cite{hdrcde}.
 It contains the daily maximum temperatures in Melbourne, Australia, from 1981-1990, excluding leap days, see figure \ref{fig:melbourne_temperature_data}.
@@
 The observed bimodality is that a hot day is likely to be followed by either an equally hot day or one much cooler.
-\begin{figure}
+\begin{figure}[!h]
     \includegraphics[width=\textwidth]{images/melbourne_temperature.png}
@@
 Hyperparameter tuning have been optimised through cross validation, see \ref{appendix:cross_validation}.
-\begin{figure}
+\begin{figure}[!h]
     \begin{subfigure}[b]{0.5\linewidth}
@@
 % tables
-\begin{table}
+\begin{table}[!h]
 \caption{Pinball loss Melbourne data}
@@
 
-\begin{table}
+\begin{table}[!h]
     \caption{Pinball loss quantile-wise Melbourne data}
@@
 \end{table}
-
 As already pointed out, the quantile regression with q=0.5 corresponds to the standard regression problem, hence we can compare the proposed methods also in terms of the mean absolute error.
-
-\begin{table}
+\begin{table}[!h]
 \caption{MAE Melbourne data}
@@
 Not only kernel quantile regression was the best in terms of total pinball loss but also the best in terms of each quantile pinball loss and mean absolute error.
-Comparison has been carried out on further dataset, yielding similar conclusion as the one of above, to know more see \ref{appendix:quantile_regressor_extensive_comparison}.
+Comparison has been carried out on further datasets, yielding similar conclusion as the one of above, to know more see \ref{appendix:quantile_regressor_extensive_comparison}.
 

### 2024-03-27 thesis/input_files/qr.tex
March 27th 2024

@@
 In this setting, you basically slice the dependent variables into quantiles and then fit a regression for each quantile.
-With standard regression we build a model for the conditional mean, conversely, with quantile regression we model the conditional quantile function for any desired quantile.
+With standard regression, we build a model for the conditional mean, conversely, with quantile regression we model the conditional quantile function for any desired quantile.
 Therefore, with quantile regression we are able to study the impact of covariates on quantiles directly.
@@
     \begin{equation}
-        Q(q)=\inf{y:F(y)\geq q}
+        Q(q)=\inf\{y:F(y)\geq q\}
     \end{equation}
 \end{definition}
-Alternatively, in order  to ease the posing of the quantile regression problem, we can formulate quantiles as the solution to a simple optimization problem.
+Alternatively, in order  to ease the posing of the quantile regression problem, we can formulate quantiles as the solutions to simple optimization problems.
 \\
-For any $0<q<1$ consider the pinball loss function \ref{pinball} $\rho_q(u)=u(q-\mathbb{1}_{u<0})$. 
-Such loss is minimized by the quantiles $Q(q)$.
-We can estimate them by minimizing the expectation of $\rho_q(Y-g(x,\beta))$ with respect to the parameter $\beta$.
+For any $0<q<1$ consider the pinball loss function from section \ref{pinball}, $\rho_q(u)=u(q-\mathbb{1}_{\{u<0\}})$. 
+Such loss is minimized by the quantile $Q(q)$.
+Thus,we can estimate quantiles by minimizing the expectation of $\rho_q(Y-g(x,\beta))$ with respect to the parameter $\beta$.
 \\
@@
 \\
-It follows that the conditional linear quantile function $Q_Y(q|X=x)=x_i\beta(q)$ can be estimated by solving
-\begin{equation}\label{}
-    \hat{\beta(q)}=\argmin{\beta}\sum \rho_q   (y_i-x \beta)
+It follows that the conditional linear quantile function $Q_Y(q|X=x)=x_i\beta(q)$, can be estimated by solving
+\begin{equation}\label{eq:linear_quantile_regression_minimisation_problem}
+    \hat{\beta}(q)=\argmin{\beta}\sum \rho_q   (y_i-x \beta)
 \end{equation}
 \\
-Notice that, the cost function is not differentiable, therefore there is no analytical solution to the quantile regression problem.
+Notice that, this cost function is not differentiable, therefore there is no analytical solution to the quantile regression problem.
 Nevertheless, we can easily solve it by employing linear programming and convex optimisation \cite{boyd2004convex}.
 \\
-Furthermore, can extend to non linear quantile regression by choosing a non linear model in place of $x\beta$ in the above equation.
-\\
-In order to illustrate the approach, let us consider applying quantile regression to the maximum daily temperature dataset in Melbourne \cite{hyndman1996estimating}; models considered will be linear, gradient boosting machine, quantile forest\cite{meinshausen2006quantile} and kernel quantile regression.
-- se si puo fare plot con stessi dati ma differenti f nel quantile regression
+Furthermore, we can extend this framework to non linear quantile regression by choosing a non linear model in place of $x\beta$ in the above equation \ref{eq:linear_quantile_regression_minimisation_problem}.
+% \\
+% In order to illustrate the approach, let us consider applying quantile regression to the maximum daily temperature dataset in Melbourne \cite{hyndman1996estimating}; models considered will be linear, gradient boosting machine, quantile forest\cite{meinshausen2006quantile} and kernel quantile regression.
+% - se si puo fare plot con stessi dati ma differenti f nel quantile regression

### 2024-03-27 dataset_applications/bash3.sh

@@ -1,4 +1,5 @@
 #!/bin/bash
+# script to apply pickled model to each task
 
 # clean
 rm /Users/luca/Desktop/ThesisKernelMethods/thesis/tables/qr_gefcom2014.txt

### 2024-03-27 dataset_applications/qr_gefcom2014_plot.py

@@ -1,0 +1,2 @@
+# take data saved as predictions 5, 50, 95 quantile
+# plot 

### 2024-03-27 dataset_applications/qr_gefcom2014_predictions.py

@@ -1,0 +1,3 @@
+# save data as csv
+# from pd dataframe with rows timestamp
+# columns quantile

### 2024-03-29 thesis/input_files/energymarket.tex
6 march
decision trees
March 29th 2024

@@
-Chapter explaining the energy market
+% Chapter explaining the energy market
+% Chapter 9 of https://pubs.naruc.org/pub.cfm?id=536E10A7-2354-D714-5191-A8AAFE45D626
 
-- How auctions work
+This chapter has two intents.
+Firstly, it summarizes several concepts for the newcomers to the field of electricity markets.
+Secondly, it discusses trends and challenges undergoing in the power sector.
+Such developments motivate the need for robust and efficient forecasting techniques in the context of electricity markets.
 
-- Difference intraday dayhead all other stuff
+\section{Electricity distribution network}
+Electricity is generated by power plants which transfer it over the so called transmission level.
+Then, through the transmission network, this energy is transported at a high voltage over long distances.
+Finally, the voltage is reduced and the energy is moved into the distribution network.
+In a nutshell, transmision lines carry electric power from stations to substations while distribution lines carry electricity from substations to load points such as businesses, industries and homes, see figure \ref{electricity_network}.
 
-- Prices can be negative
+\begin{figure}[!h]
+    \includegraphics[width=\textwidth]{images/electricity_network.png}
+    \caption{Electricity grid \cite{haben2021review}}
+    \label{electricity_network}
+\end{figure}
+
+\section{Electricity markets}
+Electricity markets have gone rapid changes all over the world during the last three decades.
+Before this revolution, the power sector was a natural monopoly.
+In particular, it was characterized by a high vertical integration and by very little market competition.
+Better technologies in both transmission, generation and distribution are the reasons for such liberal shift.
+The rationale is simple, a competitive market promotes efficiency and stimulates technical innovation more than a monopoly does.
+\\
+During the nineties various wholesale electricity markets were established; for example in Chile, Great Britain, the scandinivian block, Australia, New Zealand, the US and Canada.
+Consequently, new participants each with a specific taks act entered the power market scenario.
+\begin{itemize}
+    \item Electricity generators produce electricity to be consumed.
+    \item Electricity transmission owners 
+    move high voltage electricity to electric utility companies whilst ensuring safe and reliable supply.
+    \item Electricity grid operators are responsible for scheduling electricity over the transmission network in order to ensure supply demand balance.
+    \item Electric utilities deliver power over local lines to consumers. 
+    \item Retail energy suppliers purchase electricity in the wholesale market from electricity generators and resell it to consumers.
+\end{itemize}
+
+\subsection{The marketplace}
+We can differential between two kinds of electricity markets: power pools and power exchanges.
+In power pools, generators bid the prices at which they are willing to produce at different volumes.
+Then, the market clearing price MCP is determined by intersecting the aggregated supply curve and the estimated demand, left panel figure \ref{fig:pool_vs_echange}.
+Power pools are created on public initiatives of governments.
+Conversely, power exchanges are created through private agreements between generators, distributors and traders.
+This is the model followed by most of european countries.
+The MCP in power exchanges is determined by the intersection of the aggregate supply curve and the aggregate demand curve, right panel figure \ref{fig:pool_vs_echange}.
+\\
+It is worth to mention the two biggest power exchanges: Epex and NordPool. Epex operates throughout continental europe. NordPool covers the nordinc and baltic contries.
+\\
+Moreover, we can differentiate between two popular types of auctions: uniform-price/marginal and pay as bid/discriminatory.
+Within the uniform price setting, suppliers offering less than the clearing price are paid that price.
+Analogously, consumers bidding more than the clearing price pay that price.
+On the other hand, in pay as bid auctions, suppliers are paid the exact price they bid for.
+
+\begin{figure}[!h]
+    \includegraphics[width=\textwidth]{images/pool_vs_echange.png}
+    \caption{Pool market vs power echange \cite{weron2006modeling}}
+    \label{fig:pool_vs_echange}
+\end{figure}
+
+\subsection{How auctions work}
+We can drill down on the auction types, the two most popular forms are day ahead and intraday auctions.
+% https://medium.com/@brandonvar/day-ahead-and-intraday-electricity-markets-1eb121dcab47
+\subsection{Day ahead}
+The day ahead market is based on a blind auction taking place every day of the year.
+During these auctions, the prices of all hours for the next day are traded.
+Market participants can enter their buy sell orders until order book closure, which takes place at 12.00 PM.
+After that, the market establishes hourly prices by intersecting the demand and supply curves for each hour of the following day.
+
+\subsection{Intraday}
+The intraday market consists of continous trading 24/7.
+A trade is executed as soon as a buy and a sell order mathch.
+Electricity can be traded an hour, half-hour, quarter or also 5 minutes before delivery.
+This enables market participants to balance their positions in real time, should they need to do so.
+
+\subsection{Price peaks}
+Electricity market are characterized by spikes in their spot prices; when this happens, the system price jumps abruptly and then drops back within a very short period.
+This spikness follows from the non storability of electricity; that is, electricity has to be consumed as it is produced.
+Therefore, extreme load fluctuations combined with generation outages or transmission failures can result in price spikes.
+
+% outages=interruzione
+\subsection{Negative prices}
+It is not unusual to observe negative prices in electricity markets, even though they are rare.
+The causes of negative prices are inflexible power generation plants and low demand.
+With inflexibility, it is mean the fact that power sources cannot be switched off and restarted quickly and efficiently.
+Thus, producers are faced with the decision of either stopping and restarting their power plant or selling their energy for a negative price; that is they pay consumers for consuming their energy.
+
+\subsection{Nodal and Zonal Pricing}
+Grid configuration and physical limits on electricity lines may lead to congestion.
+Local marginal price (LMP) and zonal market clearing pricing (ZMCP) are two types of pricing schemes utilized to handle it.
+The former is adopted in EU contries while the latter is used in the US, figure \ref{fig:eu_zonal} and figure \ref{fig:us_local}.
+With ZMCP, prices may differ between zones but are the same within the same area.
+LMP is made up by summing the transmission congestion cost, generation marginal cost and the cost of marginal losses at different buses; buses is where a electicity line or several lines are connected.
+\begin{figure}[!h]
+    \includegraphics[width=\textwidth]{images/eu_zonal.png}
+    \caption{Zonal pricing EU \cite{eu_zonal}}
+    \label{fig:eu_zonal}
+\end{figure}
+
+\begin{figure}[!h]
+    \includegraphics[width=\textwidth]{images/us_local.png}
+    \caption{Local pricing US \cite{us_local}}
+    \label{fig:us_local}
+\end{figure}
+
+% pensodino \subsection{Nodal and Zonal Pricing}
+\section{Smart grids}
+A smart meter is a digital device recording energy consumption in real time and communicating this data to the supplier and the consumer.
+Deployment of smart grids and the digitalisation of the electricity network are intended to increase the energy efficiency of our networks.
+The reason is that, smart metering and the electricity network digitalisation  will enable system operators to better monitor the network, plan their investments and manage infrastructure. 
+All of this motivates for the need of efficient and robust tools to analyze the vast data resulting from the digitalisation of the power sector.
+
+\section{Renewables}
+% https://www.pnnl.gov/explainer-articles/renewable-integration
+A pressing challenge facing policymakers today is the energy transition towards cleaner energy, see figure \ref{fig:electricity_production_by_source} for a breakdown on energy production by source.
+Such move is intended to mitigate the effects of climate change and reduce pollution at the same time.
+This transition passes through renewables integration into the electric grid. 
+Nevertheless, integrating renewables into the existing electric grid involves several technical challenges.
+First, we need to develop infrastructures and technologies capable of connecting renewable power plants with the existing grid.
+For example, connecting renewables power plants located in remote and offshore locations to high voltage powerlines is not trivial.
+Additionally, renewables such as wind and solar are characterized by seasonalities and depend heavily on weather conditions.
+Therefore, balancing the electricity network and mantaining stability will become harder for grid operators; they will need more flexibility in the grid and develop new approaches in order to achieve their goals.
+
+\begin{figure}[!h]
+    \includegraphics[width=\textwidth]{images/electricity-production-by-source.png}
+    \caption{Electricity production by source}    
+    \label{fig:electricity_production_by_source}
+\end{figure}

### 2024-03-29 thesis/input_files/symbols.tex
March 20th 2024
decision trees
March 22nd 2024
March 27th 2024
March 29th 2024

@@
 %L
+\glsxtrnewsymbol[description={Load marginal price}]{LMP}{LMP}
 \glsxtrnewsymbol[description={Low carbon technologies}]{LCTs}{LCTs}
@@
 %M
+\glsxtrnewsymbol[description={Market clearing price}]{MCP}{MCP}
+\glsxtrnewsymbol[description={Market clearing volume}]{MCV}{MCV}
 \glsxtrnewsymbol[description={Mean absolute error}]{MAE}{MAE}
@@
 %P
+\glsxtrnewsymbol[description={Power exchange}]{PX}{PX}
 \glsxtrnewsymbol[description={Prediction interval}]{PI}{PI}
@@
 
+\glsxtrnewsymbol[description={Sum of squared errors}]{SSE}{SSE}
+
+\glsxtrnewsymbol[description={Support vector machines}]{SVMs}{SVMs}
+
 \glsxtrnewsymbol[description={Sustainable development goals}]{SDGs}{SDGs}
-\glsxtrnewsymbol[description={Support vector machine}]{SVM}{SVM}
 
@@
 
+% Z
+\glsxtrnewsymbol[description={Zonal market clearing price}]{ZMCP}{ZMCP}
+
 %MATHEMATICAL NOTATION
+% D
+\glsxtrnewsymbol[description={dual optimum}]{d^*}{\ensuremath{d^*}}
+
+%E
+\glsxtrnewsymbol[description={expectation}]{mathbb{E}}{\ensuremath{\mathbb{E}}}
+
+% F
+\glsxtrnewsymbol[description={feature matrix}]{Phi(x)}{\ensuremath{\Phi(x)}}
+
+\glsxtrnewsymbol[description={feature space}]{F}{\ensuremath{F}}
+
 \glsxtrnewsymbol[description={feature vector}]{phi(x)}{\ensuremath{\phi(x)}}
 
+% G
+\glsxtrnewsymbol[description={dual function}]{g(v)}{\ensuremath{g(v)}}
+
+% I
+\glsxtrnewsymbol[description={Indicator function}]{mathbb{I}}{\ensuremath{\mathbb{I}}}
+
+% K
+\glsxtrnewsymbol[description={kernel matrix}]{K}{\ensuremath{K}}
+
 \glsxtrnewsymbol[description={kernel function}]{k(x,x')}{\ensuremath{k(x,x')}}
+
+% L
+\glsxtrnewsymbol[description={lagrangian}]{L}{\ensuremath{L}}
+
+% P
+\glsxtrnewsymbol[description={pinball loss}]{rho_q}{\ensuremath{\rho_q}}
+
+\glsxtrnewsymbol[description={primal optimum}]{p^*}{\ensuremath{p^*}}
+
+% R
+\glsxtrnewsymbol[description={relative interior}]{relint}{\textrm{relint}}
+
+\glsxtrnewsymbol[description={reproducing kernel hilbert space}]{H}{\ensuremath{\mathcal{H}}}
+
+% V
+\glsxtrnewsymbol[description={vector of ones}]{1}{\ensuremath{\mathbb{1}}}

### 2024-03-29 dataset_applications/bash4.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+
+# code to extract the timestamp and zoneid from test data so that the prediction is in the same format of guidelines
+
+
+for((i=1;i<=15;++i))
+do
+    cut -d , -f 1-2 /Users/luca/Desktop/GEFCom2014\ Data/Load/Task\ $i/L$i-test.csv >> /Users/luca/Desktop/GEFCom2014\ Data/Load/Task\ $i/L$i-timestamp.csv
+done

### 2024-03-29 dataset_applications/qr_gefcom2014_plot.py

@@ -1,2 +1,0 @@
-# take data saved as predictions 5, 50, 95 quantile
-# plot 

### 2024-03-29 dataset_applications/qr_gefcom2014_plot_task1.py

@@ -1,0 +1,55 @@
+# code to create plot of 90% interval
+
+# take data saved as predictions 5, 50, 95 quantile from each task,
+# vertical stack data
+# plot 
+
+from datetime import datetime
+import numpy as np
+import matplotlib.pyplot as plt
+import pandas as pd
+
+if __name__=="__main__":
+    
+    # load test task 1
+    y_test=pd.read_csv("/Users/luca/Desktop/GEFCom2014 Data/Load/Task 1/L1-test_clean.csv")
+    
+    
+    # load prediction for task 1
+    y_predict_q=pd.read_csv("/Users/luca/Desktop/GEFCom2014 Data/Load/Task 1/L1-our_model_prediction.csv")
+
+    # select quantiles
+    y_predict_5=y_predict_q["0.05"]
+    y_predict_50=y_predict_q["0.5"]
+    y_predict_95=y_predict_q["0.95"]
+
+    # dummy x for plotting
+    x=np.arange(len(y_test))
+    # ticks for the days, (including also hours adds clutter)
+    tick_array = [i for i in range(24,744+1,24)]
+    # label ticks
+    labels=[int(i/24) for i in tick_array]
+    
+    # plot quantile range
+    plt.plot(y_test["LOAD"],color="black")
+    
+    plt.fill_between(x,y_predict_q["0.05"],y_test["LOAD"], alpha=0.4, color="green", edgecolor="red")
+    plt.fill_between(x,y_test["LOAD"],y_predict_q["0.95"], alpha=0.4, color="green", edgecolor="red")
+    
+    plt.xticks(ticks=tick_array, labels=labels, rotation=45)
+    
+    plt.xlabel("Days")
+    plt.ylabel("Load in MW")
+    plt.show()
+
+
+    # plot true vs median
+    plt.plot(x,y_test["LOAD"],color="black", label="y_test")
+    plt.plot(x,y_predict_50, "red", alpha=0.4, label="0.5 prediction")
+    plt.xticks(ticks=tick_array, labels=labels, rotation=45)
+    
+    plt.xlabel("Days")
+    plt.ylabel("Load in MW")
+    plt.legend()
+    plt.show()
+

### 2024-03-29 dataset_applications/qr_gefcom2014_predictions.py

@@ -1,3 +1,50 @@
-# save data as csv
+# code to save predictions in the same format of GEFCom challenge
+# ZONEID| TIMESTAMP| 0.01|...|0.99
+
+# save data as csv, each task in its folder
 # from pd dataframe with rows timestamp
 # columns quantile
+
+
+
+import pandas as pd
+import pickle
+
+from sklearn.preprocessing import StandardScaler
+import sys
+
+from tqdm import tqdm
+
+
+def predict_save(ith):
+    df=pd.read_csv(f"/Users/luca/Desktop/GEFCom2014 Data/Load/Task {ith}/L{ith}-test_clean.csv")
+
+    X_test=df[["DAY",  "MONTH",  "HOUR",  "DAY_OF_WEEK",  "IS_HOLIDAY",  "w_avg"]]
+    y_test=df["LOAD"]
+
+    quantiles = [i/100 for i in range(1,100)]
+
+    # we need to scale X_test
+    df_train=pd.read_csv("/Users/luca/Desktop/GEFCom2014 Data/Load/L2009-train.csv")
+    X_train=df_train[["DAY",  "MONTH",  "HOUR",  "DAY_OF_WEEK",  "IS_HOLIDAY",  "w_avg"]]
+    scaler = StandardScaler()
+    X_train_scaled = scaler.fit_transform(X_train)
+    X_test_scaled = scaler.transform(X_test)
+
+    
+    df_template_submission=pd.read_csv(f"/Users/luca/Desktop/GEFCom2014 Data/Load/Task {ith}/L{ith}-test.csv")
+    output_df=df_template_submission[["ZONEID", "TIMESTAMP"]].copy()
+    
+    for i,q in enumerate(quantiles):
+        krn_q=pickle.load(open(f"/Users/luca/Desktop/ThesisKernelMethods/dataset_applications/models/krn_qr_{i}.pkl", 'rb'))
+        y_predict_q=krn_q.predict(X_test_scaled)
+
+        # save prediction for qth quantile
+        output_df[f"{q}"]=pd.Series(y_predict_q)
+
+    output_df.to_csv(f"/Users/luca/Desktop/GEFCom2014 Data/Load/Task {ith}/L{ith}-our_model_prediction.csv", index=False)
+    return
+
+if __name__=="__main__":
+    i=int(sys.argv[1])
+    predict_save(i)

### 2024-03-31 thesis/input_files/point.tex
March 29th 2024
March 31st 2024

@@
 Besides discussing the theory underlying these models, this and the following chapter include also a couple of worked examples in order to get acquainted with the practical applications of models.
-\section{Exponential smoothing}
+
+% \section{Exponential smoothing}
+% Exponential smoothing was developed by Brown \cite{brown1959statistical}, Holt \cite{holt1957forecasting}, Winters \cite{winters1960forecasting}.
+% The idea is to forecast future values by taking a weighted average of past observations, where weights decay exponentially in time.
+% \begin{equation}
+%     \begin{aligned}
+%         \hat{y}_{t+h|t}=& l_{t}+hb_{t}+s_{t+h-m(k+1)}
+%         \\
+%         l_t=&\alpha (y_t-s_{t-m})+(1-\alpha)(l_{t-1}+b_{t-1})
+%         \\
+%         b_t=& \beta(l_t-l_{t-1})+(1-\beta)b_{t-1}
+%         \\
+%         s_t=&\gamma(y_t-l_{t-1}-b_{t-1})+(1-\gamma)s_{t-m}
+%     \end{aligned}
+% \end{equation}
+% where $0\leq \alpha \leq 1$ is the smoothing parameter and k is the integer part of $\frac{(h-1)}{m}$.%that is k encodes seasonalities
+% Optimal parameters are obtained by minimising the sum of squared errors.
 
 \section{Multiple linear regression}
+Notwithstanding, its simplicity, multiple linear regression models are still popular among the EPF literature. Notice, they are no used per se but combined with other more advanced models.
+\begin{equation}
+    L_t=\beta X_t + \epsilon_t
+\end{equation}
 
 \section{Autoregressive models}
-explain their theory
-and how the procedure how they are used
-IMPORTANT: load series is not a stationary series so before applying AR we have to perform stationary tests or differencing steps
+% explain their theory
+% and how the procedure how they are used
+% IMPORTANT: load series is not a stationary series so before applying AR we have to perform stationary tests or differencing steps
+Autoregressive models are a standard approach for modelling time series data.
+Before introducing the popular model in EPF, it is important to remember that autoregressive models assume the time series to be stationary.
+On the other hand, load and price have been observed to be a non stationary time series.
+Basically, stationarity means that the distribution of any subsequence of random variables of the stochatic process is invariant to shift along the time dimension.
+\\
+Therefore, it is essential to check stationary through unit root tests.
+Moreover, should the time series be non stationary, we have to convert it to a stationary one by a combination of either differencing, detrending and log transforming.
 \subsection{ARIMA}
+ARIMA(p,d,q) models are of the form
+\begin{equation}
+    \hat{L}_N=\sum\limits_{i=1}^{p}\phi_i L_{N-i}^{(d)}+\sum\limits_{j=1}^{q}\theta_j \epsilon_{N-j}+\epsilon_N
+\end{equation}
+where parameters p is the order of the AR part, d is the degree of integrated differencing and q is the order of the MA model.
+$\epsilon_{N-j}$ are the observed errors in the past while the differencing degree $(d)$ is defined recursively as $L_N^{(d)}=L_N^{(d-1)}-L_{N-1}^{(d-1)}$.
+ARIMA model are usually written as
+\begin{equation}
+    \phi(B)\nabla^d y_t=\theta(B)\epsilon_t
+\end{equation}
+where $B$ is the backward shift operator, $B^h y_t=y_{t-h}$ and $\phi(B)$ is shorthand for $\phi(B)=1-\phi_1B-\dots-\phi_p B^p$.
+Similarly $theta(B)=1+\theta_1B+\dots+\theta_q B^q$ and $\nabla_h$ is the lag h differencing operator $\nabla_h y_t=(1-B^h)y_t=y_t-y_{t-h}$
+\\
+Box-Jenkins is the method for estimating such models \cite{box2015time}.
+The order p is identified by taking the lags at which the sample partial autocorrelation function (PACF) falls outside its 95\% confidence interval.
+In the same way, the order q is estimated by consider the sample autocorrelation function of the observed time series.
+Finally, we compare ARIMA models with different p,d,q (in a neighbourhood of the identified orders) and choose the ones minimising either AIC or BIC.
+Notice, models are obtained through the maximum likelihood estimation framework.
+
 \subsection{ARIMAX}
-\subsection{SARIMA}
-\subsection{SARIMAX}
+ARIMAX model add explanatory variables to our ARIMA model, we have
+\begin{equation}
+    \hat{L}_N=\sum\limits_{k=1}^{h} \mu_i X_{N-k}+\sum\limits_{i=1}^{p}\phi_i L_{N-i}^{(d)}+\sum\limits_{j=1}^{q}\theta_j \epsilon_{N-j}+ \epsilon_{N}
+\end{equation}
+\subsection{SARIMA and SARIMAX}
+SARIMA and SARIMAX incorporate seasonalities into our time series models.
+The notation for these models is ARIMA$(p,d,q)(P,D,Q)_m$, the second terms represents the autoregressive structure of the seasonal patter.
+\\
+\begin{equation}
+    \phi(B)\Phi(B^s)\nabla^d\nabla_m^Dy_t=\theta(B)\Theta(B^s)\epsilon_t
+\end{equation}
+\\
+For instance, in modelling series of hourly data with daily periodicity, we set m=24.
 
 \section{Generalized additive model}
+% in standard regression the predicted y are normal since error are normal.
+But if we want to model say a y bernoulli which is the rate of success we can link the domain of bernoulli to that of gaussian with a link function
+Generalized additive models take the form
+\begin{equation}
+    g[\mathbb{E}(Y_i)]=\alpha+f_1(X_1)+\dots+f_p(X_p)
+\end{equation}
+$f_i$ are smooth functions which are fitted using cubic smoothing splines.
+g is called link function and characterizes the specific GAM model \cite{hastie2017generalized}. Approximations for $f_i$ are obtained through an iterative procedure, the backfitting algorithm.
+
 %prophet meta
-\section{K-nearest neighbors}
+Particularly successful in time series forecasting is the Prophet model \cite{taylor2018forecasting}.
+This additive model is capable of handling non linear trends, holiday effects and yearly weekly and daily seasonality.
+\section{K-nearest neighbours regression}
+K-nearest neighbours regression forecasts by averaging the most similar k instances in the training set \cite{macqueen1967some}.
+K is the algorithm hyperparameter, a small k leads to overfitting while a big k leads to underfitting.
+Since it is a metric based algorithm it is important to normalize it in order to give equal weights to features with different scales \ref{appendix:normalization}.
+The algorithm is made up of two decision: first, the choice of the metric and second, the method for combining targets.
 
@@
 \section{Artificial neural networks}
-\subsection{DNN}
+Artificial neural networks (ANN) have been successfully applied in the context of electricity markets forecasting.
+The basic building block of this class of methods is the perceptron \cite{rosenblatt1958perceptron}.
+\begin{equation}
+    y=\psi(wx+b)
+\end{equation}
+where $\psi$ is an activation function and w,b are the parameters of the neuron; such parameters are typically optimized through gradient descent.
+
+\subsection{MLP}
+In order to cover a richer space of models we can stack neurons, doing so we obtain the so called multi layer perceptrons (MLPs).
+Hyperparameters of these models are the number of hidden layers, the number of neurons in each of those layers and the kind of activation function.
+Notice, depending on the numbers of layers, MLPs are sometimes referred to as deep neural networks.
+
+\subsubsection{NARX}
+Neural networks that have as inputs past values of the target variable and other independent variables are called nonlinear autoregressive exogenous models.
+
 \subsection{LSTM}
-\subsection{DeepAr}
-\subsection{AR Net}
+% https://colah.github.io/posts/2015-08-Understanding-LSTMs/
+
+% https://rtavenar.github.io/deep_book/book_en.pdf
+LSTM extend RNN by introducing a cell state in its layers.
+In this way, the current state of a cell depndes on the current value $X_t$, on the previous activation and on the previous cell state $C_{t-1}$, see figure \ref{fig:lstm_cell} for a visual representation.
+\begin{figure}[!h]
+    \includegraphics[width=\textwidth]{images/lstm_cell.png}
+    \caption{LSTM cell}
+    \label{fig:lstm_cell}
+\end{figure}
+The cell is responsible for deciding whether to store/forget long and short term values. To do so, each cell is made of a forget gate, an input gate and an output gate.
+This forget gate layer takes $h_{t-1}$ and $x_t$, and outputs a number between 0 and 1, this corresponds to ratio of information retention for each element of the cell state $C_{t-1}$.
+\begin{equation}
+    f_t=\sigma(W_f\dot[h_{t-1},x_t]+b_f)
+\end{equation}
+The input gate layer is made up of two parts.
+First, a sigmoid layer decides which values of the cell state will be updated.
+Second, a tahn layer creates a vector of candidates $\tilde{C}_t$ to add to the current state.
+\begin{equation}
+    \begin{aligned}
+        i_t=&\sigma(W_i\dot[h_{t-1},x_t]+i_f)
+        \\
+        \tilde{C}_t=&tahn(W_c\dot[h_{t-1},x_t]+b_C)
+    \end{aligned}
+\end{equation}
+The cell state is then updated by
+\begin{equation}
+    C_t=f_t\odot C_{t-1}+i_t\odot \tilde{C}_t
+    % odot is elementwise product
+\end{equation}
+Finally we output the current cell state and the tanh activated cell state filtered by the sigmoid layer on the current input and the previous hidden state.
+\begin{equation}
+    \begin{aligned}
+    o_t=& sigma(W_o[h_{t-1},x_t]+b_o)
+    \\
+    h_t=& o_t \odot tahn(C_t)
+\end{aligned}
+\end{equation}
+% \subsection{DeepAr} deepar é probabilistic
+
+% \subsection{AR Net} no é un framework non é un metodo
 
 \section{Kernel methods}
-\subsection{Kernel regression}
+\subsection{Kernel ridge regression}
+In the setting of kernel ridge regression, we aim at estimating $f$ such that $y=f(x)+\epsilon$. Given a RKHS $\mathcal{H}$, we can estimate$\hat{f}$ by solving the optmisation problem
+\begin{equation}\label{eq:kernel ridge}
+    \hat{f}=\argmin{f\in \mathcal{H}}\frac{1}{2}\sum\limits_{i=1}^n(y_i-f(x_i))^2+\frac{\lambda}{2}\|f\|_{\mathcal{H}}^2
+\end{equation}
+Before proceeding, we need to introduce the representer theorem.
+\begin{theorem}
+    Let $k:X \times X \mapsto R$ on a non empty set $X$ with corresponding RKHS $\mathcal{H}$ and consider the regularized risk functional $Risk[f]:f \mapsto  E\left((x_{1},y_{1},f(x_{1})),\ldots ,(x_{n},y_{n},f(x_{n}))\right)+g\left(\| f\| \right)$.
+Then any minimizer of the empirical risk $Risk[f]$ can be represented as $ f^{*}(\cdot )=\sum _{i=1}^{n}\alpha _{i}k(\cdot ,x_{i})$
+\end{theorem}
+
+\begin{proof}
+    By orthogonal projection, any $f \in \mathcal{H}$ can be decomposed as the sum of two functions, with the first lying in the $\textrm{span}\{\phi(x_1),\dots, \phi(x_n)\}$ and the second staying in the orthogonal complement.
+    \begin{equation}
+        f=\sum\limits_{i=1}^{n}\alpha_i \phi(x_i)+\nu
+    \end{equation}
+    Applying $f$ to any point, by the reproducing property we have
+    \begin{equation}
+        \begin{aligned}
+            f(x_j)=&\langle \phi(x_j), \sum\limits_{i=1}^{n}\alpha_i \phi(x_i)+\nu\rangle
+            \\
+            =&\sum\limits_{i=1}^{n}\alpha_i\langle \phi(x_j), \phi(x_i)\rangle
+        \end{aligned}
+    \end{equation}
+    This means that setting $\nu=0$ does not affect $f$.
+    Next consider the regularization term; notice, the functional $g$ has to be strictly monotonic in order to penalize more complex functions $f$.
+    \begin{equation}
+        \begin{aligned}
+            g(\|f\|)=&g(\|\sum\limits_{i=1}^n \alpha_i \phi(x_i)+v\|_\mathcal{H})
+            \\
+            =&g\left(\sqrt{\|\sum\limits_{i=1}^n \alpha_i \phi(x_i)+v\|_\mathcal{H}^2}\right)
+            \\
+            =&g\left(\sqrt{\|\sum\limits_{i=1}^n \alpha_i \phi(x_i)\|_{\mathcal{H}}^2+\|v\|_\mathcal{H}^2+2\langle\sum\limits_{i=1}^n \alpha_i \phi(x_i),\nu \rangle_{\mathcal{H}}}\right)
+            \\
+            =&g\left(\sqrt{\|\sum\limits_{i=1}^n \alpha_i \phi(x_i)\|_{\mathcal{H}}^2+\|v\|_\mathcal{H}^2}\right)
+            \\
+            \geq &g\left(\sqrt{\|\sum\limits_{i=1}^n \alpha_i \phi(x_i)\|_{\mathcal{H}}^2}\right)
+        \end{aligned}
+    \end{equation}
+    We can conclude setting $\nu=0$ decreases the second terms while the first is not affected.
+Thus it follows that any minimizer of the risk functional \ref{eq:kernel ridge} must be of the following form
+    \begin{equation}
+        f^*(\cdot)=\sum\limits_{i=1}^n \alpha_i \phi(x_i)=\sum\limits_{i=1}^n \alpha_i k(\cdot, x_i)
+    \end{equation}
+\end{proof}
+
+Employing the representer theorem, we can rewrite \ref{eq:kernel ridge} in matrix notation as
+\begin{equation}
+    \begin{aligned}
+    \hat{\alpha}=&\argmin{\alpha \in R^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\lambda}{2}\left\|\sum _{i=1}^{n}\alpha _{i}k(\cdot ,x_{i})\right\|_{\mathcal{H}}^2
+    \\
+    \hat{\alpha}=&\argmin{\alpha \in R^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\lambda}{2} \langle \sum _{i}^{n}\alpha _{i}k(\cdot ,x_{i}),\sum _{j=1}^{n}\alpha _{j}k(\cdot ,x_{j})\rangle_{\mathcal{H}}^2
+    \\
+    \hat{\alpha}=&\argmin{\alpha \in R^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\lambda}{2} \sum _{i,j=1}^{n}\alpha _{i}\alpha _{j}\langle k(\cdot ,x_{i}),k(\cdot ,x_{j})\rangle_{\mathcal{H}}^2
+    \\
+    \hat{\alpha}=&\argmin{\alpha \in R^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\lambda}{2} \sum _{i,j=1}^{n}\alpha _{i}\alpha _{j}K_{ij}
+    \\
+    \hat{\alpha}=&\argmin{\alpha \in R^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\lambda}{2} \alpha^\intercal K \alpha
+    \\
+    \hat{\alpha}=&\argmin{\alpha \in R^n}\frac{1}{2}(y-K\alpha)^\intercal (y-K\alpha)+\frac{\lambda}{2} \alpha^\intercal K \alpha
+\end{aligned}
+\end{equation}
+
+Setting the gradient to zero, we have
+\begin{equation}
+    -Ky+K^2\alpha+\lambda K \alpha\overset{!}{=}0
+\end{equation}
+
+We have then, that $\hat{\alpha}=(K+\lambda I)^{-1}y$. It follows the solution takes the form
+\begin{equation}
+    \hat{f}(\cdot)=\hat{\alpha}K(\cdot,:)
+\end{equation}
+
 \subsection{Kernel support vector regression}

### 2024-03-31 thesis/input_files/symbols.tex
March 31st 2024

@@
 %A
+\glsxtrnewsymbol[description={Akaike information criterion}]{AIC}{AIC}
 \glsxtrnewsymbol[description={Artificial neural network}]{ANN}{ANN}
+\glsxtrnewsymbol[description={Autocorrelation function}]{ACF}{ACF}
 \glsxtrnewsymbol[description={Autoregressive model}]{AR}{AR}
@@
 \glsxtrnewsymbol[description={Autoregressive exogenous model}]{ARX}{ARX}
+
+%B
+\glsxtrnewsymbol[description={Bayesian information criterion}]{BIC}{BIC}
 
@@
 %G
+\glsxtrnewsymbol[description={Generalized additive models}]{GAMs}{GAMs}
 \glsxtrnewsymbol[description={Gradient boosting regression tree}]{GBRT}{GBRT}
@@
 \glsxtrnewsymbol[description={Load marginal price}]{LMP}{LMP}
+\glsxtrnewsymbol[description={Long short term memory}]{LSTM}{LSTM}
 \glsxtrnewsymbol[description={Low carbon technologies}]{LCTs}{LCTs}
@@
 \glsxtrnewsymbol[description={Minimal gated memory network}]{MGM}{MGM}
+\glsxtrnewsymbol[description={Moving average}]{MA}{MA}
 
 \glsxtrnewsymbol[description={Multiple linear regression}]{MLR}{MLR}
+
+% N
+\glsxtrnewsymbol[description={Nonlinear autoregressive exogenous models}]{NARX}{NARX}
 
@@
 %P
+\glsxtrnewsymbol[description={Partial autocorrelation function}]{PACF}{PACF}
 \glsxtrnewsymbol[description={Power exchange}]{PX}{PX}

### 2024-04-06 thesis/appendix.tex
March 31st 2024
April 6th 2024

@@
     \|
-    \phi^{new}(x)\|_{\mathcal{H}}^{2} &= \|\frac{\phi(x)}{\|\phi(x)\|_{\mathcal{H}}}
-    \|_{\mathcal{H}}^{2}
+    \phi^{new}(x)\|_{\mathcal{H}}^{2} &= \left\|\frac{\phi(x)}{\|\phi(x)\|_{\mathcal{H}}}
+    \right\|_{\mathcal{H}}^{2}
     \\
     &=
-    \|
+    \left\|
     \frac{\phi(x)}
     {\sqrt{k(x,x)}}
-    \|_{\mathcal{H}}^{2}
+    \right\|_{\mathcal{H}}^{2}
     \\
@@
 
-\section{Cross validation}\label{appendix:cross_validation}
-
 \section{Quantile regressor extensive comparison}\label{appendix:quantile_regressor_extensive_comparison}
+This section contains extensive comparison between our kernel quantile regression and other state of the art quantile regressors. We benchmark it on popular machine learning datasets.
 \subsection{Boston housing dataset}
@@
 
-\section{Cross validation}
+\section{Cross validation}\label{appendix:cross_validation}
 Cross validation directly estimates the expected test error
@@
 
-\subsection{Data normalization}
+\subsection{Data normalization}\label{appendix:normalization}
 With data normalization we transform the range of the feautures to a standard scale.

### 2024-04-06 thesis/input_files/energymarket.tex
April 6th 2024

@@
 During the nineties various wholesale electricity markets were established; for example in Chile, Great Britain, the scandinivian block, Australia, New Zealand, the US and Canada.
-Consequently, new participants each with a specific taks act entered the power market scenario.
+Consequently, new participants, each with a specific task, entered the power market scenario.
 \begin{itemize}
@@
 \end{itemize}
+Notice, these functions are not mutually exclusive; that is an entity can provide one or more of these services.
 
@@
 With ZMCP, prices may differ between zones but are the same within the same area.
-LMP is made up by summing the transmission congestion cost, generation marginal cost and the cost of marginal losses at different buses; buses is where a electicity line or several lines are connected.
+On the other hand, LMP is made up by summing the transmission congestion cost, generation marginal cost and the cost of marginal losses at different buses; buses is where a electicity line or several lines are connected.
 \begin{figure}[!h]
@@
 % https://www.pnnl.gov/explainer-articles/renewable-integration
-A pressing challenge facing policymakers today is the energy transition towards cleaner energy, see figure \ref{fig:electricity_production_by_source} for a breakdown on energy production by source.
+A pressing challenge facing policymakers today is the energy transition towards cleaner energy, see figure \ref{fig:electricity_production_by_source} for an up to date breakdown on energy production by source.
 Such move is intended to mitigate the effects of climate change and reduce pollution at the same time.

### 2024-04-06 thesis/input_files/implementation.tex
March 23rd 2024
April 6th 2024

@@
 \section{Quantile gradient boosting machine}
-The implementation of sklearn.ensemble.GradientBoostingRegressor as been used.
+The implementation of sklearn.ensemble.GradientBoostingRegressor has been used.
 \section{Quantile forest}
-The implementation of quantile\_forest.RandomForestQuantileRegressor as been used. This estimator is compatible with scikit-learn.
+The implementation of quantile\_forest.RandomForestQuantileRegressor has been used. This estimator is compatible with scikit-learn.
 \section{Kernel quantile regression}
@@
 Nevertheless they are both in R, that is there exists no python, matlab or julia open source implementation.
-Following are reported the results of a comparative study between our own implementation and the one of the R library kernlab(i guess it is in C or C++ and then binded to R).
+Following are reported the results of a comparative study between our own implementation and the one of the R library kernlab
+% (i guess it is in C or C++ and then binded to R).
+\subsection{Python vs R implementation}
+In this section, a comparison study has been carried out in order to inspect the competitiveness of our implementation with the existing one for the R programming language.

### 2024-04-06 thesis/input_files/paper1.tex
April 6th 2024

@@
 
-\section{\href{https://arxiv.org/abs/1605.09522}{Kernel Mean Embedding of Distributions: A Review and Beyond}}
+\section{Kernel Mean Embedding of Distributions: A Review and Beyond}
+% \href{https://arxiv.org/abs/1605.09522}{
 From this first paper \cite{Muandet_2017}, the notation and terms used in the theory of Reproducing Kernel Hilbert Spaces are summarized.

### 2024-04-06 thesis/input_files/paper2.tex
April 6th 2024

@@
-\section{\href{http://proceedings.mlr.press/v33/kanagawa14.pdf}{Recovering Distributions from Gaussian RKHS Embeddings}}
+\section{Recovering Distributions from Gaussian RKHS Embeddings}
+% \href{http://proceedings.mlr.press/v33/kanagawa14.pdf}
 This paper covers the RKHS embedding approach to nonparametric statistical inference \cite{pmlr}.

### 2024-04-06 thesis/input_files/point.tex
April 6th 2024

@@
 This chapter covers the theory of the most widely used method for point forecasting in the EF literature.
-Besides discussing the theory underlying these models, this and the following chapter include also a couple of worked examples in order to get acquainted with the practical applications of models.
+Besides discussing the theory underlying such models, this and the following chapter include also a couple of worked examples in order to get acquainted with the practical applications of these models.
 
@@
 \section{Multiple linear regression}
-Notwithstanding, its simplicity, multiple linear regression models are still popular among the EPF literature. Notice, they are no used per se but combined with other more advanced models.
+Notwithstanding its simplicity, multiple linear regression models are still popular among the EPF literature. Notice, they are not used per se but usually combined with other more advanced models.
 \begin{equation}
@@
 On the other hand, load and price have been observed to be a non stationary time series.
-Basically, stationarity means that the distribution of any subsequence of random variables of the stochatic process is invariant to shift along the time dimension.
+Basically, stationarity means that the distribution of any subsequence of random variables of the stochatic process is invariant to shifts along the time dimension.
 \\
-Therefore, it is essential to check stationary through unit root tests.
-Moreover, should the time series be non stationary, we have to convert it to a stationary one by a combination of either differencing, detrending and log transforming.
+Therefore, checking stationarity is a crucial step in applying this class of models. %(unit root tests)
+Furthermore, should the time series be non stationary, we have to convert it to a stationary one by a combination of either differencing, detrending and/or log transforming it.
 \subsection{ARIMA}
@@
 \end{equation}
-where parameters p is the order of the AR part, d is the degree of integrated differencing and q is the order of the MA model.
-$\epsilon_{N-j}$ are the observed errors in the past while the differencing degree $(d)$ is defined recursively as $L_N^{(d)}=L_N^{(d-1)}-L_{N-1}^{(d-1)}$.
-ARIMA model are usually written as
+where the parameter p is the order of the AR part, d is the degree of integrated differencing and q is the order of the MA model.
+$\epsilon_{N-j}$ are the observed errors in the past while the differenced term $L_N^{(d)}$ is defined recursively as $L_N^{(d)}=L_N^{(d-1)}-L_{N-1}^{(d-1)}$.
+ARIMA model are compactly written as
 \begin{equation}
@@
 where $B$ is the backward shift operator, $B^h y_t=y_{t-h}$ and $\phi(B)$ is shorthand for $\phi(B)=1-\phi_1B-\dots-\phi_p B^p$.
-Similarly $theta(B)=1+\theta_1B+\dots+\theta_q B^q$ and $\nabla_h$ is the lag h differencing operator $\nabla_h y_t=(1-B^h)y_t=y_t-y_{t-h}$
+Similarly, $\theta(B)=1+\theta_1B+\dots+\theta_q B^q$ and $\nabla_h$ is the lag h differencing operator $\nabla_h y_t=(1-B^h)y_t=y_t-y_{t-h}$
 \\
-Box-Jenkins is the method for estimating such models \cite{box2015time}.
+The most popular procedure for estimating this class of models is the Box-Jenkins method \cite{box2015time}.
 The order p is identified by taking the lags at which the sample partial autocorrelation function (PACF) falls outside its 95\% confidence interval.
-In the same way, the order q is estimated by consider the sample autocorrelation function of the observed time series.
-Finally, we compare ARIMA models with different p,d,q (in a neighbourhood of the identified orders) and choose the ones minimising either AIC or BIC.
-Notice, models are obtained through the maximum likelihood estimation framework.
+In the same way, the order q is estimated by considering the sample autocorrelation function of the observed time series.
+Finally, we compare ARIMA models with different p,d,q (in a neighbourhood of the identified orders) and choose the ones minimising either AIC or BIC criteria.
+% Notice, models are obtained through the maximum likelihood estimation framework.
 
 \subsection{ARIMAX}
-ARIMAX model add explanatory variables to our ARIMA model, we have
+ARIMAX model adds explanatory variables to our ARIMA, hence, we have
 \begin{equation}
@@
 SARIMA and SARIMAX incorporate seasonalities into our time series models.
-The notation for these models is ARIMA$(p,d,q)(P,D,Q)_m$, the second terms represents the autoregressive structure of the seasonal patter.
+The notation for these models is ARIMA$(p,d,q)(P,D,Q)_m$.
+The second term $(P,D,Q)_m$ represents the autoregressive structure of the seasonal patter, where $m$ indicates the number of seasonalities of our time series.
+For instance, in modelling series of hourly data with daily periodicity, we would set m=24.
+Writing SARIMA in compact notation we have
 \\
@@
 \end{equation}
-\\
-For instance, in modelling series of hourly data with daily periodicity, we set m=24.
 
@@
 %prophet meta
-Particularly successful in time series forecasting is the Prophet model \cite{taylor2018forecasting}.
+Particularly successful in time series forecasting is the Prophet model \cite{taylor2018forecasting}, developed by Meta.
 This additive model is capable of handling non linear trends, holiday effects and yearly weekly and daily seasonality.
+Thus, the reason we choose it as a valid benchmark to compare against.
 \section{K-nearest neighbours regression}
 K-nearest neighbours regression forecasts by averaging the most similar k instances in the training set \cite{macqueen1967some}.
-K is the algorithm hyperparameter, a small k leads to overfitting while a big k leads to underfitting.
-Since it is a metric based algorithm it is important to normalize it in order to give equal weights to features with different scales \ref{appendix:normalization}.
-The algorithm is made up of two decision: first, the choice of the metric and second, the method for combining targets.
+K is the algorithm hyperparameter, it stands for the number of neighbours; a small k leads to overfitting while a big k leads to underfitting.
+Since it is a metric based algorithm, it is important to normalize data in order to give equal weights to features with different scales, see \ref{appendix:normalization}.
+This algorithm is made up of two decisions: first, the choice of the metric and second, the method for combining targets.
 
@@
 Developed at the AT\&T Bell Laboratories by Vapnik et al.
-\cite{cortes1995support}, \cite{vapnik1997support}, support vector machines SVMs are one of the most popular techniques within the field of statistical learning.
+\cite{cortes1995support} \cite{vapnik1997support}, support vector machines SVMs are one of the most popular techniques within the field of statistical learning.
 \\
-The goal of support vector regression is finding a function f(x) with at most $\epsilon$ deviation from the actual observed data $y_i$ $\forall i$ and as flat as possible. 
+The goal of support vector regression is finding a function $f(x)$ with at most $\epsilon$ deviation from the actual observed data $y_i$ for every i and as flat as possible. 
 Put differently, we would like a model to keep error less than the $\epsilon$ threshold,  
@@
 \begin{equation}
-    f(x)=\langle w,x \rangle +b \ \textrm{with} \ w \in X, b \in R
+    f(x)=\langle w,x \rangle +b \ \textrm{with} \ w \in \mathcal{X}, b \in \mathbb{R}
 \end{equation}
-Where $X$ denotes the space of the input patters $x_i$.
-We can translate the flatness requirement into minimising the squared norm of w. Doing so we can formulate our problem as a convex optimisation problem.
+Where $\mathcal{X}$ denotes the space of the input patters $x_i$.
+We can translate the flatness requirement into minimising the squared norm of $w$. Doing so, we can formulate our problem as a convex optimisation problem.
 \begin{equation}
@@
 \end{equation}
-Next, we need to introduce the slack variables $\xi$ and $\xi^*$, in order to handle the above formulate optimisation problem.
-We obtain the following formulation
+Next, we can introduce the slack variables $\xi$ and $\xi^*$ and obtain the following equivalent formulation
 \begin{equation}
     \begin{aligned}
-        \min \quad& \frac{1}{2}\|w\|^2+C\sum\limits_{i=1}^l(\xi_i+\xi_i^*)
+        \min \quad& \frac{1}{2}\|w\|^2+C\sum\limits_{i=1}^n(\xi_i+\xi_i^*)
         \\
@@
 The C constant trades off between $\epsilon$ deviation tolerance and flatness of the function f.
+A bigger C gives more importance to error minimization while as C gets smaller, flattness gains importance.
+\\
 We seek to minimise the epsilon insensitive loss function, defined as follows
@@
         \\
-        \|\xi\|-\epsilon \quad& \textrm{if} \ \|\xi\|\leq \epsilon
+        \|\xi\|-\epsilon \quad& \textrm{if} \ \|\xi\| > \epsilon
     \end{cases}
 \end{equation}
-The model is depicted in figure \ref{fig:svm_simple}. The gry band is called the epsilon insensitive tube, only the points outside it are accounted by the loss function.
+The model is depicted in figure \ref{fig:svm_simple}. The grey band is called the epsilon insensitive tube, only the points outside it are accounted by the loss function.
 \begin{figure}
@@
     \begin{aligned}
-        \max \quad& -\frac{1}{2}\sum\limits_{i,j=1}^l(\alpha_i-\alpha_i^*)(\alpha_j-\alpha_j^*)\langle x_i,x_j\rangle +\sum\limits_{i=1}^l(\alpha_i-\alpha_i^*)(y_i-\epsilon)
+        \max_{\alpha_i, \alpha_i^*} \quad& -\frac{1}{2}\sum\limits_{i,j=1}^n(\alpha_i-\alpha_i^*)(\alpha_j-\alpha_j^*)\langle x_i,x_j\rangle +\sum\limits_{i=1}^n(\alpha_i-\alpha_i^*)(y_i-\epsilon)
         \\
-        s.t. \quad& \sum\limits_i=1^l(\alpha_i-\alpha_i^*)=0
+        s.t. \quad& \sum\limits_{i=1}^n(\alpha_i-\alpha_i^*)=0
         \\
@@
 \end{equation}
-Rearranging the gradient of the lagrangian with respect to w, we obtain the so called support vector expansion.
+Rearranging the gradient of the lagrangian with respect to $w$, we obtain the so called support vector expansion.
 \begin{equation}\label{eq:support vector expansion}
-    w=\sum\limits_{i=1}^l(\alpha_i-\alpha_i^*)x_i
+    w=\sum\limits_{i=1}^n(\alpha_i-\alpha_i^*)x_i
 \end{equation}
-That means that w can be completely described as a linear combination of the training features $x_i$.
-Notice, the complexity of the function representation is independent of the feature space dimensionaly but depends only on the number of support vectors; that is those point $i$ for which $(\alpha_i-\alpha_i^*)\neq 0$
-Notice that, we do not needd to compute w explicitly in order to evaluate f(x).
+This means that $w$ can be completely described as a linear combination of the training features $x_i$.
+Notice, the complexity of the function representation is independent of the feature space dimensionaly, but depends only on the number of support vectors; that is those point $i$ for which $(\alpha_i-\alpha_i^*)\neq 0$
+Moreover, we do not needd to compute $w$ explicitly in order to evaluate $f(x)$.
 Furthermore, equation \ref{eq:support vector expansion} implies
 \begin{equation}
-    f(x)=\sum\limits_{i=1}^l(\alpha_i-\alpha_i^*)\langle x_i, x\rangle +b
+    f(x)=\sum\limits_{i=1}^n(\alpha_i-\alpha_i^*)\langle x_i, x\rangle +b
 \end{equation}
-Employ the Karush-Kuhn-Tucker conditions, b can be retrieved easily.
-Particularly, use the condition that the product between the constraints and the dual variable has to vanish.
+Employing the Karush-Kuhn-Tucker conditions, $b$ can be retrieved easily.
+Particulalry, the condition that the product between the constraints and the dual variable has to vanish.
 These implies that the lagrange multipliers $\alpha_i, \alpha_i^*$ may be nonzero only for the samples inside the epsilon insensitive tube.
 Consequently for any of these data points, the equality $b=y_i-\langle w, x_i-\epsilon\rangle$ holds.
-To get an idea, see figure \ref{fig:svr1} for a how support vector regression handles a sinusoidal function with noise.
+\\
+To get an idea of how SVR works, see figure \ref{fig:svr1} for how a support vector regression handles a sinusoidal function with noise.
 \\
@@
 \end{equation}
-where $\psi$ is an activation function and w,b are the parameters of the neuron; such parameters are typically optimized through gradient descent.
+where $\psi$ is an activation function and $w,b$ are the parameters of the neuron; such parameters are typically optimized through gradient descent.
 
@@
 % https://rtavenar.github.io/deep_book/book_en.pdf
-LSTM extend RNN by introducing a cell state in its layers.
-In this way, the current state of a cell depndes on the current value $X_t$, on the previous activation and on the previous cell state $C_{t-1}$, see figure \ref{fig:lstm_cell} for a visual representation.
+LSTM extends RNN by introducing a cell state in its layers.
+In this way, the current state of a cell depndes on the current value $X_t$, on the previous cell activation and on the previous cell state $C_{t-1}$, see figure \ref{fig:lstm_cell} for a visual representation.
 \begin{figure}[!h]
@@
 \end{figure}
-The cell is responsible for deciding whether to store/forget long and short term values. To do so, each cell is made of a forget gate, an input gate and an output gate.
-This forget gate layer takes $h_{t-1}$ and $x_t$, and outputs a number between 0 and 1, this corresponds to ratio of information retention for each element of the cell state $C_{t-1}$.
+The cell is responsible for deciding whether to store/forget long and short term information. To do so, each cell is made of a forget gate, an input gate and an output gate.
+The forget gate takes in $h_{t-1}$ and $x_t$ and outputs a real valued vector with elements in the domain $[0,1]$, this corresponds to the ratio of information retention for each element of the cell state $C_{t-1}$.
+\\
+Its equation is given by
 \begin{equation}
@@
 \end{equation}
-The input gate layer is made up of two parts.
+The input gate is made up of two parts.
 First, a sigmoid layer decides which values of the cell state will be updated.
@@
     \begin{aligned}
-        i_t=&\sigma(W_i\dot[h_{t-1},x_t]+i_f)
+        i_t=&\sigma(W_i\dot[h_{t-1},x_t]+b_i)
         \\
@@
 \end{equation}
-Finally we output the current cell state and the tanh activated cell state filtered by the sigmoid layer on the current input and the previous hidden state.
+Finally, we output the current cell state and the tanh activated cell state filtered by the sigmoid layer on the current input and the previous hidden state.
 \begin{equation}
     \begin{aligned}
-    o_t=& sigma(W_o[h_{t-1},x_t]+b_o)
+    o_t=& \sigma(W_o[h_{t-1},x_t]+b_o)
     \\
@@
 \begin{theorem}
-    Let $k:X \times X \mapsto R$ on a non empty set $X$ with corresponding RKHS $\mathcal{H}$ and consider the regularized risk functional $Risk[f]:f \mapsto  E\left((x_{1},y_{1},f(x_{1})),\ldots ,(x_{n},y_{n},f(x_{n}))\right)+g\left(\| f\| \right)$.
-Then any minimizer of the empirical risk $Risk[f]$ can be represented as $ f^{*}(\cdot )=\sum _{i=1}^{n}\alpha _{i}k(\cdot ,x_{i})$
+    Let $k:\mathcal{X} \times \mathcal{X} \mapsto R$ on a non empty set $\mathcal{X}$ with corresponding RKHS $\mathcal{H}$ and consider the regularized risk functional $Risk[f]:f \mapsto  E\left((x_{1},y_{1},f(x_{1})),\ldots ,(x_{n},y_{n},f(x_{n}))\right)+g\left(\| f\| \right)$.
+Then any minimizer of the empirical risk $Risk[f]$ can be represented as $ f^{*}(\cdot )=\sum\limits_{i=1}^{n}\alpha _{i}k(\cdot ,x_{i})$
 \end{theorem}
@@
 \begin{proof}
-    By orthogonal projection, any $f \in \mathcal{H}$ can be decomposed as the sum of two functions, with the first lying in the $\textrm{span}\{\phi(x_1),\dots, \phi(x_n)\}$ and the second staying in the orthogonal complement.
+    By orthogonal projection, any $f \in \mathcal{H}$ can be decomposed as the sum of two functions, with the first lying in the $\textrm{span}\{\phi(x_1),\dots, \phi(x_n)\}$ and the second staying in the orthogonal complement of the former.
+Letting $\nu$ denote the function from the orthogonal complement, we have
     \begin{equation}
@@
     \end{equation}
-    This means that setting $\nu=0$ does not affect $f$.
+    This means that setting $\nu=0$ does not affect the evaluation of $f$.
     Next consider the regularization term; notice, the functional $g$ has to be strictly monotonic in order to penalize more complex functions $f$.
@@
         \begin{aligned}
-            g(\|f\|)=&g(\|\sum\limits_{i=1}^n \alpha_i \phi(x_i)+v\|_\mathcal{H})
+            g(\|f\|)=&g\left(\left\|\sum\limits_{i=1}^n \alpha_i \phi(x_i)+v\right\|_\mathcal{H}\right)
             \\
-            =&g\left(\sqrt{\|\sum\limits_{i=1}^n \alpha_i \phi(x_i)+v\|_\mathcal{H}^2}\right)
+            =&g\left(\sqrt{\left\|\sum\limits_{i=1}^n \alpha_i \phi(x_i)+v\right\|_\mathcal{H}^2}\right)
             \\
-            =&g\left(\sqrt{\|\sum\limits_{i=1}^n \alpha_i \phi(x_i)\|_{\mathcal{H}}^2+\|v\|_\mathcal{H}^2+2\langle\sum\limits_{i=1}^n \alpha_i \phi(x_i),\nu \rangle_{\mathcal{H}}}\right)
+            =&g\left(\sqrt{\left\|\sum\limits_{i=1}^n \alpha_i \phi(x_i)\right\|_{\mathcal{H}}^2+\left\|v\right\|_\mathcal{H}^2+2\langle\sum\limits_{i=1}^n \alpha_i \phi(x_i),\nu \rangle_{\mathcal{H}}}\right)
             \\
-            =&g\left(\sqrt{\|\sum\limits_{i=1}^n \alpha_i \phi(x_i)\|_{\mathcal{H}}^2+\|v\|_\mathcal{H}^2}\right)
+            =&g\left(\sqrt{\left\|\sum\limits_{i=1}^n \alpha_i \phi(x_i)\right\|_{\mathcal{H}}^2+\|v\|_\mathcal{H}^2}\right)
             \\
-            \geq &g\left(\sqrt{\|\sum\limits_{i=1}^n \alpha_i \phi(x_i)\|_{\mathcal{H}}^2}\right)
+            \geq &g\left(\sqrt{\left\|\sum\limits_{i=1}^n \alpha_i \phi(x_i)\right\|_{\mathcal{H}}^2}\right)
         \end{aligned}
     \end{equation}
-    We can conclude setting $\nu=0$ decreases the second terms while the first is not affected.
+    Hence, we can conclude that setting $\nu=0$ decreases the second term while the first term is not affected.
 Thus it follows that any minimizer of the risk functional \ref{eq:kernel ridge} must be of the following form
@@
     \begin{aligned}
-    \hat{\alpha}=&\argmin{\alpha \in R^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\lambda}{2}\left\|\sum _{i=1}^{n}\alpha _{i}k(\cdot ,x_{i})\right\|_{\mathcal{H}}^2
+    \hat{\alpha}=&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\lambda}{2}\left\|\sum _{i=1}^{n}\alpha _{i}k(\cdot ,x_{i})\right\|_{\mathcal{H}}^2
     \\
-    \hat{\alpha}=&\argmin{\alpha \in R^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\lambda}{2} \langle \sum _{i}^{n}\alpha _{i}k(\cdot ,x_{i}),\sum _{j=1}^{n}\alpha _{j}k(\cdot ,x_{j})\rangle_{\mathcal{H}}^2
+    \hat{\alpha}=&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\lambda}{2} \langle \sum _{i=1}^{n}\alpha _{i}k(\cdot ,x_{i}),\sum _{j=1}^{n}\alpha _{j}k(\cdot ,x_{j})\rangle_{\mathcal{H}}^2
     \\
-    \hat{\alpha}=&\argmin{\alpha \in R^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\lambda}{2} \sum _{i,j=1}^{n}\alpha _{i}\alpha _{j}\langle k(\cdot ,x_{i}),k(\cdot ,x_{j})\rangle_{\mathcal{H}}^2
+    \hat{\alpha}=&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\lambda}{2} \sum _{i,j=1}^{n}\alpha _{i}\alpha _{j}\langle k(\cdot ,x_{i}),k(\cdot ,x_{j})\rangle_{\mathcal{H}}^2
     \\
-    \hat{\alpha}=&\argmin{\alpha \in R^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\lambda}{2} \sum _{i,j=1}^{n}\alpha _{i}\alpha _{j}K_{ij}
+    \hat{\alpha}=&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\lambda}{2} \sum _{i,j=1}^{n}\alpha _{i}\alpha _{j}K_{ij}
     \\
-    \hat{\alpha}=&\argmin{\alpha \in R^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\lambda}{2} \alpha^\intercal K \alpha
+    \hat{\alpha}=&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\lambda}{2} \alpha^\intercal K \alpha
     \\
-    \hat{\alpha}=&\argmin{\alpha \in R^n}\frac{1}{2}(y-K\alpha)^\intercal (y-K\alpha)+\frac{\lambda}{2} \alpha^\intercal K \alpha
+    \hat{\alpha}=&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}(y-K\alpha)^\intercal (y-K\alpha)+\frac{\lambda}{2} \alpha^\intercal K \alpha
 \end{aligned}
@@
 
-Setting the gradient to zero, we have
+Setting the gradient to zero
 \begin{equation}
@@
 \subsection{Kernel support vector regression}
-Support vector regression can be kernelized by swapping the R2 euclidean dot product of data x with the dot product in the higher feature space F. Doing so, the optimisation problem can be restated as
+Support vector regression can be kernelized by swapping the $\mathbb{R}^2$ euclidean dot product of data x with the dot product in the higher feature space F.
+Doing so, the optimisation problem can be restated as
 \begin{equation}
     \begin{aligned}
-        \max \quad& -\frac{1}{2}\sum\limits_{i,j=1}^l(\alpha_i-\alpha_i^*)(\alpha_i-\alpha_i^*)k(x_i, x_j)+\sum\limits_{i=1}^l(y_i-\epsilon)(\alpha_i-\alpha_i^*)
+        \max_{\alpha_i, \alpha_i^*} \quad& -\frac{1}{2}\sum\limits_{i,j=1}^n(\alpha_i-\alpha_i^*)(\alpha_i-\alpha_i^*)k(x_i, x_j)+\sum\limits_{i=1}^n(y_i-\epsilon)(\alpha_i-\alpha_i^*)
         \\
-        s.t. \quad& \sum\limits_{i=1}^l(\alpha_i-\alpha_i^*)=0
+        s.t. \quad& \sum\limits_{i=1}^n(\alpha_i-\alpha_i^*)=0
         \\
@@
 \end{equation}
-Our regressor f is then given by
+Our regressor $f$ is then given by
 \begin{equation}
     \begin{aligned}
-        \sum\limits_{i=1}^l(\alpha_i-\alpha_i^*)k(x_i, x)+b
+        f(x)=\sum\limits_{i=1}^n(\alpha_i-\alpha_i^*)k(x_i, x)+b
     \end{aligned}
 \end{equation}
-Notice, in this setting w is no longer given explicitly and we seek for the flattest function in the feature space not the input space.
-See figure \ref{fig:svr2} and figure \ref{fig:svr3} for two examples. In the former we have used a polynomial kernel while in the latter the standard radial basis function kernel.
+Notice, in this setting $w$ is no longer given explicitly. Additionally, with kernel SVR we seek for the flattest function in the feature space not the input space.
+See figure \ref{fig:svr2} and figure \ref{fig:svr3} for two examples. In the former, we have used a polynomial kernel while in the latter the standard radial basis function kernel.
 \begin{figure}
@@
 \end{figure}
-Comparing theese pictures with figure \ref{fig:svr1}, it can be concluded that introducing kernels allows support vector regression to handle non linearities in the data.
+Comparing theese pictures with figure \ref{fig:svr1}, it can be concluded that introducing kernels allows support vector regression to handle the non linearities in the data.

### 2024-04-07 thesis/input_files/analysis.tex
April 7 2024

@@
 
+\section{Point forecasting}
+This section carries out a comparative study between the state of the art methods for point forecasting, introduced in \ref{ch:point}.
+As use case, we will consider the task of predicting the load for the next month.
+In such setting we have the following regressors
+\begin{itemize}
+    \item Day
+    \item Hour
+    \item Month
+    \item Day of the week
+    \item Is holiday
+    \item Weather temperature
+\end{itemize}
+The metric upon we agreed is the root mean squared error, see section \ref{metrics}.
+\subsection{Multiple linear regression}
+To get started, standard multiple linear regression has been applied, see fig \ref{fig:mlr_price} for a visualisation. 
+\begin{figure}
+    \includegraphics[width=\textwidth]{images/mlr_price.png}
+    \caption{Multiple linear regression prediction}
+    \label{fig:mlr_price}
+\end{figure}
+
+The resulting RMSE is 30.59.
+What can be concluded, is that multiple linear regression is capable of catching the daily seasonality. Nevertheless, it cannot catch the range of the price series properly.
+
+\subsection{Tbats}
+Next, we tried with the autoregressive approach.
+Unfortunately, AR, ARIMA and SARIMA models did not perform as expected, their output was slightly better than the one of linear regression.
+This is probably due to the fact that, the data considered entails two kinds of seasonalities, while ARIMA models can only handle one at a time.
+We remind the reader, that the price time series involves both daily and weekly seasonalities.
+Hence, the need for a more advanced time series model.
+The Tbats model is a forecasting method capable of handling complex patterns in the data.
+Its name stands for trigonometric seasonality, Box-Cox transformation, ARMA errors, trend and seasonal components.
+Tbats forecast is visualized in figure \ref{fig:tbats_price}, meanwhile it RMSE is 15.08.
+\begin{figure}[!h]
+    \includegraphics[width=\textwidth]{images/tbats_price.png}
+    \caption{Tbats prediction}
+    \label{fig:tbats_price}
+\end{figure}
+From the plot we see that Tbats is capable of catching the lows and average trend, conversely it has some difficulties handling the price peaks.
+\subsection{Prophet}
+Following, the prophet model has been considered. 
+We get started by considering the base implementation. In such setting, prophet takes in as input the time series object and learns its data generating process.
+This method achieves a RMSE of 23.96, its prediction is visualized in figure \ref{fig:prophet_price_1}.
+\begin{figure}[!h]
+    \includegraphics[width=\textwidth]{images/prophet_price_1.png}
+    \caption{Prophet prediction}
+    \label{fig:prophet_price_1}
+\end{figure}
+What can be seen is that, prophet models correctly the average trend but does not model peaks and lows precisely. 
+Next, a more complex model was trained.
+We added the weather temperature, the square of it and the categorical variable for holidays effect as regressors.
+Furthermore, we also applied a log transformation to the dependet variable.
+Doing so, RMSE went down to 10.29.
+Forecast is visualized in \ref{fig:prophet_price2}, moreover, figure \ref{fig:prophet_price2.1} and figure \ref{fig:prophet_price2.2} break down the trend and the seasonalities respectively.
+\begin{figure}[!h]
+    \includegraphics[width=\textwidth]{images/prophet_price2.png}
+    \caption{Prophet v2 prediction}
+    \label{fig:prophet_price2}
+\end{figure}
+
+\begin{figure}[!h]
+    \includegraphics[width=\textwidth]{images/prophet_price2.1.png}
+    \caption{Seasonalities breakdown}
+    \label{fig:prophet_price2.1}
+\end{figure}
+
+\begin{figure}[!h]
+    \includegraphics[width=\textwidth]{images/prophet_price2.2.png}
+    \caption{Prices trend}
+    \label{fig:prophet_price2.2}
+\end{figure}
+Prophet alongside with the right features is a good model in the context of electricity price forecasting, it is capcable of catching the trend, the peak and the lows.
+
+\subsection{K-nearest neighbours}
+Afterwards, K-nearest neighbours has been considered.
+In doing so, we cross validated the number of neighbours to get the best model instance.
+The forecast is visualized in figure \ref{fig:knn_price}, RMSE is 12.54.
+\begin{figure}[!h]
+    \includegraphics[width=\textwidth]{images/knn_price.png}
+    \caption{K-nearest neighbour regression}
+    \label{fig:knn_price}
+\end{figure}
+What it can be said is that, K-nearest neighbour is capable of predicting prices well by averaging past data.
+
+\subsection{Support vector regression}
+Coming up we have support vector regression.
+In applying this model we used gridsearch crossvalidation to search for the best regularization parameter C.
+Support vector regression achieves an RMSE of 23.24, for the prediction see figure \ref{fig:svr_price}.
+\begin{figure}[!h]
+    \includegraphics[width=\textwidth]{images/svr_price.png}
+    \caption{Support vector regression prediction}
+    \label{fig:svr_price}
+\end{figure}
+Visually, we see that the SVR performs similar to multiple linear regression as expected.
+Similarly to multiple linear regression, SVR can model the daily seasonality but the point prediction is off in terms of highs and lows.
+
+\subsection{LSTM}
+
+\subsection{Kernel ridge regression}
+Subsequently, kernel ridge regression was considered. The kernel considered is the radial basis gaussian function.
+Cross validation was carried jointly over the rbf kernel bandwith and the regularization constant.
+The RMSE achieved is 9.86, figure \ref{fig:krnridge_price} reports the prediction.
+\begin{figure}[!h]
+    \includegraphics[width=\textwidth]{images/krnridge_price.png}
+    \caption{Kernel ridge prediction}
+    \label{fig:krnridge_price}
+\end{figure}
+We can observe that, kernel ridge regression models accurately the electricity time series.
+
+\subsection{Kernel support vector regression}
+The last model we considered in this setting was the kernel support vector regression.
+RMSE achieved with this method is 9.10, forecast is reported in figure \ref{fig:krnsvr_price}.
+
+\begin{figure}[!h]
+    \includegraphics[width=\textwidth]{images/krnsvr_price.png}
+    \caption{Kernel support vector prediction}
+    \label{fig:krnsvr_price}
+\end{figure}
+We can observe that kernel support vector prediction is one of the best performing techniques between the one considered.
+
+\subsection{Results}
+This section reports the table comparing the considered methods in the context of electricity load forecasting.
+
+- table 
+|methods|tasks RMSE|
+|       |X          |       
+|       |X          |
+
 \section{Quantile forecasting}

### 2024-04-07 thesis/input_files/data_exploratory.tex
April 7 2024

--- 2024-02-24 thesis/input_files/data_exploratory.tex
@@
 not generalize well.
+
+This section covers the datasets used in our experiments, attributes and features will be thoroughly described.
+Next, we will explain the ETL pipeline that we set up in order to ease the workflow of our comparison studies.
+Finally, in order to to bettern understand patters within the data, we carry out an exploratory data analysis.
+
+\section{Dataset}
+The chosen datasets for probabilistic load forecasting, were the data from the GEFCom 2014 competition. The data is freely hosted on Dr. Hong blog \cite{hong2016probabilistic}.
+The main reason was that, these dataset are considered an a excellent test case for comparing predictive models between the EF community.
+Additionally, the scores of the competing models are freely available, this enables us to carry out a clean and transparent comparative study.
+The GEFCom 2014 consisted of four tracks: price, load, wind power and solar power forecasting. 
+The GEFCom2014 folder contains a zip file for each of the four tracks.
+This thesis work is focuesed on providing forecasts for the load and price quantities.
+\\
+\subsection{Price track}
+In the price forecasting track, the goal was to predict the electricity price for the next 24 hours of a single zone. 
+The data can be found in the GEFCom2014-P\_V2 zip file alongside a set of instructions and the benchmark forecasts.
+This data consisted of time series for the locational marginal price, for the zonal load and for the system load.
+The data covers the time interval ranging from the 1st of January 2011 to the 15th of June 2013, see figure \ref{fig:price_track_fig1}.
+Additionally, as the competition went on, the real observed data of the previous tasks were made available.
+The fifteen target days ordered by task number are: 16/06/2013, 17/06/2013, 24/06/2013, 04/07/2013, 09/07/2013, 13/07/2013, 16/07/2013, 18/07/2013, 19/07/2013, 20/07/2013, 24/07/2013, 25/07/2013, 07/12/2013, 08/12/2013.
+\begin{figure}[!h]
+    \includegraphics[width=\textwidth]{images/price_track_fig1.png}
+    \caption{Price track}
+    \label{fig:price_track_fig1}
+\end{figure}
+\subsection{Load track}
+In the load track, contestant were asked to provide one month ahead hourly probabilistic forecasts on a rolling basis for 15 consecutive rounds.
+To get started, in the first round, organizers provided 69 months of hourly load data (from 01/01/2005 to 30/10/2010) and 117 months of hourly temperature data (from 01/01/2001 to 30/10/2010).
+As for the price track, the true observed data from the previous track were made available as the competition progressed.
+\\
+The data for load forecasting is contained in the GEFCom2014-L\_V2 zip file.
+Within this subfolder, we a txt with the competition instructions and 15 folders one for each of the consecutive tasks.
+For each of those we have the prediction of the competition benchmark model and the train file upon which we will fit our model.
+Each data from the various task folders are shifted by one months between others.
+\\
+In order to compare our model performance with the winning entries of the GEFCom2014 load competition, we will refer to the Provisional\_Leaderboard\_V2 file contained also in the GEFCom2014 Data directory.
+% Notice a quick premise, within this xsls file the authors opted for various scores to compare the entries performance. This choice was motivated by the need 
+% to give a proportial weight to each task when averaging between them; that is in order to not penalise too much the tasks with higher point spreads.
+Furthermore, additional factors such as frequency of benchmark outperformance, and robustness and clarity in methodology explaination.
+\\
+Notice, the pinball scores for the load track are stored inside the subtab L-score-0/L-score-2 of Provisional\_Leaderboard\_V2.
+
+\section{ETL}
+
+\section{EDA}

### 2024-04-07 thesis/input_files/implementation.tex
April 7 2024

@@
 explain better my ideas.
-\section{Linear quantile regression}
+
+- Indicate also hyperparameters maybe in each subsection
+
+\section{Point forecasting}
+\subsection{Multiple linear regression}
+For multiple linear regression the one from the \href{https://scikit-learn.org/stable/}{sklearn} library has been used.
+
+\subsection{Tbats}
+Tbats implementation is available at \url{https://github.com/intive-DataScience/tbatsx}.
+In our application we specified as hyperparameters the lenght of seasons; 24 for the daily seasonality and 168 for the weekly seasonality.
+
+\subsection{Prophet}
+The prophet models has been applied by employing the python API provided by Meta \href{https://facebook.github.io/prophet/docs/quick_start.html}{prophet}.
+
+\subsection{K-nearest neighbours}
+The object KNeighborsRegressor of the sklearn module neighbors has been used.
+
+\subsection{Support vector regression}
+The object SVR of the sklearn module svm has been used by specifying the liner kernel.
+
+\subsection{LSTM}
+
+\subsection{Kernel ridge regression}
+The object KernelRidge of the sklearn module kernel\_ridge has been used.
+
+\subsection{Kernel support vector regression}
+The object SVR of the sklearn module svm has been used by specifying rbf as the kernel parameter.
+
+\section{Probabilistic forecasting}
+\subsection{Linear quantile regression}
 The implementation of statsmodels.regression.quantile\_regression.QuantReg has been used.
 The model is fitted through iterative reweighted least squares.
-\section{Quantile gradient boosting machine}
+\subsection{Quantile gradient boosting machine}
 The implementation of sklearn.ensemble.GradientBoostingRegressor has been used.
-\section{Quantile forest}
+\subsection{Quantile forest}
 The implementation of quantile\_forest.RandomForestQuantileRegressor has been used. This estimator is compatible with scikit-learn.
-\section{Kernel quantile regression}
+\subsection{Kernel quantile regression}
 Kernel quantile regression had no previously implemented open source library, thus the need of implementing our own version.
@@
 % (i guess it is in C or C++ and then binded to R).
-\subsection{Python vs R implementation}
+\subsubsection{Python vs R implementation}
 In this section, a comparison study has been carried out in order to inspect the competitiveness of our implementation with the existing one for the R programming language.

### 2024-04-28 thesis/input_files/analysis.tex
April 28 2024

@@
 
-- Comments
-\\
-- Comparison
-\\
-- Table of loss scores
-\\
-Plots:
-\\
-- Plots for visualizing timeseries with quantiles bounds
-\\
-- Other plots that will come up to mind
+% - Comments
+% \\
+% - Comparison
+% \\
+% - Table of loss scores
+% \\
+% Plots:
+% \\
+% - Plots for visualizing timeseries with quantiles bounds
+% \\
+% - Other plots that will come up to mind
 
@@
 This section carries out a comparative study between the state of the art methods for point forecasting, introduced in \ref{ch:point}.
-As use case, we will consider the task of predicting the load for the next month.
-In such setting we have the following regressors
+As use case, we will consider the task of predicting the electric load from the GEFCom 2014 dataset.
+In such setting we considered the following regressors
 \begin{itemize}
@@
 \end{itemize}
-The metric upon we agreed is the root mean squared error, see section \ref{metrics}.
+Methods will be compared by means of the RMSE, MAE and MAPE scores, see section \ref{metrics}.
 \subsection{Multiple linear regression}
@@
 This is probably due to the fact that, the data considered entails two kinds of seasonalities, while ARIMA models can only handle one at a time.
-We remind the reader, that the price time series involves both daily and weekly seasonalities.
+We remind the reader, that the electricity load time series involves both daily and weekly seasonalities.
 Hence, the need for a more advanced time series model.
-The Tbats model is a forecasting method capable of handling complex patterns in the data.
+The Tbats \cite{de2011forecasting} model is a forecasting method capable of handling complex patterns in the data.
 Its name stands for trigonometric seasonality, Box-Cox transformation, ARMA errors, trend and seasonal components.
-Tbats forecast is visualized in figure \ref{fig:tbats_price}, meanwhile it RMSE is 15.08.
+Tbats forecast is visualized in figure \ref{fig:tbats_price}, meanwhile its RMSE is 15.08.
 \begin{figure}[!h]
@@
 \end{figure}
-Visually, we see that the SVR performs similar to multiple linear regression as expected.
-Similarly to multiple linear regression, SVR can model the daily seasonality but the point prediction is off in terms of highs and lows.
+Visually, we see that the SVR performs similar to multiple linear regression, as expected.
+Like multiple linear regression, SVR can model the daily seasonality but the point prediction is off in terms of highs and lows.
 
 \subsection{LSTM}
+We used the torch library in order to code our LSTM regressor.
+Various hyperparameters combinations were tried during hyperparameter tuning. The final hyperparameters we set on follows.
+\begin{itemize}
+    \item hidden\_size= 64
+    \item num\_layers= 2
+    \item output\_size= 1
+    \item num\_epochs= 30
+    \item learning\_rate= 0.001
+    \item batch\_size= 32
+    \item window\_size= 24
+\end{itemize}
+LSTM forecast is reported in table \ref{fig:lstm_price}, it achieves a RMSE of 9.1782
+\begin{figure}[!h]
+    \includegraphics[width=\textwidth]{images/lstm_price.png}
+    \caption{LSTM prediction}
+    \label{fig:lstm_price}
+\end{figure}
 
@@
 \end{figure}
-We can observe that kernel support vector prediction is one of the best performing techniques between the one considered.
+We can observe that kernel support vector regression is one of the best performing techniques between the one considered.
 
 \subsection{Results}
-This section reports the table comparing the considered methods in the context of electricity load forecasting.
+This section reports the table comparing the considered methods scores 
+% in the context of electricity load forecasting.
 
-- table 
-|methods|tasks RMSE|
-|       |X          |       
-|       |X          |
+% - table 
+% |methods|tasks RMSE|
+% |       |X          |       
+% |       |X          |    
+Table \ref{tab:point_RMSE} reporst the RMSE scores for each of the considered methods.
+It can be seen that kernel based methods are the ones achieving the lowest RMSE; specifically, we have KrnSVR, KrnRidge and KNN being among the top methods on the majority of tasks.
 
-\section{Quantile forecasting}
-In this subsection we compare performance of quantile regressor considering the GEFCom2014 dataset.
+\begin{table}[!ht]
+    \caption{RMSE scores}
+    \label{tab:point_RMSE}
+    \begin{adjustbox}{width=\textwidth, height=3cm}
+        \begin{tabular}{lrrrrrrrrrrrrrrr}
+            \toprule
+            Method/RMSE & Task 1 & Task 2 & Task 3 & Task 4 & Task 5 & Task 6 & Task 7 & Task 8 & Task 9 & Task 10 & Task 11 & Task 12 & Task 13 & Task 14 & Task 15 \\
+            \midrule
+            MLR & 30.5892 & 29.0347 & 75.1682 & 63.1376 & 42.1156 & 36.6233 & 39.2496 & 38.8286 & 47.1307 & 68.0494 & 61.6005 & 30.3751 & 34.5523 & 33.4287 & 33.5581 \\
+            TBATS & 15.0886 & 31.9720 & 99.1260 & 84.4890 & 51.9740 & 34.9055 & 18.4445 & 38.3246 & 74.1117 & 98.6600 & 84.3050 & 38.8433 & 16.6080 & 28.9902 & 41.6194 \\
+            Prophet & 10.2936 & 14.4358 & 38.7551 & 63.4787 & 19.7474 & 17.5065 & 12.6926 & 14.2665 & 17.5466 & 23.5944 & 43.6666 & 20.8637 & 16.9493 & 19.1626 & 23.3889 \\
+            KNN & 12.5429 & 11.6699 & 24.5057 & 18.3310 & 13.1821 & \textbf{11.9238} & 12.0044 & 14.5165 & 16.3132 & 15.1831 & 37.6457 & 16.4690 & 12.0324 & 11.3102 & 14.1717 \\
+            SVR & 23.2421 & 25.8525 & 78.8043 & 67.6209 & 42.3748 & 32.7845 & 30.5971 & 35.3660 & 55.4213 & 77.9660 & 68.5799 & 30.2451 & 27.0345 & 28.3652 & 32.1480 \\
+            LSTM & 9.1782 & 12.0669 & \textbf{22.7048} & \textbf{16.2087} & 16.0964 & 12.7936 & \textbf{10.8559} & 14.6173 & 19.7303 & 18.0200 & 43.2051 & 17.1856 & 10.3106 & 12.1347 & 17.5849 \\
+            KrnRidge & 9.8619 & 11.6101 & 37.7824 & 35.2459 & \textbf{12.5160} & 14.5911 & 12.8791 & 17.4385 & 16.1131 & 17.1938 & 37.6961 & 14.0076 & 9.8441 & 10.7491 & 13.2975 \\
+            KrnSVR & \textbf{9.1028} & \textbf{11.3117} & 22.9281 & 19.2132 & 12.6331 & 12.6018 & 11.3537 & \textbf{12.9506} & \textbf{14.9731} & \textbf{11.7765} & \textbf{37.3797} & \textbf{13.2694} &
+            \textbf{8.8522} & \textbf{10.6185} & \textbf{13.2602} \\
+            \bottomrule
+            \end{tabular}            
+    \end{adjustbox}
+\end{table}
+
+The MAE scores are contained in table \ref{tab:point_MAE}. Similarly to above, we can conclude that kernel methods stand out for achieving also the lowest MAE score.
+\begin{table}[!ht]
+    \caption{MAE scores}
+    \label{tab:point_MAE}
+    \begin{adjustbox}{width=\textwidth, height=3cm}
+        \begin{tabular}{lrrrrrrrrrrrrrrr}
+            \toprule
+             Method/MAE & Task 1 & Task 2 & Task 3 & Task 4 & Task 5 & Task 6 & Task 7 & Task 8 & Task 9 & Task 10 & Task 11 & Task 12 & Task 13 & Task 14 & Task 15 \\
+            \midrule
+            MLR & 27.7219 & 24.3823 & 65.4016 & 52.3873 & 34.8465 & 31.2663 & 35.9420 & 34.7316 & 37.0273 & 54.7075 & 48.9327 & 26.1975 & 31.8720 & 29.2306 & 28.3428 \\
+            TBATS & 10.5408 & 23.8417 & 91.3153 & 74.5734 & 40.3258 & 26.0092 & 14.2141 & 24.7818 & 62.0452 & 86.7125 & 72.6912 & 28.4235 & 11.1184 & 21.1038 & 31.8555 \\
+            Prophet & 8.6139 & 11.4307 & 28.3509 & 46.4480 & 15.6290 & 14.1278 & 10.1574 & 11.2360 & 14.0072 & 18.9565 & 27.1142 & 16.9678 & 14.6615 & 16.1750 & 18.3043 \\
+            KNN & 9.4003 & 9.1126 & 19.6213 & 14.4906 & 10.6461 & \textbf{9.4354} & 8.7441 & 10.0065 & 12.4714 & 11.2006 & 20.3087 & 12.5416 & 9.2558 & 8.5937 & 10.5451 \\
+            SVR & 20.2443 & 21.2526 & 69.3704 & 57.1107 & 34.3651 & 27.5875 & 27.4651 & 29.4204 & 43.7009 & 64.3522 & 55.4307 & 25.1080 & 24.0294 & 24.4049 & 26.5869 \\
+            LSTM & 7.5217 & 9.4767 & \textbf{17.9851} & \textbf{13.2158} & 11.8118 & 9.8607 & \textbf{8.0561} & 10.8480 & 15.1008 & 13.8108 & 25.9906 & 13.3588 & 8.1526 & 9.8000 & 14.3186 \\
+            KrnRidge & 7.5965 & 9.2130 & 28.2484 & 24.2366 & \textbf{10.0589} & 10.7758 & 9.5788 & 11.2166 & 12.3427 & 12.3306 & 20.0213 & 10.8434 & 7.1625 & \textbf{7.9710} & 10.0767 \\
+            KrnSVR & \textbf{6.9581} & \textbf{8.6989} & 18.7739 & 15.1398 & 10.1395 & 9.5927 & 8.5731 & \textbf{9.0033} & \textbf{11.0618} & \textbf{8.5486} & \textbf{19.7383} & \textbf{10.5545} & \textbf{7.0395} & 8.1926 & \textbf{9.8125} \\
+            \bottomrule
+            \end{tabular}            
+    \end{adjustbox}            
+\end{table}
+
+Finally table \ref{tab:point_MAPE} reports the MAPE score for each method. What can be concluded is that kernel regressors are also the best in terms of MAPE score; with KrnSVR being the top one.
+\begin{table}[!ht]
+    \caption{MAPE scores}
+    \label{tab:point_MAPE}
+    \begin{adjustbox}{width=\textwidth, height=3cm}
+        \begin{tabular}{lrrrrrrrrrrrrrrr}
+            \toprule
+             Method/MAPE & Task 1 & Task 2 & Task 3 & Task 4 & Task 5 & Task 6 & Task 7 & Task 8 & Task 9 & Task 10 & Task 11 & Task 12 & Task 13 & Task 14 & Task 15 \\
+            \midrule
+            MLR & 0.2470 & 0.1935 & 0.3030 & 0.2601 & 0.2474 & 0.2667 & 0.3498 & 0.3058 & 0.1954 & 0.2432 & 0.4234 & 0.2114 & 0.2926 & 0.2502 & 0.2145 \\
+            TBATS & 0.0833 & 0.1650 & 0.4285 & 0.3731 & 0.2482 & 0.1849 & 0.1234 & 0.1634 & 0.3181 & 0.4021 & 0.4880 & 0.1761 & 0.0905 & 0.1555 & 0.2058 \\
+            Prophet & 0.0737 & 0.0854 & 0.1324 & 0.2453 & 0.1134 & 0.1169 & 0.0938 & 0.0912 & 0.0880 & 0.0992 & 0.3757 & 0.1389 & 0.1343 & 0.1356 & 0.1370 \\
+            KNN & 0.0778 & 0.0682 & 0.0924 & 0.0761 & 0.0751 & 0.0740 & 0.0784 & 0.0744 & 0.0725 & 0.0555 & \textbf{0.3115} & 0.0888 & 0.0816 & 0.0699 & 0.0765 \\
+            SVR & 0.1808 & 0.1640 & 0.3211 & 0.2829 & 0.2317 & 0.2260 & 0.2676 & 0.2451 & 0.2215 & 0.2867 & 0.4354 & 0.1870 & 0.2216 & 0.2044 & 0.1933 \\
+            LSTM & 0.0642 & 0.0706 & 0.0896 & \textbf{0.0722} & 0.0798 & 0.0772 & \textbf{0.0710} & 0.0854 & 0.0909 & 0.0741 & 0.3578 & 0.0976 & 0.0719 & 0.0778 & 0.1037 \\
+            KrnRidge & 0.0625 & 0.0685 & 0.1281 & 0.1207 & 0.0718 & 0.0797 & 0.0812 & 0.0760 & 0.0696 & 0.0587 & 0.3127 & 0.0779 & \textbf{0.0591} & \textbf{0.0637} & 0.0722 \\
+            KrnSVR & \textbf{0.0568} & \textbf{0.0642} & \textbf{0.0892} & 0.0788 & \textbf{0.0714} & \textbf{0.0724} & 0.0740 & \textbf{0.0665} & \textbf{0.0629} & \textbf{0.0428} & 0.3122 & \textbf{0.0781} & 0.0607 & 0.0646 & \textbf{0.0713} \\
+            \bottomrule
+            \end{tabular}            
+    \end{adjustbox}
+\end{table}
+
+\section{Probabilistic forecasting}
+In this subsection we compare performance of quantile regressors.
 \\

### 2024-04-29 thesis/input_files/analysis.tex
correct names in contents,list of fig and tables

@@
 
-\subsection{Tbats}
+\subsection{Trigonometric seasonality Box-Cox transformation ARMA errors trend and seasonal components}
 Next, we tried with the autoregressive approach.
@@
 
-\subsection{LSTM}
+\subsection{Long short term memory}
 We used the torch library in order to code our LSTM regressor.
@@
     \includegraphics[width=\textwidth]{images/lstm_price.png}
-    \caption{LSTM prediction}
+    \caption{Long short term memory prediction}
     \label{fig:lstm_price}
@@
 \begin{table}[!ht]
-    \caption{RMSE scores}
+    \caption{Root mean squared errors}
     \label{tab:point_RMSE}
@@
 \begin{table}[!ht]
-    \caption{MAE scores}
+    \caption{Mean absolute errors}
     \label{tab:point_MAE}
@@
 \begin{table}[!ht]
-    \caption{MAPE scores}
+    \caption{Mean absolute percentage errors}
     \label{tab:point_MAPE}

### 2024-04-29 thesis/input_files/energymarket.tex
correct names in contents,list of fig and tables

@@
     \includegraphics[width=\textwidth]{images/pool_vs_echange.png}
-    \caption{Pool market vs power echange \cite{weron2006modeling}}
+    \caption{Pool market versus power echange \cite{weron2006modeling}}
     \label{fig:pool_vs_echange}
@@
 
-\subsection{Nodal and Zonal Pricing}
+\subsection{Nodal and zonal pricing}
 Grid configuration and physical limits on electricity lines may lead to congestion.

### 2024-04-29 thesis/input_files/implementation.tex
April 28 2024
correct names in contents,list of fig and tables

@@
 - indicate computer specifics
-
-% - see whether something can be parallelized
 
@@
 \subsection{LSTM}
+The LSTM predictor has been built using the \href{https://pytorch.org}{torch} library.
 
@@
 % (i guess it is in C or C++ and then binded to R).
-\subsubsection{Python vs R implementation}
+\subsubsection{Python versus R implementation}
 In this section, a comparison study has been carried out in order to inspect the competitiveness of our implementation with the existing one for the R programming language.

### 2024-04-29 thesis/input_files/metrics.tex
correct names in contents,list of fig and tables

@@
 Additionally, note that the latter can take different forms which therefore requires different measures.
-\section{MAE}\label{mae}
+\section{Mean absolute error}\label{mae}
 Consider the time series with actual values given by $\mathrm{L}=(L_{n+1}, L_{n+2},\dots, L_{n+h})$
-and its h step ahead point forecast $\mathrm{\hat{L}}=(\hat{L}_{n+1}, \hat{L}_{n+2},\dots, \hat{L}_{n+h})$ the mean absolute error is defined as
+and its h step ahead point forecast $\mathrm{\hat{L}}=(\hat{L}_{n+1}, \hat{L}_{n+2},\dots, \hat{L}_{n+h})$ the mean absolute error (MAE) is defined as
 \begin{definition}
@@
 
-\section{RMSE}\label{rmse}
+\section{Root mean squared error}\label{rmse}
 \begin{definition}
@@
 
-MAE and RMSE posses the useful property of being expressed in the same units of the data thus enabling meaningful comparisons.
+MAE and root mean squared error (RMSE) posses the useful property of being expressed in the same units of the data thus enabling meaningful comparisons.
 However, a drawback of such measures is that we cannot use them to compare accuracy between time series wich have different magnitudes.
 For instance, a day ahead error of 1kWh is negligible when considering a daily demand of 100kWn while the same error is considerably big when daily demand is 2kWh.
-This consideration leads to relative accuracy scores, between those the MAPE is by far the most popular.
+This consideration leads to relative accuracy scores, between those the mean absolute percentage error (MAPE) is by far the most popular.
 
-\section{MAPE}\label{mape}. 
+\section{Mean absolute percentage error}\label{mape}. 
 \begin{definition}
@@
 
-\section{RMSPE}\label{rmspe}
+\section{Root mean squared percentage error}\label{rmspe}
 \begin{definition}
@@
 \end{definition}
-Note, MAPE and RMSPE may not be appropriate for series which have zero or very small values, for example, electricity demand at the household level; the result is a large score regardless of the absolute errors.
+Note, MAPE and root mean squared percentage error (RMSPE) may not be appropriate for series which have zero or very small values, for example, electricity demand at the household level; the result is a large score regardless of the absolute errors.
 Scaled errors constitute a robust family of scores.
-\section{MASE}\label{mase}
+\section{Mean absolute scaled error}\label{mase}
 \begin{definition}
@@
 In this model, the current demand makes up the prediction for the next time step; that is $\mathrm{\hat{L}^{\mathrm{naive}}}_{n+1}=\mathrm{L}_{n}$.
-\section{RMSSE}\label{rmsse}
+\section{Root mean squared scaled error}\label{rmsse}
 \begin{definition}
@@
     \includegraphics[width=\textwidth]{images/pinball_loos.png}
-    \caption{Pinball loss with q=0.2 \cite{haben2023core}}
+    \caption{Pinball loss with $q=0.2$ \cite{haben2023core}}
     \label{fig:pinball}
@@
 \end{definition}
-Where delta is the PI width, that is $\delta=Upp_t-Low_t$. This score penalizes observations falling outside the PI and rewards narrow PI.
-\section{CRPS}
-The continous ranked probability score measures the differece between the estimated cumulative distribution $\hat{F}$ and the empirical Cdf.
+Where delta is the prediction interval (PI) width, that is $\delta=Upp_t-Low_t$. This score penalizes observations falling outside the PI and rewards narrow PI.
+\section{Continous ranked probability score}
+The continous ranked probability score (CRPS) measures the differece between the estimated cumulative distribution $\hat{F}$ and the empirical Cdf.
 \begin{definition}\label{def_crps}
@@
 \\
-\section{PIT}
-The probability integral transform is a method to assess visually the quality of probabilistic forecasts.
+\section{Probability integral transform}
+The probability integral transform (PIT) is a method to assess visually the quality of probabilistic forecasts.
 PIT is obtained by applying a cdf $F$ to your data; if applying such cdf to the data results in a uniform distributed PIT then $F$ is a valid prediction.

### 2024-04-29 thesis/input_files/point.tex
April 28 2024
correct names in contents,list of fig and tables

@@
 Furthermore, should the time series be non stationary, we have to convert it to a stationary one by a combination of either differencing, detrending and/or log transforming it.
-\subsection{ARIMA}
+\subsection{Autoregressive integrated moving average}
 ARIMA(p,d,q) models are of the form
@@
 
-\subsection{ARIMAX}
+\subsection{Autoregressive integrated moving average with exogenous variables}
 ARIMAX model adds explanatory variables to our ARIMA, hence, we have
@@
 \end{equation}
-\subsection{SARIMA and SARIMAX}
+\subsection{Seasonal autoregressive integrated moving average and Seasonal autoregressive integrated moving average with exogenous variables}
 SARIMA and SARIMAX incorporate seasonalities into our time series models.
@@
 But if we want to model say a y bernoulli which is the rate of success we can link the domain of bernoulli to that of gaussian with a link function
-Generalized additive models take the form
+Generalized additive models (GAMs) take the form
 \begin{equation}
@@
 
-\subsection{MLP}
+\subsection{Multilayer perceptrons}
 In order to cover a richer space of models we can stack neurons, doing so we obtain the so called multi layer perceptrons (MLPs).
@@
 
-\subsubsection{NARX}
-Neural networks that have as inputs past values of the target variable and other independent variables are called nonlinear autoregressive exogenous models.
+% \subsubsection{NARX}
+% Neural networks that have as inputs past values of the target variable and other independent variables are called nonlinear autoregressive exogenous models.
 
-\subsection{LSTM}
+\subsection{Long short term memory}
 % https://colah.github.io/posts/2015-08-Understanding-LSTMs/
@@
     \includegraphics[width=\textwidth]{images/lstm_cell.png}
-    \caption{LSTM cell}
+    \caption{Long short term memory cell}
     \label{fig:lstm_cell}

### 2024-04-29 thesis/input_files/problemdescription.tex
correct names in contents,list of fig and tables

@@
 \\
-\section{Point vs probabilistic forecast}
+\section{Point versus probabilistic forecast}
 A distinction has to be made between two types of forecasting approaches: point forecasts and probabilistic forecasts.

### 2024-04-29 thesis/input_files/problemdescription.tex
pag 1 comments

@@
 Individuals and organizations constantly face situations of uncertainty thus the need of robust forecasting methods.
-Such methods are crucial to the process of taking informed decision and to strategic planning.
+Such methods are crucial in the process of taking informed decision and for strategic planning.
 \\
@@
 \\
-The reason of our decision to focus on the energy market is mainly motivated by the rapid changes it has undergone. %/experienced 
-Over the last decades, electricity markets have gone through an unprecedented transformation; this shift was driven by the liberalization of such markets, the development and integration of renewable energy sources, the  increase of low carbon technologies and the adoption of smart meters.
-Events like the California electricity crisis help also motivating the choice of the electricity sector as subject of our studies, see \cite{california}.
-Furthermore, the process of deregulation lead to an increasing interest in the field of electricity price forecasting (EF) within the academic community \ref{fig:epf_evolution}.
-\\
+Our decision to focus on the energy market is mainly motivated by the rapid changes it has undergone. %/experienced 
+Over the last decades, electricity markets have gone through an unprecedented transformation.
+This shift was driven by the liberalization of such markets, the development and integration of renewable energy sources, the  increase of low carbon technologies and the adoption of smart meters.
+Events like the California electricity crisis are further motivating the choice of the electricity sector as subject of our studies, see \cite{california}.
+Moreover, the process of deregulation lead to an increasing interest in the field of electricity forecasting (EF) within the academic community, see \ref{fig:epf_evolution}.
 In addtion, the United Nations have identified the right to access affordable, reliable, sustainable and modern energy as one of their 17 SDGs \cite{un_sdgs}.
-\\
 Finally, the electricity market has a set of features that make it unique: electricity cannot be stored in an efficient way and supply and demand have to be matched instantly.
@@
 \section{Motivation}
-The are mutliple reasons why the energy sector needs robust forecasting techniques.
-For power market companies, being able to predict prices with a low MAPE \ref{mape} results in increased savings \cite{savings}.
-Furthermore,the adoption of smart meters provides power market companies with a ton of consumer data; this can enable them to better model consumer preferences.
+There are mutliple reasons why the energy sector needs robust forecasting techniques.
+For power market companies, being able to predict prices with a low mean absolute percentage error (MAPE) \ref{mape} results in increased savings \cite{savings}.
+Furthermore, the adoption of smart meters provides power market companies with a ton of consumer data; this can enable them to better model consumer preferences.
 \\

### 2024-04-29 thesis/input_files/problemdescription.tex
comm 1.2

@@
 For power market companies, being able to predict prices with a low mean absolute percentage error (MAPE) \ref{mape} results in increased savings \cite{savings}.
-Furthermore, the adoption of smart meters provides power market companies with a ton of consumer data; this can enable them to better model consumer preferences.
+Furthermore, the adoption of smart meters provides power market companies with a huge amount of consumer data.
+This can enable them to better model consumer preferences.
 \\
-Transmission system operators' main goal is to match supply and deman, generally TSO do so by increasing or decreasing the generation supplied.
-Thus, from their view point forecasting is critical for balancing the electricity netwrok.
-\\
-Probabilistic forecasting may be useful to power producers, traders and consumers in order to improve their decision making process and managing risk(VaR).
-This holds in particular for traders, because probabilistic forecast enables them to simulate scenarios and carry out stress tests.
-\\
-Other possible applications are: control of storage, demand side response, anomaly detection, network design and planning, simulating inputs and handling missing data.
+Transmission system operators' (TSO) main goal is to match supply and demand, generally TSO do so by increasing or decreasing the generation.
+Thus, from their point of view forecasting is critical for balancing the electricity netwrok.
+Probabilistic forecasting may be useful to power producers, traders and consumers in order to improve their decision making process and managing risk.
+This holds in particular for traders, because probabilistic forecasts enables them to simulate scenarios and carry out stress tests.
+Other possible applications include: control of storage, demand side response, anomaly detection, network design and planning, simulating inputs and handling missing data.
 \\
@@
 A distinction has to be made between two types of forecasting approaches: point forecasts and probabilistic forecasts.
-Point prediction, also called deterministic forecasting in the literature, is all about predicting a particular value in time.
-On the other hand, with probabilistic forecasting we aim at predicting either a prediction interval, quantiles or a probability distribution for each point in time.
-For this reason, probabilistic forecasts are more informative than point forecasts; and this is why the interest of the research community is shifting towards them.
-Note that a probabilistic forecast can be turned into a point forecast by simply taking its expecatation.
+Point prediction, also called deterministic forecasting in the literature \cite{EPF_review}, is all about predicting a particular value in time.
+On the other hand, with probabilistic forecasting we aim at predicting either a prediction interval, quantiles or a probability distribution for each point in time \cite{nowotarski}.
+For this reason, probabilistic forecasts are more informative than point forecasts.
+This is why the interest of the research community is shifting towards them.
+A probabilistic forecast can be turned into a point forecast by simply taking its expecatation.
 Alternatively, a probabilistic forecast can be derived from a point one by modeling the residuals of the point prediction.

### 2024-04-29 thesis/input_files/problemdescription.tex
comm 1.4

@@
 \section{Outline}
-We start with a literature review and bibliometric analysis in section \ref{literature_review}.
-Then the theory underlying kernel methods is covered in section \ref{kernel_theory}.
-Following, section \ref{state_of_the_art} introduces the state of the art methods in the context of both point and probabilistic forecasting.
+We start with a literature review and a bibliometric analysis in Section \ref{literature_review}.
+Then, the theory underlying kernel methods is covered in Section \ref{kernel_theory}.
+Evaluation metrics necessary to rank the forecasting techniques are presented in Section \ref{metrics}.
 Section \ref{energy_market} explains the core features and terminology of the energy market and of the electricity newtork.
-Evaluation metrics necessary to rank the forecasting techniques are presented in section \ref{metrics}. Section \ref{eda} goes on with the ETL pipeline and the exploratory analysis. 
+Following, Sections \ref{ch:point} and \ref{ch:prob} introduce the state of the art methods in the context of point and probabilistic forecasting respectively.
+Section \ref{eda} goes on with the etract-load-transform (ETL) pipeline and the exploratory analysis (EDA). 
 Implementation details are included in \ref{implementation}.
-Finally section \ref{experiments} presents the experiments, the results and discusses models' strenghts, weaknesses and possible improvements.
+Finally, section \ref{analysis} presents the experiments, the results and discusses models' strenghts, weaknesses and possible improvements.

### 2024-04-29 thesis/input_files/literaturereview.tex
correct names in contents,list of fig and tables
comm 2.1

@@
 \\
-The field benefitted greatly from the increse of computing power, the greater availability of dataset and the interest in data science.
+The field benefitted greatly from the increse of computing power, the greater availability of data and the interest in data science.
 As a consequence, the forecaster's toolbox has grown in size and complexity.
 \\
-Before delving into the literature review, it is important to make clear that at this point in time there is no superior method.
+Before delving into the literature review, we stress that at this point in time there is no superior method.
 Different solutions may outperform or underperform compared to other techniques depending on the problem settings.
@@
 \\
-Finally, within this research community, it emerged the need of more homogeneity in the choice of the error valuation metrics (section \ref{metrics}), data quality and in the way of comparing model performances \cite{EPF_review}; as a solution, \cite{lago} proposes a checklist to aid evaluating the meaningfulness of new research.
-Hence, throughout this thesis work, we will stick to the proposed principles and best practices peculiar of the EF field.
+Within this research community, it emerged the need for more homogeneity in the choice of the error valuation metrics (Section \ref{metrics}), data quality and in the way of comparing model performances \cite{EPF_review}.
+As a solution, \cite{lago} proposes a checklist to aid evaluating the meaningfulness of new research.
+Throughout this thesis work, we will stick to the proposed principles and best practices peculiar of the EF field.
 %literature review
@@
 Thus, in order to introduce some clarity it is useful to classify the range of EF articles in terms of their core attributes.
-\subsection{Types}
-In the context of energy forecasting, the quantities of most interest are prices (EPF), loads (ELF) and renewables generation (mostly wind and solar).
-
-\subsection{Forecasting horizons}\label{cls_time}
-In terms of forecasting horizons, we can group EF into four major categories: very short term forcasting (VSTF), short term load forecasting (STF), medum term forecasting (MTF) and long term forecasting (LTF).
-Consensus in the literature is to use as cut off horizons one day, two weeks and three years respectilvely \cite{hong_phd}; see \ref{fig:time} for a visualisation.
+\\
+In the context of energy forecasting, the quantities of most interest are electricity prices (EPF), electricity loads (ELF) and renewables generation (mostly wind and solar).
+\\
+In terms of forecasting horizons, we can group EF into four major categories: very short term forcasting (VSTF), short term load forecasting (STF), medium term forecasting (MTF) and long term forecasting (LTF).
+Consensus in the literature is to use one day, two weeks and three years respectilvely \cite{hong_phd} as cut off horizons; see \ref{fig:time} for a visualisation.
 \begin{figure}
-  \includegraphics[width=\textwidth]{images/time.jpg}
+  \includegraphics[width=\textwidth]{images/time_2.png}
   \caption{Time classsification \cite{prob_elf}}
@@
 \end{figure}
-
-\subsection{Size}
+\\
 Forecasts can either be for the whole target electricity network (system) or for a subset of it (zonal).
-
-\subsection{Point vs Probabilistic}\label{cls_pp}
-Finally, EF literature distinguishes between point and probabilistic forecasts.
-Each of the two has its advantages and disadvantages; point forecasts are easier to generate and less computationally intensive while probabilistic forecast are more informative.
+\\
+EF literature distinguishes between point and probabilistic forecasts.
+Each of the two has its advantages and disadvantages.
+Point forecasts are easier to generate and less computationally intensive while probabilistic forecast are more informative.
 Industry and research efforts have focused primarily on point forcasting.
@@
     \includegraphics[width=\textwidth]{images/epf_evolution1.jpg}
-    \caption{EF publications}
+    \caption{Electricity forecasting publications over the past years}
     \label{fig:epf_evolution}
@@
   \includegraphics[width=\textwidth]{images/point_vs_prob.jpg}
-  \caption{Point vs probabilistic pubblications}
+  \caption{Point versus probabilistic pubblications}
   \label{fig:point_vs_prob}
@@
 \\
-\section{EF literature review}
+\section{Electricity forecasting literature review}
 To get started a few review articles were collected in order to understand conventions, best practicies and terminology of the EF community.

### 2024-04-29 thesis/input_files/literaturereview.tex
comm 2.1 2.2

@@
 This survey has been carried out by using the Scopus citation database.
-For details on the specific queries entered in Scopus, refer to \ref{src_code}.
+For details on the specific queries entered in Scopus, refer to Subsection 5 of the Appendix \ref{src_code}.
 \\
 To get started, let us consider the evolution of the EF field over the years.
-This is visualized in figure \ref{fig:epf_evolution}, with results grouped by category; note, articles prior to the 2000 have been aggregated together due to their small number.
+This is visualized in Figure \ref{fig:epf_evolution}, with results grouped by category.
+Articles prior to the 2000 have been aggregated together due to their small number.
 Figure \ref{fig:epf_evolution} shows the trend of an increasing interest in EF.
@@
 \\
-The next question was to compare the state of point versus probabilistic forecasting, this is visualized in figure \ref{fig:point_vs_prob}.
-What it can be concluded is that probabilistic is less developed than point forecasting.
-To our mind this is due to the complexity of probabilistic forecast.
+The next question is to compare the state of point versus probabilistic forecasting, this is visualized in Figure \ref{fig:point_vs_prob}.
+What can be concluded is that probabilistic is less developed than point forecasting.
+To our mind this is due to the complexity of probabilistic forecasts.
 Nevertheless, we can see a trend that suggests researchers are making an effort to fill this gap.
@@
   \includegraphics[width=\textwidth]{images/point_vs_prob.jpg}
-  \caption{Point versus probabilistic pubblications}
+  \caption{Point versus probabilistic pubblications over the past years}
   \label{fig:point_vs_prob}
@@
 \\
-The EF literature is dominated by statistical and computational intelligence methods as can be seen from figure \ref{fig:cs_stat_both}, with CI methods beign slightly preferred. 
+The EF literature is dominated by statistical and computational intelligence (CI) methods as can be seen from Figure \ref{fig:cs_stat_both}, with CI methods being slightly preferred. 
 \begin{figure}
   \includegraphics[width=\textwidth]{images/cs_stat_both.jpg}
-  \caption{Pubblications by method}
+  \caption{Pubblications by method over the past years}
   \label{fig:cs_stat_both}
@@
 \\
-EF is an heterogenous field of research, its researchers come from a wide array of backgrounds, with electrical engineers and statisticians making up the top contributors; their different educational training may explain why the split between statistical and computational intelligence methods is so marked.
+EF is a heterogenous field of research, its researchers come from a wide array of backgrounds, with electrical engineers and statisticians making up the top contributors; their different educational training may explain why the split between statistical and computational intelligence methods is so marked.
 Figure \ref{fig:subject_area} depitcs the EF publications by subject area.
-What can be concluded, is that the bulk of pubblications come from engineering, computer science, mathematics and econometrics.
+What can be concluded, is that the bulk of pubblication come from engineering, computer science, mathematics and econometrics.
 \begin{figure}
-  \includegraphics[width=\textwidth]{images/subject_area.jpg}
+  \includegraphics[width=\textwidth]{images/subject_area.png}
   \caption{Pubblications by subject area}
@@
 \\
-Finally, in order to refer to the most relevant source in the field, EF outlets have been ranked by popularity and plotted in figure \ref{fig:src_title}.
+In order to refer to the most relevant source in the field, EF outlets have been ranked by popularity and plotted in Figure \ref{fig:src_title}.
 \begin{figure}
-  \includegraphics[width=\textwidth]{images/src_title.jpg}
+  \includegraphics[width=\textwidth]{images/src_title.png}
   \caption{Most popolar sources/outlets}

### 2024-04-29 thesis/input_files/literaturereview.tex
comm 2.3

@@
 To get started a few review articles were collected in order to understand conventions, best practicies and terminology of the EF community.
-Weron \cite{EPF_review} reviews the state of the art for electricity price forecasting; beside analysing complexity of available solutions, strenghts and weaknesses it also stresses the need for objective comparative EPF studies.
+Weron \cite{EPF_review} reviews the state of the art for electricity price forecasting.
+Besides analysing complexity of available solutions, strenghts and weaknesses it also stresses the need for objective comparative EPF studies.
 Specifically, it advocates for studies using similar datasets, using the same error evaluation metric and statistical testing model's outperformance.
 Hong et al.
-\cite{prob_elf} discusses the state of the art in probabilistic electric load forecasting; it differentiates between techniques and methodologies.
+\cite{prob_elf} discusses the state of the art in probabilistic electric load forecasting.
+It differentiates between techniques and methodologies.
 With techniques they refer to a family of models, like multiple linear regression or artificial neural networks.
@@
 Also this paper stresses the need for some guidelines to standardize research in the field.
-Nowotarski et al. \cite{nowotarski} also carried out a thorough review of probabilistic forecasting.
+Nowotarski et al. \cite{nowotarski} carried out a review of probabilistic forecasting.
 Weron et al. \cite{lago} offer a set of best practices when forecasting electricity prices in order to have a common framework to evaluate and compare future research.
 Zhang et al. \cite{zhang2014review} considers state of the art methods in wind power probabilistic forecasting and describes current challenges and possible future developments.
-Ziel et al. \cite{ziel2018probabilistic} provides detailed tables, grouping research papers by the time dimension \ref{cls_time} and objective \ref{cls_pp} and reporting dataset, model and accuracy 
+Ziel et al. \cite{ziel2018probabilistic} provides detailed tables, groups research papers by the time dimension and objective and reports dataset, model and accuracy 
 measures adopted.
 David et al.
-\cite{david2016probabilistic} adopt a combination of ARMA and GARCH in probabilistic forecasts of solar irradiance.
+\cite{david2016probabilistic} adopt a combination of autoregressive moving average (ARMA) and generalized autoregressive conditional heteroskedasticity (GARCH) in probabilistic forecasts of solar irradiance.
 Furthermore, they propose a recursive framework for parameter estimation.
@@
 He et al.
-\cite{he2022cooperative} models  multistep wind speed probabilistic forecasting by mixing CEEDMAN, LASSO and QRNN.
-In this work, CEEMDAN is used to decompose the wind speed time series, LASSO compresses high dimensional features, QR is used for obtaining quantile forecasts, finally KDE converts quantile forecasts into density estimates.
-Wan et al. \cite{wan2016direct} proposes a combination of QR and ELM to generate non parametric probabilistic forecasts of wind generation.
+\cite{he2022cooperative} models  multistep wind speed probabilistic forecasting by mixing complementary ensemble empirical mode decomposition(CEEDMAN), least absolute shrinkage and selection operator (LASSO) and quantile regression (QR).
+In this work, CEEMDAN is used to decompose the wind speed time series, LASSO compresses high dimensional features, QR is used for obtaining quantile forecasts, finally kernel density estimation (KDE) converts quantile forecasts into density estimates.
+Wan et al. \cite{wan2016direct} proposes a combination of QR and extreme learning machine (ELM) to generate non parametric probabilistic forecasts of wind generation.
 Zhang et al. \cite{zhang2019wind} forecasts wind speed by adopting QRMGM, which is a combination of QR with a minimal gated memory network.
@@
 implementations, and applications \cite{gm_22}.
-Marcjasz et al. \cite{probablistic_electricity_forecast2} uses distributional neural networks to create probabilistic forecasts for the day ahead electricity prices in the german market.
+Marcjasz et al. \cite{probablistic_electricity_forecast2} uses distributional neural networks to create probabilistic forecasts for the day ahead electricity prices in the German market.
 Nowotarski et al.
-\cite{nowotarski2015computing} introduce a method for constructing PI and call it quantile regression averaging.
-Their idea is weighting a set of models' prediction such that the pinball loss of the weighted model is minimized.
-The observed results is that QRA performed better compared to twelve individual models.
+\cite{nowotarski2015computing} introduce a method for constructing prediction intervals (PI) and call it quantile regression averaging.
+Their idea is weighting a set of model predictions such that the pinball loss of the weighted model is minimized.
+The observed result is that QRA performed better compared to twelve individual models.
 Arora et al.
-\cite{arora2016forecasting} focus on modelling electricity smart meter data by proposing a non parametric probabilistic technique based on kernel density estimation and conditional kernel density estimation \cite{rosenblatt1969conditional} \cite{hyndman1996estimating}.
+\cite{arora2016forecasting} focus on modelling electricity smart meter data by proposing a non parametric probabilistic technique based on kernel density estimation and conditional kernel density estimation \cite{rosenblatt1969conditional, hyndman1996estimating}.
 Their conclusion is that kernel density methods are competitive against exponential smoothing when forecasting residential data.
-Conversely, exponential smoothing has still an edge in predicting SME data.
+Conversely, exponential smoothing has still an edge in predicting small to medium-sized enterprises data (SMEs).
 Zhang et al.
 \cite{zhang2020probability} introduce a framework based on quantile regression and kernel density estimation in the context of short term wind forecasting.
-The proposed methods behave well compared to an ARMA(1,1).
-Haben et al. \cite{haben2018short} analyses a variety of techniques in terms of both probabilistic and point forecasting, within this study they focus on load forecasting at the low voltage level.
+The proposed methods behave well compared to an autoregressive ARMA(1,1) model.
+Haben et al. \cite{haben2018short} analyses a variety of techniques in terms of both probabilistic and point forecasting. Within this study, they focus on load forecasting at the low voltage level.
 %Metrics Start
@@
 Gneiting et al. \cite{gneiting2007strictly}
-provides a thorough overview of the theory of proper scoring rules for interval and density forecasts.
+provides a thorough overview of the theory of scoring rules for interval and density forecasts.
 Gneiting et al. \cite{gneiting2014probabilistic}
-covers theory and techniques state of the art in probabilistic forecasting.
+covers theory and state of the art techniques in probabilistic forecasting.
 %Metrics End
 Zhang et al \cite{zhang2020two} proposes a two stage bootstrap sampling framework for probabilistic load forecasting.
-They test it for different regression models such as RF, GBRT, linear regression, and LSSVM regression.
+They test it for different regression models such as random forest (RF), gradient boosting regression tree (GBRT), linear regression, and least squares support vector regression (LSSVM).
 Jónsson et al. \cite{jonsson2014predictive}
@@
 Fatema et al.
-\cite{fatema2023probabilistic} considers gaussian process regression for point forecasting and prediction intervals prediction.
-Then, it inputs PI to KDE in order to estimate a probability distribution.
-Dudek \cite{dudek2018probabilistic} proposes a probabilistic forecasting model based on the Nadaraya Watson estimator \cite{nadaraya1964estimating} \cite{watson1964smooth}.
+\cite{fatema2023probabilistic} considers Gaussian process regression for point forecasting and prediction intervals prediction.
+Then, it inputs prediction intervals to kernel density estimation in order to estimate a probability distribution.
+Dudek \cite{dudek2018probabilistic} proposes a probabilistic forecasting model based on the Nadaraya Watson estimator \cite{nadaraya1964estimating,watson1964smooth}.
 Huurman et al.
-\cite{huurman2012power} surveys the predictive power of weather variables for electricity prices in the danish market.
+\cite{huurman2012power} surveys the predictive power of weather variables for electricity prices in the Danish market.
 Their empirical results suggest that weather is central for point forecasting day ahead prices.
@@
 \\
-Lately, the idea of combining forecasts has gained popularity in the forecasting community \cite{forecasting_big}; in the literature, combined forecasts are called ensemble \cite{gneiting_weather_ensemble}.
+Lately, the idea of combining forecasts has gained popularity in the forecasting community \cite{forecasting_big}.
+In the literature, combined forecasts are called ensemble \cite{gneiting_weather_ensemble}.
 Experimental results have shown ensemble methods to outperform their component forecasts.
-Note that, the more the errors of the combined models are not correlated the more we can benefit from ensembles.
-It is also worth noting that older and simpler methods are still valuable (in combination with other models or on their own); these being less subject to overfitting than complex models.
+Note that, the more the errors of the combined models are not correlated, the more we can benefit from ensembles.
+It is also worth noting that older and simpler methods are still valuable (in combination with other models or on their own). These being less subject to overfitting than complex models.
 \\
@@
 Addressing these issues was the motivation behind the creation of GEFCom by the IEEE working group on energy forecasting.
-The EF field was positively affected by this competition; a number of ideas were tested on the same setting with only the best ones being published and it also contributed bridging the gap between industry practice and academic research.
+The EF field was positively affected by this competition.
+A number of ideas were tested on the same setting with only the best ones being published and it also contributed bridging the gap between industry practice and academic research.
 GEFCom 2012 had two tracks; the former about hierarchical load forecasting, the latter about wind power forecasting, see \cite{hong2014global} for a comprehensive review.
@@
 Xie et al.
-\cite{xie2016gefcom2014} propose a two stage approach; in the first stage they use MLR to build a point forecast, then in the second stage they try different approaches for modelling the MLR residuals, among other they tried ESM, ANN and ARIMA.
-Maciejowska et al. \cite{maciejowska2016hybrid} proposes a new probabilistic model extending the idea of QRA \cite{nowotarski2015computing}.
-Haben et al. \cite{haben2016hybrid} mixes CKD and QR in their competition entry.
-Gaillard et al. \cite{gaillard2016additive} \cite{gaillardasemi} combines quantile regression with generalized additive models \cite{hastie2017generalized}.
+\cite{xie2016gefcom2014} propose a two stage approach; in the first stage they use multiple linear regression (MLR) to build a point forecast, then in the second stage they try different approaches for modelling the MLR residuals, among other they tried exponential smoothing (ESM), artificial neural networks (ANN) and autoregressive integrated moving average (ARIMA).
+Maciejowska et al. \cite{maciejowska2016hybrid} proposes a new probabilistic model extending the idea of quantile regression averaging (QRA) \cite{nowotarski2015computing}.
+Haben et al. \cite{haben2016hybrid} mixes conditional kernel density (CKD) and quantile regression in their competition entry.
+Gaillard et al. \cite{gaillard2016additive,gaillardasemi} combines quantile regression with generalized additive models \cite{hastie2017generalized}.
 Ziel et al. \cite{ziel2016lasso} estimates an AR model through the LASSO \cite{tibshirani1996regression} instead of the standard OLS.
@@
 A couple of considerations can be drawn from the above literature review.
-As already pointed out, the EF field is characterized by heterogeneity in its forecasting techniques; methods come from statisitics, mathematics, econometrics, electrical engineering and the artificial intelligence communities.
+The EF field is characterized by heterogeneity in its forecasting techniques; methods come from statisitics, mathematics, econometrics, electrical engineering and the artificial intelligence communities.
 \\
 Every paper uses different datasets.
-Therefore, it is not possible to compare directly results from one paper to another without implementing the paper specific algorithms and then applying them to your dataset.
+Therefore, it is not possible to compare directly results from one paper to another without implementing the paper specific algorithms and then applying them to the respective dataset.
 An additional hurdle is that some datasets are not freely accessible.

### 2024-04-29 thesis/input_files/literaturereview.tex
comm 2.4

@@
 %INTRO KERNEL METHODS E ENERGY PRICE MODELLING
-Kernel methods are a class of algorithms for patter analysis.
+Kernel methods are a class of algorithms for pattern analysis.
 With kernel methods we are able to apply linear methods with predictors in a high dimensional space, without having to explicitly evaluate the involved dot products of the features.
@@
 %- Kernel history evolution.
-Their name comes from the german word kern, which translates to core in english.
-Such term was first used by David Hilbert in his paper on integral equations \cite{hilbert} where he introduced the term of definite kernels.
-Following, Hilbert's and Schmidt's \cite{schmidt} work lead to the introduction of a new space, the Hilbert space.
+Their name comes from the German word kern, which translates to core in english.
+Such term was first used by David Hilbert in his paper on integral equations \cite{hilbert} where he introduced the term "definite kernels".
+Following, Hilbert's and Schmidt's work \cite{schmidt} lead to the introduction of a new space, the Hilbert space.
 In 1909, James Mercer improved Hilbert's work by proposing his theorem \cite{mercer}. This theorem underlies the power of kernel methods, that is the kernel trick.
@@
 In 1941, Kolmogorov \cite{kolmogorov} carried out stuides on representing kernel in linear spaces.
-In 1950, Aronszajn \cite{aronszajn} published the first work on RKHS; developing the general method for representing kernels in linear spaces.
+In 1950, Aronszajn \cite{aronszajn} published the first work on reproducing kernel hilbert spaces (RKHS); developing the general method for representing kernels in linear spaces.
 In 1964, Aizerman \cite{aizerman} further improved the theory of RKHS.
@@
 \\
-The kernel theory needed for this thesis work is covered in section \ref{kernel_theory}. 
-For an introduction to kernel methods, we referred to \cite{learning_with_kernels} \cite{hofmann2006review}.
+The kernel theory needed for this thesis work is covered in Section \ref{kernel_theory}. 
+For an introduction to kernel methods, we refer to \cite{learning_with_kernels, hofmann2006review}.
 \\
@@
 \cite{2022nystrom} proposes an efficient approximation procedure based on the Nyström method \cite{nystrom}, providing also an upper bound for the approximation error.
-Article \cite{supersamples} presents kernel herding; basically, Smola et al.
-used the kernel trick to extend the herding algorithm to continous spaces.
-The result is an infinite memory deterministic process that takes in a collection of samples and learns to approximate a pdf.
+Smola et al.
+\cite{supersamples} presents kernel herding; basically, the authores used the kernel trick to extend the herding algorithm to continous spaces.
+The result is an infinite memory deterministic process that takes in a collection of samples and learns to approximate a probability density function (PDF).
 \\

### 2024-04-29 thesis/input_files/probabilistic.tex
March 29th 2024
April 6th 2024
correct names in contents,list of fig and tables
update melbourne plot

@@
       \centering
-      \includegraphics[width=1.1\textwidth]{images/melborune_kernel_quantile_regression.png} 
+      \includegraphics[width=1.1\textwidth]{images/melbourne_kernel_quantile_regression.png}
       \caption{Kernel quantile regression} 
@@
 \begin{table}[!h]
-\caption{MAE Melbourne data}
+\caption{Mean absolute error Melbourne data}
 \begin{tabular}{lllll}
@@
 
-\section{Kernel density estimation}
-\input{input_files/kde.tex}
+% \section{Kernel density estimation}
+% \input{input_files/kde.tex}
 
-\section{Ensemble methods}
-\input{input_files/ensemble.tex}
+% \section{Ensemble methods}
+% \input{input_files/ensemble.tex}
 
-\section{DMLP}
-%gluon nn
-% train a beta distribution to learn its parameters
-\section{DeepAR}
+% \section{DMLP}
+% %gluon nn
+% % train a beta distribution to learn its parameters
+% \section{DeepAR}
 
-\subsection{Kernel herding}
-Select the best point forecasting method and create a probabilistic forecast by modelling its model errors with kernel herding (do something like the residual bootstrap ensembles if it is meaningful and possible to implement).
+% \subsection{Kernel herding}
+% Select the best point forecasting method and create a probabilistic forecast by modelling its model errors with kernel herding (do something like the residual bootstrap ensembles if it is meaningful and possible to implement).

### 2024-04-29 thesis/input_files/paper1.tex
correct names in contents,list of fig and tables
comm 3.1

@@
 
-\section{Kernel Mean Embedding of Distributions: A Review and Beyond}
+\section{Kernel mean embedding of distributions: A review and beyond}
 % \href{https://arxiv.org/abs/1605.09522}{
@@
 \begin{align*}
-\phi(x):\mathcal{X} &\longrightarrow \mathcal{F}
+\phi(x)\colon \mathcal{X} &\to \mathcal{F}
     \\
-    x&\mapsto \phi(x)
+    x&\to \phi(x)
 \end{align*}
 
-Then we take the inner product in the high dimensional space $\mathcal{F}$ mapped by $\phi(x)$.
+We take the inner product in the high dimensional space $\mathcal{F}$ mapped by $\phi(x)$, i.e.\
 \\
@@
 \\
-$\phi(x)$ is referred as feature map while $k$ as kernel function.
+where $\phi(x)$ is referred to as feature map while $k$ is the kernel function.
 \\
@@
 \\
-One would expect constructing the feature maps explicitly and then evaluate their inner product in $\mathcal{F}$ to be computationally expensive, and indeed it is.
+One would expect constructing the feature maps explicitly and then evaluating their inner product in $\mathcal{F}$ to be computationally expensive, and indeed it is.
 However, we do not have to explicitly perform such calculations.
-This is because of the existence of the kernel trick.
+This is because of the kernel trick.
 To illustrate the idea behind the kernel trick consider the following example. 
 \\
-Suppose $x \in \mathbb{R}^2$ and assume to select $\phi(x)=(x_{1}^{2},x_{2}^{2},\sqrt{2}x_{1}x_{2})$, then the inner product in the feature space is $x_{1}^{2}x_{1}^{'2},x_{2}^{2}x_{2}^{'2}+2x_{1}x_{2}x'_{1}x'_{2}$.
+Suppose $x \in \mathbb{R}^2$ and assume $\phi(x)=(x_{1}^{2},x_{2}^{2},\sqrt{2}x_{1}x_{2})$, then the inner product in the feature space is $x_{1}^{2}x_{1}^{'2},x_{2}^{2}x_{2}^{'2}+2x_{1}x_{2}x'_{1}x'_{2}$.
 Notice that this is the same of $\langle \phi(x), \phi(x') \rangle$; thus the kernel trick consists of just using $k(x,x')=:(x^Tx')^2$.
 
-\subsection{RKHS}
-Following are the definitions that make up the basis for the theory of kernel methods.
+\subsection{Reproducing kernel hilbert space}
+Subsequently we introduce the definitions that make up the basis for the theory of kernel methods.
 
 \begin{definition}
-    A sequence $\{v_n\}_{n=1}^{\infty}$ of elements of a normed space $\mathcal{V}$ is a Cauchy sequence if for every $\epsilon>0$, there exist $N=N(\epsilon) \in \mathbf{N}$ such that $\|v_n-v_m\|_{\mathcal{V}}<\epsilon \ \ \forall m,n\geq N$
+    A sequence $\{v_n\}_{n=1}^{\infty}$ of elements of a normed space $\mathcal{V}$ is a Cauchy sequence if for every $\epsilon>0$, there exist $N=N(\epsilon) \in \mathbb{N}$ such that $\|v_n-v_m\|_{\mathcal{V}}<\epsilon \ \ \forall m,n\geq N$
 \end{definition}
@@
 \begin{definition}
-    A Hilbert space is a vector space $\mathcal{H}$ with an inner product $\langle f, g \rangle$ such that the norm defined by $\|f\|=\sqrt{\langle f, f \rangle}$
+    A Hilbert space is a vector space $\mathcal{H}$ with an inner product $\langle \cdot, \cdot \rangle$ such that the norm defined by $\|f\|=\sqrt{\langle f, f \rangle}$
 turns $\mathcal{H}$ into a complete metric space.
@@
 \begin{definition}
-    RKHS.
-    A Reproducing Kernel Hilbert Space is an Hilbert space with the evaluation functionals $\mathcal{F}_{x}(f):=f(x)$ bounded, i.e.
-$\forall x \in \mathcal{X}$ there exists some $C>0$ such that $\| \mathcal{F}_{x}(f)\|=\|f(x)\| \leq C \|f\|_{\mathcal{H}} \ \forall f \in \mathcal{H}$
+    Let $(\mathcal{H}, \langle \cdot, \cdot \rangle_\mathcal{H})$ be a Hilbert space of real-valued functions on $\mathcal{X}$.
+A function $k: \mathcal{X} \times \mathcal{X} \to \mathbb{R}$ is called a reproducing kernel of $\mathcal{H}$ if and only if
+     \begin{align}
+        k(x, \cdot) \in \mathcal{H} &\quad \text{for all   } x \in \mathcal{X}    \\
+        \langle  h, k(x, \cdot) \rangle_\mathcal{H} = h(x) &\quad \text{for all   } h \in \mathcal{H}, x \in \mathcal{X}
+    \end{align}
 \end{definition}
 
+% \begin{definition}
+%     RKHS.
+%     A Reproducing Kernel Hilbert Space is a Hilbert space with the evaluation functionals $\mathcal{F}_{x}(f):=f(x)$ bounded, i.e.
+$\forall x \in \mathcal{X}$ there exists some $C>0$ such that $\| \mathcal{F}_{x}(f)\|=\|f(x)\| \leq C \|f\|_{\mathcal{H}} \ \forall f \in \mathcal{H}$
+% \end{definition}
+
 \begin{theorem}
-    Riesz Representation.
-If $A : \mathcal{H} \rightarrow \mathbf{R}$ is a bounded linear operator in a Hilbert space $\mathcal{H}$ , there exists some $g_{A} \in \mathcal{H}$ such that $A(f) = \langle f,g_A\rangle_\mathcal{H}, \ \forall f \in \mathcal{H}$.
+    [Riesz Representation].
+If $A : \mathcal{H} \rightarrow \mathbb{R}$ is a bounded linear operator on a Hilbert space $\mathcal{H}$ , there exists some $g_{A} \in \mathcal{H}$ such that $A(f) = \langle f,g_A\rangle_\mathcal{H}, \ \forall f \in \mathcal{H}$.
 \end{theorem}
 
-The Riesz representation theorem results in the following proposition for RKHS.
+The Riesz representation theorem results in the following proposition for reproducing kernel hilbert spaces.
 \begin{proposition}
@@
 
-The function $k_{x}$ is the reproducing kernel for the point $x$.
-Furthermore, note that $k_{x}$ is itself a function lying on $\mathcal{X}$
+The function $k_{x}$ is the reproducing kernel evaluated at point $x$.
+Furthermore, note that $k_{x}$ is itself a function lying in it.
 
 \begin{align*}
-    k_{x}(y)=\mathcal{F}_{y}(k_{x})=\langle k_{x}, k_{y}\rangle_{\mathcal{H}=\langle \phi(x), \phi(x)\rangle_{\mathcal{H}}}
+    k_{x}(y)=\mathcal{F}_{y}(k_{x})=\langle k_{x}, k_{y}\rangle_{\mathcal{H}=\langle \phi(x), \phi(y)\rangle_{\mathcal{H}}}
 \end{align*}
@@
        Periodic &  $k(x_1,x_2)=e^{\frac{-2 sin^2\left(\frac{\pi}{p}|x_1-x_2| \right)}{l^2}} $ &   p, l\\
+       Chi-squared kernel &  $k(x_1,x_2)=e^{-\gamma \sum\limits_i \frac{(x_i - y_i)^2}{x_i+y_i}} $ &   $\gamma$\\
+        Cosine  &  $k(x_1,x_2)=\frac{xy^\intercal}{\|x\|_{L_2}\|y\|_{L_2}}$ &   \\
+       Matern  &  $k(x_1, x_2) =  \frac{1}{\gamma(\nu)2^{\nu-1}}\Bigg(
+        \frac{\sqrt{2\nu}}{l} d(x_1 , x_2 )
+        \Bigg)^\nu K_\nu\Bigg(
+        \frac{\sqrt{2\nu}}{l} d(x_1 , x_2 )\Bigg)$ &  $\gamma, \nu$ \\
+
         \bottomrule

### 2024-05-01 thesis/input_files/probabilistic.tex
melbourne kernels

@@
 
+We conclude this chapter with reporting the same kind of tables \ref{tab:kernel pinball comparison}, \ref{tab:kernel pinball comparison quantile-wise}, \ref{tab:kernel mae comparison} and figures \ref{fig:kernel quantile regressors comparison} comparing various kernel functions.
+
+% tables
+\begin{table}[!h]
+    \caption{Kernels comparison pinball loss Melbourne data}
+    \label{tab:kernel pinball comparison}
+    \begin{center}
+    \begin{tabular}{lll}
+        \toprule
+        & Kernel & Pinball loss
+        \\
+        \midrule
+        & Gauss RBF &  10.031708 \\
+        & Laplacian &  10.056884 \\
+        & Gauss RBFxLaplacian & 10.150826 \\
+        & Cosine & 16.253973    \\
+        & Linear & 10.463867    \\
+        & Polynomial & 11.238393     \\
+        & Sigmoid & 16.253973            \\
+        & Chi squared & 10.023732       \\
+        & Matern & \textbf{10.021369}  \\
+        & Periodic  & 15.946272\\
+        \bottomrule
+        \end{tabular}
+    \end{center}
+    \end{table}
+
+\begin{table}[!h]
+    \caption{Kernels comparison pinball loss quantile-wise Melbourne data}
+    \label{tab:kernel pinball comparison quantile-wise}
+    \begin{adjustbox}{width=\textwidth}
+    \begin{tabular}{lllllllllll}
+    \toprule
+    Kernel/Quantiles & 0.1 & 0.2 & 0.3 & 0.4 & 0.5 & 0.6 & 0.7 & 0.8 & 0.9 \\
+    \midrule
+    Gauss RBF &  \textbf{0.540145} &
+    \textbf{0.903193} &
+    1.173627 &
+    1.368022 &
+    1.456053 &
+    1.447470 &
+    \textbf{1.331375} &
+    1.096649 &
+    0.715013 \\
+    Laplacian &  0.542499 &
+    0.906517 &
+    1.177309 &
+    1.371198 &
+    1.462111 &
+    1.453119 &
+    1.340193 &
+    \textbf{1.095991} &
+    0.707947 \\
+    Gauss RBFxLaplacian &0.545787 &
+    0.915559 &
+    1.189965 &
+    1.387109 &
+    1.479901 &
+    1.475541 &
+    1.350803 &
+    1.101151 &
+    \textbf{0.704922} \\
+    Cosine & 0.755479 &
+    1.343973 &
+    1.802603 &
+    2.123425 &
+    2.307123 &
+    2.367123 &
+    2.262603 &
+    1.971644 &
+    1.320411 \\
+    Linear &  0.565663 &
+    0.947889 &
+    1.250009 &
+    1.437863 &
+    1.506723 &
+    1.478379 &
+    1.358865 &
+    1.135702 &
+    0.782660   \\
+    Polynomial &   0.542573 &
+    0.908681 &
+    1.178357 &
+    \textbf{1.362351} &
+    1.450249 &
+    1.455228 &
+    1.339822 &
+    1.107768 &
+    0.744118  \\
+    Sigmoid &   0.755479 &
+    1.343973 &
+    1.802603 &
+    2.123425 &
+    2.307123 &
+    2.367123 &
+    2.262603 &
+    1.971644 &
+    1.320411          \\
+    Chi squared &   0.541430 &
+    0.906352 &
+    1.173164 &
+    1.358862 &
+    \textbf{1.444130} &
+    \textbf{1.440915} &
+    1.332140 &
+    1.100262 &
+    0.726625     \\
+    Matern &  0.541192 &
+    0.903209 &
+    \textbf{1.169543} &
+    1.364430 &
+    1.451248 &
+    1.442511 &
+    1.332522 &
+    1.096750 &
+    0.716226 \\
+    Periodic  & 0.759452 &
+    1.352747 &
+    1.809921 &
+    2.125765 &
+    2.314718 &
+    2.379657 &
+    2.274711 &
+    1.976768 &
+    1.328630 \\
+    \bottomrule
+    \end{tabular}
+    \end{adjustbox}
+\end{table}
+
+\begin{table}[!h]
+    \caption{Kernels comparison mean absolute error Melbourne data}
+    \label{tab:kernel mae comparison}
+    \begin{center}
+    \begin{tabular}{lll}
+        \toprule
+         & Kernel & MAE \\
+        \midrule
+        & Gauss RBF &  3.024336 \\
+        & Laplacian &  3.038633 \\
+        & Gauss RBFxLaplacian &  3.070812\\
+        & Cosine &    4.849589 \\
+        & Linear &     3.142352\\
+        & Polynomial &     3.064472 \\
+        & Sigmoid &      4.849589       \\
+        & Chi squared &    \textbf{3.022778}    \\
+        & Matern &  3.024517 \\
+        & Periodic  & 4.880673 \\
+        \bottomrule
+        \end{tabular}
+    \end{center}
+    \end{table}
+
+% kernels figure
+
+\begin{figure}[!h]
+    \begin{subfigure}[b]{0.5\linewidth}
+        \centering
+        \includegraphics[width=1.1\textwidth]{images/melborune_gaussian_rbf_kernel_quantile_regression.png} 
+        \caption{Gauss RBF} 
+        \label{} 
+        \vspace{4ex}
+    \end{subfigure}%% 
+    \begin{subfigure}[b]{0.5\linewidth}
+        \centering
+        \includegraphics[width=1.1\textwidth]{images/melborune_laplacian_kernel_quantile_regression.png} 
+        \caption{Laplacian} 
+        \label{} 
+        \vspace{4ex}
+    \end{subfigure} 
+    \begin{subfigure}[b]{0.5\linewidth}
+        \centering
+        \includegraphics[width=1.1\textwidth]{images/melborune_gaussian_rbf_x_laplacian_kernel_quantile_regression.png} 
+        \caption{Gauss RBFxLaplacian} 
+        \label{} 
+        \vspace{4ex}
+    \end{subfigure}%%
+    \begin{subfigure}[b]{0.5\linewidth}
+        \centering
+        \includegraphics[width=1.1\textwidth]{images/melborune_cosine_kernel_quantile_regression.png}
+        \caption{Cosine} 
+        \label{} 
+        \vspace{4ex}
+    \end{subfigure} 
+    \begin{subfigure}[b]{0.5\linewidth}
+        \centering
+        \includegraphics[width=1.1\textwidth]{images/melborune_linear_kernel_quantile_regression.png}
+        \caption{Linear} 
+        \label{} 
+        \vspace{4ex}
+    \end{subfigure} 
+    \begin{subfigure}[b]{0.5\linewidth}
+        \centering
+        \includegraphics[width=1.1\textwidth]{images/melborune_polynomial_kernel_quantile_regression.png}
+        \caption{Polynomial} 
+        \label{} 
+        \vspace{4ex}
+    \end{subfigure} 
+    \end{figure}
+
+\begin{figure}[!h]\ContinuedFloat
+    \begin{subfigure}[b]{0.5\linewidth}
+        \centering
+        \includegraphics[width=1.1\textwidth]{images/melborune_sigmoid_kernel_quantile_regression.png}
+        \caption{Sigmoid} 
+        \label{} 
+        \vspace{4ex}
+    \end{subfigure} 
+    \begin{subfigure}[b]{0.5\linewidth}
+        \centering
+        \includegraphics[width=1.1\textwidth]{images/melborune_chi_squared_kernel_quantile_regression.png}
+        \caption{Chi squared} 
+        \label{} 
+        \vspace{4ex}
+    \end{subfigure} 
+    \begin{subfigure}[b]{0.5\linewidth}
+        \centering
+        \includegraphics[width=1.1\textwidth]{images/melborune_matern_kernel_quantile_regression.png}
+        \caption{Matern} 
+        \label{} 
+    \end{subfigure} 
+    \begin{subfigure}[b]{0.5\linewidth}
+        \centering
+        \includegraphics[width=1.1\textwidth]{images/melborune_periodic_kernel_quantile_regression.png}
+        \caption{Periodic} 
+        \label{} 
+    \end{subfigure} 
+    \caption{Kernel quantile regressors comparison}
+    \label{fig:kernel quantile regressors comparison} 
+\end{figure}
+
 % \section{Kernel density estimation}

### 2024-05-01 experiments/kqr_kernels_comp.py

@@ -1,0 +1,110 @@
+from kernel_quantile_regression.kqr import KQR
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas as pd
+from pprint import pprint
+
+from quantile_forest import RandomForestQuantileRegressor as rfr
+
+from sklearn.ensemble import GradientBoostingRegressor as gbr
+
+from sklearn.experimental import enable_halving_search_cv  # noqa
+from sklearn.model_selection import HalvingRandomSearchCV
+from sklearn.metrics import make_scorer
+from sklearn.metrics import mean_absolute_error
+from sklearn.metrics import mean_pinball_loss
+from sklearn.model_selection import train_test_split
+
+import statsmodels.regression.quantile_regression as qr 
+import sys
+
+from tqdm import tqdm
+
+
+
+if __name__=="__main__":
+    # load data
+    df=pd.read_csv("Data/temperatures_melbourne.csv", sep=";", decimal=".")
+
+    # quantiles
+    quantiles = [0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9]
+    # table for output results
+    
+    # train, test split
+    X_train, X_test, y_train, y_test = train_test_split(df["Yt-1"], df["Yt"], test_size=0.20, random_state=0)
+
+    # equally space dataset for plotting
+    eval_set=np.linspace(df["Yt-1"].min(), df["Yt-1"].max(), 100).T
+
+    # reshape data to be predicted
+    X_train= X_train.values.reshape(-1, 1)
+    X_test = X_test.values.reshape(-1, 1)
+    eval_set=eval_set.reshape(-1,1)
+
+
+    # define loss to tune
+    neg_mean_pinball_loss_scorer_05 = make_scorer(
+    mean_pinball_loss,
+    alpha=0.5,
+    greater_is_better=False,  # maximize the negative of the loss
+    )
+    
+    # kernel quantile regression
+    qr_krn_models=[]
+    y_test_pred_qr_krn=[]
+    ktype="periodic"
+
+    # gamma=[1e-1,1e-2,1,5,10,20]
+    # sigma=[1e-1,1e-2,1,5,10,20]
+
+    # polynomial c=[0, 1, 10, 100],
+    # d=[2,3,4,5,8]
+
+    # sigmoid
+    # c=[0, 1, 10, 100],
+    # gamma=[1e-1,1e-2,1,5,10,20]
+
+    # periodic
+    # p=[0,1,10,0.5]
+
+    # matern
[…]

### 2024-05-14 experiments/kqr_kernels_comp.py
clean

@@ -1,4 +1,5 @@
 from kernel_quantile_regression.kqr import KQR
+import math
 import matplotlib.pyplot as plt
 import numpy as np
 import pandas as pd
@@ -52,7 +53,7 @@ if __name__=="__main__":
     # kernel quantile regression
     qr_krn_models=[]
     y_test_pred_qr_krn=[]
-    ktype="periodic"
+    ktype="matern_2.5"
 
     # gamma=[1e-1,1e-2,1,5,10,20]
     # sigma=[1e-1,1e-2,1,5,10,20]
@@ -73,7 +74,7 @@ if __name__=="__main__":
 
     param_grid_krn = dict(
     C=[0.1,1, 5, 10],
-    p=[0,1,10,0.5]
+    gamma=[1/(1e-1),1/(1e-2),1/(1),1/(5),1/(10),1/(20)]
     )
     krn_blueprint=KQR(alpha=0.5, kernel_type=ktype)
     best_hyperparameters_krn=HalvingRandomSearchCV(
@@ -97,9 +98,13 @@ if __name__=="__main__":
     # mae score
     print("mae", f"{mean_absolute_error(y_test, y_test_pred_qr_krn[5]):.6f}")
 
+    pinball=0
     for i,q in enumerate(quantiles):
         print(f"{mean_pinball_loss(y_test,qr_krn_models[i].predict(X_test), alpha=q):.6f}", "&")
+        pinball+=mean_pinball_loss(y_test,qr_krn_models[i].predict(X_test), alpha=q)
 
+print("total pinball loss: ", pinball)
+print("best hyperparameters: ", best_hyperparameters_krn)
 #    Linear qr     Gbm qr      Quantile forest  Kernel qr rbf gaussian   Laplacian   Rbf gaussian x laplacian
 #    11.278895     10.317612   10.370558        10.031708                10.056884   10.150826       
 

### 2024-05-14 experiments/eda/energy_charts.py
eda

@@ -1,0 +1,94 @@
+import datetime
+import matplotlib.pyplot as plt
+
+import pandas as pd
+
+import seaborn as sns
+
+
+country="CH"
+country_name="Switzerland"
+year=2021
+
+df=pd.read_csv(f"/Users/luca/Desktop/kernel_quantile_regression/Data/{country}/{year}/clean/{country.lower()}.csv")
+
+dates=pd.to_datetime(df.Time, format='%Y-%m-%d %H:%M:%S')
+
+
+# load plot
+plt.figure(figsize=(15,5))
+plt.plot(dates, df["Load"], color="black")
+
+plt.xlabel("Time")
+plt.ylabel("Load(MW)")
+# plt.title(f"Energy load {country_name} ({year})")
+plt.savefig(f"/Users/luca/Desktop/ThesisKernelMethods/thesis/images/{country}_historical_load_{year}.png")
+plt.show()
+
+# load vs temperature
+plt.plot(df["Temperature"], df["Load"], "o", alpha=0.4, markersize=3)
+
+plt.ylabel("Load (MW)")
+plt.xlabel("Temperature")
+# plt.title(f"Load versus temperature {country_name} ({year})")
+plt.savefig(f"/Users/luca/Desktop/ThesisKernelMethods/thesis/images/{country}_load_vs_temperature_{year}.png")
+plt.show()
+
+# load vs wind_speed
+plt.plot(df["Wind_speed"], df["Load"], "o", alpha=0.4, markersize=3)
+
+plt.ylabel("Load (MW)")
+plt.xlabel("Wind speed")
+# plt.title(f"Load versus wind speed {country_name} ({year})")
+plt.savefig(f"/Users/luca/Desktop/ThesisKernelMethods/thesis/images/{country}_load_vs_wind_speed_{year}.png")
+plt.show()
+
+
+# box plot load vs day of the week
+sns.set_theme(style='whitegrid')
+
+fig, ax = plt.subplots(figsize=(12, 9))
+xvalues = ["Workday", "Holiday"]
+palette = ['r', 'b']
+
+
+sns.boxplot(data=df, x='Is_holiday', y='Load', palette=palette,
+            width=0.7, dodge=False, ax=ax)
+
+ax.xaxis.set_ticklabels(xvalues)
+
+ax.set_xlabel('Day type')
+ax.set_ylabel("Load(MW)", fontsize=14)
[…]

### 2024-05-14 experiments/eda/gefcom.py

@@ -1,0 +1,64 @@
+import matplotlib.pyplot as plt
+
+import pandas as pd
+
+# script for eda gefcom load and gefcom price
+
+# load
+df=pd.read_csv("/Users/luca/Desktop/GEFCom2014 Data/Load/L-train_clean.csv")
+
+# load starts from 2005
+df=df[df["YEAR"]>=2005][1:]
+df["TIME"]=pd.to_datetime(df[['YEAR', 'MONTH', 'DAY', 'HOUR']])
+
+
+# # plot load
+plt.figure(figsize=(15,5))
+plt.plot(df["TIME"], df["LOAD"], color="black")
+
+plt.xlabel("Time")
+plt.ylabel("Load (MW)")
+
+plt.savefig("/Users/luca/Desktop/ThesisKernelMethods/thesis/images/gefcom_load_historical.png")
+
+plt.show()
+
+
+# # plot w_avg
+plt.figure(figsize=(15,5))
+plt.plot(df["TIME"], df["w_avg"], color="black")
+
+plt.xlabel("Time")
+plt.ylabel("Average weather temperature")
+
+plt.savefig("/Users/luca/Desktop/ThesisKernelMethods/thesis/images/gefcom_w_avg_historical.png")
+
+plt.show()
+
+# # load vs w_avg
+plt.plot(df["w_avg"], df["LOAD"], "o", alpha=0.4, markersize=3)
+plt.xlabel("Average weather temperature")
+plt.ylabel("Load (MW)")
+
+plt.savefig("/Users/luca/Desktop/ThesisKernelMethods/thesis/images/gefcom_load_vs_w_avg.png")
+plt.show()
+
+
+# price
+
+df=pd.read_csv("/Users/luca/Desktop/ThesisKernelMethods/experiments/gefcom_price/Price/Task 1/Task1_P_train.csv")
+
+# # zonal price vs total load
+plt.plot(df["Forecasted Total Load"], df["Zonal Price"], "o", alpha=0.4, markersize=3)
+plt.xlabel("Forecsated total load")
+plt.ylabel("Price $/MW")
+plt.savefig("/Users/luca/Desktop/ThesisKernelMethods/thesis/images/gefcom_zonal_price_vs_total_load.png")
+plt.show()
+
+
+# # zonal price vs zonal load
+plt.plot(df["Forecasted Zonal Load"], df["Zonal Price"], "o", alpha=0.4, markersize=3)
[…]

### 2024-05-14 thesis/appendix.tex
correct names in contents,list of fig and tables
May 13th 2024

@@
 \chapter{Appendix}
-\section{Feature Map Normalization}\label{appendix:new_feature}
+\section{Feature map normalization}\label{appendix:new_feature}
 \begin{proof}
@@
 \begin{table}
-\caption{MAE Boston data}    
+\caption{Mean absolute error Boston data}    
 \begin{tabular}{lllll}
@@
 \begin{table}
-    \caption{MAE Abalone data}
+    \caption{Mean absolute error Abalone data}
     \begin{tabular}{lllll}
@@
 \begin{table}
-    \caption{MAE Vehicle data}
+    \caption{Mean absolute error Vehicle data}
     \begin{tabular}{lllll}
@@
 \newpage
-\section{Src code}\label{src_code}
+\section{Source code}\label{src_code}
 The whole code for the project is hosted on
@@
     \item kqr.py: file implementing our custom kernel quantile regression
+    \item thesis: folder containing tex files
+    \item experiments: folder containing train, test script for each of the experiments carried out in the thesis alongside with the binary files of the pickled models.
 \end{itemize}

### 2024-05-14 thesis/input_files/data_exploratory.tex
May 13th 2024

@@
-- Explain how data has been retrieved.
+% - Explain how data has been retrieved.
 
-- Data provider EPEX(entsoe retrieves its data)
+% - Data provider Energycharts(entsoe retrieves its data)
 
-- Explain the ETL(Extract Transform Load) pipeline
-I set up.
+% - Explain the ETL(Extract Transform Load) pipeline
+% I set up.
 
-- Explore data in order to get useful insights for
-how to tune the models to get the most of them.
+% - Explore data in order to get useful insights for
+% how to tune the models to get the most of them.
 
-- correlation between temperatures and load
+% - correlation between temperatures and load, wind speeds and load
 
-- Correlation and auto correlation plots
+% - Split train and test dataset
+% Explain carefully why it is important to carry
+% out an out of sample test and not in sample.
+% In sample test involves look ahead bias because we are
+% fitting the model on the data we want to predict,
+% thus it overfits on the data considered but it does
+% not generalize well.
 
-- Split train and test dataset
-Explain carefully why it is important to carry
-out an out of sample test and not in sample.
-In sample test involves look ahead bias because we are
-fitting the model on the data we want to predict,
-thus it overfits on the data considered but it does
-not generalize well.
-
-This section covers the datasets used in our experiments, attributes and features will be thoroughly described.
+This section covers the datasets used in our experiments, in addition, attributes and features will be thoroughly described.
 Next, we will explain the ETL pipeline that we set up in order to ease the workflow of our comparison studies.
@@
 
-\section{Dataset}
-The chosen datasets for probabilistic load forecasting, were the data from the GEFCom 2014 competition. The data is freely hosted on Dr. Hong blog \cite{hong2016probabilistic}.
-The main reason was that, these dataset are considered an a excellent test case for comparing predictive models between the EF community.
+\section{GEFCom2014}
+The first class of datasets we identified were the ones from the GEFCom 2014 competition. The data is freely hosted on Dr. Hong blog \cite{hong2016probabilistic}.
+The main reason is that, these dataset are considered an a excellent test case for comparing predictive models between the EF community.
 Additionally, the scores of the competing models are freely available, this enables us to carry out a clean and transparent comparative study.
@@
 \\
-\subsection{Price track}
+\subsection{GEFCom2014 Price track}
 In the price forecasting track, the goal was to predict the electricity price for the next 24 hours of a single zone. 
@@
 \end{figure}
+\subsubsection{EDA}
+In exploring this data we started off by plotting the time series of the zonal prices alongside with the toal load and zonal load, see figure \ref{fig:price_track_fig1}.
+After that, we plotted the scatter plot for inspecting the correlation between the predictors and the dependent variable, figure \ref{fig:gefcom_zonal_price_vs_zonal_load} and figure \ref{fig:gefcom_zonal_price_vs_total_load}.
+From these visualisation, we can see their goodness for predicting zonal price since there exist a clear correlation between the candidate predictors and the target variable.
+
+\begin{figure}[!h]
+    \includegraphics[width=\textwidth]{images/gefcom_zonal_price_vs_zonal_load.png}
+    \caption{Zonal price versus zonal load}
+    \label{fig:gefcom_zonal_price_vs_zonal_load}
+\end{figure}
+
+\begin{figure}[!h]
+    \includegraphics[width=\textwidth]{images/gefcom_zonal_price_vs_total_load.png}
+    \caption{Zonal price versus total load}
+    \label{fig:gefcom_zonal_price_vs_total_load}
+\end{figure}
+
 \subsection{Load track}
@@
 The data for load forecasting is contained in the GEFCom2014-L\_V2 zip file.
-Within this subfolder, we a txt with the competition instructions and 15 folders one for each of the consecutive tasks.
+Within this subfolder, we have a txt with the competition instructions and 15 subfolders one for each of the consecutive tasks.
 For each of those we have the prediction of the competition benchmark model and the train file upon which we will fit our model.
@@
 Notice, the pinball scores for the load track are stored inside the subtab L-score-0/L-score-2 of Provisional\_Leaderboard\_V2.
+\subsubsection{EDA}
+We applied the same kind of EDA to the GEFCom load data.
+Figure \ref{fig:gefcom_load_historical} depicts the hisorical load while figure \ref{fig:gefcom_w_avg_historical} reports the historical weather temperature.
+Finally figure \ref{fig:gefcom_load_vs_w_avg} shows the scatter plot between the the load and the weather.
+These plots motivate the selection of weather temperature as predictor for the load.
 
-\section{ETL}
+\begin{figure}[!h]
+    \includegraphics[width=\textwidth]{images/gefcom_load_historical.png}
+    \caption{Historical load}
+    \label{fig:gefcom_load_historical}
+\end{figure}
 
-\section{EDA}
+\begin{figure}[!h]
+    \includegraphics[width=\textwidth]{images/gefcom_w_avg_historical.png}
+    \caption{Historical weather}
+    \label{fig:gefcom_w_avg_historical}
+\end{figure}
+
+\begin{figure}[!h]
+    \includegraphics[width=\textwidth]{images/gefcom_load_vs_w_avg.png}
+    \caption{Load versus weather}
+    \label{fig:gefcom_load_vs_w_avg}
+\end{figure}
+
+\subsection{GEFCom ETL}
+The provided data is raw, hence the need for a standardized extract transform and load pipeline. 
+Firstly, we drop all Nans and zeros.
+Notice, the provided data for the challenge has a timestamp column not coherent to any time format.
+Thus, we first pass the data through a method extracting time information: hour, day, month and year.
+This info is then combined into a datetime object, we selected the ISO format.
+Next, since the data does not provide the location for the target utility, the best we can do is selecting the average of weather temperatures as predictor variables.
+This independent variable is named w\_avg.
+The same thing applies in the cleaning of the test data with the addition of an if clause for handling the different naming convention and formatting used in the last track of the challenge.
+The python etl pipeline has been then automated to loop every task folder through bash scripting.
+
+\section{Germany Switzerland Austria data}
+Next, we identified various data source for Germany, Switzerland and Austria in order to further
+investigate the techniques studied in this thesis work.
+Taking inspiration from the GEFCom2014 format, we firstly created a load track for Germany and Switzerland. 
+The source of data has been the \href{https://www.energy-charts.info/index.html?l=en&c=DE}{energy charts} website.
+It provides data for data for electricity power, electricity potential capacity, energy generation, market prices, environmental measurements and also scenario simulations.
+Notice, this data is retrieved from the European network of transmission system operators for electricity \href{https://www.entsoe.eu/}{ENTSOE} and made available thanks to the \href{https://ise.fraunhofer.de/en.html}{Fraunhofer Institute for Solar Energy Systems ISE}.
+\\
+In this experiment, the goal is comparing kernel quantile regression to the other state of the art quantile regressors, see section \ref{ch:prob}.
+The target dependent variable is the national load while as predictors we took
+the weather temperature, the wind speed, the date expressed in terms of the three attributes day, month and hour, a categorical variable for the day of the week and a binary categorical variable accounting for the holiday effect.
+
+In the second experiment we compared various kernels in the context of load foreacsting.
+For this study we considered the data of \href{https://zenodo.org/records/7907883}{SECURES-Met}.
+This data consists of various metereological attributes relevant for modelling electricity demand.
+The dataset is made up of historical data until end of 2020, while from 2021 onwards the data consists of forecasts made by a metereological model developed by the European Centre for Medium-Range Weather Forecasts (ECMWF).
+Notice, the forecasted data is splitted in two files according to the scenarios RCP4.5 and RCP8.5.
+In our experiment we considered the former.
+In this setting the independent variables available are 
+direct irradiation, global radiation, power from reservoir plants, power from run of the river plants (ROR), weather temperature and wind potential.
+
+\subsubsection{ETL}
+From energy charts we downloaded the historical load along with the weather temperatures and wind speeds forecasts.
+Notice, the data request to the environmental variables at hourly frequency for a whole year is broken.
+Therefore, we downloaded each month separately and then transformed into a single csv for the whole year.
+
+For what concers the data from SECURES-Met we first identified the data interesting to us.
+Next, we formatted all the csv to a coherent format, gave coloumns meaningful names, sliced for the years 2020 (train) and 2021 (test) and finally we merged all single csv together.
+
+All the manipulations involved in the ETL pipeline have been carried out by combining python and bash scripting.
+
+\subsubsection{EDA}
+Following are reported the EDA plots alongside with insights that can be extracted from them.
+We choose to focus the following analysis on the swiss data only given that the plots for the german data are pretty similar.
+To get started we visualized the historical load data in figure \ref{fig:CH_historical_load_2021}.
+As expected, electricity load exhibits a seasonal patter with highs in the winter and lows during the summer season.
+\begin{figure}[!h]
+    \includegraphics[width=\textwidth]{images/CH_historical_load_2021.png}
+    \caption{Switzerland historical load}
+    \label{fig:CH_historical_load_2021}
+\end{figure}
+Next we built scatter plots of the independent continous variables against the load as means of visualizing correlation between them.
+Figure \ref{fig:CH_load_vs_temperature_2021} plots load against temperatures while figure \ref{fig:CH_load_vs_wind_speed_2021} depicts load against wind speed.
+From these plots we can clearly see there exist a relationship that ties wind speeds and temperatures to load.
+
+\begin{figure}[!h]
+    \includegraphics[width=\textwidth]{images/CH_load_vs_temperature_2021.png}
+    \caption{Load vs temperature Switzerland 2021}
+    \label{fig:CH_load_vs_temperature_2021}
+\end{figure}
+
+\begin{figure}[!h]
+    \includegraphics[width=\textwidth]{images/CH_load_vs_wind_speed_2021.png}
+    \caption{Load vs wind speed Switzerland 2021}
+    \label{fig:CH_load_vs_wind_speed_2021}
+\end{figure}
+
+Following, we used box plots in order to visualize the relations between categorical independent variables and the dependent one.
+Such visualisation are reported in figure \ref{fig:CH_day_of_week_boxplot_2021} and figure \ref{fig:CH_is_holiday_boxplot_2021}.
+We can see that weekdays and workdays involve a higher load and posses a wider range.
+
+\begin{figure}[!h]
+    \includegraphics[width=\textwidth]{images/CH_day_of_week_boxplot_2021.png}
+    \caption{Load versus day of week boxplot}
+    \label{fig:CH_day_of_week_boxplot_2021}
+\end{figure}
+
+\begin{figure}[!h]
+    \includegraphics[width=\textwidth]{images/CH_is_holiday_boxplot_2021.png}
+    \caption{Load versus day type boxplot}
+    \label{fig:CH_is_holiday_boxplot_2021}
+\end{figure}
+
+We conclude this EDA by reporting the heatmap of independent variables alongside with the load in figure \ref{fig:CH_heatmap_2021}.
+Temperature, wind speed and day of week show a meaningful correlation.
+This is a clear indication of their suitability as predictors for forecasting the load.
+
+\begin{figure}[!h]
+    \includegraphics[width=\textwidth]{images/CH_heatmap_2021.png}
+    \caption{Switzerland load heatmap (2021)}
+    \label{fig:CH_heatmap_2021}
+\end{figure}

### 2024-05-14 thesis/input_files/implementation.tex
May 13th 2024

@@
 This estimator is compatible with useful scikit-learn in built methods like gridsearch, crossvalidation and scoring rules.
+Moreover, this code is compatible with sklearn in built kernel functions \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.metrics.pairwise}{sklearn.metrics.pairwise} and \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.gaussian_process.kernels}{sklearn.gaussian\_process.kernels}, functions supported by our kernel quantile regression include: gaussian rbf, laplacian, mater 1/2, matern 3/2, matern 5/2, linear, cosine, sigmoid, periodic, polynomial and custom composition of kernels.
+% NOTICE, these kernel implementations differ in the form of passed parameters in the former we pass in the numerator gamma while in the latter we pass l, which stands for length scale in the denominator.
+To cross validate overt the same range of parameter across the different kernels we made sura that gamma=1/l.
 \\
@@
 % (i guess it is in C or C++ and then binded to R).
-\subsubsection{Python versus R implementation}
-In this section, a comparison study has been carried out in order to inspect the competitiveness of our implementation with the existing one for the R programming language.
+% \subsubsection{Python versus R implementation}
+% In this section, a comparison study has been carried out in order to inspect the competitiveness of our implementation with the existing one for the R programming language.

### 2024-05-14 thesis/input_files/paper1.tex
May 13th 2024

@@
 % Some commands used in this file
+The following section covers the theory of kernel methods and introduces the building blocks of kernel theory.
+% \section{Kernel mean embedding of distributions: A review and beyond}
 
-\section{Kernel mean embedding of distributions: A review and beyond}
 % \href{https://arxiv.org/abs/1605.09522}{
-From this first paper \cite{Muandet_2017}, the notation and terms used in the theory of Reproducing Kernel Hilbert Spaces are summarized.
-\\
+% From this first paper \cite{Muandet_2017}, the notation and terms used in the theory of Reproducing Kernel Hilbert Spaces are summarized.
+% \\
+
 Many algorithms use the inner product as similarity measure between data instances $x, x' \in \mathcal{X}$. However, this inner product spans only the class of linear similarity measures. 
@@
 Notice that this is the same of $\langle \phi(x), \phi(x') \rangle$; thus the kernel trick consists of just using $k(x,x')=:(x^Tx')^2$.
+For example
+\\
+$\langle \phi([2,4]),\phi([8,9]) \rangle=  \left[ {\begin{array}{cccc}
+    4 & 16 &     \sqrt{2}\cdot 2 \cdot 4\\
+  \end{array} } \right]
+  \left[ {\begin{array}{cccc}
+    64\\
+    81\\
+    \sqrt{2}\cdot8 \cdot 9
+    \\
+  \end{array} } \right]\\=4\cdot 64 +16\cdot 81 +2 \cdot 2 \cdot 4 \cdot 8 \cdot 9=2704
+$
+\\
+\\
+While $k([2,4], [8,9])=(16+36)^2=2704$
 
-\subsection{Reproducing kernel hilbert space}
+This gives a taste of how powerful kernels are thanks to the kernel trick.
+Indeed, the idea of the the kernel trick can be extended to feature maps
+$\phi$ involving an infinite feature space. In these cases, calculating the
+respective kernel is equal to calculating the inner product between an infinite
+number of features of the data points.
+
+\section{Theory of kernels}
 Subsequently we introduce the definitions that make up the basis for the theory of kernel methods.
@@
 \begin{definition}
+    A function $k: \mathcal{X} \times \mathcal{X} \to \mathbb{R}$ is said to be finitely positive semi definite if it is a symmetric function for which the matrices formed with it are positive semi definite.
+\end{definition}
+
+\begin{theorem}[Characterisation of kernels]    
+    A function 
+    \\
+    $k:\mathcal{X} \times \mathcal{X} \to \mathbb{R}$
+    \\
+    which is either continous or has finite domain can be decomposed as
+    \\
+    $k(x,y)=\langle \phi(x), \phi(y) \rangle_{\mathcal{H}}$
+    \\
+    that is the inner product in the Hilbert space $\mathcal{H}$ in which the features of the arguments $\phi(x)$ lie if and only if $k$ satisfy the positive semi definite property.
+\end{theorem}
+
+Next, we make this characterisation explicit by showing how to construct a feature mapping $\phi$ of a kernel $k$ satisfying the finitely positive semi definite property.
+Let the feature space be the set of functions 
+\\
+$\mathcal{F} =\{ \sum\limits_{i=1}^n a_i k(x_i, \cdot): n \in \mathbb{N}, x_i \in \mathcal{X}, a_i \in \mathbb{R}, i=1,\cdots, n\}$
+This space is clearly closed under addition of functions and scalar multiplication, therefore it is a vector space.
+Letting $f(x)=\sum\limits_{i=1}^n a_i k_i(x_i,x)$ and $g(x)=\sum\limits_{i=1}^m \beta_i k_i(z_i,x)$,  we can introduce an inner product on $\mathcal{F}$ as 
+\\
+$\langle f,g \rangle_{\mathcal{F}}=:\sum\limits_{i=1}^n\sum\limits_{j=1}^m a_i \beta_j k_i(x_i,z_j)=\sum\limits_{i=1}^n a_i g(x_i)= \sum\limits_{j=1}^m \beta_j f(z_j)$
+\\
+Notice, $\langle f, f \rangle= a^\intercal K a \geq 0 \forall f \in \mathcal{F}$, where $K$ is the kernel matrix.
+Moreover, $\langle f, g \rangle$ is real valued, symmetric and bilinear. All these facts implies that the just introduced inner product is a valid metric.
+\\
+Additionally, letting $g(x)=k(\cdot, x)$, we have that $\langle f, k(\cdot, x)\rangle=\sum\limits_{i=1}^n a_i k(x_i,x)=f(x)$. This is known as the reproducing property.
+\\
+Finally, to show the characterisation of kernels we have to check that the inner product makes $\mathcal{F}$ a complete metric space.
+Consider a fixed input $x$ and a Cauchy sequence $(f_n)_{n=1}^{\infty}$ we then have by means of the Cauchy-Schwarz inequality that 
+\\
+$(f_n(x)- f_m(x))^2=\langle f_n-f_m, k(\cdot, x)\rangle^2 \leq \| f_n - f_m \|^2 k(x,x)$
+\\
+We have that $f_n(x)$ is a Cauchy sequence of real numbers.
+Thus, we can let $g(x)=\lim_{n\to \infty} f_n(x)$ and include all such limits functions in $\mathcal{F}$, doing so we obtain the Hilbert space $\mathcal{H}$ associated to the kernel $k$.
+Having constructed the feature space $\mathcal{H}$, the mapping $\phi(x)$ is defined as follows
+\\
+$\phi: x \in \mathcal{X} \to \phi(x)=k(\cdot, x) \in \mathcal{H}$
+\\
+Notice, that by the reproducing property, $f$ can be represented as the inner product with itself in the feature space, that is $f(x)=\langle f, \phi(x)\rangle=\langle f, k(\cdot, x)\rangle=f(x)$
+
+\begin{definition}
     Let $(\mathcal{H}, \langle \cdot, \cdot \rangle_\mathcal{H})$ be a Hilbert space of real-valued functions on $\mathcal{X}$.
@@
     \end{align}
+    Notice, the finitely positive semi definite property is a sufficient condition for $\mathcal{H}$ to be an RKHS
 \end{definition}
+
+The analysis of above applies to any kernel, that is given a kernel function we can always construct the RKHS associated to it.
+\begin{theorem}
+    If a symmetric function $k(\cdot, \cdot)$ satisfies the reproducing property in a Hilbert space $\mathcal{H}$, then $k$ satisfies the finitely positive semi definite property.
+\end{theorem}
+
+\begin{proof}
+    \begin{align}
+        \sum\limits_{i,j=1}^n a_i a_j k(x_i, x_j)=& \sum\limits_{i,j=1}^n a_i a_j \langle k(\cdot, x_i), k(\cdot, x_j)\rangle_{\mathcal{H}}
+        \\
+        =& \langle \sum\limits_{i=1}^n a_i k(\cdot, x_i), \sum\limits_{j=1}^n a_j k(\cdot, x_j) \rangle \\
+        =& \| \sum\limits_{i=1}^n a_i k(\cdot, x_i)\|_{\mathcal{H}}^2 \geq 0
+    \end{align}
+\end{proof}
 
@@
 
+\begin{proof}
+    It is known that null space of $L$, call it $M$, is a closed linear subspace in $\mathcal{H}$.
+This implies that the Hilbert space direct sum $M \bigoplus M^\bot$ is an isomorphism of $\mathcal{H}$.
+Letting $L$ be different that the zero operator we have that $M^\bot \neq {0}$.
+    \\
+    Choose now a $z \in M^\bot$ with norm 1.
+By linearity of $L$, for any $h \in \mathcal{H}$ we have $L(zLh- hLz)=LzLh- LhLz=0$.
+This means that $zLh- hLz$ belongs to the null space of $L$, that is  $zLh- hLz \in M$.
+    \\
+    Consider now 
+    \begin{align}
+        Lh=& Lh \cdot 1
+        \\
+        =& Lh \langle z,z\rangle
+        \\
+        =& \langle zLh, z\rangle \\
+        =& \langle zLh -hLz+hLz, z\rangle
+        \\
+        =& \langle zLh -hLz, z\rangle +\langle hLz, z\rangle \\
+        =& \langle hLz, z\rangle \\
+        =& \langle h, z\overline{(Lz)}\rangle \\
+    \end{align}
+    Therefore, $Lh=\langle h, h_0\rangle$  with  $h_0=z\overline{(Lz)}$
+    \\
+    Finally, notice that $h_0$ is unique.
+    Assuming $h_0$ and $h_1$ to both satisfy the above equation for all $h \in \mathcal{H}$ we have that
+    \begin{align}
+        \langle h_0,h \rangle=&\langle h_1, h \rangle \\
+        \langle h_0,h \rangle-\langle h_1, h \rangle=& 0 \\
+        \langle h_0 -h_1,h \rangle=& 0
+    \end{align}
+    Setting $h=h_0 -h_1$, we have the uniqueness of $h_0$.
+    % because inner product is zero, inner product has to satisfy positive definitness property to be such >=0, if it is zero it means that h_0=h_1
+
+\end{proof}
+%conjugate simmetry is because in complex analysis when you swap in the inner product, you have to take the conjugate
+
 The Riesz representation theorem results in the following proposition for reproducing kernel hilbert spaces.
@@
 \begin{align*}
-    k_{x}(y)=\mathcal{F}_{y}(k_{x})=\langle k_{x}, k_{y}\rangle_{\mathcal{H}=\langle \phi(x), \phi(y)\rangle_{\mathcal{H}}}
+    k_{x}(y)=\mathcal{F}_{y}(k_{x})=\langle k_{x}, k_{y}\rangle_{\mathcal{H}}=\langle \phi(x), \phi(y)\rangle_{\mathcal{H}}
 \end{align*}
 
-\subsection{Kernel families}
+\begin{definition}
+    Considering a set of vectors $x_1, \cdots, x_n$, the Gram matrix is defined as the $n\times n$ matrix whose entries are $\langle x_i, x_j \rangle$.    
+\end{definition}
+
+Notice, it follows from the definition that Gram matrix is symmetric.
+In the kernel setting we evaluate the inner products in the feature space generated by the feature map $\phi$, therefore the Gram matrix will have entris $\langle \phi(x_i), \phi(x_j) \rangle_{\mathcal{H}}= k(x_i, x_j)$.
+Such matrix is also referred to the kernel matrix $K$.
+
+\[
+K_{n\times n}=
+\begin{bmatrix}
+    k(x_1,x_1)       &  \dots & k(x_1,x_n) \\
+    \vdots       & \ddots & \vdots \\
+    k(x_n, x_1)       & \dots & k(x_n, x_n)
+\end{bmatrix}
+\]
+All kernel methods take as input the kernel matrix, it is the central data type for all 
+kernel methods.
+
+\begin{proposition}
+    In addition to being symmetric, $K$ possesses also the property of being positive semi definite.    
+\end{proposition}
+
+\begin{proof}
+    \begin{align}
+        z^\intercal K z =& \sum_{i=1,j=1}^n z_i z_j K_{i,j}
+        \\
+        =& \sum_{i=1,j=1}^n z_i z_j \langle \phi(x_i), \phi(x_j) \rangle_{\mathcal{H}}
+        \\
+        =& \langle \sum_{i=1}^n z_i \phi(x_i), \sum_{j=1}^n z_j \phi(x_j)\rangle_{\mathcal{H}}
+        \\
+        =& \|\sum_{i=1}^n z_i \phi(x_i) \|_{\mathcal{H}}^2 \geq 0
+    \end{align}
+\end{proof}
+We know that positive semidefinite matrices form a cone in the vector space $n\times n$.
+Because of this we have that such matrices are convex, this implies that all the nice theory of convex analysis can be used when optimising over them.
+
+Kernel functions are characterized by a couple of properties that make them a closure. Thus, there exists ways of combining known kernels in order to obtain new kernels.
+
+\begin{proposition}
+    Let $k_1$ and $k_2$ be kernels over $\mathcal{X} \times \mathcal{X}, \mathcal{X} \subset \mathbb{R}^n$, $a \in \mathbb{R}^+ and \phi: \mathcal{X} \to \mathbb{R}^N$ with $k_3$ a kernel over $\mathbb{R}^N \times \mathbb{R}^N$.
+Then the following functions are kernels:
+    \begin{align}
+        k(x,z)=&k_1(x,z)+k_2(x,Y)
+        \\
+        k(x,z)=& ak_1(x,z)
+        \\
+        k(x,z)=& k_1(x,z)k_2(x,z)
+        \\
+        k(x,z)=& k_3(\phi(x),\phi(z))
+    \end{align}.
+\end{proposition}
+
+\begin{proof}
+    Consider a finite set of points $x_1, \cdots, x_n$, a vector $z \in \mathbb{R}^n$ and let $K_1$ and $K_2$ be the kernel matrices associated to $k_1$ and $k_2$ evaluted on these points.
+    \\
+    We have:
+    \\
+    \begin{align}
+    z^\intercal (K_1+K_2) z= z^\intercal K_1 z + z^\intercal K_1 z \geq 0
+    \\
+    z^\intercal aK z= a z^\intercal K z \geq 0
+    \end{align}
+    Let $K=K_1 \otimes K_2$, we have that the eigenvalues of $K$ are made up by the product pairs of the eigenvalues of $K_1$ and $K_2$.
+Thus, $K$ is positive semidefinite.
+Next, matrix corresponding to the kernel $k_1 \cdot k_2$ corresponds to the Schur product $H$ of $K_1$ and $K_2$.
+Notice, $H$ is a principal submatrix of $K$, hence for any $z in \mathbb{R}^n$ there exist a $t \in \mathbb{R}^{n^2}$ such that $z^\intercal H z= t^\intercal K t$.
+Using the fact that K is positive semidefinite we have that for any $z$, it holds
+    \begin{align}
+        z^\intercal H z= t^\intercal K t\geq 0
+    \end{align}
+\end{proof}
+
+\section{Kernel families}
+The selection of kernel determines the class of functions that will be searcehd by the learning algorithm.
+Prior knowledge of the problem domain can help restricting the candidate families of kernel functions.
 Table \ref{tab:kernel types} contains popular kernel families in literature and applications.
@@
     \caption{Kernel types}
+    \resizebox{\textwidth}{!}{
     \begin{tabular}{lll}
@@
        Polynomial &  $k(x_1,x_2)=(x_1^\intercal x_2+c)^d$ &   c, d\\
-       Gaussian RBF &  $k(x_1,x_2)=e^{-\frac{\|x_1-x_2|\|^2}{2\sigma^2}}$ &   $\sigma$\\
-       Exponential RBF/Laplacian& $k(x_1,x_2)=e^{-\frac{|x_1-x_2|}{\gamma}}$ &   $\gamma$\\
+       Gaussian RBF &  $k(x_1,x_2)=e^{-\frac{\|x_1-x_2|\|_{2}^2}{2\sigma^2}}$ &   $\sigma$\\
+       Exponential RBF/Laplacian& $k(x_1,x_2)=e^{-\frac{\|x_1-x_2\|_{1}}{\gamma}}$ &   $\gamma$\\
        Hyperbolic/Sigmoid Kernel &  $k(x_1,x_2)=\tanh(\gamma x_1^\intercal x_2+r)$ &  $\gamma$, r \\
-       Periodic &  $k(x_1,x_2)=e^{\frac{-2 sin^2\left(\frac{\pi}{p}|x_1-x_2| \right)}{l^2}} $ &   p, l\\
+       Periodic &  $k(x_1,x_2)=e^{\frac{-2 sin^2\left(\frac{\pi}{p}\|x_1-x_2\|_{2} \right)}{l^2}} $ &   p, l\\
        Chi-squared kernel &  $k(x_1,x_2)=e^{-\gamma \sum\limits_i \frac{(x_i - y_i)^2}{x_i+y_i}} $ &   $\gamma$\\
-        Cosine  &  $k(x_1,x_2)=\frac{xy^\intercal}{\|x\|_{L_2}\|y\|_{L_2}}$ &   \\
-       Matern  &  $k(x_1, x_2) =  \frac{1}{\gamma(\nu)2^{\nu-1}}\Bigg(
+        Cosine  &  $k(x_1,x_2)=\frac{xy^\intercal}{\|x\|_{2}\|y\|_{2}}$ &   \\
+       Matern  &  $k(x_1, x_2) =  \frac{1}{\Gamma(\nu)2^{\nu-1}}\Bigg(
         \frac{\sqrt{2\nu}}{l} d(x_1 , x_2 )
         \Bigg)^\nu K_\nu\Bigg(
-        \frac{\sqrt{2\nu}}{l} d(x_1 , x_2 )\Bigg)$ &  $\gamma, \nu$ \\
+        \frac{\sqrt{2\nu}}{l} d(x_1 , x_2 )\Bigg)$ &  $l, \nu$ \\
 
         \bottomrule
-    \end{tabular}
+    \end{tabular}}
     \label{tab:kernel types}
 \end{table}
+
+When $nu=p+\frac{1}{2}$, $p \in \mathbb{N}^+$, the Matern kernel can be rewritten as a product of an exponential and a polynomial of degree $p$. 
+$k(x_1,x_2)=\exp \left(-{\frac {{\sqrt {2p+1}}d(x_1,x_2)}{l }}\right){\frac {p!}{(2p)!}}\sum _{i=0}^{p}{\frac {(p+i)!}{i!(p-i)!}}\left({\frac {2{\sqrt {2p+1}}d(x_1,x_2)}{l }}\right)^{p-i}$
+For example:
+\begin{itemize}
+    \item $p=0,nu=\frac{1}{2} \implies k(x_1,x_2)=\exp\left(\frac{-d(x_1,x_2)}{l}\right)$
+    \item $p=1,\nu=\frac{3}{2} \implies 
+    k(x_1,x_2)=\left(1+{\frac {{\sqrt {3}}d(x_1,x_2)}{l }}\right)\exp \left(-{\frac {{\sqrt {3}}d(x_1,x_2)}{l }}\right)$
+    \item $p=2,\nu=\frac{5}{2} \implies k(x_1,x_2)=\left(1+{\frac {{\sqrt {5}}d}{l }}+{\frac {5d(x_1,x_2)^{2}}{3 l ^{2}}}\right)\exp \left(-{\frac {{\sqrt {5}}d(x_1, x_2)}{l }}\right)$
+\end{itemize}
+
+Furthermore, as $\nu \to \infty$ we have that the Matern kernel converges to the Gaussian RBF kernel.
+$k(x_1,x_2)=\exp\left(\frac{-d^2(x_1,x_2)}{2l^2}\right)$
+
+% CONDITIONS upon matern is equivalent to rbf
+% KERNEL THEORY FROM BOOKS AND KERNEL COOKBOOK
+
+Any algorithm involving inner products between inputs can be combined with a kernel methods.
+The chosen kernel function will calculate the inner product between the images of two inputs in the feature space.
+This makes the orginal algorithm operate in a high dimesional space.
+Kernel methods show good modularity because they work the same for any kernel and any data type.
+This concept is illustrated in figure \ref{fig:workflow_kernels}, where an abstract kernel methods workflow is visualized.
+First we build the kernel matrix by using a kernel function to process the input data.
+Next, the kernel matrix is passed to the pattern analysis algorithm, which returns the learned function.
+
+\begin{figure}[!ht]
+    \includegraphics[width=\textwidth]{images/workflow_kernels.png}
+    \caption{Workflow for the application of kernel methods \cite{shawe2004kernel}}
+    \label{fig:workflow_kernels}
+\end{figure}

### 2024-05-14 thesis/input_files/probabilistic.tex
May 13th 2024

@@
 Hereafter, the results of the four presented methods on the Melbourne dataset are reported, see figure \ref{fig:melbourne_quantiles_comparison} for a visualisation.
+Notice, the kernel considered here is the gaussian rbf.
 Hyperparameter tuning have been optimised through cross validation, see \ref{appendix:cross_validation}.
@@
     \midrule
-    0.100000 & 0.710644 & 0.549232 & 0.562888 & 0.540235 \\
-    0.200000 & 1.155014 & 0.938561 & 0.946712 & 0.903213 \\
-    0.300000 & 1.417805 & 1.212671 & 1.222407 & 1.173803 \\
-    0.400000 & 1.540108 & 1.399925 & 1.409293 & 1.367943 \\
-    0.500000 & 1.574957 & 1.517281 & 1.484589 & 1.455849 \\
-    0.600000 & 1.525114 & 1.498608 & 1.495474 & 1.447470 \\
-    0.700000 & 1.397918 & 1.372183 & 1.362173 & 1.331503 \\
-    0.800000 & 1.170140 & 1.115077 & 1.123195 & 1.096606 \\
-    0.900000 & 0.787195 & 0.714075 & 0.734112 & 0.715085 \\
+    0.100000 & 0.710644 & 0.549232 & 0.562888 & \textbf{0.540235} \\
+    0.200000 & 1.155014 & 0.938561 & 0.946712 & \textbf{0.903213} \\
+    0.300000 & 1.417805 & 1.212671 & 1.222407 & \textbf{1.173803} \\
+    0.400000 & 1.540108 & 1.399925 & 1.409293 & \textbf{1.367943} \\
+    0.500000 & 1.574957 & 1.517281 & 1.484589 & \textbf{1.455849} \\
+    0.600000 & 1.525114 & 1.498608 & 1.495474 & 1.\textbf{447470} \\
+    0.700000 & 1.397918 & 1.372183 & 1.362173 & \textbf{1.331503} \\
+    0.800000 & 1.170140 & 1.115077 & 1.123195 & \textbf{1.096606} \\
+    0.900000 & 0.787195 & \textbf{0.714075} & 0.734112 & 0.715085 \\
     \bottomrule
@@
     \midrule
-    MAE & 3.253882 & 3.134805 & 3.095041 & 3.024336 \\
+    MAE & 3.253882 & 3.134805 & 3.095041 & \textbf{3.024336} \\
     \bottomrule
@@
         & Sigmoid & 16.253973            \\
-        & Chi squared & 10.023732       \\
-        & Matern & \textbf{10.021369}  \\
+        & Chi squared & \textbf{10.023732}       \\
+        & Matern 0.5 & 10.056883  \\
+        & Matern 1.5 & 10.040288  \\
+        & Matern 2.5 & 10.024495    \\
         & Periodic  & 15.946272\\
@@
     Gauss RBF &  \textbf{0.540145} &
-    \textbf{0.903193} &
+    0.903193 &
     1.173627 &
@@
     1.447470 &
-    \textbf{1.331375} &
+    1.331375 &
     1.096649 &
@@
     1.340193 &
-    \textbf{1.095991} &
+    1.095991 &
     0.707947 \\
@@
     1.178357 &
-    \textbf{1.362351} &
+    1.362351 &
     1.450249 &
@@
     1.173164 &
-    1.358862 &
+    \textbf{1.358862} &
     \textbf{1.444130} &
@@
     0.726625     \\
-    Matern &  0.541192 &
-    0.903209 &
-    \textbf{1.169543} &
-    1.364430 &
-    1.451248 &
-    1.442511 &
-    1.332522 &
-    1.096750 &
-    0.716226 \\
+    Matern 0.5 &  0.542499 &
+    0.906517 &
+    1.177309 &
+    1.371198 &
+    1.462111 &
+    1.453119 &
+    1.340193 &
+    1.095991 &
+    0.707947 \\
+    Matern 1.5 &  0.541410 &
+    0.902901 &
+    1.173943 &
+    1.371218 &
+    1.462212 &
+    1.449296 &
+    1.333655 &
+    \textbf{1.095092} &
+    0.710560 \\
+    Matern 2.5 &  0.541233 &
+    \textbf{0.902231} &
+    \textbf{1.172001} &
+    1.366449 &
+    1.456505 &
+    1.444124 &
+    \textbf{1.331251} &
+    1.095674 &
+    0.715029 \\
     Periodic  & 0.759452 &
@@
         & Sigmoid &      4.849589       \\
-        & Chi squared &    \textbf{3.022778}    \\
-        & Matern &  3.024517 \\
+        & Chi squared &    3.022778    \\
+        & Matern 0.5 &  3.038633 \\
+        & Matern 1.5 &  3.023060 \\
+        & Matern 2.5 &  \textbf{3.021949} \\
         & Periodic  & 4.880673 \\
@@
     \end{subfigure}
+
     \begin{subfigure}[b]{0.5\linewidth}
         \centering
-        \includegraphics[width=1.1\textwidth]{images/melborune_matern_kernel_quantile_regression.png}
-        \caption{Matern} 
+        \includegraphics[width=1.1\textwidth]{images/melborune_matern_0.5_kernel_quantile_regression.png}
+        \caption{Matern 0.5} 
+        \label{} 
+        \vspace{4ex}
+    \end{subfigure} 
+    \begin{subfigure}[b]{0.5\linewidth}
+        \centering
+        \includegraphics[width=1.1\textwidth]{images/melborune_matern_1.5_kernel_quantile_regression.png}
+        \caption{Matern 1.5} 
+        \label{} 
+        \vspace{4ex}
+    \end{subfigure}
+
+    \begin{subfigure}[b]{0.5\linewidth}
+        \centering
+        \includegraphics[width=1.1\textwidth]{images/melborune_matern_2.5_kernel_quantile_regression.png}
+        \caption{Matern 2.5} 
         \label{} 

### 2024-05-14 thesis/input_files/symbols.tex
April 6th 2024
May 13th 2024

@@
 %E
-\glsxtrnewsymbol[description={Exploratory data analysis}]{EDA}{EDA}
+\glsxtrnewsymbol
+[description={European Centre for Medium-Range Weather Forecasts}]{ECMWF}{ECMWF}
+[description={Exploratory data analysis}]{EDA}{EDA}
 \glsxtrnewsymbol[description={Electricity forecasting}]{EF}{EF}
@@
 \glsxtrnewsymbol[description={Empirical mode decomposition}]{EMD}{EMD}
+\glsxtrnewsymbol[description={European network of transmission system operators for electricity}]{ENTSOE}{ENTSOE}
 \glsxtrnewsymbol[description={Electricity load forecasting}]{EPF}{EPF}
@@
 
+% T
+\glsxtrnewsymbol[description={Hadamard product}]{odot}{\ensuremath{\odot}}
+
 % V

### 2024-05-14 thesis/input_files/energymarket.tex
May 14th 2024

@@
 \subsection{The marketplace}
-We can differential between two kinds of electricity markets: power pools and power exchanges.
+We can differentiate between two kinds of electricity markets: power pools and power exchanges.
 In power pools, generators bid the prices at which they are willing to produce at different volumes.
-Then, the market clearing price MCP is determined by intersecting the aggregated supply curve and the estimated demand, left panel figure \ref{fig:pool_vs_echange}.
+Then, the market clearing price (MCP) is determined by intersecting the aggregated supply curve and the estimated demand, left panel figure \ref{fig:pool_vs_echange}.
 Power pools are created on public initiatives of governments.
@@
 % https://medium.com/@brandonvar/day-ahead-and-intraday-electricity-markets-1eb121dcab47
-\subsection{Day ahead}
+
+% \subsection{Day ahead}
 The day ahead market is based on a blind auction taking place every day of the year.
 During these auctions, the prices of all hours for the next day are traded.
-Market participants can enter their buy sell orders until order book closure, which takes place at 12.00 PM.
+Market participants can enter their buy and sell orders until order book closure, which takes place at 12.00 PM.
 After that, the market establishes hourly prices by intersecting the demand and supply curves for each hour of the following day.
 
-\subsection{Intraday}
+% \subsection{Intraday}
 The intraday market consists of continous trading 24/7.
@@
 The former is adopted in EU contries while the latter is used in the US, figure \ref{fig:eu_zonal} and figure \ref{fig:us_local}.
-With ZMCP, prices may differ between zones but are the same within the same area.
-On the other hand, LMP is made up by summing the transmission congestion cost, generation marginal cost and the cost of marginal losses at different buses; buses is where a electicity line or several lines are connected.
+With zonal market clearing pricing, prices may differ between zones but are the same within the same area.
+On the other hand, local marginal price is made up by summing the transmission congestion cost, generation marginal cost and the cost of marginal losses at different buses; buses is where a electicity line or several lines are connected.
 \begin{figure}[!h]

### 2024-05-14 thesis/input_files/metrics.tex
May 14th 2024

--- 2024-03-06 thesis/input_files/metrics.tex
@@
-Proper evaluation methods guide researchers in choosing the model that best fits their needs; thus, this chapter is dedicated to the most common evaluation metrics adopted by academics in the field of EF.
-\section{MAPE}\label{mape}
+Proper evaluation methods guide researchers in choosing the model that best fits their needs; thus, this chapter is dedicated to the most common evaluation metrics adopted by academics in the field of electricity forecasting.
+Error metrics and measures vary depending on whethter we are concerned with point or probabilistic forecasts.
+Additionally, note that the latter can take different forms which therefore requires different measures.
+\section{Mean absolute error}\label{mae}
+Consider the time series with actual values given by $y=(y_{n+1}, y_{n+2},\dots, y_{n+h})$
+and its h step ahead point forecast $\hat{y}=(\hat{y}_{n+1}, \hat{y}_{n+2},\dots, \hat{y}_{n+h})$ the mean absolute error (MAE) is defined as
+\begin{definition}
+    $\mathrm{MAE}(y,\hat{y})=\frac{1}{h}\| y- \hat{y}\|_{1}=\frac{1}{h}\sum\limits_{k=1}^{h}|y_{n+k}-\hat{y}_{n+k}|$
+\end{definition}
 
-\section{CRPS}
-Nevertheless, we can evaluate the integral in closed form. 
-As pointed out in \cite{proper_scores}, we can use lemma 2.2 of \cite{new_multi_tes1} or identity 17 of \cite{new_multi_test2}.
+\section{Root mean squared error}\label{rmse}
+\begin{definition}
+    $\mathrm{RMSE(y, \hat{y})}=\frac{1}{\sqrt{h}}\|y-\hat{y}\|_{2}=\sqrt{\frac{\sum\limits_{k=1}^{h}(y_{n+k}- \hat{y}_{n+k})^2}{h}}$
+\end{definition}
+
+Mean absolute error and root mean squared error posses the useful property of being expressed in the same units of the data, thus enabling meaningful comparisons.
+However, a drawback of such measures is that we cannot use them to compare accuracy between time series which have different magnitudes.
+For instance, a day ahead error of 1kWh is negligible when considering a daily demand of 100kWn while the same error is considerably big when daily demand is 2kWh.
+This consideration leads to relative accuracy scores, between those the mean absolute percentage error (MAPE) is by far the most popular.
+
+\section{Mean absolute percentage error}\label{mape}. 
+\begin{definition}
+    $\mathrm{MAPE}(y,\hat{y})=\frac{100}{h}\sum\limits_{k=1}^{h}\frac{|y_{n+k}-\hat{y}_{n+k}|}{|y_{n+k}|}$
+\end{definition}
+
+\section{Root mean squared percentage error}\label{rmspe}
+\begin{definition}
+    $\mathrm{RMSPE}(y,\hat{y})=100\cdot\sqrt{\frac{1}{h}\sum\limits_{k=1}^{h} \left(\frac{|y_{n+k}-\hat{y}_{n+k}|}{|y_{n+k}|}\right)^2}$
+\end{definition}
+Note, mean absolute percentage error and root mean squared percentage error may not be appropriate for series which have zero or very small values, for example, electricity demand at the household level; the result is a large score regardless of the absolute errors.
+Scaled errors constitute a robust family of scores.
+\section{Mean absolute scaled error}\label{mase}
+\begin{definition}
+    $\mathrm{MASE}(y,\hat{L})=\frac{1}{h}\sum\limits_{k=1}^N\frac{|y_{n+k}-\hat{y}_{n+k}|}{\frac{1}{h-1}\sum\limits_{k=2}^{h}|y_{k}-y_{k-1}|}$
+\end{definition}
+In the denominator we have the error of the naïve/persistence model. 
+In this model, the current demand makes up the prediction for the next time step; that is $\hat{y}^{\mathrm{naive}}_{n+1}=y_{n}$.
+\section{Root mean squared scaled error}\label{rmsse}
+\begin{definition}
+    $\mathrm{RMSSE}(y,\hat{y})=\sqrt{\frac{1}{h}\sum\limits_{k=1}^N\left(\frac{|y_{n+k}-\hat{y}_{n+k}|}{\frac{1}{h-1}\sum\limits_{k=2}^{h}|y_{k}-y_{k-1}|}\right)^2}$
+\end{definition}
+
+\section{Pinball}\label{pinball}
+The pinball score or quantile score is used to measure the accuracy of a quantile forecast.
+\begin{definition}
+    $\mathrm{Pinball}(y_{t},\hat{y}_{t,q},q)=
+\begin{cases}
+(q-1)(y_{t}-\hat{y}_{t,q}) & y_t > \hat{y}_{t,q} \\
+q(\hat{y}_{t,q}-y_t) & y_t \leq \hat{y}_{t,q}
+\end{cases}$
+\end{definition}
+The pinball loss is an asymmetric function, it weights its score differently depending on the error sign and on the quantile considered, see figure \ref{fig:pinball}.
+\begin{figure}
+    \includegraphics[width=\textwidth]{images/pinball_loss.png}
+    \caption{Pinball loss}
+    \label{fig:pinball}
+  \end{figure}
+By averaging all the pinball losses over all quantiles and over the whole forecast horizon, we obtain the pinball loss of the probabilistic forecast.
+\section{Winkler}
+\begin{definition}
+    $\mathrm{Winkler}(y_t, \hat{y}_t,q)=\begin{cases}
+        \delta & l_{t}\leq y_{t}\leq u_{t}\\
+        \delta+2(l_{t}-y_{t})/\alpha & y_{t}< l_{t}\\
+        \delta+2(l_{t}-u_{t})/\alpha & y_{t} \geq u_{t}
+    \end{cases}$
+\end{definition}
+Where delta is the prediction interval (PI) width, that is $\delta=b_t-a_t$ where $u$ is the prediction interval upper threshold and $l$ is the prediction interval lower threshold.
+This score penalizes observations falling outside the prediction interval and rewards narrow prediction intervals.
+\section{Continous ranked probability score}
+The continous ranked probability score (CRPS) measures the differece between the estimated cumulative distribution $\hat{F}$ and the empirical cumulative density function (CDF).
+\begin{definition}\label{def_crps}
+    $\mathrm{CRPS}(y, \hat{F})=\int\limits_{-\infty}^{\infty}\left(\hat{F}(x)-\mathbb{1}(x-y) \right)^2 dx$
+\end{definition}
+% Nevertheless, we can evaluate the integral in closed form. 
+Where the indicator function is defined as 
+    $\mathbb{1}(z)=
+\begin{cases}
+0 & z<0\\
+1 & z \geq 0
+\end{cases}$
 \\
-The lemma states that ...
+For a visualisation see, \ref{fig:crps}. The grey area is what contributes toward the CRPS score.
+The better the estimated cumulative density function is the smaller the total CRPS score will be.
+\begin{figure}
+    \includegraphics[width=\textwidth]{images/crps.png}
+    \caption{CRPS integral \cite{haben2023core}}
+    \label{fig:crps}
+  \end{figure}
   \\
-Note that in our case the distribution G of Y1 and Y2 is degenerate, with all probability mass on a single point (x notation will need to be cleaned here). Since $G(t)=:$
+It is worth noting, that the CRPS integral can be rewritten in terms of expectations.
+This makes its evaluation easier, since we know that the sample mean converges to the expectation by the law of large numbers.
+This was first pointed out by \cite{proper_scores}, where authors take advantage of lemma 2.2 of \cite{new_multi_test2} or equivalently identity 17 of \cite{new_multi_tes1}.
 \\
-It follows that the third addend in the summation is zero since the expectation of Y1 Y2 with distribution G means the difference of two equal constant numbers.
+\begin{lemma}
+    Let $X_1$, $X_2$, $Y_1$, $Y_2$ be independent real random variables with finite expectations.
+Let $X_1,X_2$ be identically distributed with distribution function $F$ and let $Y_1,Y_2$ be identically distributed with distribution function $G$.
+Then
+   \begin{equation}
+    \mathbb{E}(|X_1-Y_1|)-\frac{1}{2}\mathbb{E}(|X_1-X_2|)-\frac{1}{2}\mathbb{E}(|Y_1-Y_2|)=\int\limits_{-\infty}^{\infty}\left(F(x)-G(x)\right)^2dx
+\end{equation}
+\end{lemma}
+Notice that, in our case, equation \ref{def_crps}, the distribution $G$ of $Y_1$ and $Y_2$ is degenerate, with all probability mass on a single point.
+%  (x notation will need to be cleaned here). Since $G(t)=:$
+It follows that, the third addend in the summation is zero.
+That is because the fact of $Y_1$ and $Y_2$ both following distribution $G$ implies that $\mathbb{E}(|Y_1-Y_2|)$ corresponds to the difference of two equal constant numbers.
 \\
-Additionally, since x is constant in the first term we have Y=x.
+Additionally, since $Y_1$ is just a constant, we have $Y_1=y$.
+\\
+Putting everything together we have obtained an alternative way of computing the CRPS score.
+\begin{equation}
+    \int\limits_{-\infty}^{\infty}\left(\hat{F}(x)-\mathbb{1}(x-L) \right)^2 dx=\mathbb{E}(|X_1-y|)-\frac{1}{2}\mathbb{E}(|X_1-X_2|)
+\end{equation}
+\\
+\section{Probability integral transform}
+The probability integral transform (PIT) is a method to assess visually the quality of probabilistic a forecasts.
+PIT is obtained by applying the predicted cumulative density function $\hat{F}$ to your data; if applying such CDF to the data results in a uniform distributed PIT, then $\hat{F}$ is a valid prediction.
+If not, $\hat{F}$ is not the suited CDF for the considered data.
+Figure \ref{fig:pit} provides an example, applying the true CDF results in a well calibrated PIT (left).
+Alternatively, applying a bad CDF results in either a overdispersed (middle) or underdispersed (right) PIT.
+\begin{figure}
+    \includegraphics[width=\textwidth]{images/pit.png}
+    \caption{PIT types \cite{haben2023core}}
+    \label{fig:pit}
+  \end{figure}
+\\
+%RELIABILITY PLOT
+%PICP,NMPIW,CWC,unconditional coverage(arora)-->Probability density forecasting of wind speed based on quantile regression and kernel density estimation Lei Zhang
+
+% Reliability and sharpness \section{Criteria}-->Recent advances in electricity price forecsating a review of probabilistic forecasting Weron
+% Two criteria are used in literature to evaluate probabilistic forecasting: sharpness and reliability. Reliability means

### 2024-05-14 thesis/input_files/paper1.tex
May 14th 2024

--- 2024-04-29 thesis/input_files/paper1.tex
@@
 % Some commands used in this file
+The following section covers the theory of kernel methods and introduces the building blocks underlying it.
+% \section{Kernel mean embedding of distributions: A review and beyond}
 
-\section{Kernel mean embedding of distributions: A review and beyond}
 % \href{https://arxiv.org/abs/1605.09522}{
-From this first paper \cite{Muandet_2017}, the notation and terms used in the theory of Reproducing Kernel Hilbert Spaces are summarized.
-\\
-Many algorithms use the inner product as similarity measure between data instances $x, x' \in \mathcal{X}$. However, this inner product spans only the class of linear similarity measures. 
+% From this first paper \cite{Muandet_2017}, the notation and terms used in the theory of Reproducing Kernel Hilbert Spaces are summarized.
+% \\
+
+Many algorithms use the inner product as similarity measure between the data instances $x, y \in \mathcal{X}$. However, this inner product spans only the class of linear similarity measures. 
 \\
@@
 \begin{align*}
-\phi(x)\colon \mathcal{X} &\to \mathcal{F}
+\phi(x)\colon \mathcal{X} &\to \mathcal{H}
     \\
@@
 
-We take the inner product in the high dimensional space $\mathcal{F}$ mapped by $\phi(x)$, i.e.\
+We take then the inner product in the high dimensional space $\mathcal{H}$ mapped by $\phi(x)$, i.e.\
 \\
 \begin{align*}
-k(x,x'):=&\langle \phi(x), \phi(x') \rangle_{\mathcal{F}}
+k(x,y):=&\langle \phi(x), \phi(y) \rangle_{\mathcal{H}}
 \end{align*}
@@
 \\
-Therefore, we can kernelize any algorithm involving a dot product by substituting $\langle x, x' \rangle_{\mathcal{X}}$ with $\langle \phi(x), \phi(x') \rangle_{\mathcal{F}}$
+Therefore, we can kernelize any algorithm involving a dot product by substituting $\langle x, y \rangle_{\mathcal{X}}$ with $\langle \phi(x), \phi(y) \rangle_{\mathcal{H}}$
 \\
-One would expect constructing the feature maps explicitly and then evaluating their inner product in $\mathcal{F}$ to be computationally expensive, and indeed it is.
+One would expect constructing the feature maps explicitly and then evaluating their inner product in $\mathcal{H}$ to be computationally expensive, and indeed it is.
 However, we do not have to explicitly perform such calculations.
@@
 \\
-Suppose $x \in \mathbb{R}^2$ and assume $\phi(x)=(x_{1}^{2},x_{2}^{2},\sqrt{2}x_{1}x_{2})$, then the inner product in the feature space is $x_{1}^{2}x_{1}^{'2},x_{2}^{2}x_{2}^{'2}+2x_{1}x_{2}x'_{1}x'_{2}$.
-Notice that this is the same of $\langle \phi(x), \phi(x') \rangle$; thus the kernel trick consists of just using $k(x,x')=:(x^Tx')^2$.
+Suppose $x,  \in \mathbb{R}^2$ and assume $\phi(x)=(x_{1}^{2},x_{2}^{2},\sqrt{2}x_{1}x_{2})$, then the inner product in the feature space is $x_{1}^{2}y_{1}^{2},x_{2}^{2}y_{2}^{2}+2x_{1}x_{2}y_{1}y_{2}$.
+Notice that this is the same of $\langle \phi(x), \phi(y) \rangle$; thus the kernel trick consists of just using $k(x,y)=:(x^\intercal y)^2$.
+For example
+\\
+$\langle \phi([2,4]),\phi([8,9]) \rangle=  \left[ {\begin{array}{cccc}
+    4 & 16 &     \sqrt{2}\cdot 2 \cdot 4\\
+  \end{array} } \right]
+  \left[ {\begin{array}{cccc}
+    64\\
+    81\\
+    \sqrt{2}\cdot8 \cdot 9
+    \\
+  \end{array} } \right]\\=4\cdot 64 +16\cdot 81 +2 \cdot 2 \cdot 4 \cdot 8 \cdot 9=2704
+$
+\\
+\\
+While $k([2,4], [8,9])=(16+36)^2=2704$
 
-\subsection{Reproducing kernel hilbert space}
+This gives a taste of how powerful kernels are thanks to the kernel trick.
+Indeed, the idea of the the kernel trick can be extended to feature maps
+$\phi$ involving an infinite feature space. In these cases, calculating the
+respective kernel is equal to calculating the inner product between an infinite
+number of features of the data points.
+
+\section{Theory of kernels}
 Subsequently we introduce the definitions that make up the basis for the theory of kernel methods.
@@
 \begin{definition}
+    A sequence $\{v_n\}_{n=1}^{\infty}$ is convergent if and only if there exists a point $ v \in \mathcal{V}$ such that for every $\epsilon>0$ there exists $N=N(\epsilon) \in \mathbb{N}$ such that  $\|v_n-v\|_{\mathcal{V}}$ $\forall n\geq N$.
+\end{definition}
+
+\begin{definition}
     A complete metric space is a metric space in which every Cauchy sequence is convergent.
@@
 \begin{definition}
-    A Hilbert space is a vector space $\mathcal{H}$ with an inner product $\langle \cdot, \cdot \rangle$ such that the norm defined by $\|f\|=\sqrt{\langle f, f \rangle}$
+    A Hilbert space is a vector space $\mathcal{H}$ with an inner product $\langle \cdot, \cdot \rangle$ such that the norm defined by $\|f\|_{\mathcal{H}}=\sqrt{\langle f, f \rangle}_{\mathcal{H}}$
 turns $\mathcal{H}$ into a complete metric space.
 \end{definition}
+
+\begin{definition}
+    A function $k: \mathcal{X} \times \mathcal{X} \to \mathbb{R}$ is said to be finitely positive semidefinite if it is a symmetric function for which the matrices formed with it are positive semidefinite.
+\end{definition}
+
+\begin{theorem}[Characterisation of kernels]    
+    A function $k:\mathcal{X} \times \mathcal{X} \to \mathbb{R}$ which is either continous or has finite domain can be decomposed as $k(x,y)=\langle \phi(x), \phi(y) \rangle_{\mathcal{H}}$ lie if and only if $k$ satisfy the positive semidefinite property.
+\end{theorem}
+
+That is the function $k$ can be written as the inner product in the Hilbert space $\mathcal{H}$ of the features evaluated at the arguments $\phi(x)$.
+
+Next, we make this characterisation explicit by showing how to construct a feature mapping $\phi$ of a kernel $k$ satisfying the finitely positive semidefinite property.
+Let the feature space be the set of functions 
+\\
+$\mathcal{F} =\{ \sum\limits_{i=1}^n \alpha_i k(x_i, \cdot): n \in \mathbb{N}, x_i \in \mathcal{X}, a_i \in \mathbb{R}, i=1,\cdots, n\}$.
+This space is clearly closed under addition of functions and scalar multiplication, therefore it is a vector space.
+Letting $f(x)=\sum\limits_{i=1}^n \alpha_i k_i(x_i,x)$ and $g(x)=\sum\limits_{i=1}^m \beta_i k_i(z_i,x)$,  we can introduce an inner product on $\mathcal{F}$ as 
+\\
+$\langle f,g \rangle_{\mathcal{F}}=:\sum\limits_{i=1}^n\sum\limits_{j=1}^m a_i \beta_j k_i(x_i,z_j)=\sum\limits_{i=1}^n a_i g(x_i)= \sum\limits_{j=1}^m \beta_j f(z_j)$
+\\
+Notice, $\langle f, f \rangle_{\mathcal{F}}= \alpha^\intercal K \alpha \geq 0 \ \forall f \in \mathcal{F}$, where $K$ is the kernel matrix.
+Moreover, $\langle f, g \rangle_{\mathcal{F}}$ is real valued, symmetric and bilinear. All these facts implies that the just introduced inner product is a valid metric.
+\\
+Additionally, letting $g(x)=k(\cdot, x)$, we have that 
+\begin{equation}
+    \langle f, k(\cdot, x)\rangle_{\mathcal{F}}=\sum\limits_{i=1}^n a_i k(x_i,x)=f(x)    
+\end{equation}
+This is known as the reproducing property.
+\\
+Finally, to show the characterisation of kernels we have to check that the inner product makes $\mathcal{F}$ a complete metric space.
+Consider a fixed input $x$ and a Cauchy sequence $\{f_n\}_{n=1}^{\infty}$, we then have by means of the Cauchy-Schwarz inequality that 
+\\
+\begin{equation}
+    (f_n(x)- f_m(x))^2=\langle f_n-f_m, k(\cdot, x)\rangle_{\mathcal{F}}^2 \leq \| f_n - f_m \|_{\mathcal{F}}^2 k(x,x)    
+\end{equation}
+\\
+We have then, that $f_n(x)$ is a Cauchy sequence of real numbers.
+Thus, we can let $g(x)=\lim_{n\to \infty} f_n(x)$ and include all such limits functions in $\mathcal{F}$, doing so we complete $\mathcal{F}$ and we obtain the Hilbert space $\mathcal{H}$ associated to the kernel $k$.
+Having constructed the feature space $\mathcal{H}$, the mapping $\phi(x)$ is defined as follows
+\\
+\begin{equation}
+    \phi: x \in \mathcal{X} \to \phi(x)=k(\cdot, x) \in \mathcal{H}
+\end{equation}
+\\
+Notice, that by the reproducing property, $f$ can be represented as the inner product with itself in the feature space, that is 
+\begin{equation}
+    f(x)=\langle f, \phi(x)\rangle_{\mathcal{H}}=\langle f, k(\cdot, x)\rangle_{\mathcal{H}}=f(x)
+\end{equation}
 
@@
      \begin{align}
-        k(x, \cdot) \in \mathcal{H} &\quad \text{for all   } x \in \mathcal{X}    \\
-        \langle  h, k(x, \cdot) \rangle_\mathcal{H} = h(x) &\quad \text{for all   } h \in \mathcal{H}, x \in \mathcal{X}
+        k(x, \cdot) \in \mathcal{H} &\quad \ \forall x \in \mathcal{X}    \\
+        \langle  f, k(x, \cdot) \rangle_\mathcal{H} = f(x) &\quad \ \forall f\in \mathcal{H}, x \in \mathcal{X}
     \end{align}
+    Notice, the finitely positive semidefinite property of kernel $k$ is a sufficient condition for $\mathcal{H}$ to have an associated RKHS.
 \end{definition}
+
+The analysis of above applies to any kernel, that is given a kernel function we can always construct its RKHS.
+\begin{theorem}
+    If a symmetric function $k(\cdot, \cdot)$ satisfies the reproducing property in a Hilbert space $\mathcal{H}$, then $k$ satisfies the finitely positive semidefinite property.
+\end{theorem}
+
+\begin{proof}
+    \begin{align}
+        \sum\limits_{i,j=1}^n a_i a_j k(x_i, x_j)=& \sum\limits_{i,j=1}^n a_i a_j \langle k(\cdot, x_i), k(\cdot, x_j)\rangle_{\mathcal{H}}
+        \\
+        =& \left\langle \sum\limits_{i=1}^n a_i k(\cdot, x_i), \sum\limits_{j=1}^n a_j k(\cdot, x_j) \right\rangle_{\mathcal{H}} \\
+        =& \left\| \sum\limits_{i=1}^n a_i k(\cdot, x_i)\right\|_{\mathcal{H}}^2 \geq 0
+    \end{align}
+\end{proof}
 
@@
     [Riesz Representation].
-If $A : \mathcal{H} \rightarrow \mathbb{R}$ is a bounded linear operator on a Hilbert space $\mathcal{H}$ , there exists some $g_{A} \in \mathcal{H}$ such that $A(f) = \langle f,g_A\rangle_\mathcal{H}, \ \forall f \in \mathcal{H}$.
+If $L : \mathcal{H} \rightarrow \mathbb{R}$ is a bounded linear operator on a Hilbert space $\mathcal{H}$ , there exists some $h_0 \in \mathcal{H}$ such that $L(f) = \langle f,h_0\rangle_\mathcal{H}, \ \forall f \in \mathcal{H}$.
 \end{theorem}
+
+\begin{proof}
+    It is known that null space of $L$, call it $M$, is a closed linear subspace in $\mathcal{H}$.
+This implies that the Hilbert space direct sum $M \bigoplus M^\bot$ is an isomorphism of $\mathcal{H}$.
+Letting $L$ be different that the zero operator we have that $M^\bot \neq {0}$.
+    \\
+    Choose now a $z \in M^\bot$ with norm 1.
+By linearity of $L$, for any $f \in \mathcal{H}$ we have $L(zL(f)- fL(z))=L(z)L(f)- L(f)L(z)=0$.
+This means that $zL(f)- fL(z)$ belongs to the null space of $L$, that is  $zL(f)- fL(z) \in M$.
+    \\
+    Consider now 
+    \begin{align}
+        L(f)=& L(f) \cdot 1
+        \\
+        =& L(f) \langle z,z\rangle_{\mathcal{H}}
+        \\
+        =& \langle zL(f), z\rangle_{\mathcal{H}} \\
+        =& \langle zL(f) -fL(z)+fL(z), z\rangle_\mathcal{H}
+        \\
+        =& \langle zL(f) -fL(z), z\rangle_{\mathcal{H}}+\langle fL(z), z\rangle_{\mathcal{H}} \\
+        % this is zero zL(f) -fL(z), so
+        =& \langle fL(z), z\rangle_{\mathcal{H}} \\
+        =& \langle f, z\overline{L(f)}\rangle_{\mathcal{H}} \\
+    \end{align}
+    Therefore, $L(f)=\langle f, h_0\rangle$  with  $h_0=z\overline{L(z)}$
+    \\
+    Finally, notice that $h_0$ is unique.
+    Assuming $h_0$ and $h_1$ to both satisfy the above equation for all $f \in \mathcal{H}$ we have that
+    \begin{align}
+        \langle h_0,f \rangle=&\langle h_1, f \rangle \\
+        \langle h_0,f \rangle-\langle h_1, f \rangle=& 0 \\
+        \langle h_0 -h_1,f \rangle=& 0
+    \end{align}
+    Setting $f=h_0 -h_1$, we have the uniqueness of $h_0$.
+    % because inner product is zero, inner product has to satisfy positive definitness property to be such >=0, if it is zero it means that its argument x in <x,x> is zero => h_0=h_1
+
+\end{proof}
+%conjugate simmetry is because in complex analysis when you swap in the inner product, you have to take the conjugate
 
@@
 \begin{proposition}
-For each $x \in \mathcal{X}$ there exists a function $k_{x} \in \mathcal{H}$ such that $\mathcal{F}_{x}(f)=\langle k_{x}, f\rangle_{\mathcal{H}}=f(x)$    
+For each $x \in \mathcal{X}$ there exists a function $k_{x} \in \mathcal{H}$ such that the evaluation functional $F_{x}(f)=\langle k_{x}, f\rangle_{\mathcal{H}}=f(x)$    
 \end{proposition}
@@
 \begin{align*}
-    k_{x}(y)=\mathcal{F}_{y}(k_{x})=\langle k_{x}, k_{y}\rangle_{\mathcal{H}=\langle \phi(x), \phi(y)\rangle_{\mathcal{H}}}
+    k_{x}(y)=F_{y}(k_{x})=\langle k_{x}, k_{y}\rangle_{\mathcal{H}}=\langle \phi(x), \phi(y)\rangle_{\mathcal{H}}
 \end{align*}
 
-\subsection{Kernel families}
+\begin{definition}
+    Considering a set of vectors $x_1, \cdots, x_n$, the Gram matrix is defined as the $n\times n$ matrix whose entries are $\langle x_i, x_j \rangle$.    
+\end{definition}
+
+Notice, it follows from the definition that Gram matrix is symmetric.
+In the kernel setting we evaluate the inner products in the feature space generated by the feature map $\phi$, therefore the Gram matrix will have entries $\langle \phi(x_i), \phi(x_j) \rangle_{\mathcal{H}}= k(x_i, x_j)$.
+Such matrix is also referred as the kernel matrix $K$.
+
+\[
+K_{n\times n}=
+\begin{bmatrix}
+    k(x_1,x_1)       &  \dots & k(x_1,x_n) \\
+    \vdots       & \ddots & \vdots \\
+    k(x_n, x_1)       & \dots & k(x_n, x_n)
+\end{bmatrix}
+\]
+All kernel methods take as input the kernel matrix, it is the central data type for all 
+such techniques.
+
+\begin{proposition}
+    In addition to being symmetric, $K$ possesses also the property of being positive semidefinite.    
+\end{proposition}
+
+\begin{proof}
+    \begin{align}
+        z^\intercal K z =& \sum_{i,j=1}^n z_i z_j K_{i,j}
+        \\
+        =& \sum_{i,j=1}^n z_i z_j \langle \phi(x_i), \phi(x_j) \rangle_{\mathcal{H}}
+        \\
+        =& \left\langle \sum_{i=1}^n z_i \phi(x_i), \sum_{j=1}^n z_j \phi(x_j)\right\rangle_{\mathcal{H}}
+        \\
+        =& \left\|\sum_{i=1}^n z_i \phi(x_i) \right\|_{\mathcal{H}}^2 \geq 0
+    \end{align}
+\end{proof}
+We know that positive semidefinite matrices form a cone in the vector space $n\times n$ matrices.
+Because of this we have that such matrices are convex, this implies that all the nice theory of convex analysis can be used when optimising over them.
+
+Kernel functions are characterized by a couple of properties that make them a closure. Thus, there exists ways of combining known kernels in order to obtain new kernels.
+
+\begin{proposition}
+    Let $k_1$ and $k_2$ be kernels over $\mathcal{X} \times \mathcal{X}, \mathcal{X} \subset \mathbb{R}^n$, $a \in \mathbb{R}^+$ and $\phi: \mathcal{X} \to \mathbb{R}^N$ with $k_3$ a kernel over $\mathbb{R}^N \times \mathbb{R}^N$.
+Then the following functions are valid kernels:
+    \begin{align}
+        k(x,z)=&k_1(x,z)+k_2(x,Y)
+        \\
+        k(x,z)=& ak_1(x,z)
+        \\
+        k(x,z)=& k_1(x,z)k_2(x,z)
+        \\
+        k(x,z)=& k_3(\phi(x),\phi(z))
+    \end{align}.
+\end{proposition}
+
+\begin{proof}
+    Consider a finite set of points $x_1, \cdots, x_n$, a vector $z \in \mathbb{R}^n$ and let $K_1$ and $K_2$ be the kernel matrices associated to $k_1$ and $k_2$ evaluted on these points.
+    \\
+    We have:
+    \\
+    \begin{align}
+    z^\intercal (K_1+K_2) z= z^\intercal K_1 z + z^\intercal K_1 z \geq 0
+    \\
+    z^\intercal aK z= a z^\intercal K z \geq 0
+    \end{align}
+    Let $K=K_1 \otimes K_2$, we have that the eigenvalues of $K$ are made up by the product pairs of the eigenvalues of $K_1$ and $K_2$.
+Thus, $K$ is positive semidefinite.
+Next, the matrix given by $k_1 \cdot k_2$ corresponds to the Schur product of $K_1$ and $K_2$, denote if $H$.
+Notice, $H$ is a principal submatrix of $K$, hence for any $z \in \mathbb{R}^n$ there exist a $t \in \mathbb{R}^{n^2}$ such that $z^\intercal H z= t^\intercal K t$.
+Using the fact that K is positive semidefinite we have that for any $z$, it holds
+    \begin{align}
+        z^\intercal H z= t^\intercal K t\geq 0
+    \end{align}
+\end{proof}
+
+\section{Kernel families}
+The selection of kernel determines the class of functions that will be searcehd by the learning algorithm.
+Prior knowledge of the problem domain can help restricting the candidate families of kernel functions.
 Table \ref{tab:kernel types} contains popular kernel families in literature and applications.
@@
        \midrule
-       Linear &  $k(x_1,x_2)=x_1x_2$ &   \\
-       Polynomial &  $k(x_1,x_2)=(x_1^\intercal x_2+c)^d$ &   c, d\\
-       Gaussian RBF &  $k(x_1,x_2)=e^{-\frac{\|x_1-x_2|\|^2}{2\sigma^2}}$ &   $\sigma$\\
-       Exponential RBF/Laplacian& $k(x_1,x_2)=e^{-\frac{|x_1-x_2|}{\gamma}}$ &   $\gamma$\\
-       Hyperbolic/Sigmoid Kernel &  $k(x_1,x_2)=\tanh(\gamma x_1^\intercal x_2+r)$ &  $\gamma$, r \\
-       Periodic &  $k(x_1,x_2)=e^{\frac{-2 sin^2\left(\frac{\pi}{p}|x_1-x_2| \right)}{l^2}} $ &   p, l\\
-       Chi-squared kernel &  $k(x_1,x_2)=e^{-\gamma \sum\limits_i \frac{(x_i - y_i)^2}{x_i+y_i}} $ &   $\gamma$\\
-        Cosine  &  $k(x_1,x_2)=\frac{xy^\intercal}{\|x\|_{L_2}\|y\|_{L_2}}$ &   \\
-       Matern  &  $k(x_1, x_2) =  \frac{1}{\gamma(\nu)2^{\nu-1}}\Bigg(
-        \frac{\sqrt{2\nu}}{l} d(x_1 , x_2 )
-        \Bigg)^\nu K_\nu\Bigg(
-        \frac{\sqrt{2\nu}}{l} d(x_1 , x_2 )\Bigg)$ &  $\gamma, \nu$ \\
+       Linear &  $k(x,y)=xy$ &   \\
+       Polynomial &  $k(x,y)=(x^\intercal y+c)^d$ &   c, d\\
+       Gaussian RBF &  $k(x,y)=\exp\left(-\frac{\|x-y|\|_{2}^2}{2\sigma^2}\right)$ &   $\sigma$\\
+       Exponential RBF/Laplacian& $k(x,y)=\exp\left(-\frac{\|x-y\|_{1}}{\gamma}\right)$ &   $\gamma$\\
+       Hyperbolic/Sigmoid Kernel &  $k(x,y2)=\tanh(\gamma x^\intercal y+r)$ &  $\gamma$, r \\
+       Periodic &  $k(x,y)=\exp\left(\frac{-2 sin^2\left(\frac{\pi}{p}\|x-y\|_{2} \right)}{l^2}\right) $ &   $p, l$\\
+       Chi-squared kernel &  $k(x,y)=\exp\left(-\gamma \sum\limits_i \frac{(x_i - y_i)^2}{x_i+y_i}\right) $ &   $\gamma$\\
+        Cosine  &  $k(x,y)=\frac{xy^\intercal}{\|x\|_{2}\|y\|_{2}}$ &   \\
+       Matern  &  $k(x, y) =  \frac{1}{\Gamma(\nu)2^{\nu-1}}\Bigg(
+        \frac{\sqrt{2\nu}}{l} \|x-y|\|_{2}
+        \Bigg)^\nu Y_\nu\Bigg(
+        \frac{\sqrt{2\nu}}{l} \|x-y|\|_{2}\Bigg)$ &  $l, \nu$ \\
 
         \bottomrule
-    \end{tabular}
+    \end{tabular}}
     \label{tab:kernel types}
 \end{table}
+
+Notice, in the Matern family $\Gamma(\cdot)$ is the gamma function while $Y(\cdot)$ is the Bessel function of the second kind.
+\\
+When $\nu=p+\frac{1}{2}$, $p \in \mathbb{N}^+$, the Matern kernel can be rewritten as a product of an exponential and a polynomial of degree $p$. 
+\\
+$k(x,y)=\exp \left(-{\frac {{\sqrt {2p+1}}\|x-y|\|_{2}}{l }}\right){\frac {p!}{(2p)!}}\sum _{i=0}^{p}{\frac {(p+i)!}{i!(p-i)!}}\left({\frac {2{\sqrt {2p+1}}\|x-y|\|_{2}}{l }}\right)^{p-i}$
+\\
+For example:
+\begin{itemize}
+    \item $p=0,\nu=\frac{1}{2} \implies k(x,y)=\exp\left(\frac{-\|x-y|\|_{2}}{l}\right)$
+    \item $p=1,\nu=\frac{3}{2} \implies 
+    k(x,y)=\left(1+{\frac {{\sqrt {3}}\|x-y|\|_{2}}{l }}\right)\exp \left(-{\frac {{\sqrt {3}}\|x-y|\|_{2}}{l }}\right)$
+    \item $p=2,\nu=\frac{5}{2} \implies k(x,y)=\left(1+{\frac {{\sqrt {5}}\|x-y|\|_{2}}{l }}+{\frac {5\|x-y|\|_{2}^{2}}{3 l ^{2}}}\right)\exp \left(-{\frac {{\sqrt {5}}\|x-y|\|_{2}}{l }}\right)$
+\end{itemize}
+
+Furthermore, as $\nu \to \infty$ we have that the Matern kernel converges to the Gaussian RBF kernel.
+$k(x,y)=\exp\left(\frac{-\|x-y|\|_{2}^2}{2l^2}\right)$
+
+% CONDITIONS upon matern is equivalent to rbf
+% KERNEL THEORY FROM BOOKS AND KERNEL COOKBOOK
+
+Any algorithm involving inner products between inputs can be combined with a kernel methods.
+The chosen kernel function will calculate the inner product between the images of two inputs in the feature space.
+This makes the orginal algorithm operate in a high dimesional space.
+Kernel methods show good modularity because they work the same for any kernel and any data type.
+This concept is illustrated in figure \ref{fig:workflow_kernels}, where an abstract kernel methods workflow is visualized.
+First we build the kernel matrix by using a kernel function to process the input data.
+Next, the kernel matrix is passed to the pattern analysis algorithm, which returns the learned function.
+
+\begin{figure}[!ht]
+    \includegraphics[width=\textwidth]{images/workflow_kernels.png}
+    \caption{Workflow for the application of kernel methods \cite{shawe2004kernel}}
+    \label{fig:workflow_kernels}
+\end{figure}

### 2024-05-14 thesis/input_files/point.tex
May 14th 2024

@@
-This chapter covers the theory of the most widely used method for point forecasting in the EF literature.
+This chapter covers the theory of the most widely used method for point prediction in the electricity forecasting literature.
 Besides discussing the theory underlying such models, this and the following chapter include also a couple of worked examples in order to get acquainted with the practical applications of these models.
@@
 \section{Multiple linear regression}
-Notwithstanding its simplicity, multiple linear regression models are still popular among the EPF literature. Notice, they are not used per se but usually combined with other more advanced models.
+Notwithstanding its simplicity, multiple linear regression models are still popular among the electricity point forecasting literature.
+Notice, they are not used per se but usually combined with other more advanced models.
 \begin{equation}
-    L_t=\beta X_t + \epsilon_t
+    y_t=\beta X_t + \epsilon_t
 \end{equation}
@@
 Autoregressive models are a standard approach for modelling time series data.
-Before introducing the popular model in EPF, it is important to remember that autoregressive models assume the time series to be stationary.
+Before introducing the popular model in electricity point forecasting, it is important to remember that autoregressive models assume the time series to be stationary.
 On the other hand, load and price have been observed to be a non stationary time series.
 Basically, stationarity means that the distribution of any subsequence of random variables of the stochatic process is invariant to shifts along the time dimension.
-\\
 Therefore, checking stationarity is a crucial step in applying this class of models. %(unit root tests)
@@
 \begin{equation}
-    \hat{L}_N=\sum\limits_{i=1}^{p}\phi_i L_{N-i}^{(d)}+\sum\limits_{j=1}^{q}\theta_j \epsilon_{N-j}+\epsilon_N
+    \hat{y}_n=\sum\limits_{i=1}^{p}\psi_i y_{n-i}^{(d)}+\sum\limits_{j=1}^{q}\theta_j \epsilon_{n-j}+\epsilon_n
 \end{equation}
-where the parameter p is the order of the AR part, d is the degree of integrated differencing and q is the order of the MA model.
-$\epsilon_{N-j}$ are the observed errors in the past while the differenced term $L_N^{(d)}$ is defined recursively as $L_N^{(d)}=L_N^{(d-1)}-L_{N-1}^{(d-1)}$.
+where the parameter p is the order of the autoregressive (AR) part, d is the degree of integrated differencing and q is the order of the moving average (MA) model.
+$\epsilon_{n-j}$ are the observed errors in the past while the differenced term $y_n^{(d)}$ is defined recursively as $y_n^{(d)}=y_n^{(d-1)}-y_{n-1}^{(d-1)}$.
+\\
 ARIMA model are compactly written as
 \begin{equation}
-    \phi(B)\nabla^d y_t=\theta(B)\epsilon_t
+    \psi(B)\nabla^d y_t=\theta(B)\epsilon_t
 \end{equation}
-where $B$ is the backward shift operator, $B^h y_t=y_{t-h}$ and $\phi(B)$ is shorthand for $\phi(B)=1-\phi_1B-\dots-\phi_p B^p$.
+where $B$ is the backward shift operator, $B^h y_t=y_{t-h}$ and $\psi(B)$ is shorthand for $\psi(B)=1-\psi_1B-\dots-\psi_p B^p$.
 Similarly, $\theta(B)=1+\theta_1B+\dots+\theta_q B^q$ and $\nabla_h$ is the lag h differencing operator $\nabla_h y_t=(1-B^h)y_t=y_t-y_{t-h}$
@@
 \begin{equation}
-    \hat{L}_N=\sum\limits_{k=1}^{h} \mu_i X_{N-k}+\sum\limits_{i=1}^{p}\phi_i L_{N-i}^{(d)}+\sum\limits_{j=1}^{q}\theta_j \epsilon_{N-j}+ \epsilon_{N}
+    \hat{y}_n=\sum\limits_{k=1}^{h} \mu_i X_{n-k}+\sum\limits_{i=1}^{p}\psi_i y_{n-i}^{(d)}+\sum\limits_{j=1}^{q}\theta_j \epsilon_{n-j}+ \epsilon_{n}
 \end{equation}
@@
 \begin{equation}
-    \phi(B)\Phi(B^s)\nabla^d\nabla_m^Dy_t=\theta(B)\Theta(B^s)\epsilon_t
+    \psi(B)\Psi(B^s)\nabla^d\nabla_m^Dy_t=\theta(B)\Theta(B^s)\epsilon_t
 \end{equation}
-
+where $\Psi$ and $\Theta$ account for the $P$ and $Q$ of the seasonal patter $m$ respectively.
 \section{Generalized additive model}
@@
 \\
-The goal of support vector regression is finding a function $f(x)$ with at most $\epsilon$ deviation from the actual observed data $y_i$ for every i and as flat as possible. 
+The goal of support vector regression is finding a function $f(x)$ with at most $\epsilon$ deviation from the actual observed data $y_i$ for every $i$ and as flat as possible. 
 Put differently, we would like a model to keep error less than the $\epsilon$ threshold,  
@@
     \|\xi\|_\epsilon:=\begin{cases}
-        0 \quad& \textrm{if} \ \|\xi\|\leq \epsilon
+        0 \quad& \textrm{if} \ |\xi|\leq \epsilon
         \\
-        \|\xi\|-\epsilon \quad& \textrm{if} \ \|\xi\| > \epsilon
+        |\xi|-\epsilon \quad& \textrm{if} \ |\xi| > \epsilon
     \end{cases}
@@
 \end{figure}
-Smola et al. \cite{smola2004tutorial} point out that considering the dual formulation makes our optimisation problem easier to solve. Doing so we have
+Smola et al. \cite{smola2004tutorial} point out that considering the dual formulation makes our optimisation problem easier to solve. Doing so we have the equivalent optimisation problem
 \begin{equation}
@@
 This means that $w$ can be completely described as a linear combination of the training features $x_i$.
-Notice, the complexity of the function representation is independent of the feature space dimensionaly, but depends only on the number of support vectors; that is those point $i$ for which $(\alpha_i-\alpha_i^*)\neq 0$
+Notice, the complexity of the function representation is independent of the feature space dimensionality, but depends only on the number of support vectors; that is those point $i$ for which $(\alpha_i-\alpha_i^*)\neq 0$
 Moreover, we do not needd to compute $w$ explicitly in order to evaluate $f(x)$.
@@
 % https://rtavenar.github.io/deep_book/book_en.pdf
-LSTM extends RNN by introducing a cell state in its layers.
-In this way, the current state of a cell depndes on the current value $X_t$, on the previous cell activation and on the previous cell state $C_{t-1}$, see figure \ref{fig:lstm_cell} for a visual representation.
+Long short term memory network (LSTM) extends MLP by introducing a cell state in its layers.
+In this way, the current state of a cell depends on the current value $X_t$, on the previous cell activation and on the previous cell state $C_{t-1}$, see figure \ref{fig:lstm_cell} for a visual representation.
 \begin{figure}[!h]
@@
 \subsection{Kernel ridge regression}
-In the setting of kernel ridge regression, we aim at estimating $f$ such that $y=f(x)+\epsilon$. Given a RKHS $\mathcal{H}$, we can estimate$\hat{f}$ by solving the optmisation problem
+In the setting of kernel ridge regression, we aim at estimating $f$ such that $y=f(x)+\epsilon$. Given a RKHS $\mathcal{H}$, we can estimate $\hat{f}$ by solving the optmisation problem
 \begin{equation}\label{eq:kernel ridge}
@@
 \begin{theorem}
-    Let $k:\mathcal{X} \times \mathcal{X} \mapsto R$ on a non empty set $\mathcal{X}$ with corresponding RKHS $\mathcal{H}$ and consider the regularized risk functional $Risk[f]:f \mapsto  E\left((x_{1},y_{1},f(x_{1})),\ldots ,(x_{n},y_{n},f(x_{n}))\right)+g\left(\| f\| \right)$.
+    Let $k:\mathcal{X} \times \mathcal{X} \mapsto R$ on a non empty set $\mathcal{X}$ with corresponding RKHS $\mathcal{H}$,
+    a strictly increasing real valued function $g:[0,\infty) \to \mathbb{R}$, and $E:(\mathcal{X}\times \mathbb{R}^2)^n \to \mathbb{R} \cup \{\infty\}$.
+    Consider the regularized risk functional $Risk[f]:f \mapsto  E\left((x_{1},y_{1},f(x_{1})),\ldots, (x_{n},y_{n},f(x_{n}))\right)+g\left(\| f\| \right)$.
 Then any minimizer of the empirical risk $Risk[f]$ can be represented as $ f^{*}(\cdot )=\sum\limits_{i=1}^{n}\alpha _{i}k(\cdot ,x_{i})$
@@
     By orthogonal projection, any $f \in \mathcal{H}$ can be decomposed as the sum of two functions, with the first lying in the $\textrm{span}\{\phi(x_1),\dots, \phi(x_n)\}$ and the second staying in the orthogonal complement of the former.
-Letting $\nu$ denote the function from the orthogonal complement, we have
+Letting $\nu$ denote a function from the orthogonal complement, we have
     \begin{equation}
@@
     This means that setting $\nu=0$ does not affect the evaluation of $f$.
-    Next consider the regularization term; notice, the functional $g$ has to be strictly monotonic in order to penalize more complex functions $f$.
+    Next consider the regularization term; notice, the functional $g$ is defined to be strictly monotonic in order to penalize more complex functions $f$.
     \begin{equation}
@@
     Hence, we can conclude that setting $\nu=0$ decreases the second term while the first term is not affected.
-Thus it follows that any minimizer of the risk functional \ref{eq:kernel ridge} must be of the following form
+Thus it follows that any minimizer of the risk functional of equation \ref{eq:kernel ridge} must be of the following form
     \begin{equation}
@@
     \\
-    \hat{\alpha}=&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\lambda}{2} \langle \sum _{i=1}^{n}\alpha _{i}k(\cdot ,x_{i}),\sum _{j=1}^{n}\alpha _{j}k(\cdot ,x_{j})\rangle_{\mathcal{H}}^2
+    =&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\lambda}{2} \left\langle \sum _{i=1}^{n}\alpha _{i}k(\cdot ,x_{i}),\sum _{j=1}^{n}\alpha _{j}k(\cdot ,x_{j})\right\rangle_{\mathcal{H}}^2
     \\
-    \hat{\alpha}=&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\lambda}{2} \sum _{i,j=1}^{n}\alpha _{i}\alpha _{j}\langle k(\cdot ,x_{i}),k(\cdot ,x_{j})\rangle_{\mathcal{H}}^2
+    =&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\lambda}{2} \sum _{i,j=1}^{n}\alpha _{i}\alpha _{j}\langle k(\cdot ,x_{i}),k(\cdot ,x_{j})\rangle_{\mathcal{H}}^2
     \\
-    \hat{\alpha}=&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\lambda}{2} \sum _{i,j=1}^{n}\alpha _{i}\alpha _{j}K_{ij}
+    =&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\lambda}{2} \sum _{i,j=1}^{n}\alpha _{i}\alpha _{j}K_{ij}
     \\
-    \hat{\alpha}=&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\lambda}{2} \alpha^\intercal K \alpha
+    =&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\lambda}{2} \alpha^\intercal K \alpha
     \\
-    \hat{\alpha}=&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}(y-K\alpha)^\intercal (y-K\alpha)+\frac{\lambda}{2} \alpha^\intercal K \alpha
+    =&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}(y-K\alpha)^\intercal (y-K\alpha)+\frac{\lambda}{2} \alpha^\intercal K \alpha
 \end{aligned}
@@
 
-We have then, that $\hat{\alpha}=(K+\lambda I)^{-1}y$. It follows the solution takes the form
+We have then, that $\hat{\alpha}=(K+\lambda I)^{-1}y$. It follows that the solution takes the form
 \begin{equation}
@@
 \subsection{Kernel support vector regression}
-Support vector regression can be kernelized by swapping the $\mathbb{R}^2$ euclidean dot product of data x with the dot product in the higher feature space F.
+Support vector regression (SVR) can be kernelized by swapping the $\mathbb{R}^n$ euclidean dot product of data $x$ with the dot product in the higher feature space $\mathcal{F}$.
 Doing so, the optimisation problem can be restated as
@@
 \end{equation}
-Notice, in this setting $w$ is no longer given explicitly. Additionally, with kernel SVR we seek for the flattest function in the feature space not the input space.
+Notice, in this setting $w$ is no longer given explicitly. Additionally, with kernel support vector regression we seek for the flattest function in the feature space not the input space.
 See figure \ref{fig:svr2} and figure \ref{fig:svr3} for two examples. In the former, we have used a polynomial kernel while in the latter the standard radial basis function kernel.
@@
     \includegraphics[width=\textwidth]{images/svr2.png}
-    \caption{polynomial support vector regression}
+    \caption{Polynomial support vector regression}
     \label{fig:svr2}
@@
     \includegraphics[width=\textwidth]{images/svr3.png}
-    \caption{rbf support vector regression}
+    \caption{Gaussian RBF support vector regression}
     \label{fig:svr3}

### 2024-05-14 thesis/input_files/symbols.tex
May 14th 2024

--- 2024-03-31 thesis/input_files/symbols.tex
@@
 \glsxtrnewsymbol[description={Continous ranked probability score}]{CRPS}{CRPS}
+\glsxtrnewsymbol[description={Cumulative density function}]{CDF}{CDF}
 
@@
 %E
+\glsxtrnewsymbol
+[description={European Centre for Medium-Range Weather Forecasts}]{ECMWF}{ECMWF}
 \glsxtrnewsymbol[description={Exploratory data analysis}]{EDA}{EDA}
@@
 
+\glsxtrnewsymbol[description={Probability density function}]{PDF}{PDF}
+
 %Q
@@
 
+% H
+\glsxtrnewsymbol[description={Hadamard product}]{odot}{\ensuremath{\odot}}
+
 % I
@@
 
+% T
+\glsxtrnewsymbol[description={Tensor product}]{otimes}{\ensuremath{\otimes}}
+
 % V

### 2024-05-16 thesis/input_files/data_exploratory.tex
May 15th 2024

@@
 This section covers the datasets used in our experiments, in addition, attributes and features will be thoroughly described.
-Next, we will explain the ETL pipeline that we set up in order to ease the workflow of our comparison studies.
-Finally, in order to to bettern understand patters within the data, we carry out an exploratory data analysis.
+Next, we will explain the extract transform load pipeline (ETL) that we set up in order to ease the workflow of our comparison studies.
+Finally, in order to to better understand patters within the data, we carry out an exploratory data analysis.
 
@@
 The GEFCom2014 folder contains a zip file for each of the four tracks.
-This thesis work is focuesed on providing forecasts for the load and price quantities.
+This thesis work is focused on providing forecasts for the load and price quantities.
 \\
@@
 \subsubsection{EDA}
-In exploring this data we started off by plotting the time series of the zonal prices alongside with the toal load and zonal load, see figure \ref{fig:price_track_fig1}.
+In exploring this data we started off by plotting the time series of the zonal prices alongside with the total load and zonal load, see figure \ref{fig:price_track_fig1}.
 After that, we plotted the scatter plot for inspecting the correlation between the predictors and the dependent variable, figure \ref{fig:gefcom_zonal_price_vs_zonal_load} and figure \ref{fig:gefcom_zonal_price_vs_total_load}.
-From these visualisation, we can see their goodness for predicting zonal price since there exist a clear correlation between the candidate predictors and the target variable.
+From these visualisation, we can see their goodness for predicting zonal price since there exists a clear correlation between the candidate predictors and the target variable.
 
@@
 The data for load forecasting is contained in the GEFCom2014-L\_V2 zip file.
-Within this subfolder, we have a txt with the competition instructions and 15 subfolders one for each of the consecutive tasks.
+Within this subfolder, we have a txt file with the competition instructions and 15 subfolders one for each of the consecutive tasks.
 For each of those we have the prediction of the competition benchmark model and the train file upon which we will fit our model.
@@
 \\
-In order to compare our model performance with the winning entries of the GEFCom2014 load competition, we will refer to the Provisional\_Leaderboard\_V2 file contained also in the GEFCom2014 Data directory.
+In order to compare our model performance with the winning entries of the GEFCom2014 load competition, we will refer to the Provisional\_Leaderboard\_V2.csv file contained also in the GEFCom2014 Data directory.
 % Notice a quick premise, within this xsls file the authors opted for various scores to compare the entries performance. This choice was motivated by the need 
@@
 Figure \ref{fig:gefcom_load_historical} depicts the hisorical load while figure \ref{fig:gefcom_w_avg_historical} reports the historical weather temperature.
-Finally figure \ref{fig:gefcom_load_vs_w_avg} shows the scatter plot between the the load and the weather.
-These plots motivate the selection of weather temperature as predictor for the load.
+Finally, figure \ref{fig:gefcom_load_vs_w_avg} shows the scatter plot between the the load and the weather.
+These plots motivate the selection of weather temperature as predictors for the load.
 
@@
 \section{Germany Switzerland Austria data}
-Next, we identified various data source for Germany, Switzerland and Austria in order to further
+Next, we identified various data sources for Germany, Switzerland and Austria in order to further
 investigate the techniques studied in this thesis work.
@@
 The source of data has been the \href{https://www.energy-charts.info/index.html?l=en&c=DE}{energy charts} website.
-It provides data for data for electricity power, electricity potential capacity, energy generation, market prices, environmental measurements and also scenario simulations.
+It provides data for electricity power, electricity potential capacity, energy generation, market prices, environmental measurements and also scenario simulations.
 Notice, this data is retrieved from the European network of transmission system operators for electricity \href{https://www.entsoe.eu/}{ENTSOE} and made available thanks to the \href{https://ise.fraunhofer.de/en.html}{Fraunhofer Institute for Solar Energy Systems ISE}.
@@
 For what concers the data from SECURES-Met we first identified the data interesting to us.
-Next, we formatted all the csv to a coherent format, gave coloumns meaningful names, sliced for the years 2020 (train) and 2021 (test) and finally we merged all single csv together.
+Next, we formatted all the csv to a coherent format, gave coloumns meaningful names, sliced for the years 2020 (train) and 2021 (test) and finally we merged all single csv\textsubscript{s} together.
 
@@
 Following are reported the EDA plots alongside with insights that can be extracted from them.
-We choose to focus the following analysis on the swiss data only given that the plots for the german data are pretty similar.
+We choose to focus the following analysis on the Swiss data only given that the plots for the German data are pretty similar.
 To get started we visualized the historical load data in figure \ref{fig:CH_historical_load_2021}.

### 2024-05-16 thesis/input_files/energymarket.tex
May 15th 2024

@@
 \section{Electricity markets}
-Electricity markets have gone rapid changes all over the world during the last three decades.
+Electricity markets have undergone rapid changes all over the world during the last three decades.
 Before this revolution, the power sector was a natural monopoly.
@@
 Moreover, we can differentiate between two popular types of auctions: uniform-price/marginal and pay as bid/discriminatory.
-Within the uniform price setting, suppliers offering less than the clearing price are paid that price.
+Within the uniform price setting, suppliers offering for less than the clearing price are paid that price.
 Analogously, consumers bidding more than the clearing price pay that price.
@@
 With inflexibility, it is mean the fact that power sources cannot be switched off and restarted quickly and efficiently.
-Thus, producers are faced with the decision of either stopping and restarting their power plant or selling their energy for a negative price; that is they pay consumers for consuming their energy.
+Thus, producers are faced with the decision of either stopping and then restarting their power plant or selling their energy for a negative price; that is they pay consumers for consuming their energy.
 
@@
 Grid configuration and physical limits on electricity lines may lead to congestion.
-Local marginal price (LMP) and zonal market clearing pricing (ZMCP) are two types of pricing schemes utilized to handle it.
+Zonal market clearing pricing (ZMCP) and local marginal price (LMP) are two types of pricing schemes utilized to handle it.
 The former is adopted in EU contries while the latter is used in the US, figure \ref{fig:eu_zonal} and figure \ref{fig:us_local}.
@@
 The reason is that, smart metering and the electricity network digitalisation  will enable system operators to better monitor the network, plan their investments and manage infrastructure. 
-All of this motivates for the need of efficient and robust tools to analyze the vast data resulting from the digitalisation of the power sector.
+All of this motivates for the need of efficient and robust tools to analyze the vast data resulting from the digitalisation of the power sector see \cite{declean,andreadou2016telecommunication}.
 
@@
     \includegraphics[width=\textwidth]{images/electricity-production-by-source.png}
-    \caption{Electricity production by source}    
+    \caption{Electricity production by source \cite{energy2023statistical}}    
     \label{fig:electricity_production_by_source}

### 2024-05-16 thesis/input_files/implementation.tex
May 15th 2024

@@
 \\
-For the list of python packages needed see the requirement.txt file.
+For the list of python packages needed see the requirement.txt file on \url{https://github.com/luca-pernigo/ThesisKernelMethods}\label{github_repo}.
 
@@
 \subsection{K-nearest neighbours}
-The object KNeighborsRegressor of the sklearn module neighbors has been used.
+The object KNeighborsRegressor of the \href{https://scikit-learn.org/stable/}{sklearn} module neighbors has been used.
 
 \subsection{Support vector regression}
-The object SVR of the sklearn module svm has been used by specifying the liner kernel.
+The object SVR of the \href{https://scikit-learn.org/stable/}{sklearn} module svm has been used by specifying the liner kernel.
 
@@
 \subsection{Kernel ridge regression}
-The object KernelRidge of the sklearn module kernel\_ridge has been used.
+The object KernelRidge of the \href{https://scikit-learn.org/stable/}{sklearn} module kernel\_ridge has been used.
 
 \subsection{Kernel support vector regression}
-The object SVR of the sklearn module svm has been used by specifying rbf as the kernel parameter.
+The object SVR of the \href{https://scikit-learn.org/stable/}{sklearn} module svm has been used by specifying rbf as the kernel parameter.
 
@@
 \subsection{Linear quantile regression}
-The implementation of statsmodels.regression.quantile\_regression.QuantReg has been used.
+The implementation of \href{https://www.statsmodels.org/dev/generated/statsmodels.regression.quantile_regression.QuantReg.html}{quantile\_regression.QuantReg} from the regression module of statsmodels has been used.
 The model is fitted through iterative reweighted least squares.
 \subsection{Quantile gradient boosting machine}
-The implementation of sklearn.ensemble.GradientBoostingRegressor has been used.
+The implementation of \href{https://scikit-learn.org/stable/modules/generated/sklearn.ensemble.GradientBoostingRegressor.html}{GradientBoostingRegressor} from the sklearn.ensemble submodule has been used.
 \subsection{Quantile forest}
-The implementation of quantile\_forest.RandomForestQuantileRegressor has been used. This estimator is compatible with scikit-learn.
+The implementation of \href{https://pypi.org/project/quantile-forest/}{quantile\_forest.RandomForestQuantileRegressor} has been used.
+This estimator is compatible with scikit-learn API \cite{Johnson2024}.
 \subsection{Kernel quantile regression}
@@
 First this library assumes the quadratic term of the optimisation problem to be multiplied by the 0.5 factor, thus we just have to provide the Q matrix with no 0.5 in front.
-Secondly, in order to specify multiple inequalities we have to stack them an provide as a unique matrix.
+Secondly, in order to specify multiple inequalities we have to stack them and provide them as a unique matrix.
 \\
@@
 \\
-In the predict method, we pass a matrix X\_eval of independent variables, next we compute the kernel matrix between X\_train and X\_eval and obtain y\_eval with the formula $y=\alpha^\intercal K+b$.
+In the predict method, we pass a matrix $X\_eval$ of independent variables, next we compute the kernel matrix between $X\_train$ and $X\_eval$ and obtain $y\_eval$ with the formula $y=\alpha^\intercal K+b$.
 \\
 This estimator is compatible with useful scikit-learn in built methods like gridsearch, crossvalidation and scoring rules.
-Moreover, this code is compatible with sklearn in built kernel functions \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.metrics.pairwise}{sklearn.metrics.pairwise} and \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.gaussian_process.kernels}{sklearn.gaussian\_process.kernels}, functions supported by our kernel quantile regression include: gaussian rbf, laplacian, mater 1/2, matern 3/2, matern 5/2, linear, cosine, sigmoid, periodic, polynomial and custom composition of kernels.
+Moreover, this code is compatible with sklearn in built kernel functions \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.metrics.pairwise}{sklearn.metrics.pairwise} and \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.gaussian_process.kernels}{sklearn.gaussian\_process.kernels}, functions supported by our kernel quantile regression include: gaussian rbf, absolute laplacian, mater 1/2, matern 3/2, matern 5/2, linear, cosine, sigmoid, periodic, polynomial and custom composition of kernels.
 % NOTICE, these kernel implementations differ in the form of passed parameters in the former we pass in the numerator gamma while in the latter we pass l, which stands for length scale in the denominator.
@@
 \\
-Up to now, there are only two open source implementation of the quantile kernel regression.
-Nevertheless they are both in R, that is there exists no python, matlab or julia open source implementation.
-Following are reported the results of a comparative study between our own implementation and the one of the R library kernlab
+Up to now, there are only two open source implementation of the quantile kernel regression. Nevertheless they are both in R, that is there exists no python, matlab or julia open source implementation.
+
+% Following are reported the results of a comparative study between our own implementation and the one of the R library kernlab
+
 % (i guess it is in C or C++ and then binded to R).

### 2024-05-16 thesis/input_files/metrics.tex
May 15th 2024

@@
 Consider the time series with actual values given by $y=(y_{n+1}, y_{n+2},\dots, y_{n+h})$
-and its h step ahead point forecast $\hat{y}=(\hat{y}_{n+1}, \hat{y}_{n+2},\dots, \hat{y}_{n+h})$ the mean absolute error (MAE) is defined as
+and its $h$ step ahead point forecast $\hat{y}=(\hat{y}_{n+1}, \hat{y}_{n+2},\dots, \hat{y}_{n+h})$ the mean absolute error (MAE) is defined as
 \begin{definition}
@@
 \begin{definition}
-    $\mathrm{RMSE(y, \hat{y})}=\frac{1}{\sqrt{h}}\|y-\hat{y}\|_{2}=\sqrt{\frac{\sum\limits_{k=1}^{h}(y_{n+k}- \hat{y}_{n+k})^2}{h}}$
+    $\mathrm{RMSE}(y, \hat{y})=\frac{1}{\sqrt{h}}\|y-\hat{y}\|_{2}=\sqrt{\frac{\sum\limits_{k=1}^{h}(y_{n+k}- \hat{y}_{n+k})^2}{h}}$
 \end{definition}
@@
 \begin{definition}
-    $\mathrm{MASE}(y,\hat{L})=\frac{1}{h}\sum\limits_{k=1}^N\frac{|y_{n+k}-\hat{y}_{n+k}|}{\frac{1}{h-1}\sum\limits_{k=2}^{h}|y_{k}-y_{k-1}|}$
+    $\mathrm{MASE}(y,\hat{y})=\frac{1}{h}\sum\limits_{k=1}^N\frac{|y_{n+k}-\hat{y}_{n+k}|}{\frac{1}{h-1}\sum\limits_{k=2}^{h}|y_{k}-y_{k-1}|}$
 \end{definition}
@@
 \begin{cases}
-(q-1)(y_{t}-\hat{y}_{t,q}) & y_t > \hat{y}_{t,q} \\
+(q-1)(\hat{y}_{t,q}-y_{t}) & y_t > \hat{y}_{t,q} \\
 q(\hat{y}_{t,q}-y_t) & y_t \leq \hat{y}_{t,q}
@@
 \end{definition}
-Where delta is the prediction interval (PI) width, that is $\delta=b_t-a_t$ where $u$ is the prediction interval upper threshold and $l$ is the prediction interval lower threshold.
+Where delta is the prediction interval (PI) width, that is $\delta=u_t-l_t$ where $u_t$ is the prediction interval upper threshold and $l_t$ is the prediction interval lower threshold at time step $t$.
 This score penalizes observations falling outside the prediction interval and rewards narrow prediction intervals.
@@
 \begin{definition}\label{def_crps}
-    $\mathrm{CRPS}(y, \hat{F})=\int\limits_{-\infty}^{\infty}\left(\hat{F}(x)-\mathbb{1}(x-y) \right)^2 dx$
+    $\mathrm{CRPS}(y, \hat{F})=\int\limits_{-\infty}^{\infty}\left(\hat{F}(x)-\mathbb{I}_{\{x-y\}} \right)^2 dx$
 \end{definition}
@@
 Where the indicator function is defined as 
-    $\mathbb{1}(z)=
+    $\mathbb{I}_{\{z\}}=
 \begin{cases}
@@
 \end{lemma}
-Notice that, in our case, equation \ref{def_crps}, the distribution $G$ of $Y_1$ and $Y_2$ is degenerate, with all probability mass on a single point.
+Notice that, in our case, equation \ref{def_crps}, the distribution $G$ of $Y_1$ and $Y_2$ is degenerate, with all probability mass on a single point $y$.
 %  (x notation will need to be cleaned here). Since $G(t)=:$
@@
 \begin{equation}
-    \int\limits_{-\infty}^{\infty}\left(\hat{F}(x)-\mathbb{1}(x-L) \right)^2 dx=\mathbb{E}(|X_1-y|)-\frac{1}{2}\mathbb{E}(|X_1-X_2|)
+    \int\limits_{-\infty}^{\infty}\left(\hat{F}(x)-\mathbb{I}_{\{x-y\}} \right)^2 dx=\mathbb{E}(|X_1-y|)-\frac{1}{2}\mathbb{E}(|X_1-X_2|)
 \end{equation}
@@
 \section{Probability integral transform}
-The probability integral transform (PIT) is a method to assess visually the quality of probabilistic a forecasts.
+The probability integral transform (PIT) is a method to assess visually the quality of a probabilistic forecast.
 PIT is obtained by applying the predicted cumulative density function $\hat{F}$ to your data; if applying such CDF to the data results in a uniform distributed PIT, then $\hat{F}$ is a valid prediction.
@@
     \includegraphics[width=\textwidth]{images/pit.png}
-    \caption{PIT types \cite{haben2023core}}
+    \caption{Probability integral transform types \cite{haben2023core}}
     \label{fig:pit}

### 2024-05-16 thesis/input_files/paper1.tex
May 15th 2024

@@
 This is because of the kernel trick.
+\\
 To illustrate the idea behind the kernel trick consider the following example.
+Suppose $x, y \in \mathbb{R}^2$ and assume $\phi(x)=(x_{1}^{2},x_{2}^{2},\sqrt{2}x_{1}x_{2})$, then the inner product in the feature space is $x_{1}^{2}y_{1}^{2},x_{2}^{2}y_{2}^{2}+2x_{1}x_{2}y_{1}y_{2}$.
+Notice that this is the same of $\langle \phi(x), \phi(y) \rangle_{\mathcal{H}}$; thus the kernel trick consists of just using $k(x,y):=(x^\intercal y)^2$.
+For instance:
 \\
-Suppose $x,  \in \mathbb{R}^2$ and assume $\phi(x)=(x_{1}^{2},x_{2}^{2},\sqrt{2}x_{1}x_{2})$, then the inner product in the feature space is $x_{1}^{2}y_{1}^{2},x_{2}^{2}y_{2}^{2}+2x_{1}x_{2}y_{1}y_{2}$.
-Notice that this is the same of $\langle \phi(x), \phi(y) \rangle$; thus the kernel trick consists of just using $k(x,y)=:(x^\intercal y)^2$.
-For example
-\\
-$\langle \phi([2,4]),\phi([8,9]) \rangle=  \left[ {\begin{array}{cccc}
+$\langle \phi([2,4]),\phi([8,9]) \rangle_{\mathcal{H}}=  \left[ {\begin{array}{cccc}
     4 & 16 &     \sqrt{2}\cdot 2 \cdot 4\\
@@
 While $k([2,4], [8,9])=(16+36)^2=2704$
-
+We can see that the latter calculation is much quicker and compact.
 This gives a taste of how powerful kernels are thanks to the kernel trick.
@@
 \begin{definition}
-    A sequence $\{v_n\}_{n=1}^{\infty}$ is convergent if and only if there exists a point $ v \in \mathcal{V}$ such that for every $\epsilon>0$ there exists $N=N(\epsilon) \in \mathbb{N}$ such that  $\|v_n-v\|_{\mathcal{V}}$ $\forall n\geq N$.
+    A sequence $\{v_n\}_{n=1}^{\infty}$ is convergent if for every $\epsilon>0$ there exists $N=N(\epsilon) \in \mathbb{N}$ and a point $ v \in \mathcal{V}$ such that  $\|v_n-v\|_{\mathcal{V}} < \epsilon$ $\forall n\geq N$.
 \end{definition}
@@
 \begin{definition}
-    A Hilbert space is a vector space $\mathcal{H}$ with an inner product $\langle \cdot, \cdot \rangle$ such that the norm defined by $\|f\|_{\mathcal{H}}=\sqrt{\langle f, f \rangle}_{\mathcal{H}}$
+    A Hilbert space is a vector space $\mathcal{H}$ with an inner product $\langle \cdot, \cdot \rangle_{\mathcal{H}}$ such that the norm defined by $\|f\|_{\mathcal{H}}=\sqrt{\langle f, f \rangle}_{\mathcal{H}}$
 turns $\mathcal{H}$ into a complete metric space.
@@
 \begin{theorem}[Characterisation of kernels]    
-    A function $k:\mathcal{X} \times \mathcal{X} \to \mathbb{R}$ which is either continous or has finite domain can be decomposed as $k(x,y)=\langle \phi(x), \phi(y) \rangle_{\mathcal{H}}$ lie if and only if $k$ satisfy the positive semidefinite property.
+    A function $k:\mathcal{X} \times \mathcal{X} \to \mathbb{R}$ which is either continous or has finite domain can be decomposed as $k(x,y)=\langle \phi(x), \phi(y) \rangle_{\mathcal{H}}$ if and only if $k$ satisfy the positive semidefinite property.
 \end{theorem}
@@
 \\
-$\mathcal{F} =\{ \sum\limits_{i=1}^n \alpha_i k(x_i, \cdot): n \in \mathbb{N}, x_i \in \mathcal{X}, a_i \in \mathbb{R}, i=1,\cdots, n\}$.
+$\mathcal{F} =\{ \sum\limits_{i=1}^n \alpha_i k(x_i, \cdot): n \in \mathbb{N}, x_i, y_i \in \mathcal{X}, a_i \in \mathbb{R}, i=1,\cdots, n\}$.
 This space is clearly closed under addition of functions and scalar multiplication, therefore it is a vector space.
-Letting $f(x)=\sum\limits_{i=1}^n \alpha_i k_i(x_i,x)$ and $g(x)=\sum\limits_{i=1}^m \beta_i k_i(z_i,x)$,  we can introduce an inner product on $\mathcal{F}$ as 
+Letting $f(x)=\sum\limits_{i=1}^n \alpha_i k_i(x_i,x)$ and $g(x)=\sum\limits_{i=1}^m \beta_i k_i(y_i,x)$,  we can introduce an inner product on $\mathcal{F}$ as 
 \\
-$\langle f,g \rangle_{\mathcal{F}}=:\sum\limits_{i=1}^n\sum\limits_{j=1}^m a_i \beta_j k_i(x_i,z_j)=\sum\limits_{i=1}^n a_i g(x_i)= \sum\limits_{j=1}^m \beta_j f(z_j)$
+$\langle f,g \rangle_{\mathcal{F}}=:\sum\limits_{i=1}^n\sum\limits_{j=1}^m a_i \beta_j k_i(x_i,y_j)=\sum\limits_{i=1}^n a_i g(x_i)= \sum\limits_{j=1}^m \beta_j f(y_j)$
 \\
@@
 
-The analysis of above applies to any kernel, that is given a kernel function we can always construct its RKHS.
+The analysis of above applies to any kernel, that is given a valid kernel function we can always construct its RKHS in the way we just illustrated.
 \begin{theorem}
@@
         =& \left\langle \sum\limits_{i=1}^n a_i k(\cdot, x_i), \sum\limits_{j=1}^n a_j k(\cdot, x_j) \right\rangle_{\mathcal{H}} \\
-        =& \left\| \sum\limits_{i=1}^n a_i k(\cdot, x_i)\right\|_{\mathcal{H}}^2 \geq 0
+        =& \left\| \sum\limits_{i=1}^n a_i k(x_i, x_i)\right\|_{\mathcal{H}}^2 \geq 0
     \end{align}
@@
 \begin{proof}
-    It is known that null space of $L$, call it $M$, is a closed linear subspace in $\mathcal{H}$.
+    It is known that the null space of $L$, call it $M$, is a closed linear subspace in $\mathcal{H}$.
 This implies that the Hilbert space direct sum $M \bigoplus M^\bot$ is an isomorphism of $\mathcal{H}$.
-Letting $L$ be different that the zero operator we have that $M^\bot \neq {0}$.
+When $L(f) \neq 0$ we have that $L(f)=L(m +w)$ with $m \in M$ and $w \in M^\bot$, it follows that $L(f)=0+L(w) \neq 0 \implies$ $M^\bot \neq {0}$.
     \\
-    Choose now a $z \in M^\bot$ with norm 1.
+    Using this fact, we can now choose a $z \in M^\bot$ with norm 1.
 By linearity of $L$, for any $f \in \mathcal{H}$ we have $L(zL(f)- fL(z))=L(z)L(f)- L(f)L(z)=0$.
@@
     \begin{align}
-        \langle h_0,f \rangle=&\langle h_1, f \rangle \\
-        \langle h_0,f \rangle-\langle h_1, f \rangle=& 0 \\
-        \langle h_0 -h_1,f \rangle=& 0
+        \langle h_0,f \rangle_{\mathcal{H}}=&\langle h_1, f \rangle_{\mathcal{H}} \\
+        \langle h_0,f \rangle_{\mathcal{H}}-\langle h_1, f \rangle_{\mathcal{H}}=& 0 \\
+        \langle h_0 -h_1,f \rangle_{\mathcal{H}}=& 0
     \end{align}
@@
 \begin{proposition}
-For each $x \in \mathcal{X}$ there exists a function $k_{x} \in \mathcal{H}$ such that the evaluation functional $F_{x}(f)=\langle k_{x}, f\rangle_{\mathcal{H}}=f(x)$    
+For each $x \in \mathcal{X}$ there exists a function $k(\cdot, x) \in \mathcal{H}$ such that the evaluation functional $F_{x}(f)=\langle k(\cdot, x), f\rangle_{\mathcal{H}}=f(x)$    
 \end{proposition}
 
-The function $k_{x}$ is the reproducing kernel evaluated at point $x$.
+The function $k(\cdot, x)$ is the reproducing kernel evaluated at point $x$.
 Furthermore, note that $k_{x}$ is itself a function lying in it.
@@
 \begin{align*}
-    k_{x}(y)=F_{y}(k_{x})=\langle k_{x}, k_{y}\rangle_{\mathcal{H}}=\langle \phi(x), \phi(y)\rangle_{\mathcal{H}}
+    k(y, x)=F_{y}(k(\cdot, x))=\langle k(\cdot, x), k(\cdot, y)\rangle_{\mathcal{H}}=\langle \phi(x), \phi(y)\rangle_{\mathcal{H}}
 \end{align*}
@@
 \begin{definition}
-    Considering a set of vectors $x_1, \cdots, x_n$, the Gram matrix is defined as the $n\times n$ matrix whose entries are $\langle x_i, x_j \rangle$.    
+    Considering a set of vectors $x_1, \cdots, x_n$, the Gram matrix is defined as the $n\times n$ matrix whose entries are $\langle x_i, x_j \rangle_{\mathcal{X}}$.    
 \end{definition}
@@
     \begin{align}
-        z^\intercal K z =& \sum_{i,j=1}^n z_i z_j K_{i,j}
+        z^\intercal K z =& \sum_{i,j=1}^n z_i z_j k(x_i,x_j)
         \\
@@
 \end{proof}
-We know that positive semidefinite matrices form a cone in the vector space $n\times n$ matrices.
+We know that positive semidefinite matrices form a cone in the vector space of $n\times n$ matrices.
 Because of this we have that such matrices are convex, this implies that all the nice theory of convex analysis can be used when optimising over them.
@@
     \begin{align}
-        k(x,z)=&k_1(x,z)+k_2(x,Y)
+        k(x,y)=&k_1(x,y)+k_2(x,y)
         \\
-        k(x,z)=& ak_1(x,z)
+        k(x,y)=& ak_1(x,y)
         \\
-        k(x,z)=& k_1(x,z)k_2(x,z)
+        k(x,y)=& k_1(x,y)k_2(x,y)
         \\
-        k(x,z)=& k_3(\phi(x),\phi(z))
+        k(x,y)=& k_3(\phi(x),\phi(y))
     \end{align}.
@@
     \begin{align}
-    z^\intercal (K_1+K_2) z= z^\intercal K_1 z + z^\intercal K_1 z \geq 0
+    z^\intercal (K_1+K_2) z= z^\intercal K_1 z + z^\intercal K_2 z \geq 0
     \\
@@
 Thus, $K$ is positive semidefinite.
-Next, the matrix given by $k_1 \cdot k_2$ corresponds to the Schur product of $K_1$ and $K_2$, denote if $H$.
+Next, the matrix given by $k_1 \cdot k_2$ corresponds to the Schur product of $K_1$ and $K_2$, denote it $H$.
 Notice, $H$ is a principal submatrix of $K$, hence for any $z \in \mathbb{R}^n$ there exist a $t \in \mathbb{R}^{n^2}$ such that $z^\intercal H z= t^\intercal K t$.
@@
 \\
-$k(x,y)=\exp \left(-{\frac {{\sqrt {2p+1}}\|x-y|\|_{2}}{l }}\right){\frac {p!}{(2p)!}}\sum _{i=0}^{p}{\frac {(p+i)!}{i!(p-i)!}}\left({\frac {2{\sqrt {2p+1}}\|x-y|\|_{2}}{l }}\right)^{p-i}$
+$k(x,y)=\exp \left(-{\frac {{\sqrt {2p+1}}\|x-y|\|_{2}}{l }}\right){\frac {p!}{(2p)!}}\sum\limits_{i=0}^{p}{\frac {(p+i)!}{i!(p-i)!}}\left({\frac {2{\sqrt {2p+1}}\|x-y|\|_{2}}{l }}\right)^{p-i}$
 \\
@@
 
-Any algorithm involving inner products between inputs can be combined with a kernel methods.
+Any algorithm involving inner products between inputs can be combined with a kernel method.
 The chosen kernel function will calculate the inner product between the images of two inputs in the feature space.

### 2024-05-16 thesis/input_files/probabilistic.tex
May 15th 2024

@@
-When it comes to probabilistic forecasting, there are a couple of standard practical approaches; conceptually they can be grouped into two main categories.
-The former class of approaches tries modelling the distribution of the observed data directly.
-The latter family of approaches constructs first a point forecast and then learns the distribution of the model's errors.
+We saw in the previous chapter how point forecasting relates to predicting the expected value of the target variable.
+Conversely, in probabilistic forecasting we are interested in modelling a probability distribution over the target variable.
+This approach enables us to account for prediction uncertainty and to make more informed decision.
 \section{Quantile regression}
@@
 
-% \section{Copula models}
 \section{Quantile forest}
@@
 \\
-In order to understand this algorithm, it is first necessary to first cover the theory of decision trees and random forests.
+In order to understand this algorithm, it is first necessary to cover the theory of decision trees and random forests.
 \subsection{Decision trees}
@@
 Our goal is selecting the best features and best split point such that to achieve the best generalizing fit.
-For an example/visualisation consider the pictures \ref{fig:elements_statistical_learning1} and \ref{fig:elements_statistical_learning2} taken from \cite{hastie2009elements}, where the decision tree algorithm is visualised for a regression problem with two independent variables $X_2$ and $X_1$.
+For a visualisation consider the pictures \ref{fig:elements_statistical_learning1} and \ref{fig:elements_statistical_learning2} taken from \cite{hastie2009elements}, where the decision tree algorithm is visualised for a regression problem with two independent variables $X_2$ and $X_1$.
 \begin{figure}
     \includegraphics[width=0.5\textwidth]{images/elsii1.png}
-    \caption{two-dimensional feature space partitioned by recursive binary splitting}
+    \caption{two-dimensional feature space partitioned by recursive binary splitting \cite{hastie2009elements}}
     \label{fig:elements_statistical_learning1}
@@
 \includegraphics[width=1\textwidth]{images/elsii2.png}
-\caption{partition tree and regression model}
+\caption{partition tree and regression model \cite{hastie2009elements}}
 \label{fig:elements_statistical_learning2}
@@
 \\
-Suppose to partition the feature space into M regions $R_1,\dots,R_M$, then the model reads as follows.
+Suppose now to partition the feature space into M regions $R_1,\dots,R_M$, then the model reads as follows.
 \begin{equation}
@@
 \begin{equation}
-    R_1(j,s)=\{X|X_j\leq s\} \quad R_2(j,s)=\{X|X_j\>s\}
+    R_1(j,s)=\{X|X_j\leq s\} \quad R_2(j,s)=\{X|X_j >s\}
 \end{equation}
@@
 \end{equation}
-The inner problem is easy, as already pointed out we will have 
+The inner problem is easy, as already pointed out, we will have 
 \begin{equation}
@@
 The idea behind boostrapping is to randomly sample from the training set with replacement B times and then fitting B models to each of the "artificial" datasets.
-Boostrapping can serve different tasks; we can use it to assess the accuracy of a parameter estimate or of a prediction but also to improve their estimates.
+Boostrapping can serve different tasks; we can use it to assess the accuracy of a parameter %estimate
+or of a prediction but also to improve their estimates.
 
@@
 Bagging stands for boostrap aggregation. 
-The bagging estimate is dafined by $\mathbb{E}_P[\hat{f}^*]$ where $P$ is the empirical distribution putting equal probability on each data point of the training set.
+The bagging estimate is defined by $\mathbb{E}_P[\hat{f}^*]$ where $P$ is the empirical distribution putting equal probability on each data point of the training set.
 Basically, for each bootstrap fitted model $\hat{f}^{*b}(x)$, we compute the bagging estimate by
@@
 \end{equation}
-Bagging is particularly useful in reducing the variance of decision trees, resulting in an improved prediction(tradeoff bias variance).
+Bagging is particularly useful in reducing the variance of decision trees, resulting in an improved prediction (bias-variance tradeoff).
 Note, this improvement comes from the fact that averaging reduces variance and leaves biases unchanged.
@@
 The second term disappears as B increases, while the first term depends heavily on the correlation between bagged trees.
-Random forest consists in is reducing the correlation between the trees by randomly selecting m of the p features as candidates for splitting; m tipically takes a value in the order of $\sqrt{p}$ or even 1 with default value $m=\left\lfloor \frac{p}{3} \right\rfloor$, while a good minimum node size is around five.
+Random forest consists in reducing the correlation between the trees by randomly selecting m of the p features as candidates for splitting; m tipically takes a value in the order of $\sqrt{p}$ or even 1 with default value $m=\left\lfloor \frac{p}{3} \right\rfloor$, while a good minimum node size is around five.
 \\
@@
 \subsection{Quantile forest}
-The key observation here is noting that random forests approximates the conditional mean $\mathbb{E}(Y|X=x)$ by a weighted average over the observed $Y$.
+The key observation here is noting that random forests approximates the conditional mean $\mathbb{E}(Y|X=x)$ by a weighted average over the observed $y$.
+% expected value is the weighted averaged of realizations with weights their respective probabilities.
 Hence, we can extend this idea to the full conditional distribution by
@@
 \subsection{Boosted trees}
-We combine several trees obtaining the boosted tree model
+Combining several trees $T(\cdot, \Theta_m)$, we obtain the boosted tree model
 \begin{equation}
@@
 \begin{equation}\label{boosting_minimisation_problem}
-    \hat{\Theta}_m=\argmin{\Theta_m}\sum\limits_{i=1}^{N}L(y_i, f_{m-1}(x_i)+T(x_i;\Theta_m))
+    \hat{\Theta}_m=\argmin{\Theta_m}\sum\limits_{i=1}^{n}L(y_i, f_{m-1}(x_i)+T(x_i;\Theta_m))
 \end{equation}
@@
 \begin{equation}
-    \tilde{\Theta}_m=\argmin{\Theta}\sum\limits_{i=1}^{N}\left(-g_{im}-T(x_i;\Theta)\right)^2
+    \hat{\Theta}_m=\argmin{\Theta_m}\sum\limits_{i=1}^{N}\left(-g_{im}-T(x_i;\Theta)\right)^2
 \end{equation}
-where $g_{im}$ is the gradient of $L(f)=\sum\limits_{i=1}^N L(y_i, f(x_i))$ evaluated at $f_{m-1}$.
+where $g_{im}$ is the gradient of $L(f)=\sum\limits_{i=1}^n L(y_i, f(x_i))$ evaluated at $f_{m-1}$.
 Put simply, we are fitting the m-th tree to the negative of the gradient values of f through least squares.
@@
 
-\section{Kernel methods}
-\subsection{Kernel quantile regression}
+% \section{Kernel methods}
+\section{Kernel quantile regression}
 The idea of quantile regression has been extended to kernel methods by Takeuchi et al. \cite{takeuchi2006nonparametric}.
-There, they minimize a risk plus regularizer defined as follows.
+There, they minimize a risk functional plus regularizer defined as follows.
 \begin{equation}\label{eq:kqr_min1}
@@
 \end{equation}
-where $f=g+b$, $g \in \mathcal{H}$ and $b \in R$.
-Using the link between RKHS and feature spaces, we can rewrite $f(x)=\langle w, \phi(x) \rangle+b$.
-Moreover, note that the RKHS norm is defined as follows $\|f\|_{\mathcal{H}}=\inf\{\|w\|_{F}:w\in F,f(x)=\langle w,\varphi (x)\rangle _{F},\forall x\in X\}$, where $F$ is the feature space.
+where $f=g+b$, $g \in \mathcal{H}$ and $b \in \mathbb{R}$.
+Using the link between RKHS and feature spaces, we can rewrite $f(x)=\langle w, \phi(x) \rangle_{\mathcal{H}}+b$.
+Moreover, note that the RKHS norm is defined as follows $\|f\|_{\mathcal{H}}=\inf\{\|w\|_{\mathcal{F}}:w\in \mathcal{F},f(x)=\langle w,\varphi (x)\rangle _{\mathcal{F}},\forall x\in X\}$, where $\mathcal{F}$ is the feature space.
  Doing so we obtain a minimization problem equivalent to minimizing equation \ref{eq:kqr_min1}.
@@
     \min_{w,b} \quad & C \sum \limits_{i=1}^{m}
-    q(y_i-\langle w, \phi(x) \rangle-b)+ (1-q)(-y_i+\langle w, \phi(x) \rangle+b)+ \frac{1}{2}\|w\|^2\\
+    q(y_i-\langle w, \phi(x) \rangle_{\mathcal{H}}-b)+ (1-q)(-y_i+\langle w, \phi(x) \rangle_{\mathcal{H}}+b)+ \frac{1}{2}\|w\|_{\mathcal{H}}^2\\
     \end{aligned}
@@
         \min_{w,b,\xi_i,\xi_i^*} \quad & C \sum \limits_{i=1}^{m}
-        q \xi_i+ (1-q)\xi_i^*+ \frac{1}{2}\|w\|^2\\
-    \textrm{s.t.} \quad & y_i-\langle w, \phi(x) \rangle-b \leq \xi_i\\
-    & -y_i+\langle w, \phi(x) \rangle+b \leq \xi_i^*\\
+        q \xi_i+ (1-q)\xi_i^*+ \frac{1}{2}\|w\|_{\mathcal{H}}^2\\
+    \textrm{s.t.} \quad & y_i-\langle w, \phi(x) \rangle_{\mathcal{H}}-b \leq \xi_i\\
+    & -y_i+\langle w, \phi(x) \rangle_{\mathcal{H}}+b \leq \xi_i^*\\
       &\xi_i\geq0    \\
@@
     \end{equation}
-Consider now, the lagrangian $L$ associated to \ref{eq:kqr_min4}
+Consider now, the lagrangian $\mathcal{L}$ associated to \ref{eq:kqr_min4}
 \begin{equation}\label{eq:kqr_min5}
@@
 \begin{definition}
-    The dual function associated to the lagrangian $L(x,\lambda)$ is given by $g(v)=\inf_x L(x,\lambda)$
+    The dual function associated to the lagrangian $\mathcal{L}(x,\lambda)$ is given by $g(v)=\inf_x \mathcal{L}(x,\lambda)$
 \end{definition}
@@
 where $p^*$ is the optimal value of your optimization problem.
-Consider a simple lagrangian $L(x,\lambda)=f(x)+\sum \lambda_i r_i(x) +\sum v_i h_i(x)$, where $r_i(x)$ are inequality constraints while $h_i(x)$ are equality constraints of the problem.
+Consider a simple lagrangian $\mathcal{L}(x,\lambda)=f(x)+\sum\limits_{i=1}^n \lambda_i r_i(x) +\sum\limits_{i=1}^{n} v_i h_i(x)$, where $r_i(x)$ are inequality constraints while $h_i(x)$ are equality constraints of the problem.
 Then it can be noted that, the lower bound on $p^*$ is non trivial only when the lagrange multiplier $\lambda \succeq 0$.
@@
 \end{equation}
-In our case, equation \ref{eq:kqr_min5}, we have $w^\intercal w$, thus $Q$ is just the identity matrix which satisfies the positive definiteness requirement. Therefore, our problem is convex.
+% In our case, equation \ref{eq:kqr_min5}, we have $w^\intercal w$, thus $Q$ is just the identity matrix which satisfies the positive definiteness requirement. 
+This fact is easily checked by the fact that kernel matrices are by definition positive semidefinite.
+% $\|w\|_{\mathcal{H}}^2= w^\intercal w= \alpha^\intercal K \alpha$, the last element is positive semidefinitive which in turn implies also $w^\intercal w$ is also positive semidefinite.
+Therefore, our problem is convex.
 Next we check that Slater's condition holds.
@@
     \begin{aligned}
-        g(\alpha, \alpha^*, \nu, \nu^*)=& \inf_x L(x,\lambda)\\
-    = & \inf_{\xi, \xi^*, w, b} L(w,b,\xi,\xi^*)
+        g(\alpha, \alpha^*, \nu, \nu^*)=& \inf_x \mathcal{L}(x,\lambda)\\
+    = & \inf_{\xi, \xi^*, w, b} \mathcal{L}(w,b,\xi,\xi^*)
 \end{aligned}
@@
 \begin{equation}
-    f(x)=\sum\limits_i \textrm{a}_i k(x_i, x)+b
+    f(x)=\sum\limits_{i=1}^{n} \textrm{a}_i k(x_i, x)+b
 \end{equation}
@@
     \begin{aligned}
-    \lambda^* r_i(x^*)=&0, \quad i=1,\dots,m
+    \lambda^* r_i(x^*)=&0, \quad i=1,\dots,n
     \\
@@
     \begin{aligned}
-    b=&y_i-\sum_i \textrm{a}_i k(x_i,x)
+    b=&y_i-\sum\limits_{i=1}^n \textrm{a}_i k(x_i,x)
     \end{aligned}
@@
 Hereafter, the results of the four presented methods on the Melbourne dataset are reported, see figure \ref{fig:melbourne_quantiles_comparison} for a visualisation.
-Notice, the kernel considered here is the gaussian rbf.
-Hyperparameter tuning have been optimised through cross validation, see \ref{appendix:cross_validation}.
+Notice, the kernel considered here is the Gaussian RBF.
+Hyperparameters have been tuned through cross validation, see \ref{appendix:cross_validation}.
 \begin{figure}[!h]
@@
       \includegraphics[width=1.1\textwidth]{images/melbourne_kernel_quantile_regression.png}
-      \caption{Kernel quantile regression} 
+      \caption{Kernel quantile regression, Gaussian RBF} 
       \label{fig:melborune_kernel_quantile_regression} 
     \end{subfigure} 
-    \caption{Quantile regressors}
+    \caption{Quantile regressors comparison Melbourne weather data}
     \label{fig:melbourne_quantiles_comparison} 
@@
     \midrule
-    Pinball loss & 11.278895 & 10.317612 & 10.340842 & 10.031708 \\
+    Pinball loss & 11.278895 & 10.317612 & 10.340842 & \textbf{10.031708} \\
     \bottomrule
@@
         \midrule
-        & Gauss RBF &  10.031708 \\
-        & Laplacian &  10.056884 \\
-        & Gauss RBFxLaplacian & 10.150826 \\
+        & Gaussian RBF &  10.031708 \\
+        & Absolute Laplacian &  10.056884 \\
+        & Gauss RBFx Absolute Laplacian & 10.150826 \\
         & Cosine & 16.253973    \\
@@
     \end{subfigure} 
-    \caption{Kernel quantile regressors comparison}
+    \caption{Kernel quantile regressors comparison Melbourne data}
     \label{fig:kernel quantile regressors comparison} 

### 2024-05-16 thesis/input_files/qr.tex
May 15th 2024

@@
     \begin{equation}
-        Q(q)=\inf\{y:F(y)\geq q\}
+        Q_q=\inf\{y:F(y)\geq q\}
     \end{equation}
 \end{definition}
-Alternatively, in order  to ease the posing of the quantile regression problem, we can formulate quantiles as the solutions to simple optimization problems.
+Alternatively, in order  to ease the posing of the quantile regression problem, we can formulate quantiles as the solutions to the following optimization problems.
 \\
-For any $0<q<1$ consider the pinball loss function from section \ref{pinball}, $\rho_q(u)=u(q-\mathbb{1}_{\{u<0\}})$. 
+For any $0<q<1$ consider the pinball loss function from section \ref{pinball}, $\rho_q(u)=u(q-\mathbb{I}_{\{u<0\}})$. 
 Such loss is minimized by the quantile $Q(q)$.
-Thus,we can estimate quantiles by minimizing the expectation of $\rho_q(Y-g(x,\beta))$ with respect to the parameter $\beta$.
+Thus,we can estimate quantiles by minimizing the expectation of $\rho_q(y-g(x,\beta))$ with respect to the parameter $\beta$.
 \\
@@
 \\
-It follows that the conditional linear quantile function $Q_Y(q|X=x)=x_i\beta(q)$, can be estimated by solving
+It follows that the conditional linear quantile function $\hat{Q}_q=x_i\beta(q)$, can be estimated by solving
 \begin{equation}\label{eq:linear_quantile_regression_minimisation_problem}
-    \hat{\beta}(q)=\argmin{\beta}\sum \rho_q   (y_i-x \beta)
+    \hat{\beta}(q)=\argmin{\beta}\sum\limits_{i=1}^n \rho_q   (y_i-x_i \beta)
 \end{equation}

### 2024-05-16 thesis/input_files/data_exploratory.tex
May 16th 2024

@@
 This section covers the datasets used in our experiments, in addition, attributes and features will be thoroughly described.
-Next, we will explain the extract transform load pipeline (ETL) that we set up in order to ease the workflow of our comparison studies.
+Next, we will explain the extract transform load pipeline (ETL) that we set up in order to ease the workflow of our studies.
 Finally, in order to to better understand patters within the data, we carry out an exploratory data analysis.
@@
     \includegraphics[width=\textwidth]{images/price_track_fig1.png}
-    \caption{Price track}
+    \caption{Price track GEFCom2014}
     \label{fig:price_track_fig1}
@@
     \includegraphics[width=\textwidth]{images/gefcom_zonal_price_vs_zonal_load.png}
-    \caption{Zonal price versus zonal load}
+    \caption{Zonal price versus zonal load GEFCom2014}
     \label{fig:gefcom_zonal_price_vs_zonal_load}
@@
     \includegraphics[width=\textwidth]{images/gefcom_zonal_price_vs_total_load.png}
-    \caption{Zonal price versus total load}
+    \caption{Zonal price versus total load GEFCom2014}
     \label{fig:gefcom_zonal_price_vs_total_load}
@@
     \includegraphics[width=\textwidth]{images/gefcom_load_historical.png}
-    \caption{Historical load}
+    \caption{Historical load GEFCom2014}
     \label{fig:gefcom_load_historical}
@@
     \includegraphics[width=\textwidth]{images/gefcom_w_avg_historical.png}
-    \caption{Historical weather}
+    \caption{Historical weather GEFCom2014}
     \label{fig:gefcom_w_avg_historical}
@@
     \includegraphics[width=\textwidth]{images/gefcom_load_vs_w_avg.png}
-    \caption{Load versus weather}
+    \caption{Load versus weather GEFCom2014}
     \label{fig:gefcom_load_vs_w_avg}
@@
 
-\subsection{GEFCom ETL}
+% \subsection{GEFCom ETL}
+\subsubsection{GEFCom ETL}
 The provided data is raw, hence the need for a standardized extract transform and load pipeline. 
@@
 For this study we considered the data of \href{https://zenodo.org/records/7907883}{SECURES-Met}.
-This data consists of various metereological attributes relevant for modelling electricity demand.
+This data consists of various metereological attributes relevant for modelling electricity demand \cite{Formayer2023}.
 The dataset is made up of historical data until end of 2020, while from 2021 onwards the data consists of forecasts made by a metereological model developed by the European Centre for Medium-Range Weather Forecasts (ECMWF).
@@
 Notice, the data request to the environmental variables at hourly frequency for a whole year is broken.
-Therefore, we downloaded each month separately and then transformed into a single csv for the whole year.
+Therefore, we downloaded each month separately and then transformed them into a single csv for the whole year.
 
@@
 Following are reported the EDA plots alongside with insights that can be extracted from them.
-We choose to focus the following analysis on the Swiss data only given that the plots for the German data are pretty similar.
+We choose to focus the following analysis on the Swiss data only, given that the plots for the German data are pretty similar.
 To get started we visualized the historical load data in figure \ref{fig:CH_historical_load_2021}.
@@
     \includegraphics[width=\textwidth]{images/CH_historical_load_2021.png}
-    \caption{Switzerland historical load}
+    \caption{Switzerland historical load (2021)}
     \label{fig:CH_historical_load_2021}
@@
     \includegraphics[width=\textwidth]{images/CH_load_vs_temperature_2021.png}
-    \caption{Load vs temperature Switzerland 2021}
+    \caption{Load vs temperature Switzerland (2021)}
     \label{fig:CH_load_vs_temperature_2021}
@@
     \includegraphics[width=\textwidth]{images/CH_load_vs_wind_speed_2021.png}
-    \caption{Load vs wind speed Switzerland 2021}
+    \caption{Load vs wind speed Switzerland (2021)}
     \label{fig:CH_load_vs_wind_speed_2021}
@@
     \includegraphics[width=\textwidth]{images/CH_day_of_week_boxplot_2021.png}
-    \caption{Load versus day of week boxplot}
+    \caption{Load versus day of week boxplot Switzerland (2021)}
     \label{fig:CH_day_of_week_boxplot_2021}
@@
     \includegraphics[width=\textwidth]{images/CH_is_holiday_boxplot_2021.png}
-    \caption{Load versus day type boxplot}
+    \caption{Load versus day type boxplot Switzerland (2021)}
     \label{fig:CH_is_holiday_boxplot_2021}

### 2024-05-16 thesis/input_files/implementation.tex
May 16th 2024

@@
 For the list of python packages needed see the requirement.txt file on \url{https://github.com/luca-pernigo/ThesisKernelMethods}\label{github_repo}.
+% - indicate computer specifics
+All experiments have been carried out on a 3.2 GHz 16GB Apple M1 Pro.
 
 Section documenting code
-- indicate computer specifics
 
@@
 \subsection{Kernel quantile regression}
-Kernel quantile regression had no previously implemented open source library, thus the need of implementing our own version.
+Kernel quantile regression had no previously implemented python open source library, thus the need of implementing our own version.
 \\
@@
 Therefore, the KQR class is derived from the scikit-learn BaseEstimator and the mixin class RegressorMixin.
-Our KQR class is initialized by providing a quantile, the regularization term C and the rbf kernel bandwidth gamma; the latter are the two hyperparameters of our custom estimator.
+Our KQR class is initialized by providing a quantile, the regularization term C, the kernel family and its corresponding hyperparameters.
 \\
-In the fit method, we set up and solve the convex optimisation problem by using the cvxopt library. 
+In the fit method, we set up and solve the convex optimisation problem by using the cvxopt library, see offical manual \cite{vandenberghe2010cvxopt} for a reference.
 When using this library, it is important to keep two things in mind.
@@
 To cross validate overt the same range of parameter across the different kernels we made sura that gamma=1/l.
-\\
-Up to now, there are only two open source implementation of the quantile kernel regression. Nevertheless they are both in R, that is there exists no python, matlab or julia open source implementation.
+
+% \\
+% Up to now, there are only two open source implementation of the quantile kernel regression.
+Nevertheless they are both in R, that is there exists no python, matlab or julia open source implementation.
 

### 2024-05-16 thesis/input_files/probabilistic.tex
May 16th 2024

@@
 Conversely, in probabilistic forecasting we are interested in modelling a probability distribution over the target variable.
-This approach enables us to account for prediction uncertainty and to make more informed decision.
+This approach enables us to account for prediction uncertainty and to make more informed decisions.
 \section{Quantile regression}
@@
 For the outer problem, we scan through all the (j,s) tuples and pick the best pair. 
-Next, one or both of these regions from the previous step are split into two more regions; we recurse this process until some stopping condition is triggered (max number of branches, max deth of tree, threshold on the MSE, minimum number of observation in each leaf node).
+Next, one or both of these regions from the previous step are split into two more regions; we recurse this process until some stopping condition is triggered (max number of branches, max deth of tree, threshold on the mean squared error (MSE), minimum number of observations in each leaf node).
 Notice, this being a greedy algorithm implies that our final solution is guaranteed to be just a local optimum not a global one.
@@
 \begin{equation}\label{eq:kqr_min1}
-    R[f]:=\frac{1}{m}\sum\limits_{i=1}^{m}\rho_q(y_i-f(x_i))+\frac{\lambda}{2}\|g\|_\mathcal{H}^2
+    R[f]:=\frac{1}{m}\sum\limits_{i=1}^{m}\rho_q(y_i-f(x_i))+\frac{\lambda}{2}\|w\|_\mathcal{H}^2
 \end{equation}
-where $f=g+b$, $g \in \mathcal{H}$ and $b \in \mathbb{R}$.
+where $f=w+b$, $w \in \mathcal{H}$ and $b \in \mathbb{R}$.
 Using the link between RKHS and feature spaces, we can rewrite $f(x)=\langle w, \phi(x) \rangle_{\mathcal{H}}+b$. 
-Moreover, note that the RKHS norm is defined as follows $\|f\|_{\mathcal{H}}=\inf\{\|w\|_{\mathcal{F}}:w\in \mathcal{F},f(x)=\langle w,\varphi (x)\rangle _{\mathcal{F}},\forall x\in X\}$, where $\mathcal{F}$ is the feature space.
+% Moreover, note that the RKHS norm is defined as follows $\|f\|_{\mathcal{H}}=\inf\{\|w\|_{\mathcal{F}}:w\in \mathcal{F},f(x)=\langle w,\varphi (x)\rangle _{\mathcal{F}},\forall x\in X\}$, where $\mathcal{F}$ is the feature space.
  Doing so we obtain a minimization problem equivalent to minimizing equation \ref{eq:kqr_min1}.
@@
     \min_{w,b} \quad & C \sum \limits_{i=1}^{m}
-    q(y_i-\langle w, \phi(x) \rangle_{\mathcal{H}}-b)+ (1-q)(-y_i+\langle w, \phi(x) \rangle_{\mathcal{H}}+b)+ \frac{1}{2}\|w\|_{\mathcal{H}}^2\\
+    q(y_i-\langle w, \phi(x_i) \rangle_{\mathcal{H}}-b)+ (1-q)(-y_i+\langle w, \phi(x_i) \rangle_{\mathcal{H}}+b)+ \frac{1}{2}\|w\|_{\mathcal{H}}^2\\
     \end{aligned}
     \end{equation}
-Note the division by $\lambda$ so that $C=\frac{1}{\lambda m}$.
+Note the division by $\lambda$ so that $C:=\frac{1}{\lambda m}$.
 \\
@@
         q \xi_i+ (1-q)\xi_i^*+ \frac{1}{2}\|w\|_{\mathcal{H}}^2\\
-    \textrm{s.t.} \quad & y_i-\langle w, \phi(x) \rangle_{\mathcal{H}}-b \leq \xi_i\\
-    & -y_i+\langle w, \phi(x) \rangle_{\mathcal{H}}+b \leq \xi_i^*\\
+    \textrm{s.t.} \quad & y_i-\langle w, \phi(x_i) \rangle_{\mathcal{H}}-b \leq \xi_i\\
+    & -y_i+\langle w, \phi(x_i) \rangle_{\mathcal{H}}+b \leq \xi_i^*\\
       &\xi_i\geq0    \\
@@
     \begin{aligned}
-    L(w,b,\xi,\xi^*)= & C q \xi^\intercal \mathbb{1}+ C q (\xi^*)^\intercal \mathbb{1}+ \frac{1}{2}w^\intercal w- \alpha^\intercal(\xi - y+\Phi^\intercal w +b)
-    - (\alpha^*)^\intercal(\xi^* +y-\Phi^\intercal w -b)
+    L(w,b,\xi,\xi^*)= & C q \xi^\intercal \mathbb{1}+ C q (\xi^*)^\intercal \mathbb{1}+ \frac{1}{2}w^\intercal w- \lambda^\intercal(\xi - y+\Phi^\intercal w +b)
+    - (\lambda^*)^\intercal(\xi^* +y-\Phi^\intercal w -b)
     \\
@@
 \begin{definition}
-    The dual function associated to the lagrangian $\mathcal{L}(x,\lambda)$ is given by $g(v)=\inf_x \mathcal{L}(x,\lambda)$
+    The dual function associated to the lagrangian $\mathcal{L}(x,\lambda, \nu)$ is given by $g(\lambda, \nu)=\underset{x}\inf \ \mathcal{L}(x,\lambda, \nu)$
 \end{definition}
-where $\lambda$ is called the lagrange multiplier of the optimization problem. Such dual formulation has an useful property, that is \begin{equation}\label{weak_duality}
-    g(\lambda)\leq p^*
+where $\lambda$ is called the lagrange multiplier of the optimization problem. Such dual formulation has an useful property, which is \begin{equation}\label{weak_duality}
+    g(\lambda, \nu)\leq p^*
 \end{equation}
-where $p^*$ is the optimal value of your optimization problem.
-Consider a simple lagrangian $\mathcal{L}(x,\lambda)=f(x)+\sum\limits_{i=1}^n \lambda_i r_i(x) +\sum\limits_{i=1}^{n} v_i h_i(x)$, where $r_i(x)$ are inequality constraints while $h_i(x)$ are equality constraints of the problem.
+where $p^*$ is the optimal value of your optimization problem. In other words $g(\lambda, \nu)$ is a lower bound for the optimal $p^*$.
+Consider now a simple lagrangian $\mathcal{L}(x,\lambda, \nu)=f(x)+\sum\limits_{i=1}^n \lambda_i r_i(x) +\sum\limits_{i=1}^{n} \nu_i h_i(x)$, where $r_i(x)$ are inequality constraints while $h_i(x)$ are equality constraints of the problem.
 Then it can be noted that, the lower bound on $p^*$ is non trivial only when the lagrange multiplier $\lambda \succeq 0$.
+% when lagrange multiplier is negative, we have that g > p
 Therefore, the idea is that by maximizing the dual function subject to the contraint $\lambda \succeq 0$, we can obtain an approximate or perfect solution to the primal problem.
@@
 Moreover, note that the optimal duality gap is always nonnegative.
-\\
 We say that strong duality holds, when the optimal duality gap is zero; in other words, the lagrange dual bound is tight.
@@
 Constraint qualifications are condtions under which strong duality holds; one of the most popular is Slater's condition.
+\begin{proposition}
+    Slater's condition reads as
     \begin{equation}\label{slater_condition}
@@
     \end{equation}
-Where relint $D$ is the relative interior of $ D:=\cap _{i=0}^{m}\operatorname {dom} (r_{i})$
-\\
+    Where relint $D$ is the relative interior of $ D:=\underset{i=0}{\overset{m}{\cap}} \operatorname {dom} (r_{i})$
+\end{proposition}
 Slater's theorem naturally follows.
@@
 In our case we don't have any equality constraint, so we do not have to worry about the $h_i(x)=0$ term in \ref{slater_condition}.
-All we have to check is the convexity of our problem and that there exist a $x$ such that $r_i(x)<0$.
+All we have to check is the convexity of our problem and that there exist an $x$ such that $r_i(x)<0$.
 For convexity, a sufficient condition is the positive definiteness of $Q$ in the quadratic programming problem 
@@
 % In our case, equation \ref{eq:kqr_min5}, we have $w^\intercal w$, thus $Q$ is just the identity matrix which satisfies the positive definiteness requirement. 
-This fact is easily checked by the fact that kernel matrices are by definition positive semidefinite.
+This condition is easily checked by the fact that kernel matrices are by definition positive semidefinite.
 % $\|w\|_{\mathcal{H}}^2= w^\intercal w= \alpha^\intercal K \alpha$, the last element is positive semidefinitive which in turn implies also $w^\intercal w$ is also positive semidefinite.
@@
 Considering first the two non negative constraints on $\xi$ and $\xi^*$, we conclude that $\xi$  and $\xi^*$ have to be greater or equal to zero for the existence of an $x$ satisfying Slater's condition.
-Thus, let us suppose that $0 \leq \xi \leq \alpha$ and $0 \leq \xi^* \leq \alpha$.
+Thus, let us suppose that $0 \leq \xi \leq \mu$ and $0 \leq \xi^* \leq \mu$.
 \\
@@
     \begin{aligned}
-        w=& \Phi^\intercal(\Phi \Phi^\intercal)^{-1} y\\
-        b<& \alpha
+        w=& \Phi^\intercal(\Phi \Phi^\intercal)^{-1} (y-b)
     \end{aligned}
 \end{equation}
-We then have
+We then have for any $\xi > 0$ and $\xi^* > 0$  that
 \begin{equation}
     \begin{aligned}
-        -\xi + y -\Phi\Phi^\intercal(\Phi \Phi^\intercal)^{-1}y-b<&0
+        -\xi + y -\Phi\Phi^\intercal(\Phi \Phi^\intercal)^{-1}(y-b)-b<&0
         \\
-        -\xi^* - y +\Phi\Phi^\intercal(\Phi \Phi^\intercal)^{-1}y+b<&0
+        -\xi^* - y +\Phi\Phi^\intercal(\Phi \Phi^\intercal)^{-1}(y-b)+b<&0
     \end{aligned}
@@
 \\
-We end this section with the derivation of the dual problem; that is the convex problem, we will solve in order to get the quantiles prdeiction of our quantile kernel algorithm.
+We end this section with the derivation of the dual problem; that is the convex problem, which we will solve in order to get the quantiles prediction of our quantile kernel algorithm.
 \\
@@
     \begin{aligned}
-        g(\alpha, \alpha^*, \nu, \nu^*)=& \inf_x \mathcal{L}(x,\lambda)\\
-    = & \inf_{\xi, \xi^*, w, b} \mathcal{L}(w,b,\xi,\xi^*)
+        g(\lambda, \lambda^*, \nu, \nu^*)= & \inf_{\xi, \xi^*, w, b} \mathcal{L}(w,b,\xi,\xi^*,\lambda, \lambda^*, \nu, \nu^*)
 \end{aligned}
@@
     \begin{cases}
-        \frac{\partial L}{\partial w}=0 \implies w=\Phi^\intercal(\alpha-\alpha^*)
+        \frac{\partial \mathcal{L}}{\partial w}=0 \implies w=\Phi^\intercal(\lambda-\lambda^*)
         \\
-        \frac{\partial L}{\partial b}=0 \implies (\alpha-\alpha^*)^\intercal\mathbb{1}=0
+        \frac{\partial \mathcal{L}}{\partial b}=0 \implies (\lambda-\lambda^*)^\intercal\mathbb{1}=0
         \\
-        \frac{\partial L}{\partial \xi}=0 \implies Cq \mathbb{1}-\alpha- \nu=0
+        \frac{\partial \mathcal{L}}{\partial \xi}=0 \implies Cq \mathbb{1}-\lambda- \nu=0
         \\
-        \frac{\partial L}{\partial \xi^*}=0 \implies C(1-q)\mathbb{1} -\alpha^* -\nu^*=0
+        \frac{\partial \mathcal{L}}{\partial \xi^*}=0 \implies C(1-q)\mathbb{1} -\lambda^* -\nu^*=0
     \end{cases}
 \end{equation}
-As pointed out previously, the lower bound resulting from the dual formulation is non trivial only when the lagrange multipliers are $\succeq 0$.
-Looking at the last two equations of the system \ref{eq:lagrange_derivatives}, this implies the following two constraints $\alpha \in [0, Cq\mathbb{1}]$ and $\alpha^*\in [0, C(1-q)\mathbb{1}]$.
+As pointed out previously, the lower bound resulting from the dual formulation is non trivial only when the lagrange multipliers $\lambda$ are $\succeq 0$.
+Looking at the last two equations of the system \ref{eq:lagrange_derivatives}, this implies the following two constraints $\lambda \in [0, Cq\mathbb{1}]$ and $\lambda^*\in [0, C(1-q)\mathbb{1}]$.
 \\
 Substitute the conditions for an optimum into \ref{eq:kqr_min5}, we obtain the dual formulation.
+% nu and nu^* go away because they are multipliers accounting for equality constraint, which in the optimum have to be satisfied hence are zero => h_i \times \nu=0
 \begin{equation}
     \begin{aligned}
-        g(\alpha, \alpha^*)=& \xi^\intercal(Cq\mathbb{1}-\alpha -\nu)+(\xi^*)^\intercal(C(1-q)\mathbb{1}-\alpha^*-\nu^*)-(\alpha-\alpha^*)^\intercal \Phi\Phi^\intercal(\alpha-\alpha^*)
+        g(\lambda, \lambda^*)=& \xi^\intercal(Cq\mathbb{1}-\lambda -\nu)+(\xi^*)^\intercal(C(1-q)\mathbb{1}-\lambda^*-\nu^*)-(\lambda-\lambda^*)^\intercal \Phi\Phi^\intercal(\lambda-\lambda^*)
         \\
-        & +(\alpha-\alpha^*)^\intercal y-(\alpha-\alpha^*)^\intercal b+\frac{1}{2}(\alpha-\alpha^*)^\intercal \Phi\Phi^\intercal(\alpha-\alpha^*)
+        & +(\lambda-\lambda^*)^\intercal y-(\lambda-\lambda^*)^\intercal b+\frac{1}{2}(\lambda-\lambda^*)^\intercal \Phi\Phi^\intercal(\lambda-\lambda^*)
         \\
         \\
-        g(\alpha, \alpha^*)=& 0+0-\frac{1}{2}(\alpha-\alpha^*)^\intercal \Phi\Phi^\intercal(\alpha-\alpha^*)+(\alpha-\alpha^*)^\intercal y-0
+        g(\lambda, \lambda^*)=& 0+0-\frac{1}{2}(\lambda-\lambda^*)^\intercal \Phi\Phi^\intercal(\lambda-\lambda^*)+(\lambda-\lambda^*)^\intercal y-0
         \\
-        g(\alpha, \alpha^*)=& -\frac{1}{2}(\alpha-\alpha^*)^\intercal \Phi\Phi^\intercal(\alpha-\alpha^*)+(\alpha-\alpha^*)^\intercal y
+        g(\lambda, \lambda^*)=& -\frac{1}{2}(\lambda-\lambda^*)^\intercal \Phi\Phi^\intercal(\lambda-\lambda^*)+(\lambda-\lambda^*)^\intercal y
     \end{aligned}
 \end{equation}
-Defining $\textrm{a}=(\alpha-\alpha^*)$ and letting $K$ the kernel matrix, we have that the dual optimisation problem reads as follows
+Defining $\alpha=(\lambda-\lambda^*)$ and letting $K$ the kernel matrix, we have that the dual optimisation problem reads as follows
 \begin{equation}\label{eq:kqr_min6}
     \begin{aligned}
-        \max_{\textrm{a}} \quad & -\frac{1}{2}\textrm{a}^\intercal K\textrm{a}+\textrm{a}^\intercal y\\
+        \max_{\alpha} \quad & -\frac{1}{2}\alpha^\intercal K\alpha+\alpha^\intercal y\\
     \textrm{s.t.} \quad & 
-    C(q-1)\mathbb{1}\preceq \textrm{a}\preceq Cq\mathbb{1}\\
-    &\textrm{a}^\intercal\mathbb{1}=1
+    C(q-1)\mathbb{1}\preceq \alpha \preceq Cq\mathbb{1}\\
+    &\alpha^\intercal\mathbb{1}=1
     \end{aligned}
@@
     \begin{aligned}
-        \min_{\textrm{a}} \quad & +\frac{1}{2}\textrm{a}^\intercal K\textrm{a}-\textrm{a}^\intercal y\\
+        \min_{\alpha} \quad & +\frac{1}{2}\alpha^\intercal K \alpha-\alpha^\intercal y\\
     \textrm{s.t.} \quad & 
-    C(q-1)\mathbb{1}\preceq \textrm{a} \preceq Cq\mathbb{1}\\
-    &\textrm{a}^\intercal\mathbb{1}=1
+    C(q-1)\mathbb{1}\preceq \alpha \preceq Cq\mathbb{1}\\
+    &\alpha^\intercal\mathbb{1}=1
     \end{aligned}
@@
 \begin{equation}
-    f(x)=\sum\limits_{i=1}^{n} \textrm{a}_i k(x_i, x)+b
+    f(x)=\sum\limits_{i=1}^{n} \alpha_i k(x_i, x)+b
 \end{equation}
-Since our optimisation problem possesses strong duality and it is differentiable in both the objective and the constraint, we have that it must satisfy the Karush Kuhn Tucker conditions, see section 5.5.3 \cite{boyd2004convex}.
-Thanks to the KKT conditions on the primal optimisation problem we have that $f(x_i)=y_i \ \mathrm{for} \ \textrm{a}_i \not \in \{C(q-1), Cq\}$. 
+Since our optimisation problem possesses strong duality and it is differentiable in both the objective and the constraint, we have that it must satisfy the Karush Kuhn Tucker conditions (KKT), see section 5.5.3 \cite{boyd2004convex}.
+Thanks to the KKT conditions on the primal optimisation problem we have that $f(x_i)=y_i \ \mathrm{for} \ \alpha_i \not \in \{C(q-1), Cq\}$. 
 To see this, we have to consider the KKT conditions.
+% optimal lambda times optimal x
 \begin{equation}
     \begin{aligned}
-    \lambda^* r_i(x^*)=&0, \quad i=1,\dots,n
+    \lambda r_i(x)=&0, \quad i=1,\dots,n
     \\
-    \nabla L(x^*)=&0
+    \nabla \mathcal{L}(x)=&0
 \end{aligned}
@@
     \begin{aligned}
-        \alpha_i(\xi_i-y_i+r_i)=&0
+        \lambda_i(\xi_i-y_i+r_i)=&0
         \\
-        \alpha_i^*(\xi_i^*+y_i-r_i)=&0
+        \lambda_i^*(\xi_i^*+y_i-r_i)=&0
         \\
@@
         \\
-        \nabla L =&0
+        \nabla \mathcal{L} =&0
     \end{aligned}
@@
     \begin{aligned}
-        \alpha_i(\xi_i-y_i+r_i)=&0
+        \lambda_i(\xi_i-y_i+r_i)=&0
         \\
-        \alpha_i^*(\xi_i^*+y_i-r_i)=&0
+        \lambda_i^*(\xi_i^*+y_i-r_i)=&0
         \\
-        (Cq-\alpha_i) \xi_i=&0
+        (Cq-\lambda_i) \xi_i=&0
         \\
-        (C(1-q)-\alpha_i) \xi_i^*=&0
+        (C(1-q)-\lambda_i) \xi_i^*=&0
     \end{aligned}
@@
 \begin{cases}
-    \alpha_i=Cq, \ \alpha_i^*=0 & \implies \alpha_i-\alpha_i^*=Cq, \xi_i\leq 0, \xi^*=0\implies \xi_i-y_i+f_i=0 \\
-    \alpha_i=0, \ \alpha_i^*=C(1-q) & \implies \alpha_i-\alpha_i^*=C(q-1), \xi_i= 0, \xi^*\leq 0 \implies \xi_i^*+y_i-f_i=0 \\
-    0\leq \alpha_i\leq Cq, \ 0\leq \alpha_i^*\leq C(1-q) & \implies \xi_i=0,  \xi^*_i= 0 \implies -y_i+f_i=0, \ y_i-f_i=0 \\
+    \lambda_i=Cq, \ \lambda_i^*=0 & \implies \lambda_i-\lambda_i^*=Cq, \xi_i\leq 0, \xi^*=0\implies \xi_i-y_i+f_i+b=0 \\
+    \lambda_i=0, \ \lambda_i^*=C(1-q) & \implies \lambda_i-\lambda_i^*=C(q-1), \xi_i= 0, \xi^*\leq 0 \implies \xi_i^*+y_i-f_i-b=0 \\
+    0\leq \lambda_i< Cq, \ 0\leq \lambda_i^*< C(1-q) & \implies \xi_i=0,  \xi^*_i= 0 \implies -y_i+f_i+b=0, \ y_i-f_i-b=0 \\
 \end{cases}
@@
 
-Therefore in order to retrieve b we simply have to choose an index $i$ such that  $\textrm{a}_i \not \in \{C(q-1), Cq\}$
+Therefore in order to retrieve b we simply have to choose an index $i$ such that  $\alpha_i \not \in \{C(q-1), Cq\}$
 and let
@@
     \begin{aligned}
-    b=&y_i-\sum\limits_{i=1}^n \textrm{a}_i k(x_i,x)
+    b=&y_i-\sum\limits_{i=1}^n \alpha_i k(x_i,x)
     \end{aligned}
@@
     \toprule
-     & Linear qr & Gbm qr & Quantile forest & Kernel qr \\
+     & Linear qr & Gbm qr & Quantile forest & KQR \\
     \midrule
@@
     \toprule
-    Quantiles & Linear qr & Gbm qr & Quantile forest & Kernel qr \\
+    Quantiles & Linear qr & Gbm qr & Quantile forest & KQR \\
     \midrule
@@
 Not only kernel quantile regression was the best in terms of total pinball loss but also the best in terms of each quantile pinball loss and mean absolute error.
-Comparison has been carried out on further datasets, yielding similar conclusion as the one of above, to know more see \ref{appendix:quantile_regressor_extensive_comparison}.
+Comparison has been carried out on further datasets, yielding similar conclusions as the one of above, to know more see \ref{appendix:quantile_regressor_extensive_comparison}.
 
-We conclude this chapter with reporting the same kind of tables \ref{tab:kernel pinball comparison}, \ref{tab:kernel pinball comparison quantile-wise}, \ref{tab:kernel mae comparison} and figures \ref{fig:kernel quantile regressors comparison} comparing various kernel functions.
+We conclude this chapter with reporting the same kind of tables \ref{tab:kernel pinball comparison}, \ref{tab:kernel pinball comparison quantile-wise}, \ref{tab:kernel mae comparison} and figure \ref{fig:kernel quantile regressors comparison} comparing various kernel functions.
 
@@
         & Linear & 10.463867    \\
-        & Polynomial & 11.238393     \\
+        & Polynomial & 10.089147     \\
         & Sigmoid & 16.253973            \\
@@
         \includegraphics[width=1.1\textwidth]{images/melborune_gaussian_rbf_kernel_quantile_regression.png} 
-        \caption{Gauss RBF} 
+        \caption{Gaussian RBF} 
         \label{} 

### 2024-05-17 thesis/input_files/data_exploratory.tex
May 17th 2024

@@
 Next, we will explain the extract transform load pipeline (ETL) that we set up in order to ease the workflow of our studies.
-Finally, in order to to better understand patters within the data, we carry out an exploratory data analysis.
+Finally, in order to to better understand patterns within the data, we carry out an exploratory data analysis.
 
 \section{GEFCom2014}
-The first class of datasets we identified were the ones from the GEFCom 2014 competition. The data is freely hosted on Dr. Hong blog \cite{hong2016probabilistic}.
-The main reason is that, these dataset are considered an a excellent test case for comparing predictive models between the EF community.
+The first class of datasets we identified were the ones from the GEFCom 2014 competition. The data is freely hosted on Dr.\ Hong blog \cite{hong2016probabilistic}.
+The main reason is that, these datasets are considered an excellent test case for comparing predictive models between the EF community.
 Additionally, the scores of the competing models are freely available, this enables us to carry out a clean and transparent comparative study.
@@
 In exploring this data we started off by plotting the time series of the zonal prices alongside with the total load and zonal load, see figure \ref{fig:price_track_fig1}.
-After that, we plotted the scatter plot for inspecting the correlation between the predictors and the dependent variable, figure \ref{fig:gefcom_zonal_price_vs_zonal_load} and figure \ref{fig:gefcom_zonal_price_vs_total_load}.
-From these visualisation, we can see their goodness for predicting zonal price since there exists a clear correlation between the candidate predictors and the target variable.
+After that, we plotted the scatter plot for inspecting the correlation between the predictors and the dependent variables, figure \ref{fig:gefcom_zonal_price_vs_zonal_load} and figure \ref{fig:gefcom_zonal_price_vs_total_load}.
+From these visualisations, we can see their goodness for predicting zonal price since there exists a clear correlation between the candidate predictors and the target variables.
 
@@
 For each of those we have the prediction of the competition benchmark model and the train file upon which we will fit our model.
-Each data from the various task folders are shifted by one months between others.
+Any data from the various task folders are shifted by one month between eachother.
 \\
@@
 Finally, figure \ref{fig:gefcom_load_vs_w_avg} shows the scatter plot between the the load and the weather.
-These plots motivate the selection of weather temperature as predictors for the load.
+These plots motivate the selection of weather temperature as one of the predictors for the load.
 

### 2024-05-17 thesis/input_files/point.tex
May 17th 2024

@@
-This chapter covers the theory of the most widely used method for point prediction in the electricity forecasting literature.
+This chapter covers the theory of the most widely used methods for point prediction in the electricity forecasting literature.
 Besides discussing the theory underlying such models, this and the following chapter include also a couple of worked examples in order to get acquainted with the practical applications of these models.
@@
 Autoregressive models are a standard approach for modelling time series data.
-Before introducing the popular model in electricity point forecasting, it is important to remember that autoregressive models assume the time series to be stationary.
+Before introducing this popular model in electricity point forecasting, it is important to remember that autoregressive models assume the time series to be stationary.
 On the other hand, load and price have been observed to be a non stationary time series.
-Basically, stationarity means that the distribution of any subsequence of random variables of the stochatic process is invariant to shifts along the time dimension.
+In a nutshell, stationarity means that the distribution of any subsequence of random variables of the stochastic process is invariant to shifts along the time dimension.
 Therefore, checking stationarity is a crucial step in applying this class of models. %(unit root tests)
-Furthermore, should the time series be non stationary, we have to convert it to a stationary one by a combination of either differencing, detrending and/or log transforming it.
+Furthermore, should the time series be non stationary, we can convert it to a stationary one by a combination of either differencing, detrending and/or log transforming it.
 \subsection{Autoregressive integrated moving average}
@@
 The order p is identified by taking the lags at which the sample partial autocorrelation function (PACF) falls outside its 95\% confidence interval.
-In the same way, the order q is estimated by considering the sample autocorrelation function of the observed time series.
+In the same way, the order q is estimated by considering the sample autocorrelation function (ACF) of the observed time series.
 Finally, we compare ARIMA models with different p,d,q (in a neighbourhood of the identified orders) and choose the ones minimising either AIC or BIC criteria.
@@
 \end{equation}
-\subsection{Seasonal autoregressive integrated moving average and Seasonal autoregressive integrated moving average with exogenous variables}
+\subsection{Seasonal autoregressive integrated moving average and seasonal autoregressive integrated moving average with exogenous variables}
 SARIMA and SARIMAX incorporate seasonalities into our time series models.
 The notation for these models is ARIMA$(p,d,q)(P,D,Q)_m$.
-The second term $(P,D,Q)_m$ represents the autoregressive structure of the seasonal patter, where $m$ indicates the number of seasonalities of our time series.
+The second term $(P,D,Q)_m$ represents the autoregressive structure of the seasonal pattern, where $m$ indicates the number of seasonalities of our time series.
 For instance, in modelling series of hourly data with daily periodicity, we would set m=24.
@@
 \begin{equation}
-    \psi(B)\Psi(B^s)\nabla^d\nabla_m^Dy_t=\theta(B)\Theta(B^s)\epsilon_t
+    \psi(B)\Psi(B^P)\nabla^d\nabla_m^Dy_t=\theta(B)\Theta(B^Q)\epsilon_t
 \end{equation}
-where $\Psi$ and $\Theta$ account for the $P$ and $Q$ of the seasonal patter $m$ respectively.
+where $\Psi$ and $\Theta$ account for the $P$ and $Q$ of the seasonal pattern $m$ respectively.
 \section{Generalized additive model}
@@
 Particularly successful in time series forecasting is the Prophet model \cite{taylor2018forecasting}, developed by Meta.
-This additive model is capable of handling non linear trends, holiday effects and yearly weekly and daily seasonality.
-Thus, the reason we choose it as a valid benchmark to compare against.
+This additive model is capable of handling non linear trends, holiday effects and yearly weekly and daily seasonalities.
+Thus, the reason why we choose it as a valid benchmark to compare against.
 \section{K-nearest neighbours regression}
@@
 Developed at the AT\&T Bell Laboratories by Vapnik et al.
-\cite{cortes1995support} \cite{vapnik1997support}, support vector machines SVMs are one of the most popular techniques within the field of statistical learning.
+\cite{cortes1995support} \cite{vapnik1997support}, support vector machines (SVMs) are one of the most popular techniques within the field of statistical learning.
 \\
 The goal of support vector regression is finding a function $f(x)$ with at most $\epsilon$ deviation from the actual observed data $y_i$ for every $i$ and as flat as possible. 
-Put differently, we would like a model to keep error less than the $\epsilon$ threshold,  
+Put differently, we would like a model to keep error less than the $\epsilon$ threshold.  
 %https://stats.stackexchange.com/questions/5945/understanding-svm-regression-objective-function-and-flatness
-In standard SVR we have
+In standard support vector regression (SVR) we have
 \begin{equation}
-    f(x)=\langle w,x \rangle +b \ \textrm{with} \ w \in \mathcal{X}, b \in \mathbb{R}
+    f(x)=\langle w,x \rangle_{\mathbb{R}^n} +b \ \textrm{with} \ w \in \mathcal{X} \subseteq \mathbb{R}^n, b \in \mathbb{R}
 \end{equation}
-Where $\mathcal{X}$ denotes the space of the input patters $x_i$.
+Where $\mathcal{X}$ denotes the space of the input data $x_i$.
 We can translate the flatness requirement into minimising the squared norm of $w$. Doing so, we can formulate our problem as a convex optimisation problem.
@@
 \end{equation}
-The C constant trades off between $\epsilon$ deviation tolerance and flatness of the function f.
-A bigger C gives more importance to error minimization while as C gets smaller, flattness gains importance.
+The $C$ constant trades off between $\epsilon$ deviation tolerance and flatness of the function $f$.
+A bigger $C$ gives more importance to error minimization while as $C$ gets smaller, flattness gains importance.
 \\
@@
     \begin{aligned}
-        \max_{\alpha_i, \alpha_i^*} \quad& -\frac{1}{2}\sum\limits_{i,j=1}^n(\alpha_i-\alpha_i^*)(\alpha_j-\alpha_j^*)\langle x_i,x_j\rangle +\sum\limits_{i=1}^n(\alpha_i-\alpha_i^*)(y_i-\epsilon)
+        \max_{\lambda_i, \lambda_i^*} \quad& -\frac{1}{2}\sum\limits_{i,j=1}^n(\lambda_i-\lambda_i^*)(\lambda_j-\lambda_j^*)\langle x_i,x_j\rangle +\sum\limits_{i=1}^n(\lambda_i-\lambda_i^*)(y_i-\epsilon)
         \\
-        s.t. \quad& \sum\limits_{i=1}^n(\alpha_i-\alpha_i^*)=0
+        s.t. \quad& \sum\limits_{i=1}^n(\lambda_i-\lambda_i^*)=0
         \\
-        \quad& \alpha_i, \alpha_i^* \in [0, C]
+        \quad& \lambda_i, \lambda_i^* \in [0, C]
     \end{aligned}
 \end{equation}
-Rearranging the gradient of the lagrangian with respect to $w$, we obtain the so called support vector expansion.
+Rearranging the gradient of the primal lagrangian with respect to $w$, we obtain the so called support vector expansion.
 \begin{equation}\label{eq:support vector expansion}
-    w=\sum\limits_{i=1}^n(\alpha_i-\alpha_i^*)x_i
+    w=\sum\limits_{i=1}^n(\lambda_i-\lambda_i^*)x_i
 \end{equation}
 This means that $w$ can be completely described as a linear combination of the training features $x_i$.
-Notice, the complexity of the function representation is independent of the feature space dimensionality, but depends only on the number of support vectors; that is those point $i$ for which $(\alpha_i-\alpha_i^*)\neq 0$
+Notice, the complexity of the function representation is independent of the feature space dimensionality, but depends only on the number of support vectors; that is those point $i$ for which $(\lambda_i-\lambda_i^*)\neq 0$
 Moreover, we do not needd to compute $w$ explicitly in order to evaluate $f(x)$.
-Furthermore, equation \ref{eq:support vector expansion} implies
+That is because, equation \ref{eq:support vector expansion} implies
 \begin{equation}
-    f(x)=\sum\limits_{i=1}^n(\alpha_i-\alpha_i^*)\langle x_i, x\rangle +b
+    f(x)=\sum\limits_{i=1}^n(\lambda_i-\lambda_i^*)\langle x_i, x\rangle +b
 \end{equation}
-Employing the Karush-Kuhn-Tucker conditions, $b$ can be retrieved easily.
-Particulalry, the condition that the product between the constraints and the dual variable has to vanish.
-These implies that the lagrange multipliers $\alpha_i, \alpha_i^*$ may be nonzero only for the samples inside the epsilon insensitive tube.
-Consequently for any of these data points, the equality $b=y_i-\langle w, x_i-\epsilon\rangle$ holds.
+Employing the Karush-Kuhn-Tucker conditions, $b$ can be retrieved easily. Particulalry, the condition that the product between the constraints and the dual variable has to vanish.
+Consequently for any of the data points with associated $\lambda_i, \lambda^*_i \in (0,C)$, the equality $b=y_i-\langle w, x_i-\epsilon\rangle$ has to hold.
 \\
-To get an idea of how SVR works, see figure \ref{fig:svr1} for how a support vector regression handles a sinusoidal function with noise.
+Furthermore, this implies that the Lagrange multipliers $\lambda_i, \lambda_i^*$ may be nonzero only for the samples inside the epsilon insensitive tube, that is only those points may be support vectors.
+\\
+To get an idea of how the algorithm works, see figure \ref{fig:svr1} for how a support vector regression handles a sinusoidal function with noise.
 \\
@@
     \includegraphics[width=\textwidth]{images/svr1.png}
-    \caption{support vector regression}
+    \caption{Linear support vector regression}
     \label{fig:svr1}
@@
 \begin{equation}\label{eq:kernel ridge}
-    \hat{f}=\argmin{f\in \mathcal{H}}\frac{1}{2}\sum\limits_{i=1}^n(y_i-f(x_i))^2+\frac{\lambda}{2}\|f\|_{\mathcal{H}}^2
+    \hat{f}=\argmin{f\in \mathcal{H}}\frac{1}{2}\sum\limits_{i=1}^n(y_i-f(x_i))^2+\frac{\tau}{2}\|f\|_{\mathcal{H}}^2
 \end{equation}
@@
 \begin{theorem}
-    Let $k:\mathcal{X} \times \mathcal{X} \mapsto R$ on a non empty set $\mathcal{X}$ with corresponding RKHS $\mathcal{H}$,
-    a strictly increasing real valued function $g:[0,\infty) \to \mathbb{R}$, and $E:(\mathcal{X}\times \mathbb{R}^2)^n \to \mathbb{R} \cup \{\infty\}$.
+    Let $k:\mathcal{X} \times \mathcal{X} \mapsto \mathbb{R}$ on a non empty set $\mathcal{X}$ with corresponding RKHS $\mathcal{H}$,
+    a strictly increasing real valued function $g:[0,\infty) \to \mathbb{R}$, and the loss function $E:(\mathcal{X}\times \mathbb{R}^2)^n \to \mathbb{R} \cup \{\infty\}$.
     Consider the regularized risk functional $Risk[f]:f \mapsto  E\left((x_{1},y_{1},f(x_{1})),\ldots, (x_{n},y_{n},f(x_{n}))\right)+g\left(\| f\| \right)$.
@@
         \begin{aligned}
-            f(x_j)=&\langle \phi(x_j), \sum\limits_{i=1}^{n}\alpha_i \phi(x_i)+\nu\rangle
+            f(x_j)=&\langle \phi(x_j), \sum\limits_{i=1}^{n}\alpha_i \phi(x_i)+\nu\rangle_{\mathcal{H}}
             \\
-            =&\sum\limits_{i=1}^{n}\alpha_i\langle \phi(x_j), \phi(x_i)\rangle
+            =&\sum\limits_{i=1}^{n}\alpha_i\langle \phi(x_j), \phi(x_i)\rangle_{\mathcal{H}}
         \end{aligned}
@@
         \begin{aligned}
-            g(\|f\|)=&g\left(\left\|\sum\limits_{i=1}^n \alpha_i \phi(x_i)+v\right\|_\mathcal{H}\right)
+            g(\|f\|_{\mathcal{H}})=&g\left(\left\|\sum\limits_{i=1}^n \alpha_i \phi(x_i)+v\right\|_\mathcal{H}\right)
             \\
@@
             \\
-            =&g\left(\sqrt{\left\|\sum\limits_{i=1}^n \alpha_i \phi(x_i)\right\|_{\mathcal{H}}^2+\left\|v\right\|_\mathcal{H}^2+2\langle\sum\limits_{i=1}^n \alpha_i \phi(x_i),\nu \rangle_{\mathcal{H}}}\right)
+            =&g\left(\sqrt{\left\|\sum\limits_{i=1}^n \alpha_i \phi(x_i)\right\|_{\mathcal{H}}^2+\left\|v\right\|_\mathcal{H}^2+2\left\langle\sum\limits_{i=1}^n \alpha_i \phi(x_i),\nu \right\rangle_{\mathcal{H}}}\right)
             \\
@@
     \begin{aligned}
-    \hat{\alpha}=&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\lambda}{2}\left\|\sum _{i=1}^{n}\alpha _{i}k(\cdot ,x_{i})\right\|_{\mathcal{H}}^2
+    \hat{\alpha}=&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\tau}{2}\left\|\sum _{i=1}^{n}\alpha _{i}k(\cdot ,x_{i})\right\|_{\mathcal{H}}^2
     \\
-    =&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\lambda}{2} \left\langle \sum _{i=1}^{n}\alpha _{i}k(\cdot ,x_{i}),\sum _{j=1}^{n}\alpha _{j}k(\cdot ,x_{j})\right\rangle_{\mathcal{H}}^2
+    =&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\tau}{2} \left\langle \sum _{i=1}^{n}\alpha _{i}k(\cdot ,x_{i}),\sum _{j=1}^{n}\alpha _{j}k(\cdot ,x_{j})\right\rangle_{\mathcal{H}}^2
     \\
-    =&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\lambda}{2} \sum _{i,j=1}^{n}\alpha _{i}\alpha _{j}\langle k(\cdot ,x_{i}),k(\cdot ,x_{j})\rangle_{\mathcal{H}}^2
+    =&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\tau}{2} \sum _{i,j=1}^{n}\alpha _{i}\alpha _{j}\langle k(\cdot ,x_{i}),k(\cdot ,x_{j})\rangle_{\mathcal{H}}^2
     \\
-    =&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\lambda}{2} \sum _{i,j=1}^{n}\alpha _{i}\alpha _{j}K_{ij}
+    =&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\tau}{2} \sum _{i,j=1}^{n}\alpha _{i}\alpha _{j}K_{ij}
     \\
-    =&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\lambda}{2} \alpha^\intercal K \alpha
+    =&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\tau}{2} \alpha^\intercal K \alpha
     \\
-    =&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}(y-K\alpha)^\intercal (y-K\alpha)+\frac{\lambda}{2} \alpha^\intercal K \alpha
+    =&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}(y-K\alpha)^\intercal (y-K\alpha)+\frac{\tau}{2} \alpha^\intercal K \alpha
 \end{aligned}
@@
 \begin{equation}
-    -Ky+K^2\alpha+\lambda K \alpha\overset{!}{=}0
+    -Ky+K^2\alpha+\tau K \alpha\overset{!}{=}0
 \end{equation}
 
-We have then, that $\hat{\alpha}=(K+\lambda I)^{-1}y$. It follows that the solution takes the form
+We have then, that $\hat{\alpha}=(K+\tau I)^{-1}y$. It follows that the solution takes the form
 \begin{equation}
@@
 \subsection{Kernel support vector regression}
-Support vector regression (SVR) can be kernelized by swapping the $\mathbb{R}^n$ euclidean dot product of data $x$ with the dot product in the higher feature space $\mathcal{F}$.
+Support vector regression (SVR) can be kernelized by swapping the $\mathbb{R}^n$ euclidean dot product of data $x$ with the dot product in the higher feature space $\mathcal{H}$.
 Doing so, the optimisation problem can be restated as
@@
     \begin{aligned}
-        \max_{\alpha_i, \alpha_i^*} \quad& -\frac{1}{2}\sum\limits_{i,j=1}^n(\alpha_i-\alpha_i^*)(\alpha_i-\alpha_i^*)k(x_i, x_j)+\sum\limits_{i=1}^n(y_i-\epsilon)(\alpha_i-\alpha_i^*)
+        \max_{\lambda_i, \lambda_i^*} \quad& -\frac{1}{2}\sum\limits_{i,j=1}^n(\lambda_i-\lambda_i^*)(\lambda_i-\lambda_i^*)k(x_i, x_j)+\sum\limits_{i=1}^n(y_i-\epsilon)(\lambda_i-\lambda_i^*)
         \\
-        s.t. \quad& \sum\limits_{i=1}^n(\alpha_i-\alpha_i^*)=0
+        s.t. \quad& \sum\limits_{i=1}^n(\lambda_i-\lambda_i^*)=0
         \\
-        \quad& \alpha_i, \alpha_i^* \in [0,C]
+        \quad& \lambda_i, \lambda_i^* \in [0,C]
     \end{aligned}
 \end{equation}
-Our regressor $f$ is then given by
+Letting $\alpha_i=\lambda_i - \lambda_i^*$, our regressor $f$ is then given by
 \begin{equation}
     \begin{aligned}
-        f(x)=\sum\limits_{i=1}^n(\alpha_i-\alpha_i^*)k(x_i, x)+b
+        f(x)=\sum\limits_{i=1}^n \alpha_i k(x_i, x)+b
     \end{aligned}

### 2024-05-17 thesis/input_files/probabilistic.tex
May 17th 2024

@@
 For the outer problem, we scan through all the (j,s) tuples and pick the best pair. 
-Next, one or both of these regions from the previous step are split into two more regions; we recurse this process until some stopping condition is triggered (max number of branches, max deth of tree, threshold on the mean squared error (MSE), minimum number of observations in each leaf node).
+Next, one or both of these regions from the previous step are split into two more regions; we recurse this process until some stopping condition is triggered (max number of branches, max depth of tree, threshold on the mean squared error (MSE), minimum number of observations in each leaf node).
 Notice, this being a greedy algorithm implies that our final solution is guaranteed to be just a local optimum not a global one.
 Even though their simplicity, these models have proved themselves to be really powerful.
-See figure \ref{fig:decision_tree} for a example where decision tree with different hyperparameters are fitted to a sine wave plus noise.
+See figure \ref{fig:decision_tree} for an example where decision trees with different hyperparameters are fitted to a sine wave plus noise.
 The most popular among these model is the CART \cite{breiman2017classification} tree, its name comes from the fact that it can handle both classification and regression problems.
@@
 \subsection{Boosting}
-With boosting we fit an additive expansion of elementary basis functions; with M basis functions we have 
+With boosting we fit an additive expansion of elementary basis functions; with $M$ basis functions we have 
 \begin{equation}
@@
 where $b(x;\gamma)$ are the basis functions while $\beta_m$ are the coefficients of the expansion.
-Boosted models are fitted by minimizing a loss function over the training data
+Boosted models are fitted by minimizing a loss function $L$ over the training data
 \begin{equation}
@@
 \begin{equation}\label{eq:kqr_min1}
-    R[f]:=\frac{1}{m}\sum\limits_{i=1}^{m}\rho_q(y_i-f(x_i))+\frac{\lambda}{2}\|w\|_\mathcal{H}^2
+    Risk[f]:=\frac{1}{m}\sum\limits_{i=1}^{m}\rho_q(y_i-f(x_i))+\frac{\tau}{2}\|w\|_\mathcal{H}^2
 \end{equation}
@@
     \end{equation}
-Note the division by $\lambda$ so that $C:=\frac{1}{\lambda m}$.
+Note the division by $\tau$ so that $C:=\frac{1}{\tau m}$.
 \\
@@
 % when lagrange multiplier is negative, we have that g > p
-Therefore, the idea is that by maximizing the dual function subject to the contraint $\lambda \succeq 0$, we can obtain an approximate or perfect solution to the primal problem.
+Therefore, the idea is that by maximizing the dual function subject to the constraint $\lambda \succeq 0$, we can obtain an approximate or perfect solution to the primal problem.
 \\
@@
 
-Therefore in order to retrieve b we simply have to choose an index $i$ such that  $\alpha_i \not \in \{C(q-1), Cq\}$
+Therefore in order to retrieve $b$ we simply have to choose an index $i$ such that  $\alpha_i \not \in \{C(q-1), Cq\}$
 and let

### 2024-05-19 thesis/input_files/analysis.tex
added conc sec
May 14th 2024
May 18th 2024

@@
 Doing so, RMSE went down to 10.29.
-Forecast is visualized in \ref{fig:prophet_price2}, moreover, figure \ref{fig:prophet_price2.1} and figure \ref{fig:prophet_price2.2} break down the trend and the seasonalities respectively.
+Forecast is visualized in \ref{fig:prophet_price2}, moreover.
+% figure \ref{fig:prophet_price2.1} and figure \ref{fig:prophet_price2.2} break down the trend and the seasonalities respectively.
 \begin{figure}[!h]
@@
 
-\begin{figure}[!h]
-    \includegraphics[width=\textwidth]{images/prophet_price2.1.png}
-    \caption{Seasonalities breakdown}
-    \label{fig:prophet_price2.1}
-\end{figure}
+% \begin{figure}[!h]
+%     \includegraphics[width=\textwidth]{images/prophet_price2.1.png}
+%     \caption{Seasonalities breakdown}
+%     \label{fig:prophet_price2.1}
+% \end{figure}
 
-\begin{figure}[!h]
-    \includegraphics[width=\textwidth]{images/prophet_price2.2.png}
-    \caption{Prices trend}
-    \label{fig:prophet_price2.2}
-\end{figure}
+% \begin{figure}[!h]
+%     \includegraphics[width=\textwidth]{images/prophet_price2.2.png}
+%     \caption{Prices trend}
+%     \label{fig:prophet_price2.2}
+% \end{figure}
 Prophet alongside with the right features is a good model in the context of electricity price forecasting, it is capcable of catching the trend, the peak and the lows.
@@
     \label{tab:point_RMSE}
-    \begin{adjustbox}{width=\textwidth, height=3cm}
+    \begin{adjustbox}{width=\textwidth}
         \begin{tabular}{lrrrrrrrrrrrrrrr}
@@
     \label{tab:point_MAE}
-    \begin{adjustbox}{width=\textwidth, height=3cm}
+    \begin{adjustbox}{width=\textwidth}
         \begin{tabular}{lrrrrrrrrrrrrrrr}
@@
     \label{tab:point_MAPE}
-    \begin{adjustbox}{width=\textwidth, height=3cm}
+    \begin{adjustbox}{width=\textwidth}
         \begin{tabular}{lrrrrrrrrrrrrrrr}
@@
 \section{Probabilistic forecasting}
-In this subsection we compare performance of quantile regressors.
-\\
-- experiment for quantile estimator on gefcom2014 and we are good
+% In this subsection we compare performance of quantile regressors.
+% \\
+% - experiment for quantile estimator on gefcom2014 and we are good
+
+\section{Conclusions}

### 2024-05-19 thesis/input_files/analysis.tex
May 19th 2024

@@
-Analysis of experiments and results
-\\
-Building on the theory introduced in \ref{ch:point} and \ref{ch:prob}, this section covers the experiments carried out and the results obtained.
+% Analysis of experiments and results
+% \\
+Building on the theory introduced in \ref{ch:point} and \ref{ch:prob}, this section covers the experiments carried out and the results obtained respectively.
 
@@
 \section{Probabilistic forecasting}
+We continue our experimental analysis by considering the probabilistic framework.
+The goal of this section is comparing between different quantile regressors and study the choiche of the kernel function.
+
+% quantile regressors
+\subsection{Quantile regressors comparison}
+Using the load data from \href{https://www.energy-charts.info/index.html?l=en&c=DE}{energy charts} for Germany and Switzerland, we proceeded to compare between quantile regression algorithms.
+In this study, we trained our models on the entire 2021 data and then we tested them and measured their score on the 2022 data.
+Notice, for kernel quantile regression (KQR), the kernel of choice in this setting is the absolute Laplacian.
+Finally notice, the loss function utilized in this section is the normalized pinball loss.
+% normalized by the magnitude of the average load of each country.
+Here, we built two models. In the first, the national load is forecasted by taking as predictor the following variables
+\begin{itemize}
+    \item Weather temperature
+    \item Wind speed
+\end{itemize}
+Results for Germany are reported in table \ref{tab:energy_chart_de} and figure \ref{fig:DE_load_CI} while results for Switzerland can be found in table \ref{tab:energy_chart_ch} and figure \ref{fig:CH_load_CI}.
+
+\begin{table}[!ht]
+    \centering
+    \caption{Pinball loss for load in Germany (2022)}
+    \label{tab:energy_chart_de}
+    \begin{tabular}{lrrrr}
+    \toprule
+    Quantile & LQR & GBMQR & QF & KQR \\
+    \midrule
+    0.1 & 0.04517 & 0.02890 & 0.02891 & \textbf{0.02767} \\
+    0.2 & 0.08086 & 0.04745 & 0.04906 & \textbf{0.04670} \\
+    0.3 & 0.10820 & 0.06130 & 0.06254 & \textbf{0.05981} \\
+    0.4 & 0.12722 & 0.07031 & 0.07136 & \textbf{0.06834} \\
+    0.5 & 0.13886 & 0.07371 & 0.07428 & \textbf{0.07247} \\
+    0.6 & 0.14182 & 0.07129 & 0.07381 & \textbf{0.07054} \\
+    0.7 & 0.13591 & 0.06227 & 0.06237 & \textbf{0.06215} \\
+    0.8 & 0.11869 & 0.04741 & \textbf{0.04723} & 0.04730 \\
+    0.9 & 0.08458 & 0.02687 & \textbf{0.02668} & 0.02698 \\
+    \midrule
+    CRPS & 0.10904 & 0.05439 & 0.05514 & \textbf{0.05355} \\
+    \bottomrule
+    \end{tabular}
+    \end{table}
+
+\begin{table}[!ht]
+\centering
+    \caption{Pinball loss for load in Switzerland (2022)}
+    \label{tab:energy_chart_ch}
+    \begin{tabular}{lrrrr}
+    \toprule
+    Quantile & LQR & GBMQR & QF & KQR\\
+    \midrule
+    0.1 & 0.04258 & 0.01909 & 0.01899 & \textbf{0.01896} \\
+    0.2 & 0.07460 & 0.03076 & 0.03046 & \textbf{0.03041} \\
+    0.3 & 0.10012 & 0.03838 & 0.03903 & \textbf{0.03814} \\
+    0.4 & 0.11910 & 0.04311 & 0.04378 & \textbf{0.04294} \\
+    0.5 & 0.13095 & 0.04499 & 0.04489 & \textbf{0.04466} \\
+    0.6 & 0.13426 & 0.04387 & 0.04397 & \textbf{0.04347} \\
+    0.7 & 0.12855 & 0.03928 & 0.03948 & \textbf{0.03894} \\
+    0.8 & 0.11201 & 0.03083 & 0.03166 & \textbf{0.03066} \\
+    0.9 & 0.07884 & 0.01878 & 0.01878 & \textbf{0.01841} \\
+    \midrule
+    CRPS & 0.10233 & 0.03434 & 0.03456 & \textbf{0.03407} \\
+    \bottomrule
+    \end{tabular}
+    \end{table}
+
+    \begin{figure}[!ht]
+        \makebox[\linewidth]{\includegraphics[width = 1.5\textwidth]{images/DE_load_CI.png}}
+        \caption[Probabilistic forecast for load in Germany]{Load 90\% confidence interval for Germany Energy charts data using KQR Laplacian: \small \textmd{Electric load probabilistic forecast Germany 2022.
+The black line is the observed path for the load.
+The 90\% confidence interval bands are plotted in green.
+Lower and upper red lines denote the 95\% and 5\% quantile forecast respectively.}}
+        \label{fig:DE_load_CI}
+    \end{figure}
+
+    \begin{figure}[!ht]
+        \makebox[\linewidth]{\includegraphics[width = 1.5\textwidth]{images/CH_load_CI.png}}
+        \caption[Probabilistic forecast for load in Switzerland]{Load 90\% confidence interval for Switzerland Energy charts data using KQR Laplacian: \small \textmd{Electric load probabilistic forecast for Switzerland 2022.
+The black line is the observed path for the load.
+The 90\% confidence interval bands are plotted in green.
+Lower and upper red lines denote the 95\% and 5\% quantile forecast respectively.}}
+        \label{fig:CH_load_CI}
+    \end{figure}
+
+Next, we extended the models to take into account the impact of two additional categorical variables
+\begin{itemize}
+    \item Is holiday: a binary variable for holidays where holiday$=1$ and working day$=0$
+    \item Day of week: an ordinal categorical variable corresponding to the day of the week , e.i.\ Monday$=0$,$\dots$, Sunday$=6$
+\end{itemize}
+Table \ref{tab:energy_chart_de_additional_predictors} reports the model score on the German data while table \ref{tab:energy_chart_ch_additional_predictors} is the one corresponding to the Suisse data.
+
+\begin{table}[!ht]
+    \centering
+    \caption{PF pinball loss for load in Germany (2022) model\_2}
+    \label{tab:energy_chart_de_additional_predictors}
+    \begin{tabular}{lrrrr}
+    \toprule
+    Quantile & LQR & GBMQR & QF & KQR \\
+    \midrule
+    0.1 & 0.04522 & 0.02580 & 0.02796 & \textbf{0.02491} \\
+    0.2 & 0.08048 & 0.04326 & 0.04625 & \textbf{0.04209} \\
+    0.3 & 0.10677 & 0.05570 & 0.05706 & \textbf{0.05424} \\
+    0.4 & 0.12610 & 0.06182 & 0.06346 & \textbf{0.06071}\\
+    0.5 & 0.13767 & 0.06214 & 0.06544 & \textbf{0.06150} \\
+    0.6 & 0.14077 & 0.05752 & 0.06001 & \textbf{0.05726} \\
+    0.7 & 0.13409 & 0.04893 & 0.05122 & \textbf{0.04869} \\
+    0.8 & 0.11628 & 0.03687 & 0.03822 & \textbf{0.03665} \\
+    0.9 & 0.08204 & 0.02114 & 0.02324 & \textbf{0.02108} \\
+    \midrule
+    CRPS & 0.10771 & 0.04591 & 0.04810 & \textbf{0.04524} \\
+    \bottomrule
+    \end{tabular}
+    \end{table}
+
+\begin{table}[!ht]
+    \centering
+    \caption{Pinball loss for load in Switzerland (2022) model\_2}
+    \label{tab:energy_chart_ch_additional_predictors}
+    \begin{tabular}{lrrrr}
+    \toprule
+    Quantile & LQR & GBMQR & QF & KQR \\
+    \midrule
+    0.1 & 0.04118 & 0.01881 & 0.01909 & \textbf{0.01876} \\
+    0.2 & 0.07215 & 0.03004 & 0.03040 & \textbf{0.02979} \\
+    0.3 & 0.09603 & 0.03749 & 0.03819 & \textbf{0.03693} \\
+    0.4 & 0.11329 & 0.04230 & 0.04311 & \textbf{0.04165} \\
+    0.5 & 0.12403 & 0.04404 & 0.04465 & \textbf{0.04358} \\
+    0.6 & 0.12748 & 0.04282 & 0.04363 & \textbf{0.04249} \\
+    0.7 & 0.12223 & 0.03789 & 0.03891 & \textbf{0.03779} \\
+    0.8 & 0.10639 & \textbf{0.02961} & 0.03097 & 0.02972 \\
+    0.9 & 0.07531 & \textbf{0.01847} & 0.01957 & 0.01862 \\
+    \midrule
+    CRPS & 0.09756 & 0.03350 & 0.03428 & \textbf{0.03326} \\
+    \bottomrule
+    \end{tabular}
+    \end{table}
+
+% now comment tables
+What can be concluded from this first study is that kernel quantile regression equipped with the absolute laplacian kernel outperformed almost always its contenders.
+
+% kernel comparison
+\subsection{Kernel function choice}
+This section analyzes which kernel function best fits to the characteristics of the load data.
+The kernel functions considered are: Gaussian RBF, absolute Laplacian, Matern 1.5, Matern 2.5,Linear, Periodic, Polynomial, Sigmoid, and Cosine. 
+Details concerning the hyperparameter selection for the different kernels can be found in \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.metrics.pairwise}{sklearn.metrics.pairwise} and \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.gaussian_process}{sklearn.gaussian\_process}.
+The pool of data for this study has been create by combining \href{https://zenodo.org/records/7907883}{SECURES-Met} data (predictors) \cite{Formayer2023} and the load data from \href{https://www.energy-charts.info/index.html?l=en&c=DE}{energy charts}.
+SECURES-Met data consists of historical data up to the end of $2020$ while from 2021 onward, the data consists of forecasts modeled by the the European Centre for Medium-Range Weather Forecasts (ECMWF).
+Therefore, in this experimental study, we used the whole data of $2020$ as the training set and then tested our kernels on the $2021$ data.
+Note that there are two kinds of predictions for data from $2021$ onward, one for each of the emission scenarios RCP$4.5$ and RCP$8.5$.
+In this study, we considered the more conservative RCP$4.5$ scenario data.
+
+The predictors making up the dataset follow:
+\begin{itemize}
+    \item Direct irradiation: direct normal irradiation
+    \item Global radiation: mean global radiation
+    \item Hydro reservoir: daily mean power from reservoir plants in MW
+    \item Hydro river: daily mean power from run of river plants in MW
+    \item Temperature: air temperature
+    \item Wind potential: potential wind power production
+\end{itemize}
+Table \ref{tab:secures_met_de} shows results for Germany, table \ref{tab:secures_met_ch} is the one for Switzerland and table corresponds to Austria \ref{tab:secures_met_at}.
+
+\begin{table}[!ht]
+    \centering
+    \caption{Pinball loss kernel comparison for load in Switzerland (2021)}
+    \label{tab:secures_met_ch}
+    \begin{adjustbox}{width=\textwidth}
+        \begin{tabular}{llllllllll}
+    \toprule
+    Quantile & Gaussian RBF & Laplacian & Matern 1.5 & Matern 2.5 & Linear & Periodic & Polynomial & Sigmoid & Cosine \\
+    \midrule
+    0.1 & 0.018861 & \textbf{0.018798} & 0.019285 & 0.019510 & 0.019128 & 0.024474 & 0.021105 & 0.024299 & 0.018996 \\
+    0.2 & 0.030924 & \textbf{0.030430} & 0.031038 & 0.031438 & 0.031223 & 0.040274 & 0.033892 & 0.039774 & 0.031008 \\
+    0.3 & 0.039107 & \textbf{0.038510} & 0.039295 & 0.039751 & 0.039637 & 0.051023 & 0.042683 & 0.050242 & 0.039363 \\
+    0.4 & 0.044285 & \textbf{0.043718} & 0.044763 & 0.045143 & 0.045130 & 0.057785 & 0.048232 & 0.056815 & 0.044896 \\
+    0.5 & 0.046840 & \textbf{0.046261} & 0.047502 & 0.047957 & 0.047739 & 0.060671 & 0.051024 & 0.059697 & 0.047302 \\
+    0.6 & 0.046138 & \textbf{0.045582} & 0.047239 & 0.047686 & 0.047416 & 0.059627 & 0.050456 & 0.058824 & 0.046834 \\
+    0.7 & 0.041910 & \textbf{0.041615} & 0.043404 & 0.043964 & 0.043735 & 0.054688 & 0.046437 & 0.054037 & 0.042897 \\
+    0.8 & 0.034032 & \textbf{0.033852} & 0.035332 & 0.035997 & 0.035518 & 0.044591 & 0.037918 & 0.044112 & 0.034557 \\
+    0.9 & \textbf{0.021817} & 0.021923 & 0.022712 & 0.023082 & 0.022721 & 0.028070 & 0.024161 & 0.027920 & 0.022606 \\
+    \midrule
+    CRPS & 0.035990 & \textbf{0.035632} & 0.036730 & 0.037170 & 0.036916 & 0.046800 & 0.039545 & 0.046191 & 0.036495 \\
+    \bottomrule
+    \end{tabular}
+\end{adjustbox}
+    \end{table}
+
+\begin{table}[!ht]
+    \centering
+\caption{Pinball loss kernel comparison for load in German (2021)}
+\label{tab:secures_met_de}
+\begin{adjustbox}{width=\textwidth}
+    \begin{tabular}{llllllllll}
+\toprule
+Quantile & Gaussian RBF & Laplacian & Matern 1.5 & Matern 2.5 & Linear & Periodic & Polynomial & Sigmoid & Cosine \\
+\midrule
+0.1 & 0.028057 & \textbf{0.027483} & 0.028236 & 0.028257 & 0.028261 & 0.028277 & 0.028276 & 0.028146 & 0.028004 \\
+0.2 & 0.047118 & \textbf{0.046051} & 0.047531 & 0.047544 & 0.047569 & 0.047595 & 0.047596 & 0.047353 & 0.047025 \\
+0.3 & 0.060366 & \textbf{0.058739} & 0.060892 & 0.060940 & 0.060962 & 0.060988 & 0.060984 & 0.060654 & 0.060209 \\
+0.4 & 0.068344 & \textbf{0.066347} & 0.068952 & 0.069022 & 0.069054 & 0.069104 & 0.069101 & 0.068634 & 0.068202 \\
+0.5 & 0.071450 & \textbf{0.069118} & 0.072188 & 0.072250 & 0.072285 & 0.072331 & 0.072330 & 0.071807 & 0.071359 \\
+0.6 & 0.068821 & \textbf{0.067098} & 0.069454 & 0.069526 & 0.069571 & 0.069615 & 0.069613 & 0.069157 & 0.068780 \\
+0.7 & 0.061150 & \textbf{0.059972} & 0.061382 & 0.061410 & 0.061427 & 0.061442 & 0.061442 & 0.061281 & 0.060990 \\
+0.8 & 0.047147 & \textbf{0.046457} & 0.047325 & 0.047331 & 0.047340 & 0.047348 & 0.047349 & 0.047247 & 0.047056 \\
+0.9 & 0.026646 & 0.026343 & 0.026749 & 0.026757 & 0.026763 & 0.026772 & 0.026772 & \textbf{0.026282} & 0.026628 \\ 
+\midrule
+CRPS & 0.053233 & \textbf{0.051956} & 0.053634 & 0.053671 & 0.053693 & 0.053719 & 0.053718 & 0.053395 & 0.053139 \\
+\bottomrule
+\end{tabular}
+\end{adjustbox}
+\end{table}
+
+\begin{table}[!ht]
+    \centering
+\caption{Pinball loss kernel comparison for load in Austria (2021)}
+\label{tab:secures_met_at}
+\begin{adjustbox}{width=\textwidth}
+    \begin{tabular}{llllllllll}
+\toprule
+Quantile & Gaussian RBF & Laplacian & Matern 1.5 & Matern 2.5 & Linear & Periodic & Polynomial & Sigmoid & Cosine \\
+\midrule
+0.1 & 0.024309 & \textbf{0.023882} & 0.024316 & 0.024611 & 0.029367 & 0.029462 & 0.027319 & 0.029459 & 0.027816 \\
+0.2 & 0.041028 & \textbf{0.040318} & 0.040987 & 0.041487 & 0.048510 & 0.048716 & 0.046164 & 0.048714 & 0.045879 \\
+0.3 & 0.052597 & \textbf{0.051643} & 0.052548 & 0.052972 & 0.062242 & 0.062561 & 0.059133 & 0.062548 & 0.058454 \\
+0.4 & 0.058803 & \textbf{0.057676} & 0.058841 & 0.059078 & 0.070210 & 0.070577 & 0.067914 & 0.070567 & 0.066335 \\
+0.5 & 0.060846 & \textbf{0.059441} & 0.060912 & 0.061541 & 0.073399 & 0.073782 & 0.071616 & 0.073764 & 0.069574 \\
+0.6 & 0.059336 & \textbf{0.057820} & 0.059237 & 0.059930 & 0.070817 & 0.071157 & 0.070249 & 0.071134 & 0.066970 \\
+0.7 & 0.053937 & \textbf{0.052402} & 0.053949 & 0.054372 & 0.062453 & 0.062688 & 0.062970 & 0.062675 & 0.059419 \\
+0.8 & 0.043366 & \textbf{0.042427} & 0.043364 & 0.043510 & 0.049024 & 0.049150 & 0.049737 & 0.049147 & 0.046662 \\
+0.9 & 0.026727 & \textbf{0.026227} & 0.026645 & 0.026817 & 0.029516 & 0.029601 & 0.030097 & 0.029598 & 0.027986 \\
+\midrule
+CRPS & 0.046772 & \textbf{0.045760} & 0.046755 & 0.047146 & 0.055060 & 0.055299 & 0.053911 & 0.055290 & 0.052122 \\
+\bottomrule
+\end{tabular}
+\end{adjustbox}
+\end{table}
+From our tables, it can be concluded that the absolute Laplacian kernel is the one fitting best to the characteristics of loda data.
+Additionally, we can conclude that our results are in line with the ones published in \cite{he2017short}.
+In that study, similarly to what we did in this subsection, the authors carried out a comparitive study concerning the choice of kernels in KQR.
+Nevertheless notice, that the set of kernel we considered in our study is larger than the one presented in \cite{he2017short}.
+Their paper showed evidence of the superiority of the Gaussian RBF among their set of kernel functions.
+Analogously, we can see from tables \ref{tab:secures_met_de}, \ref{tab:secures_met_ch} and \ref{tab:secures_met_at} that the Gaussian RBF is the second best in terms of pinball loss.
+
+% gefcom final case
+\subsection{GEFCom2014}
+The follwing subsection compares our kernel quantile regression against the price and load track entries of the GEFCom2014 competiton.
+Error measure of the competition was the pinball loss, see section \ref{metrics}, averaged over the $99$ quantiles, $q \in \{i/100\}_{i=1}^{n=99}$.
+Backed by the conclusion from the previous subsection we selected the absolute Laplacian and the Gaussian RBF as kernels in our application of KQR.
+
+\subsubsection{GEFCom2014 price track}
+In this setting, the predictors fed to our KQR models are
+\begin{itemize}
+    \item Day
+    \item Hour
+    \item Forecasted total load
+    \item Forecasted zonal load
+\end{itemize}
+Since target days vary all over the year, we trained a model for any month associated to them.
+
+Our results for this track are reported in table \ref{tab:pinball loss gefcom2014 price data}.
+In this track, the top entries came from the teams: Tololo, Team Poland, GMD, and C3 Green Team; for a breakdown of each method's attributes, see \cite[Table 8]{hong2016probabilistic}.
+Finally, the Gaussian RBF probabilistic prediction for the $13$th July $2013$ zonal price at the $90\%$ confidence interval is visualized in Figure \ref{fig:price_task_6_gaussian_rbf}.
+
+\begin{figure}[!ht]
+    \centering
+    \includegraphics[width=\linewidth]{images/price_task_6_gaussian_rbf.png}
+    \caption[[Prediction price track task 6, Gaussian RBF kernel]]{Price 90\% confidence interval task 6: \small \textmd{Electricity price probabilistic forecast for the $13^{\text{th}}$ July $2013$.
+The black line is the observed path for the price.
+The $90\%$ confidence interval bands are plotted in green.
+Lower and upper red lines denote the $95\%$ and $5\%$ quantile forecast respectively.}}
+    % \Description{}
+    \label{fig:price_task_6_gaussian_rbf}
+\end{figure}
+
+\begin{table}[!ht]
+    \centering
+    \caption{Pinball loss GEFCom2014 price data}
+    \label{tab:pinball loss gefcom2014 price data}
+    \begin{adjustbox}{width=\textwidth}
+    \begin{tabular}{lllllllllllll}
+      \toprule
+      \midrule
+      Team name\textbackslash Task number                       & 1                               & 2                                  & 3                               & 4                              & 5                              & 6       & 7                               & 8       & 9       & 10                             & 11                             & 12               \\
+  KQR laplacian
+  &
+  \textbf{1.3249}
+  &
+  2.9497
+  &
+  0.8868
+  &
+  7.3780
+  &
+  5.4314
+  &
+  8.2166
+  &
+  4.9091
+  &
+  1.5245
+  &
+  1.9601
+  &
+  2.7324
+  &
+  2.4733
+  &
+  17.5371
+  \\
+  KQR gaussian rbf
+  &
+  2.0614
+  &
+  3.2247
+  &
+  \textbf{0.7050}
+  &
+  6.1407
+  &
+  6.1134
+  &
+  8.5969
+  &
+  2.9360
+  &
+  1.5731
+  &
+  1.8830
+  &
+  2.6419
+  &
+  2.4814
+  &
+  16.1921
+  \\
+  \\
+
+  Arkadiy Strelnikov         & 1,92899          & 3,46853          & 4,35373          & 7,16274          & 7,42959          & 6,32601          & 5,03883          & 2,02658          & \textbf{0,66422} & 2,35183          & 2,08614          & 7,23447          \\
+  Benchmark - Price          & 4,02875          & 7,97208          & 5,70395          & 12,15104         & 38,33541         & 44,22979         & 18,22395         & 31,56729         & 42,94958         & 2,85583          & 3,20395          & 22,38333         \\
+  C3 Green Team              & 1,85897          & 3,27786          & 1,2593           & 5,08886          & 6,87674          & 6,1505           & 4,42379          & 1,32639          & 1,25915          & 3,08224          & 1,55811          & 6,58123          \\
+  E.S.
+Mangalova             & 2,05693          & 7,97208          & 0,87971          & 7,04219          & 11,0464          & 6,57565          & 6,02388          & \textbf{0,69721} & 2,77446          & 2,78586          & 2,14229          & 7,35955          \\
+  EPSteam                    & 4,02875          & 7,97208          & 5,70395          & 3,46235          & 29,5226          & 27,73186         & 17,0324          & 2,49354          & 1,25639          & \textbf{1,79504} & 1,63482          & 10,357           \\
+  Florencio Gonzalez         & 4,02875          & 4,32597          & 4,35373          & 9,72491          & 10,51628         & 44,17405         & 4,00812          & 4,6924           & 7,36395          & 1,79865          & 2,1575           & 4,42008          \\
+  GMD                        & 3,7271           & 1,783            & 0,92191          & 5,08886          & 6,21331          & \textbf{3,82599} & 4,9342           & 1,47858          & 1,65933          & 2,06134          & 2,1235           & 6,84571          \\
+  Manuel Oviedo de la Fuente & 1,62605 & 3,27786          & 13,73529         & 6,68756          & 23,55608         & 10,01475         & 6,61296          & 2,2259           & 1,48028          & 3,65169          & 3,20396          & 4,7237           \\
+  NimNid                     & 3,44735          & 10,59679         & 2,76634          & 24,32776         & 23,55608         & 7,96082          & 3,33627          & 1,8191           & 1,60593          & 2,5711           & 2,32578          & 8,41167          \\
+  San/Saini                  & 2,47163          & 1,95533          & 0,84071 & 5,31786          & 9,61338          & 8,28547          & 3,0475           & 2,86903          & 3,60395          & 4,37704          & 1,82957          & 16,81896         \\
+  Team Poland                & 1,97477          & 1,81898          & 1,19162          & 2,82318          & 7,55914          & 4,20773          & 2,59715          & 1,04693          & 1,24193          & 4,06012          & \textbf{1,08458} & \textbf{3,06512} \\
+  Tololo                     & 1,70734          & \textbf{1,45173} & 1,10384          & \textbf{2,01694} & 9,15596          & 4,6821           & \textbf{1,59517} & 0,75352          & 2,45935          & 2,9614           & 1,34614          & 3,55819          \\
+  Xiaorong (Iris) Sun        & 1,45661          & 1,59212          & 0,98985          & 3,0349           & 4,73309          & 4,52459          & 3,63208          & 2,30481          & 0,90781          & 5,00935          & 1,18223          & 4,5302           \\
+  Yanghai Cong               & 1,69815          & 5,84             & 4,83261          & 3,17957          & 10,21729         & 6,44717          & 5,55374          & 3,80812          & 4,38485          & 1,45195          & 1,51106          & 14,61798         \\
+  dmlab                      & 2,30162          & 1,92804          & 1,2593           & 2,58646          & 14,09957         & 7,5589           & 4,13365          & 0,80748          & 1,5149           & 3,71509          & 3,43097          & 10,22129         \\
+  pat1                       & 2,36615          & 1,98567          & 1,07248          & 2,79465          & \textbf{4,23269} & 4,70614          & 8,40506          & 1,25376          & 2,23991          & 3,67952          & 1,06139          & 6,27517
+
+  \end{tabular}
+    \end{adjustbox}
+  \end{table}
+
+What can be concluded from table \ref{tab:pinball loss gefcom2014 price data} is that KQR performed consistently among the top algorithms.
+Furthermore, It ranked first in two out of the thirteen tasks of the price track.
+\subsubsection{GEFCom2014 load track}
+The GEFCom2014 load track constitutes a good setting for analyzing the performance of KQR in the context of medium term load forecasting.
+In this track we were faced with the challenge of predicting the load for the next month without the availability of weather temperature forecasts. 
+Therefore, the primary task was to first accurately predict the weather and temperatures, and then model the load accordingly.
+Since there were no attributes available for humidity or wind speed, we chose to predict weather temperatures by aggregating historical temperature data across different dimensions such as day, month, and hour.
+Then we proceeded with building KQR models for the load; we chose the following predictors
+\begin{itemize}
+    \item Day: the number of the day
+    \item Hour
+    \item Day of week: an ordinal categorical variable for the day of the week
+    \item Is holiday: a dummy variable for holidays
+    \item w avg: average of weather temperatures across all the $25$ stations
+\end{itemize}
+We built $12$ models, one for each month, with each task model trained on the historical data of the month associated with it.
+
+Table \ref{tab:pinball loss gefcom2014 load data} reports our results for the load track.
+The top teams for the load forecasting track were Tololo, Adada, Jingrui(Rain) Xie, OxMAth, E.S.
+Managalova, Ziel Florian, and Bidong Liu; for a breakdown of the attributes of each method, see \cite[Table 6]{hong2016probabilistic}.
+Finally, the $90\%$ confidence interval forecast by our model equipped with the absolute Laplacian kernel for task number $9$, that is the prediction for June, is reported in Figure \ref{fig:load_task_9}.
+Similarly to the price track, we have that KQR was among the best algorithms also in the load forecasting track.
+
+\begin{figure*}[!ht]
+    \centering
+    \makebox[\linewidth]{\includegraphics[width = 1.5\textwidth]{images/load_task_9_laplacian.png}}
+    \caption[Prediction load track task 9, Laplacian kernel]{Load 90\% confidence interval task 9 using KQR laplacian: \small \textmd{Electric load probabilistic forecast for June 2011.
+The black line is the observed path for the load.
+The 90\% confidence interval bands are plotted in green.
+Lower and upper red lines denote the 95\% and 5\% quantile forecast respectively.}}
+    \label{fig:load_task_9}
+\end{figure*}
+
+\begin{figure*}[!ht]
+    \centering
+    \makebox[\linewidth]{\includegraphics[width = 1.5\textwidth]{images/sub_boxplot_task_9_laplacian_.png}}
+    \caption{Boxplot for first two days of probabilistic forecast task 9 using KQR laplacian}
+    \label{fig:load_task_9}
+\end{figure*}
+
+\begin{table}[!ht]
+    \caption{Pinball loss GEFCom2014 load data}
+    \label{tab:pinball loss gefcom2014 load data}
+    \begin{adjustbox}{width=\textwidth}
+    \begin{tabular}{lllllllllllll}
+      \toprule
+      \midrule
+      Team name\textbackslash Task number                       & 1                               & 2                                  & 3                               & 4                              & 5                              & 6       & 7                               & 8       & 9       & 10                             & 11                             & 12                              \\
+
+  KQR laplacian
+  & 
+  12.0552
+  &
+  12.0573
+  &
+  8.8213
+  &
+  5.1721
+  &
+  6.9133
+  &
+  7.8080
+  &
+  11.5559
+  &
+  11.8250
+  &
+  6.8941
+  &
+  3.9680
+  &
+  7.4931
+  &
+  10.8869
+  \\
+  KQR gaussian rbf
+  & 
+  12.4665
+  &
+  12.1780
+  &
+  9.3326
+  &
+  5.1713
+  &
+  6.8703
+  &
+  6.7482
+  &
+  11.02473
+  &
+  11.93156
+  &
+  6.6019
+  &
+  4.3111
+  &
+  7.2207
+  &
+  10.8840
+  \\
+  \\
+  ACE                        & 12.1330                         & 14.4846                         & \textbf{7.3933}                          & 4.8213                         & 6.8048                         & 7.0566  & 9.5921                          & 11.6316 & 5.9859  & 5.0730 & \textbf{5.6028}                         & 8.9699                          \\
+  Adada                      & 10.5093                         & 10.0801                         & 7.6238                          & 4.7289                         & \textbf{5.3936}                         & 6.6242  & 8.0144                          & 11.1366 & 5.7779  & 3.6379                         & 7.0096                         & 8.9109                          \\
+  Alastair Muir              & 13.6118                         & 16.6750                         & 10.0759                         & 7.7078 & 9.5386 & 8.3615  & 9.8152                          & 13.1363 & 8.9715  & 5.4082                         & 8.5881 & 17.5325                         \\
+  Andrew Landgraf            & 14.3650                         & 10.1090                         & 8.4180                          & 6.2522                         & 7.2248                         & 11.1638 & 9.9403                          & 11.0204 & 5.6920  & 6.1176                         & 11.0677                        & 13.3985                         \\
+  Benchmark - Load           & 18.7384                         & 22.7585                         & 13.2163                         & 8.3626                         & 10.9162                        & 16.9937 & 13.4038                         & 17.3151 & 13.8374 & 6.4237                         & 10.9380                        & 34.0685                         \\
+  Bidong Liu                 & 16.4215                         & 11.8655                         & 9.3733                          & 5.6212                         & 7.7387                         & 6.5536  & 9.1406                          & 11.3485 & 6.5096  & 4.8031                         & 6.9697                         & 10.8935                         \\
+  C3 Green Team              & 18.7384 & 19.2208 & 7.9637                          & 4.6370                         & 6.4543                         & 8.3799  & 10.5546                         & 10.6609 & 5.8867  & 4.4866                         & 5.9396                         & 10.3917                         \\
+  Christopher Benfield       & 18.7384 & 11.1324                         & 9.4377                          & 4.9097                         & 7.4184                         & 19.7325 & 9.2215                          & 11.4385 & 6.6395  & 4.5966                         & 6.5002                         & 10.8633                         \\
+  Dao Vu                     & 33.3711                         & 13.3340                         & 10.4300                         & 6.0815                         & 9.0706                         & 8.7098  & 11.2808 & 18.4869 & 7.4065  & 5.7414                         & 10.3431                        & 23.5659                         \\
+  E.S.
+Mangalova             & 18.7384 & 13.3340 & 7.8025                          & \textbf{4.4096}                         & 6.6330                         & 6.2306  & 10.1511                         & 10.9294 & 6.2224  & 4.2382                         & 6.5464                         & 8.8080                          \\
+  Jingrui (Rain) Xie         & 11.8700                         & 10.9250                         & 8.4938                          & 4.9611                         & 7.4442                         & 6.9921  & 9.0523                          & 11.2600 & 5.4864  & \textbf{3.3602}                         & 5.9011                         & 9.7316                          \\
+  Manuel Oviedo de la Fuente & 12.5502                         & 21.4591                         & 10.1593                         & 5.4647                         & 9.3072                         & 7.6102  & 8.8361                          & 12.6340 & 12.3969 & 15.5225                        & 6.5221                         & 18.7754                         \\
+  OxMath                     & 14.4091 & \textbf{8.9136}                          & 7.6059                          & 4.4548                         & 7.2944                         & 7.4551  & 7.9527                          & \textbf{10.2444} & 5.4551  & 4.2111                         & 6.4054                         & 9.5520                          \\
+  Rasmus Paivarinta          & 18.7384 & 11.7474                         & 9.7230                          & 7.7078                         & 8.3074 & 7.1793  & 8.4391                          & 10.9357 & 6.3855  & 4.3796                         & 6.3794                         & 11.7871                         \\
+  SAOR                       & 18.7384 & 21.4591 & 10.4300 & 6.8512                         & 8.6899                         & 6.7653  & 12.1427                         & 11.6261 & 6.7310  & 6.2850 & 5.9699                         & 21.2711                         \\
+  Sniper                     & 18.7384 & 10.3056                         & 8.3493                          & 5.6639                         & 8.3074                         & 6.9599  & 10.8848                         & 13.4937 & 7.8877  & 5.3514                         & 6.4770                         & 11.1951                         \\
+  Tololo                     & \textbf{10.4369}                         & 12.5232                         & 8.2695                          & 4.4220                         & 5.8976                         & \textbf{6.1878}  & \textbf{7.3182}                          & 10.8032 & \textbf{5.4469}  & 3.9613                         & 6.3173                         & \textbf{8.4787}                          \\
+  Trevor Maynor              & 18.6934                         & 19.4422                         & 12.6071                         & 5.5631                         & 10.9298                        & 13.9594 & 10.8657                         & 14.4647 & 7.2859  & 6.2850                         & 8.5881                         & 17.4703                         \\
+  Xiaorong (Iris) Sun        & 13.8779                         & 12.3979                         & 10.3672                         & 8.3626 & 8.1979                         & 20.5364 & 11.2808                         & 12.6316 & 6.6821  & 4.4907                         & 6.2363                         & 10.1624                         \\
+  dmlab                      & 14.4091                         & 14.0059                         & 9.2128                          & 5.9099                         & 6.6675                         & 6.9046  & 10.0130                         & 11.0201 & 7.8508  & 3.8128                         & 6.6474                         & 26.6655                         \\
+  nikolina                   & 18.7384 & 19.2208                         & 10.6214                         & 5.8257                         & 9.5386                         & 7.1234  & 10.0948                         & 12.5375 & 5.8139  & 4.8060                         & 8.4956                         & 21.2711 \\
+  pat1                       & 19.3898                         & 10.5097                         & 9.1477                          & 9.7023                         & 7.0428                         & 7.8794  & 8.8610                          & 12.8179 & 5.5443  & 5.0730                         & 6.5846                         & 11.2142
+  \end{tabular}
+  \end{adjustbox}
+  \end{table}
+
 % In this subsection we compare performance of quantile regressors.
@@
 \section{Conclusions}
+% repeat the problem description  and motivation quickly
+This section concludes this thesis work with some final remarks.
+This thesis was concerned with electricity forecsating, in particular we focused on the probabilistic framework.
+Probabilistic forecasting is still a relatively new research area, lately research efforts are shifting towards this setting.
+This is due to the fact that probabilistic forecasts are more informative than point ones and better suited for the current and future electricity landscape; think for example about the higher uncertainty resulting from market liberalization or renewables integration requirements.
+
+Among the contributions of this thesis work we can list
+% list contributions:
+% - study of kernel methods in EF, both point and probabilistic
+% - contribution to less developed area of EF, which is probabilistic forecasting
+% - provide open source python package for KQR PyPi, github repo, reproducibility purposes
+% - study case for DACH region, along with cleaned dataset for easing collaboration and replication in research
+% - extensive comparison among kernel types
+% - comparison against state of the art algorithms from GEFCom2014, particularly KQR in medium term load forecasting which has not been deeply analyzed in the literature.
+\begin{itemize}
+    \item Study of kernel methods in electricity forecasting both in the point and probabilistic framework.
+    \item Provided Python open source implementation for kernel quantile regression compatible with the sklearn API.
+The code has been packaged and uploaded to the Python Package Index (PyPI) with the name \href{https://pypi.org/project/kernel-quantile-regression/#2}{kernel-quantile-regression}.
+The \href{https://github.com/luca-pernigo/kernel_quantile_regression}{github repo} hosting the source code includes also the script implementing the experiments along with the cleaned datasets; this contribution is intended to forster reproducibility in research.
+    \item Achieved superior performance of kernel quantile regression compared to standard quantile regressor algorithm. 
+    \item Created and made available datasets suitable for algorithms benchmarking considering data from the DACH region (Germany, Switzerland and Austria).
+The format of these data takes inspiration from the popular GEFCom competitions.
+    \item Kernel quantile regression extensive comparison between kernel function types.
+    \item Compared kernel quantile regression against state of the art probabilistic algorithms in the literature by means of the GEFCom2014 competition.
+\end{itemize}
+
+% results achieved:
+% - KQR beats other quantile regressor algorithms
+% - Laplacian is the best kernel
+% - our KQR implementations demostrated good performance compared to the top entries of the GEFCom2014 competition, this suggests it is a valid method.
+We observe that KQR shows favorably and can forecast both the short and medium-term horizon.
+To conclude, we proceed with summarizing the conclusion we drew from our experimental analysis.
+Kernel methods showed good performance both in the context of point and probabilistic forecasting.
+Specifically, kernel quantile regression resulted almost always in superior performance compared to standard quantile regressors.
+The absolute Laplacian kernel turned out to be the most suitable kernel according to the characteristic of load.
+Finally, we proved the validity of our kernel quantile regression implementation.
+We achieved this by showing how it compared favourably to the top entries of the GEFCom2014 competition.
+In particular, up to nowx research work focused on the short term horizon capabilities of kernel quantile regression.
+In this thesis work, we were also able to show its suitability on the medium term horizon.

### 2024-05-19 thesis/input_files/data_exploratory.tex
May 19th 2024

@@
 \\
-\subsection{GEFCom2014 Price track}
+\subsection{GEFCom2014 price track}
 In the price forecasting track, the goal was to predict the electricity price for the next 24 hours of a single zone. 
@@
 
-\subsection{Load track}
+\subsection{GEFCom2014 load track}
 In the load track, contestant were asked to provide one month ahead hourly probabilistic forecasts on a rolling basis for 15 consecutive rounds.

### 2024-05-19 thesis/input_files/implementation.tex
May 19th 2024

@@
 \\
-In the fit method, we set up and solve the convex optimisation problem by using the cvxopt library, see offical manual \cite{vandenberghe2010cvxopt} for a reference.
+In the fit method, we set up and solve the convex optimisation problem through the interior point method.
+This algorithm is taken from the cvxopt library, see its offical manual \cite{vandenberghe2010cvxopt} for a reference.
 When using this library, it is important to keep two things in mind.

### 2024-05-21 thesis/abstract.tex
May 21st 2024

--- 2024-02-02 thesis/abstract.tex
@@
 \begin{abstract}
-    The theory of kernel methods will be applied to the problem of probabilistic forecasting taking as input some data.
-    Conceptually, the proposed analysis could be applied to any kind of data. In this study we considered the electricity market, because of the interesting implications  on risk management tasks.
+  % introduction
+  This thesis is concerned with electricity forecasting. 
+  Specifically, we concentrated on the probabilistic framework.
+This choice was motivated by its importance in systems planning and operations as a consequence of the inception of competitive power markets, smart grids and renewable integration requirements.
+  % methods
+  In doing so, we focused on the family of kernel methods. In addition, we also compared them against several statistical and machine learning techniques.
+  % results
+  Our results showed the goodness of kernel methods in the field of electricity forecasting both in terms of point and probabilistic forecasting.
+In particular, in the probabilistic context, our experiments showed the validity of kernel quantile regression equipped with the absolute Laplacian kernel.
+  % discussion
+  These findings indicate kernel methods are well suited to the characteristic of electricity.
+  Anyone interested in forecasting energy should consider them when faced with the choice of the model.
+  They can be employed stand alone or combined with other valid methods into ensembles.
 
@@
 \end{abstract}
+  % The theory of kernel methods will be applied to the problem of point and probabilistic forecasting for the energy sector.
+  %   Such choice is motivated by its interesting implications.

### 2024-05-21 thesis/input_files/analysis.tex
May 21st 2024

@@
 We added the weather temperature, the square of it and the categorical variable for holidays effect as regressors.
-Furthermore, we also applied a log transformation to the dependet variable.
+Furthermore, we also applied a log transformation to the dependent variable.
 Doing so, RMSE went down to 10.29.
-Forecast is visualized in \ref{fig:prophet_price2}, moreover.
-% figure \ref{fig:prophet_price2.1} and figure \ref{fig:prophet_price2.2} break down the trend and the seasonalities respectively.
+Forecast is visualized in \ref{fig:prophet_price2}.
+% , moreover figure \ref{fig:prophet_price2.1} and figure \ref{fig:prophet_price2.2} break down the trend and the seasonalities respectively.
 \begin{figure}[!h]
     \includegraphics[width=\textwidth]{images/prophet_price2.png}
-    \caption{Prophet v2 prediction}
+    \caption{Prophet model\_2 prediction}
     \label{fig:prophet_price2}
@@
 We used the torch library in order to code our LSTM regressor.
-Various hyperparameters combinations were tried during hyperparameter tuning. The final hyperparameters we set on follows.
+Various hyperparameters combinations were tried during the tuning phase. The final architecture we set on follows
 \begin{itemize}
@@
 \end{itemize}
-LSTM forecast is reported in table \ref{fig:lstm_price}, it achieves a RMSE of 9.1782
+LSTM forecast is reported in figure \ref{fig:lstm_price}, it achieves a RMSE of 9.1782
 \begin{figure}[!h]

### 2024-05-21 thesis/input_files/implementation.tex
May 21st 2024

@@
-This section is inteneded to explain and aid for reproducibilty studies. Hereafter the specific libraries used and the custom implementatios are thoroughly documented.
+This section is inteneded to explain and aid for reproducibilty studies. Hereafter, the specific libraries used and the custom implementatios are thoroughly documented.
 \\
@@
 
-Section documenting code
+% Section documenting code
 
-- Explain how methods' implementation has been
-adapted to my specific setting.
+% - Explain how methods' implementation has been
+% adapted to my specific setting.
 
-- Explain in detail how to my src code has been implemented
-its rationale and how to use it.
+% - Explain in detail how to my src code has been implemented
+% its rationale and how to use it.
 
-- As I explain code scripts go over the test, to 
-explain better my ideas.
+% - As I explain code scripts go over the test, to 
+% explain better my ideas.
 
-- Indicate also hyperparameters maybe in each subsection
+% - Indicate also hyperparameters maybe in each subsection
 
@@
 
-\subsection{Tbats}
+\subsection{Trigonometric seasonality Box-Cox transformation ARMA errors trend and seasonal components}
 Tbats implementation is available at \url{https://github.com/intive-DataScience/tbatsx}.
@@
 
-\subsection{LSTM}
+\subsection{Long short term memory}
 The LSTM predictor has been built using the \href{https://pytorch.org}{torch} library.
@@
 \subsection{Kernel support vector regression}
-The object SVR of the \href{https://scikit-learn.org/stable/}{sklearn} module svm has been used by specifying rbf as the kernel parameter.
+The object SVR of the \href{https://scikit-learn.org/stable/}{sklearn} module svm has been used.
+% by specifying rbf as the kernel parameter.
 

### 2024-05-21 experiments/qr_abalone.py
dirclean

@@ -5,7 +5,7 @@ from qr import qr_tests
 
 if __name__=="__main__":
     # load data
-    df=pd.read_csv("/Users/luca/Desktop/ThesisKernelMethods/dataset_applications/abalone.csv", sep=",", decimal=".")
+    df=pd.read_csv("/Users/luca/Desktop/ThesisKernelMethods/extradata/abalone.csv", sep=",", decimal=".")
 
     # map categorical regressor to integer form
     key = {"I":0, "M":1, "F":2}

### 2024-05-21 experiments/qr_boston.py

@@ -5,7 +5,7 @@ from qr import qr_tests
 
 if __name__=="__main__":
     # load data
-    df=pd.read_csv("/Users/luca/Desktop/ThesisKernelMethods/dataset_applications/BostonHousing.csv", sep=",", decimal=".")
+    df=pd.read_csv("/Users/luca/Desktop/ThesisKernelMethods/extradata/BostonHousing.csv", sep=",", decimal=".")
 
     X=df.iloc[:,0:-1]
     y=df.iloc[:,-1]

### 2024-05-21 experiments/qr_cars.py

@@ -5,7 +5,7 @@ from qr import qr_tests
 
 if __name__=="__main__":
     # load data
-    df=pd.read_csv("/Users/luca/Desktop/ThesisKernelMethods/dataset_applications/car_data.csv", sep=",", decimal=".")
+    df=pd.read_csv("/Users/luca/Desktop/ThesisKernelMethods/extradata/car_data.csv", sep=",", decimal=".")
 
     # map categorical regressor to integer form
     key6 = {"Petrol":0, "Diesel":1, "CNG":2}

### 2024-05-21 experiments/qr_melbourne.py

@@ -25,7 +25,7 @@ from tqdm import tqdm
 
 if __name__=="__main__":
     # load data
-    df=pd.read_csv("/Users/luca/Desktop/ThesisKernelMethods/dataset_applications/temperatures_melbourne.csv", sep=";", decimal=".")
+    df=pd.read_csv("/Users/luca/Desktop/ThesisKernelMethods/extradata/temperatures_melbourne.csv", sep=";", decimal=".")
 
     # quantiles
     quantiles = [0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9]

### 2024-05-21 experiments/gefcom_load/bash1.sh
dirclean

@@ -1,12 +1,0 @@
-#!/bin/bash
-# code for moving test data in respective folder
-# train data of previous task is the tail of the test data for current task
-for((i=2;i<=15;i++))
-do
-    i_minus_1=$((i - 1))
-    i_minus_1=$((i-1))
-    # get test data
-    cp "/Users/luca/Desktop/GEFCom2014 Data/Load/Task ${i}/L${i}-train.csv" "/Users/luca/Desktop/GEFCom2014 Data/Load/Task ${i_minus_1}/L${i_minus_1}-test.csv"
-done
-
-cp "/Users/luca/Desktop/GEFCom2014 Data/Load/Solution to Task 15/solution15_L_temperature.csv" "/Users/luca/Desktop/GEFCom2014 Data/Load/Task 15/L15-test.csv"

### 2024-05-21 experiments/gefcom_load/bash2.sh

@@ -1,8 +1,0 @@
-#!/bin/bash
-
-# for each task clean test dataset
-for ((i=1;i<=15;i++))
-do
-    python /Users/luca/Desktop/ThesisKernelMethods/dataset_applications/gefcom_clean_test.py $i
-
-done

### 2024-05-21 experiments/gefcom_load/bash3.sh

@@ -1,11 +1,0 @@
-#!/bin/bash
-# script to apply pickled model to each task
-
-# clean
-rm /Users/luca/Desktop/ThesisKernelMethods/thesis/tables/qr_gefcom2014.txt
-
-for((i=1;i<=15;++i))
-do
-    python /Users/luca/Desktop/ThesisKernelMethods/dataset_applications/qr_gefcom2014_test.py $i
-
-done

### 2024-05-21 experiments/gefcom_load/bash4.sh

@@ -1,9 +1,0 @@
-#!/bin/bash
-
-# code to extract the timestamp and zoneid from test data so that the prediction is in the same format of guidelines
-
-
-for((i=1;i<=15;++i))
-do
-    cut -d , -f 1-2 /Users/luca/Desktop/GEFCom2014\ Data/Load/Task\ $i/L$i-test.csv >> /Users/luca/Desktop/GEFCom2014\ Data/Load/Task\ $i/L$i-timestamp.csv
-done

### 2024-05-21 experiments/gefcom_load/gefcom_clean_test.py

@@ -1,91 +1,0 @@
-import pandas as pd
-import sys
-
-import holidays
-
-def clean_test(ith):
-    if ith==15:
-        df=pd.read_csv(f"/Users/luca/Desktop/GEFCom2014 Data/Load/Task {ith}/L{ith}-test.csv")
-        # dataset 15 does not stick to convention used so far in the competition, thus we have an if conditional to handle it
-        df['TIMESTAMP'] = pd.to_datetime(df['date'], format='%m/%d/%Y')
-
-        # to be coherent hours go from 0 to 23
-        df['HOUR'] =df["hour"].subtract(1)
-
-        df['DAY'] = df['TIMESTAMP'].dt.day
-        df['MONTH'] = df['TIMESTAMP'].dt.month
-        df['YEAR'] = df['TIMESTAMP'].dt.year
-
-        # add info weekday; saturday, sunday
-        df['DAY_OF_WEEK'] = df['TIMESTAMP'].dt.dayofweek
-
-        # holidays
-        hol=holidays.US(years=range(2001, 2011))
-        df["IS_HOLIDAY"]=df['TIMESTAMP'].isin(hol)
-        # convert it to categorical variable
-        df["IS_HOLIDAY"]=pd.Categorical(df["IS_HOLIDAY"].astype(int))
-
-        # drop not useful columns
-        df.drop(columns=['TIMESTAMP', 'date', 'hour'], inplace=True)
-
-
-        # average temperatures
-        weather_stat=df.filter(regex=("w.*")).columns.to_list()
-        # average them 
-        df['w_avg'] = df[weather_stat].mean(axis=1)
-
-        # drop columns
-        df.drop(columns=weather_stat, inplace=True)
-        # df.drop(columns=["YEAR"], inplace=True)
-        df["DS"]=pd.to_datetime(df[["YEAR","MONTH","DAY","HOUR"]] , format='%Y%m%d')
-
-
-        df=df[["LOAD","DAY","MONTH","HOUR","DAY_OF_WEEK","IS_HOLIDAY","w_avg","DS"]]
-        
-        # save to csv
-        df.to_csv(f"/Users/luca/Desktop/GEFCom2014 Data/Load/Task {ith}/L{ith}-test_clean.csv", index=False)
-
-        # print(df)
-
-    else:    
[…]

### 2024-05-21 experiments/gefcom_load/gefcom_clean_test2.py

@@ -1,88 +1,0 @@
-import pandas as pd
-import sys
-
-import holidays
-
-def clean_test(ith):
-    if ith==15:
-        df=pd.read_csv(f"/Users/luca/Desktop/GEFCom2014 Data/Load/Task {ith}/L{ith}-test.csv")
-        # dataset 15 does not stick to convention used so far in the competition, thus we have an if conditional to handle it
-        df['TIMESTAMP'] = pd.to_datetime(df['date'], format='%m/%d/%Y')
-
-        # to be coherent hours go from 0 to 23
-        df['HOUR'] =df["hour"]-1
-
-        df['DAY'] = df['TIMESTAMP'].dt.day
-        df['MONTH'] = df['TIMESTAMP'].dt.month
-        df['HOUR'] = df['TIMESTAMP'].dt.hour
-        df['YEAR'] = df['TIMESTAMP'].dt.year
-
-        # add info weekday; saturday, sunday
-        df['DAY_OF_WEEK'] = df['TIMESTAMP'].dt.dayofweek
-
-        # holidays
-        hol=holidays.US(years=range(2001, 2011))
-        df["IS_HOLIDAY"]=df['TIMESTAMP'].isin(hol)
-        # convert it to categorical variable
-        df["IS_HOLIDAY"]=pd.Categorical(df["IS_HOLIDAY"].astype(int))
-
-
-        # drop not useful columns
-        df.drop(columns=['TIMESTAMP', 'date', 'hour'], inplace=True)
-
-
-        # average temperatures
-        weather_stat=df.filter(regex=("w.*")).columns.to_list()
-        # average them 
-        df['w_avg'] = df[weather_stat].mean(axis=1)
-
-        # drop columns
-        df.drop(columns=["YEAR"], inplace=True)
-
-        df=df[["LOAD","DAY","MONTH","HOUR","DAY_OF_WEEK","IS_HOLIDAY"]+[[f"w{i}"] for i in range(1,25)]]
-        
-        # save to csv
-        df.to_csv(f"/Users/luca/Desktop/GEFCom2014 Data/Load/Task {ith}/L{ith}-test_clean_2.csv", index=False)
-
-        # print(df)
-
-    else:    
-        df=pd.read_csv(f"/Users/luca/Desktop/GEFCom2014 Data/Load/Task {ith}/L{ith}-test.csv")
-
[…]

### 2024-05-21 experiments/gefcom_load/gefcom_clean_train.py

@@ -1,60 +1,0 @@
-import numpy as np
-import pandas as pd
-from sklearn.metrics import mean_pinball_loss
-
-from qr import qr_tests
-
-import holidays
-
-
-if __name__=="__main__":
-    # load data
-    df=pd.read_csv("/Users/luca/Desktop/GEFCom2014 Data/Load/Task 1/L1-train.csv", sep=",", decimal=".")
-    
-    # clean data dates
-    dates = pd.date_range(start='2001-01-01 01:00:00', periods=len(df), freq='H')
-    df['DAY'] = dates.day
-    df['MONTH'] = dates.month
-    df['HOUR'] = dates.hour
-    df['YEAR'] = dates.year
-    
-    # add info weekday; saturday, sunday
-    df['DAY_OF_WEEK'] = dates.dayofweek
-
-    # holidays
-    hol=holidays.US(years=range(2001, 2011))
-    df["IS_HOLIDAY"]=pd.Series(dates.date).isin(hol)
-    # convert it to categorical variable
-    df["IS_HOLIDAY"]=pd.Categorical(df["IS_HOLIDAY"].astype(int))
-    
-
-    # drop not useful columns
-    # zoneid is dropped because are all 1, so it is not a meaningful independent variable
-    # the information in timestamp has been decomposed in four columns so we can drop it
-    df.drop(columns=['ZONEID', 'TIMESTAMP'], inplace=True)
-
-
-    # slice for training, from when we have load informations 2005
-    # slice from 2009 so it is less computations heavy
-    idx_start = df.loc[(df['DAY'] == 1) & (df["MONTH"]==1) & (df["YEAR"]==2009) & (df["HOUR"])==1].index[0]
-    
-    
-    idx_end = df.loc[(df['DAY'] == 1) & (df["MONTH"]==1) & (df["YEAR"]==2010) & (df["HOUR"])==1].index[0]
-
-
-    df = df.loc[idx_start-1:idx_end-1]
-
-    # average temperatures
-    # select weather stations data using regex
-    weather_stat=df.filter(regex=("w.*")).columns.to_list()
-    # average them 
-    df['w_avg'] = df[weather_stat].mean(axis=1)
-
-
-    df.drop(columns=weather_stat, inplace=True)
-    df.drop(columns=["YEAR"], inplace=True)
-
-    df.to_csv("/Users/luca/Desktop/GEFCom2014 Data/Load/L2009-train.csv",index=False)
-
-    
-    

### 2024-05-21 experiments/gefcom_load/gefcom_clean_train2.py

@@ -1,53 +1,0 @@
-import numpy as np
-import pandas as pd
-from sklearn.metrics import mean_pinball_loss
-
-from qr import qr_tests
-
-import holidays
-
-
-if __name__=="__main__":
-    # load data
-    df=pd.read_csv("/Users/luca/Desktop/GEFCom2014 Data/Load/Task 1/L1-train.csv", sep=",", decimal=".")
-    
-    # clean data dates
-    dates = pd.date_range(start='2001-01-01 01:00:00', periods=len(df), freq='H')
-    df['DAY'] = dates.day
-    df['MONTH'] = dates.month
-    df['HOUR'] = dates.hour
-    df['YEAR'] = dates.year
-    
-    # add info weekday; saturday, sunday
-    df['DAY_OF_WEEK'] = dates.dayofweek
-
-    # holidays
-    hol=holidays.US(years=range(2001, 2011))
-    df["IS_HOLIDAY"]=pd.Series(dates.date).isin(hol)
-    # convert it to categorical variable
-    df["IS_HOLIDAY"]=pd.Categorical(df["IS_HOLIDAY"].astype(int))
-    
-
-    # drop not useful columns
-    # zoneid is dropped because are all 1, so it is not a meaningful independent variable
-    # the information in timestamp has been decomposed in four columns so we can drop it
-    df.drop(columns=['ZONEID', 'TIMESTAMP'], inplace=True)
-
-
-    # average temperatures
-    # select weather stations data using regex
-    weather_stat=df.filter(regex=("w.*")).columns.to_list()
-    # average them 
-    df['w_avg'] = df[weather_stat].mean(axis=1)
-
-    
-    df=df[df["YEAR"]>=2008]
-    df=df[df["YEAR"]<2010]
-    # drop first row
-    # df.drop(df.index[0], inplace=True)
-    df.drop(columns=["YEAR"], inplace=True)
-
-    # L-train_clean.csv
-    df.to_csv("/Users/luca/Desktop/GEFCom2014 Data/Load/Task 1/L-train_clean_full_2.csv",index=False)
-
-    

### 2024-05-21 experiments/gefcom_load/models_conditioned_months/kqr.py

@@ -1,283 +1,0 @@
-
-import cvxopt as opt
-from cvxopt.solvers import qp, options
-from cvxopt import matrix, spmatrix, sparse
-from cvxopt import blas
- 
-import matplotlib.pyplot as plt
-import numpy as np
-from numpy import asarray
-import pandas as pd
-from pprint import pprint
-from operator import itemgetter
-
-from scipy.stats import uniform
-from sklearn.base import BaseEstimator, RegressorMixin
-
-from sklearn.experimental import enable_halving_search_cv
-
-from sklearn.metrics import make_scorer
-from sklearn.metrics import mean_pinball_loss
-from sklearn.metrics.pairwise import rbf_kernel
-
-from sklearn.model_selection import HalvingRandomSearchCV
-from sklearn.model_selection import GridSearchCV
-from sklearn.model_selection import train_test_split
-
-from sklearn.preprocessing import StandardScaler
-from sklearn.preprocessing import MinMaxScaler
-from sklearn.preprocessing import RobustScaler
-
-from sklearn.utils.validation import check_X_y, check_array, check_is_fitted
-from sklearn.utils.multiclass import unique_labels
-
-import rpy2.robjects as robjects
-from rpy2.robjects.packages import importr
-
-import sys
-import time
-from tqdm import tqdm
-
-
-class KQR(RegressorMixin, BaseEstimator):
-    """ Code implementing kernel quantile regression.
-
-    Parameters
-    ----------
-    alpha : int, default='0.5'
-        quantile under consideration
-
-    gamma : int, default='1'
-        Bandwith of rbf kernel
-
-    C : int, default='0.5'
-        the cost regularization parameter. This parameter controls the smoothness of the fitted function, essentially higher values for C lead to less smooth functions
-
-    Attributes
-    ----------
-    X_ : ndarray, shape (n_samples, n_features)
-        The input passed during :meth:`fit`.
-    y_ : ndarray, shape (n_samples,)
-        The dependent variable, our target :meth:`fit`.
-    """
-    def __init__(self, alpha, C=1, gamma=1):
-        self.gamma = gamma
-        self.C=C
-        self.alpha=alpha
-
-    def fit(self, X, y):
[…]

### 2024-05-21 experiments/gefcom_load/models_conditioned_months/qr_gefcom2014.py

@@ -1,68 +1,0 @@
-import numpy as np
-import pandas as pd
-import pickle
-from pprint import pprint
-from sklearn.experimental import enable_halving_search_cv  # noqa
-from sklearn.model_selection import HalvingRandomSearchCV
-from sklearn.metrics import make_scorer
-from sklearn.metrics import mean_absolute_error
-from sklearn.metrics import mean_pinball_loss
-from sklearn.model_selection import train_test_split
-from sklearn.preprocessing import StandardScaler
-import sys
-from tqdm import tqdm
-
-from kqr import KQR
-
-from cvxopt import matrix, spmatrix, sparse
-
-
-
-
-if __name__=="__main__":
-    # load data
-    df=pd.read_csv("/Users/luca/Desktop/GEFCom2014 Data/Load/L-train.csv")
-    df=df[df["MONTH"]==1]
-    X_train=df[["DAY",  "MONTH",  "HOUR",  "DAY_OF_WEEK",  "IS_HOLIDAY",  "w_avg"]]
-    y_train=df["LOAD"]
-
-    # 99 quantiles
-    quantiles = [i/100 for i in range(1,100)]
-
-    # define loss to tune
-    neg_mean_pinball_loss_scorer_05 = make_scorer(
-    mean_pinball_loss,
-    alpha=0.5,
-    greater_is_better=False,  # maximize the negative of the loss
-    )
-
-    # kernel quantile regression
-    scaler = StandardScaler()
-
-    X_train_scaled = scaler.fit_transform(X_train)
-
-    qr_krn_models=[]
-    y_test_pred_qr_krn=[]
-
-    param_grid_krn = dict(
-    C=[1e-1,1e-2,1, 5, 10,1e2,1e4],
-    gamma=[1e-1,1e-2,0.5,1, 5, 10, 20]   
-    )
-    
-    krn_blueprint=KQR(alpha=0.5)
-    best_hyperparameters_krn=HalvingRandomSearchCV(
-            krn_blueprint,
-            param_grid_krn,
-            scoring=neg_mean_pinball_loss_scorer_05,
-            n_jobs=2,
-            random_state=0,
-        ).fit(X_train_scaled, y_train).best_params_
-    
-    for i,q in enumerate(tqdm(quantiles)):
-
-        # fit data for specific quantile
-        qr_krn_models+=[KQR(alpha=q, **best_hyperparameters_krn).fit(X_train_scaled, y_train)]
-
[…]

### 2024-05-21 experiments/gefcom_load/models_conditioned_months/qr_gefcom2014_cv.py

@@ -1,70 +1,0 @@
-import numpy as np
-import pandas as pd
-import pickle
-from pprint import pprint
-from sklearn.experimental import enable_halving_search_cv  # noqa
-from sklearn.model_selection import HalvingRandomSearchCV
-from sklearn.metrics import make_scorer
-from sklearn.metrics import mean_absolute_error
-from sklearn.metrics import mean_pinball_loss
-from sklearn.model_selection import train_test_split
-from sklearn.preprocessing import StandardScaler
-import sys
-from tqdm import tqdm
-
-from kqr import KQR
-
-from cvxopt import matrix, spmatrix, sparse
-
-
-
-
-if __name__=="__main__":
-    ith=15
-    # load data
-    df=pd.read_csv("/Users/luca/Desktop/GEFCom2014 Data/Load/L-train.csv")
-    df=df[df["MONTH"]==(ith-3)][-1000:]
-    X_train=df[["DAY",  "MONTH",  "HOUR",  "DAY_OF_WEEK",  "IS_HOLIDAY",  "w_avg"]]
-    y_train=df["LOAD"]
-
-    # 99 quantiles
-    quantiles = [i/100 for i in range(1,100)]
-
-    # kernel quantile regression
-    scaler = StandardScaler()
-
-    X_train_scaled = scaler.fit_transform(X_train)
-
-    qr_krn_models=[]
-    y_test_pred_qr_krn=[]
-
-    param_grid_krn = dict(
-    C=[1e-1,1e-2,1, 5, 10,1e2,1e4],
-    gamma=[1e-1,1e-2,0.5,1, 5, 10, 20]   
-    )
-    
-    
-    for i,q in enumerate(tqdm(quantiles)):
-        
-        # define loss to tune
-        neg_mean_pinball_loss_scorer = make_scorer(
-        mean_pinball_loss,
-        alpha=q,
-        greater_is_better=False,  # maximize the negative of the loss
-        )
-
-        krn_blueprint=KQR(alpha=q)
-        best_hyperparameters_krn=HalvingRandomSearchCV(
-                krn_blueprint,
-                param_grid_krn,
-                scoring=neg_mean_pinball_loss_scorer,
-                n_jobs=2,
-                random_state=0,
-            ).fit(X_train_scaled, y_train).best_params_
-
-        # fit data for specific quantile
[…]

### 2024-05-21 experiments/gefcom_load/models_conditioned_months/qr_gefcom2014_test.py

@@ -1,54 +1,0 @@
-
-import pandas as pd
-import pickle
-
-from sklearn.metrics import mean_absolute_error
-from sklearn.metrics import mean_pinball_loss
-from sklearn.preprocessing import StandardScaler
-import sys
-
-from tqdm import tqdm
-
-
-def test(ith):
-    df=pd.read_csv(f"/Users/luca/Desktop/GEFCom2014 Data/Load/Task {ith}/L{ith}-avg_test_clean.csv")
-
-    X_test=df[["DAY",  "MONTH",  "HOUR",  "DAY_OF_WEEK",  "IS_HOLIDAY",  "w_avg"]]
-    y_test=df["LOAD"]
-
-    quantiles = [i/100 for i in range(1,100)]
-
-    # we need to scale X_test
-    df_train=pd.read_csv("/Users/luca/Desktop/GEFCom2014 Data/Load/L-train.csv")
-    df_train=df_train[df_train["MONTH"]==(ith-3)]
-    X_train=df_train[["DAY",  "MONTH",  "HOUR",  "DAY_OF_WEEK",  "IS_HOLIDAY",  "w_avg"]]
-    scaler = StandardScaler()
-    X_train_scaled = scaler.fit_transform(X_train)
-    X_test_scaled = scaler.transform(X_test)
-
-    pinball_tot=0
-    
-    for i,q in enumerate(quantiles):
-        krn_q=pickle.load(open(f"/Users/luca/Desktop/ThesisKernelMethods/dataset_applications/models_conditioned_months/task {ith} cv/krn_qr_{i}.pkl", 'rb'))
-        pinball_q=mean_pinball_loss(y_test,krn_q.predict(X_test_scaled), alpha=q)
-        print(f"pinball loss quantile {q}: ", pinball_q)
-        pinball_tot+=pinball_q
-
-        print(pinball_tot/(i+1))
-    print("total quantile: ", pinball_tot/len(quantiles))
-    ans=pinball_tot/len(quantiles)
-    return ans
-
-if __name__=="__main__":
-    i=int(sys.argv[1])
-    ans=test(i)
-
-
-    # original_stdout=sys.stdout
-    # with open(f"/Users/luca/Desktop/ThesisKernelMethods/thesis/tables/qr_gefcom2014.txt", "a") as f:
-    #     sys.stdout=f
-    #     # print(f"Average pinball loss, task n. {i}: ")
-    #     print(f"{ans:.4f}")
-    #     print("&")
-
-    #     sys.stdout=original_stdout

### 2024-05-21 experiments/gefcom_load/qr_gefcom2014.py

@@ -1,136 +1,0 @@
-import numpy as np
-import pandas as pd
-import pickle
-from pprint import pprint
-from sklearn.experimental import enable_halving_search_cv  # noqa
-from sklearn.model_selection import HalvingRandomSearchCV
-from sklearn.metrics import make_scorer
-from sklearn.metrics import mean_absolute_error
-from sklearn.metrics import mean_pinball_loss
-from sklearn.model_selection import train_test_split
-from sklearn.preprocessing import StandardScaler
-import sys
-from tqdm import tqdm
-
-from kqr import KQR
-
-from cvxopt import matrix, spmatrix, sparse
-
-
-
-
-if __name__=="__main__":
-    # load data 2009
-    df=pd.read_csv("/Users/luca/Desktop/GEFCom2014 Data/Load/L2009-train.csv")
-    
-    X_train=df[["DAY",  "MONTH",  "HOUR",  "DAY_OF_WEEK",  "IS_HOLIDAY",  "w_avg"]]
-    y_train=df["LOAD"]
-
-    # 99 quantiles
-    quantiles = [i/100 for i in range(1,100)]
-
-    # define loss to tune
-    neg_mean_pinball_loss_scorer_05 = make_scorer(
-    mean_pinball_loss,
-    alpha=0.5,
-    greater_is_better=False,  # maximize the negative of the loss
-    )
-
-    # kernel quantile regression
-    scaler = StandardScaler()
-
-    X_train_scaled = scaler.fit_transform(X_train)
-
-    qr_krn_models=[]
-    y_test_pred_qr_krn=[]
-
-    param_grid_krn = dict(
-    C=[1e-1,1e-2,1, 5, 10,1e2,1e4],
-    gamma=[1e-1,1e-2,0.5,1, 5, 10, 20]   
-    )
-    
-    krn_blueprint=KQR(alpha=0.5)
-    best_hyperparameters_krn=HalvingRandomSearchCV(
-            krn_blueprint,
-            param_grid_krn,
-            scoring=neg_mean_pinball_loss_scorer_05,
-            n_jobs=2,
-            random_state=0,
-        ).fit(X_train_scaled, y_train).best_params_
-    
-    for i,q in enumerate(tqdm(quantiles)):
-
-        # fit data for specific quantile
-        qr_krn_models+=[KQR(alpha=q, **best_hyperparameters_krn).fit(X_train_scaled, y_train)]
-
-        # save models to pickle
[…]

### 2024-05-21 experiments/gefcom_load/qr_gefcom2014_2.py

@@ -1,68 +1,0 @@
-import numpy as np
-import pandas as pd
-import pickle
-from pprint import pprint
-from sklearn.experimental import enable_halving_search_cv  # noqa
-from sklearn.model_selection import HalvingRandomSearchCV
-from sklearn.metrics import make_scorer
-from sklearn.metrics import mean_absolute_error
-from sklearn.metrics import mean_pinball_loss
-from sklearn.model_selection import train_test_split
-from sklearn.preprocessing import StandardScaler
-import sys
-from tqdm import tqdm
-
-from kqr import KQR
-
-from cvxopt import matrix, spmatrix, sparse
-
-
-
-
-if __name__=="__main__":
-    # load data 2009
-    df=pd.read_csv("/Users/luca/Desktop/GEFCom2014 Data/Load/L2009-train_2.csv")
-    
-    X_train=df[[f"w{i}" for i in range(1,26)]+["DAY",  "MONTH",  "HOUR",  "DAY_OF_WEEK",  "IS_HOLIDAY",  "w_avg"]]
-    y_train=df["LOAD"]
-
-    # 99 quantiles
-    quantiles = [i/100 for i in range(1,100)]
-
-    # define loss to tune
-    neg_mean_pinball_loss_scorer_05 = make_scorer(
-    mean_pinball_loss,
-    alpha=0.5,
-    greater_is_better=False,  # maximize the negative of the loss
-    )
-
-    # kernel quantile regression
-    scaler = StandardScaler()
-
-    X_train_scaled = scaler.fit_transform(X_train)
-
-    qr_krn_models=[]
-    y_test_pred_qr_krn=[]
-
-    param_grid_krn = dict(
-    C=[1e-1,1e-2,1, 5, 10,1e2,1e4],
-    gamma=[1e-1,1e-2,0.5,1, 5, 10, 20]   
-    )
-    
-    krn_blueprint=KQR(alpha=0.5)
-    best_hyperparameters_krn=HalvingRandomSearchCV(
-            krn_blueprint,
-            param_grid_krn,
-            scoring=neg_mean_pinball_loss_scorer_05,
-            n_jobs=2,
-            random_state=0,
-        ).fit(X_train_scaled, y_train).best_params_
-    
-    for i,q in enumerate(tqdm(quantiles)):
-
-        # fit data for specific quantile
[…]

### 2024-05-21 experiments/gefcom_load/qr_gefcom2014_2_big.py

@@ -1,49 +1,0 @@
-import numpy as np
-import pandas as pd
-import pickle
-from pprint import pprint
-from sklearn.experimental import enable_halving_search_cv  # noqa
-from sklearn.model_selection import HalvingRandomSearchCV
-from sklearn.metrics import make_scorer
-from sklearn.metrics import mean_absolute_error
-from sklearn.metrics import mean_pinball_loss
-from sklearn.model_selection import train_test_split
-from sklearn.preprocessing import StandardScaler
-import sys
-from tqdm import tqdm
-
-from kqr import KQR
-
-from cvxopt import matrix, spmatrix, sparse
-
-
-
-
-if __name__=="__main__":
-    # load data start 2008- start 2010
-    df=pd.read_csv("/Users/luca/Desktop/GEFCom2014 Data/Load/Task 1/L-train_clean_full_2.csv")
-    
-    X_train=df[[f"w{i}" for i in range(1,26)]+["DAY",  "MONTH",  "HOUR",  "DAY_OF_WEEK",  "IS_HOLIDAY",  "w_avg"]]
-    y_train=df["LOAD"]
-
-    # 99 quantiles
-    quantiles = [i/100 for i in range(0,10)]
-
-
-    # kernel quantile regression
-    scaler = StandardScaler()
-
-    X_train_scaled = scaler.fit_transform(X_train)
-
-    qr_krn_models=[]
-    y_test_pred_qr_krn=[]
-    # cross validated on smaller subset=> C=100, gamma=0.5=> I use these on the full dataset
-
-    for i,q in enumerate(tqdm(quantiles)):
-
-        # fit data for specific quantile
-        qr_krn_models+=[KQR(alpha=q, C=100, gamma=0.5).fit(X_train_scaled, y_train)]
-
-        # save models to pickle
-        pickle.dump(qr_krn_models[i], open(f'/Users/luca/Desktop/ThesisKernelMethods/dataset_applications/models_full_2/krn_qr_{i}.pkl', 'wb'))
-    

### 2024-05-21 experiments/gefcom_load/qr_gefcom2014_plot_task1.py

@@ -1,55 +1,0 @@
-# code to create plot of 90% interval
-
-# take data saved as predictions 5, 50, 95 quantile from each task,
-# vertical stack data
-# plot 
-
-from datetime import datetime
-import numpy as np
-import matplotlib.pyplot as plt
-import pandas as pd
-
-if __name__=="__main__":
-    
-    # load test task 1
-    y_test=pd.read_csv("/Users/luca/Desktop/GEFCom2014 Data/Load/Task 1/L1-test_clean.csv")
-    
-    
-    # load prediction for task 1
-    y_predict_q=pd.read_csv("/Users/luca/Desktop/GEFCom2014 Data/Load/Task 1/L1-our_model_prediction.csv")
-
-    # select quantiles
-    y_predict_5=y_predict_q["0.05"]
-    y_predict_50=y_predict_q["0.5"]
-    y_predict_95=y_predict_q["0.95"]
-
-    # dummy x for plotting
-    x=np.arange(len(y_test))
-    # ticks for the days, (including also hours adds clutter)
-    tick_array = [i for i in range(24,744+1,24)]
-    # label ticks
-    labels=[int(i/24) for i in tick_array]
-    
-    # plot quantile range
-    plt.plot(y_test["LOAD"],color="black")
-    
-    plt.fill_between(x,y_predict_q["0.05"],y_test["LOAD"], alpha=0.4, color="green", edgecolor="red")
-    plt.fill_between(x,y_test["LOAD"],y_predict_q["0.95"], alpha=0.4, color="green", edgecolor="red")
-    
-    plt.xticks(ticks=tick_array, labels=labels, rotation=45)
-    
-    plt.xlabel("Days")
-    plt.ylabel("Load in MW")
-    plt.show()
-
-
-    # plot true vs median
-    plt.plot(x,y_test["LOAD"],color="black", label="y_test")
-    plt.plot(x,y_predict_50, "red", alpha=0.4, label="0.5 prediction")
-    plt.xticks(ticks=tick_array, labels=labels, rotation=45)
-    
-    plt.xlabel("Days")
-    plt.ylabel("Load in MW")
-    plt.legend()
-    plt.show()
-

### 2024-05-21 experiments/gefcom_load/qr_gefcom2014_predictions.py

@@ -1,58 +1,0 @@
-# code to save predictions in the same format of GEFCom challenge
-# ZONEID| TIMESTAMP| 0.01|...|0.99
-
-# save data as csv, each task in its folder
-# from pd dataframe with rows timestamp
-# columns quantile
-
-
-
-import pandas as pd
-import pickle
-
-from sklearn.preprocessing import StandardScaler
-import sys
-
-from tqdm import tqdm
-
-
-def predict_save(ith):
-    # df=pd.read_csv(f"/Users/luca/Desktop/GEFCom2014 Data/Load/Task {ith}/L{ith}-test_clean.csv")
-    df=df=pd.read_csv(f"/Users/luca/Desktop/GEFCom2014 Data/Load/Task {ith}/L{ith}-avg_test_clean.csv")
-    df["w_avg"]=pd.read_csv(f"/Users/luca/Desktop/GEFCom2014 Data/Load/Task {ith}/temp.csv")
-
-    X_test=df[["DAY",  "MONTH",  "HOUR",  "DAY_OF_WEEK",  "IS_HOLIDAY",  "w_avg"]]
-    y_test=df["LOAD"]
-
-    quantiles = [i/100 for i in range(1,100)]
-
-    # we need to scale X_test
-    df_train=pd.read_csv("/Users/luca/Desktop/GEFCom2014 Data/Load/L2009-train.csv")
-    X_train=df_train[["DAY",  "MONTH",  "HOUR",  "DAY_OF_WEEK",  "IS_HOLIDAY",  "w_avg"]]
-    scaler = StandardScaler()
-    X_train_scaled = scaler.fit_transform(X_train)
-    X_test_scaled = scaler.transform(X_test)
-
-    
-    df_template_submission=pd.read_csv(f"/Users/luca/Desktop/GEFCom2014 Data/Load/Task {ith}/L{ith}-test.csv")
-    
-    if ith==15:
-        df_template_submission=pd.read_csv(f"/Users/luca/Desktop/GEFCom2014 Data/Load/Task {ith}/L15-benchmark.csv")
-        output_df=df_template_submission[["ZONEID", "TIMESTAMP"]].copy()
-        
-    else:
-        output_df=df_template_submission[["ZONEID", "TIMESTAMP"]].copy()
-    
-    for i,q in enumerate(quantiles):
[…]

### 2024-05-21 experiments/gefcom_load/qr_gefcom2014_test.py

@@ -1,52 +1,0 @@
-
-import pandas as pd
-import pickle
-
-from sklearn.metrics import mean_absolute_error
-from sklearn.metrics import mean_pinball_loss
-from sklearn.preprocessing import StandardScaler
-import sys
-
-from tqdm import tqdm
-
-
-def test(ith):
-    df=pd.read_csv(f"/Users/luca/Desktop/GEFCom2014 Data/Load/Task {ith}/L{ith}-test_clean.csv")
-
-    X_test=df[["DAY",  "MONTH",  "HOUR",  "DAY_OF_WEEK",  "IS_HOLIDAY",  "w_avg"]]
-    y_test=df["LOAD"]
-
-    quantiles = [i/100 for i in range(1,100)]
-
-    # we need to scale X_test
-    df_train=pd.read_csv("/Users/luca/Desktop/GEFCom2014 Data/Load/L2009-train.csv")
-    X_train=df_train[["DAY",  "MONTH",  "HOUR",  "DAY_OF_WEEK",  "IS_HOLIDAY",  "w_avg"]]
-    scaler = StandardScaler()
-    X_train_scaled = scaler.fit_transform(X_train)
-    X_test_scaled = scaler.transform(X_test)
-
-    pinball_tot=0
-
-    for i,q in enumerate(quantiles):
-        krn_q=pickle.load(open(f"/Users/luca/Desktop/ThesisKernelMethods/dataset_applications/models/krn_qr_{i}.pkl", 'rb'))
-        pinball_q=mean_pinball_loss(y_test,krn_q.predict(X_test_scaled), alpha=q)
-        print(f"pinball loss quantile {q}: ", pinball_q)
-        pinball_tot+=pinball_q
-    
-    print("total quantile: ", pinball_tot/len(quantiles))
-    ans=pinball_tot/len(quantiles)
-    return ans
-
-if __name__=="__main__":
-    i=int(sys.argv[1])
-    ans=test(i)
-
-
-    original_stdout=sys.stdout
-    with open(f"/Users/luca/Desktop/ThesisKernelMethods/thesis/tables/qr_gefcom2014.txt", "a") as f:
-        sys.stdout=f
-        # print(f"Average pinball loss, task n. {i}: ")
-        print(f"{ans:.4f}")
-        print("&")
-
-        sys.stdout=original_stdout

### 2024-05-21 experiments/gefcom_load/qr_gefcom2014_test_2.py

@@ -1,52 +1,0 @@
-
-import pandas as pd
-import pickle
-
-from sklearn.metrics import mean_absolute_error
-from sklearn.metrics import mean_pinball_loss
-from sklearn.preprocessing import StandardScaler
-import sys
-
-from tqdm import tqdm
-
-
-def test(ith):
-    df=pd.read_csv(f"/Users/luca/Desktop/GEFCom2014 Data/Load/Task {ith}/L{ith}-test_clean_2.csv")
-
-    X_test=df[[f"w{i}" for i in range(1,26)]+["DAY",  "MONTH",  "HOUR",  "DAY_OF_WEEK",  "IS_HOLIDAY",  "w_avg"]]
-    y_test=df["LOAD"]
-
-    quantiles = [i/100 for i in range(1,100)]
-
-    # we need to scale X_test
-    df_train=pd.read_csv("/Users/luca/Desktop/GEFCom2014 Data/Load/L2009-train_2.csv")
-    X_train=df_train[[f"w{i}" for i in range(1,26)]+["DAY",  "MONTH",  "HOUR",  "DAY_OF_WEEK",  "IS_HOLIDAY",  "w_avg"]]
-    scaler = StandardScaler()
-    X_train_scaled = scaler.fit_transform(X_train)
-    X_test_scaled = scaler.transform(X_test)
-
-    pinball_tot=0
-
-    for i,q in enumerate(quantiles):
-        krn_q=pickle.load(open(f"/Users/luca/Desktop/ThesisKernelMethods/dataset_applications/models_2/krn_qr_{i}.pkl", 'rb'))
-        pinball_q=mean_pinball_loss(y_test,krn_q.predict(X_test_scaled), alpha=q)
-        print(f"pinball loss quantile {q}: ", pinball_q)
-        pinball_tot+=pinball_q
-    
-    print("total quantile: ", pinball_tot/len(quantiles))
-    ans=pinball_tot/len(quantiles)
-    return ans
-
-if __name__=="__main__":
-    i=int(sys.argv[1])
-    ans=test(i)
-
-
-    original_stdout=sys.stdout
-    with open(f"/Users/luca/Desktop/ThesisKernelMethods/thesis/tables/qr_gefcom2014_2.txt", "a") as f:
-        sys.stdout=f
-        # print(f"Average pinball loss, task n. {i}: ")
-        print("Task ", f"{i}: ",f"{ans:.4f}")
-        print("&")
-
-        sys.stdout=original_stdout

### 2024-05-21 experiments/gefcom_load/qr_gefcom2014_test_2_big.py

@@ -1,52 +1,0 @@
-
-import pandas as pd
-import pickle
-
-from sklearn.metrics import mean_absolute_error
-from sklearn.metrics import mean_pinball_loss
-from sklearn.preprocessing import StandardScaler
-import sys
-
-from tqdm import tqdm
-
-
-def test(ith):
-    df=pd.read_csv(f"/Users/luca/Desktop/GEFCom2014 Data/Load/Task {ith}/L{ith}-test_clean_2.csv")
-
-    X_test=df[[f"w{i}" for i in range(1,26)]+["DAY",  "MONTH",  "HOUR",  "DAY_OF_WEEK",  "IS_HOLIDAY",  "w_avg"]]
-    y_test=df["LOAD"]
-
-    quantiles = [i/100 for i in range(1,100)]
-
-    # we need to scale X_test
-    df_train=pd.read_csv("/Users/luca/Desktop/GEFCom2014 Data/Load/Task 1/L-train_clean_full_2.csv")
-    X_train=df_train[[f"w{i}" for i in range(1,26)]+["DAY",  "MONTH",  "HOUR",  "DAY_OF_WEEK",  "IS_HOLIDAY",  "w_avg"]]
-    scaler = StandardScaler()
-    X_train_scaled = scaler.fit_transform(X_train)
-    X_test_scaled = scaler.transform(X_test)
-
-    pinball_tot=0
-
-    for i,q in enumerate(quantiles):
-        krn_q=pickle.load(open(f"/Users/luca/Desktop/ThesisKernelMethods/dataset_applications/models_2/krn_qr_{i}.pkl", 'rb'))
-        pinball_q=mean_pinball_loss(y_test,krn_q.predict(X_test_scaled), alpha=q)
-        print(f"pinball loss quantile {q}: ", pinball_q)
-        pinball_tot+=pinball_q
-    
-    print("total quantile: ", pinball_tot/len(quantiles))
-    ans=pinball_tot/len(quantiles)
-    return ans
-
-if __name__=="__main__":
-    i=int(sys.argv[1])
-    ans=test(i)
-
-
-    original_stdout=sys.stdout
-    with open(f"/Users/luca/Desktop/ThesisKernelMethods/thesis/tables/qr_gefcom2014_2_big.txt", "a") as f:
-        sys.stdout=f
-        # print(f"Average pinball loss, task n. {i}: ")
-        print("Task ", f"{i}: ",f"{ans:.4f}")
-        print("&")
-
-        sys.stdout=original_stdout

### 2024-05-21 experiments/gefcom_load/qr_gefcom2014_test_avg.py

@@ -1,51 +1,0 @@
-
-import pandas as pd
-import pickle
-
-from sklearn.metrics import mean_absolute_error
-from sklearn.metrics import mean_pinball_loss
-from sklearn.preprocessing import StandardScaler
-import sys
-
-from tqdm import tqdm
-
-
-def test(ith):
-    df=pd.read_csv(f"/Users/luca/Desktop/GEFCom2014 Data/Load/Task {ith}/L{ith}-avg_test_clean.csv")
-
-    X_test=df[["DAY",  "MONTH",  "HOUR",  "DAY_OF_WEEK",  "IS_HOLIDAY",  "w_avg"]]
-    y_test=df["LOAD"]
-
-    quantiles = [i/100 for i in range(1,100)]
-
-    # we need to scale X_test
-    df_train=pd.read_csv("/Users/luca/Desktop/GEFCom2014 Data/Load/L2009-train.csv")
-    X_train=df_train[["DAY",  "MONTH",  "HOUR",  "DAY_OF_WEEK",  "IS_HOLIDAY",  "w_avg"]]
-    scaler = StandardScaler()
-    X_train_scaled = scaler.fit_transform(X_train)
-    X_test_scaled = scaler.transform(X_test)
-
-    pinball_tot=0
-
-    for i,q in enumerate(quantiles):
-        krn_q=pickle.load(open(f"/Users/luca/Desktop/ThesisKernelMethods/dataset_applications/models/krn_qr_{i}.pkl", 'rb'))
-        pinball_q=mean_pinball_loss(y_test,krn_q.predict(X_test_scaled), alpha=q)
-        print(f"pinball loss quantile {q}: ", pinball_q)
-        pinball_tot+=pinball_q
-    
-    print("total quantile: ", pinball_tot/len(quantiles))
-    ans=pinball_tot/len(quantiles)
-    return ans
-
-if __name__=="__main__":
-    i=int(sys.argv[1])
-    ans=test(i)
-
-    original_stdout=sys.stdout
-    with open(f"/Users/luca/Desktop/ThesisKernelMethods/thesis/tables/qr_gefcom2014_w_avg.txt", "a") as f:
-        sys.stdout=f
-        # print(f"Average pinball loss, task n. {i}: ")
-        print(f"{ans:.4f}")
-        print("&")
-
-        sys.stdout=original_stdout

### 2024-05-21 experiments/gefcom_load/qr_gefcom2014_test_avg2.py

@@ -1,50 +1,0 @@
-import pandas as pd
-import pickle
-
-from sklearn.metrics import mean_absolute_error
-from sklearn.metrics import mean_pinball_loss
-from sklearn.preprocessing import StandardScaler
-import sys
-
-from tqdm import tqdm
-
-
-def test(ith):
-    df=pd.read_csv(f"/Users/luca/Desktop/GEFCom2014 Data/Load/Task {ith}/L{ith}-avg_test_clean2.csv")
-
-    X_test=df[["DAY",  "MONTH",  "HOUR",  "DAY_OF_WEEK",  "IS_HOLIDAY",  "w_avg"]]
-    y_test=df["LOAD"]
-
-    quantiles = [i/100 for i in range(1,100)]
-
-    # we need to scale X_test
-    df_train=pd.read_csv("/Users/luca/Desktop/GEFCom2014 Data/Load/L2009-train.csv")
-    X_train=df_train[["DAY",  "MONTH",  "HOUR",  "DAY_OF_WEEK",  "IS_HOLIDAY",  "w_avg"]]
-    scaler = StandardScaler()
-    X_train_scaled = scaler.fit_transform(X_train)
-    X_test_scaled = scaler.transform(X_test)
-
-    pinball_tot=0
-
-    for i,q in enumerate(quantiles):
-        krn_q=pickle.load(open(f"/Users/luca/Desktop/ThesisKernelMethods/dataset_applications/models/krn_qr_{i}.pkl", 'rb'))
-        pinball_q=mean_pinball_loss(y_test,krn_q.predict(X_test_scaled), alpha=q)
-        print(f"pinball loss quantile {q}: ", pinball_q)
-        pinball_tot+=pinball_q
-    
-    print("total quantile: ", pinball_tot/len(quantiles))
-    ans=pinball_tot/len(quantiles)
-    return ans
-
-if __name__=="__main__":
-    i=int(sys.argv[1])
-    ans=test(i)
-
-    original_stdout=sys.stdout
-    with open(f"/Users/luca/Desktop/ThesisKernelMethods/thesis/tables/qr_gefcom2014_w_avg2.txt", "a") as f:
-        sys.stdout=f
-        # print(f"Average pinball loss, task n. {i}: ")
-        print(f"{ans:.4f}")
-        print("&")
-
-        sys.stdout=original_stdout

### 2024-05-21 experiments/gefcom_load/weather_prediction_avg.py

@@ -1,39 +1,0 @@
-# this script creates the testdataset with weather predicted as the average of the past 9 years
-import pandas as pd
-
-df_w=pd.read_csv(f"/Users/luca/Desktop/GEFCom2014 Data/Load/L-weather-train.csv")
-
-# drop leap year
-df_w=df_w[~((df_w["MONTH"]==2) & (df_w["DAY"]==29))]
-
-w_avg=df_w.groupby(["MONTH","DAY","HOUR"]).mean("w_avg").reset_index()
-
-
-# print(len(w_avg))
-task_month={1:10,2:11,3:12,4:1,5:2,6:3,7:4,8:5,9:6,10:7,11:8,12:9,13:10,14:11,15:12}
-
-for t,m in task_month.items():
-    
-    w_avg_month=w_avg[w_avg["MONTH"]==m]
-    
-    if m!=12:
-        idx_next=w_avg_month.index[-1]
-    else:
-        idx_next=0
-    # print(end)
-    # print(w_avg.iloc[end-1])
-    
-    # add last row because test data ends at 00:00 of the next month
-    w_avg_month.loc[len(w_avg_month)] = w_avg.iloc[idx_next+1]
-    
-    # drop first row because test data starts from hour 01:00, thus drop data for hour 00:00
-    w_avg_month.drop(index=w_avg_month.index[0], axis=0, inplace=True)
-    # print(w_avg_month)
-
-    df_test=pd.read_csv(f"/Users/luca/Desktop/GEFCom2014 Data/Load/Task {t}/L{t}-test_clean.csv")
-    
-    df_test["w_avg"]=w_avg_month["w_avg"].values
-
-    df_test.to_csv(f"/Users/luca/Desktop/GEFCom2014 Data/Load/Task {t}/L{t}-avg_test_clean.csv", index=False)
-
-

### 2024-05-21 experiments/gefcom_load/weather_prediction_avg2.py

@@ -1,84 +1,0 @@
-# same as weather_prediction_avg.py but selecting only weather stations "w6", "w10", "w22", "w25"
-import numpy as np
-import pandas as pd
-from sklearn.metrics import mean_pinball_loss
-
-from qr import qr_tests
-
-import holidays
-
-
-df=pd.read_csv("/Users/luca/Desktop/GEFCom2014 Data/Load/Task 1/L1-train.csv", sep=",", decimal=".")
-    
-# clean data dates
-dates = pd.date_range(start='2001-01-01 01:00:00', periods=len(df), freq='H')
-df['DAY'] = dates.day
-df['MONTH'] = dates.month
-df['HOUR'] = dates.hour
-df['YEAR'] = dates.year
-
-# add info weekday; saturday, sunday
-df['DAY_OF_WEEK'] = dates.dayofweek
-
-# holidays
-hol=holidays.US(years=range(2001, 2011))
-df["IS_HOLIDAY"]=pd.Series(dates.date).isin(hol)
-# convert it to categorical variable
-df["IS_HOLIDAY"]=pd.Categorical(df["IS_HOLIDAY"].astype(int))
-
-
-# drop not useful columns
-# zoneid is dropped because are all 1, so it is not a meaningful independent variable
-# the information in timestamp has been decomposed in four columns so we can drop it
-df.drop(columns=['ZONEID', 'TIMESTAMP'], inplace=True)
-
-
-# average temperatures
-w_avg_sub = df[["w6", "w10", "w22", "w25"]].mean(axis=1)
-
-weather_stat=df.filter(regex=("w.*")).columns.to_list()
-df.drop(columns=weather_stat, inplace=True)
-df['w_avg']=w_avg_sub
-
-df.drop(columns=["YEAR"], inplace=True)
-
-df.to_csv("/Users/luca/Desktop/GEFCom2014 Data/Load/L-weather-train2.csv",index=False)
-
-
-
-df_w=pd.read_csv(f"/Users/luca/Desktop/GEFCom2014 Data/Load/L-weather-train2.csv")
-
-# drop leap year
-df_w=df_w[~((df_w["MONTH"]==2) & (df_w["DAY"]==29))]
-
-w_avg=df_w.groupby(["MONTH","DAY","HOUR"]).mean("w_avg").reset_index()
-
-
-# print(len(w_avg))
[…]

### 2024-05-21 experiments/gefcom_price/qr_gefcom2014.py

@@ -1,45 +1,0 @@
-import numpy as np
-import pandas as pd
-import pickle
-from pprint import pprint
-from sklearn.experimental import enable_halving_search_cv  # noqa
-from sklearn.model_selection import HalvingRandomSearchCV
-from sklearn.metrics import make_scorer
-from sklearn.metrics import mean_absolute_error
-from sklearn.metrics import mean_pinball_loss
-from sklearn.model_selection import train_test_split
-from sklearn.preprocessing import StandardScaler
-import sys
-from tqdm import tqdm
-
-from kqr import KQR
-
-from cvxopt import matrix, spmatrix, sparse
-
-if __name__=="__main__":
-    # load data
-    df_train=pd.read_csv("/Users/luca/Desktop/ThesisKernelMethods/dataset_applications/gefcom_price/Price/Task 4/Task4_P_train.csv")
-    
-    df_train=df_train[(df_train["MONTH"]==7) | (df_train["MONTH"]==6)]
-    # define train
-    X_train=df_train[["MONTH","DAY","HOUR","Forecasted Total Load","Forecasted Zonal Load"]]
-    y_train=df_train["Zonal Price"]
-    
-    scaler = StandardScaler()
-    X_train_scaled = scaler.fit_transform(X_train)
-    
-    # 99 quantiles
-    quantiles = [i/100 for i in range(1,100)]
-
-    # kernel quantile regression
-    qr_krn_models=[]
-    y_test_pred_qr_krn=[]
-    
-    for i,q in enumerate(tqdm(quantiles)):
-
-        # fit data for specific quantile
-        qr_krn_models+=[KQR(alpha=q, C=100, gamma=0.5).fit(X_train_scaled, y_train)]
-
-        # save models to pickle
-        pickle.dump(qr_krn_models[i], open(f'/Users/luca/Desktop/ThesisKernelMethods/dataset_applications/gefcom_price/models/krn_qr_{i}.pkl', 'wb'))
-        

### 2024-05-21 experiments/gefcom_price/qr_gefcom2014_cv.py

@@ -1,72 +1,0 @@
-import numpy as np
-import pandas as pd
-import pickle
-from pprint import pprint
-from sklearn.experimental import enable_halving_search_cv  # noqa
-from sklearn.model_selection import HalvingRandomSearchCV
-from sklearn.metrics import make_scorer
-from sklearn.metrics import mean_absolute_error
-from sklearn.metrics import mean_pinball_loss
-from sklearn.model_selection import train_test_split
-from sklearn.preprocessing import StandardScaler
-import sys
-from tqdm import tqdm
-
-from kqr import KQR
-
-from cvxopt import matrix, spmatrix, sparse
-
-
-
-
-if __name__=="__main__":
-    ith=15
-    task_month={4:7,5:7,6:7,7:7,8:7,9:7,10:7,11:7,12:7,13:12,14:12,15:12}
-    # load data
-    df_train=pd.read_csv(f"/Users/luca/Desktop/ThesisKernelMethods/dataset_applications/gefcom_price/Price/Task {ith}/Task{ith}_P_train.csv")
-    df_train=df_train[(df_train["MONTH"]==task_month[ith])][-1000:]
-    # define train
-    X_train=df_train[["MONTH","DAY","HOUR","Forecasted Total Load","Forecasted Zonal Load"]]
-    y_train=df_train["Zonal Price"]
-
-    # 99 quantiles
-    quantiles = [i/100 for i in range(1,100)]
-
-    # kernel quantile regression
-    scaler = StandardScaler()
-
-    X_train_scaled = scaler.fit_transform(X_train)
-
-    qr_krn_models=[]
-    y_test_pred_qr_krn=[]
-
-    param_grid_krn = dict(
-    C=[1e-1,1e-2,1, 5, 10,1e2,1e4],
-    gamma=[1e-1,1e-2,0.5,1, 5, 10, 20]   
-    )
-    
-    
-    for i,q in enumerate(tqdm(quantiles)):
-        
-        # define loss to tune
-        neg_mean_pinball_loss_scorer = make_scorer(
-        mean_pinball_loss,
-        alpha=q,
-        greater_is_better=False,  # maximize the negative of the loss
-        )
-
-        krn_blueprint=KQR(alpha=q)
-        best_hyperparameters_krn=HalvingRandomSearchCV(
-                krn_blueprint,
-                param_grid_krn,
-                scoring=neg_mean_pinball_loss_scorer,
-                n_jobs=2,
[…]

### 2024-05-21 experiments/gefcom_price/qr_gefcom2014_test.py

@@ -1,69 +1,0 @@
-
-import pandas as pd
-import pickle
-
-from sklearn.metrics import mean_absolute_error
-from sklearn.metrics import mean_pinball_loss
-from sklearn.preprocessing import StandardScaler
-import sys
-
-from tqdm import tqdm
-
-
-def test(ith):
-
-    task_month={4:7,5:7,6:7,7:7,8:7,9:7,10:7,11:7,12:7,13:12,14:12,15:12}
-
-    df=pd.read_csv(f"/Users/luca/Desktop/ThesisKernelMethods/dataset_applications/gefcom_price/Price/Task {ith}/Task{ith}_P_test.csv")
-
-    X_test=df[["MONTH","DAY","HOUR","Forecasted Total Load","Forecasted Zonal Load"]]
-    y_test=df["Zonal Price"]
-
-    quantiles = [i/100 for i in range(1,100)]
-
-    # we need to scale X_test
-    df_train=pd.read_csv(f"/Users/luca/Desktop/ThesisKernelMethods/dataset_applications/gefcom_price/Price/Task {ith}/Task{ith}_P_train.csv")
-    df_train=df_train[(df_train["MONTH"]==task_month[ith])]
-
-    X_train=df_train[["MONTH","DAY","HOUR","Forecasted Total Load","Forecasted Zonal Load"]]
-    scaler = StandardScaler()
-    X_train_scaled = scaler.fit_transform(X_train)
-    X_test_scaled = scaler.transform(X_test)
-
-    pinball_tot=0
-
-    # predict
-    df_template_submission=pd.read_csv(f"/Users/luca/Desktop/ThesisKernelMethods/dataset_applications/gefcom_price/Price/Task {ith}/Benchmark{ith}_P.csv")
-
-    df_predict=df_template_submission[["ZONEID", "timestamp"]].copy()
-
-    for i,q in enumerate(quantiles):
-        krn_q=pickle.load(open(f"/Users/luca/Desktop/ThesisKernelMethods/dataset_applications/gefcom_price/models cv/task {ith}/krn_qr_{i}.pkl", "rb"))
-        y_predict_q=krn_q.predict(X_test_scaled)
-        pinball_q=mean_pinball_loss(y_test,y_predict_q, alpha=q)
-        print(f"pinball loss quantile {q}: ", pinball_q)
-        pinball_tot+=pinball_q
-
-        df_predict[f"{q}"]=pd.Series(y_predict_q)
-
[…]

### 2024-05-21 experiments/gefcom_price/utils/create_test.py

@@ -1,48 +1,0 @@
-
-import re
-import pandas as pd
-import sys
-
-import holidays
-
-import miscellaneous
-
-# bash script
-# for ((i=1;i<=15;i++)); do python create_test.py "/Users/luca/Desktop/ThesisKernelMethods/dataset_applications/gefcom_price/Price/Task $i/Task${i}_P.csv"; done
-
-
-
-def create_test(file):
-    df=miscellaneous.clean_time(file)
-
-    df_test=df[-24:]
-    # print(df_test)
-
-    # get new file containing prices we have to test for task i
-    file_new=miscellaneous.get_test(file)
-    df_new=pd.read_csv(f"{file_new}", sep=",", decimal=".")
-        
-    # join on df_test
-    df_test.fillna(0)
-    df_merged=df_test.merge(df_new, on="timestamp", how="left",suffixes=('_x', ''))
-    
-    df_merged.drop(df_merged.filter(regex='_x$').columns, axis=1, inplace=True)
-    
-    
-    # order columns
-    df_merged=miscellaneous.order_columns(df_merged)
-    # save
-    n=miscellaneous.get_task_number(file)
-    
-    
-    if miscellaneous.get_task_number(file)==15:
-        df_merged["ZONEID"]=df_new["ZONEID"]
-        df_merged["Zonal Price"]=df_new["Zonal Price"]
-        
-    df_merged.to_csv(f"/Users/luca/Desktop/ThesisKernelMethods/dataset_applications/gefcom_price/Price/Task {n}/Task{n}_P_test.csv", index=False)
-
-
-if __name__=="__main__":
-    file=sys.argv[1]
-    # clean passed file
-    create_test(f"{file}")

### 2024-05-21 experiments/gefcom_price/utils/create_train.py

@@ -1,26 +1,0 @@
-
-
-import pandas as pd
-import sys
-
-import holidays
-
-import miscellaneous
-# bash script
-# for ((i=1;i<=15;i++)); do python create_train.py "/Users/luca/Desktop/ThesisKernelMethods/dataset_applications/gefcom_price/Price/Task $i/Task${i}_P.csv"; done
-
-def create_train(file):
-    df=miscellaneous.clean_time(file)
-
-    df_train=df[0:-24]
-
-    # order columns
-    df_train=miscellaneous.order_columns(df_train)
-    # save
-    n=miscellaneous.get_task_number(file)
-    df_train.to_csv(f"/Users/luca/Desktop/ThesisKernelMethods/dataset_applications/gefcom_price/Price/Task {n}/Task{n}_P_train.csv", index=False)
-
-if __name__=="__main__":
-    file=sys.argv[1]
-    # clean passed file
-    create_train(file)

### 2024-05-21 experiments/gefcom_price/utils/miscellaneous.py

@@ -1,53 +1,0 @@
-import pandas as pd
-import re
-
-import holidays
-
-def clean_time(file):
-    # read file as pandas df
-    df=pd.read_csv(file, sep=",", decimal=".")
-    
-    ## clean data dates
-    dates = pd.date_range(start='2011-01-01 00:00:00', periods=len(df), freq='H')
-    df['DAY'] = dates.day
-    df['MONTH'] = dates.month
-    df['HOUR'] = dates.hour
-    df['YEAR'] = dates.year
-    
-    # add info weekday; saturday, sunday
-    df['DAY_OF_WEEK'] = dates.dayofweek
-
-    # holidays
-    hol=holidays.US(years=range(2011, 2014))
-    df["IS_HOLIDAY"]=pd.Series(dates.date).isin(hol).astype(int)  
-    # print(df)
-    return df
-
-def order_columns(df):
-    df=df[["ZONEID","timestamp","MONTH","DAY","HOUR","Forecasted Total Load","Forecasted Zonal Load","Zonal Price"]]
-    return df
-
-def get_task_number(file):
-    pattern = r'(\d+)'
-
-    matches = re.findall(pattern, file)
-
-    if matches:
-        n = int(matches[0])
-        
-    return n
-
-
-
-def get_test(file):
-    pattern = r'(\d+)'
-
-    matches = re.findall(pattern, file)
-
-    if matches:
-        n = int(matches[0])
-        n_new = n + 1
-
-        file_new = re.sub(pattern, str(n_new), file)
-    
-    return file_new

### 2024-05-21 experiments/gefcom_price/utils/qr_gefcom2014.py

@@ -1,44 +1,0 @@
-import numpy as np
-import pandas as pd
-import pickle
-from pprint import pprint
-from sklearn.experimental import enable_halving_search_cv  # noqa
-from sklearn.model_selection import HalvingRandomSearchCV
-from sklearn.metrics import make_scorer
-from sklearn.metrics import mean_absolute_error
-from sklearn.metrics import mean_pinball_loss
-from sklearn.model_selection import train_test_split
-from sklearn.preprocessing import StandardScaler
-import sys
-from tqdm import tqdm
-
-from kqr import KQR
-
-from cvxopt import matrix, spmatrix, sparse
-
-if __name__=="__main__":
-    # load data
-    df=pd.read_csv("/Users/luca/Desktop/ThesisKernelMethods/dataset_applications/gefcom_price/Price/Task 4/Task4_P_train.csv")
-    
-    # define train
-    X_train=df[["MONTH","DAY","HOUR","Forecasted Total Load","Forecasted Zonal Load"]][-1000:]
-    y_train=df["Zonal Price"][-1000:]
-    
-    scaler = StandardScaler()
-    X_train_scaled = scaler.fit_transform(X_train)
-
-    # 99 quantiles
-    quantiles = [i/100 for i in range(1,100)]
-
-    # kernel quantile regression
-    qr_krn_models=[]
-    y_test_pred_qr_krn=[]
-    
-    for i,q in enumerate(tqdm(quantiles)):
-
-        # fit data for specific quantile
-        qr_krn_models+=[KQR(alpha=q, C=100, gamma=0.5).fit(X_train_scaled, y_train)]
-
-        # save models to pickle
-        pickle.dump(qr_krn_models[i], open(f'/Users/luca/Desktop/ThesisKernelMethods/dataset_applications/gefcom_price/models/krn_qr_{i}.pkl', 'wb'))
-        

### 2024-05-21 thesis/appendix.tex
May 21st 2024
thesis update

@@
 \section{Feature map normalization}\label{appendix:new_feature}
+
+\begin{lemma}
+    Associated to any RKHS $\mathcal{H}$, there exists a feature vector $\phi(x)$ such that $\|\phi(x)\|_{\mathcal{H}}=R \quad \forall x \in \mathcal{X}$.
+    That is the Hilbert space norm of the feature vector is equal to a constant R for all states in the set $\mathcal{X}$.
+\end{lemma}
 \begin{proof}
+    Taking the new feature vector as $\phi^{new} (x):=\frac{\phi(x)}{\|\phi(x)\|_{\mathcal{H}}}$ we have
 \begin{align*}
@@
     &=
-    \langle
+    \left\langle
     \frac{\phi(x)}
@@
     {\sqrt{k(x,x)}}
-    \rangle_{\mathcal{H}}
+    \right\rangle_{\mathcal{H}}
     \\
@@
 \midrule
-0 & 13.785678 & 11.418540 & 10.587686 & 10.297572 \\
+0 & 13.785678 & 11.418540 & 10.587686 & \textbf{10.297572} \\
 \bottomrule
@@
 \midrule
-0.100000 & 0.729749 & 0.771714 & 0.588441 & 0.578898 \\
-0.200000 & 1.122582 & 1.033442 & 0.932824 & 0.869145 \\
-0.300000 & 1.479486 & 1.170642 & 1.153765 & 1.142783 \\
-0.400000 & 1.712577 & 1.436263 & 1.352667 & 1.331955 \\
-0.500000 & 1.911385 & 1.344361 & 1.408333 & 1.396300 \\
-0.600000 & 1.989514 & 1.448885 & 1.464902 & 1.431705 \\
-0.700000 & 1.938362 & 1.508741 & 1.427912 & 1.382772 \\
-0.800000 & 1.658058 & 1.497901 & 1.275059 & 1.245288 \\
-0.900000 & 1.243965 & 1.206591 & 0.983784 & 0.918725 \\
+0.100000 & 0.729749 & 0.771714 & 0.588441 & \textbf{0.578898} \\
+0.200000 & 1.122582 & 1.033442 & 0.932824 & \textbf{0.869145} \\
+0.300000 & 1.479486 & 1.170642 & 1.153765 & \textbf{1.142783} \\
+0.400000 & 1.712577 & 1.436263 & 1.352667 & \textbf{1.331955} \\
+0.500000 & 1.911385 & \textbf{1.344361} & 1.408333 & 1.396300 \\
+0.600000 & 1.989514 & 1.448885 & 1.464902 & \textbf{1.431705} \\
+0.700000 & 1.938362 & 1.508741 & 1.427912 & \textbf{1.382772} \\
+0.800000 & 1.658058 & 1.497901 & 1.275059 & \textbf{1.245288} \\
+0.900000 & 1.243965 & 1.206591 & 0.983784 & \textbf{0.918725} \\
 \bottomrule
@@
 \midrule
-0 & 3.826326 & 2.845989 & 2.965490 & 2.810494 \\
+0 & 3.826326 & 2.845989 & 2.965490 & \textbf{2.810494} \\
 \bottomrule
@@
     \midrule
-    0 & 5.613975 & 5.531938 & 5.212990 & 5.252491 \\
+    0 & 5.613975 & 5.531938 & \textbf{5.212990} & 5.252491 \\
     \bottomrule
@@
     \midrule
-    0.100000 & 0.277903 & 0.290531 & 0.274079 & 0.269287 \\
-    0.200000 & 0.469361 & 0.488079 & 0.453110 & 0.457286 \\
-    0.300000 & 0.621961 & 0.625633 & 0.580766 & 0.596791 \\
-    0.400000 & 0.729757 & 0.715875 & 0.689904 & 0.691310 \\
-    0.500000 & 0.794695 & 0.766185 & 0.735945 & 0.740834 \\
-    0.600000 & 0.810691 & 0.785769 & 0.744928 & 0.746636 \\
-    0.700000 & 0.769587 & 0.730318 & 0.700287 & 0.710392 \\
-    0.800000 & 0.667776 & 0.656913 & 0.609378 & 0.608334 \\
-    0.900000 & 0.472244 & 0.472635 & 0.424593 & 0.431621 \\
+    0.100000 & 0.277903 & 0.290531 & 0.274079 & \textbf{0.269287} \\
+    0.200000 & 0.469361 & 0.488079 & \textbf{0.453110} & 0.457286 \\
+    0.300000 & 0.621961 & 0.625633 & \textbf{0.580766} & 0.596791 \\
+    0.400000 & 0.729757 & 0.715875 & \textbf{0.689904} & 0.691310 \\
+    0.500000 & 0.794695 & 0.766185 & \textbf{0.735945} & 0.740834 \\
+    0.600000 & 0.810691 & 0.785769 & \textbf{0.744928} & 0.746636 \\
+    0.700000 & 0.769587 & 0.730318 & \textbf{0.700287} & 0.710392 \\
+    0.800000 & 0.667776 & 0.656913 & 0.609378 & \textbf{0.608334} \\
+    0.900000 & 0.472244 & 0.472635 & \textbf{0.424593} & 0.431621 \\
     \bottomrule
@@
     \midrule
-    0 & 1.627627 & 1.574179 & 1.499522 & 1.498583 \\
+    0 & 1.627627 & 1.574179 & 1.499522 & \textbf{1.498583} \\
     \bottomrule
@@
     \midrule
-    0 & 4.054449 & 2.289554 & 2.844410 & 2.204343 \\
+    0 & 4.054449 & 2.289554 & 2.844410 & \textbf{2.204343} \\
     \bottomrule
@@
     \midrule
-    0.100000 & 0.254649 & 0.139849 & 0.170489 & 0.182490 \\
-    0.200000 & 0.403772 & 0.236339 & 0.285875 & 0.223165 \\
-    0.300000 & 0.548820 & 0.244086 & 0.357375 & 0.242963 \\
-    0.400000 & 0.576918 & 0.263169 & 0.389305 & 0.262835 \\
-    0.500000 & 0.554367 & 0.306123 & 0.410738 & 0.295878 \\
-    0.600000 & 0.563046 & 0.335363 & 0.398125 & 0.306062 \\
-    0.700000 & 0.516019 & 0.287490 & 0.326572 & 0.283716 \\
-    0.800000 & 0.407742 & 0.261882 & 0.313698 & 0.235793 \\
-    0.900000 & 0.229115 & 0.215253 & 0.192233 & 0.171440 \\
+    0.100000 & 0.254649 & \textbf{0.139849} & 0.170489 & 0.182490 \\
+    0.200000 & 0.403772 & 0.236339 & 0.285875 & \textbf{0.223165} \\
+    0.300000 & 0.548820 & 0.244086 & 0.357375 & \textbf{0.242963} \\
+    0.400000 & 0.576918 & 0.263169 & 0.389305 & \textbf{0.262835} \\
+    0.500000 & 0.554367 & 0.306123 & 0.410738 & \textbf{0.295878} \\
+    0.600000 & 0.563046 & 0.335363 & 0.398125 & \textbf{0.306062} \\
+    0.700000 & 0.516019 & 0.287490 & 0.326572 & \textbf{0.283716} \\
+    0.800000 & 0.407742 & 0.261882 & 0.313698 & \textbf{0.235793} \\
+    0.900000 & 0.229115 & 0.215253 & 0.192233 & \textbf{0.171440} \\
     \bottomrule
@@
     \midrule
-    0 & 1.117714 & 0.606971 & 0.752197 & 0.594292 \\
+    0 & 1.117714 & 0.606971 & 0.752197 & \textbf{0.594292} \\
     \bottomrule
@@
 The rationale is that K-fold will randomize the order of the data, thus it might happen to use future data to predict the past; we want to avoid such behaviour in a time series setting.
-When carryout out any kind of cross validation, we must keep consistency in the way we evaluate our predictors during model selection and in the way we perform evaluation of the test data.
+When carrying out out any kind of cross validation, we must keep consistency in the way we evaluate our predictors during model selection and in the way we perform evaluation of the test data.
 Hence, for time series crossvalidation we have the timeseries split procedure, see figure \ref{fig:crossvalidationtimeseries} for a visualisation; the blue observations make up the training sets while the orange observations form the test sets.
@@
 The same concept applies for multi step ahead forecasting, see picture \ref{fig:crossvalidationtimeseries2}.
-That is in predicting $\hat{L}_{N+m}$ we use as inputs $L_1, L_2, \dots, \hat{L}_{N},\hat{L}_{N+1},\dots, \hat{L}_{N+m-1}$; where $\hat{L}_{N},\hat{L}_{N+1},\dots, \hat{L}_{N+m-1}$ are one step ahead forecasts.
+That is in predicting $\hat{y}_{N+m}$ we use as inputs $y_1, y_2, \dots, \hat{y}_{N},\hat{y}_{N+1},\dots, \hat{y}_{N+m-1}$; where $\hat{y}_{N},\hat{y}_{N+1},\dots, \hat{y}_{N+m-1}$ are one step ahead forecasts.
 \begin{figure}
@@
 With data normalization we transform the range of the feautures to a standard scale.
-Such preprocessing step is essential when employing distance based algorithms like svm or k-nearest neighbors.
+Such preprocessing step is essential when employing distance based algorithms like SVM or K-nearest neighbors.
 The rationale behind it is that by normalizing data we give an uniform weight to each feature in the learning process; in this way we do not favour larger scale features.
@@
 We conclude this subsection with a custom example that motivates the need of feature scaling \cite{scikitscale_example}.
-The idea is to compare the results of modelling the data with k nearest neighbors on the unscaled data against the scaled data.
+The idea is to compare the results of modelling the data with K-nearest neighbors on the unscaled data against the scaled data.
 The considered data is the wine recognition dataset \href{https://archive.ics.uci.edu/dataset/109/wine}{https://archive.ics.uci.edu/dataset/109/wine}.
@@
 The first feature has values in the [0,1000] range while the second feature has values contained in [1,10].
-The unscaled and scaled version are compared in figure \ref{fig:feature_scaler_example}
+The unscaled and scaled version are compared in figure \ref{fig:feature_scaler_example}.
 \begin{figure}[!h]
@@
 On the left, we can see that distances between categories are impacted solely by the larger scale feature.
-Conversely, on the right, we have that the two features contributing equally in determining the neighbors.
+Conversely, on the right, we have that the two features contribute equally in determining the neighbors.
 
@@
 \subsubsection{Nystrom decomposition}
-The Nyström approximation involves storing a submatrix of the whole kernel matrix. Thus, storage and computational cost are reduced to $O(nm)$ and $O(nm^2)$ respectively.
-Nyström works by selecting $m<n$ points, callled representative points; it approximates K as
+The Nyström approximation involves storing a submatrix of the whole kernel matrix. Thus, storage and computational costs are reduced to $O(nm)$ and $O(nm^2)$ respectively.
+Nyström works by selecting $m<n$ points, callled representative points; it approximates $K$ as
 \begin{equation}
@@
 \begin{itemize}
-    \item query: folder containing Scopus data and scripts to generate bibliometric survey plots in section \ref{literature_review}
-    \item kqr.py: file implementing our custom kernel quantile regression
-    \item thesis: folder containing tex files
+    \item query/: folder containing Scopus data and scripts to generate bibliometric survey plots in section \ref{literature_review}.
+    \item thesis/: folder containing tex files for thesis.
+    \item beamer/: folder containing tex files for pptx thesis defence.
     \item experiments: folder containing train, test script for each of the experiments carried out in the thesis alongside with the binary files of the pickled models.
+    \item experiments/src/kernel\_quantile\_regression/kqr.py: file implementing our custom kernel quantile regression.
+    \item experiments/Data: cleaned data used in our experimental studies.
+    \item experiment/plots: folder storing plots resulting from experiments.
+    \item experiment/eda/: scripts for exploratory data analysis.
+    \item experiments/point/: scripts for experiments point framework.
+    \item experiments/train\_test/: scripts for training, testing and generating tables for experiments probabilistic framework.
+    \item experiments/utils/: utility functions used in extracting, transforming and cleaning raw data.
+    \item extradata/: folder containing data for experiments in section \ref{appendix:quantile_regressor_extensive_comparison}.
+    \item requirements.txt: pip freeze of python packages used.
 \end{itemize}

### 2024-05-21 thesis/input_files/analysis.tex
thesis update

@@
     \item Day of the week
-    \item Is holiday
+    \item Is holiday: a binary variable for holidays, where holiday$=1$ and working day$=0$
     \item Weather temperature
@@
 \subsection{Multiple linear regression}
-To get started, standard multiple linear regression has been applied, see fig \ref{fig:mlr_price} for a visualisation. 
+To get started, standard multiple linear regression has been applied, see figure \ref{fig:mlr_price} for a visualisation. 
 \begin{figure}
@@
 % \end{figure}
-Prophet alongside with the right features is a good model in the context of electricity price forecasting, it is capcable of catching the trend, the peak and the lows.
+Prophet alongside with the right features is a good model in the context of electricity load forecasting, it is capcable of catching the trend, the peak and the lows.
 
@@
 \subsection{Long short term memory}
-We used the torch library in order to code our LSTM regressor.
+We used the \href{https://pytorch.org}{torch} library in order to code our LSTM regressor.
 Various hyperparameters combinations were tried during the tuning phase. The final architecture we set on follows
@@
 \subsection{Kernel ridge regression}
-Subsequently, kernel ridge regression was considered. The kernel considered is the radial basis gaussian function.
-Cross validation was carried jointly over the rbf kernel bandwith and the regularization constant.
+Subsequently, kernel ridge regression was considered. The kernel considered is the radial basis Gaussian function.
+Cross validation was carried jointly over the RBF kernel bandwith and the regularization constant.
 The RMSE achieved is 9.86, figure \ref{fig:krnridge_price} reports the prediction.
@@
 \subsection{Results}
-This section reports the table comparing the considered methods scores 
+This section reports the tables comparing the considered methods scores.
 % in the context of electricity load forecasting.
@@
 
-The MAE scores are contained in table \ref{tab:point_MAE}. Similarly to above, we can conclude that kernel methods stand out for achieving also the lowest MAE score.
+The MAE scores are contained in table \ref{tab:point_MAE}. Similarly to above, we can conclude that kernel methods stand out for achieving also the lowest MAE score; with KrnSVR being the top one.
 \begin{table}[!ht]
@@
 
-Finally table \ref{tab:point_MAPE} reports the MAPE score for each method. What can be concluded is that kernel regressors are also the best in terms of MAPE score; with KrnSVR being the top one.
+Finally table \ref{tab:point_MAPE} reports the MAPE score for each method. Its results are in line with the other tables, hence suggesting the goodness of kernel methods.
+% What can be concluded is that kernel regressors are also the best in terms of MAPE score; with KrnSVR being the top one.
 \begin{table}[!ht]
@@
 We continue our experimental analysis by considering the probabilistic framework.
-The goal of this section is comparing between different quantile regressors and study the choiche of the kernel function.
+The goal of this section is comparing between different quantile regressors and study the choice of the kernel function.
 
@@
 Notice, for kernel quantile regression (KQR), the kernel of choice in this setting is the absolute Laplacian.
-Finally notice, the loss function utilized in this section is the normalized pinball loss.
+Additionally notice, the loss function utilized in this section is the normalized pinball loss.
 % normalized by the magnitude of the average load of each country.
@@
         \makebox[\linewidth]{\includegraphics[width = 1.5\textwidth]{images/DE_load_CI.png}}
-        \caption[Probabilistic forecast for load in Germany]{Load 90\% confidence interval for Germany Energy charts data using KQR Laplacian: \small \textmd{Electric load probabilistic forecast Germany 2022.
+        \caption[Probabilistic forecast for load in Germany]{Load 90\% confidence interval for Germany Energy charts (2022) data using KQR Laplacian.
 The black line is the observed path for the load.
 The 90\% confidence interval bands are plotted in green.
-Lower and upper red lines denote the 95\% and 5\% quantile forecast respectively.}}
+Lower and upper red lines denote the 95\% and 5\% quantile forecast respectively.}
         \label{fig:DE_load_CI}
@@
         \makebox[\linewidth]{\includegraphics[width = 1.5\textwidth]{images/CH_load_CI.png}}
-        \caption[Probabilistic forecast for load in Switzerland]{Load 90\% confidence interval for Switzerland Energy charts data using KQR Laplacian: \small \textmd{Electric load probabilistic forecast for Switzerland 2022.
+        \caption[Probabilistic forecast for load in Switzerland]{Load 90\% confidence interval for Switzerland Energy charts data (2022) using KQR Laplacian.
 The black line is the observed path for the load.
 The 90\% confidence interval bands are plotted in green.
-Lower and upper red lines denote the 95\% and 5\% quantile forecast respectively.}}
+Lower and upper red lines denote the 95\% and 5\% quantile forecast respectively.}
         \label{fig:CH_load_CI}
@@
 \begin{itemize}
-    \item Is holiday: a binary variable for holidays where holiday$=1$ and working day$=0$
-    \item Day of week: an ordinal categorical variable corresponding to the day of the week , e.i.\ Monday$=0$,$\dots$, Sunday$=6$
+    \item Is holiday: a binary variable for holidays, where holiday$=1$ and working day$=0$
+    \item Day of week: an ordinal categorical variable corresponding to the day of the week, e.i.\ Monday$=0$,$\dots$, Sunday$=6$
 \end{itemize}
-Table \ref{tab:energy_chart_de_additional_predictors} reports the model score on the German data while table \ref{tab:energy_chart_ch_additional_predictors} is the one corresponding to the Suisse data.
+Table \ref{tab:energy_chart_de_additional_predictors} reports the model scores on the German data while table \ref{tab:energy_chart_ch_additional_predictors} is the one corresponding to the Suisse data.
 
@@
     \centering
-    \caption{PF pinball loss for load in Germany (2022) model\_2}
+    \caption{Pinball loss for load in Germany (2022) model\_2}
     \label{tab:energy_chart_de_additional_predictors}
@@
 % now comment tables
-What can be concluded from this first study is that kernel quantile regression equipped with the absolute laplacian kernel outperformed almost always its contenders.
+What can be concluded from this first study is that kernel quantile regression equipped with the absolute Laplacian kernel outperformed almost always its contenders.
 
@@
 This section analyzes which kernel function best fits to the characteristics of the load data.
-The kernel functions considered are: Gaussian RBF, absolute Laplacian, Matern 1.5, Matern 2.5,Linear, Periodic, Polynomial, Sigmoid, and Cosine. 
+The kernel functions considered are: Gaussian RBF, absolute Laplacian, Matern 1.5, Matern 2.5, Linear, Periodic, Polynomial, Sigmoid, and Cosine. 
 Details concerning the hyperparameter selection for the different kernels can be found in \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.metrics.pairwise}{sklearn.metrics.pairwise} and \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.gaussian_process}{sklearn.gaussian\_process}.
@@
 
-The predictors making up the dataset follow:
+The predictors making up the datasets follow:
 \begin{itemize}
@@
 \end{itemize}
-Table \ref{tab:secures_met_de} shows results for Germany, table \ref{tab:secures_met_ch} is the one for Switzerland and table corresponds to Austria \ref{tab:secures_met_at}.
+Table \ref{tab:secures_met_de} shows results for Germany, table \ref{tab:secures_met_ch} is the one for Switzerland and table \ref{tab:secures_met_at} corresponds to Austria.
 
@@
     \centering
-\caption{Pinball loss kernel comparison for load in German (2021)}
+\caption{Pinball loss kernel comparison for load in Germany (2021)}
 \label{tab:secures_met_de}
@@
     \includegraphics[width=\linewidth]{images/price_task_6_gaussian_rbf.png}
-    \caption[[Prediction price track task 6, Gaussian RBF kernel]]{Price 90\% confidence interval task 6: \small \textmd{Electricity price probabilistic forecast for the $13^{\text{th}}$ July $2013$.
+    \caption[Prediction price track task 6, Gaussian RBF kernel]{Price 90\% confidence interval task 6.
+Electricity price probabilistic forecast for the $13^{\text{th}}$ July $2013$.
 The black line is the observed path for the price.
 The $90\%$ confidence interval bands are plotted in green.
-Lower and upper red lines denote the $95\%$ and $5\%$ quantile forecast respectively.}}
+Lower and upper red lines denote the $95\%$ and $5\%$ quantile forecast respectively.}
     % \Description{}
@@
 \subsubsection{GEFCom2014 load track}
-The GEFCom2014 load track constitutes a good setting for analyzing the performance of KQR in the context of medium term load forecasting.
+The GEFCom2014 load track constitutes a good setting for analyzing the performance of KQR in the context of medium term load forecasting (MTLF).
 In this track we were faced with the challenge of predicting the load for the next month without the availability of weather temperature forecasts. 
-Therefore, the primary task was to first accurately predict the weather and temperatures, and then model the load accordingly.
+Therefore, the primary task was to first accurately predict the weather temperatures, and then model the load accordingly.
 Since there were no attributes available for humidity or wind speed, we chose to predict weather temperatures by aggregating historical temperature data across different dimensions such as day, month, and hour.
@@
     \item Day of week: an ordinal categorical variable for the day of the week
-    \item Is holiday: a dummy variable for holidays
+    \item Is holiday: a binary variable for holidays, where holiday$=1$ and working day$=0$
     \item w avg: average of weather temperatures across all the $25$ stations
@@
     \makebox[\linewidth]{\includegraphics[width = 1.5\textwidth]{images/load_task_9_laplacian.png}}
-    \caption[Prediction load track task 9, Laplacian kernel]{Load 90\% confidence interval task 9 using KQR laplacian: \small \textmd{Electric load probabilistic forecast for June 2011.
+    \caption[Prediction load track task 9, Laplacian kernel]{Load 90\% confidence interval task 9 using KQR laplacian.
 The black line is the observed path for the load.
 The 90\% confidence interval bands are plotted in green.
-Lower and upper red lines denote the 95\% and 5\% quantile forecast respectively.}}
+Lower and upper red lines denote the 95\% and 5\% quantile forecast respectively.}
     \label{fig:load_task_9}
@@
 The \href{https://github.com/luca-pernigo/kernel_quantile_regression}{github repo} hosting the source code includes also the script implementing the experiments along with the cleaned datasets; this contribution is intended to forster reproducibility in research.
-    \item Achieved superior performance of kernel quantile regression compared to standard quantile regressor algorithm. 
+    \item Achieved superior performance of kernel quantile regression compared to standard quantile regressor algorithms. 
     \item Created and made available datasets suitable for algorithms benchmarking considering data from the DACH region (Germany, Switzerland and Austria).
@@
 We observe that KQR shows favorably and can forecast both the short and medium-term horizon.
-To conclude, we proceed with summarizing the conclusion we drew from our experimental analysis.
+To finish, we proceed with summarizing the conclusions we drew from our experimental analysis.
 Kernel methods showed good performance both in the context of point and probabilistic forecasting.
 Specifically, kernel quantile regression resulted almost always in superior performance compared to standard quantile regressors.
-The absolute Laplacian kernel turned out to be the most suitable kernel according to the characteristic of load.
+The absolute Laplacian kernel turned out to be the most suitable kernel according to the characteristic of electricity forecasting.
 Finally, we proved the validity of our kernel quantile regression implementation.
 We achieved this by showing how it compared favourably to the top entries of the GEFCom2014 competition.
-In particular, up to nowx research work focused on the short term horizon capabilities of kernel quantile regression.
+In particular, up to now research work focused on the short term horizon capabilities of kernel quantile regression.
 In this thesis work, we were also able to show its suitability on the medium term horizon.

### 2024-05-21 thesis/input_files/implementation.tex
thesis update

@@
 \subsection{Prophet}
-The prophet models has been applied by employing the python API provided by Meta \href{https://facebook.github.io/prophet/docs/quick_start.html}{prophet}.
+The \href{https://facebook.github.io/prophet/docs/quick_start.html}{prophet} model has been applied by employing the python API provided by Meta.
 
@@
 This estimator is compatible with useful scikit-learn in built methods like gridsearch, crossvalidation and scoring rules.
-Moreover, this code is compatible with sklearn in built kernel functions \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.metrics.pairwise}{sklearn.metrics.pairwise} and \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.gaussian_process.kernels}{sklearn.gaussian\_process.kernels}, functions supported by our kernel quantile regression include: gaussian rbf, absolute laplacian, mater 1/2, matern 3/2, matern 5/2, linear, cosine, sigmoid, periodic, polynomial and custom composition of kernels.
+Moreover, this code is compatible with sklearn in built kernel functions \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.metrics.pairwise}{sklearn.metrics.pairwise} and \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.gaussian_process.kernels}{sklearn.gaussian\_process.kernels}, functions supported by our kernel quantile regression include: gaussian rbf, absolute laplacian, matern 3/2, matern 5/2, linear, cosine, sigmoid, periodic, polynomial and custom composition of kernels.
 % NOTICE, these kernel implementations differ in the form of passed parameters in the former we pass in the numerator gamma while in the latter we pass l, which stands for length scale in the denominator.

### 2024-05-21 src/CDF.py
dirclean

@@ -1,75 +1,0 @@
-import numpy as np
-from src.miscellaneous import find
-
-def cdf_funct(x):
-    ans=np.arange(len(x)) / float(len(x)-1)
-    return ans
-
-
-def inv_cdf_funct(y,x):
-    # x has to be in format np([])
-    x=x.reshape(x.size,)
-
-    sorted_x_axis=np.sort(x)
-    idx_pos= y*float(len(sorted_x_axis))
-    # print(idx_pos)
-    # -1 because when calculating cdf_funct
-    # the function appends an additional value to
-    # the sorted array. However, when it comes to the inverse
-    # we deal with the original sorted array. That is the one
-    # without the additional x
-    ans=  sorted_x_axis[int(idx_pos)-1]
-    return ans
-
-
-
-
-def Phi(xn, xcdf, ycdf):
-    i=find(xn, xcdf)
-    if(xn<xcdf.min()):
-        ans=0
-        return ans
-    if(xn>xcdf.max()):
-        ans=1
-        return ans
-    
-
-    ans=ycdf[i]*(xcdf[i+1]-xn)/(xcdf[i+1]-xcdf[i])+ycdf[i+1]*(xn-xcdf[i])/(xcdf[i+1]-xcdf[i])
-    return ans
-
-
-
-
-def phi(xn, xpdf, ypdf):
-    i=find(xn, xpdf)
-    if(xn<xpdf.min()):
-        ans=ypdf[0]
-        return ans
-    if(xn>xpdf.max()):
-        ans=ypdf[-1]
-        return ans
-    
-
-    ans=ypdf[i]*(xpdf[i+1]-xn)/(xpdf[i+1]-xpdf[i])+ypdf[i+1]*(xn-xpdf[i])/(xpdf[i+1]-xpdf[i])
-    
-    return ans
-
-
-
-
-def phi_inv(u, xpdf,ypdf, xcdf, ycdf, tol=1e-3):
-    xn=0
-    bol=True
-    
-    if(u==1):
-        return xpdf[-1]
-    
-    while(bol):
-        xnew=xn-(Phi(xn,xcdf, ycdf)-u)/phi(xn, xpdf, ypdf)
-        if(abs(xn-xnew)<tol):
-            bol=False
-    
-        xn=xnew
-        # print(xn)
-    
-    return xn

### 2024-05-21 src/__init__.py

@@ -1,12 +1,0 @@
-import sys
-import os
-
-# Get the directory of the current file (__init__.py)
-current_dir = os.path.dirname(os.path.abspath(__file__))
-
-# Get the parent directory of the current directory (src)
-parent_dir = os.path.abspath(os.path.join(current_dir, os.pardir))
-
-# Add the parent directory to the Python path
-sys.path.append(current_dir)
-sys.path.append(parent_dir)

### 2024-05-21 src/bsm.py

@@ -1,46 +1,0 @@
-# Beasley Springer Moro algorithm
-
-import numpy as np
-
-def inverse_normal_cdf(u_array):
-    approximate_normal_array = np.vectorize(inverse_normal_cdf_u)
-    arr = approximate_normal_array(u_array)
-    return arr
-
-
-def inverse_normal_cdf_u(u):
-  a0= 2.50662823884
-  a1= -18.61500062529
-  a2= 41.39119773534
-  a3= -25.44106049637
-
-  b0= -8.47351093090 
-  b1= 23.08336743743
-  b2= -21.06224101826
-  b3= 3.13082909833
-
-  c0= 0.3374754822726147
-  c1= 0.9761690190917186
-  c2= 0.1607979714918209
-  c3= 0.0276438810333863
-  c4= 0.0038405729373609
-  c5= 0.0003951896511919
-  c6= 0.0000321767881768
-  c7= 0.0000002888167364
-  c8= 0.0000003960315187
-  
-  
-  y=u-0.5
-  if abs(y)<0.42:
-    r=y*y
-    x=y*(((a3*r+a2)*r+a1)*r+a0)/((((b3*r+b2)*r+b1)*r+b0)*r+1)
-  else:
-    r=u
-    if(y>0):
-      r=1-u
-    r=np.log(-np.log(r))
-    x=c0+r*(c1+r*(c2+r*(c3+r*(c4+r*(c5+r*(c6+r*(c7+r*c8)))))))
-    if (y<0):
-      x= -x   
-  return x
-

### 2024-05-21 src/kde.py

@@ -1,156 +1,0 @@
-from sklearn.metrics.pairwise import pairwise_distances
-import numpy as np
-from src.piv_cholesky import piv_chol
-from src.gramian import eval_kernel, gram
-import FMCA
-import matplotlib.pyplot as plt
-from mpl_toolkits.axes_grid1 import make_axes_locatable
-
-# kde computes approximate the probability density function at a specified x point
-# suppose x=2, the kde function will return the kernel mean embedding estimate of p(2)
-# x_sample is the dataset you have, whose underlying distribution is unknown
-# kernel is just a kernel object from sklearn.gaussian_process.kernels
-def kde(x_sample, x,kernel):
-    ans=[eval_kernel(x_sample[i], x, kernel) for i in range(len(x_sample))]
-    ans=np.sum(ans)/len(x_sample)
-    return ans
-
-
-
-
-def nyström_kde(x_sample, x_tilde, x,kernel, Kmt, Kmn):
-    # compute kernel feature map, that is k(x_tilde, .) where the dot in this notebook is the variable x
-    ans=[eval_kernel(x_tilde[i], x, kernel) for i in range(len(x_tilde))]
-
-    # compute alphas https://proceedings.mlr.press/v162/chatalic22a/chatalic22a.pdf
-    alphas=1/len(x_sample)*Kmt@Kmn@np.ones(len(x_sample))
-
-    # compute summation
-    ans=alphas.T@ans
-    return ans
-
-
-# K_x_sample=gramian(x_sample, x_sample)
-# tol=piv cholesky tolerance
-# otherwise you can define a maximum number of steps=piv cholesky steps
-
-# K_x_sample is the kernel matrix computed on x_sample array of data
-def piv_cholesky_kde(x_sample, x,kernel, K_x_sample, tol, steps):
-    B,L,m, indices= piv_chol(np.matrix(K_x_sample), tol,steps)
-    # get the pivots xs in the pivoted cholesky decomposition
-    x_pivots=x_sample[indices]
-
-    Km=gram(x_pivots, x_pivots)
-    
-    Kmn=gram(x_pivots, x_sample)
-    ans=[eval_kernel(x_pivots[i], x, kernel) for i in range(len(x_pivots))]
-
-    # compute alphas https://proceedings.mlr.press/v162/chatalic22a/chatalic22a.pdf
[…]

### 2024-05-21 src/miscellaneous.py

@@ -1,24 +1,0 @@
-
-# find index i s.t. vec[i]<x<vec[i+1]
-def find(x, vec):
-    l=0
-    r=vec.size-1
-    while True:
-        m=int((l+r)/2)
-        
-        if x>vec[m]:
-            l=m
-        elif x<vec[m]:
-            r=m
-        
-        if r-l==1:
-            break
-
-    return l
-
-# def find(x, vec):
-#     for i in range(len(vec)):
-#         if x>vec[i] and x<vec[i+1]:
-#             ans=i
-#             break
-#     return i

### 2024-05-21 src/mmd.py

@@ -1,18 +1,0 @@
-import numpy as np
-import FMCA
-
-def mmd(x,y, kernel_type, kernel_prm):
-    m=x.size
-    n=y.size
-    cov=FMCA.CovarianceKernel(f"{kernel_type}", kernel_prm)
-    Kx=cov.eval(x,x)
-    Ky=cov.eval(y,y)
-    Kxy=cov.eval(x,y)
-
-    A=(np.sum(Kx))/(m*m)
-    B= (np.sum(Kxy))/(m*n)
-    C=(np.sum(Ky))/(n*n)
-
-    mmd2=A-2*B+C
-
-    return np.sqrt(mmd2)

### 2024-05-21 src/pinball_crps.py

@@ -1,41 +1,0 @@
-# function to compute pinball score
-
-from sklearn.metrics import mean_pinball_loss
-import numpy as np
-
-def pinball(tau, Yt_pred, Yt):
-    """
-    - quantile tau
-    - predicted Yt_pred
-    - true Yt
-    """
-    # use sklearn to compute pinball at a certain quantile tau
-    pinball_tau=mean_pinball_loss(Yt, Yt_pred, alpha=tau)
-    return pinball_tau
-
-
-def CRPS(Y_pred_test, Y_test):
-    """
-    Y_test= observed Y for t in test:=[T, T']
-    Y_pred_test=predicted for Y_test
-    """
-    # pinball requires input to be passed as np([[]])
-    Y_pred_test=np.array([Y_pred_test])
-    Y_test=np.array([Y_test])
-
-
-    CRPS=0
-    # for every t in test
-    for t in range(Y_test):
-        # for every tau in [0.01, ..., 0.99]
-        for tau in range(1,100):
-            # discrete approximation of CPRS
-            CRPS+=pinball(tau/100, Y_pred_test[t], Y_test[t])
-            print("tau: ", tau, "t: ", t, "CRPS: ", CRPS)
-            print("------------------")
-    
-    CRPS/=Y_test.size
-
-    return CRPS
-
-

### 2024-05-21 src/piv_cholesky.py

@@ -1,53 +1,0 @@
-import numpy as np
-
-def piv_chol(K, tol, steps=0):
-# K=square symmetric kernel matrix
-    m=1
-    d=np.diag(K)
-    # e=np.linalg.eig(K)[1]
-    e=np.matrix(np.zeros(shape=(K.shape[0],1)))
-
-    L=np.matrix(np.zeros(shape=(K.shape[0],K.shape[0])))
-    B=np.matrix(np.zeros(shape=(K.shape[0],K.shape[0])))
-    err=np.sum(np.abs(d))
-    i=0
-    indices=[]
-    N=K.shape[0]-1
-# steps=0 by default, if nothing is specified the maximum number of steps will be K.shape[0]-1
-# otherwise if any steps variable is specified, the number of maximum steps can be set to be smaller
-    if steps!=0:
-        N=steps-1
-    while err>tol and i<=N:
-            
-        j=np.argmax(d)
-        indices+=[j]
-        l_m_hat=K[:,j]-L@L.T[:,j]
-
-
-        ej=e.copy() 
-        ej[j]=1
-        
-        b_m_hat=ej-B@L.T[:,j]
-    
-        
-        lm=l_m_hat/np.sqrt(d[j]+1e-10)
-        
-        bm = b_m_hat/np.sqrt(d[j]+1e-10)
-                  
-        L[:,i]=lm
-        B[:,i]=bm
-
-        
-        d=d-np.squeeze(np.asarray(lm))*np.squeeze(np.asarray(lm))
-
-        err=np.sum(np.abs(d))
-        
-        m+=1
-        i+=1
-
-    L=L[:,0:(m+1)]
-    B=B[:,0:(m+1)]
-
-    # print(err)
-    ans=[B, L, m, indices]
-    return ans

### 2024-05-21 src/poly_fit.py

@@ -1,159 +1,0 @@
-import numpy as np
-from scipy.optimize import curve_fit
-
-import math
-
-from src.cdf import cdf_funct
-from src.miscellaneous import find
-
-def spl_polyfit(x,y):
-    s=3
-    # take subset to make interpolation smoother
-    x=np.sort(x)
-    x_sub=x[0:-1:s]
-    # append the max as end point to be interpolated
-    if x_sub[-1]!=x.max:
-        x_sub=np.append([x_sub], [x.max()])
-
-    while x_sub.size%s!=0:
-        # print("*")
-        i=np.random.random_integers(0,x_sub.size-1)
-        x_sub=np.delete(x_sub, i)
-
-    y_sub=cdf_funct(x_sub)
-
-    fit_spl_rpol=np.zeros(y_sub.size)
-    for i in range(int(y_sub.size/s)):
-        fit_spl_rpol[i*s:(i+1)*s]=polyfit(y_sub[i*s:(i+1)*s],beta_rat_polyfit(y_sub[i*s:(i+1)*s], np.sort(x_sub)[i*s:(i+1)*s], order=1))
-    
-    return [x_sub, y_sub,fit_spl_rpol]
-
-def simulate_spl_polyfit(unif,y, fit_spl_rpol):
-    
-    sim_spl_rpol=np.zeros(unif.size)
-
-    for i in range(len(unif)):
-        idx=find(unif[i], y)
-        t1=(y[idx+1]-unif[i])/(y[idx+1]-y[idx])
-        t2=(unif[i]-y[idx])/(y[idx+1]-y[idx])
-        sim_spl_rpol[i]=fit_spl_rpol[idx]*t1+fit_spl_rpol[idx+1]*t2
-    
-    return sim_spl_rpol
-
-
-def beta_rat_polyfit(x, y, order):
-    n=y.size
-    A=np.zeros([n, order*2+1])
-    A[:,0]=np.ones(y.size)
-    for i in range(order):
-        A[:,i+1]=x**(i+1)
-    for i in range(order):
-        A[:,order+i+1]=-x**(i+1)*y
-    
-    # b=np.linalg.inv(A.T@A)@A.T@y
-    
-    # # solve with QR decomposition
-    Q, R = np.linalg.qr(A)
-    p = np.dot(Q.T, y)
-    b=np.dot(np.linalg.inv(R), p)
-    
-
-    return b
-
[…]

### 2024-05-21 src/spline.py

@@ -1,29 +1,0 @@
-import numpy as np
-from scipy import interpolate
-
-
-def spline_evaluate(t, tck):
-
-    t_vec=tck[0]
-    cntr_points=tck[1]
-    n=tck[2]
-
-    e_vec=np.zeros(cntr_points.size)
-    k=0
-    while t_vec[n+k+1]<t:
-        k+=1
-    
-    for i in range(n+1):
-        e_vec[i]=cntr_points[i+k]
-
-
-    for j in range(1,n+1,1):
-        for i in range(0,n-j+1,1):
-            e_vec[i]=(t_vec[n+i+k+1]-t)/(t_vec[n+i+k+1]-t_vec[i+k+j])*e_vec[i]+(t-t_vec[i+k+j])/(t_vec[n+i+k+1]-t_vec[i+k+j])*e_vec[i+1]
-    
-    return e_vec[0]
-
-
-def control_points(x,y,n):
-    ans=interpolate.splrep(x,y, k=n)[1]
-    return ans

### 2024-05-21 thesis/input_files/acknowledgements.tex
acknowledgements

\chapter*{Acknowledgements}
\thispagestyle{empty}

% Thank:
% - supervisor, co
First and foremost I would like to express my deepest gratitude to my advisor and co-advisor, Prof. Multerer and Dr. Baroli for their invaluable feedback, continuous support and patience during my Master's Thesis. Furthermore, I thank them for having given me the opportunity to collaborate through this thesis work.
\\
% - master director
I could not have undertaken this journey without my Master directors Prof. Schenk and Prof. Wit. I am extremely grateful for their guidance, mentorship and advice throughout my Master studies.
\\
% - professors of faculty
% From Polya How to solve it
% forster independent thinking  by challenging curiosity, stimulating questions and delving into the motives behind theory.
% tutti i professori che ho avuto, shaped my interest, saziato sete curiosità
I would like to extend my deepest appreciation to all professors I had during my studies at Usi. Their stimulating questions and push to enquire motives behind theory challenged my curiosity, shaped my interests and fostered my independent thinking.
\\
% - research group 
% discussions, suggestions
% listening the challenges I faced and insightful suggestions
I also had great pleasure of discussing with the research group of Prof. Multerer the challenges I faced during this thesis work. I thank them for their affability  and insightful suggestions.
\\
% - classmates
% friends I made along the way
% mutual support and good times we had together
% be it during a group revision session before an exam of during light hearted moments.
Special thanks to my cohort members and to the friends I made along the way. I will never forget the good times we had together.
\clearpage

### 2024-05-21 thesis/abstract.tex
corrected abstract

@@
   % methods
-  In doing so, we focused on the family of kernel methods. In addition, we also compared them against several statistical and machine learning techniques.
+  In doing so, we focused on the family of kernel methods. We have compared them against several other statistical and machine learning techniques.
   % results
-  Our results showed the goodness of kernel methods in the field of electricity forecasting both in terms of point and probabilistic forecasting.
-In particular, in the probabilistic context, our experiments showed the validity of kernel quantile regression equipped with the absolute Laplacian kernel.
+  Our results showed the feasibility of kernel methods in the field of electricity forecasting both in terms of point and probabilistic forecasting.
+In particular, in the probabilistic context, our experiments show the validity of kernel quantile regression equipped with the Laplacian kernel.
   % discussion
-  These findings indicate kernel methods are well suited to the characteristic of electricity.
+  These findings indicate that kernel methods are well suited to the characteristics of electricity.
   Anyone interested in forecasting energy should consider them when faced with the choice of the model.
-  They can be employed stand alone or combined with other valid methods into ensembles.
+  They can be employed stand-alone or combined with other valid methods into ensembles.
 

### 2024-05-21 thesis/input_files/problemdescription.tex
corrected intro

@@
-Individuals and organizations constantly face situations of uncertainty thus the need of robust forecasting methods.
-Such methods are crucial in the process of taking informed decision and for strategic planning.
+Individuals and organizations constantly face situations of uncertainty, thus the need for robust forecasting methods.
+Such methods are crucial in the process of taking informed decisions and for strategic planning.
 \\
@@
 Events like the California electricity crisis are further motivating the choice of the electricity sector as subject of our studies, see \cite{california}.
-Moreover, the process of deregulation lead to an increasing interest in the field of electricity forecasting (EF) within the academic community, see \ref{fig:epf_evolution}.
-In addtion, the United Nations have identified the right to access affordable, reliable, sustainable and modern energy as one of their 17 SDGs \cite{un_sdgs}.
+Moreover, the process of deregulation lead to an increasing interest in the field of electricity forecasting (EF) within the academic community, see figure \ref{fig:epf_evolution}.
+In addtion, the United Nations have identified the right to access affordable, reliable, sustainable and modern energy as one of their 17 sustainable development goals (SDGs) \cite{un_sdgs}.
 Finally, the electricity market has a set of features that make it unique: electricity cannot be stored in an efficient way and supply and demand have to be matched instantly.
@@
 Furthermore, the adoption of smart meters provides power market companies with a huge amount of consumer data.
-This can enable them to better model consumer preferences.
+This enables them to better model consumer preferences.
 \\
 Transmission system operators' (TSO) main goal is to match supply and demand, generally TSO do so by increasing or decreasing the generation.
-Thus, from their point of view forecasting is critical for balancing the electricity netwrok.
+Thus, from their point of view forecasting is critical for balancing the electricity network.
 Probabilistic forecasting may be useful to power producers, traders and consumers in order to improve their decision making process and managing risk.
-This holds in particular for traders, because probabilistic forecasts enables them to simulate scenarios and carry out stress tests.
+This holds in particular for traders, because probabilistic forecasts enable them to simulate scenarios and carry out stress tests.
 Other possible applications include: control of storage, demand side response, anomaly detection, network design and planning, simulating inputs and handling missing data.
@@
 Point prediction, also called deterministic forecasting in the literature \cite{EPF_review}, is all about predicting a particular value in time.
-On the other hand, with probabilistic forecasting we aim at predicting either a prediction interval, quantiles or a probability distribution for each point in time \cite{nowotarski}.
+On the other hand, with probabilistic forecasting we aim at predicting either an interval, quantiles or a probability distribution for each point in time \cite{nowotarski}.
 For this reason, probabilistic forecasts are more informative than point forecasts.
 This is why the interest of the research community is shifting towards them.
-A probabilistic forecast can be turned into a point forecast by simply taking its expecatation.
+A probabilistic forecast can be turned into a point forecast by simply taking its expectation.
 Alternatively, a probabilistic forecast can be derived from a point one by modeling the residuals of the point prediction.
@@
 Following, Sections \ref{ch:point} and \ref{ch:prob} introduce the state of the art methods in the context of point and probabilistic forecasting respectively.
-Section \ref{eda} goes on with the etract-load-transform (ETL) pipeline and the exploratory analysis (EDA). 
-Implementation details are included in \ref{implementation}.
-Finally, section \ref{analysis} presents the experiments, the results and discusses models' strenghts, weaknesses and possible improvements.
+Section \ref{eda} goes on with the extract-load-transform (ETL) pipeline and the exploratory data analysis (EDA). 
+Implementation details are included in Section \ref{implementation}.
+Finally, Section \ref{analysis} presents experiments results and discusses models' strenghts, weaknesses and possible improvements.

### 2024-05-21 thesis/input_files/literaturereview.tex
May 13th 2024
corrected literature review

@@
 \\
-Within this research community, it emerged the need for more homogeneity in the choice of the error valuation metrics (Section \ref{metrics}), data quality and in the way of comparing model performances \cite{EPF_review}.
+Within this research community, the need for more homogeneity in the choice of the error valuation metrics emerged (Section \ref{metrics}), data quality and in the way of comparing model performances \cite{EPF_review}.
 As a solution, \cite{lago} proposes a checklist to aid evaluating the meaningfulness of new research.
@@
 \section{Electricity forecasting classification}
-Electricity forecasting is a vague term and it is used in the literature to refer to the whole field.
-Thus, in order to introduce some clarity it is useful to classify the range of EF articles in terms of their core attributes.
+Electricity forecasting is a vague term and is used in the literature to refer to the whole field.
+Thus, in order to introduce some clarity, it is useful to classify the range of EF articles in terms of their core attributes.
 \\
@@
 In terms of forecasting horizons, we can group EF into four major categories: very short term forcasting (VSTF), short term load forecasting (STF), medium term forecasting (MTF) and long term forecasting (LTF).
-Consensus in the literature is to use one day, two weeks and three years respectilvely \cite{hong_phd} as cut off horizons; see \ref{fig:time} for a visualisation.
+Consensus in the literature is to use one day, two weeks and three years respectilvely \cite{hong_phd} as cut off horizons; see figure \ref{fig:time} for a visualisation.
 \begin{figure}
   \includegraphics[width=\textwidth]{images/time_2.png}
-  \caption{Time classsification \cite{prob_elf}}
+  \caption{Time classification \cite{prob_elf}}
   \label{fig:time}
@@
 Each of the two has its advantages and disadvantages.
-Point forecasts are easier to generate and less computationally intensive while probabilistic forecast are more informative.
+Point forecasts are easier to generate and less computationally intensive while probabilistic forecasts are more informative.
 Industry and research efforts have focused primarily on point forcasting.
-Nevertheless, interest in probabilistic forecasting has risen considerably over the last years due to renewable integration requirements, introduction of smart grids and increased market competitivness.
+Nevertheless, interest in probabilistic forecasting has risen considerably over the last years due to renewable integration requirements, introduction of smart grids and increased market competitiveness.
 \section{Bibliographic analysis}
@@
   \includegraphics[width=\textwidth]{images/point_vs_prob.jpg}
-  \caption{Point versus probabilistic pubblications over the past years}
+  \caption{Point versus probabilistic publications over the past years}
   \label{fig:point_vs_prob}
@@
   \includegraphics[width=\textwidth]{images/cs_stat_both.jpg}
-  \caption{Pubblications by method over the past years}
+  \caption{Publications by method over the past years}
   \label{fig:cs_stat_both}
@@
 Figure \ref{fig:subject_area} depitcs the EF publications by subject area.
-What can be concluded, is that the bulk of pubblication come from engineering, computer science, mathematics and econometrics.
+What can be concluded, is that the bulk of publication come from engineering, computer science, mathematics and econometrics.
 \begin{figure}
   \includegraphics[width=\textwidth]{images/subject_area.png}
-  \caption{Pubblications by subject area}
+  \caption{Publications by subject area}
   \label{fig:subject_area}
@@
 The kernel theory needed for this thesis work is covered in Section \ref{kernel_theory}. 
-For an introduction to kernel methods, we refer to \cite{learning_with_kernels, hofmann2006review}.
+For an introduction to kernel methods, we refer to \cite{learning_with_kernels, hofmann2006review} and \cite{shawe2004kernel}.
 \\

### 2024-05-22 thesis/appendix.tex
thesis update
22 May 2024

@@
 \\
-The scikit-learn team addressed this disadvantage of gridsearch by introducing the halving gridsearch method \cite{scikithalvinggridsearch} (2020).
+The scikit-learn team addressed this disadvantage of gridsearch by introducing the halving gridsearch method \citeW{scikithalvinggridsearch} (2020).
 Such technique has proved itself to greatly speed up hyperparameter tuning.
@@
 \end{itemize}
-Many other data scaling algorithms exists, we refer the reader to a thorough comparison \cite{scikitscalers}.
+Many other data scaling algorithms exists, we refer the reader to a thorough comparison \citeW{scikitscalers}.
 \\
-We conclude this subsection with a custom example that motivates the need of feature scaling \cite{scikitscale_example}.
+We conclude this subsection with a custom example that motivates the need of feature scaling \citeW{scikitscale_example}.
 The idea is to compare the results of modelling the data with K-nearest neighbors on the unscaled data against the scaled data.
@@
     \item experiments/Data: cleaned data used in our experimental studies.
-    \item experiment/plots: folder storing plots resulting from experiments.
+    \item experiment/plots/: folder storing plots resulting from experiments.
     \item experiment/eda/: scripts for exploratory data analysis.

### 2024-05-22 thesis/input_files/energymarket.tex
22 May 2024

@@
 Finally, the voltage is reduced and the energy is moved into the distribution network.
-In a nutshell, transmision lines carry electric power from stations to substations while distribution lines carry electricity from substations to load points such as businesses, industries and homes, see figure \ref{electricity_network}.
+In a nutshell, transmission lines carry electric power from stations to substations while distribution lines carry electricity from substations to load points such as businesses, industries and homes, see figure \ref{electricity_network}.
 
@@
 \\
-It is worth to mention the two biggest power exchanges: Epex and NordPool. Epex operates throughout continental europe. NordPool covers the nordinc and baltic contries.
+It is worth to mention the two biggest power exchanges: Epex and NordPool. Epex operates throughout continental europe. NordPool covers the nordic and baltic countries.
 \\
@@
 The intraday market consists of continous trading 24/7.
-A trade is executed as soon as a buy and a sell order mathch.
+A trade is executed as soon as a buy and a sell order match.
 Electricity can be traded an hour, half-hour, quarter or also 5 minutes before delivery.
@@
 \subsection{Price peaks}
-Electricity market are characterized by spikes in their spot prices; when this happens, the system price jumps abruptly and then drops back within a very short period.
-This spikness follows from the non storability of electricity; that is, electricity has to be consumed as it is produced.
+Electricity markets are characterized by spikes in their spot prices; when this happens, the system price jumps abruptly and then drops back within a very short period.
+This spikeness follows from the non storability of electricity; that is, electricity has to be consumed as it is produced.
 Therefore, extreme load fluctuations combined with generation outages or transmission failures can result in price spikes.
@@
 With zonal market clearing pricing, prices may differ between zones but are the same within the same area.
-On the other hand, local marginal price is made up by summing the transmission congestion cost, generation marginal cost and the cost of marginal losses at different buses; buses is where a electicity line or several lines are connected.
+On the other hand, local marginal price is made up by summing the transmission congestion cost, generation marginal cost and the cost of marginal losses at different buses; buses is where a electricity line or several lines are connected.
 \begin{figure}[!h]
     \includegraphics[width=\textwidth]{images/eu_zonal.png}
-    \caption{Zonal pricing EU \cite{eu_zonal}}
+    \caption[Zonal pricing EU]{Zonal pricing EU \citeW{eu_zonal}}
     \label{fig:eu_zonal}
@@
     \includegraphics[width=\textwidth]{images/us_local.png}
-    \caption{Local pricing US \cite{us_local}}
+    \caption[Local pricing US]{Local pricing US \citeW{us_local}}
     \label{fig:us_local}
@@
 The reason is that, smart metering and the electricity network digitalisation  will enable system operators to better monitor the network, plan their investments and manage infrastructure. 
-All of this motivates for the need of efficient and robust tools to analyze the vast data resulting from the digitalisation of the power sector see \cite{declean,andreadou2016telecommunication}.
+All of this motivates for the need of efficient and robust tools to analyze the vast data resulting from the digitalisation of the power sector, see \cite{declean,andreadou2016telecommunication}.
 

### 2024-05-22 thesis/input_files/metrics.tex
22 May 2024

@@
 Proper evaluation methods guide researchers in choosing the model that best fits their needs; thus, this chapter is dedicated to the most common evaluation metrics adopted by academics in the field of electricity forecasting.
-Error metrics and measures vary depending on whethter we are concerned with point or probabilistic forecasts.
+Error metrics and measures vary depending on whether we are concerned with point or probabilistic forecasts.
 Additionally, note that the latter can take different forms which therefore requires different measures.
@@
 However, a drawback of such measures is that we cannot use them to compare accuracy between time series which have different magnitudes.
-For instance, a day ahead error of 1kWh is negligible when considering a daily demand of 100kWn while the same error is considerably big when daily demand is 2kWh.
+For instance, a day ahead error of 1kWh is negligible when considering a daily demand of 100h while the same error is considerably big when daily demand is 2kWh.
 This consideration leads to relative accuracy scores, between those the mean absolute percentage error (MAPE) is by far the most popular.
@@
 \end{definition}
-Where delta is the prediction interval (PI) width, that is $\delta=u_t-l_t$ where $u_t$ is the prediction interval upper threshold and $l_t$ is the prediction interval lower threshold at time step $t$.
+Where $\delta$ is the prediction interval (PI) width, that is $\delta=u_t-l_t$, $u_t$ is the prediction interval upper threshold and $l_t$ is the prediction interval lower threshold at time step $t$.
 This score penalizes observations falling outside the prediction interval and rewards narrow prediction intervals.
 \section{Continous ranked probability score}
-The continous ranked probability score (CRPS) measures the differece between the estimated cumulative distribution $\hat{F}$ and the empirical cumulative density function (CDF).
+The continous ranked probability score (CRPS) measures the difference between the estimated cumulative distribution $\hat{F}$ and the empirical cumulative density function (CDF).
 \begin{definition}\label{def_crps}

### 2024-05-22 thesis/input_files/paper1.tex
22 May 2024

@@
 To illustrate the idea behind the kernel trick consider the following example.
-Suppose $x, y \in \mathbb{R}^2$ and assume $\phi(x)=(x_{1}^{2},x_{2}^{2},\sqrt{2}x_{1}x_{2})$, then the inner product in the feature space is $x_{1}^{2}y_{1}^{2},x_{2}^{2}y_{2}^{2}+2x_{1}x_{2}y_{1}y_{2}$.
+Suppose $x, y \in \mathbb{R}^2$ and assume $\phi(x)=(x_{1}^{2},x_{2}^{2},\sqrt{2}x_{1}x_{2})$, then the inner product in the feature space is $x_{1}^{2}y_{1}^{2}+x_{2}^{2}y_{2}^{2}+2x_{1}x_{2}y_{1}y_{2}$.
 Notice that this is the same of $\langle \phi(x), \phi(y) \rangle_{\mathcal{H}}$; thus the kernel trick consists of just using $k(x,y):=(x^\intercal y)^2$.
@@
 \\
-$\langle f,g \rangle_{\mathcal{F}}=:\sum\limits_{i=1}^n\sum\limits_{j=1}^m a_i \beta_j k_i(x_i,y_j)=\sum\limits_{i=1}^n a_i g(x_i)= \sum\limits_{j=1}^m \beta_j f(y_j)$
+$\langle f,g \rangle_{\mathcal{F}}:=\sum\limits_{i=1}^n\sum\limits_{j=1}^m a_i \beta_j k_i(x_i,y_j)=\sum\limits_{i=1}^n a_i g(x_i)= \sum\limits_{j=1}^m \beta_j f(y_j)$
 \\
 Notice, $\langle f, f \rangle_{\mathcal{F}}= \alpha^\intercal K \alpha \geq 0 \ \forall f \in \mathcal{F}$, where $K$ is the kernel matrix.
-Moreover, $\langle f, g \rangle_{\mathcal{F}}$ is real valued, symmetric and bilinear. All these facts implies that the just introduced inner product is a valid metric.
+Moreover, $\langle f, g \rangle_{\mathcal{F}}$ is real valued, symmetric and bilinear. All these facts imply that the just introduced inner product is a valid metric.
 \\
@@
 We have then, that $f_n(x)$ is a Cauchy sequence of real numbers.
-Thus, we can let $g(x)=\lim_{n\to \infty} f_n(x)$ and include all such limits functions in $\mathcal{F}$, doing so we complete $\mathcal{F}$ and we obtain the Hilbert space $\mathcal{H}$ associated to the kernel $k$.
+Thus, we can let $g(x):=\underset{n\to \infty}\lim f_n(x)$ and include all such limits functions in $\mathcal{F}$, doing so we complete $\mathcal{F}$ and we obtain the Hilbert space $\mathcal{H}$ associated to the kernel $k$.
 Having constructed the feature space $\mathcal{H}$, the mapping $\phi(x)$ is defined as follows
@@
         k(x, \cdot) \in \mathcal{H} &\quad \ \forall x \in \mathcal{X}    \\
-        \langle  f, k(x, \cdot) \rangle_\mathcal{H} = f(x) &\quad \ \forall f\in \mathcal{H}, x \in \mathcal{X}
+        \langle  f, k(x, \cdot) \rangle_\mathcal{H} = f(x) &\quad \ \forall f\in \mathcal{H}, \ x \in \mathcal{X}
     \end{align}
@@
 The function $k(\cdot, x)$ is the reproducing kernel evaluated at point $x$.
-Furthermore, note that $k_{x}$ is itself a function lying in it.
+Furthermore, note that $k(\cdot, x)$ is itself a function lying in it.
 
@@
 \begin{proof}
-    Consider a finite set of points $x_1, \cdots, x_n$, a vector $z \in \mathbb{R}^n$ and let $K_1$ and $K_2$ be the kernel matrices associated to $k_1$ and $k_2$ evaluted on these points.
+    Consider a finite set of points $x_1, \cdots, x_n$, a vector $z \in \mathbb{R}^n$ and let $K_1$ and $K_2$ be the kernel matrices associated to $k_1$ and $k_2$ evaluated on these points.
     \\
@@
     \begin{align}
-    z^\intercal (K_1+K_2) z= z^\intercal K_1 z + z^\intercal K_2 z \geq 0
+    z^\intercal (K_1+K_2) z&= z^\intercal K_1 z + z^\intercal K_2 z \geq 0
     \\
-    z^\intercal aK z= a z^\intercal K z \geq 0
+    z^\intercal aK z&= a z^\intercal K z \geq 0
     \end{align}
@@
 
-Notice, in the Matern family $\Gamma(\cdot)$ is the gamma function while $Y(\cdot)$ is the Bessel function of the second kind.
+Notice, in the Matern family, $\Gamma(\cdot)$ is the gamma function while $Y(\cdot)$ is the Bessel function of the second kind.
 \\
@@
 \begin{itemize}
-    \item $p=0,\nu=\frac{1}{2} \implies k(x,y)=\exp\left(\frac{-\|x-y|\|_{2}}{l}\right)$
+    \item $p=0,\nu=\frac{1}{2} \implies k(x,y)=\exp\left(-\frac{\|x-y|\|_{2}}{l}\right)$
     \item $p=1,\nu=\frac{3}{2} \implies 

### 2024-05-22 thesis/input_files/problemdescription.tex
22 May 2024

@@
 Moreover, the process of deregulation lead to an increasing interest in the field of electricity forecasting (EF) within the academic community, see figure \ref{fig:epf_evolution}.
-In addtion, the United Nations have identified the right to access affordable, reliable, sustainable and modern energy as one of their 17 sustainable development goals (SDGs) \cite{un_sdgs}.
+In addtion, the United Nations have identified the right to access affordable, reliable, sustainable and modern energy as one of their 17 sustainable development goals (SDGs) \citeW{un_sdgs}.
 Finally, the electricity market has a set of features that make it unique: electricity cannot be stored in an efficient way and supply and demand have to be matched instantly.

### 2024-05-22 thesis/introduction.tex
22 May 2024

% Some commands used in this file
\newcommand{\package}{\emph}

\chapter{Introduction}
This is version \verb-v1.4- of the template.

We assume that you found this template on our institute's website, so
we do not repeat everything stated there.  Consult the website again
for pointers to further reading about \LaTeX{}.  This chapter only
gives a brief overview of the files you are looking at.

\section{Features}
\label{sec:features}

The rest of this document shows off a few features of the template
files.  Look at the source code to see which macros we used!

The template is divided into \TeX{} files as follows:
\begin{enumerate}
\item \texttt{thesis.tex} is the main file.
\item \texttt{extrapackages.tex} holds extra package includes.
\item \texttt{layoutsetup.tex} defines the style used in this document.
\item \texttt{theoremsetup.tex} declares the theorem-like environments.
\item \texttt{macrosetup.tex} defines extra macros that you may find
  useful.
\item \texttt{introduction.tex} contains this text.
\item \texttt{sections.tex} is a quick demo of each sectioning level
  available.
\item \texttt{refs.bib} is an example bibliography file.  You can use
  Bib\TeX{} to quote references.  For example, read
  \cite{bringhurst1996ets} if you can get a hold of it.
\end{enumerate}

\subsection{Extra package includes}

The file \texttt{extrapackages.tex} lists some packages that usually
come in handy.  Simply have a look at the source code.  We have
added the following comments based on our experiences:
\begin{description}
\item[REC] This package is recommended.
\item[OPT] This package is optional.  It usually solves a specific
  problem in a clever way.
\item[ADV] This package is for the advanced user, but solves a problem
  frequent enough that we mention it. Consult the package's
  documentation.
\end{description}

As a small example, here is a reference to the Section \emph{Features}
typeset with the recommended \package{varioref} package:
\begin{quote}
  See Section~\vref{sec:features}.
\end{quote}

\subsection{Layout setup}

This defines the overall look of the document -- for example, it
changes the chapter and section heading appearance.  We consider this
a `do not touch' area.  Take a look at the excellent \emph{Memoir}
documentation before changing it.

In fact, take a look at the excellent \emph{Memoir} documentation,
full stop.

\subsection{Theorem setup}

This file defines a bunch of theorem-like environments.

\begin{theorem}
  An example theorem.
\end{theorem}

\begin{proof}
  Proof text goes here.
\end{proof}

Note that the q.e.d.\ symbol moves to the correct place automatically
if you end the proof with an \texttt{enumerate} or
\texttt{displaymath}.  You do not need to use \verb-\qedhere- as with
\package{amsthm}.

\begin{theorem}[Some Famous Guy]
  Another example theorem.
\end{theorem}

\begin{proof}
  This proof
  \begin{enumerate}
  \item ends in an enumerate.
  \end{enumerate}
\end{proof}

\begin{proposition}
  Note that all theorem-like environments are by default numbered on
  the same counter.
\end{proposition}

\begin{proof}
  This proof ends in a display like so:
  \begin{displaymath}
    f(x) = x^2.
  \end{displaymath}
\end{proof}

\subsection{Macro setup}

For now the macro setup only shows how to define some basic macros,
and how to use a neat feature of the \package{mathtools} package:
\begin{displaymath}
  \abs{a}, \quad \abs*{\frac{a}{b}}, \quad \abs[\big]{\frac{a}{b}}.
\end{displaymath}

### 2024-05-23 thesis/input_files/energymarket.tex
May 23rd 2024

@@
 \\
-During the nineties various wholesale electricity markets were established; for example in Chile, Great Britain, the scandinivian block, Australia, New Zealand, the US and Canada.
+During the nineties various wholesale electricity markets were established; for example in Chile, Great Britain, the Scandinivian block, Australia, New Zealand, the US and Canada.
 Consequently, new participants, each with a specific task, entered the power market scenario.
@@
 Conversely, power exchanges are created through private agreements between generators, distributors and traders.
-This is the model followed by most of european countries.
+This is the model followed by most of European countries.
 The MCP in power exchanges is determined by the intersection of the aggregate supply curve and the aggregate demand curve, right panel figure \ref{fig:pool_vs_echange}.
 \\
-It is worth to mention the two biggest power exchanges: Epex and NordPool. Epex operates throughout continental europe. NordPool covers the nordic and baltic countries.
+It is worth to mention the two biggest power exchanges: Epex and NordPool. Epex operates throughout continental Europe. NordPool covers the nordic and baltic countries.
 \\
-Moreover, we can differentiate between two popular types of auctions: uniform-price/marginal and pay as bid/discriminatory.
+Moreover, we can differentiate between two popular types of auctions: uniform-price/marginal and pay-as-bid/discriminatory.
 Within the uniform price setting, suppliers offering for less than the clearing price are paid that price.
 Analogously, consumers bidding more than the clearing price pay that price.
-On the other hand, in pay as bid auctions, suppliers are paid the exact price they bid for.
+On the other hand, in pay-as-bid auctions, suppliers are paid the exact price they bid for.
 
@@
 The causes of negative prices are inflexible power generation plants and low demand.
-With inflexibility, it is mean the fact that power sources cannot be switched off and restarted quickly and efficiently.
+With inflexibility, it is meant the fact that power sources cannot be switched off and restarted quickly and efficiently.
 Thus, producers are faced with the decision of either stopping and then restarting their power plant or selling their energy for a negative price; that is they pay consumers for consuming their energy.
@@
 Zonal market clearing pricing (ZMCP) and local marginal price (LMP) are two types of pricing schemes utilized to handle it.
-The former is adopted in EU contries while the latter is used in the US, figure \ref{fig:eu_zonal} and figure \ref{fig:us_local}.
+The former is adopted in European contries while the latter is used in the United States, figure \ref{fig:eu_zonal} and figure \ref{fig:us_local}.
 With zonal market clearing pricing, prices may differ between zones but are the same within the same area.

### 2024-05-23 thesis/input_files/point.tex
corrected intro
May 23rd 2024

@@
 where the parameter p is the order of the autoregressive (AR) part, d is the degree of integrated differencing and q is the order of the moving average (MA) model.
-$\epsilon_{n-j}$ are the observed errors in the past while the differenced term $y_n^{(d)}$ is defined recursively as $y_n^{(d)}=y_n^{(d-1)}-y_{n-1}^{(d-1)}$.
+$\epsilon_{n-j}$ are the observed errors in the past while the differenced term $y_n^{(d)}$ is defined recursively as $y_n^{(d)}:=y_n^{(d-1)}-y_{n-1}^{(d-1)}$.
 \\
-ARIMA model are compactly written as
+ARIMA models are compactly written as
 \begin{equation}
@@
 where $B$ is the backward shift operator, $B^h y_t=y_{t-h}$ and $\psi(B)$ is shorthand for $\psi(B)=1-\psi_1B-\dots-\psi_p B^p$.
-Similarly, $\theta(B)=1+\theta_1B+\dots+\theta_q B^q$ and $\nabla_h$ is the lag h differencing operator $\nabla_h y_t=(1-B^h)y_t=y_t-y_{t-h}$
+Similarly, $\theta(B)=1+\theta_1B+\dots+\theta_q B^q$ and $\nabla_h$ is the lag $h$ differencing operator $\nabla_h y_t=(1-B^h)y_t=y_t-y_{t-h}$
 \\
 The most popular procedure for estimating this class of models is the Box-Jenkins method \cite{box2015time}.
-The order p is identified by taking the lags at which the sample partial autocorrelation function (PACF) falls outside its 95\% confidence interval.
-In the same way, the order q is estimated by considering the sample autocorrelation function (ACF) of the observed time series.
-Finally, we compare ARIMA models with different p,d,q (in a neighbourhood of the identified orders) and choose the ones minimising either AIC or BIC criteria.
+The order $p$ is identified by taking the lags at which the sample partial autocorrelation function (PACF) falls outside its 95\% confidence interval.
+In the same way, the order $q$ is estimated by considering the sample autocorrelation function (ACF) of the observed time series.
+Finally, we compare ARIMA models with different $p,d,q$ (in a neighbourhood of the identified orders) and choose the ones minimising either AIC or BIC criteria.
 % Notice, models are obtained through the maximum likelihood estimation framework.
@@
 The second term $(P,D,Q)_m$ represents the autoregressive structure of the seasonal pattern, where $m$ indicates the number of seasonalities of our time series.
-For instance, in modelling series of hourly data with daily periodicity, we would set m=24.
+For instance, in modelling series of hourly data with daily periodicity, we would set $m=24$.
 Writing SARIMA in compact notation we have
@@
 $f_i$ are smooth functions which are fitted using cubic smoothing splines.
-g is called link function and characterizes the specific GAM model \cite{hastie2017generalized}. Approximations for $f_i$ are obtained through an iterative procedure, the backfitting algorithm.
+$g$ is called link function and characterizes the specific GAM model \cite{hastie2017generalized}. Approximations for $f_i$ are obtained through an iterative procedure, the backfitting algorithm.
 
@@
 Particularly successful in time series forecasting is the Prophet model \cite{taylor2018forecasting}, developed by Meta.
-This additive model is capable of handling non linear trends, holiday effects and yearly weekly and daily seasonalities.
+This additive model is capable of handling non linear trends, holiday effects and yearly weekly, and daily seasonalities.
 Thus, the reason why we choose it as a valid benchmark to compare against.
 \section{K-nearest neighbours regression}
-K-nearest neighbours regression forecasts by averaging the most similar k instances in the training set \cite{macqueen1967some}.
-K is the algorithm hyperparameter, it stands for the number of neighbours; a small k leads to overfitting while a big k leads to underfitting.
+K-nearest neighbours regression forecasts by averaging the most similar $k$ instances in the training set \cite{macqueen1967some}.
+$k$ is the algorithm hyperparameter, it stands for the number of neighbours; a small $k$ leads to overfitting while a big $k$ leads to underfitting.
 Since it is a metric based algorithm, it is important to normalize data in order to give equal weights to features with different scales, see \ref{appendix:normalization}.
@@
 \section{Support vector regression}
-Developed at the AT\&T Bell Laboratories by Vapnik et al.
-\cite{cortes1995support} \cite{vapnik1997support}, support vector machines (SVMs) are one of the most popular techniques within the field of statistical learning.
+Developed at the AT\&T Bell Laboratories by Vapnik et al.\ \cite{cortes1995support} \cite{vapnik1997support}, support vector machines (SVMs) are one of the most popular techniques within the field of statistical learning.
 \\
 The goal of support vector regression is finding a function $f(x)$ with at most $\epsilon$ deviation from the actual observed data $y_i$ for every $i$ and as flat as possible. 
-Put differently, we would like a model to keep error less than the $\epsilon$ threshold.  
+Put differently, we would like a model to keep errors less than the $\epsilon$ threshold.  
 %https://stats.stackexchange.com/questions/5945/understanding-svm-regression-objective-function-and-flatness
@@
     \includegraphics[width=\textwidth]{images/svm_simple.png}    
-    \caption{Support vector regression,\cite{learning_with_kernels}}
+    \caption{Support vector regression \cite{learning_with_kernels}}
     \label{fig:svm_simple}
 \end{figure}
-Smola et al. \cite{smola2004tutorial} point out that considering the dual formulation makes our optimisation problem easier to solve. Doing so we have the equivalent optimisation problem
+Smola et al.\ \cite{smola2004tutorial} point out that considering the dual formulation makes our optimisation problem easier to solve. Doing so we have the equivalent optimisation problem
 \begin{equation}
@@
 \end{equation}
-Rearranging the gradient of the primal lagrangian with respect to $w$, we obtain the so called support vector expansion.
+Rearranging the gradient of the primal Lagrangian with respect to $w$, we obtain the so called support vector expansion.
 \begin{equation}\label{eq:support vector expansion}
@@
 This means that $w$ can be completely described as a linear combination of the training features $x_i$.
-Notice, the complexity of the function representation is independent of the feature space dimensionality, but depends only on the number of support vectors; that is those point $i$ for which $(\lambda_i-\lambda_i^*)\neq 0$
-Moreover, we do not needd to compute $w$ explicitly in order to evaluate $f(x)$.
+Notice, the complexity of the function representation is independent of the feature space dimensionality, but depends only on the number of support vectors; that is those points $i$ for which $(\lambda_i-\lambda_i^*)\neq 0$
+Moreover, we do not need to compute $w$ explicitly in order to evaluate $f(x)$.
 That is because, equation \ref{eq:support vector expansion} implies
@@
 \end{equation}
-Employing the Karush-Kuhn-Tucker conditions, $b$ can be retrieved easily. Particulalry, the condition that the product between the constraints and the dual variable has to vanish.
+Employing the Karush-Kuhn-Tucker conditions, $b$ can be retrieved easily. Particularly, the condition that the product between the constraints and the dual variable has to vanish.
 Consequently for any of the data points with associated $\lambda_i, \lambda^*_i \in (0,C)$, the equality $b=y_i-\langle w, x_i-\epsilon\rangle$ has to hold.
@@
         \\
-        \tilde{C}_t=&tahn(W_c\dot[h_{t-1},x_t]+b_C)
+        \tilde{C}_t=&tahn(W_C\dot[h_{t-1},x_t]+b_C)
     \end{aligned}
@@
     a strictly increasing real valued function $g:[0,\infty) \to \mathbb{R}$, and the loss function $E:(\mathcal{X}\times \mathbb{R}^2)^n \to \mathbb{R} \cup \{\infty\}$.
-    Consider the regularized risk functional $Risk[f]:f \mapsto  E\left((x_{1},y_{1},f(x_{1})),\ldots, (x_{n},y_{n},f(x_{n}))\right)+g\left(\| f\| \right)$.
-Then any minimizer of the empirical risk $Risk[f]$ can be represented as $ f^{*}(\cdot )=\sum\limits_{i=1}^{n}\alpha _{i}k(\cdot ,x_{i})$
+    Consider the regularized risk functional $Risk[f]:f \mapsto  E\left((x_{1},y_{1},f(x_{1})),\ldots, (x_{n},y_{n},f(x_{n}))\right)+g\left(\| f\|_{\mathcal{H}} \right)$.
+Then any minimizer of the empirical risk can be represented as $ f^{*}(\cdot )=\sum\limits_{i=1}^{n}\alpha _{i}k(\cdot ,x_{i})$
 \end{theorem}
@@
     Hence, we can conclude that setting $\nu=0$ decreases the second term while the first term is not affected.
-Thus it follows that any minimizer of the risk functional of equation \ref{eq:kernel ridge} must be of the following form
+Thus, it follows that any minimizer of the risk functional of equation \ref{eq:kernel ridge} must be of the following form
     \begin{equation}
@@
 \end{figure}
-Comparing theese pictures with figure \ref{fig:svr1}, it can be concluded that introducing kernels allows support vector regression to handle the non linearities in the data.
+Comparing these pictures with figure \ref{fig:svr1}, it can be concluded that introducing kernels allows support vector regression to handle the non linearities in the data.

### 2024-05-25 thesis/input_files/analysis.tex
May 25th 2024

@@
 % \\
-Building on the theory introduced in \ref{ch:point} and \ref{ch:prob}, this section covers the experiments carried out and the results obtained respectively.
+Building on the theory introduced in section \ref{ch:point} and \ref{ch:prob}, this section covers the experiments carried out and the results obtained respectively.
 
@@
 This section carries out a comparative study between the state of the art methods for point forecasting, introduced in \ref{ch:point}.
-As use case, we will consider the task of predicting the electric load from the GEFCom 2014 dataset.
+As use case, we will consider the task of predicting the electric load from the GEFCom2014 dataset.
 In such setting we considered the following regressors
@@
 We added the weather temperature, the square of it and the categorical variable for holidays effect as regressors.
-Furthermore, we also applied a log transformation to the dependent variable.
+Furthermore, we also applied a log transformation to the dependent variables.
 Doing so, RMSE went down to 10.29.
-Forecast is visualized in \ref{fig:prophet_price2}.
+Forecast is visualised in figure \ref{fig:prophet_price2}.
 % , moreover figure \ref{fig:prophet_price2.1} and figure \ref{fig:prophet_price2.2} break down the trend and the seasonalities respectively.
@@
 % \end{figure}
-Prophet alongside with the right features is a good model in the context of electricity load forecasting, it is capcable of catching the trend, the peak and the lows.
+Prophet alongside with the right features is a good model in the context of electricity load forecasting, it is capable of catching the trend, the peak and the lows.
 
@@
 In doing so, we cross validated the number of neighbours to get the best model instance.
-The forecast is visualized in figure \ref{fig:knn_price}, RMSE is 12.54.
+The forecast is visualised in figure \ref{fig:knn_price}, RMSE is 12.54.
 \begin{figure}[!h]
@@
 \end{figure}
-What it can be said is that, K-nearest neighbour is capable of predicting prices well by averaging past data.
+What it can be said is that, K-nearest neighbour is capable of predicting load well by averaging past data.
 
@@
 Coming up we have support vector regression.
-In applying this model we used gridsearch crossvalidation to search for the best regularization parameter C.
+In applying this model we used gridsearch crossvalidation to search for the best regularisation parameter C.
 Support vector regression achieves an RMSE of 23.24, for the prediction see figure \ref{fig:svr_price}.
@@
 Subsequently, kernel ridge regression was considered. The kernel considered is the radial basis Gaussian function.
-Cross validation was carried jointly over the RBF kernel bandwith and the regularization constant.
+Cross validation was carried jointly over the RBF kernel bandwith and the regularisation constant.
 The RMSE achieved is 9.86, figure \ref{fig:krnridge_price} reports the prediction.
@@
 % |       |X          |    
-Table \ref{tab:point_RMSE} reporst the RMSE scores for each of the considered methods.
+Table \ref{tab:point_RMSE} reports the RMSE scores for each of the considered methods.
 It can be seen that kernel based methods are the ones achieving the lowest RMSE; specifically, we have KrnSVR, KrnRidge and KNN being among the top methods on the majority of tasks.
@@
 Details concerning the hyperparameter selection for the different kernels can be found in \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.metrics.pairwise}{sklearn.metrics.pairwise} and \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.gaussian_process}{sklearn.gaussian\_process}.
-The pool of data for this study has been create by combining \href{https://zenodo.org/records/7907883}{SECURES-Met} data (predictors) \cite{Formayer2023} and the load data from \href{https://www.energy-charts.info/index.html?l=en&c=DE}{energy charts}.
-SECURES-Met data consists of historical data up to the end of $2020$ while from 2021 onward, the data consists of forecasts modeled by the the European Centre for Medium-Range Weather Forecasts (ECMWF).
+The pool of data for this study has been created by combining \href{https://zenodo.org/records/7907883}{SECURES-Met} data (predictors) \cite{Formayer2023} and the load data from \href{https://www.energy-charts.info/index.html?l=en&c=DE}{energy charts}.
+SECURES-Met data consists of historical data up to the end of $2020$ while from 2021 onward, the data consists of forecasts modeled by the European Centre for Medium-Range Weather Forecasts (ECMWF).
 Therefore, in this experimental study, we used the whole data of $2020$ as the training set and then tested our kernels on the $2021$ data.
@@
 \subsection{GEFCom2014}
-The follwing subsection compares our kernel quantile regression against the price and load track entries of the GEFCom2014 competiton.
+The following subsection compares our kernel quantile regression against the price and load track entries of the GEFCom2014 competition.
 Error measure of the competition was the pinball loss, see section \ref{metrics}, averaged over the $99$ quantiles, $q \in \{i/100\}_{i=1}^{n=99}$.
@@
 In this track, the top entries came from the teams: Tololo, Team Poland, GMD, and C3 Green Team; for a breakdown of each method's attributes, see \cite[Table 8]{hong2016probabilistic}.
-Finally, the Gaussian RBF probabilistic prediction for the $13$th July $2013$ zonal price at the $90\%$ confidence interval is visualized in Figure \ref{fig:price_task_6_gaussian_rbf}.
+Finally, the Gaussian RBF probabilistic prediction for the $13^{th}$ July $2013$ zonal price at the $90\%$ confidence interval is visualized in Figure \ref{fig:price_task_6_gaussian_rbf}.
 
@@
 \subsubsection{GEFCom2014 load track} 
-The GEFCom2014 load track constitutes a good setting for analyzing the performance of KQR in the context of medium term load forecasting (MTLF).
+The GEFCom2014 load track constitutes a good setting for analysing the performance of KQR in the context of medium term load forecasting (MTLF).
 In this track we were faced with the challenge of predicting the load for the next month without the availability of weather temperature forecasts. 
@@
     \makebox[\linewidth]{\includegraphics[width = 1.5\textwidth]{images/load_task_9_laplacian.png}}
-    \caption[Prediction load track task 9, Laplacian kernel]{Load 90\% confidence interval task 9 using KQR laplacian.
+    \caption[Prediction load track task 9, Laplacian kernel]{Load 90\% confidence interval task 9 using KQR Laplacian.
 The black line is the observed path for the load.
@@
     \makebox[\linewidth]{\includegraphics[width = 1.5\textwidth]{images/sub_boxplot_task_9_laplacian_.png}}
-    \caption{Boxplot for first two days of probabilistic forecast task 9 using KQR laplacian}
+    \caption{Boxplot for first two days of probabilistic forecast task 9 using KQR Laplacian}
     \label{fig:load_task_9}
@@
 This section concludes this thesis work with some final remarks.
-This thesis was concerned with electricity forecsating, in particular we focused on the probabilistic framework.
+This thesis was concerned with electricity forecasting, in particular we focused on the probabilistic framework.
 Probabilistic forecasting is still a relatively new research area, lately research efforts are shifting towards this setting.
-This is due to the fact that probabilistic forecasts are more informative than point ones and better suited for the current and future electricity landscape; think for example about the higher uncertainty resulting from market liberalization or renewables integration requirements.
+This is due to the fact that probabilistic forecasts are more informative than point ones and better suited for the current and future electricity landscape; think for example about the higher uncertainty resulting from market liberalisation or renewables integration requirements.
 
@@
 We observe that KQR shows favorably and can forecast both the short and medium-term horizon.
-To finish, we proceed with summarizing the conclusions we drew from our experimental analysis.
+To finish, we proceed with summarising the conclusions we drew from our experimental analysis.
 Kernel methods showed good performance both in the context of point and probabilistic forecasting.
 Specifically, kernel quantile regression resulted almost always in superior performance compared to standard quantile regressors.
-The absolute Laplacian kernel turned out to be the most suitable kernel according to the characteristic of electricity forecasting.
+The absolute Laplacian kernel turned out to be the most suitable kernel according to the characteristics of electricity forecasting.
 Finally, we proved the validity of our kernel quantile regression implementation.

### 2024-05-25 thesis/input_files/data_exploratory.tex
May 25th 2024

@@
 Next, we will explain the extract transform load pipeline (ETL) that we set up in order to ease the workflow of our studies.
-Finally, in order to to better understand patterns within the data, we carry out an exploratory data analysis.
+Finally, in order to better understand patterns within the data, we carry out an exploratory data analysis.
 
 \section{GEFCom2014}
-The first class of datasets we identified were the ones from the GEFCom 2014 competition. The data is freely hosted on Dr.\ Hong blog \cite{hong2016probabilistic}.
+The first class of datasets we identified were the ones from the GEFCom2014 competition. The data is freely hosted on Dr.\ Hong blog \cite{hong2016probabilistic}.
 The main reason is that, these datasets are considered an excellent test case for comparing predictive models between the EF community.
 Additionally, the scores of the competing models are freely available, this enables us to carry out a clean and transparent comparative study.
-The GEFCom 2014 consisted of four tracks: price, load, wind power and solar power forecasting. 
+The GEFCom2014 consisted of four tracks: price, load, wind power and solar power forecasting. 
 The GEFCom2014 folder contains a zip file for each of the four tracks.
@@
 This data consisted of time series for the locational marginal price, for the zonal load and for the system load.
-The data covers the time interval ranging from the 1st of January 2011 to the 15th of June 2013, see figure \ref{fig:price_track_fig1}.
+The data covers the time interval ranging from the $1^{st}$ of January 2011 to the $15^{th}$ of June 2013, see figure \ref{fig:price_track_fig1}.\
 Additionally, as the competition went on, the real observed data of the previous tasks were made available.
@@
 \subsection{GEFCom2014 load track}
-In the load track, contestant were asked to provide one month ahead hourly probabilistic forecasts on a rolling basis for 15 consecutive rounds.
+In the load track, contestants were asked to provide one month ahead hourly probabilistic forecasts on a rolling basis for 15 consecutive rounds.
 To get started, in the first round, organizers provided 69 months of hourly load data (from 01/01/2005 to 30/10/2010) and 117 months of hourly temperature data (from 01/01/2001 to 30/10/2010).
@@
 For each of those we have the prediction of the competition benchmark model and the train file upon which we will fit our model.
-Any data from the various task folders are shifted by one month between eachother.
+Any data from the various task folders are shifted by one month between each other.
 \\
@@
 We applied the same kind of EDA to the GEFCom load data.
-Figure \ref{fig:gefcom_load_historical} depicts the hisorical load while figure \ref{fig:gefcom_w_avg_historical} reports the historical weather temperature.
-Finally, figure \ref{fig:gefcom_load_vs_w_avg} shows the scatter plot between the the load and the weather.
+Figure \ref{fig:gefcom_load_historical} depicts the historical load while figure \ref{fig:gefcom_w_avg_historical} reports the historical weather temperature.
+Finally, figure \ref{fig:gefcom_load_vs_w_avg} shows the scatter plot between the load and the weather.
 These plots motivate the selection of weather temperature as one of the predictors for the load.
@@
 Next, since the data does not provide the location for the target utility, the best we can do is selecting the average of weather temperatures as predictor variables.
-This independent variable is named w\_avg.
+This independent variable is named w avg.
 The same thing applies in the cleaning of the test data with the addition of an if clause for handling the different naming convention and formatting used in the last track of the challenge.
-The python etl pipeline has been then automated to loop every task folder through bash scripting.
+The Python ETL pipeline has been then automated to loop every task folder through Bash scripting.
 
@@
 It provides data for electricity power, electricity potential capacity, energy generation, market prices, environmental measurements and also scenario simulations.
-Notice, this data is retrieved from the European network of transmission system operators for electricity \href{https://www.entsoe.eu/}{ENTSOE} and made available thanks to the \href{https://ise.fraunhofer.de/en.html}{Fraunhofer Institute for Solar Energy Systems ISE}.
+Notice, this data is retrieved from the European network of transmission system operators for electricity \href{https://www.entsoe.eu/}{(ENTSOE)} and made available thanks to the \href{https://ise.fraunhofer.de/en.html}{Fraunhofer Institute for Solar Energy Systems ISE}.
 \\
@@
 
-In the second experiment we compared various kernels in the context of load foreacsting.
+In the second experiment we compared various kernels in the context of load forecasting.
 For this study we considered the data of \href{https://zenodo.org/records/7907883}{SECURES-Met}.
@@
 Notice, the forecasted data is splitted in two files according to the scenarios RCP4.5 and RCP8.5.
-In our experiment we considered the former.
+In our experiments we considered the former.
 In this setting the independent variables available are 
@@
 
-For what concers the data from SECURES-Met we first identified the data interesting to us.
-Next, we formatted all the csv to a coherent format, gave coloumns meaningful names, sliced for the years 2020 (train) and 2021 (test) and finally we merged all single csv\textsubscript{s} together.
+For what concerns the data from SECURES-Met we first identified the data interesting to us.
+Next, we formatted all the csv to a coherent format, gave columns meaningful names, sliced for the years 2020 (train) and 2021 (test) and finally we merged all single csv\textsubscript{s} together.
 
-All the manipulations involved in the ETL pipeline have been carried out by combining python and bash scripting.
+All the manipulations involved in the ETL pipeline have been carried out by combining Python and Bash scripting.
 
@@
 To get started we visualized the historical load data in figure \ref{fig:CH_historical_load_2021}.
-As expected, electricity load exhibits a seasonal patter with highs in the winter and lows during the summer season.
+As expected, electricity load exhibits a seasonal pattern with highs in the winter and lows during the summer season.
 \begin{figure}[!h]
@@
 \end{figure}
-Next we built scatter plots of the independent continous variables against the load as means of visualizing correlation between them.
+Next, we built scatter plots of the independent continous variables against the load as means of visualizing correlation between them.
 Figure \ref{fig:CH_load_vs_temperature_2021} plots load against temperatures while figure \ref{fig:CH_load_vs_wind_speed_2021} depicts load against wind speed.
-From these plots we can clearly see there exist a relationship that ties wind speeds and temperatures to load.
+From these plots we can clearly see there exists a relationship that ties wind speeds and temperatures to load.
 
@@
 Such visualisation are reported in figure \ref{fig:CH_day_of_week_boxplot_2021} and figure \ref{fig:CH_is_holiday_boxplot_2021}.
-We can see that weekdays and workdays involve a higher load and posses a wider range.
+We can see that weekdays and workdays involve a higher load and possess a wider range.
 

### 2024-05-25 thesis/input_files/implementation.tex
22 May 2024
May 25th 2024

@@
-This section is inteneded to explain and aid for reproducibilty studies. Hereafter, the specific libraries used and the custom implementatios are thoroughly documented.
+This section is intended to explain and aid for reproducibility studies. Hereafter, the specific libraries used and the custom implementatios are thoroughly documented.
 \\
-For the list of python packages needed see the requirement.txt file on \url{https://github.com/luca-pernigo/ThesisKernelMethods}\label{github_repo}.
+For the list of Python packages needed see the requirement.txt file on \url{https://github.com/luca-pernigo/ThesisKernelMethods}\label{github_repo}.
 % - indicate computer specifics
@@
 Tbats implementation is available at \url{https://github.com/intive-DataScience/tbatsx}.
-In our application we specified as hyperparameters the lenght of seasons; 24 for the daily seasonality and 168 for the weekly seasonality.
+In our application we specified as hyperparameters the length of seasons; 24 for the daily seasonality and 168 for the weekly seasonality.
 
 \subsection{Prophet}
-The \href{https://facebook.github.io/prophet/docs/quick_start.html}{prophet} model has been applied by employing the python API provided by Meta.
+The \href{https://facebook.github.io/prophet/docs/quick_start.html}{prophet} model has been applied by employing the Python API provided by Meta.
 
@@
 \subsection{Support vector regression}
-The object SVR of the \href{https://scikit-learn.org/stable/}{sklearn} module svm has been used by specifying the liner kernel.
+The object SVR of the \href{https://scikit-learn.org/stable/}{sklearn} module svm has been used by specifying the linear kernel.
 
@@
 \subsection{Kernel quantile regression}
-Kernel quantile regression had no previously implemented python open source library, thus the need of implementing our own version.
+Kernel quantile regression had no previously implemented Python open source library, thus the need of implementing our own version.
 \\
@@
 In the fit method, we set up and solve the convex optimisation problem through the interior point method.
-This algorithm is taken from the cvxopt library, see its offical manual \cite{vandenberghe2010cvxopt} for a reference.
+This algorithm is taken from the cvxopt library, see its official manual \citeW{vandenberghe2010cvxopt} for a reference.
 When using this library, it is important to keep two things in mind.
-First this library assumes the quadratic term of the optimisation problem to be multiplied by the 0.5 factor, thus we just have to provide the Q matrix with no 0.5 in front.
+First this library assumes the quadratic term of the optimisation problem to be multiplied by the 0.5 factor, thus we just have to provide the $Q$ matrix with no 0.5 in front.
 Secondly, in order to specify multiple inequalities we have to stack them and provide them as a unique matrix.
@@
 This estimator is compatible with useful scikit-learn in built methods like gridsearch, crossvalidation and scoring rules.
-Moreover, this code is compatible with sklearn in built kernel functions \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.metrics.pairwise}{sklearn.metrics.pairwise} and \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.gaussian_process.kernels}{sklearn.gaussian\_process.kernels}, functions supported by our kernel quantile regression include: gaussian rbf, absolute laplacian, matern 3/2, matern 5/2, linear, cosine, sigmoid, periodic, polynomial and custom composition of kernels.
+Moreover, this code is compatible with sklearn in built kernel functions \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.metrics.pairwise}{sklearn.metrics.pairwise} and \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.gaussian_process.kernels}{sklearn.gaussian\_process.kernels}, functions supported by our kernel quantile regression include: Gaussian RBF, absolute Laplacian, Matern 3/2, Matern 5/2, Linear, Cosine, Sigmoid, Periodic, Polynomial and custom composition of kernels.
 % NOTICE, these kernel implementations differ in the form of passed parameters in the former we pass in the numerator gamma while in the latter we pass l, which stands for length scale in the denominator.

### 2024-05-25 thesis/input_files/literaturereview.tex
22 May 2024
May 25th 2024

@@
 This survey has been carried out by using the Scopus citation database.
-For details on the specific queries entered in Scopus, refer to Subsection 5 of the Appendix \ref{src_code}.
+For details on the specific queries entered in Scopus, refer to Appendix \ref{src_code}.
 \\
@@
 \\
-The next question is to compare the state of point versus probabilistic forecasting, this is visualized in Figure \ref{fig:point_vs_prob}.
-What can be concluded is that probabilistic is less developed than point forecasting.
+The next question is to compare the state of point-versus probabilistic forecasting, this is visualized in Figure \ref{fig:point_vs_prob}.
+What can be concluded is that probabilistic-is less developed than point forecasting.
 To our mind this is due to the complexity of probabilistic forecasts.
@@
 \\
-The EF literature is dominated by statistical and computational intelligence (CI) methods as can be seen from Figure \ref{fig:cs_stat_both}, with CI methods being slightly preferred. 
+The EF literature is dominated by statistical-and computational intelligence (CI) methods as can be seen from Figure \ref{fig:cs_stat_both}, with CI methods being slightly preferred. 
 \begin{figure}
@@
 EF is a heterogenous field of research, its researchers come from a wide array of backgrounds, with electrical engineers and statisticians making up the top contributors; their different educational training may explain why the split between statistical and computational intelligence methods is so marked.
-Figure \ref{fig:subject_area} depitcs the EF publications by subject area.
-What can be concluded, is that the bulk of publication come from engineering, computer science, mathematics and econometrics.
+Figure \ref{fig:subject_area} depicts the EF publications by subject area.
+What can be concluded is that the bulk of publication come from engineering, computer science, mathematics and econometrics.
 \begin{figure}
@@
 The IEEE Power and Energy Society provides also insightful lecture notes on probabilistic energy forecasting methodologies,
-implementations, and applications \cite{gm_22}.
+implementations, and applications \citeW{gm_22}.
 Marcjasz et al. \cite{probablistic_electricity_forecast2} uses distributional neural networks to create probabilistic forecasts for the day ahead electricity prices in the German market.

### 2024-05-25 thesis/input_files/probabilistic.tex
22 May 2024
May 25th 2024

@@
 Our goal is selecting the best features and best split point such that to achieve the best generalizing fit.
-For a visualisation consider the pictures \ref{fig:elements_statistical_learning1} and \ref{fig:elements_statistical_learning2} taken from \cite{hastie2009elements}, where the decision tree algorithm is visualised for a regression problem with two independent variables $X_2$ and $X_1$.
+For a visualisation consider the pictures \ref{fig:elements_statistical_learning1} and \ref{fig:elements_statistical_learning2}, where the decision tree algorithm is visualised for a regression problem with two independent variables $X_2$ and $X_1$.
 \begin{figure}
     \includegraphics[width=0.5\textwidth]{images/elsii1.png}
-    \caption{two-dimensional feature space partitioned by recursive binary splitting \cite{hastie2009elements}}
+    \caption{Two-dimensional feature space partitioned by recursive binary splitting \cite{hastie2009elements}}
     \label{fig:elements_statistical_learning1}
@@
 \includegraphics[width=1\textwidth]{images/elsii2.png}
-\caption{partition tree and regression model \cite{hastie2009elements}}
+\caption{Partition tree and regression model \cite{hastie2009elements}}
 \label{fig:elements_statistical_learning2}
@@
 Finding the best split in terms of minimum sum of squares is computationally infeasible in practice. Therefore we approximate a solution by approaching the problem in a greedy fashion.
-Let $j$ denote the splitting variable and $s$ be the the split point we define the two half planes
+Let $j$ denote the splitting variable and $s$ be the split point we define the two half planes
 \begin{equation}
@@
 See figure \ref{fig:decision_tree} for an example where decision trees with different hyperparameters are fitted to a sine wave plus noise.
-The most popular among these model is the CART \cite{breiman2017classification} tree, its name comes from the fact that it can handle both classification and regression problems.
+The most popular among these models is the CART \cite{breiman2017classification} tree, its name comes from the fact that it can handle both classification and regression problems.
 \begin{figure}
     \includegraphics[width=\textwidth]{images/decision_tree.png}
-    \caption{decision tree regression}
+    \caption{Decision tree regression}
     \label{fig:decision_tree}
@@
 \subsection{Bootstrap}
-The idea behind boostrapping is to randomly sample from the training set with replacement B times and then fitting B models to each of the "artificial" datasets.
-Boostrapping can serve different tasks; we can use it to assess the accuracy of a parameter %estimate
+The idea behind bootstrapping is to randomly sample from the training set with replacement $B$ times and then fitting $B$ models to each of the "artificial" datasets.
+Bootstrapping can serve different tasks; we can use it to assess the accuracy of a parameter %estimate
 or of a prediction but also to improve their estimates.
@@
 \subsection{Bagging}
-Bagging stands for boostrap aggregation. 
+Bagging stands for bootstrap aggregation. 
 The bagging estimate is defined by $\mathbb{E}_P[\hat{f}^*]$ where $P$ is the empirical distribution putting equal probability on each data point of the training set.
@@
 Decision trees are characterised by high variance and low bias, thus, they can benefit extremely from bagging.
-Furthermore, every decision tree generated through bagging will be identically distributed (i.d.), thus the expectation of an average of B trees is probabilistically equivalent to the expectation of any such trees.
+Furthermore, every decision tree generated through bagging will be identically distributed (i.d.), thus the expectation of an average of $B$ trees is probabilistically equivalent to the expectation of any such tree.
 As a consequence, the bias will stay fixed since the bias of the bagged estimator is the same as that of each individual tree.
 \\
-Consider positively correlated i.d. random variables, then the variance of their average is 
+Consider positively correlated i.d.\ random variables, then the variance of their average is 
 \begin{equation}
@@
 \end{equation}
-The second term disappears as B increases, while the first term depends heavily on the correlation between bagged trees.
-Random forest consists in reducing the correlation between the trees by randomly selecting m of the p features as candidates for splitting; m tipically takes a value in the order of $\sqrt{p}$ or even 1 with default value $m=\left\lfloor \frac{p}{3} \right\rfloor$, while a good minimum node size is around five.
+The second term disappears as $B$ increases, while the first term depends heavily on the correlation between bagged trees.
+Random forest consists in reducing the correlation between the trees by randomly selecting $m$ of the $p$ features as candidates for splitting; $m$ tipically takes a value in the order of $\sqrt{p}$ or even 1 with default value $m=\left\lfloor \frac{p}{3} \right\rfloor$, while a good minimum node size is around five.
 \\
-Letting $T(x;\Theta_b)$ be the bth bagged tree where $\Theta_b$ denotes the randomness characterizing its splits, cutpoints and terminal node values, we have that the random forest regressor is given by
+Letting $T(x;\Theta_b)$ be the $b^{th}$ bagged tree where $\Theta_b$ denotes the randomness characterising its splits, cutpoints and terminal node values, we have that the random forest regressor is given by
 \begin{equation}
@@
 \end{equation}
-For a simple visualisation compare figure \ref{fig:random_forest} with \ref{fig:decision_tree}
+For a simple visualisation compare figure \ref{fig:random_forest} with \ref{fig:decision_tree}.
 \begin{figure}
     \includegraphics[width=\textwidth]{images/random_forest.png}
-    \caption{random forest regression}
+    \caption{Random forest regression}
     \label{fig:random_forest}
@@
 \subsection{Quantile forest}
-The key observation here is noting that random forests approximates the conditional mean $\mathbb{E}(Y|X=x)$ by a weighted average over the observed $y$.
+The key observation here is noting that random forest approximates the conditional expected value $\mathbb{E}(Y|X=x)$ by a weighted average over the observed $y$.
 % expected value is the weighted averaged of realizations with weights their respective probabilities.
@@
 \end{equation}
-By swapping $F(y|X=x)$ with $\hat{F}(y|X=x)$ in the defintion of conditional quantiles we obtain their respective random forest estimator
+By swapping $F(y|X=x)$ with $\hat{F}(y|X=x)$ in the definition of conditional quantiles we obtain their respective random forest estimator
 \begin{equation}
@@
 where $b(x;\gamma)$ are the basis functions while $\beta_m$ are the coefficients of the expansion.
-Boosted models are fitted by minimizing a loss function $L$ over the training data
+Boosted models are fitted by minimising a loss function $L$ over the training data
 \begin{equation}
@@
 However, such problem is highly intensive in terms of computation.
-Therefore, what is done in the literature is approximating its solution by iteratively adding new basis function to the current expansion.
+Therefore, what is done in the literature is approximating its solution by iteratively adding new basis functions to the current expansion.
 That is, we construct $f_m$ by solving for the optimal basis function and coefficients to add to $f_{m-1}$.
@@
 \end{equation}
-Remember, $\Theta_m$ referes to the parameters of the mth tree, $\Theta_m=\{R_{jm}, \gamma_{jm}\}_1^{J_m}$
+Remember, $\Theta_m$ refers to the parameters of the $m^{th}$ tree, $\Theta_m=\{R_{jm}, \gamma_{jm}\}_1^{J_m}$
 \subsection{Gradient boosting}
@@
 \begin{equation}
-    \hat{\Theta}_m=\argmin{\Theta_m}\sum\limits_{i=1}^{N}\left(-g_{im}-T(x_i;\Theta)\right)^2
+    \hat{\Theta}_m=\argmin{\Theta_m}\sum\limits_{i=1}^{N}\left(-g_{im}-T(x_i;\Theta_m)\right)^2
 \end{equation}
 where $g_{im}$ is the gradient of $L(f)=\sum\limits_{i=1}^n L(y_i, f(x_i))$ evaluated at $f_{m-1}$.
-Put simply, we are fitting the m-th tree to the negative of the gradient values of f through least squares.
+Put simply, we are fitting the $m^{th}$ tree to the negative of the gradient values of $f$ through least squares.
 In order to solve our quantile regression tasks through gradient boosting, all we need to do is specifying the pinball loss as our criterion to guide the minimisation algorithm.
-See figure \ref{fig:gradient_boosting} for a visualisation
+See figure \ref{fig:gradient_boosting} for a visualisation.
 \begin{figure}
     \includegraphics[width=\textwidth]{images/gradient_boosting.png}
-    \caption{gradient boosting regression}
+    \caption{Gradient boosting regression}
     \label{fig:gradient_boosting}
@@
 The idea of quantile regression has been extended to kernel methods by Takeuchi et al. \cite{takeuchi2006nonparametric}.
-There, they minimize a risk functional plus regularizer defined as follows.
+There, they minimise a risk functional plus regulariser defined as follows.
 \begin{equation}\label{eq:kqr_min1}
@@
 % Moreover, note that the RKHS norm is defined as follows $\|f\|_{\mathcal{H}}=\inf\{\|w\|_{\mathcal{F}}:w\in \mathcal{F},f(x)=\langle w,\varphi (x)\rangle _{\mathcal{F}},\forall x\in X\}$, where $\mathcal{F}$ is the feature space.
- Doing so we obtain a minimization problem equivalent to minimizing equation \ref{eq:kqr_min1}.
+ Doing so we obtain a minimisation problem equivalent to minimising equation \ref{eq:kqr_min1}.
 \begin{equation}\label{eq:kqr_min2}
@@
     \end{equation}
-Consider now, the lagrangian $\mathcal{L}$ associated to \ref{eq:kqr_min4}
+Consider now, the Lagrangian $\mathcal{L}$ associated to \ref{eq:kqr_min4}
 \begin{equation}\label{eq:kqr_min5}
@@
 \begin{definition}
-    The dual function associated to the lagrangian $\mathcal{L}(x,\lambda, \nu)$ is given by $g(\lambda, \nu)=\underset{x}\inf \ \mathcal{L}(x,\lambda, \nu)$
+    The dual function associated to the Lagrangian $\mathcal{L}(x,\lambda, \nu)$ is given by $g(\lambda, \nu)=\underset{x}\inf \ \mathcal{L}(x,\lambda, \nu)$
 \end{definition}
-where $\lambda$ is called the lagrange multiplier of the optimization problem. Such dual formulation has an useful property, which is \begin{equation}\label{weak_duality}
+where $\lambda$ is called the Lagrange multiplier of the optimisation problem. Such dual formulation has an useful property, which is \begin{equation}\label{weak_duality}
     g(\lambda, \nu)\leq p^*
 \end{equation}
-where $p^*$ is the optimal value of your optimization problem. In other words $g(\lambda, \nu)$ is a lower bound for the optimal $p^*$.
-Consider now a simple lagrangian $\mathcal{L}(x,\lambda, \nu)=f(x)+\sum\limits_{i=1}^n \lambda_i r_i(x) +\sum\limits_{i=1}^{n} \nu_i h_i(x)$, where $r_i(x)$ are inequality constraints while $h_i(x)$ are equality constraints of the problem.
-Then it can be noted that, the lower bound on $p^*$ is non trivial only when the lagrange multiplier $\lambda \succeq 0$.
+where $p^*$ is the optimal value of your optimisation problem. In other words $g(\lambda, \nu)$ is a lower bound for the optimal $p^*$.
+Consider now a simple Lagrangian $\mathcal{L}(x,\lambda, \nu)=f(x)+\sum\limits_{i=1}^n \lambda_i r_i(x) +\sum\limits_{i=1}^{n} \nu_i h_i(x)$, where $r_i(x)$ are inequality constraints while $h_i(x)$ are equality constraints of the problem.
+Then it can be noted that, the lower bound on $p^*$ is non trivial only when the Lagrange multiplier $\lambda \succeq 0$.
 % when lagrange multiplier is negative, we have that g > p
-Therefore, the idea is that by maximizing the dual function subject to the constraint $\lambda \succeq 0$, we can obtain an approximate or perfect solution to the primal problem.
+Therefore, the idea is that by maximising the dual function subject to the constraint $\lambda \succeq 0$, we can obtain an approximate or perfect solution to the primal problem.
 \\
-To explain why we may or not be able to attain the best solution to the primal problem by maximizing its dual, we have to introduce the concept of duality.
+To explain why we may or not be able to attain the best solution to the primal problem by maximising its dual, we have to introduce the concept of duality.
 \\
-We use $d^*$ to denote the optimal value of the lagrange dual problem; we can think of it as the best lower bound on $p^*$. 
+We use $d^*$ to denote the optimal value of the Lagrange dual problem; we can think of it as the best lower bound on $p^*$. 
 \\
@@
 Moreover, note that the optimal duality gap is always nonnegative.
-We say that strong duality holds, when the optimal duality gap is zero; in other words, the lagrange dual bound is tight.
+We say that strong duality holds, when the optimal duality gap is zero; in other words, the Lagrange dual bound is tight.
 \\
-Constraint qualifications are condtions under which strong duality holds; one of the most popular is Slater's condition.
+Constraint qualifications are conditions under which strong duality holds; one of the most popular is Slater's condition.
 \begin{proposition}
@@
 \begin{theorem}
-    If Slater's condtion holds and the problem is convex then strong duality holds.
+    If Slater's condition holds and the problem is convex then strong duality holds.
 \end{theorem}
-We can now check that our optimization problem posseses strong duality by checking Slater's condition.
+We can now check that our optimisation problem possesses strong duality by checking Slater's condition.
 \\
 In our case we don't have any equality constraint, so we do not have to worry about the $h_i(x)=0$ term in \ref{slater_condition}.
-All we have to check is the convexity of our problem and that there exist an $x$ such that $r_i(x)<0$.
+All we have to check is the convexity of our problem and that there exists an $x$ such that $r_i(x)<0$.
 For convexity, a sufficient condition is the positive definiteness of $Q$ in the quadratic programming problem 
@@
 \end{equation}
-Hence, we conclude that our problem satisfies Slater's condition. Therefore the solution of the dual and primal problem are equivalent.
+Hence, we conclude that our problem satisfies Slater's condition. Therefore, the solution of the dual and primal problems are equivalent.
 \\
@@
 \end{equation}
-Setting its derivates to zero
+Setting its derivatives to zero
 \begin{equation}\label{eq:lagrange_derivatives}
@@
 \end{equation}
-As pointed out previously, the lower bound resulting from the dual formulation is non trivial only when the lagrange multipliers $\lambda$ are $\succeq 0$.
+As pointed out previously, the lower bound resulting from the dual formulation is non trivial only when the Lagrange multipliers $\lambda$ are $\succeq 0$.
 Looking at the last two equations of the system \ref{eq:lagrange_derivatives}, this implies the following two constraints $\lambda \in [0, Cq\mathbb{1}]$ and $\lambda^*\in [0, C(1-q)\mathbb{1}]$.
@@
 \end{equation}
-Using the gradient of the lagrangian of equation \ref{eq:lagrange_derivatives} we end up with
+Using the gradient of the Lagrangian of equation \ref{eq:lagrange_derivatives} we end up with
 \begin{equation}
@@
 
-Therefore in order to retrieve $b$ we simply have to choose an index $i$ such that  $\alpha_i \not \in \{C(q-1), Cq\}$
+Therefore, in order to retrieve $b$ we simply have to choose an index $i$ such that  $\alpha_i \not \in \{C(q-1), Cq\}$
 and let
@@
 \\
-The dataset used is the Melbourne daily maximum temperatures \cite{hdrcde}.
+The dataset used is the Melbourne daily maximum temperatures \citeW{hdrcde}.
 It contains the daily maximum temperatures in Melbourne, Australia, from 1981-1990, excluding leap days, see figure \ref{fig:melbourne_temperature_data}.
@@
 \end{table}
-As already pointed out, the quantile regression with q=0.5 corresponds to the standard regression problem, hence we can compare the proposed methods also in terms of the mean absolute error.
+As already pointed out, the quantile regression with $q=0.5$ corresponds to the standard regression problem, hence we can compare the proposed methods also in terms of the mean absolute error.
 \begin{table}[!h]

### 2024-05-25 thesis/input_files/qr.tex
May 25th 2024

@@
 \end{definition}
-Alternatively, in order  to ease the posing of the quantile regression problem, we can formulate quantiles as the solutions to the following optimization problems.
+Alternatively, in order  to ease the posing of the quantile regression problem, we can formulate quantiles as the solutions to the following optimisation problems.
 \\
 For any $0<q<1$ consider the pinball loss function from section \ref{pinball}, $\rho_q(u)=u(q-\mathbb{I}_{\{u<0\}})$. 
-Such loss is minimized by the quantile $Q(q)$.
-Thus,we can estimate quantiles by minimizing the expectation of $\rho_q(y-g(x,\beta))$ with respect to the parameter $\beta$.
+Such loss is minimized by the quantile $Q_q$.
+ we can estimate quantiles by minimizing the expectation of $\rho_q(y-g(x,\beta))$ with respect to the parameter $\beta$.
 \\

### 2024-05-25 thesis/input_files/data_exploratory.tex
May 25 2024

@@
 
-This section covers the datasets used in our experiments, in addition, attributes and features will be thoroughly described.
+This section covers the datasets used in our experiments. In addition, attributes and features will be thoroughly described.
 Next, we will explain the extract transform load pipeline (ETL) that we set up in order to ease the workflow of our studies.
@@
 \section{GEFCom2014}
-The first class of datasets we identified were the ones from the GEFCom2014 competition. The data is freely hosted on Dr.\ Hong blog \cite{hong2016probabilistic}.
+The first class of datasets we identified are the ones from the GEFCom2014 competition. The data is freely hosted on Dr.\ Hong's blog \cite{hong2016probabilistic}.
 The main reason is that, these datasets are considered an excellent test case for comparing predictive models between the EF community.
@@
 The GEFCom2014 folder contains a zip file for each of the four tracks.
-This thesis work is focused on providing forecasts for the load and price quantities.
+We provide forecasts for the load and price quantities.
 \\
@@
 The data can be found in the GEFCom2014-P\_V2 zip file alongside a set of instructions and the benchmark forecasts.
-This data consisted of time series for the locational marginal price, for the zonal load and for the system load.
-The data covers the time interval ranging from the $1^{st}$ of January 2011 to the $15^{th}$ of June 2013, see figure \ref{fig:price_track_fig1}.\
+This data consists of time series for the locational marginal price, for the zonal load and for the system load.
+The data covers the time interval ranging from the $1^{st}$ of January 2011 to the $15^{th}$ of June 2013, see Figure \ref{fig:price_track_fig1}.\
 Additionally, as the competition went on, the real observed data of the previous tasks were made available.
@@
 \subsubsection{EDA}
-In exploring this data we started off by plotting the time series of the zonal prices alongside with the total load and zonal load, see figure \ref{fig:price_track_fig1}.
-After that, we plotted the scatter plot for inspecting the correlation between the predictors and the dependent variables, figure \ref{fig:gefcom_zonal_price_vs_zonal_load} and figure \ref{fig:gefcom_zonal_price_vs_total_load}.
-From these visualisations, we can see their goodness for predicting zonal price since there exists a clear correlation between the candidate predictors and the target variables.
+In exploring this data we started off by plotting the time series of the zonal prices alongside with the total load and zonal load, see Figure \ref{fig:price_track_fig1}.
+After that, we plotted the scatter plot for inspecting the correlation between the predictors and the dependent variables, Figure \ref{fig:gefcom_zonal_price_vs_zonal_load} and Figure \ref{fig:gefcom_zonal_price_vs_total_load}.
+From these visualisations, we can see their goodness for predicting zonal prices since there exists a clear correlation between the candidate predictors and the target variables.
 
@@
 \\
-In order to compare our model performance with the winning entries of the GEFCom2014 load competition, we will refer to the Provisional\_Leaderboard\_V2.csv file contained also in the GEFCom2014 Data directory.
+In order to compare our model performance with the winning entries of the GEFCom2014 load competition, we will refer to the Provisional\_Leaderboard\_\-V2.csv file contained also in the GEFCom2014 Data directory.
 % Notice a quick premise, within this xsls file the authors opted for various scores to compare the entries performance. This choice was motivated by the need 
@@
 \\
-Notice, the pinball scores for the load track are stored inside the subtab L-score-0/L-score-2 of Provisional\_Leaderboard\_V2.
+The pinball scores for the load track are stored inside the subtab L-score-0/L-score-2 of Provisional\_Leaderboard\_V2.
 \subsubsection{EDA}

### 2024-05-25 thesis/input_files/energymarket.tex
May 25 2024

@@
 Firstly, it summarizes several concepts for the newcomers to the field of electricity markets.
-Secondly, it discusses trends and challenges undergoing in the power sector.
-Such developments motivate the need for robust and efficient forecasting techniques in the context of electricity markets.
+Secondly, it discusses trends and challenges in the power sector.
+Such developments motivate the need for robust and efficient forecasting techniques.
 
@@
 Finally, the voltage is reduced and the energy is moved into the distribution network.
-In a nutshell, transmission lines carry electric power from stations to substations while distribution lines carry electricity from substations to load points such as businesses, industries and homes, see figure \ref{electricity_network}.
+Transmission lines carry electric power from stations to substations while distribution lines carry electricity from substations to load points such as businesses, industries and homes, see Figure \ref{electricity_network}.
 
@@
 In particular, it was characterized by a high vertical integration and by very little market competition.
-Better technologies in both transmission, generation and distribution are the reasons for such liberal shift.
+Better technologies in transmission, generation and distribution are the reasons for such liberal shift.
 The rationale is simple, a competitive market promotes efficiency and stimulates technical innovation more than a monopoly does.
 \\
-During the nineties various wholesale electricity markets were established; for example in Chile, Great Britain, the Scandinivian block, Australia, New Zealand, the US and Canada.
-Consequently, new participants, each with a specific task, entered the power market scenario.
+During the 1990s, various wholesale electricity markets were established; for example in Chile, Great Britain, the Scandinivian block, Australia, New Zealand, the US and Canada.
+New participants, each with a specific task, entered the power market scenario.
 \begin{itemize}
@@
 \end{itemize}
-Notice, these functions are not mutually exclusive; that is an entity can provide one or more of these services.
+These functions are not mutually exclusive. That is, an entity can provide one or more of these services.
 
@@
 In power pools, generators bid the prices at which they are willing to produce at different volumes.
-Then, the market clearing price (MCP) is determined by intersecting the aggregated supply curve and the estimated demand, left panel figure \ref{fig:pool_vs_echange}.
+Then, the market clearing price (MCP) is determined by intersecting the aggregated supply curve and the estimated demand, left panel Figure \ref{fig:pool_vs_echange}.
 Power pools are created on public initiatives of governments.
@@
 This is the model followed by most of European countries.
-The MCP in power exchanges is determined by the intersection of the aggregate supply curve and the aggregate demand curve, right panel figure \ref{fig:pool_vs_echange}.
+The MCP in power exchanges is determined by the intersection of the aggregate supply curve and the aggregate demand curve, right panel Figure \ref{fig:pool_vs_echange}.
 \\
@@
 \subsection{Price peaks}
-Electricity markets are characterized by spikes in their spot prices; when this happens, the system price jumps abruptly and then drops back within a very short period.
-This spikeness follows from the non storability of electricity; that is, electricity has to be consumed as it is produced.
+Electricity markets are characterized by spikes in their spot prices. When this happens, the system price jumps abruptly and then drops back within a very short period.
+This spikeness follows from the non storability of electricity, that is, electricity has to be consumed as it is produced.
 Therefore, extreme load fluctuations combined with generation outages or transmission failures can result in price spikes.
@@
 The causes of negative prices are inflexible power generation plants and low demand.
-With inflexibility, it is meant the fact that power sources cannot be switched off and restarted quickly and efficiently.
+With inflexibility, is meant the fact that power sources cannot be switched off and restarted quickly and efficiently.
 Thus, producers are faced with the decision of either stopping and then restarting their power plant or selling their energy for a negative price; that is they pay consumers for consuming their energy.
@@
 Grid configuration and physical limits on electricity lines may lead to congestion.
-Zonal market clearing pricing (ZMCP) and local marginal price (LMP) are two types of pricing schemes utilized to handle it.
-The former is adopted in European contries while the latter is used in the United States, figure \ref{fig:eu_zonal} and figure \ref{fig:us_local}.
+Zonal market clearing pricing (ZMCP) and local marginal price (LMP) are two types of pricing schemes utilized to address this.
+The former is adopted in European contries while the latter is used in the United States, Figure \ref{fig:eu_zonal} and Figure \ref{fig:us_local}.
 With zonal market clearing pricing, prices may differ between zones but are the same within the same area.
-On the other hand, local marginal price is made up by summing the transmission congestion cost, generation marginal cost and the cost of marginal losses at different buses; buses is where a electricity line or several lines are connected.
+On the other hand, local marginal price is made up by summing the transmission congestion cost, generation marginal cost and the cost of marginal losses at different buses.
+Buses is where a electricity line or several lines are connected.
 \begin{figure}[!h]
     \includegraphics[width=\textwidth]{images/eu_zonal.png}
-    \caption[Zonal pricing EU]{Zonal pricing EU \citeW{eu_zonal}}
+    \caption[Zonal pricing in the EU]{Zonal pricing in the EU \citeW{eu_zonal}}
     \label{fig:eu_zonal}
@@
     \includegraphics[width=\textwidth]{images/us_local.png}
-    \caption[Local pricing US]{Local pricing US \citeW{us_local}}
+    \caption[Local pricing for the US]{Local pricing for the US \citeW{us_local}}
     \label{fig:us_local}
@@
 Deployment of smart grids and the digitalisation of the electricity network are intended to increase the energy efficiency of our networks.
-The reason is that, smart metering and the electricity network digitalisation  will enable system operators to better monitor the network, plan their investments and manage infrastructure. 
-All of this motivates for the need of efficient and robust tools to analyze the vast data resulting from the digitalisation of the power sector, see \cite{declean,andreadou2016telecommunication}.
+The reason is that smart metering and the electricity network digitalisation  will enable system operators to better monitor the network, plan their investments and manage infrastructure. 
+All of this motivates the need for efficient and robust tools to analyze the vast data resulting from the digitalisation of the power sector, see \cite{declean,andreadou2016telecommunication}.
 
@@
 % https://www.pnnl.gov/explainer-articles/renewable-integration
-A pressing challenge facing policymakers today is the energy transition towards cleaner energy, see figure \ref{fig:electricity_production_by_source} for an up to date breakdown on energy production by source.
+A pressing challenge facing policymakers today is the energy transition towards cleaner energy, see Figure \ref{fig:electricity_production_by_source} for an up to date breakdown on energy production by source.
 Such move is intended to mitigate the effects of climate change and reduce pollution at the same time.
@@
 Additionally, renewables such as wind and solar are characterized by seasonalities and depend heavily on weather conditions.
-Therefore, balancing the electricity network and mantaining stability will become harder for grid operators; they will need more flexibility in the grid and develop new approaches in order to achieve their goals.
+Therefore, balancing the electricity network and mantaining stability will become harder for grid operators.
+They will need more flexibility in the grid and develop new approaches in order to achieve their goals.
 

### 2024-05-25 thesis/input_files/literaturereview.tex
May 25 2024

@@
 In terms of forecasting horizons, we can group EF into four major categories: very short term forcasting (VSTF), short term load forecasting (STF), medium term forecasting (MTF) and long term forecasting (LTF).
-Consensus in the literature is to use one day, two weeks and three years respectilvely \cite{hong_phd} as cut off horizons; see figure \ref{fig:time} for a visualisation.
+Consensus in the literature is to use one day, two weeks and three years respectilvely \cite{hong_phd} as cut off horizons; see Figure \ref{fig:time} for a visualisation.
 \begin{figure}
@@
 Figure \ref{fig:subject_area} depicts the EF publications by subject area.
-What can be concluded is that the bulk of publication come from engineering, computer science, mathematics and econometrics.
+What can be concluded is that the bulk of publications come from engineering, computer science, mathematics and econometrics.
 \begin{figure}
@@
 \section{Electricity forecasting literature review}
-To get started a few review articles were collected in order to understand conventions, best practicies and terminology of the EF community.
+To get started, a few review articles were collected in order to understand conventions, best practicies and terminology of the EF community.
 Weron \cite{EPF_review} reviews the state of the art for electricity price forecasting.
-Besides analysing complexity of available solutions, strenghts and weaknesses it also stresses the need for objective comparative EPF studies.
+Besides analysing complexity of available solutions, strenghts and weaknesses, it also stresses the need for objective comparative EPF studies.
 Specifically, it advocates for studies using similar datasets, using the same error evaluation metric and statistical testing model's outperformance.
@@
 He et al.
-\cite{he2022cooperative} models  multistep wind speed probabilistic forecasting by mixing complementary ensemble empirical mode decomposition(CEEDMAN), least absolute shrinkage and selection operator (LASSO) and quantile regression (QR).
+\cite{he2022cooperative} models  multistep wind speed probabilistic forecasting by mixing complementary ensemble empirical mode decomposition (CEEDMAN), least absolute shrinkage and selection operator (LASSO) and quantile regression (QR).
 In this work, CEEMDAN is used to decompose the wind speed time series, LASSO compresses high dimensional features, QR is used for obtaining quantile forecasts, finally kernel density estimation (KDE) converts quantile forecasts into density estimates.
@@
 \cite{zhang2020probability} introduce a framework based on quantile regression and kernel density estimation in the context of short term wind forecasting.
-The proposed methods behave well compared to an autoregressive ARMA(1,1) model.
+The proposed methods behave well compared to an autoregressive ARMA model.
+% ARMA(p=1,q=1) 
 Haben et al. \cite{haben2018short} analyses a variety of techniques in terms of both probabilistic and point forecasting. Within this study, they focus on load forecasting at the low voltage level.
@@
 Fatema et al.
-\cite{fatema2023probabilistic} considers Gaussian process regression for point forecasting and prediction intervals prediction.
+\cite{fatema2023probabilistic} considers Gaussian process regression for point forecasting and prediction intervals estimation.
 Then, it inputs prediction intervals to kernel density estimation in order to estimate a probability distribution.
@@
 Every paper uses different datasets.
-Therefore, it is not possible to compare directly results from one paper to another without implementing the paper specific algorithms and then applying them to the respective dataset.
+Therefore, it is not possible to directly compare results from one paper to another without implementing the paper specific algorithms and then applying them to the respective dataset.
 An additional hurdle is that some datasets are not freely accessible.
 \\
-Thus, a good understanding of state of the art methods in EF is required to carry out a rigorous comparison between methods.
+Thus, a good understanding of the state of the art methods in EF is required to carry out a rigorous comparison between methods.
 That is why the following chapters are devoted to summarising the mathematical theory underlying such techniques.
@@
 With kernel methods we are able to apply linear methods with predictors in a high dimensional space, without having to explicitly evaluate the involved dot products of the features.
-Throughout this thesis work, we will address the performance of kernel methods in the context of EF.
+Throughout this thesis, we will address the performance of kernel methods in the context of EF.
 \\
@@
 Such term was first used by David Hilbert in his paper on integral equations \cite{hilbert} where he introduced the term "definite kernels".
-Following, Hilbert's and Schmidt's work \cite{schmidt} lead to the introduction of a new space, the Hilbert space.
+Following, Hilbert's and Schmidt's work, \cite{schmidt} lead to the introduction of a new space, the Hilbert space.
 In 1909, James Mercer improved Hilbert's work by proposing his theorem \cite{mercer}. This theorem underlies the power of kernel methods, that is the kernel trick.
 In 1938, Schoenberg \cite{schoenberg} developed the mathematical results that allow us to find the kernel associated to a specific feature space metric.
-In 1941, Kolmogorov \cite{kolmogorov} carried out stuides on representing kernel in linear spaces.
+In 1941, Kolmogorov \cite{kolmogorov} carried out stuides on representing kernels in linear spaces.
 In 1950, Aronszajn \cite{aronszajn} published the first work on reproducing kernel hilbert spaces (RKHS); developing the general method for representing kernels in linear spaces.
 In 1964, Aizerman \cite{aizerman} further improved the theory of RKHS.
-It was in the nineties that theory of kernel methods got popular, particularly in the field of machine learning.
-Kernels have been used in various different tasks such as SVM \cite{vapnik1} \cite{vapnik2}, Gaussian process classifiers \cite{williams}, spline methods \cite{wahba}, neural networks \cite{poggio} and principal component analysis \cite{pca_scholkopf}.
+It was in the 1990s that the theory of kernel methods got popular, particularly in the field of machine learning.
+Kernels have been used in various different tasks such as support vector machines (SVM) \cite{vapnik1} \cite{vapnik2}, Gaussian process classifiers \cite{williams}, spline methods \cite{wahba}, neural networks \cite{poggio} and principal component analysis \cite{pca_scholkopf}.
 Nevertheless kernel methods received very little attention in the specific setting of EF literature.
@@
 \cite{2022nystrom} proposes an efficient approximation procedure based on the Nyström method \cite{nystrom}, providing also an upper bound for the approximation error.
-Smola et al.
+Another example of the applicability of kernel methods is \cite{supersamples}.
+In that paper, Smola et al.\ kernelize the herding algorithm introduced in \cite{welling2009herding,welling2009herding,Welling2010}.
+The original herding algorithm converts empirical moments into a sequence of pseudo-samples approximating the probability density function (PDF) of the observed random variable.
+Its weakness is that it can only take a finite number of features and this limits its expressive power.
+Smola et al.\ overcome this by translating the algorithm into the framework of kernels.
+
+% Smola et al.
 \cite{supersamples} presents kernel herding; basically, the authores used the kernel trick to extend the herding algorithm to continous spaces.
 The result is an infinite memory deterministic process that takes in a collection of samples and learns to approximate a probability density function (PDF).
-\\
+
 %QUESTO DA VEDERE

### 2024-05-25 thesis/input_files/metrics.tex
May 25 2024

@@
-Proper evaluation methods guide researchers in choosing the model that best fits their needs; thus, this chapter is dedicated to the most common evaluation metrics adopted by academics in the field of electricity forecasting.
+Proper evaluation methods guide researchers in choosing the model that best fits their needs.
+This chapter is dedicated to the most common evaluation metrics adopted by academics in the field of electricity forecasting.
 Error metrics and measures vary depending on whether we are concerned with point or probabilistic forecasts.
@@
 \begin{definition}
-    $\mathrm{MAE}(y,\hat{y})=\frac{1}{h}\| y- \hat{y}\|_{1}=\frac{1}{h}\sum\limits_{k=1}^{h}|y_{n+k}-\hat{y}_{n+k}|$
+    The mean absolute error is defined as
+    $$
+    \mathrm{MAE}(y,\hat{y})=\frac{1}{h}\| y- \hat{y}\|_{1}=\frac{1}{h}\sum\limits_{k=1}^{h}|y_{n+k}-\hat{y}_{n+k}|
+    $$
 \end{definition}
@@
 \begin{definition}
-    $\mathrm{RMSE}(y, \hat{y})=\frac{1}{\sqrt{h}}\|y-\hat{y}\|_{2}=\sqrt{\frac{\sum\limits_{k=1}^{h}(y_{n+k}- \hat{y}_{n+k})^2}{h}}$
+    The root mean squared error is defined as
+    $$
+    \mathrm{RMSE}(y, \hat{y})=\frac{1}{\sqrt{h}}\|y-\hat{y}\|_{2}=\sqrt{\frac{\sum\limits_{k=1}^{h}(y_{n+k}- \hat{y}_{n+k})^2}{h}}
+    $$
 \end{definition}
@@
 However, a drawback of such measures is that we cannot use them to compare accuracy between time series which have different magnitudes.
-For instance, a day ahead error of 1kWh is negligible when considering a daily demand of 100h while the same error is considerably big when daily demand is 2kWh.
-This consideration leads to relative accuracy scores, between those the mean absolute percentage error (MAPE) is by far the most popular.
+For instance, a day ahead error of 1kWh is negligible when considering a daily demand of 100kWh while the same error is considerably big when daily demand is 2kWh.
+This consideration leads to relative accuracy scores.
+Between those the mean absolute percentage error (MAPE) is by far the most popular.
 
@@
 \begin{definition}
-    $\mathrm{MAPE}(y,\hat{y})=\frac{100}{h}\sum\limits_{k=1}^{h}\frac{|y_{n+k}-\hat{y}_{n+k}|}{|y_{n+k}|}$
+    The mean absolute percentage error is defined as
+    $$
+    \mathrm{MAPE}(y,\hat{y})=\frac{100}{h}\sum\limits_{k=1}^{h}\frac{|y_{n+k}-\hat{y}_{n+k}|}{|y_{n+k}|}$$
 \end{definition}
@@
 \begin{definition}
-    $\mathrm{RMSPE}(y,\hat{y})=100\cdot\sqrt{\frac{1}{h}\sum\limits_{k=1}^{h} \left(\frac{|y_{n+k}-\hat{y}_{n+k}|}{|y_{n+k}|}\right)^2}$
+    The root mean squared percentage error is defined as
+    $$
+    \mathrm{RMSPE}(y,\hat{y})=100\cdot\sqrt{\frac{1}{h}\sum\limits_{k=1}^{h} \left(\frac{|y_{n+k}-\hat{y}_{n+k}|}{|y_{n+k}|}\right)^2}
+    $$
 \end{definition}
-Note, mean absolute percentage error and root mean squared percentage error may not be appropriate for series which have zero or very small values, for example, electricity demand at the household level; the result is a large score regardless of the absolute errors.
+Mean absolute percentage error and root mean squared percentage error may not be appropriate for series which have zero or very small values, for example, electricity demand at the household level.
+The result is a large score regardless of the absolute errors.
 Scaled errors constitute a robust family of scores.
@@
 \begin{definition}
-    $\mathrm{MASE}(y,\hat{y})=\frac{1}{h}\sum\limits_{k=1}^N\frac{|y_{n+k}-\hat{y}_{n+k}|}{\frac{1}{h-1}\sum\limits_{k=2}^{h}|y_{k}-y_{k-1}|}$
+    The mean absolute scaled error is defined as
+    $$
+    \mathrm{MASE}(y,\hat{y})=\frac{1}{h}\sum\limits_{k=1}^N\frac{|y_{n+k}-\hat{y}_{n+k}|}{\frac{1}{h-1}\sum\limits_{k=2}^{h}|y_{k}-y_{k-1}|}
+    $$
 \end{definition}
 In the denominator we have the error of the naïve/persistence model. 
-In this model, the current demand makes up the prediction for the next time step; that is $\hat{y}^{\mathrm{naive}}_{n+1}=y_{n}$.
+In this model, the current demand makes up the prediction for the next time step, that is $\hat{y}^{\mathrm{naive}}_{n+1}=y_{n}$.
 \section{Root mean squared scaled error}\label{rmsse}
 \begin{definition}
-    $\mathrm{RMSSE}(y,\hat{y})=\sqrt{\frac{1}{h}\sum\limits_{k=1}^N\left(\frac{|y_{n+k}-\hat{y}_{n+k}|}{\frac{1}{h-1}\sum\limits_{k=2}^{h}|y_{k}-y_{k-1}|}\right)^2}$
+    The root mean squared scaled error is defined as
+    $$
+    \mathrm{RMSSE}(y,\hat{y})=\sqrt{\frac{1}{h}\sum\limits_{k=1}^N\left(\frac{|y_{n+k}-\hat{y}_{n+k}|}{\frac{1}{h-1}\sum\limits_{k=2}^{h}|y_{k}-y_{k-1}|}\right)^2}
+    $$
 \end{definition}
@@
 \begin{definition}
-    $\mathrm{Pinball}(y_{t},\hat{y}_{t,q},q)=
+    The pinball loss is defined as
+    $$
+    \mathrm{Pinball}(y_{t},\hat{y}_{t,q},q)=
 \begin{cases}
@@
 q(\hat{y}_{t,q}-y_t) & y_t \leq \hat{y}_{t,q}
-\end{cases}$
+\end{cases}
+$$
 \end{definition}
-The pinball loss is an asymmetric function, it weights its score differently depending on the error sign and on the quantile considered, see figure \ref{fig:pinball}.
+The pinball loss is an asymmetric function, it weights its score differently depending on the error sign and on the quantile considered, see Figure \ref{fig:pinball}.
 \begin{figure}
@@
 \begin{definition}
-    $\mathrm{Winkler}(y_t, \hat{y}_t,q)=\begin{cases}
+    The Winkler score is defined as
+    $$
+    \mathrm{Winkler}(y_t, \hat{y}_t,q)=\begin{cases}
         \delta & l_{t}\leq y_{t}\leq u_{t}\\
@@
         \delta+2(l_{t}-u_{t})/\alpha & y_{t} \geq u_{t}
-    \end{cases}$
+    \end{cases}
+    $$
 \end{definition}
@@
 \begin{definition}\label{def_crps}
-    $\mathrm{CRPS}(y, \hat{F})=\int\limits_{-\infty}^{\infty}\left(\hat{F}(x)-\mathbb{I}_{\{x-y\}} \right)^2 dx$
+    The continous ranked probability score is defined as
+    $$
+    \mathrm{CRPS}(y, \hat{F})=\int\limits_{-\infty}^{\infty}\left(\hat{F}(x)-\mathbb{I}_{\{x-y\}} \right)^2 dx
+    $$
 \end{definition}
@@
 Where the indicator function is defined as 
-    $\mathbb{I}_{\{z\}}=
+    $$\mathbb{I}_{\{z\}}=
 \begin{cases}
-0 & z<0\\
-1 & z \geq 0
-\end{cases}$
+0, & z<0\\
+1, & z \geq 0
+\end{cases}$$
 \\
-For a visualisation see, \ref{fig:crps}. The grey area is what contributes toward the CRPS score.
+For a visualisation see Figure \ref{fig:crps}. The grey area is what contributes toward the CRPS score.
 The better the estimated cumulative density function is the smaller the total CRPS score will be.
@@
 It is worth noting, that the CRPS integral can be rewritten in terms of expectations.
-This makes its evaluation easier, since we know that the sample mean converges to the expectation by the law of large numbers.
-This was first pointed out by \cite{proper_scores}, where authors take advantage of lemma 2.2 of \cite{new_multi_test2} or equivalently identity 17 of \cite{new_multi_tes1}.
+This makes its evaluation easier, since we know that the sample mean converges to the expectation for sufficiently big sample sizes by the law of large numbers.
+This was first pointed out by \cite{proper_scores}, where the authors take advantage of Lemma 2.2 of \cite{new_multi_test2} or equivalently identity (17) of \cite{new_multi_tes1}.
 \\
@@
 Then
-   \begin{equation}
+   \begin{equation}\label{eq:crps_expectation}
     \mathbb{E}(|X_1-Y_1|)-\frac{1}{2}\mathbb{E}(|X_1-X_2|)-\frac{1}{2}\mathbb{E}(|Y_1-Y_2|)=\int\limits_{-\infty}^{\infty}\left(F(x)-G(x)\right)^2dx
@@
 \end{lemma}
-Notice that, in our case, equation \ref{def_crps}, the distribution $G$ of $Y_1$ and $Y_2$ is degenerate, with all probability mass on a single point $y$.
+Notice that, in our case, Equation \ref{eq:crps_expectation}, the distribution $G$ of $Y_1$ and $Y_2$ is degenerate, with all probability mass on a single point $y$.
 %  (x notation will need to be cleaned here). Since $G(t)=:$
-It follows that, the third addend in the summation is zero.
-That is because the fact of $Y_1$ and $Y_2$ both following distribution $G$ implies that $\mathbb{E}(|Y_1-Y_2|)$ corresponds to the difference of two equal constant numbers.
+It follows that the third addend in the summation is zero.
+That is because $Y_1$ and $Y_2$ both following distribution $G$ implies that $\mathbb{E}(|Y_1-Y_2|)$ corresponds to the difference of two equal constant numbers.
 \\
@@
 The probability integral transform (PIT) is a method to assess visually the quality of a probabilistic forecast.
-PIT is obtained by applying the predicted cumulative density function $\hat{F}$ to your data; if applying such CDF to the data results in a uniform distributed PIT, then $\hat{F}$ is a valid prediction.
-If not, $\hat{F}$ is not the suited CDF for the considered data.
+PIT is obtained by applying the predicted cumulative density function $\hat{F}$ to the data.
+If applying such CDF to the data results in a uniform distributed PIT, then $\hat{F}$ is a valid prediction.
+If not, $\hat{F}$ is not a good estimate of the CDF for the considered data.
 Figure \ref{fig:pit} provides an example, applying the true CDF results in a well calibrated PIT (left).

### 2024-05-25 thesis/input_files/paper1.tex
May 25 2024

@@
 \\
-$\langle \phi([2,4]),\phi([8,9]) \rangle_{\mathcal{H}}=  \left[ {\begin{array}{cccc}
+$$
+\begin{aligned}
+\langle \phi([2,4]),\phi([8,9]) \rangle_{\mathcal{H}}&=  \left[ {\begin{array}{cccc}
     4 & 16 &     \sqrt{2}\cdot 2 \cdot 4\\
@@
     \\
-  \end{array} } \right]\\=4\cdot 64 +16\cdot 81 +2 \cdot 2 \cdot 4 \cdot 8 \cdot 9=2704
-$
+  \end{array} } \right]\\ &=4\cdot 64 +16\cdot 81 +2 \cdot 2 \cdot 4 \cdot 8 \cdot 9=2704
+\end{aligned}
+$$
 \\
@@
 \begin{definition}
-    A sequence $\{v_n\}_{n=1}^{\infty}$ of elements of a normed space $\mathcal{V}$ is a Cauchy sequence if for every $\epsilon>0$, there exist $N=N(\epsilon) \in \mathbb{N}$ such that $\|v_n-v_m\|_{\mathcal{V}}<\epsilon \ \ \forall m,n\geq N$
+    A sequence $\{v_n\}_{n=1}^{\infty}$ of elements of a normed vector space $\mathcal{V}$ is a Cauchy sequence if for every $\epsilon>0$, there exist $N=N(\epsilon) \in \mathbb{N}$ such that $\|v_n-v_m\|_{\mathcal{V}}<\epsilon \ \ \forall m,n\geq N$
 \end{definition}
@@
 \begin{definition}
-    A complete metric space is a metric space in which every Cauchy sequence is convergent.
+    A complete normed space is a normed vector space in which every Cauchy sequence is convergent.
 \end{definition}
@@
     A Hilbert space is a vector space $\mathcal{H}$ with an inner product $\langle \cdot, \cdot \rangle_{\mathcal{H}}$ such that the norm defined by $\|f\|_{\mathcal{H}}=\sqrt{\langle f, f \rangle}_{\mathcal{H}}$
-turns $\mathcal{H}$ into a complete metric space.
+turns $\mathcal{H}$ into a complete normed vector space.
 \end{definition}
@@
 \begin{definition}
-    A function $k: \mathcal{X} \times \mathcal{X} \to \mathbb{R}$ is said to be finitely positive semidefinite if it is a symmetric function for which the matrices formed with it are positive semidefinite.
+    Considering a set of vectors $x_1, \cdots, x_m \in \mathcal{X} \subseteq \mathbb{R}^n$, the Gram matrix is defined as the $n\times n$ matrix whose entries are $\langle x_i, x_j \rangle_{\mathcal{X}}$.
 \end{definition}
 
-\begin{theorem}[Characterisation of kernels]    
-    A function $k:\mathcal{X} \times \mathcal{X} \to \mathbb{R}$ which is either continous or has finite domain can be decomposed as $k(x,y)=\langle \phi(x), \phi(y) \rangle_{\mathcal{H}}$ if and only if $k$ satisfy the positive semidefinite property.
+It follows from the definition that the Gram matrix is symmetric.
+In the kernel setting we evaluate the inner products in the feature space generated by the feature map $\phi$, therefore the Gram matrix will have entries $\langle \phi(x_i), \phi(x_j) \rangle_{\mathcal{H}}= k(x_i, x_j)$.
+Such matrix is also referred to as the kernel matrix $K$.
+
+\[
+K_{m\times m}=
+\begin{bmatrix}
+    k(x_1,x_1)       &  \dots & k(x_1,x_n) \\
+    \vdots       & \ddots & \vdots \\
+    k(x_n, x_1)       & \dots & k(x_n, x_n)
+\end{bmatrix}
+\]
+% All kernel methods take as input the kernel matrix, it is the central data type for all 
+% such techniques.
+
+\begin{proposition}
+    In addition to being symmetric, $K$ possesses also the property of being positive semidefinite.    
+\end{proposition}
+
+\begin{proof}
+    $$
+    \begin{aligned}
+        z^\intercal K z =& \sum_{i,j=1}^n z_i z_j k(x_i,x_j)
+        \\
+        =& \sum_{i,j=1}^n z_i z_j \langle \phi(x_i), \phi(x_j) \rangle_{\mathcal{H}}
+        \\
+        =& \left\langle \sum_{i=1}^n z_i \phi(x_i), \sum_{j=1}^n z_j \phi(x_j)\right\rangle_{\mathcal{H}}
+        \\
+        =& \left\|\sum_{i=1}^n z_i \phi(x_i) \right\|_{\mathcal{H}}^2 \geq 0
+    \end{aligned}
+    $$
+\end{proof}
+We know that positive semidefinite matrices form a cone in the vector space of $n\times n$ matrices.
+Because of this, we have that the set of such matrices is convex, this implies that the theory of convex analysis can be used when optimising over them.
+
+\begin{definition}
+    A function $k: \mathcal{X} \times \mathcal{X} \to \mathbb{R}$ is said to be finitely positive semidefinite if it is a symmetric function for which the matrices formed by restriction to any finite subset of the space $\mathcal{X}$ are positive semidefinite.
+\end{definition}
+
+\begin{theorem}[Characterisation of kernels \cite{shawe2004kernel}]    
+    A function $k:\mathcal{X} \times \mathcal{X} \to \mathbb{R}$ which is either continous or has finite domain can be decomposed as $k(x,y)=\langle \phi(x), \phi(y) \rangle_{\mathcal{H}}$ if and only if $k$ is positive semidefinite.
 \end{theorem}
 
-That is the function $k$ can be written as the inner product in the Hilbert space $\mathcal{H}$ of the features evaluated at the arguments $\phi(x)$.
+The theorem implies that the function $k$ can be written as the inner product in the Hilbert space $\mathcal{H}$ of the features evaluated at the arguments $\phi(x)$.
 
-Next, we make this characterisation explicit by showing how to construct a feature mapping $\phi$ of a kernel $k$ satisfying the finitely positive semidefinite property.
+Next, we make this characterisation explicit by showing how to construct a feature mapping $\phi$ of a kernel $k$ such that it is positive semidefinite.
 Let the feature space be the set of functions 
 \\
-$\mathcal{F} =\{ \sum\limits_{i=1}^n \alpha_i k(x_i, \cdot): n \in \mathbb{N}, x_i, y_i \in \mathcal{X}, a_i \in \mathbb{R}, i=1,\cdots, n\}$.
-This space is clearly closed under addition of functions and scalar multiplication, therefore it is a vector space.
+$$ 
+\begin{aligned}
+    \mathcal{F} =\bigg\{ \sum\limits_{i=1}^n \alpha_i k(x_i, \cdot): n \in \mathbb{N}, x_i, y_i \in \mathcal{X}, a_i \in \mathbb{R}, i=1,\cdots, n \bigg\}    
+\end{aligned}
+$$.
+This space is clearly closed under addition of functions and scalar multiplication. Therefore, it is a vector space.
 Letting $f(x)=\sum\limits_{i=1}^n \alpha_i k_i(x_i,x)$ and $g(x)=\sum\limits_{i=1}^m \beta_i k_i(y_i,x)$,  we can introduce an inner product on $\mathcal{F}$ as 
 \\
-$\langle f,g \rangle_{\mathcal{F}}:=\sum\limits_{i=1}^n\sum\limits_{j=1}^m a_i \beta_j k_i(x_i,y_j)=\sum\limits_{i=1}^n a_i g(x_i)= \sum\limits_{j=1}^m \beta_j f(y_j)$
+$$
+\begin{aligned}
+    \langle f,g \rangle_{\mathcal{F}}:=\sum\limits_{i=1}^n\sum\limits_{j=1}^m a_i \beta_j k_i(x_i,y_j)=\sum\limits_{i=1}^n a_i g(x_i)= \sum\limits_{j=1}^m \beta_j f(y_j)
+\end{aligned}
+$$
 \\
 Notice, $\langle f, f \rangle_{\mathcal{F}}= \alpha^\intercal K \alpha \geq 0 \ \forall f \in \mathcal{F}$, where $K$ is the kernel matrix.
-Moreover, $\langle f, g \rangle_{\mathcal{F}}$ is real valued, symmetric and bilinear. All these facts imply that the just introduced inner product is a valid metric.
+Moreover, $\langle f, g \rangle_{\mathcal{F}}$ is real valued, symmetric and bilinear and, hence, an inner product.
 \\
@@
 \\
-Finally, to show the characterisation of kernels we have to check that the inner product makes $\mathcal{F}$ a complete metric space.
+Finally, we extend $\mathcal{F}$ to a complete vector space.
 Consider a fixed input $x$ and a Cauchy sequence $\{f_n\}_{n=1}^{\infty}$, we then have by means of the Cauchy-Schwarz inequality that 
@@
 We have then, that $f_n(x)$ is a Cauchy sequence of real numbers.
-Thus, we can let $g(x):=\underset{n\to \infty}\lim f_n(x)$ and include all such limits functions in $\mathcal{F}$, doing so we complete $\mathcal{F}$ and we obtain the Hilbert space $\mathcal{H}$ associated to the kernel $k$.
-Having constructed the feature space $\mathcal{H}$, the mapping $\phi(x)$ is defined as follows
+Thus, we can let $g(x):=\underset{n\to \infty}\lim f_n(x)$ and add all such limits functions to $\mathcal{F}$, doing so we complete $\mathcal{F}$ and we obtain the Hilbert space $\mathcal{H}$ associated to the kernel $k$.
+Having constructed the feature space $\mathcal{H}$, the mapping $\phi(x)$ is defined as
 \\
@@
     \end{align}
-    Notice, the finitely positive semidefinite property of kernel $k$ is a sufficient condition for $\mathcal{H}$ to have an associated RKHS.
+    Notice, positive semidefinitenees property of kernel $k$ is a sufficient condition for $\mathcal{H}$ to have an associated RKHS.
 \end{definition}
 
-The analysis of above applies to any kernel, that is given a valid kernel function we can always construct its RKHS in the way we just illustrated.
+The analysis above applies to any kernel, that is, given a valid kernel function we can always construct its RKHS in the way we just illustrated.
 \begin{theorem}
-    If a symmetric function $k(\cdot, \cdot)$ satisfies the reproducing property in a Hilbert space $\mathcal{H}$, then $k$ satisfies the finitely positive semidefinite property.
+    If a symmetric function $k(\cdot, \cdot)$ satisfies the reproducing property in a Hilbert space $\mathcal{H}$, then $k$ is positive semidefinite.
 \end{theorem}
@@
 \begin{theorem}
-    [Riesz Representation].
-If $L : \mathcal{H} \rightarrow \mathbb{R}$ is a bounded linear operator on a Hilbert space $\mathcal{H}$ , there exists some $h_0 \in \mathcal{H}$ such that $L(f) = \langle f,h_0\rangle_\mathcal{H}, \ \forall f \in \mathcal{H}$.
+    [Riesz Representation \cite{Conway1994-kh}] If $L : \mathcal{H} \rightarrow \mathbb{R}$ is a bounded linear operator on a Hilbert space $\mathcal{H}$ , there exists some $h_0 \in \mathcal{H}$ such that $L(f) = \langle f,h_0\rangle_\mathcal{H}, \ \forall f \in \mathcal{H}$.
 \end{theorem}
 
-\begin{proof}
-    It is known that the null space of $L$, call it $M$, is a closed linear subspace in $\mathcal{H}$.
+% \begin{proof}
+%     It is known that the null space of $L$, call it $M$, is a closed linear subspace in $\mathcal{H}$.
 This implies that the Hilbert space direct sum $M \bigoplus M^\bot$ is an isomorphism of $\mathcal{H}$.
 When $L(f) \neq 0$ we have that $L(f)=L(m +w)$ with $m \in M$ and $w \in M^\bot$, it follows that $L(f)=0+L(w) \neq 0 \implies$ $M^\bot \neq {0}$.
-    \\
-    Using this fact, we can now choose a $z \in M^\bot$ with norm 1.
+%     \\
+%     Using this fact, we can now choose a $z \in M^\bot$ with norm 1.
 By linearity of $L$, for any $f \in \mathcal{H}$ we have $L(zL(f)- fL(z))=L(z)L(f)- L(f)L(z)=0$.
 This means that $zL(f)- fL(z)$ belongs to the null space of $L$, that is  $zL(f)- fL(z) \in M$.
-    \\
-    Consider now 
-    \begin{align}
-        L(f)=& L(f) \cdot 1
-        \\
-        =& L(f) \langle z,z\rangle_{\mathcal{H}}
-        \\
-        =& \langle zL(f), z\rangle_{\mathcal{H}} \\
-        =& \langle zL(f) -fL(z)+fL(z), z\rangle_\mathcal{H}
-        \\
-        =& \langle zL(f) -fL(z), z\rangle_{\mathcal{H}}+\langle fL(z), z\rangle_{\mathcal{H}} \\
-        % this is zero zL(f) -fL(z), so
-        =& \langle fL(z), z\rangle_{\mathcal{H}} \\
-        =& \langle f, z\overline{L(f)}\rangle_{\mathcal{H}} \\
-    \end{align}
-    Therefore, $L(f)=\langle f, h_0\rangle$  with  $h_0=z\overline{L(z)}$
-    \\
-    Finally, notice that $h_0$ is unique.
-    Assuming $h_0$ and $h_1$ to both satisfy the above equation for all $f \in \mathcal{H}$ we have that
-    \begin{align}
-        \langle h_0,f \rangle_{\mathcal{H}}=&\langle h_1, f \rangle_{\mathcal{H}} \\
-        \langle h_0,f \rangle_{\mathcal{H}}-\langle h_1, f \rangle_{\mathcal{H}}=& 0 \\
-        \langle h_0 -h_1,f \rangle_{\mathcal{H}}=& 0
-    \end{align}
-    Setting $f=h_0 -h_1$, we have the uniqueness of $h_0$.
-    % because inner product is zero, inner product has to satisfy positive definitness property to be such >=0, if it is zero it means that its argument x in <x,x> is zero => h_0=h_1
+%     \\
+%     Consider now 
+%     \begin{align}
+%         L(f)=& L(f) \cdot 1
+%         \\
+%         =& L(f) \langle z,z\rangle_{\mathcal{H}}
+%         \\
+%         =& \langle zL(f), z\rangle_{\mathcal{H}} \\
+%         =& \langle zL(f) -fL(z)+fL(z), z\rangle_\mathcal{H}
+%         \\
+%         =& \langle zL(f) -fL(z), z\rangle_{\mathcal{H}}+\langle fL(z), z\rangle_{\mathcal{H}} \\
+%         % this is zero zL(f) -fL(z), so
+%         =& \langle fL(z), z\rangle_{\mathcal{H}} \\
+%         =& \langle f, z\overline{L(f)}\rangle_{\mathcal{H}} \\
+%     \end{align}
+%     Therefore, $L(f)=\langle f, h_0\rangle$  with  $h_0=z\overline{L(z)}$
+%     \\
+%     Finally, notice that $h_0$ is unique.
+%     Assuming $h_0$ and $h_1$ to both satisfy the above equation for all $f \in \mathcal{H}$ we have that
+%     \begin{align}
+%         \langle h_0,f \rangle_{\mathcal{H}}=&\langle h_1, f \rangle_{\mathcal{H}} \\
+%         \langle h_0,f \rangle_{\mathcal{H}}-\langle h_1, f \rangle_{\mathcal{H}}=& 0 \\
+%         \langle h_0 -h_1,f \rangle_{\mathcal{H}}=& 0
+%     \end{align}
+%     Setting $f=h_0 -h_1$, we have the uniqueness of $h_0$.
+%     % because inner product is zero, inner product has to satisfy positive definitness property to be such >=0, if it is zero it means that its argument x in <x,x> is zero => h_0=h_1
 
-\end{proof}
+% \end{proof}
+
 %conjugate simmetry is because in complex analysis when you swap in the inner product, you have to take the conjugate
 
-The Riesz representation theorem results in the following proposition for reproducing kernel hilbert spaces.
+The Riesz representation theorem results in the following proposition for reproducing kernel Hilbert spaces.
 \begin{proposition}
@@
 The function $k(\cdot, x)$ is the reproducing kernel evaluated at point $x$.
-Furthermore, note that $k(\cdot, x)$ is itself a function lying in it.
-
+Furthermore, note that $k(\cdot, x)$ is itself a function lying in it
 \begin{align*}
@@
 
-\begin{definition}
-    Considering a set of vectors $x_1, \cdots, x_n$, the Gram matrix is defined as the $n\times n$ matrix whose entries are $\langle x_i, x_j \rangle_{\mathcal{X}}$.    
-\end{definition}
-
-Notice, it follows from the definition that Gram matrix is symmetric.
-In the kernel setting we evaluate the inner products in the feature space generated by the feature map $\phi$, therefore the Gram matrix will have entries $\langle \phi(x_i), \phi(x_j) \rangle_{\mathcal{H}}= k(x_i, x_j)$.
-Such matrix is also referred as the kernel matrix $K$.
-
-\[
-K_{n\times n}=
-\begin{bmatrix}
-    k(x_1,x_1)       &  \dots & k(x_1,x_n) \\
-    \vdots       & \ddots & \vdots \\
-    k(x_n, x_1)       & \dots & k(x_n, x_n)
-\end{bmatrix}
-\]
-All kernel methods take as input the kernel matrix, it is the central data type for all 
-such techniques.
-
-\begin{proposition}
-    In addition to being symmetric, $K$ possesses also the property of being positive semidefinite.    
-\end{proposition}
-
-\begin{proof}
-    \begin{align}
-        z^\intercal K z =& \sum_{i,j=1}^n z_i z_j k(x_i,x_j)
-        \\
-        =& \sum_{i,j=1}^n z_i z_j \langle \phi(x_i), \phi(x_j) \rangle_{\mathcal{H}}
-        \\
-        =& \left\langle \sum_{i=1}^n z_i \phi(x_i), \sum_{j=1}^n z_j \phi(x_j)\right\rangle_{\mathcal{H}}
-        \\
-        =& \left\|\sum_{i=1}^n z_i \phi(x_i) \right\|_{\mathcal{H}}^2 \geq 0
-    \end{align}
-\end{proof}
-We know that positive semidefinite matrices form a cone in the vector space of $n\times n$ matrices.
-Because of this we have that such matrices are convex, this implies that all the nice theory of convex analysis can be used when optimising over them.
-
-Kernel functions are characterized by a couple of properties that make them a closure. Thus, there exists ways of combining known kernels in order to obtain new kernels.
+% Kernel functions are characterised by a couple of properties that make them a closure. Thus, there exists ways of combining known kernels in order to obtain new kernels.
 
@@
 Then the following functions are valid kernels:
-    \begin{align}
+    $$
+    \begin{aligned}
         k(x,y)=&k_1(x,y)+k_2(x,y)
@@
         k(x,y)=& k_3(\phi(x),\phi(y))
-    \end{align}.
+    \end{aligned}
+    $$.
 \end{proposition}
@@
     \\
-    \begin{align}
+    $$  
+    \begin{aligned}
     z^\intercal (K_1+K_2) z&= z^\intercal K_1 z + z^\intercal K_2 z \geq 0
@@
     z^\intercal aK z&= a z^\intercal K z \geq 0
-    \end{align}
+    \end{aligned}
+    $$
     Let $K=K_1 \otimes K_2$, we have that the eigenvalues of $K$ are made up by the product pairs of the eigenvalues of $K_1$ and $K_2$.
@@
 Using the fact that K is positive semidefinite we have that for any $z$, it holds
-    \begin{align}
+    $$
+    \begin{aligned}
         z^\intercal H z= t^\intercal K t\geq 0
-    \end{align}
+    \end{aligned}
+    $$
 \end{proof}
@@
 \section{Kernel families}
-The selection of kernel determines the class of functions that will be searcehd by the learning algorithm.
+The selection of the kernel determines the class of functions that will be searched by the learning algorithm.
 Prior knowledge of the problem domain can help restricting the candidate families of kernel functions.
@@
        Polynomial &  $k(x,y)=(x^\intercal y+c)^d$ &   c, d\\
-       Gaussian RBF &  $k(x,y)=\exp\left(-\frac{\|x-y|\|_{2}^2}{2\sigma^2}\right)$ &   $\sigma$\\
-       Exponential RBF/Laplacian& $k(x,y)=\exp\left(-\frac{\|x-y\|_{1}}{\gamma}\right)$ &   $\gamma$\\
+       Gaussian RBF &  $k(x,y)=\exp\left(-\frac{\|x-y\|_{2}^2}{2\l^2}\right)$ &   $l$\\
+       Exponential RBF/Laplacian& $k(x,y)=\exp\left(-\frac{\|x-y\|_{2}}{l}\right)$ &   $l$\\
        Hyperbolic/Sigmoid Kernel &  $k(x,y2)=\tanh(\gamma x^\intercal y+r)$ &  $\gamma$, r \\
@@
         Cosine  &  $k(x,y)=\frac{xy^\intercal}{\|x\|_{2}\|y\|_{2}}$ &   \\
-       Matern  &  $k(x, y) =  \frac{1}{\Gamma(\nu)2^{\nu-1}}\Bigg(
+       Matérn  &  $k(x, y) =  \frac{1}{\Gamma(\nu)2^{\nu-1}}\Bigg(
         \frac{\sqrt{2\nu}}{l} \|x-y|\|_{2}
@@
 
-Notice, in the Matern family, $\Gamma(\cdot)$ is the gamma function while $Y(\cdot)$ is the Bessel function of the second kind.
+Notice, in the Matérn family, $\Gamma(\cdot)$ is the gamma function while $Y(\cdot)$ is the Bessel function of the second kind.
 \\
-When $\nu=p+\frac{1}{2}$, $p \in \mathbb{N}^+$, the Matern kernel can be rewritten as a product of an exponential and a polynomial of degree $p$. 
+When $\nu=p+\frac{1}{2}$, $p \in \mathbb{N}^+$, the Matérn kernel can be rewritten as a product of an exponential and a polynomial of degree $p$. 
 \\
-$k(x,y)=\exp \left(-{\frac {{\sqrt {2p+1}}\|x-y|\|_{2}}{l }}\right){\frac {p!}{(2p)!}}\sum\limits_{i=0}^{p}{\frac {(p+i)!}{i!(p-i)!}}\left({\frac {2{\sqrt {2p+1}}\|x-y|\|_{2}}{l }}\right)^{p-i}$
+$$k(x,y)=\exp \left(-{\frac {{\sqrt {2p+1}}\|x-y|\|_{2}}{l }}\right){\frac {p!}{(2p)!}}\sum\limits_{i=0}^{p}{\frac {(p+i)!}{i!(p-i)!}}\left({\frac {2{\sqrt {2p+1}}\|x-y|\|_{2}}{l }}\right)^{p-i}$$
 \\
@@
 \begin{itemize}
-    \item $p=0,\nu=\frac{1}{2} \implies k(x,y)=\exp\left(-\frac{\|x-y|\|_{2}}{l}\right)$
+    \item $p=0,\nu=\frac{1}{2} \implies k(x,y)=\exp\left(-\frac{\|x-y\|_{2}}{l}\right)$
     \item $p=1,\nu=\frac{3}{2} \implies 
-    k(x,y)=\left(1+{\frac {{\sqrt {3}}\|x-y|\|_{2}}{l }}\right)\exp \left(-{\frac {{\sqrt {3}}\|x-y|\|_{2}}{l }}\right)$
-    \item $p=2,\nu=\frac{5}{2} \implies k(x,y)=\left(1+{\frac {{\sqrt {5}}\|x-y|\|_{2}}{l }}+{\frac {5\|x-y|\|_{2}^{2}}{3 l ^{2}}}\right)\exp \left(-{\frac {{\sqrt {5}}\|x-y|\|_{2}}{l }}\right)$
+    k(x,y)=\left(1+{\frac {{\sqrt {3}}\|x-y\|_{2}}{l }}\right)\exp \left(-{\frac {{\sqrt {3}}\|x-y\|_{2}}{l }}\right)$
+    \item $p=2,\nu=\frac{5}{2} \implies k(x,y)=\left(1+{\frac {{\sqrt {5}}\|x-y\|_{2}}{l }}+{\frac {5\|x-y\|_{2}^{2}}{3 l ^{2}}}\right)\exp \left(-{\frac {{\sqrt {5}}\|x-y\|_{2}}{l }}\right)$
 \end{itemize}
 
-Furthermore, as $\nu \to \infty$ we have that the Matern kernel converges to the Gaussian RBF kernel.
-$k(x,y)=\exp\left(\frac{-\|x-y|\|_{2}^2}{2l^2}\right)$
+When $\nu=\frac{1}{2}$ the Matérn kernel corresponds to the Laplacian kernel.
+Furthermore, as $\nu \to \infty$ we have that the Matérn kernel converges to the Gaussian RBF kernel.
+$k(x,y)=\exp\left(\frac{-\|x-y\|_{2}^2}{2l^2}\right)$
 
@@
 
-Any algorithm involving inner products between inputs can be combined with a kernel method.
+% Any algorithm involving inner products between inputs can be combined with a kernel method.
 The chosen kernel function will calculate the inner product between the images of two inputs in the feature space.
@@
 Kernel methods show good modularity because they work the same for any kernel and any data type.
-This concept is illustrated in figure \ref{fig:workflow_kernels}, where an abstract kernel methods workflow is visualized.
+This concept is illustrated in Figure \ref{fig:workflow_kernels}, where an abstract kernel methods workflow is visualized.
 First we build the kernel matrix by using a kernel function to process the input data.
-Next, the kernel matrix is passed to the pattern analysis algorithm, which returns the learned function.
+Next, the kernel matrix is passed to the pattern analysis (PA) algorithm, which returns the learned function.
 

### 2024-05-25 thesis/input_files/point.tex
May 25 2024

@@
 This chapter covers the theory of the most widely used methods for point prediction in the electricity forecasting literature.
-Besides discussing the theory underlying such models, this and the following chapter include also a couple of worked examples in order to get acquainted with the practical applications of these models.
+Besides discussing the theory underlying such models, this and the following chapter include also a couple of examples in order to get acquainted with the practical applications of these models.
 
@@
 Notwithstanding its simplicity, multiple linear regression models are still popular among the electricity point forecasting literature.
-Notice, they are not used per se but usually combined with other more advanced models.
+Notice that they are not used per se but usually combined with other more advanced models.
+Letting $y_t$ be the dependent variable, $X_t$ be the matrix made up of independent variables and $\epsilon_t$ be the vector of the errors random variable, multiple linear regression takes the form
 \begin{equation}
@@
 Before introducing this popular model in electricity point forecasting, it is important to remember that autoregressive models assume the time series to be stationary.
-On the other hand, load and price have been observed to be a non stationary time series.
-In a nutshell, stationarity means that the distribution of any subsequence of random variables of the stochastic process is invariant to shifts along the time dimension.
+However, load and price have been observed to be a non stationary time series.
+Stationarity means that the distribution of any subsequence of random variables of the stochastic process is invariant to shifts along the time dimension.
 Therefore, checking stationarity is a crucial step in applying this class of models. %(unit root tests)
@@
 \end{equation}
-where the parameter p is the order of the autoregressive (AR) part, d is the degree of integrated differencing and q is the order of the moving average (MA) model.
-$\epsilon_{n-j}$ are the observed errors in the past while the differenced term $y_n^{(d)}$ is defined recursively as $y_n^{(d)}:=y_n^{(d-1)}-y_{n-1}^{(d-1)}$.
+where the parameter $p$ is the order of the autoregressive (AR) part, $d$ is the degree of integrated differencing and $q$ is the order of the moving average (MA) model.
+The quantities $\epsilon_{n-j}$ are the observed errors in the past, while the differenced term $y_n^{(d)}$ is defined recursively as $y_n^{(d)}:=y_n^{(d-1)}-y_{n-1}^{(d-1)}$.
 \\
@@
 \end{equation}
-where $B$ is the backward shift operator, $B^h y_t=y_{t-h}$ and $\psi(B)$ is shorthand for $\psi(B)=1-\psi_1B-\dots-\psi_p B^p$.
+where $B$ is the backward shift operator, $B^h y_t=y_{t-h}$ and $\psi(B)$ is shorthand for $\psi(B)=$1-$\psi_1B-\dots-\psi_p B^p$.
 Similarly, $\theta(B)=1+\theta_1B+\dots+\theta_q B^q$ and $\nabla_h$ is the lag $h$ differencing operator $\nabla_h y_t=(1-B^h)y_t=y_t-y_{t-h}$
@@
 In the same way, the order $q$ is estimated by considering the sample autocorrelation function (ACF) of the observed time series.
-Finally, we compare ARIMA models with different $p,d,q$ (in a neighbourhood of the identified orders) and choose the ones minimising either AIC or BIC criteria.
+Finally, we compare ARIMA models with different $p,d,q$ (in a neighbourhood of the identified orders) and choose the ones minimising either Akaike information criterion (AIC) or Bayesian information criterion (BIC) criteria.
 % Notice, models are obtained through the maximum likelihood estimation framework.
@@
 \subsection{Autoregressive integrated moving average with exogenous variables}
-ARIMAX model adds explanatory variables to our ARIMA, hence, we have
+The ARIMAX model adds explanatory variables to ARIMA. Hence, we have
 \begin{equation}
@@
 \subsection{Seasonal autoregressive integrated moving average and seasonal autoregressive integrated moving average with exogenous variables}
-SARIMA and SARIMAX incorporate seasonalities into our time series models.
+SARIMA and SARIMAX incorporate seasonalities into time series models.
 The notation for these models is ARIMA$(p,d,q)(P,D,Q)_m$.
-The second term $(P,D,Q)_m$ represents the autoregressive structure of the seasonal pattern, where $m$ indicates the number of seasonalities of our time series.
+The second term $(P,D,Q)_m$ represents the autoregressive structure of the seasonal pattern, where $m$ indicates the number of seasonalities of the time series.
 For instance, in modelling series of hourly data with daily periodicity, we would set $m=24$.
@@
 $f_i$ are smooth functions which are fitted using cubic smoothing splines.
-$g$ is called link function and characterizes the specific GAM model \cite{hastie2017generalized}. Approximations for $f_i$ are obtained through an iterative procedure, the backfitting algorithm.
+Herein, $g$ is called link function and characterizes the specific GAM model \cite{hastie2017generalized}.
+Approximations for $f_i$ are obtained through an iterative procedure, the backfitting algorithm.
 
@@
 This additive model is capable of handling non linear trends, holiday effects and yearly weekly, and daily seasonalities.
-Thus, the reason why we choose it as a valid benchmark to compare against.
+This is, the reason why we choose it as a benchmark to compare against.
 \section{K-nearest neighbours regression}
@@
 $k$ is the algorithm hyperparameter, it stands for the number of neighbours; a small $k$ leads to overfitting while a big $k$ leads to underfitting.
-Since it is a metric based algorithm, it is important to normalize data in order to give equal weights to features with different scales, see \ref{appendix:normalization}.
+Since it is a metric based algorithm, it is important to normalize data in order to give equal weights to features with different scales, see Appendix \ref{appendix:normalization}.
 This algorithm is made up of two decisions: first, the choice of the metric and second, the method for combining targets.
@@
 \section{Support vector regression}
-Developed at the AT\&T Bell Laboratories by Vapnik et al.\ \cite{cortes1995support} \cite{vapnik1997support}, support vector machines (SVMs) are one of the most popular techniques within the field of statistical learning.
+Developed at the AT\&T Bell Laboratories by Vapnik et al.\ \cite{cortes1995support, vapnik1997support}, support vector machines (SVMs) are one of the most popular techniques within the field of statistical learning.
 \\
 The goal of support vector regression is finding a function $f(x)$ with at most $\epsilon$ deviation from the actual observed data $y_i$ for every $i$ and as flat as possible. 
-Put differently, we would like a model to keep errors less than the $\epsilon$ threshold.  
+Put differently, we would like a model to keep the errors less than an $\epsilon$ threshold.  
 %https://stats.stackexchange.com/questions/5945/understanding-svm-regression-objective-function-and-flatness
@@
 \end{equation}
-Where $\mathcal{X}$ denotes the space of the input data $x_i$.
+Where $\mathcal{X}$ denotes the set of input data $x_i$.
 We can translate the flatness requirement into minimising the squared norm of $w$. Doing so, we can formulate our problem as a convex optimisation problem.
@@
 \\
-We seek to minimise the epsilon insensitive loss function, defined as follows
+We seek to minimise the $\epsilon$ insensitive loss function, defined as follows
 \begin{equation}
+    \begin{aligned}
         \|\xi\|_\epsilon:=\begin{cases}
-        0 \quad& \textrm{if} \ |\xi|\leq \epsilon
+            0, \quad& \textrm{if} \ |\xi|\leq \epsilon,
             \\
-        |\xi|-\epsilon \quad& \textrm{if} \ |\xi| > \epsilon
+            |\xi|-\epsilon, \quad& \textrm{if} \ |\xi| > \epsilon
         \end{cases}
+    \end{aligned}
 \end{equation}
-The model is depicted in figure \ref{fig:svm_simple}. The grey band is called the epsilon insensitive tube, only the points outside it are accounted by the loss function.
+The model is depicted in Figure \ref{fig:svm_simple}. The grey band is called the epsilon insensitive tube, only the points outside it are accounted by the loss function.
 \begin{figure}
@@
 \end{figure}
-Smola et al.\ \cite{smola2004tutorial} point out that considering the dual formulation makes our optimisation problem easier to solve. Doing so we have the equivalent optimisation problem
+Smola et al.\ \cite{smola2004tutorial} point out that considering the dual formulation makes our optimisation problem easier to solve. Doing so, we have the equivalent optimisation problem
 \begin{equation}
@@
 \end{equation}
-Rearranging the gradient of the primal Lagrangian with respect to $w$, we obtain the so called support vector expansion.
+Rearranging the gradient of the primal Lagrangian with respect to $w$, we obtain the so called support vector expansion
 \begin{equation}\label{eq:support vector expansion}
@@
 This means that $w$ can be completely described as a linear combination of the training features $x_i$.
-Notice, the complexity of the function representation is independent of the feature space dimensionality, but depends only on the number of support vectors; that is those points $i$ for which $(\lambda_i-\lambda_i^*)\neq 0$
+Notice, the complexity of the function representation is independent of the feature space dimensionality, but depends only on the number of support vectors, that is, those points $i$ for which $(\lambda_i-\lambda_i^*)\neq 0$
 Moreover, we do not need to compute $w$ explicitly in order to evaluate $f(x)$.
-That is because, equation \ref{eq:support vector expansion} implies
+That is because, Equation \ref{eq:support vector expansion} implies
 \begin{equation}
@@
 \end{equation}
-Employing the Karush-Kuhn-Tucker conditions, $b$ can be retrieved easily. Particularly, the condition that the product between the constraints and the dual variable has to vanish.
+Employing the Karush-Kuhn-Tucker conditions, $b$ can be retrieved easily. 
+% Particularly, the condition that the product between the constraints and the dual variable has to vanish.
 Consequently for any of the data points with associated $\lambda_i, \lambda^*_i \in (0,C)$, the equality $b=y_i-\langle w, x_i-\epsilon\rangle$ has to hold.
 \\
-Furthermore, this implies that the Lagrange multipliers $\lambda_i, \lambda_i^*$ may be nonzero only for the samples inside the epsilon insensitive tube, that is only those points may be support vectors.
+This implies that the Lagrange multipliers $\lambda_i, \lambda_i^*$ may be nonzero only for the samples inside the $\epsilon$ insensitive tube, that is only those points may be support vectors. 
 \\
-To get an idea of how the algorithm works, see figure \ref{fig:svr1} for how a support vector regression handles a sinusoidal function with noise.
+To get an idea of how the algorithm works, see Figure \ref{fig:svr1} for how a support vector regression handles a sinusoidal function with noise.
 \\
@@
 The basic building block of this class of methods is the perceptron \cite{rosenblatt1958perceptron}.
+This model is defined as follows
 \begin{equation}
@@
 \end{equation}
-where $\psi$ is an activation function and $w,b$ are the parameters of the neuron; such parameters are typically optimized through gradient descent.
+where $\psi$ is an activation function and $w,b$ are the parameters of the neuron. The parameters are typically optimized through gradient descent.
 
 \subsection{Multilayer perceptrons}
-In order to cover a richer space of models we can stack neurons, doing so we obtain the so called multi layer perceptrons (MLPs).
+In order to cover a richer space of models we can stack neurons.
+Doing so we obtain the so called multi layer perceptrons (MLPs).
 Hyperparameters of these models are the number of hidden layers, the number of neurons in each of those layers and the kind of activation function.
@@
 % https://rtavenar.github.io/deep_book/book_en.pdf
-Long short term memory network (LSTM) extends MLP by introducing a cell state in its layers.
-In this way, the current state of a cell depends on the current value $X_t$, on the previous cell activation and on the previous cell state $C_{t-1}$, see figure \ref{fig:lstm_cell} for a visual representation.
+Long short term memory network (LSTM) extends MLP by introducing a cell state in its layers \cite{hochreiter1997long}.
+In this way, the current state of a cell depends on the current value $X_t$, on the previous cell activation and on the previous cell state $C_{t-1}$, see Figure \ref{fig:lstm_cell} for a visual representation.
 \begin{figure}[!h]
@@
 The cell is responsible for deciding whether to store/forget long and short term information. To do so, each cell is made of a forget gate, an input gate and an output gate.
-The forget gate takes in $h_{t-1}$ and $x_t$ and outputs a real valued vector with elements in the domain $[0,1]$, this corresponds to the ratio of information retention for each element of the cell state $C_{t-1}$.
+The forget gate takes in $h_{t-1}$ and $x_t$ and outputs a real valued vector with elements in $[0,1]$.
+This corresponds to the ratio of information retention for each element of the cell state $C_{t-1}$.
 \\
@@
 \begin{equation}
-    f_t=\sigma(W_f\dot[h_{t-1},x_t]+b_f)
+    f_t=\sigma(W_f[h_{t-1},x_t]+b_f)
 \end{equation}
@@
 First, a sigmoid layer decides which values of the cell state will be updated.
-Second, a tahn layer creates a vector of candidates $\tilde{C}_t$ to add to the current state.
+Second, a tanh layer creates a vector of candidates $\tilde{C}_t$ to add to the current state.
 \begin{equation}
     \begin{aligned}
-        i_t=&\sigma(W_i\dot[h_{t-1},x_t]+b_i)
+        i_t=&\sigma(W_i [h_{t-1},x_t]+b_i)
         \\
-        \tilde{C}_t=&tahn(W_C\dot[h_{t-1},x_t]+b_C)
+        \tilde{C}_t=&tanh(W_C [h_{t-1},x_t]+b_C)
     \end{aligned}
@@
     \\
-    h_t=& o_t \odot tahn(C_t)
+    h_t=& o_t \odot tanh(C_t)
 \end{aligned}
@@
 \subsection{Kernel ridge regression}
-In the setting of kernel ridge regression, we aim at estimating $f$ such that $y=f(x)+\epsilon$. Given a RKHS $\mathcal{H}$, we can estimate $\hat{f}$ by solving the optmisation problem
+In the setting of kernel ridge regression, we aim at estimating $f$ such that $y=f(x)+\epsilon$. Given an RKHS $\mathcal{H}$, we can estimate $f$ by solving the optmisation problem
 \begin{equation}\label{eq:kernel ridge}
@@
     Consider the regularized risk functional $Risk[f]:f \mapsto  E\left((x_{1},y_{1},f(x_{1})),\ldots, (x_{n},y_{n},f(x_{n}))\right)+g\left(\| f\|_{\mathcal{H}} \right)$.
-Then any minimizer of the empirical risk can be represented as $ f^{*}(\cdot )=\sum\limits_{i=1}^{n}\alpha _{i}k(\cdot ,x_{i})$
+Then any minimizer of the empirical risk can be represented as $$ f^{*}(\cdot )=\sum\limits_{i=1}^{n}\alpha _{i}k(\cdot ,x_{i})$$
 \end{theorem}
@@
     This means that setting $\nu=0$ does not affect the evaluation of $f$.
-    Next consider the regularization term; notice, the functional $g$ is defined to be strictly monotonic in order to penalize more complex functions $f$.
+    Next consider the regularization term. The functional $g$ is defined to be strictly monotonic in order to penalize more complex functions $f$.
     \begin{equation}
@@
     Hence, we can conclude that setting $\nu=0$ decreases the second term while the first term is not affected.
-Thus, it follows that any minimizer of the risk functional of equation \ref{eq:kernel ridge} must be of the following form
+Thus, it follows that any minimizer of the risk functional of Equation \ref{eq:kernel ridge} must be of the following form
     \begin{equation}
@@
 
-Employing the representer theorem, we can rewrite \ref{eq:kernel ridge} in matrix notation as
+Employing the representer theorem, we can rewrite Equation \ref{eq:kernel ridge} in matrix notation as
 \begin{equation}
@@
 \subsection{Kernel support vector regression}
-Support vector regression (SVR) can be kernelized by swapping the $\mathbb{R}^n$ euclidean dot product of data $x$ with the dot product in the higher feature space $\mathcal{H}$.
+Support vector regression (SVR) can be kernelized by swapping the $\mathbb{R}^n$ euclidean inner product of $x$ with the inner product in the higher feature space $\mathcal{H}$.
 Doing so, the optimisation problem can be restated as
@@
 \end{equation}
-Notice, in this setting $w$ is no longer given explicitly. Additionally, with kernel support vector regression we seek for the flattest function in the feature space not the input space.
-See figure \ref{fig:svr2} and figure \ref{fig:svr3} for two examples. In the former, we have used a polynomial kernel while in the latter the standard radial basis function kernel.
+Notice, in this setting $w$ is no longer given explicitly. Additionally, with kernel support vector regression we seek for the flattest function in the feature space instead of the input space.
+See Figure \ref{fig:svr2} and Figure \ref{fig:svr3} for two examples. In the former, we have used a polynomial kernel and the standard radial basis function kernel in the latter.
 \begin{figure}
@@
 \end{figure}
-Comparing these pictures with figure \ref{fig:svr1}, it can be concluded that introducing kernels allows support vector regression to handle the non linearities in the data.
+Comparing these pictures with Figure \ref{fig:svr1}, it can be concluded that introducing kernels allows support vector regression to handle non linearities in the data.

### 2024-05-25 thesis/input_files/probabilistic.tex
May 25 2024

@@
 \subsection{Decision trees}
-Decision trees methods partition recursively the feature space in a set of binary rectangles and then fit a simple model in each of those partitions (the most straightforward is fitting just a constant).
+Decision tree methods partition recursively the feature space in a set of binary rectangles and then fit a simple model in each of those partitions (the most straightforward is fitting just a constant).
 To get started, we first split the space into two disjoint regions, then we model the response variable by the mean of the observed predicted variables with associated features falling in that specific region.
-Our goal is selecting the best features and best split point such that to achieve the best generalizing fit.
-For a visualisation consider the pictures \ref{fig:elements_statistical_learning1} and \ref{fig:elements_statistical_learning2}, where the decision tree algorithm is visualised for a regression problem with two independent variables $X_2$ and $X_1$.
+Our goal is selecting the best features and best split point to achieve the best generalizing fit.
+For a visualisation consider the Figures \ref{fig:elements_statistical_learning1} and \ref{fig:elements_statistical_learning2}, where the decision tree algorithm is visualised for a regression problem with two independent variables $X_1$ and $X_2$.
 \begin{figure}
@@
 \end{equation}
-For the outer problem, we scan through all the (j,s) tuples and pick the best pair. 
-Next, one or both of these regions from the previous step are split into two more regions; we recurse this process until some stopping condition is triggered (max number of branches, max depth of tree, threshold on the mean squared error (MSE), minimum number of observations in each leaf node).
+For the outer problem, we scan through all the $(j,s)$ tuples and pick the best pair. 
+Next, one or both of these regions from the previous step are split into two more regions.
+We recurse this process until some stopping condition is triggered (max number of branches, max depth of tree, threshold on the mean squared error (MSE), minimum number of observations in each leaf node).
 Notice, this being a greedy algorithm implies that our final solution is guaranteed to be just a local optimum not a global one.
 Even though their simplicity, these models have proved themselves to be really powerful.
-See figure \ref{fig:decision_tree} for an example where decision trees with different hyperparameters are fitted to a sine wave plus noise.
+See Figure \ref{fig:decision_tree} for an example where decision trees with different hyperparameters are fitted to a sine wave plus noise.
 The most popular among these models is the CART \cite{breiman2017classification} tree, its name comes from the fact that it can handle both classification and regression problems.
@@
 The idea behind bootstrapping is to randomly sample from the training set with replacement $B$ times and then fitting $B$ models to each of the "artificial" datasets.
-Bootstrapping can serve different tasks; we can use it to assess the accuracy of a parameter %estimate
+Bootstrapping can serve different tasks.
+We can use it to assess the accuracy of a parameter %estimate
 or of a prediction but also to improve their estimates.
@@
 \end{equation}
-For a simple visualisation compare figure \ref{fig:random_forest} with \ref{fig:decision_tree}.
+For a simple visualisation compare Figure \ref{fig:random_forest} with Figure \ref{fig:decision_tree}.
 \begin{figure}
@@
 \subsection{Gradient boosting}
-In order to robustly solve \ref{boosting_minimisation_problem}, gradient boosting considers the following problem
+In order to robustly solve Equation \ref{boosting_minimisation_problem}, gradient boosting considers the following problem
 \begin{equation}
@@
 In order to solve our quantile regression tasks through gradient boosting, all we need to do is specifying the pinball loss as our criterion to guide the minimisation algorithm.
-See figure \ref{fig:gradient_boosting} for a visualisation.
+See Figure \ref{fig:gradient_boosting} for a visualisation.
 \begin{figure}
@@
 \begin{equation}\label{eq:kqr_min2}
-    \begin{aligned}
-    \min_{w,b} \quad & C \sum \limits_{i=1}^{m}
-    q(y_i-\langle w, \phi(x_i) \rangle_{\mathcal{H}}-b)+ (1-q)(-y_i+\langle w, \phi(x_i) \rangle_{\mathcal{H}}+b)+ \frac{1}{2}\|w\|_{\mathcal{H}}^2\\
-    \end{aligned}
+    \begin{split}
+        \min_{w,b} \quad & C \sum \limits_{i=1}^{m} \hbox{$q(y_i - \langle w, \phi(x_i) \rangle_{\mathcal{H}} - b)$} \\
+        & + \hbox{$(1-q)(-y_i + \langle w, \phi(x_i) \rangle_{\mathcal{H}} + b)$} \\
+        & + \frac{1}{2}\|w\|_{\mathcal{H}}^2
+    \end{split}
     \end{equation}
@@
 \\
-We can next rephrase the optimisation in \ref{eq:kqr_min2} by introducing the slack variables $\xi_i$ and $\xi_i^*$.
+We can next rephrase the optimisation in Equation \ref{eq:kqr_min2} by introducing the slack variables $\xi_i$ and $\xi_i^*$.
 \begin{equation}\label{eq:kqr_min3}
@@
     \end{equation}
-In order to make it more compact, we rewrite equation \ref{eq:kqr_min3} in matrix notation.
+In order to make it more compact, we rewrite Equation \ref{eq:kqr_min3} in matrix notation.
 \begin{equation}\label{eq:kqr_min4}
     \begin{aligned}
-        \min_{w,b\xi,\xi^*} \quad & C q \xi^\intercal \mathbb{1}+ C q (\xi^*)^\intercal \mathbb{1}+ \frac{1}{2}w^\intercal w\\
+        \min_{w,b\xi,\xi^*} \quad & C q \xi^\intercal \mathbb{1}+ C (1-q) (\xi^*)^\intercal \mathbb{1}+ \frac{1}{2}w^\intercal w\\
     \textrm{s.t.} \quad & y-\Phi^\intercal w -b \preceq \xi\\
@@
     \end{equation}
-Consider now, the Lagrangian $\mathcal{L}$ associated to \ref{eq:kqr_min4}
+Consider now, the Lagrangian $\mathcal{L}$ associated to Equation \ref{eq:kqr_min4}
 \begin{equation}\label{eq:kqr_min5}
-    \begin{aligned}
-    L(w,b,\xi,\xi^*)= & C q \xi^\intercal \mathbb{1}+ C q (\xi^*)^\intercal \mathbb{1}+ \frac{1}{2}w^\intercal w- \lambda^\intercal(\xi - y+\Phi^\intercal w +b)
-    - (\lambda^*)^\intercal(\xi^* +y-\Phi^\intercal w -b)
-    \\
-    & -\nu^\intercal \xi - (\nu^*)^\intercal \xi^*
-\end{aligned}
+    \begin{split}
+        L(w, b, \xi, \xi^*) = \quad & \hbox{$C q \xi^\intercal \mathbb{1} + C (1-q) (\xi^*)^\intercal \mathbb{1} + \frac{1}{2} w^\intercal w$} \\
+        & \hbox{$- \lambda^\intercal (\xi - y + \Phi^\intercal w + b)$} \\
+        & \hbox{$- (\lambda^*)^\intercal (\xi^* + y - \Phi^\intercal w - b)$} \\
+        & \hbox{$- \nu^\intercal \xi - (\nu^*)^\intercal \xi^*$}
+    \end{split}
 \end{equation}
@@
 where $p^*$ is the optimal value of your optimisation problem. In other words $g(\lambda, \nu)$ is a lower bound for the optimal $p^*$.
-Consider now a simple Lagrangian $\mathcal{L}(x,\lambda, \nu)=f(x)+\sum\limits_{i=1}^n \lambda_i r_i(x) +\sum\limits_{i=1}^{n} \nu_i h_i(x)$, where $r_i(x)$ are inequality constraints while $h_i(x)$ are equality constraints of the problem.
+Consider now a simple Lagrangian $$\mathcal{L}(x,\lambda, \nu)=f(x)+\sum\limits_{i=1}^n \lambda_i r_i(x) +\sum\limits_{i=1}^{n} \nu_i h_i(x)$$
+where $r_i(x)$ are inequality constraints while $h_i(x)$ are equality constraints of the problem.
 Then it can be noted that, the lower bound on $p^*$ is non trivial only when the Lagrange multiplier $\lambda \succeq 0$.
@@
 \\
-To explain why we may or not be able to attain the best solution to the primal problem by maximising its dual, we have to introduce the concept of duality.
+To explain why we may or may not be able to attain the best solution to the primal problem by maximising its dual, we have to introduce the concept of duality.
 \\
@@
 \\
-The inequality \ref{weak_duality} is called weak duality.
-The difference $p^*-d^*$ is said the optimal duality gap; it is the gap between the optimal value of the primal problem and the best lower bound on it that can be obtained from the Lagrange dual function.
+The Inequality \ref{weak_duality} is called weak duality.
+The difference $p^*-d^*$ is the optimal duality gap.
+It is the gap between the optimal value of the primal problem and the best lower bound on it that can be obtained from the Lagrange dual function.
 Moreover, note that the optimal duality gap is always nonnegative.
-We say that strong duality holds, when the optimal duality gap is zero; in other words, the Lagrange dual bound is tight.
+We say that strong duality holds, when the optimal duality gap is zero. In other words, the Lagrange dual bound is tight.
 \\
-Constraint qualifications are conditions under which strong duality holds; one of the most popular is Slater's condition.
+Constraint qualifications are conditions under which strong duality holds. One of the most popular is Slater's condition.
 \begin{proposition}
@@
 \\
-In our case we don't have any equality constraint, so we do not have to worry about the $h_i(x)=0$ term in \ref{slater_condition}.
+In our case we do not have any equality constraint, so we do not have to worry about the $h_i(x)=0$ term in Equation \ref{slater_condition}.
 All we have to check is the convexity of our problem and that there exists an $x$ such that $r_i(x)<0$.
@@
 \\
-We end this section with the derivation of the dual problem; that is the convex problem, which we will solve in order to get the quantiles prediction of our quantile kernel algorithm.
+We end this section with the derivation of the dual problem, that is, the convex problem, which we will solve in order to get the quantiles prediction of our quantile kernel algorithm.
 \\
-First, derive the dual function of \ref{eq:kqr_min5}.
+First, take the dual function from Equation \ref{eq:kqr_min5}.
 \begin{equation}
@@
 \begin{equation}
-    \begin{aligned}
-        g(\lambda, \lambda^*)=& \xi^\intercal(Cq\mathbb{1}-\lambda -\nu)+(\xi^*)^\intercal(C(1-q)\mathbb{1}-\lambda^*-\nu^*)-(\lambda-\lambda^*)^\intercal \Phi\Phi^\intercal(\lambda-\lambda^*)
+    \begin{split}
+        g(\lambda, \lambda^*) = \quad & \hbox{$\xi^\intercal(Cq\mathbb{1} - \lambda - \nu) + (\xi^*)^\intercal(C(1-q)\mathbb{1} - \lambda^* - \nu^*)$} \\
+        & \hbox{$- (\lambda - \lambda^*)^\intercal \Phi \Phi^\intercal (\lambda - \lambda^*)$} \\
+        & \hbox{$+ (\lambda - \lambda^*)^\intercal y - (\lambda - \lambda^*)^\intercal b + \frac{1}{2} (\lambda - \lambda^*)^\intercal \Phi \Phi^\intercal (\lambda - \lambda^*)$} \\
         \\
-        & +(\lambda-\lambda^*)^\intercal y-(\lambda-\lambda^*)^\intercal b+\frac{1}{2}(\lambda-\lambda^*)^\intercal \Phi\Phi^\intercal(\lambda-\lambda^*)
+        g(\lambda, \lambda^*) = \quad & \hbox{$0 + 0 - \frac{1}{2} (\lambda - \lambda^*)^\intercal \Phi \Phi^\intercal (\lambda - \lambda^*)$} \\
+        & \hbox{$+ (\lambda - \lambda^*)^\intercal y - 0$} \\
         \\
-        \\
-        g(\lambda, \lambda^*)=& 0+0-\frac{1}{2}(\lambda-\lambda^*)^\intercal \Phi\Phi^\intercal(\lambda-\lambda^*)+(\lambda-\lambda^*)^\intercal y-0
-        \\
-        g(\lambda, \lambda^*)=& -\frac{1}{2}(\lambda-\lambda^*)^\intercal \Phi\Phi^\intercal(\lambda-\lambda^*)+(\lambda-\lambda^*)^\intercal y
-    \end{aligned}
+        g(\lambda, \lambda^*) = \quad & \hbox{$- \frac{1}{2} (\lambda - \lambda^*)^\intercal \Phi \Phi^\intercal (\lambda - \lambda^*)$} \\
+        & \hbox{$+ (\lambda - \lambda^*)^\intercal y$}
+    \end{split}
 \end{equation}
-Defining $\alpha=(\lambda-\lambda^*)$ and letting $K$ the kernel matrix, we have that the dual optimisation problem reads as follows
+Defining $\alpha=(\lambda-\lambda^*)$ and letting $K$ be the kernel matrix, we have that the dual optimisation problem reads as follows
 \begin{equation}\label{eq:kqr_min6}
@@
 \end{equation}
-Since our optimisation problem possesses strong duality and it is differentiable in both the objective and the constraint, we have that it must satisfy the Karush Kuhn Tucker conditions (KKT), see section 5.5.3 \cite{boyd2004convex}.
+Since our optimisation problem possesses strong duality and it is differentiable in both the objective and the constraint, we have that it must satisfy the Karush Kuhn Tucker conditions (KKT), see Section 5.5.3 \cite{boyd2004convex}.
 Thanks to the KKT conditions on the primal optimisation problem we have that $f(x_i)=y_i \ \mathrm{for} \ \alpha_i \not \in \{C(q-1), Cq\}$. 
@@
     \begin{aligned}
-    \lambda r_i(x)=&0, \quad i=1,\dots,n
+    \lambda r_i(x)=&0, \quad i=1,\dots,n,
     \\
@@
 \end{equation}
-Using the gradient of the Lagrangian of equation \ref{eq:lagrange_derivatives} we end up with
+Using the gradient of the Lagrangian of Equation \ref{eq:lagrange_derivatives}, we end up with
 \begin{equation}
@@
 \begin{cases}
-    \lambda_i=Cq, \ \lambda_i^*=0 & \implies \lambda_i-\lambda_i^*=Cq, \xi_i\leq 0, \xi^*=0\implies \xi_i-y_i+f_i+b=0 \\
-    \lambda_i=0, \ \lambda_i^*=C(1-q) & \implies \lambda_i-\lambda_i^*=C(q-1), \xi_i= 0, \xi^*\leq 0 \implies \xi_i^*+y_i-f_i-b=0 \\
-    0\leq \lambda_i< Cq, \ 0\leq \lambda_i^*< C(1-q) & \implies \xi_i=0,  \xi^*_i= 0 \implies -y_i+f_i+b=0, \ y_i-f_i-b=0 \\
+    \lambda_i=Cq, \ \lambda_i^*=0 & 
+    \hbox{$\implies \lambda_i-\lambda_i^*=Cq, \xi_i\leq 0, \xi^*=0$}
+    \\
+    &
+    \hbox{$\implies \xi_i-y_i+f_i+b=0$}
+    \\
+    \lambda_i=0, \ \lambda_i^*=C(1-q) & 
+    \hbox{$\implies \lambda_i-\lambda_i^*=C(q-1), \xi_i= 0, \xi^*\leq 0$} 
+    \\
+    & \hbox{$ \implies \xi_i^*+y_i-f_i-b=0$}
+    \\
+    0\leq \lambda_i< Cq, \ 0\leq \lambda_i^*< C(1-q) & 
+    \hbox{$\implies \xi_i=0,  \xi^*_i= 0$}
+    \\
+    & \hbox{$\implies -y_i+f_i+b=0, \ y_i-f_i-b=0$} \\
 \end{cases}
@@
 The dataset used is the Melbourne daily maximum temperatures \citeW{hdrcde}.
-It contains the daily maximum temperatures in Melbourne, Australia, from 1981-1990, excluding leap days, see figure \ref{fig:melbourne_temperature_data}.
+It contains the daily maximum temperatures in Melbourne, Australia, from 1981-1990, excluding leap days, see Figure \ref{fig:melbourne_temperature_data}.
 Due to the bimodality of the data, such dataset is commonly used to give a difficult quantile regression problem \cite{hyndman1996estimating}, thus why we chose it.
@@
 \end{figure}
-Hereafter, the results of the four presented methods on the Melbourne dataset are reported, see figure \ref{fig:melbourne_quantiles_comparison} for a visualisation.
+Hereafter, the results of the four presented methods on the Melbourne dataset are reported, see Figure \ref{fig:melbourne_quantiles_comparison} for a visualisation.
 Notice, the kernel considered here is the Gaussian RBF.
-Hyperparameters have been tuned through cross validation, see \ref{appendix:cross_validation}.
+Hyperparameters have been tuned through cross validation, see Appendix \ref{appendix:cross_validation}.
 \begin{figure}[!h]
@@
 \begin{table}[!h]
-\caption{Pinball loss Melbourne data}
+\caption{Pinball loss of Melbourne data}
 \begin{tabular}{lllll}
@@
 \begin{table}[!h]
-    \caption{Pinball loss quantile-wise Melbourne data}
+    \caption{Pinball loss quantile-wise of Melbourne data}
     \begin{tabular}{lllll}
@@
 \begin{table}[!h]
-\caption{Mean absolute error Melbourne data}
+\caption{Mean absolute error of Melbourne data}
 \begin{tabular}{lllll}
@@
 \end{table}  
-From these tables, we can see that kernel quantile regression outperformed the simple quantile regressor as well as the more complex models like quantile forest and gradient boosting quantile regression for the Melbourne temperatures dataset.
+From these tables, we can see that kernel quantile regression outperformes the simple quantile regressor as well as the more complex models like quantile forest and gradient boosting quantile regression for the Melbourne temperatures dataset.
 Not only kernel quantile regression was the best in terms of total pinball loss but also the best in terms of each quantile pinball loss and mean absolute error.
-Comparison has been carried out on further datasets, yielding similar conclusions as the one of above, to know more see \ref{appendix:quantile_regressor_extensive_comparison}.
+Comparison has been carried out on further datasets, yielding similar conclusions as the one of above, to know more see Appendix \ref{appendix:quantile_regressor_extensive_comparison}.
 
-We conclude this chapter with reporting the same kind of tables \ref{tab:kernel pinball comparison}, \ref{tab:kernel pinball comparison quantile-wise}, \ref{tab:kernel mae comparison} and figure \ref{fig:kernel quantile regressors comparison} comparing various kernel functions.
+We conclude this chapter with reporting the same kind of Tables \ref{tab:kernel pinball comparison}, \ref{tab:kernel pinball comparison quantile-wise}, \ref{tab:kernel mae comparison} and Figure \ref{fig:kernel quantile regressors comparison} comparing various kernel functions.
 
@@
 \begin{table}[!h]
-    \caption{Kernels comparison pinball loss Melbourne data}
+    \caption{Kernels comparison pinball loss for Melbourne data}
     \label{tab:kernel pinball comparison}
@@
 \begin{table}[!h]
-    \caption{Kernels comparison pinball loss quantile-wise Melbourne data}
+    \caption{Kernels comparison pinball loss quantile-wise for Melbourne data}
     \label{tab:kernel pinball comparison quantile-wise}
@@
     \end{subfigure} 
-    \caption{Kernel quantile regressors comparison Melbourne data}
+    \caption{Kernel quantile regressors comparison for Melbourne data}
     \label{fig:kernel quantile regressors comparison} 

### 2024-05-25 thesis/input_files/problemdescription.tex
May 25 2024

@@
 Events like the California electricity crisis are further motivating the choice of the electricity sector as subject of our studies, see \cite{california}.
-Moreover, the process of deregulation lead to an increasing interest in the field of electricity forecasting (EF) within the academic community, see figure \ref{fig:epf_evolution}.
+Moreover, the process of deregulation lead to an increasing interest in the field of electricity forecasting (EF) within the academic community, see Figure \ref{fig:epf_evolution}.
 In addtion, the United Nations have identified the right to access affordable, reliable, sustainable and modern energy as one of their 17 sustainable development goals (SDGs) \citeW{un_sdgs}.

### 2024-05-25 thesis/input_files/qr.tex
May 25 2024

@@
 Quantile regression can be interpreted as an extension of standard regression.
-In this setting, you basically slice the dependent variables into quantiles and then fit a regression for each quantile.
+In this setting, one slices the dependent variables into quantiles and then fit a regression for each quantile.
 With standard regression, we build a model for the conditional mean, conversely, with quantile regression we model the conditional quantile function for any desired quantile.
@@
 \\
-Furthermore, we can extend this framework to non linear quantile regression by choosing a non linear model in place of $x\beta$ in the above equation \ref{eq:linear_quantile_regression_minimisation_problem}.
+Furthermore, we can extend this framework to non linear quantile regression by choosing a non linear model in place of $x\beta$ in the above Equation \ref{eq:linear_quantile_regression_minimisation_problem}.
 % \\

### 2024-05-26 thesis/appendix.tex
corrected all comments

@@
 \item CRIM per capita crime rate by town
-\item ZN proportion of residential land zoned for lots over 25,000 sq.ft.
-\item INDUS proportion of non-retail business acres per town.
+\item ZN proportion of residential land zoned for lots over 25,000 sq.ft.\
+\item INDUS proportion of non-retail business acres per town
 \item CHAS Charles River dummy variable (1 if tract bounds river; 0 otherwise)
@@
 \begin{table}
-\caption{Pinball loss Boston housing data}
+\caption{Pinball loss for Boston housing data}
 \begin{tabular}{lllll}
 \toprule
-    & Linear qr & Gbm qr & Quantile forest & Kernel qr \\
+    & Linear qr & Gbm qr & Quantile forest & KQR \\
 \midrule
-0 & 13.785678 & 11.418540 & 10.587686 & \textbf{10.297572} \\
+CRPS & 1.3785678 & 1.1418540 & 1.0587686 & \textbf{1.0297572} \\
 \bottomrule
@@
 \begin{table}
-    \caption{Pinball loss quantile-wise Boston data}
+    \caption{Pinball loss quantile-wise for Boston data}
 \begin{tabular}{lllll}
 \toprule
-    & Linear qr & Gbm qr & Quantile forest & Kernel qr \\
+    & Linear qr & Gbm qr & Quantile forest & KQR \\
 \midrule
-0.100000 & 0.729749 & 0.771714 & 0.588441 & \textbf{0.578898} \\
-0.200000 & 1.122582 & 1.033442 & 0.932824 & \textbf{0.869145} \\
-0.300000 & 1.479486 & 1.170642 & 1.153765 & \textbf{1.142783} \\
-0.400000 & 1.712577 & 1.436263 & 1.352667 & \textbf{1.331955} \\
-0.500000 & 1.911385 & \textbf{1.344361} & 1.408333 & 1.396300 \\
-0.600000 & 1.989514 & 1.448885 & 1.464902 & \textbf{1.431705} \\
-0.700000 & 1.938362 & 1.508741 & 1.427912 & \textbf{1.382772} \\
-0.800000 & 1.658058 & 1.497901 & 1.275059 & \textbf{1.245288} \\
-0.900000 & 1.243965 & 1.206591 & 0.983784 & \textbf{0.918725} \\
+0.1 & 0.729749 & 0.771714 & 0.588441 & \textbf{0.578898} \\
+0.2 & 1.122582 & 1.033442 & 0.932824 & \textbf{0.869145} \\
+0.3 & 1.479486 & 1.170642 & 1.153765 & \textbf{1.142783} \\
+0.4 & 1.712577 & 1.436263 & 1.352667 & \textbf{1.331955} \\
+0.5 & 1.911385 & \textbf{1.344361} & 1.408333 & 1.396300 \\
+0.6 & 1.989514 & 1.448885 & 1.464902 & \textbf{1.431705} \\
+0.7 & 1.938362 & 1.508741 & 1.427912 & \textbf{1.382772} \\
+0.8 & 1.658058 & 1.497901 & 1.275059 & \textbf{1.245288} \\
+0.9 & 1.243965 & 1.206591 & 0.983784 & \textbf{0.918725} \\
 \bottomrule
@@
 \begin{table}
-\caption{Mean absolute error Boston data}    
+\caption{Mean absolute error for Boston data}    
 \begin{tabular}{lllll}
@@
 \midrule
-0 & 3.826326 & 2.845989 & 2.965490 & \textbf{2.810494} \\
+ & 3.826326 & 2.845989 & 2.965490 & \textbf{2.810494} \\
 \bottomrule
@@
 \begin{table}
-    \caption{Pinball loss Abalone data}
+    \caption{Pinball loss for Abalone data}
 \begin{tabular}{lllll}
     \toprule
-     & Linear qr & Gbm qr & Quantile forest & Kernel qr \\
+     & Linear qr & Gbm qr & Quantile forest & KQR \\
     \midrule
-    0 & 5.613975 & 5.531938 & \textbf{5.212990} & 5.252491 \\
+    CRPS & 0.5613975 & 0.5531938 & \textbf{0.5212990} & 0.5252491 \\
     \bottomrule
@@
 \begin{table}
-    \caption{Pinball loss quantile-wise Abalone data}
+    \caption{Pinball loss quantile-wise for Abalone data}
     \begin{tabular}{lllll}
     \toprule
-     & Linear qr & Gbm qr & Quantile forest & Kernel qr \\
+     & Linear qr & Gbm qr & Quantile forest & KQR \\
     \midrule
-    0.100000 & 0.277903 & 0.290531 & 0.274079 & \textbf{0.269287} \\
-    0.200000 & 0.469361 & 0.488079 & \textbf{0.453110} & 0.457286 \\
-    0.300000 & 0.621961 & 0.625633 & \textbf{0.580766} & 0.596791 \\
-    0.400000 & 0.729757 & 0.715875 & \textbf{0.689904} & 0.691310 \\
-    0.500000 & 0.794695 & 0.766185 & \textbf{0.735945} & 0.740834 \\
-    0.600000 & 0.810691 & 0.785769 & \textbf{0.744928} & 0.746636 \\
-    0.700000 & 0.769587 & 0.730318 & \textbf{0.700287} & 0.710392 \\
-    0.800000 & 0.667776 & 0.656913 & 0.609378 & \textbf{0.608334} \\
-    0.900000 & 0.472244 & 0.472635 & \textbf{0.424593} & 0.431621 \\
+    0.1 & 0.277903 & 0.290531 & 0.274079 & \textbf{0.269287} \\
+    0.2 & 0.469361 & 0.488079 & \textbf{0.453110} & 0.457286 \\
+    0.3 & 0.621961 & 0.625633 & \textbf{0.580766} & 0.596791 \\
+    0.4 & 0.729757 & 0.715875 & \textbf{0.689904} & 0.691310 \\
+    0.5 & 0.794695 & 0.766185 & \textbf{0.735945} & 0.740834 \\
+    0.6 & 0.810691 & 0.785769 & \textbf{0.744928} & 0.746636 \\
+    0.7 & 0.769587 & 0.730318 & \textbf{0.700287} & 0.710392 \\
+    0.8 & 0.667776 & 0.656913 & 0.609378 & \textbf{0.608334} \\
+    0.9 & 0.472244 & 0.472635 & \textbf{0.424593} & 0.431621 \\
     \bottomrule
@@
 \begin{table}
-    \caption{Mean absolute error Abalone data}
+    \caption{Mean absolute error for Abalone data}
     \begin{tabular}{lllll}
     \toprule
-     & Linear qr & Gbm qr & Quantile forest & Kernel qr \\
+     & Linear qr & Gbm qr & Quantile forest & KQR \\
     \midrule
-    0 & 1.627627 & 1.574179 & 1.499522 & \textbf{1.498583} \\
+     & 1.627627 & 1.574179 & 1.499522 & \textbf{1.498583} \\
     \bottomrule
@@
 \subsection{Vehicle dataset}
-This data contains info about used cars \href{https://www.kaggle.com/datasets/nehalbirla/vehicle-dataset-from-cardekho}{https://www.kaggle.com/datasets/nehalbirla/vehicle-dataset-from-cardekho}, the predictors are:
+This data contains info about used cars \href{https://www.kaggle.com/datasets/nehalbirla/vehicle-dataset-from-cardekho}{https://www.kaggle.com/datasets/ne-halbirla/vehicle-dataset-from-cardekho}, the predictors are:
 \begin{itemize}
@@
 \begin{table}
-\caption{Pinball loss Vehicle data}
+\caption{Pinball loss for Vehicle data}
 \begin{tabular}{lllll}
     \toprule
-     & Linear qr & Gbm qr & Quantile forest & Kernel qr \\
+     & Linear qr & Gbm qr & Quantile forest & KQR \\
     \midrule
-    0 & 4.054449 & 2.289554 & 2.844410 & \textbf{2.204343} \\
+    CRPS & 0.4054449 & 0.2289554 & 0.2844410 & \textbf{0.2204343} \\
     \bottomrule
@@
 \begin{table}
-    \caption{Pinball loss quantile-wise Vehicle data}
+    \caption{Pinball loss quantile-wise for Vehicle data}
     \begin{tabular}{lllll}
     \toprule
-     & Linear qr & Gbm qr & Quantile forest & Kernel qr \\
+     & Linear qr & Gbm qr & Quantile forest & KQR \\
     \midrule
-    0.100000 & 0.254649 & \textbf{0.139849} & 0.170489 & 0.182490 \\
-    0.200000 & 0.403772 & 0.236339 & 0.285875 & \textbf{0.223165} \\
-    0.300000 & 0.548820 & 0.244086 & 0.357375 & \textbf{0.242963} \\
-    0.400000 & 0.576918 & 0.263169 & 0.389305 & \textbf{0.262835} \\
-    0.500000 & 0.554367 & 0.306123 & 0.410738 & \textbf{0.295878} \\
-    0.600000 & 0.563046 & 0.335363 & 0.398125 & \textbf{0.306062} \\
-    0.700000 & 0.516019 & 0.287490 & 0.326572 & \textbf{0.283716} \\
-    0.800000 & 0.407742 & 0.261882 & 0.313698 & \textbf{0.235793} \\
-    0.900000 & 0.229115 & 0.215253 & 0.192233 & \textbf{0.171440} \\
+    0.1 & 0.254649 & \textbf{0.139849} & 0.170489 & 0.182490 \\
+    0.2 & 0.403772 & 0.236339 & 0.285875 & \textbf{0.223165} \\
+    0.3 & 0.548820 & 0.244086 & 0.357375 & \textbf{0.242963} \\
+    0.4 & 0.576918 & 0.263169 & 0.389305 & \textbf{0.262835} \\
+    0.5 & 0.554367 & 0.306123 & 0.410738 & \textbf{0.295878} \\
+    0.6 & 0.563046 & 0.335363 & 0.398125 & \textbf{0.306062} \\
+    0.7 & 0.516019 & 0.287490 & 0.326572 & \textbf{0.283716} \\
+    0.8 & 0.407742 & 0.261882 & 0.313698 & \textbf{0.235793} \\
+    0.9 & 0.229115 & 0.215253 & 0.192233 & \textbf{0.171440} \\
     \bottomrule
@@
 \begin{table}
-    \caption{Mean absolute error Vehicle data}
+    \caption{Mean absolute error for Vehicle data}
     \begin{tabular}{lllll}
     \toprule
-     & Linear qr & Gbm qr & Quantile forest & Kernel qr \\
+     & Linear qr & Gbm qr & Quantile forest & KQR \\
     \midrule
-    0 & 1.117714 & 0.606971 & 0.752197 & \textbf{0.594292} \\
+     & 1.117714 & 0.606971 & 0.752197 & \textbf{0.594292} \\
     \bottomrule
@@
 \end{equation}
-That is $Err$ is the average over everything that is random: X,Y and the training set $\Tau$ used to learn $\hat{f}$.
+that is, $Err$ is the average over everything that is random: X,Y and the training set $\Tau$ used to learn $\hat{f}$.
 Hence, our interest lies in estimating the $Err$ quantity in order to guide model selection.
 \subsection{K-fold cross validation}
-K-fold cross validation splits the data into K roughly equally sized parts, then K models are trained.
-For k from 1 to K we train the kth model on the whole dataset except the kth partition.
+K-fold cross validation splits the data into $K$ roughly equally sized parts, then $K$ models are trained.
+For $k$ from 1 to $K$ we train the kth model on the whole dataset except the kth partition.
 Next,  the prediction error of the kth fitted model is computed.
-Finally averaging all the k prediction erros we obtain an estimate for the expected test error.
+Finally averaging all the $k$ prediction erros we obtain an estimate for the expected test error.
 We select the model which performs best in terms of expected prediction error.
 \\
-Usually, K is set equal to 5 or 10. Leave-one-out cross validation is the case when K is set equal to the size of the data.
+Usually, $K$ is set equal to 5 or 10. Leave-one-out cross validation is the case when $K$ is set equal to the size of the data.
 % REMEMBER: Parameters allow the model to learn the rules from the data while hyperparameters control how the model is training.
@@
 Essentially, gridsearch carries out hyperparameter tuning by performing a search over a predefined hyperparemeters grid.
-During its search, it tries all possible combination and evaluates the different models %(different hyperparameters define different models)
+During its search, it tries all possible combinations and evaluates the different models %(different hyperparameters define different models)
 using cross validation. For example, suppose that the grid contains 100 possible candidates and that we are doing 5-fold cross-validation, then the gridsearch algorithm will carry out 500 iterations.
-Therefore, we get an estimate of prediction error for each considered model and this guides us in hyperparameters (model) selection.
+Therefore, we get an estimate of prediction error for each considered model and this guides us in the hyperparameters selection.
 
@@
 \subsection{Cross validation for time series data}
-When data points are dependent on preceding values, we cannot use standard K-fold cross validation.
-The rationale is that K-fold will randomize the order of the data, thus it might happen to use future data to predict the past; we want to avoid such behaviour in a time series setting.
+When data points are dependent on preceding values, we cannot use standard $K$-fold cross validation.
+The rationale is that $K$-fold will randomize the order of the data, thus it might happen to use future data to predict the past; we want to avoid such behaviour in a time series setting.
 When carrying out out any kind of cross validation, we must keep consistency in the way we evaluate our predictors during model selection and in the way we perform evaluation of the test data.
-Hence, for time series crossvalidation we have the timeseries split procedure, see figure \ref{fig:crossvalidationtimeseries} for a visualisation; the blue observations make up the training sets while the orange observations form the test sets.
+Hence, for time series crossvalidation we have the timeseries split procedure, see Figure \ref{fig:crossvalidationtimeseries} for a visualisation.
+The blue observations make up the training sets while the orange observations form the test sets.
 \begin{figure}
@@
 This comes from the fact that at each iteration we push forward the origin of our forecast.
-The same concept applies for multi step ahead forecasting, see picture \ref{fig:crossvalidationtimeseries2}.
-That is in predicting $\hat{y}_{N+m}$ we use as inputs $y_1, y_2, \dots, \hat{y}_{N},\hat{y}_{N+1},\dots, \hat{y}_{N+m-1}$; where $\hat{y}_{N},\hat{y}_{N+1},\dots, \hat{y}_{N+m-1}$ are one step ahead forecasts.
+The same concept applies for multi step ahead forecasting, see Figure \ref{fig:crossvalidationtimeseries2}.
+In predicting $\hat{y}_{N+m}$ we use as inputs $y_1, y_2, \dots, \hat{y}_{N},\hat{y}_{N+1},\dots, \hat{y}_{N+m-1}$, where $\hat{y}_{N},\hat{y}_{N+1},\dots, \hat{y}_{N+m-1}$ are one step ahead forecasts.
 \begin{figure}
@@
 \begin{itemize}
-    \item Standard scaler= it computes the standard score z of a sample x
+    \item Standard scaler= it computes the standard score $z$ of a sample $x$
     \\
-    z=$\frac{x-\mu}{\sigma}$.
-    \item MinMax scaler= it maps every data sample to the range 0, 1.
+    $z=\frac{x-\mu}{\sigma}$.
+    \item MinMax scaler= it maps every data sample to the range $[0,1]$.
     \\
@@
 The idea is to compare the results of modelling the data with K-nearest neighbors on the unscaled data against the scaled data.
-The considered data is the wine recognition dataset \href{https://archive.ics.uci.edu/dataset/109/wine}{https://archive.ics.uci.edu/dataset/109/wine}.
+The considered data is the wine recognition dataset \href{https://archive.ics.uci.edu/dataset/109/wine}{https://archive.ics.uci.edu/da-taset/109/wine}.
 The goal for this dataset is recognising from whose cultivator the wine comes based on two features with a completely different scale.
 The first feature has values in the [0,1000] range while the second feature has values contained in [1,10].
-The unscaled and scaled version are compared in figure \ref{fig:feature_scaler_example}.
+The unscaled and scaled version are compared in Figure \ref{fig:feature_scaler_example}.
 \begin{figure}[!h]
     \includegraphics[width=\textwidth]{images/feature_scaler_example.png}
-    \caption{importance of feature scaling}
+    \caption{Importance of feature scaling}
     \label{fig:feature_scaler_example}
 \end{figure}
-What we can conclude from the image is that, the model trained on scaled data is greatly better than the other.
+What we can conclude from the image is that, the model trained on scaled data is much better than the other.
 On the left, we can see that distances between categories are impacted solely by the larger scale feature.
@@
 \subsection{Data compression}
-When the number of points n is large, kernel methods suffer from high computational costs. 
-Using kernel methods, we have that the storage cost is of the order of $O(n^2)$ while the computational cost for finding the solution is of the order $O(n^3)$.
+When the number of points $n$ is large, kernel methods suffer from high computational costs. 
+Using kernel methods, we have that the storage cost is of the order $O(n^2)$ while the computational cost for finding the solution is of the order $O(n^3)$.
 A significant speed up can be obtained thanks to low rank approximation.
@@
 The Nyström approximation involves storing a submatrix of the whole kernel matrix. Thus, storage and computational costs are reduced to $O(nm)$ and $O(nm^2)$ respectively.
-Nyström works by selecting $m<n$ points, callled representative points; it approximates $K$ as
+Nyström works by selecting $m<n$ points, called representative points; it approximates $K$ as
 \begin{equation}
@@
 \subsubsection{Pivoted Cholesky decomposition}
-Pivoted Cholesky approximate the Cholesky decomposition of a matrix.
-Since kernel matrices are positive definite, they can decomposed in terms of the Cholesky decomposition.
+Pivoted Cholesky approximates the Cholesky decomposition of a matrix.
+Since kernel matrices are positive definite, they can be decomposed in terms of the Cholesky decomposition.
 Hence, we have that pivoted Cholesky can be used to approximate the full kernel matrix.
@@
 \begin{itemize}
-    \item query/: folder containing Scopus data and scripts to generate bibliometric survey plots in section \ref{literature_review}.
+    \item query/: folder containing Scopus data and scripts to generate bibliometric survey plots in Section \ref{literature_review}.
     \item thesis/: folder containing tex files for thesis.

### 2024-05-26 thesis/input_files/analysis.tex
corrected all comments

@@
 % \\
-Building on the theory introduced in section \ref{ch:point} and \ref{ch:prob}, this section covers the experiments carried out and the results obtained respectively.
+Building on the theory introduced in Sections \ref{ch:point} and \ref{ch:prob}, this section covers the experiments carried out and the results obtained respectively.
 
@@
     \item Day of the week
-    \item Is holiday: a binary variable for holidays, where holiday$=1$ and working day$=0$
+    \item Is holiday: a binary variable for holidays, where holiday=1 and working day=0
     \item Weather temperature
 \end{itemize}
-Methods will be compared by means of the RMSE, MAE and MAPE scores, see section \ref{metrics}.
+Methods will be compared by means of the RMSE, MAE and MAPE scores, see Section \ref{metrics}.
 \subsection{Multiple linear regression}
-To get started, standard multiple linear regression has been applied, see figure \ref{fig:mlr_price} for a visualisation. 
+To get started, standard multiple linear regression has been applied, see Figure \ref{fig:mlr_price} for a visualisation. 
 \begin{figure}
@@
 Its name stands for trigonometric seasonality, Box-Cox transformation, ARMA errors, trend and seasonal components.
-Tbats forecast is visualized in figure \ref{fig:tbats_price}, meanwhile its RMSE is 15.08.
+Tbats forecast is visualized in Figure \ref{fig:tbats_price}, meanwhile its RMSE is 15.08.
 \begin{figure}[!h]
@@
 Following, the prophet model has been considered. 
-We get started by considering the base implementation. In such setting, prophet takes in as input the time series object and learns its data generating process.
-This method achieves a RMSE of 23.96, its prediction is visualized in figure \ref{fig:prophet_price_1}.
+We get started by considering the base implementation. In such setting, prophet takes as input the time series object and learns its data generating process.
+This method achieves a RMSE of 23.96, its prediction is visualized in Figure \ref{fig:prophet_price_1}.
 \begin{figure}[!h]
@@
 \end{figure}
-What can be seen is that, prophet models correctly the average trend but does not model peaks and lows precisely. 
+What can be seen is that prophet correctly models the average trend but does not model peaks and lows precisely. 
 Next, a more complex model was trained.
@@
 Doing so, RMSE went down to 10.29.
-Forecast is visualised in figure \ref{fig:prophet_price2}.
+Forecast is visualised in Figure \ref{fig:prophet_price2}.
 % , moreover figure \ref{fig:prophet_price2.1} and figure \ref{fig:prophet_price2.2} break down the trend and the seasonalities respectively.
@@
 In doing so, we cross validated the number of neighbours to get the best model instance.
-The forecast is visualised in figure \ref{fig:knn_price}, RMSE is 12.54.
+The forecast is visualised in Figure \ref{fig:knn_price}, RMSE is 12.54.
 \begin{figure}[!h]
@@
 \subsection{Support vector regression}
-Coming up we have support vector regression.
-In applying this model we used gridsearch crossvalidation to search for the best regularisation parameter C.
-Support vector regression achieves an RMSE of 23.24, for the prediction see figure \ref{fig:svr_price}.
+In applying support vector regression we used gridsearch crossvalidation to search for the best regularisation parameter $C$.
+Support vector regression achieves an RMSE of 23.24, for the prediction see Figure \ref{fig:svr_price}.
 \begin{figure}[!h]
@@
 \end{itemize}
-LSTM forecast is reported in figure \ref{fig:lstm_price}, it achieves a RMSE of 9.1782
+LSTM forecast is reported in Figure \ref{fig:lstm_price}, it achieves a RMSE of 9.1782
 \begin{figure}[!h]
@@
 Cross validation was carried jointly over the RBF kernel bandwith and the regularisation constant.
-The RMSE achieved is 9.86, figure \ref{fig:krnridge_price} reports the prediction.
+The RMSE achieved is 9.86, Figure \ref{fig:krnridge_price} reports the prediction.
 \begin{figure}[!h]
@@
 \end{figure}
-We can observe that, kernel ridge regression models accurately the electricity time series.
+We can observe that, kernel ridge regression accurately models the electricity time series.
 
@@
 The last model we considered in this setting was the kernel support vector regression.
-RMSE achieved with this method is 9.10, forecast is reported in figure \ref{fig:krnsvr_price}.
+RMSE achieved with this method is 9.10, forecast is reported in Figure \ref{fig:krnsvr_price}.
 
@@
 \end{figure}
-We can observe that kernel support vector regression is one of the best performing techniques between the one considered.
+We can observe that kernel support vector regression is one of the best performing techniques between the ones considered.
 
@@
 
-The MAE scores are contained in table \ref{tab:point_MAE}. Similarly to above, we can conclude that kernel methods stand out for achieving also the lowest MAE score; with KrnSVR being the top one.
+The MAE scores are contained in Table \ref{tab:point_MAE}. Similarly to above, we can conclude that kernel methods stand out for achieving also the lowest MAE score; with KrnSVR being the top one.
 \begin{table}[!ht]
@@
 
-Finally table \ref{tab:point_MAPE} reports the MAPE score for each method. Its results are in line with the other tables, hence suggesting the goodness of kernel methods.
+Finally Table \ref{tab:point_MAPE} reports the MAPE score for each method. Its results are in line with the other tables, hence suggesting the goodness of kernel methods.
 % What can be concluded is that kernel regressors are also the best in terms of MAPE score; with KrnSVR being the top one.
@@
 Using the load data from \href{https://www.energy-charts.info/index.html?l=en&c=DE}{energy charts} for Germany and Switzerland, we proceeded to compare between quantile regression algorithms.
-In this study, we trained our models on the entire 2021 data and then we tested them and measured their score on the 2022 data.
+In this study, we trained our models on the entire 2021 data and then tested them and measured their score on the 2022 data.
 Notice, for kernel quantile regression (KQR), the kernel of choice in this setting is the absolute Laplacian.
@@
 The 90\% confidence interval bands are plotted in green.
-Lower and upper red lines denote the 95\% and 5\% quantile forecast respectively.}
+Lower and upper red lines denote the 95\% and 5\% quantile forecast respectively}
         \label{fig:DE_load_CI}
@@
 The 90\% confidence interval bands are plotted in green.
-Lower and upper red lines denote the 95\% and 5\% quantile forecast respectively.}
+Lower and upper red lines denote the 95\% and 5\% quantile forecast respectively}
         \label{fig:CH_load_CI}
@@
 \end{itemize}
-Table \ref{tab:energy_chart_de_additional_predictors} reports the model scores on the German data while table \ref{tab:energy_chart_ch_additional_predictors} is the one corresponding to the Suisse data.
+Table \ref{tab:energy_chart_de_additional_predictors} reports the model scores on the German data while Table \ref{tab:energy_chart_ch_additional_predictors} is the one corresponding to the Suisse data.
 
@@
 \end{itemize}
-Table \ref{tab:secures_met_de} shows results for Germany, table \ref{tab:secures_met_ch} is the one for Switzerland and table \ref{tab:secures_met_at} corresponds to Austria.
+Table \ref{tab:secures_met_de} shows results for Germany, Table \ref{tab:secures_met_ch} is the one for Switzerland and Table \ref{tab:secures_met_at} corresponds to Austria.
 
@@
 \end{table}
-From our tables, it can be concluded that the absolute Laplacian kernel is the one fitting best to the characteristics of loda data.
+From our tables, it can be concluded that the absolute Laplacian kernel is the one fitting best to the characteristics of load data.
 Additionally, we can conclude that our results are in line with the ones published in \cite{he2017short}.
@@
 Their paper showed evidence of the superiority of the Gaussian RBF among their set of kernel functions.
-Analogously, we can see from tables \ref{tab:secures_met_de}, \ref{tab:secures_met_ch} and \ref{tab:secures_met_at} that the Gaussian RBF is the second best in terms of pinball loss.
+Analogously, we can see from Tables \ref{tab:secures_met_de}, \ref{tab:secures_met_ch} and \ref{tab:secures_met_at} that the Gaussian RBF is the second best in terms of pinball loss.
 
@@
 The following subsection compares our kernel quantile regression against the price and load track entries of the GEFCom2014 competition.
-Error measure of the competition was the pinball loss, see section \ref{metrics}, averaged over the $99$ quantiles, $q \in \{i/100\}_{i=1}^{n=99}$.
+Error measure of the competition was the pinball loss, see Section \ref{metrics}, averaged over the $99$ quantiles, $q \in \{i/100\}_{i=1}^{n=99}$.
 Backed by the conclusion from the previous subsection we selected the absolute Laplacian and the Gaussian RBF as kernels in our application of KQR.
@@
 
-Our results for this track are reported in table \ref{tab:pinball loss gefcom2014 price data}.
+Our results for this track are reported in Table \ref{tab:pinball loss gefcom2014 price data}.
 In this track, the top entries came from the teams: Tololo, Team Poland, GMD, and C3 Green Team; for a breakdown of each method's attributes, see \cite[Table 8]{hong2016probabilistic}.
@@
 The $90\%$ confidence interval bands are plotted in green.
-Lower and upper red lines denote the $95\%$ and $5\%$ quantile forecast respectively.}
+Lower and upper red lines denote the $95\%$ and $5\%$ quantile forecast respectively}
     % \Description{}
@@
       Team name\textbackslash Task number                       & 1                               & 2                                  & 3                               & 4                              & 5                              & 6       & 7                               & 8       & 9       & 10                             & 11                             & 12               \\
-  KQR laplacian
+  KQR Laplacian
   &
@@
   \\
-  KQR gaussian rbf
+  KQR Gaussian RBF
   &
@@
 
-What can be concluded from table \ref{tab:pinball loss gefcom2014 price data} is that KQR performed consistently among the top algorithms.
-Furthermore, It ranked first in two out of the thirteen tasks of the price track.
+What can be concluded from Table \ref{tab:pinball loss gefcom2014 price data} is that KQR performs consistently among the top algorithms.
+Furthermore, It ranks first in two out of the thirteen tasks of the price track.
 \subsubsection{GEFCom2014 load track} 
@@
     \item Day of week: an ordinal categorical variable for the day of the week
-    \item Is holiday: a binary variable for holidays, where holiday$=1$ and working day$=0$
-    \item w avg: average of weather temperatures across all the $25$ stations
+    \item Is holiday: a binary variable for holidays, where holiday=1 and working day=0
+    \item w avg: average of weather temperatures across 25 stations
 \end{itemize}
@@
 The 90\% confidence interval bands are plotted in green.
-Lower and upper red lines denote the 95\% and 5\% quantile forecast respectively.}
+Lower and upper red lines denote the 95\% and 5\% quantile forecast respectively}
     \label{fig:load_task_9}
@@
 
-  KQR laplacian
+  KQR Laplacian
   & 
@@
   \\
-  KQR gaussian rbf
+  KQR Gaussian RBF
   & 
@@
 This thesis was concerned with electricity forecasting, in particular we focused on the probabilistic framework.
-Probabilistic forecasting is still a relatively new research area, lately research efforts are shifting towards this setting.
+Probabilistic forecasting is still a relatively new research area.
+Lately, research efforts are shifting towards this setting.
 This is due to the fact that probabilistic forecasts are more informative than point ones and better suited for the current and future electricity landscape; think for example about the higher uncertainty resulting from market liberalisation or renewables integration requirements.
 
-Among the contributions of this thesis work we can list
+Among the contributions of this thesis work, we can list
 % list contributions:
@@
 \begin{itemize}
-    \item Study of kernel methods in electricity forecasting both in the point and probabilistic framework.
-    \item Provided Python open source implementation for kernel quantile regression compatible with the sklearn API.
+    \item We have studied kernel methods for electricity forecasting both in the point and probabilistic framework.
+    \item We provided a Python open source implementation for kernel quantile regression compatible with the sklearn API.
 The code has been packaged and uploaded to the Python Package Index (PyPI) with the name \href{https://pypi.org/project/kernel-quantile-regression/#2}{kernel-quantile-regression}.
 The \href{https://github.com/luca-pernigo/kernel_quantile_regression}{github repo} hosting the source code includes also the script implementing the experiments along with the cleaned datasets; this contribution is intended to forster reproducibility in research.
-    \item Achieved superior performance of kernel quantile regression compared to standard quantile regressor algorithms. 
-    \item Created and made available datasets suitable for algorithms benchmarking considering data from the DACH region (Germany, Switzerland and Austria).
+    \item We achieved superior performance of kernel quantile regression compared to standard quantile regressor algorithms. 
+    \item We created and made available datasets suitable for algorithms benchmarking considering data from the DACH region (Germany, Switzerland and Austria).
 The format of these data takes inspiration from the popular GEFCom competitions.
     \item Kernel quantile regression extensive comparison between kernel function types.
-    \item Compared kernel quantile regression against state of the art probabilistic algorithms in the literature by means of the GEFCom2014 competition.
+    \item We compared kernel quantile regression against state of the art probabilistic algorithms in the literature by means of the GEFCom2014 competition.
 \end{itemize}
@@
 We observe that KQR shows favorably and can forecast both the short and medium-term horizon.
-To finish, we proceed with summarising the conclusions we drew from our experimental analysis.
+To finish, we proceed summarising the conclusions we drew from our experimental analysis.
 Kernel methods showed good performance both in the context of point and probabilistic forecasting.

### 2024-05-26 thesis/input_files/data_exploratory.tex
corrected all comments

@@
     \includegraphics[width=\textwidth]{images/gefcom_zonal_price_vs_total_load.png}
-    \caption{Zonal price versus total load GEFCom2014}
+    \caption{Zonal price versus total load of GEFCom2014}
     \label{fig:gefcom_zonal_price_vs_total_load}
@@
 We applied the same kind of EDA to the GEFCom load data.
-Figure \ref{fig:gefcom_load_historical} depicts the historical load while figure \ref{fig:gefcom_w_avg_historical} reports the historical weather temperature.
-Finally, figure \ref{fig:gefcom_load_vs_w_avg} shows the scatter plot between the load and the weather.
+Figure \ref{fig:gefcom_load_historical} depicts the historical load while Figure \ref{fig:gefcom_w_avg_historical} reports the historical weather temperature.
+Finally, Figure \ref{fig:gefcom_load_vs_w_avg} shows the scatter plot between the load and the weather.
 These plots motivate the selection of weather temperature as one of the predictors for the load.
@@
     \includegraphics[width=\textwidth]{images/gefcom_load_historical.png}
-    \caption{Historical load GEFCom2014}
+    \caption{Historical load of GEFCom2014}
     \label{fig:gefcom_load_historical}
@@
     \includegraphics[width=\textwidth]{images/gefcom_w_avg_historical.png}
-    \caption{Historical weather GEFCom2014}
+    \caption{Historical weather of GEFCom2014}
     \label{fig:gefcom_w_avg_historical}
@@
     \includegraphics[width=\textwidth]{images/gefcom_load_vs_w_avg.png}
-    \caption{Load versus weather GEFCom2014}
+    \caption{Load versus weather of GEFCom2014}
     \label{fig:gefcom_load_vs_w_avg}
@@
 Firstly, we drop all Nans and zeros.
-Notice, the provided data for the challenge has a timestamp column not coherent to any time format.
+The provided data for the challenge has a timestamp column not coherent to any time format.
 Thus, we first pass the data through a method extracting time information: hour, day, month and year.
@@
 This independent variable is named w avg.
-The same thing applies in the cleaning of the test data with the addition of an if clause for handling the different naming convention and formatting used in the last track of the challenge.
-The Python ETL pipeline has been then automated to loop every task folder through Bash scripting.
+The same thing applies in the cleaning of the test data with the addition of an if clause for handling the different naming conventions and formatting used in the last track of the challenge.
+The Python ETL pipeline has since been automated to loop every task folder through Bash scripting.
 
@@
 Next, we identified various data sources for Germany, Switzerland and Austria in order to further
-investigate the techniques studied in this thesis work.
+investigate the techniques studied in this thesis.
 Taking inspiration from the GEFCom2014 format, we firstly created a load track for Germany and Switzerland. 
@@
 It provides data for electricity power, electricity potential capacity, energy generation, market prices, environmental measurements and also scenario simulations.
-Notice, this data is retrieved from the European network of transmission system operators for electricity \href{https://www.entsoe.eu/}{(ENTSOE)} and made available thanks to the \href{https://ise.fraunhofer.de/en.html}{Fraunhofer Institute for Solar Energy Systems ISE}.
+This data is retrieved from the European network of transmission system operators for electricity \href{https://www.entsoe.eu/}{(ENTSOE)} and made available thanks to the \href{https://ise.fraunhofer.de/en.html}{Fraunhofer Institute for Solar Energy Systems ISE}.
 \\
-In this experiment, the goal is comparing kernel quantile regression to the other state of the art quantile regressors, see section \ref{ch:prob}.
+In this experiment, the goal is comparing kernel quantile regression to the other state of the art quantile regressors, see Section \ref{ch:prob}.
 The target dependent variable is the national load while as predictors we took
@@
 In the second experiment we compared various kernels in the context of load forecasting.
-For this study we considered the data of \href{https://zenodo.org/records/7907883}{SECURES-Met}.
+For this study, we considered the data of \href{https://zenodo.org/records/7907883}{SECURES-Met}.
 This data consists of various metereological attributes relevant for modelling electricity demand \cite{Formayer2023}.
 The dataset is made up of historical data until end of 2020, while from 2021 onwards the data consists of forecasts made by a metereological model developed by the European Centre for Medium-Range Weather Forecasts (ECMWF).
-Notice, the forecasted data is splitted in two files according to the scenarios RCP4.5 and RCP8.5.
-In our experiments we considered the former.
+The forecasted data is split in two files according to the scenarios RCP4.5 and RCP8.5.
+In our experiments, we considered the former.
 In this setting the independent variables available are 
@@
 From energy charts we downloaded the historical load along with the weather temperatures and wind speeds forecasts.
-Notice, the data request to the environmental variables at hourly frequency for a whole year is broken.
+The data request to the environmental variables at hourly frequency for a whole year is broken.
 Therefore, we downloaded each month separately and then transformed them into a single csv for the whole year.
@@
 For what concerns the data from SECURES-Met we first identified the data interesting to us.
-Next, we formatted all the csv to a coherent format, gave columns meaningful names, sliced for the years 2020 (train) and 2021 (test) and finally we merged all single csv\textsubscript{s} together.
+Next, we formatted all the csv to a coherent format, gave columns meaningful names, sliced for the years 2020 (train) and 2021 (test) and finally merged all single csv\textsubscript{s} together.
 
@@
 \subsubsection{EDA}
-Following are reported the EDA plots alongside with insights that can be extracted from them.
+Following we report the EDA plots alongside with insights that can be extracted from them.
 We choose to focus the following analysis on the Swiss data only, given that the plots for the German data are pretty similar.
-To get started we visualized the historical load data in figure \ref{fig:CH_historical_load_2021}.
+To get started we visualise the historical load data in Figure \ref{fig:CH_historical_load_2021}.
 As expected, electricity load exhibits a seasonal pattern with highs in the winter and lows during the summer season.
@@
 Next, we built scatter plots of the independent continous variables against the load as means of visualizing correlation between them.
-Figure \ref{fig:CH_load_vs_temperature_2021} plots load against temperatures while figure \ref{fig:CH_load_vs_wind_speed_2021} depicts load against wind speed.
-From these plots we can clearly see there exists a relationship that ties wind speeds and temperatures to load.
+Figure \ref{fig:CH_load_vs_temperature_2021} plots load against temperatures while Figure \ref{fig:CH_load_vs_wind_speed_2021} depicts load against wind speed.
+From these plots, we can clearly see there exists a relationship that ties wind speeds and temperatures to load.
 
@@
     \includegraphics[width=\textwidth]{images/CH_load_vs_temperature_2021.png}
-    \caption{Load vs temperature Switzerland (2021)}
+    \caption{Load versus temperature for Switzerland (2021)}
     \label{fig:CH_load_vs_temperature_2021}
@@
     \includegraphics[width=\textwidth]{images/CH_load_vs_wind_speed_2021.png}
-    \caption{Load vs wind speed Switzerland (2021)}
+    \caption{Load versus wind speed for Switzerland (2021)}
     \label{fig:CH_load_vs_wind_speed_2021}
@@
 
-Following, we used box plots in order to visualize the relations between categorical independent variables and the dependent one.
-Such visualisation are reported in figure \ref{fig:CH_day_of_week_boxplot_2021} and figure \ref{fig:CH_is_holiday_boxplot_2021}.
+Following, we use box plots in order to visualize the relations between categorical independent variables and the dependent one.
+Such visualisation are reported in Figure \ref{fig:CH_day_of_week_boxplot_2021} and Figure \ref{fig:CH_is_holiday_boxplot_2021}.
 We can see that weekdays and workdays involve a higher load and possess a wider range.
@@
     \includegraphics[width=\textwidth]{images/CH_day_of_week_boxplot_2021.png}
-    \caption{Load versus day of week boxplot Switzerland (2021)}
+    \caption{Load versus day of week boxplot for Switzerland (2021)}
     \label{fig:CH_day_of_week_boxplot_2021}
@@
     \includegraphics[width=\textwidth]{images/CH_is_holiday_boxplot_2021.png}
-    \caption{Load versus day type boxplot Switzerland (2021)}
+    \caption{Load versus day type boxplot for Switzerland (2021)}
     \label{fig:CH_is_holiday_boxplot_2021}
@@
 
-We conclude this EDA by reporting the heatmap of independent variables alongside with the load in figure \ref{fig:CH_heatmap_2021}.
+We conclude this EDA by reporting the heatmap of independent variables alongside with the load in Figure \ref{fig:CH_heatmap_2021}.
 Temperature, wind speed and day of week show a meaningful correlation.

### 2024-05-26 thesis/input_files/implementation.tex
corrected all comments

@@
 \\
-For the list of Python packages needed see the requirement.txt file on \url{https://github.com/luca-pernigo/ThesisKernelMethods}\label{github_repo}.
+For the list of Python packages needed, see the requirement.txt file on \url{https://github.com/luca-pernigo/ThesisKernelMethods}\label{github_repo}.
 % - indicate computer specifics
@@
 \subsection{Trigonometric seasonality Box-Cox transformation ARMA errors trend and seasonal components}
-Tbats implementation is available at \url{https://github.com/intive-DataScience/tbatsx}.
-In our application we specified as hyperparameters the length of seasons; 24 for the daily seasonality and 168 for the weekly seasonality.
+The Tbats implementation is available at \href{https://github.com/intive-DataScience/tbats}{https://github.com/intive-Data-Science/tbatsx}.
+In our application we specified as hyperparameters the length of seasons, that is, 24 for the daily seasonality and 168 for the weekly seasonality.
 
@@
 Therefore, the KQR class is derived from the scikit-learn BaseEstimator and the mixin class RegressorMixin.
-Our KQR class is initialized by providing a quantile, the regularization term C, the kernel family and its corresponding hyperparameters.
+Our KQR class is initialized by providing a quantile, the regularisation term $C$, the kernel family and its corresponding hyperparameters.
 \\
@@
 First this library assumes the quadratic term of the optimisation problem to be multiplied by the 0.5 factor, thus we just have to provide the $Q$ matrix with no 0.5 in front.
-Secondly, in order to specify multiple inequalities we have to stack them and provide them as a unique matrix.
+Secondly, in order to specify multiple inequalities we have to stack them and provide them as a single matrix.
 \\
@@
 \\
-This estimator is compatible with useful scikit-learn in built methods like gridsearch, crossvalidation and scoring rules.
-Moreover, this code is compatible with sklearn in built kernel functions \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.metrics.pairwise}{sklearn.metrics.pairwise} and \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.gaussian_process.kernels}{sklearn.gaussian\_process.kernels}, functions supported by our kernel quantile regression include: Gaussian RBF, absolute Laplacian, Matern 3/2, Matern 5/2, Linear, Cosine, Sigmoid, Periodic, Polynomial and custom composition of kernels.
+This estimator is compatible with in built scikit-learn methods like gridsearch, crossvalidation and scoring rules.
+Moreover, this code is compatible with sklearn in built kernel functions \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.metrics.pairwise}{sklearn.metrics.pairwise} and \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.gaussian_process.kernels}{sklearn.gaussian\_process.kernels}.
+Functions supported by our kernel quantile regression include: Gaussian RBF, absolute Laplacian, Matern 3/2, Matern 5/2, Linear, Cosine, Sigmoid, Periodic, Polynomial and custom composition of kernels.
 % NOTICE, these kernel implementations differ in the form of passed parameters in the former we pass in the numerator gamma while in the latter we pass l, which stands for length scale in the denominator.

### 2024-05-26 thesis/input_files/literaturereview.tex
corrected all comments

@@
 In 1941, Kolmogorov \cite{kolmogorov} carried out stuides on representing kernels in linear spaces.
-In 1950, Aronszajn \cite{aronszajn} published the first work on reproducing kernel hilbert spaces (RKHS); developing the general method for representing kernels in linear spaces.
+In 1950, Aronszajn \cite{aronszajn} published the first work on reproducing kernel Hilbert spaces (RKHS); developing the general method for representing kernels in linear spaces.
 In 1964, Aizerman \cite{aizerman} further improved the theory of RKHS.

### 2024-05-26 experiments/plots/box_ci.py

@@ -5,8 +5,8 @@ import numpy as np
 import pandas as pd
 
 
-df=pd.read_csv("Data/Load/Task 9/L9-model_prediction_laplacian.csv")
-df_test=pd.read_csv(f"Data/Load/Task 9/L9-test_clean.csv")
+df=pd.read_csv("experiments/Data/Load/Task 9/L9-model_prediction_laplacian.csv")
+df_test=pd.read_csv(f"experiments/Data/Load/Task 9/L9-test_clean.csv")
 # get subset of interesting columns
 df=df.iloc[:, np.r_[0, 2:101]]
 
@@ -35,7 +35,7 @@ fig, ax = plt.subplots(figsize=(15,5))
 df_res.boxplot(column=["LOAD"], by="TIMESTAMP", showfliers=False, ax=ax)
 
 plt.suptitle("")
-plt.title("Boxplot task 9, laplacian kernel")
+plt.title("Boxplot task 9, Laplacian kernel")
 
 plt.ylabel("Load (MW)")
 plt.xlabel("Hours")

### 2024-05-26 thesis/input_files/paper1.tex
May 26 2024

@@
        Exponential RBF/Laplacian& $k(x,y)=\exp\left(-\frac{\|x-y\|_{2}}{l}\right)$ &   $l$\\
+       Absolute Laplacian& $k(x,y)=\exp\left(-\frac{\|x-y\|_{1}}{l}\right)$ &   $l$\\
        Hyperbolic/Sigmoid Kernel &  $k(x,y2)=\tanh(\gamma x^\intercal y+r)$ &  $\gamma$, r \\

### 2024-05-27 thesis/input_files/analysis.tex
May 27th 2024

@@
 \end{itemize}
-Table \ref{tab:energy_chart_de_additional_predictors} reports the model scores on the German data while Table \ref{tab:energy_chart_ch_additional_predictors} is the one corresponding to the Suisse data.
+Table \ref{tab:energy_chart_de_additional_predictors} reports the model scores on the German data while Table \ref{tab:energy_chart_ch_additional_predictors} is the one corresponding to the Swiss data.
 
@@
 \begin{adjustbox}{width=\textwidth}
-    \begin{tabular}{llllllllll}
+    \begin{tabular}{lllllllllll}
         \toprule
-Quantile & Gaussian RBF & Laplacian & Matern 1.5 & Matern 2.5 & Linear & Periodic & Polynomial & Sigmoid & Cosine \\
+        Quantile & \begin{tabular}[c]{@{}l@{}}Absolute\\ Laplacian\end{tabular} & \begin{tabular}[c]{@{}l@{}}Matern 0.5/\\ Laplacian\end{tabular} & Matern 1.5 & Matern 2.5 & \begin{tabular}[c]{@{}l@{}}Matern $\infty$/\\ Gaussian RBF\end{tabular} & Linear & Periodic & Polynomial & Sigmoid & Cosine \\
         \midrule
-0.1 & 0.028057 & \textbf{0.027483} & 0.028236 & 0.028257 & 0.028261 & 0.028277 & 0.028276 & 0.028146 & 0.028004 \\
-0.2 & 0.047118 & \textbf{0.046051} & 0.047531 & 0.047544 & 0.047569 & 0.047595 & 0.047596 & 0.047353 & 0.047025 \\
-0.3 & 0.060366 & \textbf{0.058739} & 0.060892 & 0.060940 & 0.060962 & 0.060988 & 0.060984 & 0.060654 & 0.060209 \\
-0.4 & 0.068344 & \textbf{0.066347} & 0.068952 & 0.069022 & 0.069054 & 0.069104 & 0.069101 & 0.068634 & 0.068202 \\
-0.5 & 0.071450 & \textbf{0.069118} & 0.072188 & 0.072250 & 0.072285 & 0.072331 & 0.072330 & 0.071807 & 0.071359 \\
-0.6 & 0.068821 & \textbf{0.067098} & 0.069454 & 0.069526 & 0.069571 & 0.069615 & 0.069613 & 0.069157 & 0.068780 \\
-0.7 & 0.061150 & \textbf{0.059972} & 0.061382 & 0.061410 & 0.061427 & 0.061442 & 0.061442 & 0.061281 & 0.060990 \\
-0.8 & 0.047147 & \textbf{0.046457} & 0.047325 & 0.047331 & 0.047340 & 0.047348 & 0.047349 & 0.047247 & 0.047056 \\
-0.9 & 0.026646 & 0.026343 & 0.026749 & 0.026757 & 0.026763 & 0.026772 & 0.026772 & \textbf{0.026282} & 0.026628 \\ 
+        0.1 & 0.027483 & 0.027827 & 0.028236 & 0.028257 & 0.028057 & 0.028261 & 0.028277 & 0.028276 & 0.028146 & 0.028004 \\
+        0.2 & 0.046051 & 0.046679 & 0.047531 & 0.047544 & 0.047118 & 0.047569 & 0.047595 & 0.047596 & 0.047353 & 0.047025 \\
+        0.3 & 0.058739 & 0.059835 & 0.060892 & 0.060940 & 0.060366 & 0.060962 & 0.060988 & 0.060984 & 0.060654 & 0.060209 \\
+        0.4 & 0.066347 & 0.067768 & 0.068952 & 0.069022 & 0.068344 & 0.069054 & 0.069104 & 0.069101 & 0.068634 & 0.068202 \\
+        0.5 & 0.069118 & 0.070761 & 0.072188 & 0.072250 & 0.071450 & 0.072285 & 0.072331 & 0.072330 & 0.071807 & 0.071359 \\
+        0.6 & 0.067098 & 0.068276 & 0.069454 & 0.069526 & 0.068821 & 0.069571 & 0.069615 & 0.069613 & 0.069157 & 0.068780 \\
+        0.7 & 0.059972 & 0.060783 & 0.061382 & 0.061410 & 0.061150 & 0.061427 & 0.061442 & 0.061442 & 0.061281 & 0.060990 \\
+        0.8 & 0.046457 & 0.046944 & 0.047325 & 0.047331 & 0.047147 & 0.047340 & 0.047348 & 0.047349 & 0.047247 & 0.047056 \\
+        0.9 & 0.026343 & 0.026554 & 0.026749 & 0.026757 & 0.026646 & 0.026763 & 0.026772 & 0.026772 & 0.026282 & 0.026628 \\
         \midrule
-CRPS & 0.053233 & \textbf{0.051956} & 0.053634 & 0.053671 & 0.053693 & 0.053719 & 0.053718 & 0.053395 & 0.053139 \\
+        CRPS & 0.051956 & 0.052825 & 0.053634 & 0.053671 & 0.053233 & 0.053693 & 0.053719 & 0.053718 & 0.053395 & 0.053139 \\
         \bottomrule
@@
 \begin{adjustbox}{width=\textwidth}
-    \begin{tabular}{llllllllll}
+    \begin{tabular}{lllllllllll}
         \toprule
-Quantile & Gaussian RBF & Laplacian & Matern 1.5 & Matern 2.5 & Linear & Periodic & Polynomial & Sigmoid & Cosine \\
+        Quantile & \begin{tabular}[c]{@{}l@{}}Absolute\\ Laplacian\end{tabular} & \begin{tabular}[c]{@{}l@{}}Matern 0.5/\\ Laplacian\end{tabular} & Matern 1.5 & Matern 2.5 & \begin{tabular}[c]{@{}l@{}}Matern $\infty$/\\ Gaussian RBF\end{tabular} & Linear & Periodic & Polynomial & Sigmoid & Cosine \\
         \midrule
-0.1 & 0.024309 & \textbf{0.023882} & 0.024316 & 0.024611 & 0.029367 & 0.029462 & 0.027319 & 0.029459 & 0.027816 \\
-0.2 & 0.041028 & \textbf{0.040318} & 0.040987 & 0.041487 & 0.048510 & 0.048716 & 0.046164 & 0.048714 & 0.045879 \\
-0.3 & 0.052597 & \textbf{0.051643} & 0.052548 & 0.052972 & 0.062242 & 0.062561 & 0.059133 & 0.062548 & 0.058454 \\
-0.4 & 0.058803 & \textbf{0.057676} & 0.058841 & 0.059078 & 0.070210 & 0.070577 & 0.067914 & 0.070567 & 0.066335 \\
-0.5 & 0.060846 & \textbf{0.059441} & 0.060912 & 0.061541 & 0.073399 & 0.073782 & 0.071616 & 0.073764 & 0.069574 \\
-0.6 & 0.059336 & \textbf{0.057820} & 0.059237 & 0.059930 & 0.070817 & 0.071157 & 0.070249 & 0.071134 & 0.066970 \\
-0.7 & 0.053937 & \textbf{0.052402} & 0.053949 & 0.054372 & 0.062453 & 0.062688 & 0.062970 & 0.062675 & 0.059419 \\
-0.8 & 0.043366 & \textbf{0.042427} & 0.043364 & 0.043510 & 0.049024 & 0.049150 & 0.049737 & 0.049147 & 0.046662 \\
-0.9 & 0.026727 & \textbf{0.026227} & 0.026645 & 0.026817 & 0.029516 & 0.029601 & 0.030097 & 0.029598 & 0.027986 \\
+    0.100000 & 0.023882 & 0.027901 & 0.024316 & 0.024611 & 0.024309 & 0.029367 & 0.029462 & 0.027319 & 0.029459 & 0.027816 \\
+    0.200000 & 0.040318 & 0.046043 & 0.040987 & 0.041487 & 0.041028 & 0.048510 & 0.048716 & 0.046164 & 0.048714 & 0.045879 \\
+    0.300000 & 0.051643 & 0.058812 & 0.052548 & 0.052972 & 0.052597 & 0.062242 & 0.062561 & 0.059133 & 0.062548 & 0.058454 \\
+    0.400000 & 0.057676 & 0.066704 & 0.058841 & 0.059078 & 0.058803 & 0.070210 & 0.070577 & 0.067914 & 0.070567 & 0.066335 \\
+    0.500000 & 0.059441 & 0.071401 & 0.060912 & 0.061541 & 0.060846 & 0.073399 & 0.073782 & 0.071616 & 0.073764 & 0.069574 \\
+    0.600000 & 0.057820 & 0.067650 & 0.059237 & 0.059930 & 0.059336 & 0.070817 & 0.071157 & 0.070249 & 0.071134 & 0.066970 \\
+    0.700000 & 0.052402 & 0.059061 & 0.053949 & 0.054372 & 0.053937 & 0.062453 & 0.062688 & 0.062970 & 0.062675 & 0.059419 \\
+    0.800000 & 0.042427 & 0.047954 & 0.043364 & 0.043510 & 0.043366 & 0.049024 & 0.049150 & 0.049737 & 0.049147 & 0.046662 \\
+    0.900000 & 0.026227 & 0.028459 & 0.026645 & 0.026817 & 0.026727 & 0.029516 & 0.029601 & 0.030097 & 0.029598 & 0.027986 \\
     \midrule
-CRPS & 0.046772 & \textbf{0.045760} & 0.046755 & 0.047146 & 0.055060 & 0.055299 & 0.053911 & 0.055290 & 0.052122 \\
+    CRPS & 0.045760 & 0.052665 & 0.046755 & 0.047146 & 0.046772 & 0.055060 & 0.055299 & 0.053911 & 0.055290 & 0.052122 \\
     \bottomrule

### 2024-05-27 experiments/test_krnls.py
prepared

@@ -8,14 +8,14 @@ from sklearn.gaussian_process.kernels  import Matern, RBF
 
 g=1/10
 
-# matern = 1.0 * Matern(length_scale=1/(np.sqrt(2*g)), nu=np.inf)
+matern = 1.0 * Matern(length_scale=1/(np.sqrt(2*g)), nu=np.inf)
 
-# X = [[3, 1, 2,4], [1, 2, 1,5],[1,5,32,2]]
-# Y = [[1, 12,24 ,0], [1, 1,12 ,0], [1,5,0,6]]
+X = [[3, 1, 2,4], [1, 2, 1,5],[1,5,32,2]]
+Y = [[1, 12,24 ,0], [1, 1,12 ,0], [1,5,0,6]]
 
-# print("rbf kernel","\n", rbf_kernel(X,Y, gamma=g))
+print("rbf kernel","\n", rbf_kernel(X,Y, gamma=1/1000))
 
-# print("matern_inf kernel","\n", matern(X,Y))
+print("matern_inf kernel","\n", matern(X,Y))
 
 
 # # print("*******")
@@ -27,9 +27,9 @@ matern_05 = 1.0 * Matern(length_scale=1/g, nu=0.5)
 # print("matern_05 kernel","\n", matern_05(X,Y))
 
 
-# when X is 1d a_laplacian and matern are the same
-X = [[3], [1],[1]]
-Y = [[1], [12], [16]]
-print("laplacian kernel","\n", laplacian_kernel(X,Y, gamma=g))
+# # when X is 1d a_laplacian and matern are the same
+# X = [[3], [1],[1]]
+# Y = [[1], [12], [16]]
+# print("laplacian kernel","\n", laplacian_kernel(X,Y, gamma=g))
 
-print("matern_05 kernel","\n", matern_05(X,Y))
+# print("matern_05 kernel","\n", matern_05(X,Y))

### 2024-05-27 experiments/train_test/SECURES-Met/materns.sh

@@ -1,5 +1,8 @@
 #!/usr/bin/env bash
 # script for rerunning materns
+python experiments/train_test/SECURES-Met/secures_qr.py "DE" "a_laplacian"
+python experiments/train_test/SECURES-Met/secures_qr.py "DE" "gaussian_rbf"
+python experiments/train_test/SECURES-Met/secures_qr.py "DE" "matern_0.5"
 python experiments/train_test/SECURES-Met/secures_qr.py "DE" "matern_1.5"
 python experiments/train_test/SECURES-Met/secures_qr.py "DE" "matern_2.5"
 

### 2024-05-27 experiments/train_test/SECURES-Met/secures_qr.py

@@ -8,7 +8,7 @@ import pickle
 
 from sklearn.metrics import make_scorer
 from sklearn.experimental import enable_halving_search_cv  # noqa
-from sklearn.model_selection import HalvingRandomSearchCV
+from sklearn.model_selection import HalvingRandomSearchCV, HalvingGridSearchCV
 
 from sklearn.metrics import mean_pinball_loss
 from sklearn.model_selection import train_test_split
@@ -16,64 +16,71 @@ from sklearn.preprocessing import StandardScaler
 import sys
 from tqdm import tqdm
 
-sys.path.append("/Users/luca/Desktop/kernel_quantile_regression/")
-from kernel_quantile_regression.kqr import KQR
+sys.path.append("/Users/luca/Desktop/ThesisKernelMethods/experiments/")
+from kqr import KQR
+
+if __name__=="__main__":
+
+    country=sys.argv[1]
+    ktype=sys.argv[2]
+    # load train data
+    df=pd.read_csv(f"/Users/luca/Desktop/ThesisKernelMethods/experiments/Data/SECURES-Met/{country}/clean/train/df.csv")
+
+    # quantiles
+    # quantiles = [i/100 for i in range(1,100)]
+    # quantiles = [0.05, 0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9, 0.95]
+    quantiles = [0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9]
 
 
-country="CH"
-ktype="matern_1.5"
-# load train data
-df=pd.read_csv(f"/Users/luca/Desktop/kernel_quantile_regression/Data/SECURES-Met/{country}/clean/train/df.csv")
+    # for the experiment with just temperature and wind_speed exp=""
+    # X y
+    X_train=df[["Direct_irradiation","Global_radiation","Hydro_reservoir","Hydro_river","Temperature","Wind_potential"]]
+    y_train=df["Load"]
 
-# quantiles
-# quantiles = [i/100 for i in range(1,100)]
-# quantiles = [0.05, 0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9, 0.95]
-quantiles = [0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9]
+    # scale
+    scaler = StandardScaler()
 
+    X_train_scaled = scaler.fit_transform(X_train)
[…]

### 2024-05-27 experiments/train_test/SECURES-Met/secures_qr_test.py

@@ -22,13 +22,13 @@ sys.path.append(project_directory)
 from utils import miscellaneous
 
 
-country="CH"
-ktype="matern_1.5"
+country="DE"
+ktype="a_laplacian"
 
 # load train test data
 
-train=pd.read_csv(f"/Users/luca/Desktop/kernel_quantile_regression/Data/SECURES-Met/{country}/clean/train/df.csv")
-test=pd.read_csv(f"/Users/luca/Desktop/kernel_quantile_regression/Data/SECURES-Met/{country}/clean/test/df.csv")
+train=pd.read_csv(f"/Users/luca/Desktop/ThesisKernelMethods/experiments/Data/SECURES-Met/{country}/clean/train/df.csv")
+test=pd.read_csv(f"/Users/luca/Desktop/ThesisKernelMethods/experiments/Data/SECURES-Met/{country}/clean/test/df.csv")
 
 
 # X y
@@ -50,15 +50,15 @@ df_predict=pd.DataFrame(columns=[f"{q}" for q in quantiles])
 
 for i,q in enumerate(quantiles):
     # load model
-    krn_q=pickle.load(open(f'/Users/luca/Desktop/kernel_quantile_regression/train_test/SECURES-Met/{country}/{ktype}/krn_qr_{q}.pkl', 'rb'))
+    krn_q=pickle.load(open(f'/Users/luca/Desktop/ThesisKernelMethods/experiments/train_test/SECURES-Met/{country}/{ktype}/krn_qr_{q}.pkl', 'rb'))
     # predict
     y_predict_q=krn_q.predict(X_test_scaled)
-    print(mean_pinball_loss(y_test,y_predict_q, alpha=q)/np.mean(y_test))
+    print(mean_pinball_loss(y_test,y_predict_q, alpha=q))
     # put in predict df
     df_predict[f"{q}"]=pd.Series(y_predict_q)
 
 # # save predictions to csv
-# df_predict.to_csv(f"/Users/luca/Desktop/kernel_quantile_regression/Data/SECURES-Met/{country}/clean/model_prediction_{ktype}.csv", index=False)
+df_predict.to_csv(f"/Users/luca/Desktop/ThesisKernelMethods/experiments/Data/SECURES-Met/{country}/clean/model_prediction_{ktype}.csv", index=False)
 
 
 # compute pinball loss

### 2024-05-28 beamer/input_files/ef_research.tex
slides structure

% ef research
% ch 2, 4, 5

\section{Energy forecasting research}

\subsection{Probabilistic scores}
\begin{frame}{Pinball loss}
    The pinball score or quantile score is used to measure the accuracy of a quantile forecast.
\begin{definition}
    The pinball loss is defined as
    $$
    \mathrm{Pinball}(y_{t},\hat{y}_{t,q},q)=
\begin{cases}
(q-1)(\hat{y}_{t,q}-y_{t}) & y_t > \hat{y}_{t,q} \\
q(\hat{y}_{t,q}-y_t) & y_t \leq \hat{y}_{t,q}
\end{cases}
$$
\end{definition}
\end{frame}

\begin{frame}{Pinball loss}
    The pinball loss is an asymmetric function, it weights its score differently depending on the error sign and on the quantile considered.
By averaging all the pinball losses over all quantiles and over the whole forecast horizon, we obtain the pinball loss of the probabilistic forecast.
    \begin{center}
        \includegraphics[height=0.5\textheight]{../thesis/images/pinball_loss.png}
    \end{center}
\end{frame}

\begin{frame}{Continous ranked probability score}
The continous ranked probability score (CRPS) measures the difference between the estimated cumulative distribution $\hat{F}$ and the empirical cumulative density function (CDF).
\begin{definition}\label{def_crps}
    The continous ranked probability score is defined as
    $$
    \mathrm{CRPS}(y, \hat{F})=\int\limits_{-\infty}^{\infty}\left(\hat{F}(x)-\mathbb{I}_{\{x-y\}} \right)^2 dx
    $$
\end{definition}
% Nevertheless, we can evaluate the integral in closed form. 
Where the indicator function is defined as 
    $$\mathbb{I}_{\{z\}}=
\begin{cases}
0, & z<0\\
1, & z \geq 0
\end{cases}$$
\end{frame}

\begin{frame}{Continous ranked probability score}
    For a visualisation see Figure. The grey area is what contributes toward the CRPS score.
    The better the estimated cumulative density function is the smaller the total CRPS score will be.
    \begin{figure}
        \includegraphics[width=0.5\textwidth]{../thesis/images/crps.png}
    \end{figure}
\end{frame}

### 2024-05-28 experiments/kqr.py
linear, cosine

@@ -127,7 +127,7 @@ class KQR(RegressorMixin, BaseEstimator):
             return 100.0*laplacian_kernel(X,Y, gamma=1/self.gamma)
         
         if self.kernel_type=="gaussian_rbf_":
-            return 100.0*rbf_kernel(X,Y, gamma=self.gamma)
+            return 100.0*rbf_kernel(X,Y, gamma=1/self.gamma)
 
         elif self.kernel_type=="linear":
             return linear_kernel(X,Y)
@@ -136,10 +136,10 @@ class KQR(RegressorMixin, BaseEstimator):
             return cosine_similarity(X,Y)
         
         elif self.kernel_type=="polynomial":
-            return polynomial_kernel(X,Y, coef0=self.c, degree=self.d)
+            return polynomial_kernel(X,Y, coef0=self.c, degree=self.d, gamma=1/self.gamma)
         
         elif self.kernel_type=="sigmoid":
-            return sigmoid_kernel(X,Y, coef0=self.c, gamma=self.gamma)
+            return 100.0*sigmoid_kernel(X,Y, coef0=self.c, gamma=1/self.gamma)
         
         elif self.kernel_type=="matern_0.5":
             # is the same of laplacian
@@ -155,7 +155,7 @@ class KQR(RegressorMixin, BaseEstimator):
             return matern_kernel(X,Y)
 
         elif self.kernel_type=="chi_squared":
-            return 100.0*chi2_kernel(X,Y,gamma=self.gamma)
+            return 100.0*chi2_kernel(X,Y,gamma=1/self.gamma)
         
         elif self.kernel_type=="periodic":
             periodic=100.0*ExpSineSquared(length_scale=self.gamma, periodicity=self.p)
@@ -163,26 +163,26 @@ class KQR(RegressorMixin, BaseEstimator):
         
         # class of kernels functions are closed under addition and product
         elif self.kernel_type=="gaussian_rbf_x_laplacian":
-            return rbf_kernel(X,Y, gamma=self.gamma)* laplacian_kernel(X,Y, gamma=self.sigma)
+            return rbf_kernel(X,Y, gamma=1/self.gamma)* laplacian_kernel(X,Y, gamma=1/self.sigma)
         
         
         elif self.kernel_type=="se_ard":
             se_ard=1
             for i in range(X.shape[1]-1):
                 
[…]

### 2024-05-28 experiments/test_krnls.py

@@ -3,7 +3,7 @@
 import numpy as np
 
 from sklearn.metrics.pairwise import rbf_kernel, laplacian_kernel, linear_kernel
-from sklearn.gaussian_process.kernels  import Matern, RBF
+from sklearn.gaussian_process.kernels  import Matern, RBF, PairwiseKernel
 # find when they are the same
 
 g=1/10
@@ -13,13 +13,13 @@ matern = 1.0 * Matern(length_scale=1/(np.sqrt(2*g)), nu=np.inf)
 X = [[3, 1, 2,4], [1, 2, 1,5],[1,5,32,2]]
 Y = [[1, 12,24 ,0], [1, 1,12 ,0], [1,5,0,6]]
 
-print("rbf kernel","\n", rbf_kernel(X,Y, gamma=1/1000))
+# print("rbf kernel","\n", rbf_kernel(X,Y, gamma=1/1000))
 
-print("matern_inf kernel","\n", matern(X,Y))
+# print("matern_inf kernel","\n", matern(X,Y))
 
 
 # # print("*******")
-matern_05 = 1.0 * Matern(length_scale=1/g, nu=0.5)
+# matern_05 = 1.0 * Matern(length_scale=1/g, nu=0.5)
 
 
 # print("laplacian kernel","\n", laplacian_kernel(X,Y, gamma=g))
@@ -33,3 +33,9 @@ matern_05 = 1.0 * Matern(length_scale=1/g, nu=0.5)
 # print("laplacian kernel","\n", laplacian_kernel(X,Y, gamma=g))
 
 # print("matern_05 kernel","\n", matern_05(X,Y))
+
+
+laplacian_pairwise=PairwiseKernel(metric='laplacian', gamma=g)
+
+print("laplacian kernel","\n", laplacian_kernel(X,Y, gamma=g))
+print("laplacian pairwise kernel","\n", laplacian_pairwise(X,Y))

### 2024-05-28 experiments/train_test/SECURES-Met/secures_qr.py

@@ -52,7 +52,8 @@ if __name__=="__main__":
     m=len(y_train)
     # C=[1/(m*1e-4),1/(m*1e-3),1/(m*1e-1),1/(m*1e-2),1/(m*1), 1/(m*5), 1/(m*10),1/(m*1e2),1/(m*1e4)],
     param_grid_krn = dict(
-    gamma= [1,2,4,8]
+    d= [2,3],
+    gamma=[1,2,3]
     )
 
     neg_mean_pinball_loss_scorer_05 = make_scorer(
@@ -61,7 +62,7 @@ if __name__=="__main__":
         greater_is_better=False,
         )
 
-    krn_blueprint=KQR(alpha=0.5, C=1, kernel_type=ktype)
+    krn_blueprint=KQR(alpha=0.5, C=1,c=0, kernel_type=ktype)
     cv=HalvingGridSearchCV(
             krn_blueprint,
             param_grid_krn,
@@ -71,16 +72,22 @@ if __name__=="__main__":
         ).fit(X_train, y_train)
     best_hyperparameters_krn=cv.best_params_
 
-    # cv results
+    # # cv results
     df_cv_res=pd.DataFrame(cv.cv_results_)
     df_cv_res.to_csv(f"/Users/luca/Desktop/ThesisKernelMethods/experiments/train_test/SECURES-Met/{country}/{ktype}/models_{ktype}_gridsearch.csv",index=False)
 
     # train
     for i,q in enumerate(tqdm(quantiles)):
-        print(best_hyperparameters_krn)
+        # print(best_hyperparameters_krn)
         # fit data for specific quantile
-        qr_krn_models+=[KQR(alpha=q, C=1, **best_hyperparameters_krn, kernel_type=ktype).fit(X_train_scaled, y_train)]
+        qr_krn_models+=[KQR(alpha=q, C=1,c=0,**best_hyperparameters_krn, kernel_type=ktype).fit(X_train_scaled, y_train)]
 
         # save models to pickle
         pickle.dump(qr_krn_models[i], open(f'/Users/luca/Desktop/ThesisKernelMethods/experiments/train_test/SECURES-Met/{country}/{ktype}/krn_qr_{q}.pkl', 'wb'))
         
+
+
+# parameters for kernel comparison with C=1
+# materns gamma= [1,2,4,8]
+# linear, cosine none
+# polynomial d= [2,3], gamma=[1,2,3]

### 2024-05-28 experiments/train_test/SECURES-Met/secures_qr_test.py

@@ -22,8 +22,8 @@ sys.path.append(project_directory)
 from utils import miscellaneous
 
 
-country="AT"
-ktype="matern_2.5"
+country="CH"
+ktype="polynomial"
 
 # load train test data
 

### 2024-05-28 beamer/input_files/results.tex
polynomial

--- 2024-05-26 thesis/input_files/analysis.tex
@@
-% Analysis of experiments and results
-% \\
-Building on the theory introduced in Sections \ref{ch:point} and \ref{ch:prob}, this section covers the experiments carried out and the results obtained respectively.
-
-% - Comments
-% \\
-% - Comparison
-% \\
-% - Table of loss scores
-% \\
-% Plots:
-% \\
-% - Plots for visualizing timeseries with quantiles bounds
-% \\
-% - Other plots that will come up to mind
-
-\section{Point forecasting}
-This section carries out a comparative study between the state of the art methods for point forecasting, introduced in \ref{ch:point}.
-As use case, we will consider the task of predicting the electric load from the GEFCom2014 dataset.
-In such setting we considered the following regressors
-\begin{itemize}
-    \item Day
-    \item Hour
-    \item Month
-    \item Day of the week
-    \item Is holiday: a binary variable for holidays, where holiday=1 and working day=0
-    \item Weather temperature
-\end{itemize}
-Methods will be compared by means of the RMSE, MAE and MAPE scores, see Section \ref{metrics}.
-\subsection{Multiple linear regression}
-To get started, standard multiple linear regression has been applied, see Figure \ref{fig:mlr_price} for a visualisation. 
-\begin{figure}
-    \includegraphics[width=\textwidth]{images/mlr_price.png}
-    \caption{Multiple linear regression prediction}
-    \label{fig:mlr_price}
-\end{figure}
-
-The resulting RMSE is 30.59.
-What can be concluded, is that multiple linear regression is capable of catching the daily seasonality. Nevertheless, it cannot catch the range of the price series properly.
-
-\subsection{Trigonometric seasonality Box-Cox transformation ARMA errors trend and seasonal components}
-Next, we tried with the autoregressive approach.
-Unfortunately, AR, ARIMA and SARIMA models did not perform as expected, their output was slightly better than the one of linear regression.
-This is probably due to the fact that, the data considered entails two kinds of seasonalities, while ARIMA models can only handle one at a time.
-We remind the reader, that the electricity load time series involves both daily and weekly seasonalities.
-Hence, the need for a more advanced time series model.
-The Tbats \cite{de2011forecasting} model is a forecasting method capable of handling complex patterns in the data.
-Its name stands for trigonometric seasonality, Box-Cox transformation, ARMA errors, trend and seasonal components.
-Tbats forecast is visualized in Figure \ref{fig:tbats_price}, meanwhile its RMSE is 15.08.
-\begin{figure}[!h]
-    \includegraphics[width=\textwidth]{images/tbats_price.png}
-    \caption{Tbats prediction}
-    \label{fig:tbats_price}
-\end{figure}
-From the plot we see that Tbats is capable of catching the lows and average trend, conversely it has some difficulties handling the price peaks.
-\subsection{Prophet}
-Following, the prophet model has been considered. 
-We get started by considering the base implementation. In such setting, prophet takes as input the time series object and learns its data generating process.
-This method achieves a RMSE of 23.96, its prediction is visualized in Figure \ref{fig:prophet_price_1}.
-\begin{figure}[!h]
-    \includegraphics[width=\textwidth]{images/prophet_price_1.png}
-    \caption{Prophet prediction}
-    \label{fig:prophet_price_1}
-\end{figure}
-What can be seen is that prophet correctly models the average trend but does not model peaks and lows precisely. 
-Next, a more complex model was trained.
-We added the weather temperature, the square of it and the categorical variable for holidays effect as regressors.
-Furthermore, we also applied a log transformation to the dependent variables.
-Doing so, RMSE went down to 10.29.
-Forecast is visualised in Figure \ref{fig:prophet_price2}.
-% , moreover figure \ref{fig:prophet_price2.1} and figure \ref{fig:prophet_price2.2} break down the trend and the seasonalities respectively.
-\begin{figure}[!h]
-    \includegraphics[width=\textwidth]{images/prophet_price2.png}
-    \caption{Prophet model\_2 prediction}
-    \label{fig:prophet_price2}
-\end{figure}
-
-% \begin{figure}[!h]
-%     \includegraphics[width=\textwidth]{images/prophet_price2.1.png}
-%     \caption{Seasonalities breakdown}
-%     \label{fig:prophet_price2.1}
-% \end{figure}
-
-% \begin{figure}[!h]
-%     \includegraphics[width=\textwidth]{images/prophet_price2.2.png}
-%     \caption{Prices trend}
-%     \label{fig:prophet_price2.2}
-% \end{figure}
-Prophet alongside with the right features is a good model in the context of electricity load forecasting, it is capable of catching the trend, the peak and the lows.
-
-\subsection{K-nearest neighbours}
-Afterwards, K-nearest neighbours has been considered.
-In doing so, we cross validated the number of neighbours to get the best model instance.
-The forecast is visualised in Figure \ref{fig:knn_price}, RMSE is 12.54.
-\begin{figure}[!h]
-    \includegraphics[width=\textwidth]{images/knn_price.png}
-    \caption{K-nearest neighbour regression}
-    \label{fig:knn_price}
-\end{figure}
-What it can be said is that, K-nearest neighbour is capable of predicting load well by averaging past data.
-
-\subsection{Support vector regression}
-In applying support vector regression we used gridsearch crossvalidation to search for the best regularisation parameter $C$.
-Support vector regression achieves an RMSE of 23.24, for the prediction see Figure \ref{fig:svr_price}.
-\begin{figure}[!h]
-    \includegraphics[width=\textwidth]{images/svr_price.png}
-    \caption{Support vector regression prediction}
-    \label{fig:svr_price}
-\end{figure}
-Visually, we see that the SVR performs similar to multiple linear regression, as expected.
-Like multiple linear regression, SVR can model the daily seasonality but the point prediction is off in terms of highs and lows.
-
-\subsection{Long short term memory}
-We used the \href{https://pytorch.org}{torch} library in order to code our LSTM regressor.
-Various hyperparameters combinations were tried during the tuning phase. The final architecture we set on follows
-\begin{itemize}
-    \item hidden\_size= 64
-    \item num\_layers= 2
-    \item output\_size= 1
-    \item num\_epochs= 30
-    \item learning\_rate= 0.001
-    \item batch\_size= 32
-    \item window\_size= 24
-\end{itemize}
-LSTM forecast is reported in Figure \ref{fig:lstm_price}, it achieves a RMSE of 9.1782
-\begin{figure}[!h]
-    \includegraphics[width=\textwidth]{images/lstm_price.png}
-    \caption{Long short term memory prediction}
-    \label{fig:lstm_price}
-\end{figure}
-
-\subsection{Kernel ridge regression}
-Subsequently, kernel ridge regression was considered. The kernel considered is the radial basis Gaussian function.
-Cross validation was carried jointly over the RBF kernel bandwith and the regularisation constant.
-The RMSE achieved is 9.86, Figure \ref{fig:krnridge_price} reports the prediction.
-\begin{figure}[!h]
-    \includegraphics[width=\textwidth]{images/krnridge_price.png}
-    \caption{Kernel ridge prediction}
-    \label{fig:krnridge_price}
-\end{figure}
-We can observe that, kernel ridge regression accurately models the electricity time series.
-
-\subsection{Kernel support vector regression}
-The last model we considered in this setting was the kernel support vector regression.
-RMSE achieved with this method is 9.10, forecast is reported in Figure \ref{fig:krnsvr_price}.
-
-\begin{figure}[!h]
-    \includegraphics[width=\textwidth]{images/krnsvr_price.png}
-    \caption{Kernel support vector prediction}
-    \label{fig:krnsvr_price}
-\end{figure}
-We can observe that kernel support vector regression is one of the best performing techniques between the ones considered.
-
-\subsection{Results}
-This section reports the tables comparing the considered methods scores.
-% in the context of electricity load forecasting.
-
-% - table 
-% |methods|tasks RMSE|
-% |       |X          |       
-% |       |X          |    
-Table \ref{tab:point_RMSE} reports the RMSE scores for each of the considered methods.
-It can be seen that kernel based methods are the ones achieving the lowest RMSE; specifically, we have KrnSVR, KrnRidge and KNN being among the top methods on the majority of tasks.
-
-\begin{table}[!ht]
-    \caption{Root mean squared errors}
-    \label{tab:point_RMSE}
-    \begin{adjustbox}{width=\textwidth}
-        \begin{tabular}{lrrrrrrrrrrrrrrr}
-            \toprule
-            Method/RMSE & Task 1 & Task 2 & Task 3 & Task 4 & Task 5 & Task 6 & Task 7 & Task 8 & Task 9 & Task 10 & Task 11 & Task 12 & Task 13 & Task 14 & Task 15 \\
-            \midrule
-            MLR & 30.5892 & 29.0347 & 75.1682 & 63.1376 & 42.1156 & 36.6233 & 39.2496 & 38.8286 & 47.1307 & 68.0494 & 61.6005 & 30.3751 & 34.5523 & 33.4287 & 33.5581 \\
-            TBATS & 15.0886 & 31.9720 & 99.1260 & 84.4890 & 51.9740 & 34.9055 & 18.4445 & 38.3246 & 74.1117 & 98.6600 & 84.3050 & 38.8433 & 16.6080 & 28.9902 & 41.6194 \\
-            Prophet & 10.2936 & 14.4358 & 38.7551 & 63.4787 & 19.7474 & 17.5065 & 12.6926 & 14.2665 & 17.5466 & 23.5944 & 43.6666 & 20.8637 & 16.9493 & 19.1626 & 23.3889 \\
-            KNN & 12.5429 & 11.6699 & 24.5057 & 18.3310 & 13.1821 & \textbf{11.9238} & 12.0044 & 14.5165 & 16.3132 & 15.1831 & 37.6457 & 16.4690 & 12.0324 & 11.3102 & 14.1717 \\
-            SVR & 23.2421 & 25.8525 & 78.8043 & 67.6209 & 42.3748 & 32.7845 & 30.5971 & 35.3660 & 55.4213 & 77.9660 & 68.5799 & 30.2451 & 27.0345 & 28.3652 & 32.1480 \\
-            LSTM & 9.1782 & 12.0669 & \textbf{22.7048} & \textbf{16.2087} & 16.0964 & 12.7936 & \textbf{10.8559} & 14.6173 & 19.7303 & 18.0200 & 43.2051 & 17.1856 & 10.3106 & 12.1347 & 17.5849 \\
-            KrnRidge & 9.8619 & 11.6101 & 37.7824 & 35.2459 & \textbf{12.5160} & 14.5911 & 12.8791 & 17.4385 & 16.1131 & 17.1938 & 37.6961 & 14.0076 & 9.8441 & 10.7491 & 13.2975 \\
-            KrnSVR & \textbf{9.1028} & \textbf{11.3117} & 22.9281 & 19.2132 & 12.6331 & 12.6018 & 11.3537 & \textbf{12.9506} & \textbf{14.9731} & \textbf{11.7765} & \textbf{37.3797} & \textbf{13.2694} &
-            \textbf{8.8522} & \textbf{10.6185} & \textbf{13.2602} \\
-            \bottomrule
-            \end{tabular}            
-    \end{adjustbox}
-\end{table}
-
-The MAE scores are contained in Table \ref{tab:point_MAE}. Similarly to above, we can conclude that kernel methods stand out for achieving also the lowest MAE score; with KrnSVR being the top one.
-\begin{table}[!ht]
-    \caption{Mean absolute errors}
-    \label{tab:point_MAE}
-    \begin{adjustbox}{width=\textwidth}
-        \begin{tabular}{lrrrrrrrrrrrrrrr}
-            \toprule
-             Method/MAE & Task 1 & Task 2 & Task 3 & Task 4 & Task 5 & Task 6 & Task 7 & Task 8 & Task 9 & Task 10 & Task 11 & Task 12 & Task 13 & Task 14 & Task 15 \\
-            \midrule
-            MLR & 27.7219 & 24.3823 & 65.4016 & 52.3873 & 34.8465 & 31.2663 & 35.9420 & 34.7316 & 37.0273 & 54.7075 & 48.9327 & 26.1975 & 31.8720 & 29.2306 & 28.3428 \\
-            TBATS & 10.5408 & 23.8417 & 91.3153 & 74.5734 & 40.3258 & 26.0092 & 14.2141 & 24.7818 & 62.0452 & 86.7125 & 72.6912 & 28.4235 & 11.1184 & 21.1038 & 31.8555 \\
-            Prophet & 8.6139 & 11.4307 & 28.3509 & 46.4480 & 15.6290 & 14.1278 & 10.1574 & 11.2360 & 14.0072 & 18.9565 & 27.1142 & 16.9678 & 14.6615 & 16.1750 & 18.3043 \\
-            KNN & 9.4003 & 9.1126 & 19.6213 & 14.4906 & 10.6461 & \textbf{9.4354} & 8.7441 & 10.0065 & 12.4714 & 11.2006 & 20.3087 & 12.5416 & 9.2558 & 8.5937 & 10.5451 \\
-            SVR & 20.2443 & 21.2526 & 69.3704 & 57.1107 & 34.3651 & 27.5875 & 27.4651 & 29.4204 & 43.7009 & 64.3522 & 55.4307 & 25.1080 & 24.0294 & 24.4049 & 26.5869 \\
-            LSTM & 7.5217 & 9.4767 & \textbf{17.9851} & \textbf{13.2158} & 11.8118 & 9.8607 & \textbf{8.0561} & 10.8480 & 15.1008 & 13.8108 & 25.9906 & 13.3588 & 8.1526 & 9.8000 & 14.3186 \\
-            KrnRidge & 7.5965 & 9.2130 & 28.2484 & 24.2366 & \textbf{10.0589} & 10.7758 & 9.5788 & 11.2166 & 12.3427 & 12.3306 & 20.0213 & 10.8434 & 7.1625 & \textbf{7.9710} & 10.0767 \\
-            KrnSVR & \textbf{6.9581} & \textbf{8.6989} & 18.7739 & 15.1398 & 10.1395 & 9.5927 & 8.5731 & \textbf{9.0033} & \textbf{11.0618} & \textbf{8.5486} & \textbf{19.7383} & \textbf{10.5545} & \textbf{7.0395} & 8.1926 & \textbf{9.8125} \\
-            \bottomrule
-            \end{tabular}            
-    \end{adjustbox}            
-\end{table}
-
-Finally Table \ref{tab:point_MAPE} reports the MAPE score for each method. Its results are in line with the other tables, hence suggesting the goodness of kernel methods.
-% What can be concluded is that kernel regressors are also the best in terms of MAPE score; with KrnSVR being the top one.
-\begin{table}[!ht]
-    \caption{Mean absolute percentage errors}
-    \label{tab:point_MAPE}
-    \begin{adjustbox}{width=\textwidth}
-        \begin{tabular}{lrrrrrrrrrrrrrrr}
-            \toprule
-             Method/MAPE & Task 1 & Task 2 & Task 3 & Task 4 & Task 5 & Task 6 & Task 7 & Task 8 & Task 9 & Task 10 & Task 11 & Task 12 & Task 13 & Task 14 & Task 15 \\
-            \midrule
-            MLR & 0.2470 & 0.1935 & 0.3030 & 0.2601 & 0.2474 & 0.2667 & 0.3498 & 0.3058 & 0.1954 & 0.2432 & 0.4234 & 0.2114 & 0.2926 & 0.2502 & 0.2145 \\
-            TBATS & 0.0833 & 0.1650 & 0.4285 & 0.3731 & 0.2482 & 0.1849 & 0.1234 & 0.1634 & 0.3181 & 0.4021 & 0.4880 & 0.1761 & 0.0905 & 0.1555 & 0.2058 \\
-            Prophet & 0.0737 & 0.0854 & 0.1324 & 0.2453 & 0.1134 & 0.1169 & 0.0938 & 0.0912 & 0.0880 & 0.0992 & 0.3757 & 0.1389 & 0.1343 & 0.1356 & 0.1370 \\
-            KNN & 0.0778 & 0.0682 & 0.0924 & 0.0761 & 0.0751 & 0.0740 & 0.0784 & 0.0744 & 0.0725 & 0.0555 & \textbf{0.3115} & 0.0888 & 0.0816 & 0.0699 & 0.0765 \\
-            SVR & 0.1808 & 0.1640 & 0.3211 & 0.2829 & 0.2317 & 0.2260 & 0.2676 & 0.2451 & 0.2215 & 0.2867 & 0.4354 & 0.1870 & 0.2216 & 0.2044 & 0.1933 \\
-            LSTM & 0.0642 & 0.0706 & 0.0896 & \textbf{0.0722} & 0.0798 & 0.0772 & \textbf{0.0710} & 0.0854 & 0.0909 & 0.0741 & 0.3578 & 0.0976 & 0.0719 & 0.0778 & 0.1037 \\
-            KrnRidge & 0.0625 & 0.0685 & 0.1281 & 0.1207 & 0.0718 & 0.0797 & 0.0812 & 0.0760 & 0.0696 & 0.0587 & 0.3127 & 0.0779 & \textbf{0.0591} & \textbf{0.0637} & 0.0722 \\
-            KrnSVR & \textbf{0.0568} & \textbf{0.0642} & \textbf{0.0892} & 0.0788 & \textbf{0.0714} & \textbf{0.0724} & 0.0740 & \textbf{0.0665} & \textbf{0.0629} & \textbf{0.0428} & 0.3122 & \textbf{0.0781} & 0.0607 & 0.0646 & \textbf{0.0713} \\
-            \bottomrule
-            \end{tabular}            
-    \end{adjustbox}
-\end{table}
-
-\section{Probabilistic forecasting}
-We continue our experimental analysis by considering the probabilistic framework.
-The goal of this section is comparing between different quantile regressors and study the choice of the kernel function.
-
-% quantile regressors
-\subsection{Quantile regressors comparison}
-Using the load data from \href{https://www.energy-charts.info/index.html?l=en&c=DE}{energy charts} for Germany and Switzerland, we proceeded to compare between quantile regression algorithms.
-In this study, we trained our models on the entire 2021 data and then tested them and measured their score on the 2022 data.
-Notice, for kernel quantile regression (KQR), the kernel of choice in this setting is the absolute Laplacian.
-Additionally notice, the loss function utilized in this section is the normalized pinball loss.
-% normalized by the magnitude of the average load of each country.
-Here, we built two models. In the first, the national load is forecasted by taking as predictor the following variables
-\begin{itemize}
-    \item Weather temperature
-    \item Wind speed
-\end{itemize}
-Results for Germany are reported in table \ref{tab:energy_chart_de} and figure \ref{fig:DE_load_CI} while results for Switzerland can be found in table \ref{tab:energy_chart_ch} and figure \ref{fig:CH_load_CI}.
-
-\begin{table}[!ht]
-    \centering
-    \caption{Pinball loss for load in Germany (2022)}
-    \label{tab:energy_chart_de}
-    \begin{tabular}{lrrrr}
-    \toprule
-    Quantile & LQR & GBMQR & QF & KQR \\
-    \midrule
-    0.1 & 0.04517 & 0.02890 & 0.02891 & \textbf{0.02767} \\
-    0.2 & 0.08086 & 0.04745 & 0.04906 & \textbf{0.04670} \\
-    0.3 & 0.10820 & 0.06130 & 0.06254 & \textbf{0.05981} \\
-    0.4 & 0.12722 & 0.07031 & 0.07136 & \textbf{0.06834} \\
-    0.5 & 0.13886 & 0.07371 & 0.07428 & \textbf{0.07247} \\
-    0.6 & 0.14182 & 0.07129 & 0.07381 & \textbf{0.07054} \\
-    0.7 & 0.13591 & 0.06227 & 0.06237 & \textbf{0.06215} \\
-    0.8 & 0.11869 & 0.04741 & \textbf{0.04723} & 0.04730 \\
-    0.9 & 0.08458 & 0.02687 & \textbf{0.02668} & 0.02698 \\
-    \midrule
-    CRPS & 0.10904 & 0.05439 & 0.05514 & \textbf{0.05355} \\
-    \bottomrule
-    \end{tabular}
-    \end{table}
-
-\begin{table}[!ht]
-\centering
-    \caption{Pinball loss for load in Switzerland (2022)}
-    \label{tab:energy_chart_ch}
-    \begin{tabular}{lrrrr}
-    \toprule
-    Quantile & LQR & GBMQR & QF & KQR\\
-    \midrule
-    0.1 & 0.04258 & 0.01909 & 0.01899 & \textbf{0.01896} \\
-    0.2 & 0.07460 & 0.03076 & 0.03046 & \textbf{0.03041} \\
-    0.3 & 0.10012 & 0.03838 & 0.03903 & \textbf{0.03814} \\
-    0.4 & 0.11910 & 0.04311 & 0.04378 & \textbf{0.04294} \\
-    0.5 & 0.13095 & 0.04499 & 0.04489 & \textbf{0.04466} \\
-    0.6 & 0.13426 & 0.04387 & 0.04397 & \textbf{0.04347} \\
-    0.7 & 0.12855 & 0.03928 & 0.03948 & \textbf{0.03894} \\
-    0.8 & 0.11201 & 0.03083 & 0.03166 & \textbf{0.03066} \\
-    0.9 & 0.07884 & 0.01878 & 0.01878 & \textbf{0.01841} \\
-    \midrule
-    CRPS & 0.10233 & 0.03434 & 0.03456 & \textbf{0.03407} \\
-    \bottomrule
-    \end{tabular}
-    \end{table}
-
-    \begin{figure}[!ht]
-        \makebox[\linewidth]{\includegraphics[width = 1.5\textwidth]{images/DE_load_CI.png}}
-        \caption[Probabilistic forecast for load in Germany]{Load 90\% confidence interval for Germany Energy charts (2022) data using KQR Laplacian.
-The black line is the observed path for the load.
-The 90\% confidence interval bands are plotted in green.
-Lower and upper red lines denote the 95\% and 5\% quantile forecast respectively}
-        \label{fig:DE_load_CI}
-    \end{figure}
-
-    \begin{figure}[!ht]
-        \makebox[\linewidth]{\includegraphics[width = 1.5\textwidth]{images/CH_load_CI.png}}
-        \caption[Probabilistic forecast for load in Switzerland]{Load 90\% confidence interval for Switzerland Energy charts data (2022) using KQR Laplacian.
-The black line is the observed path for the load.
-The 90\% confidence interval bands are plotted in green.
-Lower and upper red lines denote the 95\% and 5\% quantile forecast respectively}
-        \label{fig:CH_load_CI}
-    \end{figure}
-
-Next, we extended the models to take into account the impact of two additional categorical variables
-\begin{itemize}
-    \item Is holiday: a binary variable for holidays, where holiday$=1$ and working day$=0$
-    \item Day of week: an ordinal categorical variable corresponding to the day of the week, e.i.\ Monday$=0$,$\dots$, Sunday$=6$
-\end{itemize}
-Table \ref{tab:energy_chart_de_additional_predictors} reports the model scores on the German data while Table \ref{tab:energy_chart_ch_additional_predictors} is the one corresponding to the Suisse data.
-
-\begin{table}[!ht]
-    \centering
-    \caption{Pinball loss for load in Germany (2022) model\_2}
-    \label{tab:energy_chart_de_additional_predictors}
-    \begin{tabular}{lrrrr}
-    \toprule
-    Quantile & LQR & GBMQR & QF & KQR \\
-    \midrule
-    0.1 & 0.04522 & 0.02580 & 0.02796 & \textbf{0.02491} \\
-    0.2 & 0.08048 & 0.04326 & 0.04625 & \textbf{0.04209} \\
-    0.3 & 0.10677 & 0.05570 & 0.05706 & \textbf{0.05424} \\
-    0.4 & 0.12610 & 0.06182 & 0.06346 & \textbf{0.06071}\\
-    0.5 & 0.13767 & 0.06214 & 0.06544 & \textbf{0.06150} \\
-    0.6 & 0.14077 & 0.05752 & 0.06001 & \textbf{0.05726} \\
-    0.7 & 0.13409 & 0.04893 & 0.05122 & \textbf{0.04869} \\
-    0.8 & 0.11628 & 0.03687 & 0.03822 & \textbf{0.03665} \\
-    0.9 & 0.08204 & 0.02114 & 0.02324 & \textbf{0.02108} \\
-    \midrule
-    CRPS & 0.10771 & 0.04591 & 0.04810 & \textbf{0.04524} \\
-    \bottomrule
-    \end{tabular}
-    \end{table}
-
-\begin{table}[!ht]
-    \centering
-    \caption{Pinball loss for load in Switzerland (2022) model\_2}
-    \label{tab:energy_chart_ch_additional_predictors}
-    \begin{tabular}{lrrrr}
-    \toprule
-    Quantile & LQR & GBMQR & QF & KQR \\
-    \midrule
-    0.1 & 0.04118 & 0.01881 & 0.01909 & \textbf{0.01876} \\
-    0.2 & 0.07215 & 0.03004 & 0.03040 & \textbf{0.02979} \\
-    0.3 & 0.09603 & 0.03749 & 0.03819 & \textbf{0.03693} \\
-    0.4 & 0.11329 & 0.04230 & 0.04311 & \textbf{0.04165} \\
-    0.5 & 0.12403 & 0.04404 & 0.04465 & \textbf{0.04358} \\
-    0.6 & 0.12748 & 0.04282 & 0.04363 & \textbf{0.04249} \\
-    0.7 & 0.12223 & 0.03789 & 0.03891 & \textbf{0.03779} \\
-    0.8 & 0.10639 & \textbf{0.02961} & 0.03097 & 0.02972 \\
-    0.9 & 0.07531 & \textbf{0.01847} & 0.01957 & 0.01862 \\
-    \midrule
-    CRPS & 0.09756 & 0.03350 & 0.03428 & \textbf{0.03326} \\
-    \bottomrule
-    \end{tabular}
-    \end{table}
-
-% now comment tables
-What can be concluded from this first study is that kernel quantile regression equipped with the absolute Laplacian kernel outperformed almost always its contenders.
-
-% kernel comparison
-\subsection{Kernel function choice}
-This section analyzes which kernel function best fits to the characteristics of the load data.
-The kernel functions considered are: Gaussian RBF, absolute Laplacian, Matern 1.5, Matern 2.5, Linear, Periodic, Polynomial, Sigmoid, and Cosine. 
-Details concerning the hyperparameter selection for the different kernels can be found in \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.metrics.pairwise}{sklearn.metrics.pairwise} and \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.gaussian_process}{sklearn.gaussian\_process}.
-The pool of data for this study has been created by combining \href{https://zenodo.org/records/7907883}{SECURES-Met} data (predictors) \cite{Formayer2023} and the load data from \href{https://www.energy-charts.info/index.html?l=en&c=DE}{energy charts}.
-SECURES-Met data consists of historical data up to the end of $2020$ while from 2021 onward, the data consists of forecasts modeled by the European Centre for Medium-Range Weather Forecasts (ECMWF).
-Therefore, in this experimental study, we used the whole data of $2020$ as the training set and then tested our kernels on the $2021$ data.
-Note that there are two kinds of predictions for data from $2021$ onward, one for each of the emission scenarios RCP$4.5$ and RCP$8.5$.
-In this study, we considered the more conservative RCP$4.5$ scenario data.
-
-The predictors making up the datasets follow:
-\begin{itemize}
-    \item Direct irradiation: direct normal irradiation
-    \item Global radiation: mean global radiation
-    \item Hydro reservoir: daily mean power from reservoir plants in MW
-    \item Hydro river: daily mean power from run of river plants in MW
-    \item Temperature: air temperature
-    \item Wind potential: potential wind power production
-\end{itemize}
-Table \ref{tab:secures_met_de} shows results for Germany, Table \ref{tab:secures_met_ch} is the one for Switzerland and Table \ref{tab:secures_met_at} corresponds to Austria.
-
-\begin{table}[!ht]
-    \centering
-    \caption{Pinball loss kernel comparison for load in Switzerland (2021)}
-    \label{tab:secures_met_ch}
-    \begin{adjustbox}{width=\textwidth}
-        \begin{tabular}{llllllllll}
-    \toprule
-    Quantile & Gaussian RBF & Laplacian & Matern 1.5 & Matern 2.5 & Linear & Periodic & Polynomial & Sigmoid & Cosine \\
-    \midrule
-    0.1 & 0.018861 & \textbf{0.018798} & 0.019285 & 0.019510 & 0.019128 & 0.024474 & 0.021105 & 0.024299 & 0.018996 \\
-    0.2 & 0.030924 & \textbf{0.030430} & 0.031038 & 0.031438 & 0.031223 & 0.040274 & 0.033892 & 0.039774 & 0.031008 \\
-    0.3 & 0.039107 & \textbf{0.038510} & 0.039295 & 0.039751 & 0.039637 & 0.051023 & 0.042683 & 0.050242 & 0.039363 \\
-    0.4 & 0.044285 & \textbf{0.043718} & 0.044763 & 0.045143 & 0.045130 & 0.057785 & 0.048232 & 0.056815 & 0.044896 \\
-    0.5 & 0.046840 & \textbf{0.046261} & 0.047502 & 0.047957 & 0.047739 & 0.060671 & 0.051024 & 0.059697 & 0.047302 \\
-    0.6 & 0.046138 & \textbf{0.045582} & 0.047239 & 0.047686 & 0.047416 & 0.059627 & 0.050456 & 0.058824 & 0.046834 \\
-    0.7 & 0.041910 & \textbf{0.041615} & 0.043404 & 0.043964 & 0.043735 & 0.054688 & 0.046437 & 0.054037 & 0.042897 \\
-    0.8 & 0.034032 & \textbf{0.033852} & 0.035332 & 0.035997 & 0.035518 & 0.044591 & 0.037918 & 0.044112 & 0.034557 \\
-    0.9 & \textbf{0.021817} & 0.021923 & 0.022712 & 0.023082 & 0.022721 & 0.028070 & 0.024161 & 0.027920 & 0.022606 \\
-    \midrule
-    CRPS & 0.035990 & \textbf{0.035632} & 0.036730 & 0.037170 & 0.036916 & 0.046800 & 0.039545 & 0.046191 & 0.036495 \\
-    \bottomrule
-    \end{tabular}
-\end{adjustbox}
-    \end{table}
-
-\begin{table}[!ht]
-    \centering
-\caption{Pinball loss kernel comparison for load in Germany (2021)}
-\label{tab:secures_met_de}
-\begin{adjustbox}{width=\textwidth}
-    \begin{tabular}{llllllllll}
-\toprule
-Quantile & Gaussian RBF & Laplacian & Matern 1.5 & Matern 2.5 & Linear & Periodic & Polynomial & Sigmoid & Cosine \\
-\midrule
-0.1 & 0.028057 & \textbf{0.027483} & 0.028236 & 0.028257 & 0.028261 & 0.028277 & 0.028276 & 0.028146 & 0.028004 \\
-0.2 & 0.047118 & \textbf{0.046051} & 0.047531 & 0.047544 & 0.047569 & 0.047595 & 0.047596 & 0.047353 & 0.047025 \\
-0.3 & 0.060366 & \textbf{0.058739} & 0.060892 & 0.060940 & 0.060962 & 0.060988 & 0.060984 & 0.060654 & 0.060209 \\
-0.4 & 0.068344 & \textbf{0.066347} & 0.068952 & 0.069022 & 0.069054 & 0.069104 & 0.069101 & 0.068634 & 0.068202 \\
-0.5 & 0.071450 & \textbf{0.069118} & 0.072188 & 0.072250 & 0.072285 & 0.072331 & 0.072330 & 0.071807 & 0.071359 \\
-0.6 & 0.068821 & \textbf{0.067098} & 0.069454 & 0.069526 & 0.069571 & 0.069615 & 0.069613 & 0.069157 & 0.068780 \\
-0.7 & 0.061150 & \textbf{0.059972} & 0.061382 & 0.061410 & 0.061427 & 0.061442 & 0.061442 & 0.061281 & 0.060990 \\
-0.8 & 0.047147 & \textbf{0.046457} & 0.047325 & 0.047331 & 0.047340 & 0.047348 & 0.047349 & 0.047247 & 0.047056 \\
-0.9 & 0.026646 & 0.026343 & 0.026749 & 0.026757 & 0.026763 & 0.026772 & 0.026772 & \textbf{0.026282} & 0.026628 \\ 
-\midrule
-CRPS & 0.053233 & \textbf{0.051956} & 0.053634 & 0.053671 & 0.053693 & 0.053719 & 0.053718 & 0.053395 & 0.053139 \\
-\bottomrule
-\end{tabular}
-\end{adjustbox}
-\end{table}
-
-\begin{table}[!ht]
-    \centering
-\caption{Pinball loss kernel comparison for load in Austria (2021)}
-\label{tab:secures_met_at}
-\begin{adjustbox}{width=\textwidth}
-    \begin{tabular}{llllllllll}
-\toprule
-Quantile & Gaussian RBF & Laplacian & Matern 1.5 & Matern 2.5 & Linear & Periodic & Polynomial & Sigmoid & Cosine \\
-\midrule
-0.1 & 0.024309 & \textbf{0.023882} & 0.024316 & 0.024611 & 0.029367 & 0.029462 & 0.027319 & 0.029459 & 0.027816 \\
-0.2 & 0.041028 & \textbf{0.040318} & 0.040987 & 0.041487 & 0.048510 & 0.048716 & 0.046164 & 0.048714 & 0.045879 \\
-0.3 & 0.052597 & \textbf{0.051643} & 0.052548 & 0.052972 & 0.062242 & 0.062561 & 0.059133 & 0.062548 & 0.058454 \\
-0.4 & 0.058803 & \textbf{0.057676} & 0.058841 & 0.059078 & 0.070210 & 0.070577 & 0.067914 & 0.070567 & 0.066335 \\
-0.5 & 0.060846 & \textbf{0.059441} & 0.060912 & 0.061541 & 0.073399 & 0.073782 & 0.071616 & 0.073764 & 0.069574 \\
-0.6 & 0.059336 & \textbf{0.057820} & 0.059237 & 0.059930 & 0.070817 & 0.071157 & 0.070249 & 0.071134 & 0.066970 \\
-0.7 & 0.053937 & \textbf{0.052402} & 0.053949 & 0.054372 & 0.062453 & 0.062688 & 0.062970 & 0.062675 & 0.059419 \\
-0.8 & 0.043366 & \textbf{0.042427} & 0.043364 & 0.043510 & 0.049024 & 0.049150 & 0.049737 & 0.049147 & 0.046662 \\
-0.9 & 0.026727 & \textbf{0.026227} & 0.026645 & 0.026817 & 0.029516 & 0.029601 & 0.030097 & 0.029598 & 0.027986 \\
-\midrule
-CRPS & 0.046772 & \textbf{0.045760} & 0.046755 & 0.047146 & 0.055060 & 0.055299 & 0.053911 & 0.055290 & 0.052122 \\
-\bottomrule
-\end{tabular}
-\end{adjustbox}
-\end{table}
-From our tables, it can be concluded that the absolute Laplacian kernel is the one fitting best to the characteristics of load data.
-Additionally, we can conclude that our results are in line with the ones published in \cite{he2017short}.
-In that study, similarly to what we did in this subsection, the authors carried out a comparitive study concerning the choice of kernels in KQR.
-Nevertheless notice, that the set of kernel we considered in our study is larger than the one presented in \cite{he2017short}.
-Their paper showed evidence of the superiority of the Gaussian RBF among their set of kernel functions.
-Analogously, we can see from Tables \ref{tab:secures_met_de}, \ref{tab:secures_met_ch} and \ref{tab:secures_met_at} that the Gaussian RBF is the second best in terms of pinball loss.
-
-% gefcom final case
-\subsection{GEFCom2014}
-The following subsection compares our kernel quantile regression against the price and load track entries of the GEFCom2014 competition.
-Error measure of the competition was the pinball loss, see Section \ref{metrics}, averaged over the $99$ quantiles, $q \in \{i/100\}_{i=1}^{n=99}$.
-Backed by the conclusion from the previous subsection we selected the absolute Laplacian and the Gaussian RBF as kernels in our application of KQR.
-
-\subsubsection{GEFCom2014 price track}
-In this setting, the predictors fed to our KQR models are
-\begin{itemize}
-    \item Day
-    \item Hour
-    \item Forecasted total load
-    \item Forecasted zonal load
-\end{itemize}
-Since target days vary all over the year, we trained a model for any month associated to them.
-
-Our results for this track are reported in Table \ref{tab:pinball loss gefcom2014 price data}.
-In this track, the top entries came from the teams: Tololo, Team Poland, GMD, and C3 Green Team; for a breakdown of each method's attributes, see \cite[Table 8]{hong2016probabilistic}.
-Finally, the Gaussian RBF probabilistic prediction for the $13^{th}$ July $2013$ zonal price at the $90\%$ confidence interval is visualized in Figure \ref{fig:price_task_6_gaussian_rbf}.
-
-\begin{figure}[!ht]
-    \centering
-    \includegraphics[width=\linewidth]{images/price_task_6_gaussian_rbf.png}
-    \caption[Prediction price track task 6, Gaussian RBF kernel]{Price 90\% confidence interval task 6.
-Electricity price probabilistic forecast for the $13^{\text{th}}$ July $2013$.
-The black line is the observed path for the price.
-The $90\%$ confidence interval bands are plotted in green.
-Lower and upper red lines denote the $95\%$ and $5\%$ quantile forecast respectively}
-    % \Description{}
-    \label{fig:price_task_6_gaussian_rbf}
-\end{figure}
-
-\begin{table}[!ht]
-    \centering
-    \caption{Pinball loss GEFCom2014 price data}
-    \label{tab:pinball loss gefcom2014 price data}
-    \begin{adjustbox}{width=\textwidth}
-    \begin{tabular}{lllllllllllll}
-      \toprule
-      \midrule
-      Team name\textbackslash Task number                       & 1                               & 2                                  & 3                               & 4                              & 5                              & 6       & 7                               & 8       & 9       & 10                             & 11                             & 12               \\
-  KQR Laplacian
-  &
-  \textbf{1.3249}
-  &
-  2.9497
-  &
-  0.8868
-  &
-  7.3780
-  &
-  5.4314
-  &
-  8.2166
-  &
-  4.9091
-  &
-  1.5245
-  &
-  1.9601
-  &
-  2.7324
-  &
-  2.4733
-  &
-  17.5371
-  \\
-  KQR Gaussian RBF
-  &
-  2.0614
-  &
-  3.2247
-  &
-  \textbf{0.7050}
-  &
-  6.1407
-  &
-  6.1134
-  &
-  8.5969
-  &
-  2.9360
-  &
-  1.5731
-  &
-  1.8830
-  &
-  2.6419
-  &
-  2.4814
-  &
-  16.1921
-  \\
-  \\
-
-  Arkadiy Strelnikov         & 1,92899          & 3,46853          & 4,35373          & 7,16274          & 7,42959          & 6,32601          & 5,03883          & 2,02658          & \textbf{0,66422} & 2,35183          & 2,08614          & 7,23447          \\
-  Benchmark - Price          & 4,02875          & 7,97208          & 5,70395          & 12,15104         & 38,33541         & 44,22979         & 18,22395         & 31,56729         & 42,94958         & 2,85583          & 3,20395          & 22,38333         \\
-  C3 Green Team              & 1,85897          & 3,27786          & 1,2593           & 5,08886          & 6,87674          & 6,1505           & 4,42379          & 1,32639          & 1,25915          & 3,08224          & 1,55811          & 6,58123          \\
-  E.S.
-Mangalova             & 2,05693          & 7,97208          & 0,87971          & 7,04219          & 11,0464          & 6,57565          & 6,02388          & \textbf{0,69721} & 2,77446          & 2,78586          & 2,14229          & 7,35955          \\
-  EPSteam                    & 4,02875          & 7,97208          & 5,70395          & 3,46235          & 29,5226          & 27,73186         & 17,0324          & 2,49354          & 1,25639          & \textbf{1,79504} & 1,63482          & 10,357           \\
-  Florencio Gonzalez         & 4,02875          & 4,32597          & 4,35373          & 9,72491          & 10,51628         & 44,17405         & 4,00812          & 4,6924           & 7,36395          & 1,79865          & 2,1575           & 4,42008          \\
-  GMD                        & 3,7271           & 1,783            & 0,92191          & 5,08886          & 6,21331          & \textbf{3,82599} & 4,9342           & 1,47858          & 1,65933          & 2,06134          & 2,1235           & 6,84571          \\
-  Manuel Oviedo de la Fuente & 1,62605 & 3,27786          & 13,73529         & 6,68756          & 23,55608         & 10,01475         & 6,61296          & 2,2259           & 1,48028          & 3,65169          & 3,20396          & 4,7237           \\
-  NimNid                     & 3,44735          & 10,59679         & 2,76634          & 24,32776         & 23,55608         & 7,96082          & 3,33627          & 1,8191           & 1,60593          & 2,5711           & 2,32578          & 8,41167          \\
-  San/Saini                  & 2,47163          & 1,95533          & 0,84071 & 5,31786          & 9,61338          & 8,28547          & 3,0475           & 2,86903          & 3,60395          & 4,37704          & 1,82957          & 16,81896         \\
-  Team Poland                & 1,97477          & 1,81898          & 1,19162          & 2,82318          & 7,55914          & 4,20773          & 2,59715          & 1,04693          & 1,24193          & 4,06012          & \textbf{1,08458} & \textbf{3,06512} \\
-  Tololo                     & 1,70734          & \textbf{1,45173} & 1,10384          & \textbf{2,01694} & 9,15596          & 4,6821           & \textbf{1,59517} & 0,75352          & 2,45935          & 2,9614           & 1,34614          & 3,55819          \\
-  Xiaorong (Iris) Sun        & 1,45661          & 1,59212          & 0,98985          & 3,0349           & 4,73309          & 4,52459          & 3,63208          & 2,30481          & 0,90781          & 5,00935          & 1,18223          & 4,5302           \\
-  Yanghai Cong               & 1,69815          & 5,84             & 4,83261          & 3,17957          & 10,21729         & 6,44717          & 5,55374          & 3,80812          & 4,38485          & 1,45195          & 1,51106          & 14,61798         \\
-  dmlab                      & 2,30162          & 1,92804          & 1,2593           & 2,58646          & 14,09957         & 7,5589           & 4,13365          & 0,80748          & 1,5149           & 3,71509          & 3,43097          & 10,22129         \\
-  pat1                       & 2,36615          & 1,98567          & 1,07248          & 2,79465          & \textbf{4,23269} & 4,70614          & 8,40506          & 1,25376          & 2,23991          & 3,67952          & 1,06139          & 6,27517
-
-  \end{tabular}
-    \end{adjustbox}
-  \end{table}
-
-What can be concluded from Table \ref{tab:pinball loss gefcom2014 price data} is that KQR performs consistently among the top algorithms.
-Furthermore, It ranks first in two out of the thirteen tasks of the price track.
-\subsubsection{GEFCom2014 load track} 
-The GEFCom2014 load track constitutes a good setting for analysing the performance of KQR in the context of medium term load forecasting (MTLF).
-In this track we were faced with the challenge of predicting the load for the next month without the availability of weather temperature forecasts. 
-Therefore, the primary task was to first accurately predict the weather temperatures, and then model the load accordingly.
-Since there were no attributes available for humidity or wind speed, we chose to predict weather temperatures by aggregating historical temperature data across different dimensions such as day, month, and hour.
-Then we proceeded with building KQR models for the load; we chose the following predictors
-\begin{itemize}
-    \item Day: the number of the day
-    \item Hour
-    \item Day of week: an ordinal categorical variable for the day of the week
-    \item Is holiday: a binary variable for holidays, where holiday=1 and working day=0
-    \item w avg: average of weather temperatures across 25 stations
-\end{itemize}
-We built $12$ models, one for each month, with each task model trained on the historical data of the month associated with it.
-
-Table \ref{tab:pinball loss gefcom2014 load data} reports our results for the load track.
-The top teams for the load forecasting track were Tololo, Adada, Jingrui(Rain) Xie, OxMAth, E.S.
-Managalova, Ziel Florian, and Bidong Liu; for a breakdown of the attributes of each method, see \cite[Table 6]{hong2016probabilistic}.
-Finally, the $90\%$ confidence interval forecast by our model equipped with the absolute Laplacian kernel for task number $9$, that is the prediction for June, is reported in Figure \ref{fig:load_task_9}.
-Similarly to the price track, we have that KQR was among the best algorithms also in the load forecasting track.
-
-\begin{figure*}[!ht]
-    \centering
-    \makebox[\linewidth]{\includegraphics[width = 1.5\textwidth]{images/load_task_9_laplacian.png}}
-    \caption[Prediction load track task 9, Laplacian kernel]{Load 90\% confidence interval task 9 using KQR Laplacian.
-The black line is the observed path for the load.
-The 90\% confidence interval bands are plotted in green.
-Lower and upper red lines denote the 95\% and 5\% quantile forecast respectively}
-    \label{fig:load_task_9}
-\end{figure*}
-
-\begin{figure*}[!ht]
-    \centering
-    \makebox[\linewidth]{\includegraphics[width = 1.5\textwidth]{images/sub_boxplot_task_9_laplacian_.png}}
-    \caption{Boxplot for first two days of probabilistic forecast task 9 using KQR Laplacian}
-    \label{fig:load_task_9}
-\end{figure*}
-
-\begin{table}[!ht]
-    \caption{Pinball loss GEFCom2014 load data}
-    \label{tab:pinball loss gefcom2014 load data}
-    \begin{adjustbox}{width=\textwidth}
-    \begin{tabular}{lllllllllllll}
-      \toprule
-      \midrule
-      Team name\textbackslash Task number                       & 1                               & 2                                  & 3                               & 4                              & 5                              & 6       & 7                               & 8       & 9       & 10                             & 11                             & 12                              \\
-
-  KQR Laplacian
-  & 
-  12.0552
-  &
-  12.0573
-  &
-  8.8213
-  &
-  5.1721
-  &
-  6.9133
-  &
-  7.8080
-  &
-  11.5559
-  &
-  11.8250
-  &
-  6.8941
-  &
-  3.9680
-  &
-  7.4931
-  &
-  10.8869
-  \\
-  KQR Gaussian RBF
-  & 
-  12.4665
-  &
-  12.1780
-  &
-  9.3326
-  &
-  5.1713
-  &
-  6.8703
-  &
-  6.7482
-  &
-  11.02473
-  &
-  11.93156
-  &
-  6.6019
-  &
-  4.3111
-  &
-  7.2207
-  &
-  10.8840
-  \\
-  \\
-  ACE                        & 12.1330                         & 14.4846                         & \textbf{7.3933}                          & 4.8213                         & 6.8048                         & 7.0566  & 9.5921                          & 11.6316 & 5.9859  & 5.0730 & \textbf{5.6028}                         & 8.9699                          \\
-  Adada                      & 10.5093                         & 10.0801                         & 7.6238                          & 4.7289                         & \textbf{5.3936}                         & 6.6242  & 8.0144                          & 11.1366 & 5.7779  & 3.6379                         & 7.0096                         & 8.9109                          \\
-  Alastair Muir              & 13.6118                         & 16.6750                         & 10.0759                         & 7.7078 & 9.5386 & 8.3615  & 9.8152                          & 13.1363 & 8.9715  & 5.4082                         & 8.5881 & 17.5325                         \\
-  Andrew Landgraf            & 14.3650                         & 10.1090                         & 8.4180                          & 6.2522                         & 7.2248                         & 11.1638 & 9.9403                          & 11.0204 & 5.6920  & 6.1176                         & 11.0677                        & 13.3985                         \\
-  Benchmark - Load           & 18.7384                         & 22.7585                         & 13.2163                         & 8.3626                         & 10.9162                        & 16.9937 & 13.4038                         & 17.3151 & 13.8374 & 6.4237                         & 10.9380                        & 34.0685                         \\
-  Bidong Liu                 & 16.4215                         & 11.8655                         & 9.3733                          & 5.6212                         & 7.7387                         & 6.5536  & 9.1406                          & 11.3485 & 6.5096  & 4.8031                         & 6.9697                         & 10.8935                         \\
-  C3 Green Team              & 18.7384 & 19.2208 & 7.9637                          & 4.6370                         & 6.4543                         & 8.3799  & 10.5546                         & 10.6609 & 5.8867  & 4.4866                         & 5.9396                         & 10.3917                         \\
-  Christopher Benfield       & 18.7384 & 11.1324                         & 9.4377                          & 4.9097                         & 7.4184                         & 19.7325 & 9.2215                          & 11.4385 & 6.6395  & 4.5966                         & 6.5002                         & 10.8633                         \\
-  Dao Vu                     & 33.3711                         & 13.3340                         & 10.4300                         & 6.0815                         & 9.0706                         & 8.7098  & 11.2808 & 18.4869 & 7.4065  & 5.7414                         & 10.3431                        & 23.5659                         \\
-  E.S.
-Mangalova             & 18.7384 & 13.3340 & 7.8025                          & \textbf{4.4096}                         & 6.6330                         & 6.2306  & 10.1511                         & 10.9294 & 6.2224  & 4.2382                         & 6.5464                         & 8.8080                          \\
-  Jingrui (Rain) Xie         & 11.8700                         & 10.9250                         & 8.4938                          & 4.9611                         & 7.4442                         & 6.9921  & 9.0523                          & 11.2600 & 5.4864  & \textbf{3.3602}                         & 5.9011                         & 9.7316                          \\
-  Manuel Oviedo de la Fuente & 12.5502                         & 21.4591                         & 10.1593                         & 5.4647                         & 9.3072                         & 7.6102  & 8.8361                          & 12.6340 & 12.3969 & 15.5225                        & 6.5221                         & 18.7754                         \\
-  OxMath                     & 14.4091 & \textbf{8.9136}                          & 7.6059                          & 4.4548                         & 7.2944                         & 7.4551  & 7.9527                          & \textbf{10.2444} & 5.4551  & 4.2111                         & 6.4054                         & 9.5520                          \\
-  Rasmus Paivarinta          & 18.7384 & 11.7474                         & 9.7230                          & 7.7078                         & 8.3074 & 7.1793  & 8.4391                          & 10.9357 & 6.3855  & 4.3796                         & 6.3794                         & 11.7871                         \\
-  SAOR                       & 18.7384 & 21.4591 & 10.4300 & 6.8512                         & 8.6899                         & 6.7653  & 12.1427                         & 11.6261 & 6.7310  & 6.2850 & 5.9699                         & 21.2711                         \\
-  Sniper                     & 18.7384 & 10.3056                         & 8.3493                          & 5.6639                         & 8.3074                         & 6.9599  & 10.8848                         & 13.4937 & 7.8877  & 5.3514                         & 6.4770                         & 11.1951                         \\
-  Tololo                     & \textbf{10.4369}                         & 12.5232                         & 8.2695                          & 4.4220                         & 5.8976                         & \textbf{6.1878}  & \textbf{7.3182}                          & 10.8032 & \textbf{5.4469}  & 3.9613                         & 6.3173                         & \textbf{8.4787}                          \\
-  Trevor Maynor              & 18.6934                         & 19.4422                         & 12.6071                         & 5.5631                         & 10.9298                        & 13.9594 & 10.8657                         & 14.4647 & 7.2859  & 6.2850                         & 8.5881                         & 17.4703                         \\
-  Xiaorong (Iris) Sun        & 13.8779                         & 12.3979                         & 10.3672                         & 8.3626 & 8.1979                         & 20.5364 & 11.2808                         & 12.6316 & 6.6821  & 4.4907                         & 6.2363                         & 10.1624                         \\
-  dmlab                      & 14.4091                         & 14.0059                         & 9.2128                          & 5.9099                         & 6.6675                         & 6.9046  & 10.0130                         & 11.0201 & 7.8508  & 3.8128                         & 6.6474                         & 26.6655                         \\
-  nikolina                   & 18.7384 & 19.2208                         & 10.6214                         & 5.8257                         & 9.5386                         & 7.1234  & 10.0948                         & 12.5375 & 5.8139  & 4.8060                         & 8.4956                         & 21.2711 \\
-  pat1                       & 19.3898                         & 10.5097                         & 9.1477                          & 9.7023                         & 7.0428                         & 7.8794  & 8.8610                          & 12.8179 & 5.5443  & 5.0730                         & 6.5846                         & 11.2142
-  \end{tabular}
-  \end{adjustbox}
-  \end{table}
-
-% In this subsection we compare performance of quantile regressors.
-% \\
-% - experiment for quantile estimator on gefcom2014 and we are good
-
-\section{Conclusions}
-% repeat the problem description  and motivation quickly
-This section concludes this thesis work with some final remarks.
-This thesis was concerned with electricity forecasting, in particular we focused on the probabilistic framework.
-Probabilistic forecasting is still a relatively new research area.
-Lately, research efforts are shifting towards this setting.
-This is due to the fact that probabilistic forecasts are more informative than point ones and better suited for the current and future electricity landscape; think for example about the higher uncertainty resulting from market liberalisation or renewables integration requirements.
-
-Among the contributions of this thesis work, we can list
-% list contributions:
-% - study of kernel methods in EF, both point and probabilistic
-% - contribution to less developed area of EF, which is probabilistic forecasting
-% - provide open source python package for KQR PyPi, github repo, reproducibility purposes
-% - study case for DACH region, along with cleaned dataset for easing collaboration and replication in research
-% - extensive comparison among kernel types
-% - comparison against state of the art algorithms from GEFCom2014, particularly KQR in medium term load forecasting which has not been deeply analyzed in the literature.
+\begin{frame}{Results}
     \begin{itemize}
@@
         \item We provided a Python open source implementation for kernel quantile regression compatible with the sklearn API.
-The code has been packaged and uploaded to the Python Package Index (PyPI) with the name \href{https://pypi.org/project/kernel-quantile-regression/#2}{kernel-quantile-regression}.
+The code has been packaged and uploaded to the Python Package Index (PyPI) with the name \href{https://pypi.org/project/kernel-quantile-regression/\#2}{kernel-quantile-regression}.
 The \href{https://github.com/luca-pernigo/kernel_quantile_regression}{github repo} hosting the source code includes also the script implementing the experiments along with the cleaned datasets; this contribution is intended to forster reproducibility in research.
@@
     \end{itemize}
-
-% results achieved:
-% - KQR beats other quantile regressor algorithms
-% - Laplacian is the best kernel
-% - our KQR implementations demostrated good performance compared to the top entries of the GEFCom2014 competition, this suggests it is a valid method.
-We observe that KQR shows favorably and can forecast both the short and medium-term horizon.
-To finish, we proceed summarising the conclusions we drew from our experimental analysis.
-Kernel methods showed good performance both in the context of point and probabilistic forecasting.
-Specifically, kernel quantile regression resulted almost always in superior performance compared to standard quantile regressors.
-The absolute Laplacian kernel turned out to be the most suitable kernel according to the characteristics of electricity forecasting.
-Finally, we proved the validity of our kernel quantile regression implementation.
-We achieved this by showing how it compared favourably to the top entries of the GEFCom2014 competition.
-In particular, up to now research work focused on the short term horizon capabilities of kernel quantile regression.
-In this thesis work, we were also able to show its suitability on the medium term horizon.
+\end{frame}

### 2024-05-28 experiments/train_test/SECURES-Met/secures_qr.py

@@ -80,7 +80,7 @@ if __name__=="__main__":
     for i,q in enumerate(tqdm(quantiles)):
         # print(best_hyperparameters_krn)
         # fit data for specific quantile
-        qr_krn_models+=[KQR(alpha=q, C=1,c=0,d=2,gamma=1, kernel_type=ktype).fit(X_train_scaled, y_train)]
+        qr_krn_models+=[KQR(alpha=q, C=1,c=1,gamma=1,d=2, kernel_type=ktype).fit(X_train_scaled, y_train)]
 
         # save models to pickle
         pickle.dump(qr_krn_models[i], open(f'/Users/luca/Desktop/ThesisKernelMethods/experiments/train_test/SECURES-Met/{country}/{ktype}/krn_qr_{q}.pkl', 'wb'))

### 2024-05-28 experiments/train_test/SECURES-Met/secures_qr_test.py

@@ -23,7 +23,7 @@ from utils import miscellaneous
 
 
 country="AT"
-ktype="polynomial"
+ktype="linear"
 
 # load train test data
 

### 2024-05-29 experiments/train_test/SECURES-Met/scores.py
scores kernel comparison

@@ -6,7 +6,7 @@ from sklearn.metrics import mean_pinball_loss
 # script for storing scores SECURES-Met data
 
 quantiles=[0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9]
-country="CH"
+country="AT"
 
 test=pd.read_csv(f"/Users/luca/Desktop/ThesisKernelMethods/experiments/Data/SECURES-Met/{country}/clean/test/df.csv")
 y_test=test["Load"]

### 2024-05-30 experiments/utils/price_create_test.py
is holiday

@@ -8,12 +8,12 @@ import holidays
 import miscellaneous
 
 # bash script
-# for ((i=1;i<=15;i++)); do python utils/price_create_test.py "Data/Price/Task $i/Task${i}_P.csv"; done
+# for ((i=1;i<=15;i++)); do python experiments/utils/price_create_test.py "experiments/Data/Price/Task $i/Task${i}_P.csv"; done
 
 
 
 def create_test(file):
-    df=miscellaneous.clean_time(file, 2011, 2014)
+    df=miscellaneous.clean_time(file, 2011, 2014,0)
 
     df_test=df[-24:]
     # print(df_test)
@@ -30,7 +30,7 @@ def create_test(file):
     
     
     # order columns
-    df_merged=miscellaneous.order_columns(df_merged,["ZONEID","timestamp","MONTH","DAY","HOUR","Forecasted Total Load","Forecasted Zonal Load","Zonal Price"])
+    df_merged=miscellaneous.order_columns(df_merged,["ZONEID","timestamp","MONTH","DAY","HOUR","Forecasted Total Load","Forecasted Zonal Load","Zonal Price","IS_HOLIDAY"])
     # save
     n=miscellaneous.get_task_number(file)
     
@@ -39,7 +39,7 @@ def create_test(file):
         df_merged["ZONEID"]=df_new["ZONEID"]
         df_merged["Zonal Price"]=df_new["Zonal Price"]
         
-    df_merged.to_csv(f"Data/Price/Task {n}/Task{n}_P_test.csv", index=False)
+    df_merged.to_csv(f"experiments/Data/Price/Task {n}/Task{n}_P_test.csv", index=False)
 
 
 if __name__=="__main__":

### 2024-05-30 experiments/utils/price_create_train.py

@@ -13,12 +13,15 @@ def create_train(file):
     df=miscellaneous.clean_time(file, 2011, 2014,0)
 
     df_train=df[0:-24]
+    
+    print(df)
 
     # order columns
-    df_train=miscellaneous.order_columns(df_train,["ZONEID","timestamp","MONTH","DAY","HOUR","Forecasted Total Load","Forecasted Zonal Load","Zonal Price"])
+    df_train=miscellaneous.order_columns(df_train,["ZONEID","timestamp","MONTH","DAY","HOUR","Forecasted Total Load","Forecasted Zonal Load","Zonal Price", "IS_HOLIDAY"])
     # save
+    
     n=miscellaneous.get_task_number(file)
-    df_train.to_csv(f"Data/Price/Task {n}/Task{n}_P_train.csv", index=False)
+    df_train.to_csv(f"experiments/Data/Price/Task {n}/Task{n}_P_train.csv", index=False)
 
 if __name__=="__main__":
     file=sys.argv[1]

### 2024-05-30 thesis/input_files/data_exploratory.tex
corrected chapter 7

@@
 Additionally, as the competition went on, the real observed data of the previous tasks were made available.
-The fifteen target days ordered by task number are: 16/06/2013, 17/06/2013, 24/06/2013, 04/07/2013, 09/07/2013, 13/07/2013, 16/07/2013, 18/07/2013, 19/07/2013, 20/07/2013, 24/07/2013, 25/07/2013, 07/12/2013, 08/12/2013.
+The fifteen target days ordered by task number are: 16/06/2013, 17/06/2013, 24/06/2013, 04/07/2013, 09/07/2013, 13/07/2013, 16/07/2013, 18/07/2013, 19/07/2013, 20/07/2013, 24/07/2013, 25/07/2013, 07/12/2013, 08/12/2013, 17/12/2013.
 \begin{figure}[!h]

### 2024-05-30 thesis/input_files/probabilistic.tex
corrected chapter 7

@@
       \centering
-      \includegraphics[width=1.1\textwidth]{images/melbourne_kernel_quantile_regression.png}
+      \includegraphics[width=1.1\textwidth]{images/melbourne_gaussian_rbf_kernel_quantile_regression.png}
       \caption{Kernel quantile regression, Gaussian RBF} 
@@
     \midrule
-    Pinball loss & 11.278895 & 10.317612 & 10.340842 & \textbf{10.031708} \\
+    Pinball loss & 11.278895 & 10.317612 & 10.340842 & \textbf{10.031546} \\
     \bottomrule
@@
     \midrule
-    0.100000 & 0.710644 & 0.549232 & 0.562888 & \textbf{0.540235} \\
-    0.200000 & 1.155014 & 0.938561 & 0.946712 & \textbf{0.903213} \\
-    0.300000 & 1.417805 & 1.212671 & 1.222407 & \textbf{1.173803} \\
-    0.400000 & 1.540108 & 1.399925 & 1.409293 & \textbf{1.367943} \\
-    0.500000 & 1.574957 & 1.517281 & 1.484589 & \textbf{1.455849} \\
-    0.600000 & 1.525114 & 1.498608 & 1.495474 & 1.\textbf{447470} \\
-    0.700000 & 1.397918 & 1.372183 & 1.362173 & \textbf{1.331503} \\
-    0.800000 & 1.170140 & 1.115077 & 1.123195 & \textbf{1.096606} \\
-    0.900000 & 0.787195 & \textbf{0.714075} & 0.734112 & 0.715085 \\
+    0.100000 & 0.710644 & 0.549232 & 0.562888 & \textbf{0.540145} \\
+    0.200000 & 1.155014 & 0.938561 & 0.946712 & \textbf{0.903193} \\
+    0.300000 & 1.417805 & 1.212671 & 1.222407 & \textbf{1.173627} \\
+    0.400000 & 1.540108 & 1.399925 & 1.409293 & \textbf{1.368022} \\
+    0.500000 & 1.574957 & 1.517281 & 1.484589 & \textbf{1.456053} \\
+    0.600000 & 1.525114 & 1.498608 & 1.495474 & \textbf{1.447470} \\
+    0.700000 & 1.397918 & 1.372183 & 1.362173 & \textbf{1.331375} \\
+    0.800000 & 1.170140 & 1.115077 & 1.123195 & \textbf{1.096649} \\
+    0.900000 & 0.787195 & \textbf{0.714075} & 0.734112 & 0.715013 \\
     \bottomrule
@@
     \midrule
-    MAE & 3.253882 & 3.134805 & 3.095041 & \textbf{3.024336} \\
+    MAE & 3.253882 & 3.134805 & 3.095041 & \textbf{3.024305} \\
     \bottomrule
@@
         \midrule
-        & Gaussian RBF &  10.031708 \\
-        & Absolute Laplacian &  10.056884 \\
-        & Gauss RBFx Absolute Laplacian & 10.150826 \\
-        & Cosine & 16.253973    \\
-        & Linear & 10.463867    \\
-        & Polynomial & 10.089147     \\
-        & Sigmoid & 16.253973            \\
-        & Chi squared & \textbf{10.023732}       \\
-        & Matern 0.5 & 10.056883  \\
-        & Matern 1.5 & 10.040288  \\
-        & Matern 2.5 & 10.024495    \\
-        & Periodic  & 15.946272\\
+        & Absolute Laplacian &  10.136551 \\
+        & Matern 0.5/Laplacian & 10.136551  \\
+        & Matern 1.5 & 10.082668  \\
+        & Matern 2.5 & 10.0494190    \\
+
+        & Matern $\infty$/Gaussian RBF &  10.031546 \\
+
+        & Chi squared & \textbf{10.018701}       \\
+
+        & Linear & 10.463752    \\
+        & Periodic  & 10.493333\\
+        & Polynomial & 10.073935     \\
+        & Sigmoid & 10.496149            \\
+        % & Gauss RBFx Absolute Laplacian & 10.150826 \\
+        & Cosine & 16.254383    \\
         \bottomrule
@@
     \midrule
-    Gauss RBF &  \textbf{0.540145} &
-    0.903193 &
+    Absolute Laplacian & 0.545490 &
+    0.913866 &
+    1.189372 &
+    1.386753 &
+    1.477606 &
+    1.468567 &
+    1.348241 &
+    1.101059 &
+    \textbf{0.705598} &
+    \\
+    Matern 0.5/Laplacian & 0.545490 &
+    0.913866 &
+    1.189372 &
+    1.386753 &
+    1.477606 &
+    1.468567 &
+    1.348241 &
+    1.101059 &
+    \textbf{0.705598} &
+\\
+    Matern 1.5 & 0.542709 &
+    0.904605 &
+    1.184761 &
+    1.382041 &
+    1.468031 &
+    1.459695 &
+    1.338982 &
+    1.096042 &
+    0.705803 &
+  \\  
+    Matern 2.5 & 0.541228 &
+    0.902478 &
+    1.175042 &
+    1.374289 &
+    1.465790 &
+    1.453434 &
+    1.334173 &
+    \textbf{1.093902} &
+    0.709083 &
+    \\
+    Matern $\infty$/Gaussian RBF & \textbf{0.540145} &
+    \textbf{0.903193} &
     1.173627 &
@@
     1.096649 &
-    0.715013 \\
-    Laplacian &  0.542499 &
-    0.906517 &
-    1.177309 &
-    1.371198 &
-    1.462111 &
-    1.453119 &
-    1.340193 &
-    1.095991 &
-    0.707947 \\
-    Gauss RBFxLaplacian &0.545787 &
-    0.915559 &
-    1.189965 &
-    1.387109 &
-    1.479901 &
-    1.475541 &
-    1.350803 &
-    1.101151 &
-    \textbf{0.704922} \\
+    0.715013 &    
+    \\
+    Chi squared & 0.540200 &
+    0.904170 &
+    \textbf{1.170647} &
+    \textbf{1.360472} &
+    \textbf{1.449636} &
+    1.446009 &
+    \textbf{1.333024} &
+    1.096450 &
+    0.718095 &
+    \\
+    Linear & 0.565663 &
+    0.947889 &
+    1.250009 &
+    1.437863 &
+    1.506723 &
+    1.478379 &
+    1.358865 &
+    1.135702 &
+    0.782660 &
+    \\
+    Periodic & 0.554232 &
+    0.930133 &
+    1.222068 &
+    1.422040 &
+    1.518273 &
+    1.509894 &
+    1.398269 &
+    1.160021 &
+    0.778402 &
+    \\
+    Polynomial & 0.542649 &
+    0.908764 &
+    1.183810 &
+    1.370962 &
+    1.452645 &
+    \textbf{1.442749} &
+    1.331137 &
+    1.103014 &
+    0.738205 &
+    \\
+    Sigmoid & 0.572588 &
+    0.948625 &
+    1.244226 &
+    1.433785 &
+    1.506071 &
+    1.480981 &
+    1.359103 &
+    1.140366 &
+    0.810406 &
+    \\
     Cosine & 0.755479 &
@@
     1.320411 \\
-    Linear &  0.565663 &
-    0.947889 &
-    1.250009 &
-    1.437863 &
-    1.506723 &
-    1.478379 &
-    1.358865 &
-    1.135702 &
-    0.782660   \\
-    Polynomial &   0.542573 &
-    0.908681 &
-    1.178357 &
-    1.362351 &
-    1.450249 &
-    1.455228 &
-    1.339822 &
-    1.107768 &
-    0.744118  \\
-    Sigmoid &   0.755479 &
-    1.343973 &
-    1.802603 &
-    2.123425 &
-    2.307123 &
-    2.367123 &
-    2.262603 &
-    1.971644 &
-    1.320411          \\
-    Chi squared &   0.541430 &
-    0.906352 &
-    1.173164 &
-    \textbf{1.358862} &
-    \textbf{1.444130} &
-    \textbf{1.440915} &
-    1.332140 &
-    1.100262 &
-    0.726625     \\
-    Matern 0.5 &  0.542499 &
-    0.906517 &
-    1.177309 &
-    1.371198 &
-    1.462111 &
-    1.453119 &
-    1.340193 &
-    1.095991 &
-    0.707947 \\
-    Matern 1.5 &  0.541410 &
-    0.902901 &
-    1.173943 &
-    1.371218 &
-    1.462212 &
-    1.449296 &
-    1.333655 &
-    \textbf{1.095092} &
-    0.710560 \\
-    Matern 2.5 &  0.541233 &
-    \textbf{0.902231} &
-    \textbf{1.172001} &
-    1.366449 &
-    1.456505 &
-    1.444124 &
-    \textbf{1.331251} &
-    1.095674 &
-    0.715029 \\
-    Periodic  & 0.759452 &
-    1.352747 &
-    1.809921 &
-    2.125765 &
-    2.314718 &
-    2.379657 &
-    2.274711 &
-    1.976768 &
-    1.328630 \\
     \bottomrule
@@
         \midrule
-        & Gauss RBF &  3.024336 \\
-        & Laplacian &  3.038633 \\
-        & Gauss RBFxLaplacian &  3.070812\\
+        & Absolute Laplacian &  3.061553 \\
+        & Matern 0.5/Laplacian &  3.061553 \\
+        & Matern 1.5 &  3.046474 \\
+        & Matern 2.5 &  3.034608 \\
+        & Matern $\infty$/Gauss RBF &  3.024305 \\
+        & Chi squared &    3.027183    \\
+        & Linear &     3.142352\\
+        & Periodic  & 3.118420 \\
+        & Polynomial &     \textbf{3.022743} \\
+        & Sigmoid &      3.108107       \\
         & Cosine &    4.849589 \\
-        & Linear &     3.142352\\
-        & Polynomial &     3.064472 \\
-        & Sigmoid &      4.849589       \\
-        & Chi squared &    3.022778    \\
-        & Matern 0.5 &  3.038633 \\
-        & Matern 1.5 &  3.023060 \\
-        & Matern 2.5 &  \textbf{3.021949} \\
-        & Periodic  & 4.880673 \\
         \bottomrule
@@
         \centering
-        \includegraphics[width=1.1\textwidth]{images/melborune_gaussian_rbf_kernel_quantile_regression.png} 
-        \caption{Gaussian RBF} 
+        \includegraphics[width=1.1\textwidth]{images/melbourne_a_laplacian_kernel_quantile_regression.png} 
+        \caption{Absolute Laplacian} 
         \label{} 
@@
         \centering
-        \includegraphics[width=1.1\textwidth]{images/melborune_laplacian_kernel_quantile_regression.png} 
-        \caption{Laplacian} 
+        \includegraphics[width=1.1\textwidth]{images/melbourne_matern_0.5_kernel_quantile_regression.png} 
+        \caption{Matern 0.5/Laplacian} 
         \label{} 
@@
         \centering
-        \includegraphics[width=1.1\textwidth]{images/melborune_gaussian_rbf_x_laplacian_kernel_quantile_regression.png} 
-        \caption{Gauss RBFxLaplacian} 
+        \includegraphics[width=1.1\textwidth]{images/melbourne_matern_1.5_kernel_quantile_regression.png} 
+        \caption{Matern 1.5} 
         \label{} 
@@
         \centering
-        \includegraphics[width=1.1\textwidth]{images/melborune_cosine_kernel_quantile_regression.png}
-        \caption{Cosine} 
+        \includegraphics[width=1.1\textwidth]{images/melbourne_matern_2.5_kernel_quantile_regression.png}
+        \caption{Matern 2.5} 
         \label{} 
@@
         \centering
-        \includegraphics[width=1.1\textwidth]{images/melborune_linear_kernel_quantile_regression.png}
+        \includegraphics[width=1.1\textwidth]{images/melbourne_gaussian_rbf_kernel_quantile_regression.png}
+        \caption{Matern $\infty$/Gaussian RBF} 
+        \label{} 
+        \vspace{4ex}
+    \end{subfigure} 
+    \begin{subfigure}[b]{0.5\linewidth}
+        \centering
+        \includegraphics[width=1.1\textwidth]{images/melbourne_chi_squared_kernel_quantile_regression.png}
+        \caption{Chi squared} 
+        \label{} 
+        \vspace{4ex}
+    \end{subfigure} 
+    \end{figure}
+\begin{figure}[!h]\ContinuedFloat
+    \begin{subfigure}[b]{0.5\linewidth}
+        \centering
+        \includegraphics[width=1.1\textwidth]{images/melbourne_linear_kernel_quantile_regression.png}
         \caption{Linear} 
@@
         \centering
-        \includegraphics[width=1.1\textwidth]{images/melborune_polynomial_kernel_quantile_regression.png}
+        \includegraphics[width=1.1\textwidth]{images/melbourne_periodic_kernel_quantile_regression.png}
+        \caption{Periodic} 
+        \label{} 
+        \vspace{4ex}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.5\linewidth}
+        \centering
+        \includegraphics[width=1.1\textwidth]{images/melbourne_polynomial_kernel_quantile_regression.png}
         \caption{Polynomial} 
@@
     \end{subfigure} 
-    \end{figure}
-
-\begin{figure}[!h]\ContinuedFloat
     \begin{subfigure}[b]{0.5\linewidth}
         \centering
-        \includegraphics[width=1.1\textwidth]{images/melborune_sigmoid_kernel_quantile_regression.png}
+        \includegraphics[width=1.1\textwidth]{images/melbourne_sigmoid_kernel_quantile_regression.png}
         \caption{Sigmoid} 
@@
         \centering
-        \includegraphics[width=1.1\textwidth]{images/melborune_chi_squared_kernel_quantile_regression.png}
-        \caption{Chi squared} 
-        \label{} 
-        \vspace{4ex}
-    \end{subfigure}
-
-    \begin{subfigure}[b]{0.5\linewidth}
-        \centering
-        \includegraphics[width=1.1\textwidth]{images/melborune_matern_0.5_kernel_quantile_regression.png}
-        \caption{Matern 0.5} 
-        \label{} 
-        \vspace{4ex}
-    \end{subfigure} 
-    \begin{subfigure}[b]{0.5\linewidth}
-        \centering
-        \includegraphics[width=1.1\textwidth]{images/melborune_matern_1.5_kernel_quantile_regression.png}
-        \caption{Matern 1.5} 
-        \label{} 
-        \vspace{4ex}
-    \end{subfigure}
-
-    \begin{subfigure}[b]{0.5\linewidth}
-        \centering
-        \includegraphics[width=1.1\textwidth]{images/melborune_matern_2.5_kernel_quantile_regression.png}
-        \caption{Matern 2.5} 
-        \label{} 
-    \end{subfigure} 
-    \begin{subfigure}[b]{0.5\linewidth}
-        \centering
-        \includegraphics[width=1.1\textwidth]{images/melborune_periodic_kernel_quantile_regression.png}
-        \caption{Periodic} 
+        \includegraphics[width=1.1\textwidth]{images/melbourne_cosine_kernel_quantile_regression.png}
+        \caption{Cosine} 
         \label{} 

### 2024-05-30 experiments/kqr_kernels_comp.py
clean

@@ -55,9 +55,9 @@ if __name__=="__main__":
     # kernel quantile regression
     qr_krn_models=[]
     y_test_pred_qr_krn=[]
-    ktype="chi_squared"
+    ktype="gaussian_rbf"
 
-    # gamma=[1e-1,1e-2,1,5,10,20]
+    # gamma=[1e-1,1e-2,1,5,10,20], var=1
     # sigma=[1e-1,1e-2,1,5,10,20]
 
     # polynomial c=[0, 1, 10, 100],
@@ -104,7 +104,7 @@ if __name__=="__main__":
     for i,q in enumerate(tqdm(quantiles)):
 
         # fit data for specific quantile
-        qr_krn_models+=[KQR(alpha=q,kernel_type=ktype, C=10, **best_hyperparameters_krn).fit(X_train, y_train)]
+        qr_krn_models+=[KQR(alpha=q,kernel_type=ktype, C=10,**best_hyperparameters_krn).fit(X_train, y_train)]
 
         print(f"{mean_pinball_loss(y_test,qr_krn_models[i].predict(X_test), alpha=q):.6f}", "&")
         print(best_hyperparameters_krn)
@@ -141,7 +141,7 @@ plt.ylabel("Today temperature")
 plt.ylim(5, 45)
 plt.yticks(np.arange(10, 45+1, 5))
 
-plt.savefig(f"experiments/plots/melborune_{ktype}_kernel_quantile_regression.png")
+plt.savefig(f"experiments/plots/melbourne_{ktype}_kernel_quantile_regression.png")
 plt.show()
 
 

### 2024-05-30 experiments/train_test/Price/price_qr_gefcom2014.py

@@ -12,7 +12,8 @@ from sklearn.preprocessing import StandardScaler
 import sys
 from tqdm import tqdm
 
-from kernel_quantile_regression.kqr import KQR
+# from kernel_quantile_regression.kqr import KQR
+from kqr import KQR
 
 
 

### 2024-05-30 experiments/train_test/Price/price_qr_gefcom2014_2.py

@@ -1,0 +1,83 @@
+import numpy as np
+import pandas as pd
+import pickle
+from pprint import pprint
+from sklearn.experimental import enable_halving_search_cv  # noqa
+from sklearn.model_selection import GridSearchCV,HalvingGridSearchCV,HalvingRandomSearchCV
+from sklearn.metrics import make_scorer
+from sklearn.metrics import mean_absolute_error
+from sklearn.metrics import mean_pinball_loss
+from sklearn.model_selection import train_test_split
+from sklearn.preprocessing import StandardScaler
+import sys
+from tqdm import tqdm
+
+# from kernel_quantile_regression.kqr import KQR
+sys.path.append("/Users/luca/Desktop/ThesisKernelMethods/experiments/")
+from kqr import KQR
+
+
+
+if __name__=="__main__":
+    ith=int(sys.argv[1])
+    
+    ktype=sys.argv[2]
+
+    task_month={4:7,5:7,6:7,7:7,8:7,9:7,10:7,11:7,12:7,13:12,14:12,15:12}
+    # load data
+    df_train=pd.read_csv(f"experiments/Data/Price/Task {ith}/Task{ith}_P_train.csv")
+    
+    df_train=df_train[(df_train["MONTH"]==task_month[ith])]
+    # define train
+    X_train=df_train[["DAY","HOUR","Forecasted Total Load","Forecasted Zonal Load"]]
+    y_train=df_train["Zonal Price"]
+
+    m=len(y_train)
+    # print(len(y_train))
+
+    # 99 quantiles
+    quantiles = [i/100 for i in range(1,100)]
+
+    # kernel quantile regression
+    scaler = StandardScaler()
+
+    X_train_scaled = scaler.fit_transform(X_train)
+
+    qr_krn_models=[]
+    y_test_pred_qr_krn=[]
+
+    param_grid_krn = dict(
+    C=[1/(m*10e-6),1/(m*10e-5),1/(m*10e-4),1/(m*10e-3),1/(m*10e-2),1/(m*10e-1),1/(m*10e-0),1/(m*10e1)],
+    gamma=[2**-6,2**-5,2**-4, 2**-3,2**-2,2**-1,1, 2, 2**2, 2**3,2**4,2**5,2**6]   
+    )
+    
+    # # define loss to tune
+    # # greater_is_better=False, we need to maximize the negative of the loss
+    neg_mean_pinball_loss_scorer = make_scorer(
+    mean_pinball_loss,
[…]

### 2024-05-30 experiments/train_test/Price/price_qr_gefcom2014_test.py

@@ -23,8 +23,8 @@ from utils.miscellaneous import price_plot_ci
 def test(ith):
 
     task_month={4:7,5:7,6:7,7:7,8:7,9:7,10:7,11:7,12:7,13:12,14:12,15:12}
-
-    df=pd.read_csv(f"Data/Price/Task {ith}/Task{ith}_P_test.csv")
+    task_model={4:4,5:4,6:4,7:4,8:4,9:4,10:4,11:4,12:4,13:13,14:13,15:13}
+    df=pd.read_csv(f"experiments/Data/Price/Task {ith}/Task{ith}_P_test.csv")
 
     X_test=df[["DAY","HOUR","Forecasted Total Load","Forecasted Zonal Load"]]
     y_test=df["Zonal Price"]
@@ -32,7 +32,7 @@ def test(ith):
     quantiles = [i/100 for i in range(1,100)]
 
     # we need to scale X_test
-    df_train=pd.read_csv(f"Data/Price/Task {ith}/Task{ith}_P_train.csv")
+    df_train=pd.read_csv(f"experiments/Data/Price/Task {ith}/Task{ith}_P_train.csv")
     df_train=df_train[(df_train["MONTH"]==task_month[ith])]
 
     X_train=df_train[["DAY","HOUR","Forecasted Total Load","Forecasted Zonal Load"]]
@@ -45,12 +45,12 @@ def test(ith):
     ktype="laplacian"
 
     # predict
-    df_template_submission=pd.read_csv(f"Data/Price/Task {ith}/Benchmark{ith}_P.csv")
+    df_template_submission=pd.read_csv(f"experiments/Data/Price/Task {ith}/Benchmark{ith}_P.csv")
 
     df_predict=df_template_submission[["ZONEID", "timestamp"]].copy()
 
     for i,q in enumerate(quantiles):
-        krn_q=pickle.load(open(f"train_test/Price/{ktype}/task {ith}/krn_qr_{i}.pkl", "rb"))
+        krn_q=pickle.load(open(f"experiments/train_test/Price/{ktype}/task {task_model[ith]}/krn_qr_{i}.pkl", "rb"))
         y_predict_q=krn_q.predict(X_test_scaled)
         
         df_predict[f"{q}"]=pd.Series(y_predict_q)
@@ -59,7 +59,7 @@ def test(ith):
     reo=order_quantiles(df_predict)
 
     # save predictions to csv
[…]

### 2024-05-30 thesis/input_files/analysis.tex
scores kernel comparison
gefcomeprice update end

@@
     \begin{adjustbox}{width=\textwidth}
-        \begin{tabular}{llllllllll}
+        \begin{tabular}{lllllllllll}
             \toprule
-    Quantile & Gaussian RBF & Laplacian & Matern 1.5 & Matern 2.5 & Linear & Periodic & Polynomial & Sigmoid & Cosine \\
+            Quantile & \begin{tabular}[c]{@{}l@{}}Absolute\\ Laplacian\end{tabular} & \begin{tabular}[c]{@{}l@{}}Matern 0.5/\\ Laplacian\end{tabular} & Matern 1.5 & Matern 2.5 & \begin{tabular}[c]{@{}l@{}}Matern $\infty$/\\ Gaussian RBF\end{tabular} & Linear & Periodic & Polynomial & Sigmoid & Cosine \\
             \midrule
-    0.1 & 0.018861 & \textbf{0.018798} & 0.019285 & 0.019510 & 0.019128 & 0.024474 & 0.021105 & 0.024299 & 0.018996 \\
-    0.2 & 0.030924 & \textbf{0.030430} & 0.031038 & 0.031438 & 0.031223 & 0.040274 & 0.033892 & 0.039774 & 0.031008 \\
-    0.3 & 0.039107 & \textbf{0.038510} & 0.039295 & 0.039751 & 0.039637 & 0.051023 & 0.042683 & 0.050242 & 0.039363 \\
-    0.4 & 0.044285 & \textbf{0.043718} & 0.044763 & 0.045143 & 0.045130 & 0.057785 & 0.048232 & 0.056815 & 0.044896 \\
-    0.5 & 0.046840 & \textbf{0.046261} & 0.047502 & 0.047957 & 0.047739 & 0.060671 & 0.051024 & 0.059697 & 0.047302 \\
-    0.6 & 0.046138 & \textbf{0.045582} & 0.047239 & 0.047686 & 0.047416 & 0.059627 & 0.050456 & 0.058824 & 0.046834 \\
-    0.7 & 0.041910 & \textbf{0.041615} & 0.043404 & 0.043964 & 0.043735 & 0.054688 & 0.046437 & 0.054037 & 0.042897 \\
-    0.8 & 0.034032 & \textbf{0.033852} & 0.035332 & 0.035997 & 0.035518 & 0.044591 & 0.037918 & 0.044112 & 0.034557 \\
-    0.9 & \textbf{0.021817} & 0.021923 & 0.022712 & 0.023082 & 0.022721 & 0.028070 & 0.024161 & 0.027920 & 0.022606 \\
+            0.1 & \textbf{0.018783} & 0.018988 & 0.019108 & 0.019257 & 0.019547 & 0.019952 & 0.018891 & 0.019654 & 0.021456 & 0.021921 \\
+            0.2 & \textbf{0.030441} & 0.030668 & 0.030809 & 0.031028 & 0.031546 & 0.032158 & 0.030683 & 0.031857 & 0.034032 & 0.034629 \\
+            0.3 & \textbf{0.038467} & 0.038864 & 0.039050 & 0.039285 & 0.039907 & 0.040657 & 0.038968 & 0.040075 & 0.042944 & 0.043393 \\
+            0.4 & \textbf{0.043622} & 0.044186 & 0.044516 & 0.044706 & 0.045286 & 0.046153 & 0.044354 & 0.045506 & 0.048701 & 0.048977 \\
+            0.5 & \textbf{0.046160} & 0.046792 & 0.047205 & 0.047450 & 0.048116 & 0.048951 & 0.046956 & 0.047891 & 0.051423 & 0.051896 \\
+            0.6 & \textbf{0.045499} & 0.046133 & 0.046824 & 0.047177 & 0.047840 & 0.048667 & 0.046446 & 0.047496 & 0.051292 & 0.051894 \\
+            0.7 & \textbf{0.041494} & 0.042044 & 0.042928 & 0.043371 & 0.044104 & 0.044926 & 0.042458 & 0.043565 & 0.047656 & 0.048275 \\
+            0.8 & \textbf{0.033837} & 0.034128 & 0.034797 & 0.035325 & 0.036166 & 0.037055 & 0.034290 & 0.035533 & 0.039501 & 0.040007 \\
+            0.9 & \textbf{0.021883} & 0.021871 & 0.022431 & 0.022682 & 0.023169 & 0.023730 & 0.022000 & 0.022811 & 0.025561 & 0.026019 \\
             \midrule
-    CRPS & 0.035990 & \textbf{0.035632} & 0.036730 & 0.037170 & 0.036916 & 0.046800 & 0.039545 & 0.046191 & 0.036495 \\
+            CRPS & \textbf{0.035576} & 0.035964 & 0.036407 & 0.036698 & 0.037298 & 0.038028 & 0.036116 & 0.037154 & 0.040285 & 0.040779 \\
             \bottomrule
@@
         \midrule
-        0.1 & 0.027483 & 0.027827 & 0.028236 & 0.028257 & 0.028057 & 0.028261 & 0.028277 & 0.028276 & 0.028146 & 0.028004 \\
-        0.2 & 0.046051 & 0.046679 & 0.047531 & 0.047544 & 0.047118 & 0.047569 & 0.047595 & 0.047596 & 0.047353 & 0.047025 \\
-        0.3 & 0.058739 & 0.059835 & 0.060892 & 0.060940 & 0.060366 & 0.060962 & 0.060988 & 0.060984 & 0.060654 & 0.060209 \\
-        0.4 & 0.066347 & 0.067768 & 0.068952 & 0.069022 & 0.068344 & 0.069054 & 0.069104 & 0.069101 & 0.068634 & 0.068202 \\
-        0.5 & 0.069118 & 0.070761 & 0.072188 & 0.072250 & 0.071450 & 0.072285 & 0.072331 & 0.072330 & 0.071807 & 0.071359 \\
-        0.6 & 0.067098 & 0.068276 & 0.069454 & 0.069526 & 0.068821 & 0.069571 & 0.069615 & 0.069613 & 0.069157 & 0.068780 \\
-        0.7 & 0.059972 & 0.060783 & 0.061382 & 0.061410 & 0.061150 & 0.061427 & 0.061442 & 0.061442 & 0.061281 & 0.060990 \\
-        0.8 & 0.046457 & 0.046944 & 0.047325 & 0.047331 & 0.047147 & 0.047340 & 0.047348 & 0.047349 & 0.047247 & 0.047056 \\
-        0.9 & 0.026343 & 0.026554 & 0.026749 & 0.026757 & 0.026646 & 0.026763 & 0.026772 & 0.026772 & 0.026282 & 0.026628 \\
+        0.1 & \textbf{0.025734} & 0.026272 & 0.026642 & 0.026845 & 0.027090 & 0.027306 & 0.026139 & 0.026621 & 0.027954 & 0.028004 \\
+        0.2 & \textbf{0.043114} & 0.044138 & 0.044733 & 0.045057 & 0.045469 & 0.045917 & 0.044088 & 0.044716 & 0.046925 & 0.047025 \\
+        0.3 & \textbf{0.054861} & 0.056343 & 0.056930 & 0.057335 & 0.057920 & 0.058528 & 0.056229 & 0.057066 & 0.060092 & 0.060209 \\
+        0.4 & \textbf{0.061436} & 0.063490 & 0.064291 & 0.064748 & 0.065433 & 0.066136 & 0.063324 & 0.064683 & 0.067979 & 0.068202 \\
+        0.5 & \textbf{0.064144} & 0.066330 & 0.067229 & 0.067696 & 0.068275 & 0.068960 & 0.066325 & 0.071421 & 0.070957 & 0.071359 \\
+        0.6 & \textbf{0.062306} & 0.064676 & 0.065836 & 0.066299 & 0.066881 & 0.067405 & 0.064660 & 0.069160 & 0.068494 & 0.068780 \\
+        0.7 & \textbf{0.055491} & 0.057851 & 0.058848 & 0.059317 & 0.059878 & 0.060275 & 0.061442 & 0.061423 & 0.060988 & 0.060990 \\
+        0.8 & \textbf{0.044023} & 0.045011 & 0.045441 & 0.045644 & 0.046008 & 0.046374 & 0.047348 & 0.047182 & 0.047071 & 0.047056 \\
+        0.9 & 0.026178 & \textbf{0.026126} & 0.026160 & 0.026187 & 0.026207 & 0.026282 & 0.026772 & 0.026766 & 0.026581 & 0.026628 \\
         \midrule
-        CRPS & 0.051956 & 0.052825 & 0.053634 & 0.053671 & 0.053233 & 0.053693 & 0.053719 & 0.053718 & 0.053395 & 0.053139 \\
+        CRPS & \textbf{0.048587} & 0.050026 & 0.050679 & 0.051014 & 0.051462 & 0.051909 & 0.050703 & 0.052115 & 0.053005 & 0.053139 \\
         \bottomrule
@@
         \midrule
-    0.100000 & 0.023882 & 0.027901 & 0.024316 & 0.024611 & 0.024309 & 0.029367 & 0.029462 & 0.027319 & 0.029459 & 0.027816 \\
-    0.200000 & 0.040318 & 0.046043 & 0.040987 & 0.041487 & 0.041028 & 0.048510 & 0.048716 & 0.046164 & 0.048714 & 0.045879 \\
-    0.300000 & 0.051643 & 0.058812 & 0.052548 & 0.052972 & 0.052597 & 0.062242 & 0.062561 & 0.059133 & 0.062548 & 0.058454 \\
-    0.400000 & 0.057676 & 0.066704 & 0.058841 & 0.059078 & 0.058803 & 0.070210 & 0.070577 & 0.067914 & 0.070567 & 0.066335 \\
-    0.500000 & 0.059441 & 0.071401 & 0.060912 & 0.061541 & 0.060846 & 0.073399 & 0.073782 & 0.071616 & 0.073764 & 0.069574 \\
-    0.600000 & 0.057820 & 0.067650 & 0.059237 & 0.059930 & 0.059336 & 0.070817 & 0.071157 & 0.070249 & 0.071134 & 0.066970 \\
-    0.700000 & 0.052402 & 0.059061 & 0.053949 & 0.054372 & 0.053937 & 0.062453 & 0.062688 & 0.062970 & 0.062675 & 0.059419 \\
-    0.800000 & 0.042427 & 0.047954 & 0.043364 & 0.043510 & 0.043366 & 0.049024 & 0.049150 & 0.049737 & 0.049147 & 0.046662 \\
-    0.900000 & 0.026227 & 0.028459 & 0.026645 & 0.026817 & 0.026727 & 0.029516 & 0.029601 & 0.030097 & 0.029598 & 0.027986 \\
+        0.1 & \textbf{0.024365} & 0.024592 & 0.024550 & 0.024611 & 0.024831 & 0.026853 & 0.025744 & 0.026004 & 0.027672 & 0.027816 \\
+        0.2 & \textbf{0.040971} & 0.041406 & 0.041413 & 0.041487 & 0.041857 & 0.045007 & 0.043247 & 0.043720 & 0.045899 & 0.045879 \\
+        0.3 & \textbf{0.052316} & 0.052969 & 0.052888 & 0.052973 & 0.053449 & 0.057880 & 0.055231 & 0.055446 & 0.058572 & 0.058454 \\
+        0.4 & \textbf{0.058286} & 0.059018 & 0.058916 & 0.059078 & 0.059592 & 0.065770 & 0.061844 & 0.062495 & 0.066599 & 0.066335 \\
+        0.5 & 0.060344 & 0.061100 & 0.061319 & 0.061541 & 0.062338 & 0.068197 & 0.063864 & 0.071731 & 0.069761 & 0.069574 \\
+        0.6 & \textbf{0.058724} & 0.059507 & 0.059674 & 0.059930 & 0.060709 & 0.065828 & 0.062265 & 0.070269 & 0.067179 & 0.066970 \\
+        0.7 & \textbf{0.053148} & 0.054189 & 0.054249 & 0.054372 & 0.054916 & 0.058548 & 0.055830 & 0.063015 & 0.059807 & 0.059419 \\
+        0.8 & \textbf{0.042740} & 0.043368 & 0.043408 & 0.043510 & 0.043881 & 0.046017 & 0.044384 & 0.049813 & 0.047078 & 0.046662 \\
+        0.9 & \textbf{0.026430} & 0.026583 & 0.026781 & 0.026817 & 0.026908 & 0.027050 & 0.026471 & 0.030279 & 0.028031 & 0.027986 \\
         \midrule
-    CRPS & 0.045760 & 0.052665 & 0.046755 & 0.047146 & 0.046772 & 0.055060 & 0.055299 & 0.053911 & 0.055290 & 0.052122 \\
+        CRPS & \textbf{0.046369} & 0.046970 & 0.047022 & 0.047146 & 0.047609 & 0.051239 & 0.048764 & 0.052530 & 0.052289 & 0.052122 \\
         \bottomrule
@@
 In this track, the top entries came from the teams: Tololo, Team Poland, GMD, and C3 Green Team; for a breakdown of each method's attributes, see \cite[Table 8]{hong2016probabilistic}.
-Finally, the Gaussian RBF probabilistic prediction for the $13^{th}$ July $2013$ zonal price at the $90\%$ confidence interval is visualized in Figure \ref{fig:price_task_6_gaussian_rbf}.
+Finally, the Gaussian RBF probabilistic prediction for the $4^{th}$ July $2013$ zonal price at the $90\%$ confidence interval is visualized in Figure \ref{fig:price_task_4_gaussian_rbf}.
 
-\begin{figure}[!ht]
-    \centering
-    \includegraphics[width=\linewidth]{images/price_task_6_gaussian_rbf.png}
-    \caption[Prediction price track task 6, Gaussian RBF kernel]{Price 90\% confidence interval task 6.
+% \begin{figure}[!ht]
+%     \centering
+%     \includegraphics[width=\linewidth]{images/price_task_6_gaussian_rbf.png}
+%     \caption[Prediction price track task 6, Gaussian RBF kernel]{Price 90\% confidence interval task 6.
 Electricity price probabilistic forecast for the $13^{\text{th}}$ July $2013$.
@@
 Lower and upper red lines denote the $95\%$ and $5\%$ quantile forecast respectively}
+%     % \Description{}
+%     \label{fig:price_task_6_gaussian_rbf}
+% \end{figure}
+
+\begin{figure}[!ht]
+    \centering
+    \includegraphics[width=\linewidth]{images/price_task_4_gaussian_rbf.png}
+    \caption[Prediction price track task 4, Gaussian RBF kernel]{Price 90\% confidence interval task 4.
+Electricity price probabilistic forecast for the $4^{\text{th}}$ July $2013$.
+The black line is the observed path for the price.
+The $90\%$ confidence interval bands are plotted in green.
+Lower and upper red lines denote the $95\%$ and $5\%$ quantile forecast respectively}
     % \Description{}
-    \label{fig:price_task_6_gaussian_rbf}
+    \label{fig:price_task_4_gaussian_rbf}
 \end{figure}
@@
       Team name\textbackslash Task number                       & 1                               & 2                                  & 3                               & 4                              & 5                              & 6       & 7                               & 8       & 9       & 10                             & 11                             & 12               \\
-  KQR Laplacian
+  KQR Absolute Laplacian
   &
-  \textbf{1.3249}
+  \textbf{1.02492}
   &
-  2.9497
+  3.35057
   &
-  0.8868
+  4.21374
   &
-  7.3780
+  7.33987
   &
-  5.4314
+  5.00981
   &
-  8.2166
+  6.96522
   &
-  4.9091
+  3.57168
   &
-  1.5245
+  1.77610
   &
-  1.9601
+  1.28765
   &
-  2.7324
+  2.73863
   &
-  2.4733
+  2.39831
   &
-  17.5371
+  23.30234
   \\
@@
   &
-  2.0614
+  1.84673
   &
-  3.2247
+  2.81882
   &
-  \textbf{0.7050}
+  1.54608
   &
-  6.1407
+  8.31636
   &
-  6.1134
+  4.05988
   &
-  8.5969
+  6.60456
   &
-  2.9360
+  3.57818
   &
-  1.5731
+  2.02177
   &
-  1.8830
+  1.45779
   &
-  2.6419
+  2.20701
   &
-  2.4814
+  1.98184
   &
-  16.1921
+  21.41033
   \\

### 2024-05-30 experiments/train_test/Price/price_qr_gefcom2014_test.py

@@ -42,7 +42,7 @@ def test(ith):
 
     pinball_tot=0
 
-    ktype="laplacian"
+    ktype="gaussian_rbf"
 
     # predict
     df_template_submission=pd.read_csv(f"experiments/Data/Price/Task {ith}/Benchmark{ith}_P.csv")
@@ -71,15 +71,18 @@ def test(ith):
 
     # plot
     price_plot_ci(reo, y_test)
-    plt.title(f"Task {ith}, {ktype} kernel")
-    # plt.savefig(f"plots/Price/price_task_{ith}_{ktype}.png")
+    plt.title(f"Task {ith}, {title_label[ktype]} kernel")
+    plt.savefig(f"experiments/plots/Price/price_task_{ith}_{ktype}.png")
     plt.legend()
     plt.show()
 
     ans=pinball_tot/len(quantiles)
-    print("total quantile: ", ans)
+    print("total quantile: ", f"{ans:.5f}")
     
     return ans
+
+title_label={"gaussian_rbf": "Gaussian RBF",
+             "a_laplacian": "Absolute Laplacian"}
 
 if __name__=="__main__":
     i=int(sys.argv[1])

### 2024-05-30 experiments/train_test/Price/rest.sh

@@ -1,0 +1,7 @@
+#!/usr/bin/bash
+
+# python experiments/train_test/Price/price_qr_gefcom2014_2.py 13 "a_laplacian"
+
+
+python experiments/train_test/Price/price_qr_gefcom2014_2.py 4 "gaussian_rbf"
+python experiments/train_test/Price/price_qr_gefcom2014_2.py 13 "gaussian_rbf"

### 2024-05-30 thesis/input_files/analysis.tex
May 30th 2024

@@
 Their paper showed evidence of the superiority of the Gaussian RBF among their set of kernel functions.
-Analogously, we can see from Tables \ref{tab:secures_met_de}, \ref{tab:secures_met_ch} and \ref{tab:secures_met_at} that the Gaussian RBF is the second best in terms of pinball loss.
+Analogously, we can see from Tables \ref{tab:secures_met_de}, \ref{tab:secures_met_ch} and \ref{tab:secures_met_at} that the Matern kernels are the best in terms of pinball loss.
+Finally, given that the features dimension is greater than one we have that the Absolute Laplacian and the Matern 0.5 kernels differ.
+The former kernel is better than the latter and this comes from their distance function.
+This is because the Manhattan distance is more robust than the Euclidean distance in terms of measuring similarity between objects as data grows in dimensionality \cite{Aggarwal2001}.
+%(curse of dimensionality)
+Those kernel matrices for the Swiss data are compared in Figure \ref{fig:k_matrix_laplacian_comp}.
+
+\begin{figure}[!ht]
+    \includegraphics[width=\textwidth]{images/k_matrix_laplacian_comp.png}
+    \caption{Absolute Laplacian versus Matern 0.5 kernel matrix comparison}
+    \label{fig:k_matrix_laplacian_comp}
+\end{figure}
 
@@
 Managalova, Ziel Florian, and Bidong Liu; for a breakdown of the attributes of each method, see \cite[Table 6]{hong2016probabilistic}.
-Finally, the $90\%$ confidence interval forecast by our model equipped with the absolute Laplacian kernel for task number $9$, that is the prediction for June, is reported in Figure \ref{fig:load_task_9}.
+Finally, the $90\%$ confidence interval forecast by our model equipped with the Absolute Laplacian kernel for task number $9$, that is the prediction for June, are reported in Figure \ref{fig:load_task_9} and Figuree \ref{fig:boxplot_load_task_9}.
 Similarly to the price track, we have that KQR was among the best algorithms also in the load forecasting track.
@@
     \caption{Boxplot for first two days of probabilistic forecast task 9 using KQR Laplacian}
-    \label{fig:load_task_9}
+    \label{fig:boxplot_load_task_9}
 \end{figure*}
@@
 
-  KQR Laplacian
+  KQR Absolute Laplacian
   & 

### 2024-05-31 thesis/appendix.tex
May 31st 2024

@@
 \begin{table}
-\caption{Pinball loss for Boston housing data}
+\caption{Continous ranked probability score for Boston housing data}
 \begin{tabular}{lllll}
@@
 \toprule
-    & Linear qr & Gbm qr & Quantile forest & KQR \\
+  Quantile  & Linear qr & Gbm qr & Quantile forest & KQR \\
 \midrule
@@
 \begin{table}
-    \caption{Pinball loss for Abalone data}
+    \caption{Continous ranked probability score for Abalone data}
 \begin{tabular}{lllll}
@@
     \toprule
-     & Linear qr & Gbm qr & Quantile forest & KQR \\
+    Quantile & Linear qr & Gbm qr & Quantile forest & KQR \\
     \midrule
@@
 \begin{table}
-\caption{Pinball loss for Vehicle data}
+\caption{Continous ranked probability score for Vehicle data}
 \begin{tabular}{lllll}

### 2024-05-31 thesis/input_files/acknowledgements.tex
May 31st 2024

@@
 % - supervisor, co
-First and foremost I would like to express my deepest gratitude to my advisor and co-advisor, Prof.
-Multerer and Dr.
-Baroli for their invaluable feedback, continuous support and patience during my Master's Thesis.
+First and foremost I would like to express my deepest gratitude to my advisor and co-advisor, Prof.\ Multerer and Dr.\ Baroli for their invaluable feedback, continuous support and patience during my Master's Thesis.
 Furthermore, I thank them for having given me the opportunity to collaborate through this thesis work.
@@
 % - master director
-I could not have undertaken this journey without my Master directors Prof. Schenk and Prof. Wit. I am extremely grateful for their guidance, mentorship and advice throughout my Master studies.
+I could not have undertaken this journey without my Master directors Prof.\ Schenk and Prof.\ Wit. I am extremely grateful for their guidance, mentorship and advice throughout my Master studies.
 \\
@@
 % listening the challenges I faced and insightful suggestions
-I also had great pleasure of discussing with the research group of Prof. Multerer the challenges I faced during this thesis work. I thank them for their affability  and insightful suggestions.
+I also had great pleasure of discussing with the research group of Prof.\ Multerer the challenges I faced during this thesis work. I thank them for their affability  and insightful suggestions.
 \\

### 2024-05-31 thesis/input_files/analysis.tex
May 31st 2024

@@
 Analogously, we can see from Tables \ref{tab:secures_met_de}, \ref{tab:secures_met_ch} and \ref{tab:secures_met_at} that the Matern kernels are the best in terms of pinball loss.
+The Periodic equipped with period $p=24$ shows good performance. 
+After those kernels, we have the Linear and Polynomial kernel.
+The worst among all kernels considered are the Cosine and the Sigmoid.
+% , which is better suited for binary classification tasks. 
 Finally, given that the features dimension is greater than one we have that the Absolute Laplacian and the Matern 0.5 kernels differ.
@@
 In this track, the top entries came from the teams: Tololo, Team Poland, GMD, and C3 Green Team; for a breakdown of each method's attributes, see \cite[Table 8]{hong2016probabilistic}.
-Finally, the Gaussian RBF probabilistic prediction for the $4^{th}$ July $2013$ zonal price at the $90\%$ confidence interval is visualized in Figure \ref{fig:price_task_4_gaussian_rbf}.
+The Gaussian RBF probabilistic prediction for the $4^{th}$ July $2013$ zonal price at the $90\%$ confidence interval is visualized in Figure \ref{fig:price_task_4_gaussian_rbf}.
+Cross validation results are visualised in Figure.
+
+\begin{figure}
+    \includegraphics[width=\textwidth]{images/cv_scatter.png}
+    \label{fig:cv_scatter}
+    \caption{Cross validation Gaussian RBF kernel task 4 GEFCom2014 price}
+\end{figure}
 

### 2024-05-31 thesis/input_files/implementation.tex
May 31st 2024

@@
 Functions supported by our kernel quantile regression include: Gaussian RBF, absolute Laplacian, Matern 3/2, Matern 5/2, Linear, Cosine, Sigmoid, Periodic, Polynomial and custom composition of kernels.
+Kernels from these two submodules use a different convention in the hyperparameters.
+The former has the lenghtscale parameter in the numerator while the latter in the denominator.
+In order to meaningfully compare across different kernels we made them consistent in this aspect.
+For this purpose we adhere to the convention of having lenghtscale hyperparameters in the denominator.
+
 % NOTICE, these kernel implementations differ in the form of passed parameters in the former we pass in the numerator gamma while in the latter we pass l, which stands for length scale in the denominator.

### 2024-05-31 thesis/input_files/paper1.tex
May 31st 2024

@@
 
-\section{Kernel families}
+\section{Kernel families}\label{kernel families}
 The selection of the kernel determines the class of functions that will be searched by the learning algorithm.
@@
 
-When $\nu=\frac{1}{2}$ the Matérn kernel corresponds to the Laplacian kernel.
+When $\nu=\frac{1}{2}$ the Matérn kernel corresponds to the Laplacian kernel. If the predictors are one dimensional, then the Absolute Laplacian and the Matern 0.5 are equivalent. %L2 norm=L1 norm
 Furthermore, as $\nu \to \infty$ we have that the Matérn kernel converges to the Gaussian RBF kernel.

### 2024-05-31 thesis/input_files/probabilistic.tex
May 31st 2024

@@
 
-We conclude this chapter with reporting the same kind of Tables \ref{tab:kernel pinball comparison}, \ref{tab:kernel pinball comparison quantile-wise}, \ref{tab:kernel mae comparison} and Figure \ref{fig:kernel quantile regressors comparison} comparing various kernel functions.
+We conclude this chapter with reporting the same kind of Tables \ref{tab:kernel pinball comparison}, \ref{tab:kernel pinball comparison quantile-wise}, \ref{tab:kernel mae comparison} and Figure \ref{fig:kernel quantile regressors comparison} comparing various kernel functions, see Section \ref{kernel families}.
 
+From these tables, we can conclude that the Matern and the Chi squared kernels are the top performers on the Melbourne weather data.
+Finally, the equaility in the scores of the Absolute Laplacian and the Matern 0.5 follows from the fact that regressors are one dimensional for the Melbourne weather data.
 % tables

### 2024-06-02 thesis/appendix.tex
June 2nd 2024

@@
 \chapter{Appendix}
-\section{Feature map normalization}\label{appendix:new_feature}
+\section{Feature map normalisation}\label{appendix:new_feature}
 
@@
 \begin{itemize}
-\item CRIM per capita crime rate by town
-\item ZN proportion of residential land zoned for lots over 25,000 sq.ft.\
-\item INDUS proportion of non-retail business acres per town
-\item CHAS Charles River dummy variable (1 if tract bounds river; 0 otherwise)
-\item NOX nitric oxides concentration (parts per 10 million)
-\item RM average number of rooms per dwelling
-\item AGE proportion of owner-occupied units built prior to 1940
-\item DIS weighted distances to five Boston employment centres
-\item RAD index of accessibility to radial highways
-\item TAX full-value property-tax rate per 10,000
-\item PTRATIO pupil-teacher ratio by town
-\item B $1000(Bk - 0.63)^2$ where Bk is the proportion of afroamericans by town
-\item LSTAT lower status of the population
+\item CRIM: per capita crime rate by town
+\item ZN: proportion of residential land zoned for lots over 25,000 sq.ft.\
+\item INDUS: proportion of non-retail business acres per town
+\item CHAS: Charles River dummy variable (1 if tract bounds river, 0 otherwise)
+\item NOX: nitric oxides concentration (parts per 10 million)
+\item RM: average number of rooms per dwelling
+\item AGE: proportion of owner-occupied units built prior to 1940
+\item DIS: weighted distances to five Boston employment centres
+\item RAD: index of accessibility to radial highways
+\item TAX: full-value property-tax rate per 10,000
+\item PTRATIO: pupil-teacher ratio by town
+\item B: $1000(Bk - 0.63)^2$ where Bk is the proportion of afroamericans by town
+\item LSTAT: lower status of the population
 \end{itemize}
@@
 \midrule
- & 3.826326 & 2.845989 & 2.965490 & \textbf{2.810494} \\
+MAE & 3.826326 & 2.845989 & 2.965490 & \textbf{2.810494} \\
 \bottomrule
@@
 \subsection{Abalone dataset}
-The abalone data \href{https://archive.ics.uci.edu/dataset/1/abalone}{https://archive.ics.uci.edu/dataset/1/abalone} consist of measurements of abalone molluscs, the goal is predicting their age by building a model for estimating its number of rings; age is the number of rings plus 1.5
+The abalone data \href{https://archive.ics.uci.edu/dataset/1/abalone}{https://archive.ics.uci.edu/dataset/1/abalone} consistst of measurements of abalone molluscs, the goal is predicting their age by building a model for estimating its number of rings; age is the number of rings plus 1.5.
 The data has 8 attributes:
 \begin{itemize}
-    \item Sex Categorical variable either male, female or infant
+    \item Sex: categorical variable either male, female or infant
     \item Length
@@
     \midrule
-     & 1.627627 & 1.574179 & 1.499522 & \textbf{1.498583} \\
+    MAE & 1.627627 & 1.574179 & 1.499522 & \textbf{1.498583} \\
     \bottomrule
@@
     \item Year
-    \item Present\_price ex showroom price
+    \item Present\_price: ex showroom price
     \item Kms Driven
@@
     \item Transmission
-    \item Owner number of previous owners
+    \item Owner: number of previous owners
 \end{itemize}
@@
     \midrule
-     & 1.117714 & 0.606971 & 0.752197 & \textbf{0.594292} \\
+   MAE  & 1.117714 & 0.606971 & 0.752197 & \textbf{0.594292} \\
     \bottomrule
@@
 \begin{equation}
-    \mathbb{E}_{\Tau}=\mathbb{E}\left[L\left(Y,\hat{f}(X)\right)| \Tau \right]
+    Err_{\Tau}=\mathbb{E}\left[L\left(Y,\hat{f}(X)\right)| \Tau \right]
 \end{equation}
-that is, $Err$ is the average over everything that is random: X,Y and the training set $\Tau$ used to learn $\hat{f}$.
+that is, $Err$ is the average over everything that is random: $X,Y$ and the training set $\Tau$ used to learn $\hat{f}$.
 Hence, our interest lies in estimating the $Err$ quantity in order to guide model selection.
@@
 K-fold cross validation splits the data into $K$ roughly equally sized parts, then $K$ models are trained.
-For $k$ from 1 to $K$ we train the kth model on the whole dataset except the kth partition.
-Next,  the prediction error of the kth fitted model is computed.
+For $k$ from 1 to $K$ we train the $k^{th}$ model on the whole dataset except the $k^{th}$ partition.
+Next,  the prediction error of the $k^{th}$ fitted model is computed.
 Finally averaging all the $k$ prediction erros we obtain an estimate for the expected test error.
@@
     \includegraphics[width=\textwidth]{images/crossvalidationtimeseries2.png}
-    \caption{Cross validation for m step ahead timeseries data \cite{hyndman2018forecasting}}
+    \caption{Cross validation for $m$ step ahead timeseries data \cite{hyndman2018forecasting}}
     \label{fig:crossvalidationtimeseries2}
@@
 
-\subsection{Data normalization}\label{appendix:normalization}
+\subsection{Data normalisation}\label{appendix:normalization}
 With data normalization we transform the range of the feautures to a standard scale.
-Such preprocessing step is essential when employing distance based algorithms like SVM or K-nearest neighbors.
+Such preprocessing step is essential when employing distance based algorithms like SVM or K-nearest neighbours.
 The rationale behind it is that by normalizing data we give an uniform weight to each feature in the learning process; in this way we do not favour larger scale features.
@@
 \begin{itemize}
-    \item Standard scaler= it computes the standard score $z$ of a sample $x$
+    \item Standard scaler: it computes the standard score $z$ of a sample $x$
     \\
     $z=\frac{x-\mu}{\sigma}$.
-    \item MinMax scaler= it maps every data sample to the range $[0,1]$.
+    \item MinMax scaler: it maps every data sample to the range $[0,1]$.
     \\
     $\frac{x-\min(X)}{\max(X)-\min(X)}$
-    \item Robust scaler= it scales features using statistics that are robust to outliers.
+    \item Robust scaler: it scales features using statistics that are robust to outliers.
     Essentially, it subtracts the median and then scales the data according to the interquantile range.
@@
 We conclude this subsection with a custom example that motivates the need of feature scaling \citeW{scikitscale_example}.
-The idea is to compare the results of modelling the data with K-nearest neighbors on the unscaled data against the scaled data.
+The idea is to compare the results of modelling the data with K-nearest neighbours on the unscaled data against the scaled data.
 The considered data is the wine recognition dataset \href{https://archive.ics.uci.edu/dataset/109/wine}{https://archive.ics.uci.edu/da-taset/109/wine}.
@@
     \includegraphics[width=\textwidth]{images/feature_scaler_example.png}
-    \caption{Importance of feature scaling}
+    \caption[Importance of feature scaling]{Importance of feature scaling \citeW{scikitscale_example}}
     \label{fig:feature_scaler_example}
@@
 On the left, we can see that distances between categories are impacted solely by the larger scale feature.
-Conversely, on the right, we have that the two features contribute equally in determining the neighbors.
+Conversely, on the right, we have that the two features contribute equally in determining the neighbours.
 
@@
     \item query/: folder containing Scopus data and scripts to generate bibliometric survey plots in Section \ref{literature_review}.
-    \item thesis/: folder containing tex files for thesis.
-    \item beamer/: folder containing tex files for pptx thesis defence.
+    \item thesis/: folder containing tex files for thesis pdf.
+    \item beamer/: folder containing tex files for beamer pptx %pptx thesis defence.
     \item experiments: folder containing train, test script for each of the experiments carried out in the thesis alongside with the binary files of the pickled models.

### 2024-06-02 thesis/input_files/analysis.tex
June 2nd 2024

@@
 
-The resulting RMSE is 30.59.
+The resulting RMSE on task 1 is 30.59.
 What can be concluded, is that multiple linear regression is capable of catching the daily seasonality. Nevertheless, it cannot catch the range of the price series properly.
@@
 Its name stands for trigonometric seasonality, Box-Cox transformation, ARMA errors, trend and seasonal components.
-Tbats forecast is visualized in Figure \ref{fig:tbats_price}, meanwhile its RMSE is 15.08.
+Tbats forecast is visualized in Figure \ref{fig:tbats_price}, meanwhile its RMSE on task 1 is 15.08.
 \begin{figure}[!h]
@@
 We get started by considering the base implementation. In such setting, prophet takes as input the time series object and learns its data generating process.
-This method achieves a RMSE of 23.96, its prediction is visualized in Figure \ref{fig:prophet_price_1}.
+This method achieves on task 1 a RMSE of 23.96, its prediction is visualized in Figure \ref{fig:prophet_price_1}.
 \begin{figure}[!h]
@@
 Furthermore, we also applied a log transformation to the dependent variables.
-Doing so, RMSE went down to 10.29.
+Doing so, RMSE on task 1 went down to 10.29.
 Forecast is visualised in Figure \ref{fig:prophet_price2}.
@@
 In doing so, we cross validated the number of neighbours to get the best model instance.
-The forecast is visualised in Figure \ref{fig:knn_price}, RMSE is 12.54.
+The forecast is visualised in Figure \ref{fig:knn_price}, RMSE on task 1 is 12.54.
 \begin{figure}[!h]
     \includegraphics[width=\textwidth]{images/knn_price.png}
-    \caption{K-nearest neighbour regression}
+    \caption{K-nearest neighbours regression}
     \label{fig:knn_price}
 \end{figure}
-What it can be said is that, K-nearest neighbour is capable of predicting load well by averaging past data.
+What it can be said is that, K-nearest neighbours is capable of predicting load well by averaging past data.
 
@@
 In applying support vector regression we used gridsearch crossvalidation to search for the best regularisation parameter $C$.
-Support vector regression achieves an RMSE of 23.24, for the prediction see Figure \ref{fig:svr_price}.
+Support vector regression achieves on task 1 an RMSE of 23.24, for the prediction see Figure \ref{fig:svr_price}.
 \begin{figure}[!h]
@@
 
-\subsection{Long short term memory}
+\subsection{Long short term memory}\label{sec:lstm point}
 We used the \href{https://pytorch.org}{torch} library in order to code our LSTM regressor.
@@
 \end{itemize}
-LSTM forecast is reported in Figure \ref{fig:lstm_price}, it achieves a RMSE of 9.1782
+LSTM forecast is reported in Figure \ref{fig:lstm_price}, it achieves on task 1 a RMSE of 9.1782
 \begin{figure}[!h]
@@
 Cross validation was carried jointly over the RBF kernel bandwith and the regularisation constant.
-The RMSE achieved is 9.86, Figure \ref{fig:krnridge_price} reports the prediction.
+The RMSE achieved on task 1 is 9.86, Figure \ref{fig:krnridge_price} reports the prediction.
 \begin{figure}[!h]
@@
 The last model we considered in this setting was the kernel support vector regression.
-RMSE achieved with this method is 9.10, forecast is reported in Figure \ref{fig:krnsvr_price}.
+RMSE achieved with this method on task 1 is 9.10, forecast is reported in Figure \ref{fig:krnsvr_price}.
 
@@
 \subsection{Results}
-This section reports the tables comparing the considered methods scores.
+This section reports the tables comparing the considered methods' scores.
 % in the context of electricity load forecasting.
@@
 In this study, we trained our models on the entire 2021 data and then tested them and measured their score on the 2022 data.
-Notice, for kernel quantile regression (KQR), the kernel of choice in this setting is the absolute Laplacian.
+Notice, for kernel quantile regression (KQR), the kernel of choice in this setting is the Absolute Laplacian.
 Additionally notice, the loss function utilized in this section is the normalized pinball loss.
@@
         \makebox[\linewidth]{\includegraphics[width = 1.5\textwidth]{images/DE_load_CI.png}}
-        \caption[Probabilistic forecast for load in Germany]{Load 90\% confidence interval for Germany Energy charts (2022) data using KQR Laplacian.
+        \caption[Probabilistic forecast for load in Germany]{Load 90\% confidence interval for Germany Energy charts (2022) data using KQR Absolute Laplacian.
 The black line is the observed path for the load.
@@
         \makebox[\linewidth]{\includegraphics[width = 1.5\textwidth]{images/CH_load_CI.png}}
-        \caption[Probabilistic forecast for load in Switzerland]{Load 90\% confidence interval for Switzerland Energy charts data (2022) using KQR Laplacian.
+        \caption[Probabilistic forecast for load in Switzerland]{Load 90\% confidence interval for Switzerland Energy charts data (2022) using KQR Absolute Laplacian.
 The black line is the observed path for the load.
@@
 \begin{itemize}
-    \item Is holiday: a binary variable for holidays, where holiday$=1$ and working day$=0$
-    \item Day of week: an ordinal categorical variable corresponding to the day of the week, e.i.\ Monday$=0$,$\dots$, Sunday$=6$
+    \item Is holiday: a binary variable for holidays, where holiday=1 and working day=0
+    \item Day of week: an ordinal categorical variable corresponding to the day of the week, e.i.\ Monday=0,$\dots$, Sunday=6
 \end{itemize}
@@
 % now comment tables
-What can be concluded from this first study is that kernel quantile regression equipped with the absolute Laplacian kernel outperformed almost always its contenders.
+What can be concluded from this first study is that kernel quantile regression equipped with the Absolute Laplacian kernel outperformed almost always its contenders.
 
@@
 This section analyzes which kernel function best fits to the characteristics of the load data.
-The kernel functions considered are: Gaussian RBF, absolute Laplacian, Matern 1.5, Matern 2.5, Linear, Periodic, Polynomial, Sigmoid, and Cosine. 
+The kernel functions considered are: Gaussian RBF, Absolute Laplacian, Matern 0.5, Matern 1.5, Matern 2.5, Linear, Periodic, Polynomial, Sigmoid, and Cosine. 
 Details concerning the hyperparameter selection for the different kernels can be found in \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.metrics.pairwise}{sklearn.metrics.pairwise} and \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.gaussian_process}{sklearn.gaussian\_process}.
@@
             0.8 & \textbf{0.033837} & 0.034128 & 0.034797 & 0.035325 & 0.036166 & 0.037055 & 0.034290 & 0.035533 & 0.039501 & 0.040007 \\
-            0.9 & \textbf{0.021883} & 0.021871 & 0.022431 & 0.022682 & 0.023169 & 0.023730 & 0.022000 & 0.022811 & 0.025561 & 0.026019 \\
+            0.9 & 0.021883 & \textbf{0.021871} & 0.022431 & 0.022682 & 0.023169 & 0.023730 & 0.022000 & 0.022811 & 0.025561 & 0.026019 \\
             \midrule
@@
         0.4 & \textbf{0.058286} & 0.059018 & 0.058916 & 0.059078 & 0.059592 & 0.065770 & 0.061844 & 0.062495 & 0.066599 & 0.066335 \\
-        0.5 & 0.060344 & 0.061100 & 0.061319 & 0.061541 & 0.062338 & 0.068197 & 0.063864 & 0.071731 & 0.069761 & 0.069574 \\
+        0.5 & \textbf{0.060344} & 0.061100 & 0.061319 & 0.061541 & 0.062338 & 0.068197 & 0.063864 & 0.071731 & 0.069761 & 0.069574 \\
         0.6 & \textbf{0.058724} & 0.059507 & 0.059674 & 0.059930 & 0.060709 & 0.065828 & 0.062265 & 0.070269 & 0.067179 & 0.066970 \\
@@
 \end{table}
-From our tables, it can be concluded that the absolute Laplacian kernel is the one fitting best to the characteristics of load data.
+From our tables, it can be concluded that the Absolute Laplacian kernel is the one fitting best to the characteristics of load data.
 Additionally, we can conclude that our results are in line with the ones published in \cite{he2017short}.
@@
 The Periodic equipped with period $p=24$ shows good performance. 
-After those kernels, we have the Linear and Polynomial kernel.
+After those kernels, we have the Linear and Polynomial kernels.
 The worst among all kernels considered are the Cosine and the Sigmoid.
@@
 Error measure of the competition was the pinball loss, see Section \ref{metrics}, averaged over the $99$ quantiles, $q \in \{i/100\}_{i=1}^{n=99}$.
-Backed by the conclusion from the previous subsection we selected the absolute Laplacian and the Gaussian RBF as kernels in our application of KQR.
+Backed by the conclusion from the previous subsection we selected the Absolute Laplacian and the Gaussian RBF as kernels in our application of KQR.
 
@@
     \label{fig:cv_scatter}
-    \caption{Cross validation Gaussian RBF kernel task 4 GEFCom2014 price}
+    \caption{Cross validation price track task 4, Gaussian RBF kernel}
 \end{figure}
@@
     \includegraphics[width=\linewidth]{images/price_task_4_gaussian_rbf.png}
-    \caption[Prediction price track task 4, Gaussian RBF kernel]{Price 90\% confidence interval task 4.
+    \caption[Prediction price track task 4, Gaussian RBF kernel]{Price 90\% confidence interval task 4 using KQR Gaussian RBF.
 Electricity price probabilistic forecast for the $4^{\text{th}}$ July $2013$.
@@
 What can be concluded from Table \ref{tab:pinball loss gefcom2014 price data} is that KQR performs consistently among the top algorithms.
-Furthermore, It ranks first in two out of the thirteen tasks of the price track.
+Furthermore, It ranks first in one out of the twelve tasks of the price track.
 \subsubsection{GEFCom2014 load track} 
@@
     \item Is holiday: a binary variable for holidays, where holiday=1 and working day=0
-    \item w avg: average of weather temperatures across 25 stations
+    \item w avg: average of weather temperatures across 25 weather stations
 \end{itemize}
@@
     \makebox[\linewidth]{\includegraphics[width = 1.5\textwidth]{images/load_task_9_laplacian.png}}
-    \caption[Prediction load track task 9, Laplacian kernel]{Load 90\% confidence interval task 9 using KQR Laplacian.
+    \caption[Prediction load track task 9, Absolute Laplacian kernel]{Load 90\% confidence interval task 9 using KQR Absolute Laplacian.
 The black line is the observed path for the load.
@@
     \makebox[\linewidth]{\includegraphics[width = 1.5\textwidth]{images/sub_boxplot_task_9_laplacian_.png}}
-    \caption{Boxplot for first two days of probabilistic forecast task 9 using KQR Laplacian}
+    \caption{Boxplot for first two days of probabilistic forecast task 9 using KQR Absolute Laplacian}
     \label{fig:boxplot_load_task_9}
@@
 % repeat the problem description  and motivation quickly
-This section concludes this thesis work with some final remarks.
+This section concludes this thesis with some final remarks.
 This thesis was concerned with electricity forecasting, in particular we focused on the probabilistic framework.
@@
 The code has been packaged and uploaded to the Python Package Index (PyPI) with the name \href{https://pypi.org/project/kernel-quantile-regression/#2}{kernel-quantile-regression}.
-The \href{https://github.com/luca-pernigo/kernel_quantile_regression}{github repo} hosting the source code includes also the script implementing the experiments along with the cleaned datasets; this contribution is intended to forster reproducibility in research.
+The \href{https://github.com/luca-pernigo/ThesisKernelMethods}{github repo} hosting the source code includes also the script implementing the experiments along with the cleaned datasets; this contribution is intended to forster reproducibility in research.
     \item We achieved superior performance of kernel quantile regression compared to standard quantile regressor algorithms. 
@@
 Specifically, kernel quantile regression resulted almost always in superior performance compared to standard quantile regressors.
-The absolute Laplacian kernel turned out to be the most suitable kernel according to the characteristics of electricity forecasting.
+The Absolute Laplacian kernel turned out to be the most suitable kernel according to the characteristics of electricity forecasting.
 Finally, we proved the validity of our kernel quantile regression implementation.

### 2024-06-02 thesis/input_files/data_exploratory.tex
June 2nd 2024

--- 2024-05-26 thesis/input_files/data_exploratory.tex
@@
 Additionally, as the competition went on, the real observed data of the previous tasks were made available.
-The fifteen target days ordered by task number are: 16/06/2013, 17/06/2013, 24/06/2013, 04/07/2013, 09/07/2013, 13/07/2013, 16/07/2013, 18/07/2013, 19/07/2013, 20/07/2013, 24/07/2013, 25/07/2013, 07/12/2013, 08/12/2013.
+The 15 target days ordered by task number are: 16/06/2013, 17/06/2013, 24/06/2013, 04/07/2013, 09/07/2013, 13/07/2013, 16/07/2013, 18/07/2013, 19/07/2013, 20/07/2013, 24/07/2013, 25/07/2013, 07/12/2013, 08/12/2013, 17/12/2013.
+The first three tasks were designed for testing, they did not count for the competition.
 \begin{figure}[!h]
@@
     \includegraphics[width=\textwidth]{images/gefcom_zonal_price_vs_zonal_load.png}
-    \caption{Zonal price versus zonal load GEFCom2014}
+    \caption{Zonal price versus zonal load of GEFCom2014}
     \label{fig:gefcom_zonal_price_vs_zonal_load}
@@
 Any data from the various task folders are shifted by one month between each other.
+Also in this track, the first three tasks do not count towards the ranking of the various algorithms.
 \\
@@
 Such visualisation are reported in Figure \ref{fig:CH_day_of_week_boxplot_2021} and Figure \ref{fig:CH_is_holiday_boxplot_2021}.
-We can see that weekdays and workdays involve a higher load and possess a wider range.
+We can see that workdays involve a higher load and possess a wider range than weekdays and holidays.
 

### 2024-06-02 thesis/input_files/implementation.tex
June 2nd 2024

@@
 \subsection{K-nearest neighbours}
-The object KNeighborsRegressor of the \href{https://scikit-learn.org/stable/}{sklearn} module neighbors has been used.
+The object KNeighborsRegressor of the \href{https://scikit-learn.org/stable/}{sklearn} module neighbors has been used with 12 neighbours and the Euclidean distance.
 
@@
 \subsection{Long short term memory}
-The LSTM predictor has been built using the \href{https://pytorch.org}{torch} library.
+The LSTM predictor has been built using the \href{https://pytorch.org}{torch} library, see Section \ref{sec:lstm point} for architecture details.
 
 \subsection{Kernel ridge regression}
-The object KernelRidge of the \href{https://scikit-learn.org/stable/}{sklearn} module kernel\_ridge has been used.
+The object KernelRidge of the \href{https://scikit-learn.org/stable/}{sklearn} module kernel\_ridge has been used with the Gaussian RBF kernel.
 
 \subsection{Kernel support vector regression}
-The object SVR of the \href{https://scikit-learn.org/stable/}{sklearn} module svm has been used.
+The object SVR of the \href{https://scikit-learn.org/stable/}{sklearn} module svm has been used with the Gaussian RBF kernel.
 % by specifying rbf as the kernel parameter.

### 2024-06-02 thesis/input_files/literaturereview.tex
June 2nd 2024

@@
 \\
-In terms of forecasting horizons, we can group EF into four major categories: very short term forcasting (VSTF), short term load forecasting (STF), medium term forecasting (MTF) and long term forecasting (LTF).
-Consensus in the literature is to use one day, two weeks and three years respectilvely \cite{hong_phd} as cut off horizons; see Figure \ref{fig:time} for a visualisation.
+In terms of forecasting horizons, we can group EF into four major categories: very short term forcasting (VSTF), short term forecasting (STF), medium term forecasting (MTF) and long term forecasting (LTF).
+Consensus in the literature is to use one day, two weeks and three years respectilvely as cut off horizons \cite{hong_phd}; see Figure \ref{fig:time} for a visualisation.
 \begin{figure}
@@
 \section{Bibliographic analysis}
-This section presents the results of the bibliometric analysis we performed on March 6th 2024.
+This section presents the results of the bibliometric analysis we performed on March $6^th$ 2024.
 This survey has been carried out by using the Scopus citation database.
@@
 Specifically, it advocates for studies using similar datasets, using the same error evaluation metric and statistical testing model's outperformance.
-Hong et al.
-\cite{prob_elf} discusses the state of the art in probabilistic electric load forecasting.
+Hong et al.\ \cite{prob_elf} discusses the state of the art in probabilistic electric load forecasting.
 It differentiates between techniques and methodologies.
@@
 Also this paper stresses the need for some guidelines to standardize research in the field.
-Nowotarski et al. \cite{nowotarski} carried out a review of probabilistic forecasting.
-Weron et al. \cite{lago} offer a set of best practices when forecasting electricity prices in order to have a common framework to evaluate and compare future research.
-Zhang et al. \cite{zhang2014review} considers state of the art methods in wind power probabilistic forecasting and describes current challenges and possible future developments.
-Ziel et al. \cite{ziel2018probabilistic} provides detailed tables, groups research papers by the time dimension and objective and reports dataset, model and accuracy 
+Nowotarski et al.\ \cite{nowotarski} carried out a review of probabilistic forecasting.
+Weron et al.\ \cite{lago} offer a set of best practices when forecasting electricity prices in order to have a common framework to evaluate and compare future research.
+Zhang et al.\ \cite{zhang2014review} considers state of the art methods in wind power probabilistic forecasting and describes current challenges and possible future developments.
+Ziel et al.\ \cite{ziel2018probabilistic} provides detailed tables, groups research papers by the time dimension and objective and reports dataset, model and accuracy 
 measures adopted.
-David et al.
-\cite{david2016probabilistic} adopt a combination of autoregressive moving average (ARMA) and generalized autoregressive conditional heteroskedasticity (GARCH) in probabilistic forecasts of solar irradiance.
+David et al.\ \cite{david2016probabilistic} adopt a combination of autoregressive moving average (ARMA) and generalized autoregressive conditional heteroskedasticity (GARCH) in probabilistic forecasts of solar irradiance.
 Furthermore, they propose a recursive framework for parameter estimation.
 De Gooijer \cite{de200625} reviews 25 years of time series forecasting for the period 1990-2005, highlighting the most influential works.
-He et al.
-\cite{he2022cooperative} models  multistep wind speed probabilistic forecasting by mixing complementary ensemble empirical mode decomposition (CEEDMAN), least absolute shrinkage and selection operator (LASSO) and quantile regression (QR).
+He et al.\ \cite{he2022cooperative} models multistep wind speed probabilistic forecasting by mixing complementary ensemble empirical mode decomposition (CEEDMAN), least absolute shrinkage and selection operator (LASSO) and quantile regression (QR).
 In this work, CEEMDAN is used to decompose the wind speed time series, LASSO compresses high dimensional features, QR is used for obtaining quantile forecasts, finally kernel density estimation (KDE) converts quantile forecasts into density estimates.
-Wan et al. \cite{wan2016direct} proposes a combination of QR and extreme learning machine (ELM) to generate non parametric probabilistic forecasts of wind generation.
-Zhang et al. \cite{zhang2019wind} forecasts wind speed by adopting QRMGM, which is a combination of QR with a minimal gated memory network.
-Hyndman et al. \cite{hyndman1996estimating} explains kernel estimator of conditional density, analyses its asymptotic behaviour and covers an application for the daily temperatures of Melbourne.
-Kaur et al. \cite{kaur2022energy} carries out a comparative study of techniques spanning ideas from statistic and artificial intelligence.
-Van der Meer et al. \cite{van2018review} provides another thorough anlysis of the probabilistic forecasting realm by covering recent advances and identifying research gaps.
+Wan et al.\ \cite{wan2016direct} proposes a combination of QR and extreme learning machine (ELM) to generate non parametric probabilistic forecasts of wind generation.
+Zhang et al.\ \cite{zhang2019wind} forecasts wind speed by adopting QRMGM, which is a combination of QR with a minimal gated memory network.
+Hyndman et al.\ \cite{hyndman1996estimating} explains kernel estimator of conditional density, analyses its asymptotic behaviour and covers an application for the daily temperatures of Melbourne.
+Kaur et al.\ \cite{kaur2022energy} carries out a comparative study of techniques spanning ideas from statistic and artificial intelligence.
+Van der Meer et al.\ \cite{van2018review} provides another thorough anlysis of the probabilistic forecasting realm by covering recent advances and identifying research gaps.
 The IEEE Power and Energy Society provides also insightful lecture notes on probabilistic energy forecasting methodologies,
 implementations, and applications \citeW{gm_22}.
-Marcjasz et al. \cite{probablistic_electricity_forecast2} uses distributional neural networks to create probabilistic forecasts for the day ahead electricity prices in the German market.
-Nowotarski et al.
-\cite{nowotarski2015computing} introduce a method for constructing prediction intervals (PI) and call it quantile regression averaging.
+Marcjasz et al.\ \cite{probablistic_electricity_forecast2} uses distributional neural networks to create probabilistic forecasts for the day ahead electricity prices in the German market.
+Nowotarski et al.\ \cite{nowotarski2015computing} introduce a method for constructing prediction intervals (PI) and call it quantile regression averaging.
 Their idea is weighting a set of model predictions such that the pinball loss of the weighted model is minimized.
 The observed result is that QRA performed better compared to twelve individual models.
-Arora et al.
-\cite{arora2016forecasting} focus on modelling electricity smart meter data by proposing a non parametric probabilistic technique based on kernel density estimation and conditional kernel density estimation \cite{rosenblatt1969conditional, hyndman1996estimating}.
+Arora et al.\ \cite{arora2016forecasting} focus on modelling electricity smart meter data by proposing a non parametric probabilistic technique based on kernel density estimation and conditional kernel density estimation \cite{rosenblatt1969conditional, hyndman1996estimating}.
 Their conclusion is that kernel density methods are competitive against exponential smoothing when forecasting residential data.
 Conversely, exponential smoothing has still an edge in predicting small to medium-sized enterprises data (SMEs).
-Zhang et al.
-\cite{zhang2020probability} introduce a framework based on quantile regression and kernel density estimation in the context of short term wind forecasting.
+Zhang et al.\ \cite{zhang2020probability} introduce a framework based on quantile regression and kernel density estimation in the context of short term wind forecasting.
 The proposed methods behave well compared to an autoregressive ARMA model.
 % ARMA(p=1,q=1) 
-Haben et al. \cite{haben2018short} analyses a variety of techniques in terms of both probabilistic and point forecasting. Within this study, they focus on load forecasting at the low voltage level.
+Haben et al.\ \cite{haben2018short} analyses a variety of techniques in terms of both probabilistic and point forecasting. Within this study, they focus on load forecasting at the low voltage level.
 %Metrics Start
-Koochali et al. \cite{koochali2022random} reviews various existing methods for assessing probabilistic forecast models and discusses their advantages and disadvantages.
-Matheson et al. \cite{matheson1976scoring} develops classes of scoring rules for continuous probability distributions.
-Gneiting et al. \cite{gneiting2007strictly}
+Koochali et al.\ \cite{koochali2022random} reviews various existing methods for assessing probabilistic forecast models and discusses their advantages and disadvantages.
+Matheson et al.\ \cite{matheson1976scoring} develops classes of scoring rules for continuous probability distributions.
+Gneiting et al.\ \cite{gneiting2007strictly}
 provides a thorough overview of the theory of scoring rules for interval and density forecasts.
-Gneiting et al. \cite{gneiting2014probabilistic}
+Gneiting et al.\ \cite{gneiting2014probabilistic}
 covers theory and state of the art techniques in probabilistic forecasting.
 %Metrics End
-Zhang et al \cite{zhang2020two} proposes a two stage bootstrap sampling framework for probabilistic load forecasting.
+Zhang et al.\ \cite{zhang2020two} proposes a two stage bootstrap sampling framework for probabilistic load forecasting.
 They test it for different regression models such as random forest (RF), gradient boosting regression tree (GBRT), linear regression, and least squares support vector regression (LSSVM).
-Jónsson et al. \cite{jonsson2014predictive}
+Jónsson et al.\ \cite{jonsson2014predictive}
 introduces a density model for the day ahead market extending the adaptive QR framework of \cite{moller2008time} by modelling the tails of the predicted density with an exponential distribution.
-Fatema et al.
-\cite{fatema2023probabilistic} considers Gaussian process regression for point forecasting and prediction intervals estimation.
+Fatema et al.\ \cite{fatema2023probabilistic} considers Gaussian process regression for point forecasting and prediction intervals estimation.
 Then, it inputs prediction intervals to kernel density estimation in order to estimate a probability distribution.
 Dudek \cite{dudek2018probabilistic} proposes a probabilistic forecasting model based on the Nadaraya Watson estimator \cite{nadaraya1964estimating,watson1964smooth}.
-Huurman et al.
-\cite{huurman2012power} surveys the predictive power of weather variables for electricity prices in the Danish market.
+Huurman et al.\ \cite{huurman2012power} surveys the predictive power of weather variables for electricity prices in the Danish market.
 Their empirical results suggest that weather is central for point forecasting day ahead prices.
@@
 %GEFCom 2014
-The focus of GEFCom 2014 was on probabilistic forecasting, Hong et al. \cite{hong2016probabilistic} discusses the problem tracks, the data and the winning methods.
+The focus of GEFCom 2014 was on probabilistic forecasting, Hong et al.\ \cite{hong2016probabilistic} discusses the problem tracks, the data and the winning methods.
 In this paragraph some of the winning entries of the 2014 GEFCom edition are discussed.
-Xie et al.
-\cite{xie2016gefcom2014} propose a two stage approach; in the first stage they use multiple linear regression (MLR) to build a point forecast, then in the second stage they try different approaches for modelling the MLR residuals, among other they tried exponential smoothing (ESM), artificial neural networks (ANN) and autoregressive integrated moving average (ARIMA).
-Maciejowska et al. \cite{maciejowska2016hybrid} proposes a new probabilistic model extending the idea of quantile regression averaging (QRA) \cite{nowotarski2015computing}.
-Haben et al. \cite{haben2016hybrid} mixes conditional kernel density (CKD) and quantile regression in their competition entry.
-Gaillard et al. \cite{gaillard2016additive,gaillardasemi} combines quantile regression with generalized additive models \cite{hastie2017generalized}.
-Ziel et al. \cite{ziel2016lasso} estimates an AR model through the LASSO \cite{tibshirani1996regression} instead of the standard OLS.
+Xie et al.\ \cite{xie2016gefcom2014} propose a two stage approach; in the first stage they use multiple linear regression (MLR) to build a point forecast, then in the second stage they try different approaches for modelling the MLR residuals, among other they tried exponential smoothing (ESM), artificial neural networks (ANN) and autoregressive integrated moving average (ARIMA).
+Maciejowska et al.\ \cite{maciejowska2016hybrid} proposes a new probabilistic model extending the idea of quantile regression averaging (QRA) \cite{nowotarski2015computing}.
+Haben et al.\ \cite{haben2016hybrid} mixes conditional kernel density (CKD) and quantile regression in their competition entry.
+Gaillard et al.\ \cite{gaillard2016additive,gaillardasemi} combines quantile regression with generalized additive models \cite{hastie2017generalized}.
+Ziel et al.\ \cite{ziel2016lasso} estimates an AR model through the LASSO \cite{tibshirani1996regression} instead of the standard OLS.
 \\
@@
 The kernel theory needed for this thesis work is covered in Section \ref{kernel_theory}. 
-For an introduction to kernel methods, we refer to \cite{learning_with_kernels, hofmann2006review} and \cite{shawe2004kernel}.
+For an introduction to kernel methods, we refer to \cite{learning_with_kernels, hofmann2006review, shawe2004kernel}.
 \\
 Kanagawa et Fukumizu introduces to the concept of kernel mean embedding \cite{pmlr}.
-Muandet et al.
-\cite{Muandet_2017} surveys established results and new advances in the theory of Hilbert space distribution embeddings.
+Muandet et al.\ \cite{Muandet_2017} surveys established results and new advances in the theory of Hilbert space distribution embeddings.
 It has to be said that, computing and storing such embeddings becomes prohibitive for large scale settings.
-Rudi et al.
-\cite{2022nystrom} proposes an efficient approximation procedure based on the Nyström method \cite{nystrom}, providing also an upper bound for the approximation error.
+Rudi et al.\ \cite{2022nystrom} proposes an efficient approximation procedure based on the Nyström method \cite{nystrom}, providing also an upper bound for the approximation error.
 Another example of the applicability of kernel methods is \cite{supersamples}.

### 2024-06-02 thesis/input_files/paper1.tex
June 2nd 2024

@@
     \\
-    x&\to \phi(x)
+    x&\mapsto \phi(x)
 \end{align*}
@@
 \section{Theory of kernels}
-Subsequently we introduce the definitions that make up the basis for the theory of kernel methods.
+Subsequently, we introduce the definitions that make up the basis for the theory of kernel methods.
 
@@
     \end{align}
-    Notice, positive semidefinitenees property of kernel $k$ is a sufficient condition for $\mathcal{H}$ to have an associated RKHS.
+    Notice, positive semidefiniteness property of kernel $k$ is a sufficient condition for $\mathcal{H}$ to have an associated RKHS.
 \end{definition}
@@
 \begin{proposition}
-    Let $k_1$ and $k_2$ be kernels over $\mathcal{X} \times \mathcal{X}, \mathcal{X} \subset \mathbb{R}^n$, $a \in \mathbb{R}^+$ and $\phi: \mathcal{X} \to \mathbb{R}^N$ with $k_3$ a kernel over $\mathbb{R}^N \times \mathbb{R}^N$.
+    Let $k_1$ and $k_2$ be kernels over $\mathcal{X} \times \mathcal{X}, \mathcal{X} \subseteq \mathbb{R}^n$, $a \in \mathbb{R}^+$ and $\phi: \mathcal{X} \to \mathbb{R}^N$ with $k_3$ a kernel over $\mathbb{R}^N \times \mathbb{R}^N$.
 Then the following functions are valid kernels:
@@
        \midrule
-       Linear &  $k(x,y)=xy$ &   \\
+       Linear &  $k(x,y)=x^\intercal y$ &   \\
        Polynomial &  $k(x,y)=(x^\intercal y+c)^d$ &   c, d\\
@@
        Chi-squared kernel &  $k(x,y)=\exp\left(-\gamma \sum\limits_i \frac{(x_i - y_i)^2}{x_i+y_i}\right) $ &   $\gamma$\\
-        Cosine  &  $k(x,y)=\frac{xy^\intercal}{\|x\|_{2}\|y\|_{2}}$ &   \\
+        Cosine  &  $k(x,y)=\frac{x^\intercal y}{\|x\|_{2}\|y\|_{2}}$ &   \\
        Matérn  &  $k(x, y) =  \frac{1}{\Gamma(\nu)2^{\nu-1}}\Bigg(

### 2024-06-02 thesis/input_files/point.tex
June 2nd 2024

@@
 The $C$ constant trades off between $\epsilon$ deviation tolerance and flatness of the function $f$.
-A bigger $C$ gives more importance to error minimization while as $C$ gets smaller, flattness gains importance.
+A bigger $C$ gives more importance to error minimisation while as $C$ gets smaller, flattness gains importance.
 \\
@@
     \includegraphics[width=\textwidth]{images/lstm_cell.png}
-    \caption{Long short term memory cell}
+    \caption{Long short term memory cell \cite{Han2024}}
     \label{fig:lstm_cell}
@@
 \begin{theorem}
-    Let $k:\mathcal{X} \times \mathcal{X} \mapsto \mathbb{R}$ on a non empty set $\mathcal{X}$ with corresponding RKHS $\mathcal{H}$,
+    Let $k:\mathcal{X} \times \mathcal{X} \to \mathbb{R}$ on a non empty set $\mathcal{X}$ with corresponding RKHS $\mathcal{H}$,
     a strictly increasing real valued function $g:[0,\infty) \to \mathbb{R}$, and the loss function $E:(\mathcal{X}\times \mathbb{R}^2)^n \to \mathbb{R} \cup \{\infty\}$.
-    Consider the regularized risk functional $Risk[f]:f \mapsto  E\left((x_{1},y_{1},f(x_{1})),\ldots, (x_{n},y_{n},f(x_{n}))\right)+g\left(\| f\|_{\mathcal{H}} \right)$.
+    Consider the regularised risk functional $Risk[f]:f \mapsto  E\left((x_{1},y_{1},f(x_{1})),\ldots, (x_{n},y_{n},f(x_{n}))\right)+g\left(\| f\|_{\mathcal{H}} \right)$.
 Then any minimizer of the empirical risk can be represented as $$ f^{*}(\cdot )=\sum\limits_{i=1}^{n}\alpha _{i}k(\cdot ,x_{i})$$
@@
     This means that setting $\nu=0$ does not affect the evaluation of $f$.
-    Next consider the regularization term. The functional $g$ is defined to be strictly monotonic in order to penalize more complex functions $f$.
+    Next consider the regularisation term. The functional $g$ is defined to be strictly monotonic in order to penalise more complex functions $f$.
     \begin{equation}

### 2024-06-02 thesis/input_files/probabilistic.tex
June 2nd 2024

@@
 \section{Kernel quantile regression}
-The idea of quantile regression has been extended to kernel methods by Takeuchi et al. \cite{takeuchi2006nonparametric}.
+The idea of quantile regression has been extended to kernel methods by Takeuchi et al.\ \cite{takeuchi2006nonparametric}.
 There, they minimise a risk functional plus regulariser defined as follows.
@@
     \end{subfigure} 
-    \caption{Quantile regressors comparison Melbourne weather data}
+    \caption{Quantile regressors comparison for Melbourne weather data}
     \label{fig:melbourne_quantiles_comparison} 
@@
 \begin{table}[!h]
-\caption{Pinball loss of Melbourne data}
+\caption{Pinball loss for Melbourne data}
 \begin{tabular}{lllll}
@@
 \begin{table}[!h]
-    \caption{Pinball loss quantile-wise of Melbourne data}
+    \caption{Pinball loss quantile-wise for Melbourne data}
     \begin{tabular}{lllll}
@@
 \begin{table}[!h]
-\caption{Mean absolute error of Melbourne data}
+\caption{Mean absolute error for Melbourne data}
 \begin{tabular}{lllll}
@@
 \begin{table}[!h]
-    \caption{Kernels comparison mean absolute error Melbourne data}
+    \caption{Kernels comparison mean absolute error for Melbourne data}
     \label{tab:kernel mae comparison}

### 2024-06-02 thesis/input_files/problemdescription.tex
June 2nd 2024

@@
-Individuals and organizations constantly face situations of uncertainty, thus the need for robust forecasting methods.
+Individuals and organisations constantly face situations of uncertainty, thus the need for robust forecasting methods.
 Such methods are crucial in the process of taking informed decisions and for strategic planning.
@@
 Over the last decades, electricity markets have gone through an unprecedented transformation.
-This shift was driven by the liberalization of such markets, the development and integration of renewable energy sources, the  increase of low carbon technologies and the adoption of smart meters.
+This shift was driven by the liberalisation of such markets, the development and integration of renewable energy sources, the  increase of low carbon technologies and the adoption of smart meters.
 Events like the California electricity crisis are further motivating the choice of the electricity sector as subject of our studies, see \cite{california}.

### 2024-06-02 thesis/input_files/qr.tex
June 2nd 2024

@@
 Quantile regression can be interpreted as an extension of standard regression.
-In this setting, one slices the dependent variables into quantiles and then fit a regression for each quantile.
+In this setting, one slices the dependent variables into quantiles and then fits a regression for each quantile.
 With standard regression, we build a model for the conditional mean, conversely, with quantile regression we model the conditional quantile function for any desired quantile.
@@
 \\
-For any $0<q<1$ consider the pinball loss function from section \ref{pinball}, $\rho_q(u)=u(q-\mathbb{I}_{\{u<0\}})$. 
+For any $0<q<1$ consider the pinball loss function from Section \ref{pinball}, $\rho_q(u)=u(q-\mathbb{I}_{\{u<0\}})$. 
 Such loss is minimized by the quantile $Q_q$.

### 2024-06-03 thesis/appendix.tex
final

@@
 The Boston housing dataset \href{https://www.kaggle.com/datasets/altavish/boston-housing-dataset}{https://www.kaggle.com/datasets/altavish/boston-housing-dataset} contains information about various attributes for suburbs in Boston.
-There are 13 indipendent variables:
+There are 13 indipendent variables
 \begin{itemize}
@@
 \subsection{Abalone dataset}
-The abalone data \href{https://archive.ics.uci.edu/dataset/1/abalone}{https://archive.ics.uci.edu/dataset/1/abalone} consistst of measurements of abalone molluscs, the goal is predicting their age by building a model for estimating its number of rings; age is the number of rings plus 1.5.
-The data has 8 attributes:
+The abalone data \href{https://archive.ics.uci.edu/dataset/1/abalone}{https://archive.ics.uci.edu/dataset/1/abalone} consists of measurements of abalone molluscs, the goal is predicting their age by building a model for estimating its number of rings; age is the number of rings plus 1.5.
+The data has 8 attributes
 \begin{itemize}
@@
 \subsection{Vehicle dataset}
-This data contains info about used cars \href{https://www.kaggle.com/datasets/nehalbirla/vehicle-dataset-from-cardekho}{https://www.kaggle.com/datasets/ne-halbirla/vehicle-dataset-from-cardekho}, the predictors are:
+This data contains info about used cars \href{https://www.kaggle.com/datasets/nehalbirla/vehicle-dataset-from-cardekho}{https://www.kaggle.com/datasets/ne-halbirla/vehicle-dataset-from-cardekho}, the predictors are
 \begin{itemize}
@@
 
-\subsection{Randomized search}
-Randomized search controls the number of steps by choosing smartly the hyperparameters to try in each iterations.
-Let us say, there are 100 candidates and we set the number of iterations to 20 then the search will stop after the 20th iteration and return the best set among the hyperparameters observed.
+\subsection{Randomised search}
+Randomised search controls the number of steps by choosing smartly the hyperparameters to try in each iteration.
+Let us say, there are 100 candidates and we set the number of iterations to 20 then the search will stop after the $20^{th}$ iteration and return the best set among the hyperparameters observed.
 The advantage is that it is much more quicker than gridsearch, but on the other hand its performance is worse than gridsearch.
@@
 When data points are dependent on preceding values, we cannot use standard $K$-fold cross validation.
-The rationale is that $K$-fold will randomize the order of the data, thus it might happen to use future data to predict the past; we want to avoid such behaviour in a time series setting.
+The rationale is that $K$-fold will randomise the order of the data, thus it might happen to use future data to predict the past; we want to avoid such behaviour in a time series setting.
 When carrying out out any kind of cross validation, we must keep consistency in the way we evaluate our predictors during model selection and in the way we perform evaluation of the test data.
@@
 \subsection{Data normalisation}\label{appendix:normalization}
-With data normalization we transform the range of the feautures to a standard scale.
+With data normalisation we transform the range of the features to a standard scale.
 Such preprocessing step is essential when employing distance based algorithms like SVM or K-nearest neighbours.
-The rationale behind it is that by normalizing data we give an uniform weight to each feature in the learning process; in this way we do not favour larger scale features.
+The rationale behind it is that by normalising data we give an uniform weight to each feature in the learning process; in this way we do not favour larger scale features.
 Examples of the most popular features scaling are:
@@
 Using kernel methods, we have that the storage cost is of the order $O(n^2)$ while the computational cost for finding the solution is of the order $O(n^3)$.
-A significant speed up can be obtained thanks to low rank approximation.
+A significant speed up can be obtained thanks to low rank approximations.
 
-\subsubsection{Nystrom decomposition}
+\subsubsection{Nyström decomposition}
 The Nyström approximation involves storing a submatrix of the whole kernel matrix. Thus, storage and computational costs are reduced to $O(nm)$ and $O(nm^2)$ respectively.
@@
 \begin{itemize}
-    \item query/: folder containing Scopus data and scripts to generate bibliometric survey plots in Section \ref{literature_review}.
+    \item query/: folder containing Scopus data and scripts to generate bibliometric survey plots in Section \ref{literature_review}. Database queries have been taken from \cite{nowotarski}.
     \item thesis/: folder containing tex files for thesis pdf.

### 2024-06-03 thesis/input_files/analysis.tex
final

@@
 \section{Point forecasting}
-This section carries out a comparative study between the state of the art methods for point forecasting, introduced in \ref{ch:point}.
+This section carries out a comparative study between the state of the art methods for point forecasting, introduced in Section \ref{ch:point}.
 As use case, we will consider the task of predicting the electric load from the GEFCom2014 dataset.
@@
     \item Month
-    \item Day of the week
+    \item Day of week: an ordinal categorical variable corresponding to the day of the week, e.i.\ Monday=0,$\dots$, Sunday=6
     \item Is holiday: a binary variable for holidays, where holiday=1 and working day=0
@@
 Its name stands for trigonometric seasonality, Box-Cox transformation, ARMA errors, trend and seasonal components.
-Tbats forecast is visualized in Figure \ref{fig:tbats_price}, meanwhile its RMSE on task 1 is 15.08.
+Tbats forecast is visualised in Figure \ref{fig:tbats_price}, meanwhile its RMSE on task 1 is 15.08.
 \begin{figure}[!h]
@@
 We get started by considering the base implementation. In such setting, prophet takes as input the time series object and learns its data generating process.
-This method achieves on task 1 a RMSE of 23.96, its prediction is visualized in Figure \ref{fig:prophet_price_1}.
+This method achieves on task 1 a RMSE of 23.96, its prediction is visualised in Figure \ref{fig:prophet_price_1}.
 \begin{figure}[!h]
@@
 Notice, for kernel quantile regression (KQR), the kernel of choice in this setting is the Absolute Laplacian.
-Additionally notice, the loss function utilized in this section is the normalized pinball loss.
+Additionally notice, the loss function utilised in this section is the normalised pinball loss.
 % normalized by the magnitude of the average load of each country.
@@
 \subsection{Kernel function choice}
-This section analyzes which kernel function best fits to the characteristics of the load data.
+This section analyses which kernel function best fits to the characteristics of the load data.
 The kernel functions considered are: Gaussian RBF, Absolute Laplacian, Matern 0.5, Matern 1.5, Matern 2.5, Linear, Periodic, Polynomial, Sigmoid, and Cosine. 
@@
 The pool of data for this study has been created by combining \href{https://zenodo.org/records/7907883}{SECURES-Met} data (predictors) \cite{Formayer2023} and the load data from \href{https://www.energy-charts.info/index.html?l=en&c=DE}{energy charts}.
-SECURES-Met data consists of historical data up to the end of $2020$ while from 2021 onward, the data consists of forecasts modeled by the European Centre for Medium-Range Weather Forecasts (ECMWF).
+SECURES-Met data consists of historical data up to the end of $2020$ while from 2021 onward, the data consists of forecasts modelled by the European Centre for Medium-Range Weather Forecasts (ECMWF).
 Therefore, in this experimental study, we used the whole data of $2020$ as the training set and then tested our kernels on the $2021$ data.
@@
 
-The predictors making up the datasets follow:
+The predictors making up the datasets follow
 \begin{itemize}
@@
 Additionally, we can conclude that our results are in line with the ones published in \cite{he2017short}.
-In that study, similarly to what we did in this subsection, the authors carried out a comparitive study concerning the choice of kernels in KQR.
-Nevertheless notice, that the set of kernel we considered in our study is larger than the one presented in \cite{he2017short}.
+In that study, similarly to what we did in this subsection, the authors carried out a comparative study concerning the choice of kernels in KQR.
+Nevertheless, notice that the set of kernels we considered in our study is larger than the one presented in \cite{he2017short}.
 Their paper showed evidence of the superiority of the Gaussian RBF among their set of kernel functions.
@@
 In this track, the top entries came from the teams: Tololo, Team Poland, GMD, and C3 Green Team; for a breakdown of each method's attributes, see \cite[Table 8]{hong2016probabilistic}.
-The Gaussian RBF probabilistic prediction for the $4^{th}$ July $2013$ zonal price at the $90\%$ confidence interval is visualized in Figure \ref{fig:price_task_4_gaussian_rbf}.
-Cross validation results are visualised in Figure.
+The Gaussian RBF probabilistic prediction for the $4^{th}$ July $2013$ zonal price at the $90\%$ confidence interval is visualised in Figure \ref{fig:price_task_4_gaussian_rbf}.
+Cross validation results are visualised in Figure \ref{fig:cv_scatter}.
 
@@
     \includegraphics[width=\textwidth]{images/cv_scatter.png}
+    \caption{Cross validation price track task 4, Gaussian RBF kernel}
     \label{fig:cv_scatter}
-    \caption{Cross validation price track task 4, Gaussian RBF kernel}
 \end{figure}
@@
 What can be concluded from Table \ref{tab:pinball loss gefcom2014 price data} is that KQR performs consistently among the top algorithms.
-Furthermore, It ranks first in one out of the twelve tasks of the price track.
+Furthermore, it ranks first in one out of the twelve tasks of the price track.
 \subsubsection{GEFCom2014 load track} 
@@
     \item Hour
-    \item Day of week: an ordinal categorical variable for the day of the week
+    \item Day of week: an ordinal categorical variable corresponding to the day of the week, e.i.\ Monday=0,$\dots$, Sunday=6
     \item Is holiday: a binary variable for holidays, where holiday=1 and working day=0
@@
 Managalova, Ziel Florian, and Bidong Liu; for a breakdown of the attributes of each method, see \cite[Table 6]{hong2016probabilistic}.
-Finally, the $90\%$ confidence interval forecast by our model equipped with the Absolute Laplacian kernel for task number $9$, that is the prediction for June, are reported in Figure \ref{fig:load_task_9} and Figuree \ref{fig:boxplot_load_task_9}.
+Finally, the $90\%$ confidence interval forecast by our model equipped with the Absolute Laplacian kernel for task number $9$, that is the prediction for June, is reported in Figure \ref{fig:load_task_9} and Figure \ref{fig:boxplot_load_task_9}.
 Similarly to the price track, we have that KQR was among the best algorithms also in the load forecasting track.
@@
 The code has been packaged and uploaded to the Python Package Index (PyPI) with the name \href{https://pypi.org/project/kernel-quantile-regression/#2}{kernel-quantile-regression}.
-The \href{https://github.com/luca-pernigo/ThesisKernelMethods}{github repo} hosting the source code includes also the script implementing the experiments along with the cleaned datasets; this contribution is intended to forster reproducibility in research.
+The \href{https://github.com/luca-pernigo/ThesisKernelMethods}{github repo} hosting the source code includes also the scripts implementing the experiments along with the cleaned datasets; this contribution is intended to forster reproducibility in research.
     \item We achieved superior performance of kernel quantile regression compared to standard quantile regressor algorithms. 

### 2024-06-03 thesis/input_files/data_exploratory.tex
final

@@
 In the load track, contestants were asked to provide one month ahead hourly probabilistic forecasts on a rolling basis for 15 consecutive rounds.
-To get started, in the first round, organizers provided 69 months of hourly load data (from 01/01/2005 to 30/10/2010) and 117 months of hourly temperature data (from 01/01/2001 to 30/10/2010).
+To get started, in the first round, organisers provided 69 months of hourly load data (from 01/01/2005 to 30/10/2010) and 117 months of hourly temperature data (from 01/01/2001 to 30/10/2010).
 As for the price track, the true observed data from the previous track were made available as the competition progressed.
@@
 \subsubsection{GEFCom ETL}
-The provided data is raw, hence the need for a standardized extract transform and load pipeline. 
+The provided data is raw, hence the need for a standardised extract transform and load pipeline. 
 Firstly, we drop all Nans and zeros.
@@
 \end{figure}
-Next, we built scatter plots of the independent continous variables against the load as means of visualizing correlation between them.
+Next, we built scatter plots of the independent continous variables against the load as means of visualising correlation between them.
 Figure \ref{fig:CH_load_vs_temperature_2021} plots load against temperatures while Figure \ref{fig:CH_load_vs_wind_speed_2021} depicts load against wind speed.
@@
 
-Following, we use box plots in order to visualize the relations between categorical independent variables and the dependent one.
-Such visualisation are reported in Figure \ref{fig:CH_day_of_week_boxplot_2021} and Figure \ref{fig:CH_is_holiday_boxplot_2021}.
+Following, we use box plots in order to visualise the relations between categorical independent variables and the dependent one.
+Such visualisations are reported in Figure \ref{fig:CH_day_of_week_boxplot_2021} and Figure \ref{fig:CH_is_holiday_boxplot_2021}.
 We can see that workdays involve a higher load and possess a wider range than weekdays and holidays.

### 2024-06-03 thesis/input_files/energymarket.tex
final

@@
 This chapter has two intents.
-Firstly, it summarizes several concepts for the newcomers to the field of electricity markets.
+Firstly, it summarises several concepts for the newcomers to the field of electricity markets.
 Secondly, it discusses trends and challenges in the power sector.
@@
 Before this revolution, the power sector was a natural monopoly.
-In particular, it was characterized by a high vertical integration and by very little market competition.
+In particular, it was characterised by a high vertical integration and by very little market competition.
 Better technologies in transmission, generation and distribution are the reasons for such liberal shift.
@@
 \subsection{Price peaks}
-Electricity markets are characterized by spikes in their spot prices. When this happens, the system price jumps abruptly and then drops back within a very short period.
+Electricity markets are characterised by spikes in their spot prices. When this happens, the system price jumps abruptly and then drops back within a very short period.
 This spikeness follows from the non storability of electricity, that is, electricity has to be consumed as it is produced.
@@
 Grid configuration and physical limits on electricity lines may lead to congestion.
-Zonal market clearing pricing (ZMCP) and local marginal price (LMP) are two types of pricing schemes utilized to address this.
+Zonal market clearing pricing (ZMCP) and local marginal price (LMP) are two types of pricing schemes utilised to address this.
 The former is adopted in European contries while the latter is used in the United States, Figure \ref{fig:eu_zonal} and Figure \ref{fig:us_local}.
@@
 The reason is that smart metering and the electricity network digitalisation  will enable system operators to better monitor the network, plan their investments and manage infrastructure. 
-All of this motivates the need for efficient and robust tools to analyze the vast data resulting from the digitalisation of the power sector, see \cite{declean,andreadou2016telecommunication}.
+All of this motivates the need for efficient and robust tools to analyse the vast data resulting from the digitalisation of the power sector, see \cite{declean,andreadou2016telecommunication}.
 
@@
 For example, connecting renewables power plants located in remote and offshore locations to high voltage powerlines is not trivial.
-Additionally, renewables such as wind and solar are characterized by seasonalities and depend heavily on weather conditions.
+Additionally, renewables such as wind and solar are characterised by seasonalities and depend heavily on weather conditions.
 Therefore, balancing the electricity network and mantaining stability will become harder for grid operators.

### 2024-06-03 thesis/input_files/implementation.tex
final

@@
-This section is intended to explain and aid for reproducibility studies. Hereafter, the specific libraries used and the custom implementatios are thoroughly documented.
+This section is intended to explain and aid for reproducibility studies. Hereafter, the specific libraries used and the custom implementations are thoroughly documented.
 \\
@@
 \subsection{Kernel quantile regression}
-Kernel quantile regression had no previously implemented Python open source library, thus the need of implementing our own version.
+Kernel quantile regression had no previously implemented Python open source library, thus, the need of implementing our own version.
 \\
@@
 Therefore, the KQR class is derived from the scikit-learn BaseEstimator and the mixin class RegressorMixin.
-Our KQR class is initialized by providing a quantile, the regularisation term $C$, the kernel family and its corresponding hyperparameters.
+Our KQR class is initialised by providing a quantile, the regularisation term $C$, the kernel family and its corresponding hyperparameters.
 \\
-In the fit method, we set up and solve the convex optimisation problem through the interior point method.
+In the fit method, we set up and solve the convex optimisation problem through the interior point algorithm.
 This algorithm is taken from the cvxopt library, see its official manual \citeW{vandenberghe2010cvxopt} for a reference.
 When using this library, it is important to keep two things in mind.
-First this library assumes the quadratic term of the optimisation problem to be multiplied by the 0.5 factor, thus we just have to provide the $Q$ matrix with no 0.5 in front.
+First this library assumes the quadratic term of the optimisation problem to be multiplied by the 0.5 factor, thus, we just have to provide the $Q$ matrix with no 0.5 in front.
 Secondly, in order to specify multiple inequalities we have to stack them and provide them as a single matrix.
@@
 Moreover, this code is compatible with sklearn in built kernel functions \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.metrics.pairwise}{sklearn.metrics.pairwise} and \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.gaussian_process.kernels}{sklearn.gaussian\_process.kernels}.
-Functions supported by our kernel quantile regression include: Gaussian RBF, absolute Laplacian, Matern 3/2, Matern 5/2, Linear, Cosine, Sigmoid, Periodic, Polynomial and custom composition of kernels.
+Functions supported by our kernel quantile regression include: Gaussian RBF, Absolute Laplacian, Matern 3/2, Matern 5/2, Linear, Cosine, Sigmoid, Periodic, Polynomial and custom composition of kernels.
 Kernels from these two submodules use a different convention in the hyperparameters.
-The former has the lenghtscale parameter in the numerator while the latter in the denominator.
+The former has the lengthscale parameter in the numerator while the latter in the denominator.
 In order to meaningfully compare across different kernels we made them consistent in this aspect.
-For this purpose we adhere to the convention of having lenghtscale hyperparameters in the denominator.
+For this purpose we adhere to the convention of having lengthscale hyperparameters in the denominator.
 

### 2024-06-03 thesis/input_files/literaturereview.tex
final

@@
 \\
-The field benefitted greatly from the increse of computing power, the greater availability of data and the interest in data science.
+The field benefitted greatly from the increase of computing power, the greater availability of data and the interest in data science.
 As a consequence, the forecaster's toolbox has grown in size and complexity.
@@
 Different solutions may outperform or underperform compared to other techniques depending on the problem settings.
-Thus, understanding the complexity, strenghts and weaknesses of each method is crucial for fitting the right model to the right setting.
+Thus, understanding the complexity, strengths and weaknesses of each method is crucial for fitting the right model to the right setting.
 \\
@@
 \\
-In terms of forecasting horizons, we can group EF into four major categories: very short term forcasting (VSTF), short term forecasting (STF), medium term forecasting (MTF) and long term forecasting (LTF).
-Consensus in the literature is to use one day, two weeks and three years respectilvely as cut off horizons \cite{hong_phd}; see Figure \ref{fig:time} for a visualisation.
+In terms of forecasting horizons, we can group EF into four major categories: very short term forecasting (VSTF), short term forecasting (STF), medium term forecasting (MTF) and long term forecasting (LTF).
+Consensus in the literature is to use one day, two weeks and three years respectively as cut off horizons \cite{hong_phd}; see Figure \ref{fig:time} for a visualisation.
 \begin{figure}
@@
 Point forecasts are easier to generate and less computationally intensive while probabilistic forecasts are more informative.
-Industry and research efforts have focused primarily on point forcasting.
+Industry and research efforts have focused primarily on point forecasting.
 Nevertheless, interest in probabilistic forecasting has risen considerably over the last years due to renewable integration requirements, introduction of smart grids and increased market competitiveness.
 \section{Bibliographic analysis}
-This section presents the results of the bibliometric analysis we performed on March $6^th$ 2024.
+This section presents the results of the bibliometric analysis we performed on March $6^{th}$ 2024.
 This survey has been carried out by using the Scopus citation database.
@@
 To get started, let us consider the evolution of the EF field over the years.
-This is visualized in Figure \ref{fig:epf_evolution}, with results grouped by category.
+This is visualised in Figure \ref{fig:epf_evolution}, with results grouped by category.
 Articles prior to the 2000 have been aggregated together due to their small number.
@@
 \\
-The next question is to compare the state of point-versus probabilistic forecasting, this is visualized in Figure \ref{fig:point_vs_prob}.
+The next question is to compare the state of point-versus probabilistic forecasting, this is visualised in Figure \ref{fig:point_vs_prob}.
 What can be concluded is that probabilistic-is less developed than point forecasting.
@@
 \\
-EF is a heterogenous field of research, its researchers come from a wide array of backgrounds, with electrical engineers and statisticians making up the top contributors; their different educational training may explain why the split between statistical and computational intelligence methods is so marked.
+EF is a heterogeneous field of research, its researchers come from a wide array of backgrounds, with electrical engineers and statisticians making up the top contributors; their different educational training may explain why the split between statistical and computational intelligence methods is so marked.
 Figure \ref{fig:subject_area} depicts the EF publications by subject area.
@@
   \includegraphics[width=\textwidth]{images/src_title.png}
-  \caption{Most popolar sources/outlets}
+  \caption{Most popular sources/outlets}
   \label{fig:src_title}
@@
 Weron \cite{EPF_review} reviews the state of the art for electricity price forecasting.
-Besides analysing complexity of available solutions, strenghts and weaknesses, it also stresses the need for objective comparative EPF studies.
+Besides analysing complexity of available solutions, strengths and weaknesses, it also stresses the need for objective comparative EPF studies.
 Specifically, it advocates for studies using similar datasets, using the same error evaluation metric and statistical testing model's outperformance.
@@
 On the other hand, methodologies consist of general frameworks that can be incorporated into any method, for example variable selection mechanisms.
-Also this paper stresses the need for some guidelines to standardize research in the field.
+Also this paper stresses the need for some guidelines to standardise research in the field.
 Nowotarski et al.\ \cite{nowotarski} carried out a review of probabilistic forecasting.
@@
 Zhang et al.\ \cite{zhang2014review} considers state of the art methods in wind power probabilistic forecasting and describes current challenges and possible future developments.
-Ziel et al.\ \cite{ziel2018probabilistic} provides detailed tables, groups research papers by the time dimension and objective and reports dataset, model and accuracy 
+Ziel et al.\ \cite{ziel2018probabilistic} provides detailed tables, groups research papers by the time dimension and objective and reports datasets, models and accuracy 
 measures adopted.
-David et al.\ \cite{david2016probabilistic} adopt a combination of autoregressive moving average (ARMA) and generalized autoregressive conditional heteroskedasticity (GARCH) in probabilistic forecasts of solar irradiance.
+David et al.\ \cite{david2016probabilistic} adopts a combination of autoregressive moving average (ARMA) and generalised autoregressive conditional heteroskedasticity (GARCH) in probabilistic forecasts of solar irradiance.
 Furthermore, they propose a recursive framework for parameter estimation.
@@
 Kaur et al.\ \cite{kaur2022energy} carries out a comparative study of techniques spanning ideas from statistic and artificial intelligence.
-Van der Meer et al.\ \cite{van2018review} provides another thorough anlysis of the probabilistic forecasting realm by covering recent advances and identifying research gaps.
+Van der Meer et al.\ \cite{van2018review} provides another thorough analysis of the probabilistic forecasting realm by covering recent advances and identifying research gaps.
 The IEEE Power and Energy Society provides also insightful lecture notes on probabilistic energy forecasting methodologies,
@@
 Nowotarski et al.\ \cite{nowotarski2015computing} introduce a method for constructing prediction intervals (PI) and call it quantile regression averaging.
-Their idea is weighting a set of model predictions such that the pinball loss of the weighted model is minimized.
+Their idea is weighting a set of model predictions such that the pinball loss of the weighted model is minimised.
 The observed result is that QRA performed better compared to twelve individual models.
@@
 Note that, the more the errors of the combined models are not correlated, the more we can benefit from ensembles.
-It is also worth noting that older and simpler methods are still valuable (in combination with other models or on their own). These being less subject to overfitting than complex models.
+It is also worth noting that older and simpler methods are still valuable (in combination with other models or on their own), these being less subject to overfitting than complex models.
 \\
@@
 A number of ideas were tested on the same setting with only the best ones being published and it also contributed bridging the gap between industry practice and academic research.
-GEFCom 2012 had two tracks; the former about hierarchical load forecasting, the latter about wind power forecasting, see \cite{hong2014global} for a comprehensive review.
+GEFCom2012 had two tracks; the former about hierarchical load forecasting, the latter about wind power forecasting, see \cite{hong2014global} for a comprehensive review.
 \\
 %GEFCom 2014
-The focus of GEFCom 2014 was on probabilistic forecasting, Hong et al.\ \cite{hong2016probabilistic} discusses the problem tracks, the data and the winning methods.
+The focus of GEFCom2014 was on probabilistic forecasting, Hong et al.\ \cite{hong2016probabilistic} discusses the problem tracks, the data and the winning methods.
 In this paragraph some of the winning entries of the 2014 GEFCom edition are discussed.
-Xie et al.\ \cite{xie2016gefcom2014} propose a two stage approach; in the first stage they use multiple linear regression (MLR) to build a point forecast, then in the second stage they try different approaches for modelling the MLR residuals, among other they tried exponential smoothing (ESM), artificial neural networks (ANN) and autoregressive integrated moving average (ARIMA).
+Xie et al.\ \cite{xie2016gefcom2014} proposes a two stage approach; in the first stage they use multiple linear regression (MLR) to build a point forecast, then in the second stage they try different approaches for modelling the MLR residuals, among other they tried exponential smoothing (ESM), artificial neural networks (ANN) and autoregressive integrated moving average (ARIMA).
 Maciejowska et al.\ \cite{maciejowska2016hybrid} proposes a new probabilistic model extending the idea of quantile regression averaging (QRA) \cite{nowotarski2015computing}.
 Haben et al.\ \cite{haben2016hybrid} mixes conditional kernel density (CKD) and quantile regression in their competition entry.
-Gaillard et al.\ \cite{gaillard2016additive,gaillardasemi} combines quantile regression with generalized additive models \cite{hastie2017generalized}.
+Gaillard et al.\ \cite{gaillard2016additive,gaillardasemi} combines quantile regression with generalised additive models \cite{hastie2017generalized}.
 Ziel et al.\ \cite{ziel2016lasso} estimates an AR model through the LASSO \cite{tibshirani1996regression} instead of the standard OLS.
@@
 %GEFCom 2017
-The last GEFCom was held in 2017, its focus was providing probabilistic load forecasts, see \cite{hong2019global}.
-The GEFCom competition has also inspired the organization of other competitions such as the RWEnpower competition in the UK, the RTE competition in France, the Tokyo electric power company competition in Japan and the BigDEAL forecasting competitions.
+The last GEFCom was held in 2017, its focus was providing hierarchical probabilistic load forecasts, see \cite{hong2019global}.
+The GEFCom competition has also inspired the organisation of other competitions such as the RWEnpower competition in the UK, the RTE competition in France, the Tokyo electric power company competition in Japan and the BigDEAL forecasting competitions.
 \\
@@
 A couple of considerations can be drawn from the above literature review.
-The EF field is characterized by heterogeneity in its forecasting techniques; methods come from statisitics, mathematics, econometrics, electrical engineering and the artificial intelligence communities.
+The EF field is characterised by heterogeneity in its forecasting techniques; methods come from statisitics, mathematics, econometrics, electrical engineering and the artificial intelligence communities.
 \\

### 2024-06-03 thesis/input_files/metrics.tex
June 2nd 2024
final

@@
     $$
-    \mathrm{MASE}(y,\hat{y})=\frac{1}{h}\sum\limits_{k=1}^N\frac{|y_{n+k}-\hat{y}_{n+k}|}{\frac{1}{h-1}\sum\limits_{k=2}^{h}|y_{k}-y_{k-1}|}
+    \mathrm{MASE}(y,\hat{y})=\frac{1}{h}\sum\limits_{k=1}^h\frac{|y_{n+k}-\hat{y}_{n+k}|}{\frac{1}{h-1}\sum\limits_{k=2}^{h}|y_{k}-y_{k-1}|}
     $$
@@
     $$
-    \mathrm{RMSSE}(y,\hat{y})=\sqrt{\frac{1}{h}\sum\limits_{k=1}^N\left(\frac{|y_{n+k}-\hat{y}_{n+k}|}{\frac{1}{h-1}\sum\limits_{k=2}^{h}|y_{k}-y_{k-1}|}\right)^2}
+    \mathrm{RMSSE}(y,\hat{y})=\sqrt{\frac{1}{h}\sum\limits_{k=1}^h\left(\frac{|y_{n+k}-\hat{y}_{n+k}|}{\frac{1}{h-1}\sum\limits_{k=2}^{h}|y_{k}-y_{k-1}|}\right)^2}
     $$
@@
 
-\section{Pinball}\label{pinball}
+\section{Pinball score}\label{pinball}
 The pinball score or quantile score is used to measure the accuracy of a quantile forecast.
@@
 By averaging all the pinball losses over all quantiles and over the whole forecast horizon, we obtain the pinball loss of the probabilistic forecast.
-\section{Winkler}
+\section{Winkler score}
 \begin{definition}
@@
     $$
-    \mathrm{Winkler}(y_t, \hat{y}_t,q)=\begin{cases}
+    \mathrm{Winkler}(y_t,  u_{t},l_{t},\alpha)=\begin{cases}
         \delta & l_{t}\leq y_{t}\leq u_{t}\\
@@
 \end{definition}
-Where $\delta$ is the prediction interval (PI) width, that is $\delta=u_t-l_t$, $u_t$ is the prediction interval upper threshold and $l_t$ is the prediction interval lower threshold at time step $t$.
-This score penalizes observations falling outside the prediction interval and rewards narrow prediction intervals.
+Where $\delta$ is the prediction interval (PI) width, that is $\delta=u_t-l_t$, $\alpha$ is (1-coverage level), $u_t$ is the prediction interval upper threshold and $l_t$ is the prediction interval lower threshold at time step $t$.
+This score penalises observations falling outside the prediction interval and rewards narrow prediction intervals.
 \section{Continous ranked probability score}
@@
 Figure \ref{fig:pit} provides an example, applying the true CDF results in a well calibrated PIT (left).
-Alternatively, applying a bad CDF results in either a overdispersed (middle) or underdispersed (right) PIT.
+Alternatively, applying a bad CDF results in either an overdispersed (middle) or underdispersed (right) PIT.
 \begin{figure}

### 2024-06-03 thesis/input_files/paper1.tex
final

@@
         Cosine  &  $k(x,y)=\frac{x^\intercal y}{\|x\|_{2}\|y\|_{2}}$ &   \\
-       Matérn  &  $k(x, y) =  \frac{1}{\Gamma(\nu)2^{\nu-1}}\Bigg(
+       Matern  &  $k(x, y) =  \frac{1}{\Gamma(\nu)2^{\nu-1}}\Bigg(
         \frac{\sqrt{2\nu}}{l} \|x-y|\|_{2}
@@
 
-Notice, in the Matérn family, $\Gamma(\cdot)$ is the gamma function while $Y(\cdot)$ is the Bessel function of the second kind.
+Notice, in the Matern family, $\Gamma(\cdot)$ is the gamma function while $Y(\cdot)$ is the Bessel function of the second kind.
 \\
-When $\nu=p+\frac{1}{2}$, $p \in \mathbb{N}^+$, the Matérn kernel can be rewritten as a product of an exponential and a polynomial of degree $p$. 
+When $\nu=p+\frac{1}{2}$, $p \in \mathbb{N}$, the Matern kernel can be rewritten as a product of an exponential and a polynomial of degree $p$. 
 \\
@@
 
-When $\nu=\frac{1}{2}$ the Matérn kernel corresponds to the Laplacian kernel. If the predictors are one dimensional, then the Absolute Laplacian and the Matern 0.5 are equivalent. %L2 norm=L1 norm
-Furthermore, as $\nu \to \infty$ we have that the Matérn kernel converges to the Gaussian RBF kernel.
+When $\nu=\frac{1}{2}$ the Matern kernel corresponds to the Laplacian kernel. If the predictors are one dimensional, then the Absolute Laplacian and the Matern 0.5 are equivalent. %L2 norm=L1 norm
+Furthermore, as $\nu \to \infty$ we have that the Matern kernel converges to the Gaussian RBF kernel.
 $k(x,y)=\exp\left(\frac{-\|x-y\|_{2}^2}{2l^2}\right)$

### 2024-06-03 thesis/input_files/point.tex
final

@@
 where $\Psi$ and $\Theta$ account for the $P$ and $Q$ of the seasonal pattern $m$ respectively.
-\section{Generalized additive model}
+\section{Generalised additive model}
 % in standard regression the predicted y are normal since error are normal.
 But if we want to model say a y bernoulli which is the rate of success we can link the domain of bernoulli to that of gaussian with a link function
-Generalized additive models (GAMs) take the form
+Generalised additive models (GAMs) take the form
 \begin{equation}
@@
 $k$ is the algorithm hyperparameter, it stands for the number of neighbours; a small $k$ leads to overfitting while a big $k$ leads to underfitting.
-Since it is a metric based algorithm, it is important to normalize data in order to give equal weights to features with different scales, see Appendix \ref{appendix:normalization}.
+Since it is a metric based algorithm, it is important to normalise data in order to give equal weights to features with different scales, see Appendix \ref{appendix:normalization}.
 This algorithm is made up of two decisions: first, the choice of the metric and second, the method for combining targets.
@@
 \end{equation}
-where $\psi$ is an activation function and $w,b$ are the parameters of the neuron. The parameters are typically optimized through gradient descent.
+where $\psi$ is an activation function and $w,b$ are the parameters of the neuron. The parameters are typically optimised through gradient descent.
 
@@
     Consider the regularised risk functional $Risk[f]:f \mapsto  E\left((x_{1},y_{1},f(x_{1})),\ldots, (x_{n},y_{n},f(x_{n}))\right)+g\left(\| f\|_{\mathcal{H}} \right)$.
-Then any minimizer of the empirical risk can be represented as $$ f^{*}(\cdot )=\sum\limits_{i=1}^{n}\alpha _{i}k(\cdot ,x_{i})$$
+Then any minimiser of the empirical risk can be represented as $$ f^{*}(\cdot )=\sum\limits_{i=1}^{n}\alpha _{i}k(\cdot ,x_{i})$$
 \end{theorem}
@@
     Hence, we can conclude that setting $\nu=0$ decreases the second term while the first term is not affected.
-Thus, it follows that any minimizer of the risk functional of Equation \ref{eq:kernel ridge} must be of the following form
+Thus, it follows that any minimiser of the risk functional of Equation \ref{eq:kernel ridge} must be of the following form
     \begin{equation}
@@
     \\
-    =&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\tau}{2} \left\langle \sum _{i=1}^{n}\alpha _{i}k(\cdot ,x_{i}),\sum _{j=1}^{n}\alpha _{j}k(\cdot ,x_{j})\right\rangle_{\mathcal{H}}^2
+    =&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\tau}{2} \left\langle \sum _{i=1}^{n}\alpha _{i}k(\cdot ,x_{i}),\sum _{j=1}^{n}\alpha _{j}k(\cdot ,x_{j})\right\rangle_{\mathcal{H}}
     \\
-    =&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\tau}{2} \sum _{i,j=1}^{n}\alpha _{i}\alpha _{j}\langle k(\cdot ,x_{i}),k(\cdot ,x_{j})\rangle_{\mathcal{H}}^2
+    =&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\tau}{2} \sum _{i,j=1}^{n}\alpha _{i}\alpha _{j}\langle k(\cdot ,x_{i}),k(\cdot ,x_{j})\rangle_{\mathcal{H}}
     \\
@@
 \subsection{Kernel support vector regression}
-Support vector regression (SVR) can be kernelized by swapping the $\mathbb{R}^n$ euclidean inner product of $x$ with the inner product in the higher feature space $\mathcal{H}$.
+Support vector regression (SVR) can be kernelized by swapping the $\mathbb{R}^n$ Euclidean inner product of $x$ with the inner product in the higher feature space $\mathcal{H}$.
 Doing so, the optimisation problem can be restated as

### 2024-06-03 thesis/input_files/probabilistic.tex
final

@@
 To get started, we first split the space into two disjoint regions, then we model the response variable by the mean of the observed predicted variables with associated features falling in that specific region.
-Our goal is selecting the best features and best split point to achieve the best generalizing fit.
+Our goal is selecting the best features and best split point to achieve the best generalising fit.
 For a visualisation consider the Figures \ref{fig:elements_statistical_learning1} and \ref{fig:elements_statistical_learning2}, where the decision tree algorithm is visualised for a regression problem with two independent variables $X_1$ and $X_2$.
@@
 It follows that the function minimising the sum of squares is the one with $\hat{c}_m=mean(y_i|x_i \in R_m)$.
-Finding the best split in terms of minimum sum of squares is computationally infeasible in practice. Therefore we approximate a solution by approaching the problem in a greedy fashion.
+Finding the best split in terms of minimum sum of squares is computationally infeasible in practice. Therefore, we approximate a solution by approaching the problem in a greedy fashion.
 Let $j$ denote the splitting variable and $s$ be the split point we define the two half planes
@@
     C(q-1)\mathbb{1}\preceq \alpha \preceq Cq\mathbb{1}\\
-    &\alpha^\intercal\mathbb{1}=1
+    &\alpha^\intercal\mathbb{1}=0
     \end{aligned}
@@
 \end{equation}
-Since our optimisation problem possesses strong duality and it is differentiable in both the objective and the constraint, we have that it must satisfy the Karush Kuhn Tucker conditions (KKT), see Section 5.5.3 \cite{boyd2004convex}.
+Since our optimisation problem possesses strong duality and it is differentiable in both the objective and the constraint, we have that it must satisfy the Karush Kuhn Tucker conditions (KKT), see \cite[Section 5.5.3]{boyd2004convex}.
 Thanks to the KKT conditions on the primal optimisation problem we have that $f(x_i)=y_i \ \mathrm{for} \ \alpha_i \not \in \{C(q-1), Cq\}$. 
@@
     \midrule
-    0.100000 & 0.710644 & 0.549232 & 0.562888 & \textbf{0.540145} \\
-    0.200000 & 1.155014 & 0.938561 & 0.946712 & \textbf{0.903193} \\
-    0.300000 & 1.417805 & 1.212671 & 1.222407 & \textbf{1.173627} \\
-    0.400000 & 1.540108 & 1.399925 & 1.409293 & \textbf{1.368022} \\
-    0.500000 & 1.574957 & 1.517281 & 1.484589 & \textbf{1.456053} \\
-    0.600000 & 1.525114 & 1.498608 & 1.495474 & \textbf{1.447470} \\
-    0.700000 & 1.397918 & 1.372183 & 1.362173 & \textbf{1.331375} \\
-    0.800000 & 1.170140 & 1.115077 & 1.123195 & \textbf{1.096649} \\
-    0.900000 & 0.787195 & \textbf{0.714075} & 0.734112 & 0.715013 \\
+    0.1 & 0.710644 & 0.549232 & 0.562888 & \textbf{0.540145} \\
+    0.2 & 1.155014 & 0.938561 & 0.946712 & \textbf{0.903193} \\
+    0.3 & 1.417805 & 1.212671 & 1.222407 & \textbf{1.173627} \\
+    0.4 & 1.540108 & 1.399925 & 1.409293 & \textbf{1.368022} \\
+    0.5 & 1.574957 & 1.517281 & 1.484589 & \textbf{1.456053} \\
+    0.6 & 1.525114 & 1.498608 & 1.495474 & \textbf{1.447470} \\
+    0.7 & 1.397918 & 1.372183 & 1.362173 & \textbf{1.331375} \\
+    0.8 & 1.170140 & 1.115077 & 1.123195 & \textbf{1.096649} \\
+    0.9 & 0.787195 & \textbf{0.714075} & 0.734112 & 0.715013 \\
     \bottomrule
@@
 \end{table}  
-From these tables, we can see that kernel quantile regression outperformes the simple quantile regressor as well as the more complex models like quantile forest and gradient boosting quantile regression for the Melbourne temperatures dataset.
+From these tables, we can see that kernel quantile regression outperforms the simple quantile regressor as well as the more complex models like quantile forest and gradient boosting quantile regression for the Melbourne temperatures dataset.
 Not only kernel quantile regression was the best in terms of total pinball loss but also the best in terms of each quantile pinball loss and mean absolute error.
@@
 From these tables, we can conclude that the Matern and the Chi squared kernels are the top performers on the Melbourne weather data.
-Finally, the equaility in the scores of the Absolute Laplacian and the Matern 0.5 follows from the fact that regressors are one dimensional for the Melbourne weather data.
+Finally, the equality in the scores of the Absolute Laplacian and the Matern 0.5 follows from the fact that regressors are one dimensional for the Melbourne weather data.
 % tables

### 2024-06-03 thesis/input_files/problemdescription.tex
final

@@
 Events like the California electricity crisis are further motivating the choice of the electricity sector as subject of our studies, see \cite{california}.
-Moreover, the process of deregulation lead to an increasing interest in the field of electricity forecasting (EF) within the academic community, see Figure \ref{fig:epf_evolution}.
+Moreover, the process of deregulation led to an increasing interest in the field of electricity forecasting (EF) within the academic community, see Figure \ref{fig:epf_evolution}.
 In addtion, the United Nations have identified the right to access affordable, reliable, sustainable and modern energy as one of their 17 sustainable development goals (SDGs) \citeW{un_sdgs}.
@@
 There are mutliple reasons why the energy sector needs robust forecasting techniques.
-For power market companies, being able to predict prices with a low mean absolute percentage error (MAPE) \ref{mape} results in increased savings \cite{savings}.
+For power market companies, being able to predict prices with a low mean absolute percentage error (MAPE), see Section \ref{mape}, results in increased savings \cite{savings}.
 Furthermore, the adoption of smart meters provides power market companies with a huge amount of consumer data.
@@
 A probabilistic forecast can be turned into a point forecast by simply taking its expectation.
-Alternatively, a probabilistic forecast can be derived from a point one by modeling the residuals of the point prediction.
+Alternatively, a probabilistic forecast can be derived from a point one by modelling the residuals of the point prediction.
 
@@
 
-The scope of the thesis is analyzing state of the art forecasting methods in the energy market and to compare and to integrate them with ideas coming from the theory of kernel methods.
+The scope of the thesis is analysing state of the art forecasting methods in the energy market and to compare and to integrate them with ideas coming from the theory of kernel methods.
 
@@
 Evaluation metrics necessary to rank the forecasting techniques are presented in Section \ref{metrics}.
-Section \ref{energy_market} explains the core features and terminology of the energy market and of the electricity newtork.
+Section \ref{energy_market} explains the core features and terminology of the energy market and of the electricity network.
 Following, Sections \ref{ch:point} and \ref{ch:prob} introduce the state of the art methods in the context of point and probabilistic forecasting respectively.
@@
 Implementation details are included in Section \ref{implementation}.
-Finally, Section \ref{analysis} presents experiments results and discusses models' strenghts, weaknesses and possible improvements.
+Finally, Section \ref{analysis} concludes this thesis presenting and commenting experiments results.
+% and discusses models' strenghts, weaknesses and possible improvements.

### 2024-06-03 thesis/input_files/qr.tex
final

@@
 For any $0<q<1$ consider the pinball loss function from Section \ref{pinball}, $\rho_q(u)=u(q-\mathbb{I}_{\{u<0\}})$. 
-Such loss is minimized by the quantile $Q_q$.
- we can estimate quantiles by minimizing the expectation of $\rho_q(y-g(x,\beta))$ with respect to the parameter $\beta$.
+Such loss is minimised by the quantile $Q_q$.
+ we can estimate quantiles by minimising the expectation of $\rho_q(y-g(x,\beta))$ with respect to the parameter $\beta$.
 \\

### 2024-06-03 thesis/input_files/symbols.tex
May 15th 2024
thesis update
May 25 2024
corrected all comments
final

@@
 %G
-\glsxtrnewsymbol[description={Generalized additive models}]{GAMs}{GAMs}
+\glsxtrnewsymbol[description={Generalised additive models}]{GAMs}{GAMs}
 \glsxtrnewsymbol[description={Gradient boosting regression tree}]{GBRT}{GBRT}
 \glsxtrnewsymbol[description={Global energy forecasting competition}]{GEFCom}{GEFCom}
-\glsxtrnewsymbol[description={Gaussian process regression}]{GPR}{GPR}
+% \glsxtrnewsymbol[description={Gaussian process regression}]{GPR}{GPR}
 
@@
 \glsxtrnewsymbol[description={Random forest}]{RF}{RF}
-\glsxtrnewsymbol[description={Reproducing kernel hilbert space}]{RKHS}{RKHS}
+\glsxtrnewsymbol[description={Reproducing kernel Hilbert space}]{RKHS}{RKHS}
 \glsxtrnewsymbol[description={Root mean squared error}]{RMSE}{RMSE}
+\glsxtrnewsymbol[description={Run of river}]{ROR}{ROR}
 
 %P
+\glsxtrnewsymbol[description={Pattern analysis}]{PA}{PA}
 \glsxtrnewsymbol[description={Partial autocorrelation function}]{PACF}{PACF}
@@
 % D
+\glsxtrnewsymbol[description={direct sum}]{oplus}{\ensuremath{\oplus}}
 \glsxtrnewsymbol[description={dual optimum}]{d^*}{\ensuremath{d^*}}
@@
 
-\glsxtrnewsymbol[description={feature space}]{F}{\ensuremath{F}}
+\glsxtrnewsymbol[description={feature space}]{F}{\ensuremath{\mathcal{F}}}
 
@@
 % I
-\glsxtrnewsymbol[description={Indicator function}]{mathbb{I}}{\ensuremath{\mathbb{I}}}
+\glsxtrnewsymbol[description={indicator function}]{mathbb{I}}{\ensuremath{\mathbb{I}}}
 
@@
 % L
-\glsxtrnewsymbol[description={lagrangian}]{L}{\ensuremath{L}}
+\glsxtrnewsymbol[description={Lagrangian}]{L}{\ensuremath{\mathcal{L}}}
 
@@
 
-\glsxtrnewsymbol[description={reproducing kernel hilbert space}]{H}{\ensuremath{\mathcal{H}}}
+\glsxtrnewsymbol[description={reproducing kernel Hilbert space}]{H}{\ensuremath{\mathcal{H}}}
 
 % T
-\glsxtrnewsymbol[description={Tensor product}]{otimes}{\ensuremath{\otimes}}
+\glsxtrnewsymbol[description={tensor product}]{otimes}{\ensuremath{\otimes}}
 

### 2024-06-03 experiments/eda/gefcom.py

@@ -46,11 +46,11 @@ plt.show()
 
 # price
 
-df=pd.read_csv("/Users/luca/Desktop/ThesisKernelMethods/experiments/gefcom_price/Price/Task 1/Task1_P_train.csv")
+df=pd.read_csv("/Users/luca/Desktop/ThesisKernelMethods/experiments/Data/Price/Task 1/Task1_P_train.csv")
 
 # # zonal price vs total load
 plt.plot(df["Forecasted Total Load"], df["Zonal Price"], "o", alpha=0.4, markersize=3)
-plt.xlabel("Forecsated total load")
+plt.xlabel("Forecasted total load")
 plt.ylabel("Price $/MW")
 plt.savefig("/Users/luca/Desktop/ThesisKernelMethods/thesis/images/gefcom_zonal_price_vs_total_load.png")
 plt.show()
@@ -58,7 +58,7 @@ plt.show()
 
 # # zonal price vs zonal load
 plt.plot(df["Forecasted Zonal Load"], df["Zonal Price"], "o", alpha=0.4, markersize=3)
-plt.xlabel("Forecsated zonal load")
+plt.xlabel("Forecasted zonal load")
 plt.ylabel("Price $/MW")
 plt.savefig("/Users/luca/Desktop/ThesisKernelMethods/thesis/images/gefcom_zonal_price_vs_zonal_load.png")
 plt.show()

### 2024-06-03 experiments/train_test/Load/load_qr_gefcom2014_test.py

@@ -40,7 +40,7 @@ def test(ith):
 
     pinball_tot=0
 
-    ktype="gaussian_rbf"
+    ktype="a_laplacian"
     
     # predict
     df_template_submission=pd.read_csv(f"experiments/Data/Load/Task {ith}/L{ith}-benchmark.csv")
@@ -72,7 +72,7 @@ def test(ith):
     # plot
     load_plot_ci(reo, y_test)
     plt.title(f"Task {ith}, {ktype} kernel")
-    # plt.savefig(f"plots/Load/load_task_{ith}_{ktype}.png")
+    # plt.savefig(f"experiments/plots/Load/load_task_{ith}_{ktype}.png")
     plt.show()
 
 

### 2024-06-04 thesis/appendix.tex
cleaned all problems and warnings

@@
-\chapter{Appendix}
 \section{Feature map normalisation}\label{appendix:new_feature}
@@
 \end{lemma}
+
 \begin{proof}
     Taking the new feature vector as $\phi^{new} (x):=\frac{\phi(x)}{\|\phi(x)\|_{\mathcal{H}}}$ we have
+
 \begin{align*}
@@
 \subsection{Boston housing dataset}
-The Boston housing dataset \href{https://www.kaggle.com/datasets/altavish/boston-housing-dataset}{https://www.kaggle.com/datasets/altavish/boston-housing-dataset} contains information about various attributes for suburbs in Boston.
+The Boston housing dataset \url{https://www.kaggle.com/datasets/altavish/boston-housing-dataset} contains information about various attributes for suburbs in Boston.
 There are 13 indipendent variables
@@
 \subsection{Vehicle dataset}
-This data contains info about used cars \href{https://www.kaggle.com/datasets/nehalbirla/vehicle-dataset-from-cardekho}{https://www.kaggle.com/datasets/ne-halbirla/vehicle-dataset-from-cardekho}, the predictors are
+This data contains info about used cars \url{https://www.kaggle.com/datasets/nehalbirla/vehicle-dataset-from-cardekho}, the predictors are
 \begin{itemize}
@@
 The idea is to compare the results of modelling the data with K-nearest neighbours on the unscaled data against the scaled data.
-The considered data is the wine recognition dataset \href{https://archive.ics.uci.edu/dataset/109/wine}{https://archive.ics.uci.edu/da-taset/109/wine}.
+The considered data is the wine recognition dataset \url{https://archive.ics.uci.edu/dataset/109/wine}.
 The goal for this dataset is recognising from whose cultivator the wine comes based on two features with a completely different scale.
@@
 The whole code for the project is hosted on
-\url{https://github.com/luca-pernigo/ThesisKernelMethods}\label{github_repo}.
-\\
+\href{https://github.com/luca-pernigo/ThesisKernelMethods}{github repo}.
+
 \begin{itemize}

### 2024-06-04 thesis/input_files/analysis.tex
cleaned all problems and warnings

@@
 
-\subsection{Trigonometric seasonality Box-Cox transformation ARMA errors trend and seasonal components}
+\subsection{Trigonometric seasonality Box-Cox transformation AR\-MA errors trend and seasonal components}
 Next, we tried with the autoregressive approach.
@@
 Tbats forecast is visualised in Figure \ref{fig:tbats_price}, meanwhile its RMSE on task 1 is 15.08.
-\begin{figure}[!h]
+\begin{figure}[!ht]
     \includegraphics[width=\textwidth]{images/tbats_price.png}
@@
 This method achieves on task 1 a RMSE of 23.96, its prediction is visualised in Figure \ref{fig:prophet_price_1}.
-\begin{figure}[!h]
+\begin{figure}[!ht]
     \includegraphics[width=\textwidth]{images/prophet_price_1.png}
@@
 % , moreover figure \ref{fig:prophet_price2.1} and figure \ref{fig:prophet_price2.2} break down the trend and the seasonalities respectively.
-\begin{figure}[!h]
+\begin{figure}[!ht]
     \includegraphics[width=\textwidth]{images/prophet_price2.png}
@@
 The forecast is visualised in Figure \ref{fig:knn_price}, RMSE on task 1 is 12.54.
-\begin{figure}[!h]
+\begin{figure}[!ht]
     \includegraphics[width=\textwidth]{images/knn_price.png}
@@
 Support vector regression achieves on task 1 an RMSE of 23.24, for the prediction see Figure \ref{fig:svr_price}.
-\begin{figure}[!h]
+\begin{figure}[!ht]
     \includegraphics[width=\textwidth]{images/svr_price.png}
@@
 LSTM forecast is reported in Figure \ref{fig:lstm_price}, it achieves on task 1 a RMSE of 9.1782
-\begin{figure}[!h]
+\begin{figure}[!ht]
     \includegraphics[width=\textwidth]{images/lstm_price.png}
@@
 The RMSE achieved on task 1 is 9.86, Figure \ref{fig:krnridge_price} reports the prediction.
-\begin{figure}[!h]
+\begin{figure}[!ht]
     \includegraphics[width=\textwidth]{images/krnridge_price.png}
@@
 RMSE achieved with this method on task 1 is 9.10, forecast is reported in Figure \ref{fig:krnsvr_price}.
-
-\begin{figure}[!h]
+\begin{figure}[!ht]
     \includegraphics[width=\textwidth]{images/krnsvr_price.png}
@@
       \midrule
-      Team name\textbackslash Task number                       & 1                               & 2                                  & 3                               & 4                              & 5                              & 6       & 7                               & 8       & 9       & 10                             & 11                             & 12               \\
+      Team name/Task number                       & 1                               & 2                                  & 3                               & 4                              & 5                              & 6       & 7                               & 8       & 9       & 10                             & 11                             & 12               \\
   KQR Absolute Laplacian
@@
       \midrule
-      Team name\textbackslash Task number                       & 1                               & 2                                  & 3                               & 4                              & 5                              & 6       & 7                               & 8       & 9       & 10                             & 11                             & 12                              \\
+      Team name/Task number                       & 1                               & 2                                  & 3                               & 4                              & 5                              & 6       & 7                               & 8       & 9       & 10                             & 11                             & 12                              \\
 

### 2024-06-04 thesis/input_files/data_exploratory.tex
cleaned all problems and warnings

@@
-% - Explain how data has been retrieved.
-
-% - Data provider Energycharts(entsoe retrieves its data)
-
-% - Explain the ETL(Extract Transform Load) pipeline
-% I set up.
-
-% - Explore data in order to get useful insights for
-% how to tune the models to get the most of them.
-
-% - correlation between temperatures and load, wind speeds and load
-
-% - Split train and test dataset
-% Explain carefully why it is important to carry
-% out an out of sample test and not in sample.
-% In sample test involves look ahead bias because we are
-% fitting the model on the data we want to predict,
-% thus it overfits on the data considered but it does
-% not generalize well.
-
-This section covers the datasets used in our experiments. In addition, attributes and features will be thoroughly described.
-Next, we will explain the extract transform load pipeline (ETL) that we set up in order to ease the workflow of our studies.
+This section covers the datasets used in our experiments. 
+In addition, attributes and features will be thoroughly described. Next, we will explain the extract transform load pipeline (ETL) that we set up in order to ease the workflow of our studies. 
 Finally, in order to better understand patterns within the data, we carry out an exploratory data analysis.
@@
 We provide forecasts for the load and price quantities.
-\\
+
 \subsection{GEFCom2014 price track}
@@
 The first three tasks were designed for testing, they did not count for the competition.
-\begin{figure}[!h]
+\begin{figure}[!ht]
     \includegraphics[width=\textwidth]{images/price_track_fig1.png}
@@
 
-\begin{figure}[!h]
+\begin{figure}[!ht]
     \includegraphics[width=\textwidth]{images/gefcom_zonal_price_vs_zonal_load.png}
@@
 
-\begin{figure}[!h]
+\begin{figure}[!ht]
     \includegraphics[width=\textwidth]{images/gefcom_zonal_price_vs_total_load.png}
@@
 As for the price track, the true observed data from the previous track were made available as the competition progressed.
-\\
+
 The data for load forecasting is contained in the GEFCom2014-L\_V2 zip file.
@@
 Also in this track, the first three tasks do not count towards the ranking of the various algorithms.
-\\
+
 In order to compare our model performance with the winning entries of the GEFCom2014 load competition, we will refer to the Provisional\_Leaderboard\_\-V2.csv file contained also in the GEFCom2014 Data directory.
@@
 
-\begin{figure}[!h]
+\begin{figure}[!ht]
     \includegraphics[width=\textwidth]{images/gefcom_load_historical.png}
@@
 
-\begin{figure}[!h]
+\begin{figure}[!ht]
     \includegraphics[width=\textwidth]{images/gefcom_w_avg_historical.png}
@@
 
-\begin{figure}[!h]
+\begin{figure}[!ht]
     \includegraphics[width=\textwidth]{images/gefcom_load_vs_w_avg.png}
@@
 As expected, electricity load exhibits a seasonal pattern with highs in the winter and lows during the summer season.
-\begin{figure}[!h]
+\begin{figure}[!ht]
     \includegraphics[width=\textwidth]{images/CH_historical_load_2021.png}
@@
 
-\begin{figure}[!h]
+\begin{figure}[!ht]
     \includegraphics[width=\textwidth]{images/CH_load_vs_temperature_2021.png}
@@
 
-\begin{figure}[!h]
+\begin{figure}[!ht]
     \includegraphics[width=\textwidth]{images/CH_load_vs_wind_speed_2021.png}
@@
 
-\begin{figure}[!h]
+\begin{figure}[!ht]
     \includegraphics[width=\textwidth]{images/CH_day_of_week_boxplot_2021.png}
@@
 
-\begin{figure}[!h]
+\begin{figure}[!ht]
     \includegraphics[width=\textwidth]{images/CH_is_holiday_boxplot_2021.png}
@@
 
-\begin{figure}[!h]
+\begin{figure}[!ht]
     \includegraphics[width=\textwidth]{images/CH_heatmap_2021.png}

### 2024-06-04 thesis/input_files/energymarket.tex
cleaned all problems and warnings

@@
-% Chapter explaining the energy market
-% Chapter 9 of https://pubs.naruc.org/pub.cfm?id=536E10A7-2354-D714-5191-A8AAFE45D626
-
 This chapter has two intents.
@@
 Such developments motivate the need for robust and efficient forecasting techniques.
-
 \section{Electricity distribution network}
@@
 Transmission lines carry electric power from stations to substations while distribution lines carry electricity from substations to load points such as businesses, industries and homes, see Figure \ref{electricity_network}.
-
-\begin{figure}[!h]
+\begin{figure}[!ht]
     \includegraphics[width=\textwidth]{images/electricity_network.png}
@@
 \end{figure}
-
 \section{Electricity markets}
@@
 The rationale is simple, a competitive market promotes efficiency and stimulates technical innovation more than a monopoly does.
-\\
+
 During the 1990s, various wholesale electricity markets were established; for example in Chile, Great Britain, the Scandinivian block, Australia, New Zealand, the US and Canada.
@@
 The MCP in power exchanges is determined by the intersection of the aggregate supply curve and the aggregate demand curve, right panel Figure \ref{fig:pool_vs_echange}.
-\\
+
 It is worth to mention the two biggest power exchanges: Epex and NordPool. Epex operates throughout continental Europe. NordPool covers the nordic and baltic countries.
-\\
+
 Moreover, we can differentiate between two popular types of auctions: uniform-price/marginal and pay-as-bid/discriminatory.
@@
 
-\begin{figure}[!h]
+\begin{figure}[!ht]
     \includegraphics[width=\textwidth]{images/pool_vs_echange.png}
@@
 Buses is where a electricity line or several lines are connected.
-\begin{figure}[!h]
+\begin{figure}[!ht]
     \includegraphics[width=\textwidth]{images/eu_zonal.png}
@@
 
-\begin{figure}[!h]
+\begin{figure}[!ht]
     \includegraphics[width=\textwidth]{images/us_local.png}
@@
 
-\begin{figure}[!h]
+\begin{figure}[!ht]
     \includegraphics[width=\textwidth]{images/electricity-production-by-source.png}
@@
 \end{figure}
+
+% Chapter explaining the energy market
+% Chapter 9 of https://pubs.naruc.org/pub.cfm?id=536E10A7-2354-D714-5191-A8AAFE45D626

### 2024-06-04 thesis/input_files/implementation.tex
cleaned all problems and warnings

@@
 This section is intended to explain and aid for reproducibility studies. Hereafter, the specific libraries used and the custom implementations are thoroughly documented.
-\\
-For the list of Python packages needed, see the requirement.txt file on \url{https://github.com/luca-pernigo/ThesisKernelMethods}\label{github_repo}.
+
+For the list of Python packages needed, see the requirement.txt file on the \href{https://github.com/luca-pernigo/ThesisKernelMethods}{github repo}.
 % - indicate computer specifics
@@
 
-\subsection{Trigonometric seasonality Box-Cox transformation ARMA errors trend and seasonal components}
+\subsection{Trigonometric seasonality Box-Cox transformation AR\-MA errors trend and seasonal components}
 The Tbats implementation is available at \href{https://github.com/intive-DataScience/tbats}{https://github.com/intive-Data-Science/tbatsx}.

### 2024-06-04 thesis/input_files/literaturereview.tex
cleaned all problems and warnings

@@
-%Electricity Forecasting 
-%intro
-During the past 25 years %because paper of 2014 says 15, so now it is 25
-a wide range of new ideas have been proposed for electricity point forecasting and for probabilistic forecasting.
-\\
+During the past 25 years a wide range of new ideas have been proposed for electricity point forecasting and for probabilistic forecasting.
 The field benefitted greatly from the increase of computing power, the greater availability of data and the interest in data science.
 As a consequence, the forecaster's toolbox has grown in size and complexity.
-\\
 Before delving into the literature review, we stress that at this point in time there is no superior method.
@@
 Thus, understanding the complexity, strengths and weaknesses of each method is crucial for fitting the right model to the right setting.
-\\
 Within this research community, the need for more homogeneity in the choice of the error valuation metrics emerged (Section \ref{metrics}), data quality and in the way of comparing model performances \cite{EPF_review}.
@@
 Thus, in order to introduce some clarity, it is useful to classify the range of EF articles in terms of their core attributes.
-\\
 In the context of energy forecasting, the quantities of most interest are electricity prices (EPF), electricity loads (ELF) and renewables generation (mostly wind and solar).
-\\
 In terms of forecasting horizons, we can group EF into four major categories: very short term forecasting (VSTF), short term forecasting (STF), medium term forecasting (MTF) and long term forecasting (LTF).
@@
 \end{figure}
-\\
 Forecasts can either be for the whole target electricity network (system) or for a subset of it (zonal).
-\\
 EF literature distinguishes between point and probabilistic forecasts.
@@
 For details on the specific queries entered in Scopus, refer to Appendix \ref{src_code}.
-\\
+
 To get started, let us consider the evolution of the EF field over the years.
@@
   \end{figure}
-\\
+
 The next question is to compare the state of point-versus probabilistic forecasting, this is visualised in Figure \ref{fig:point_vs_prob}.
@@
 Nevertheless, we can see a trend that suggests researchers are making an effort to fill this gap.
+
 \begin{figure}
@@
 \end{figure}
-\\
+
 The EF literature is dominated by statistical-and computational intelligence (CI) methods as can be seen from Figure \ref{fig:cs_stat_both}, with CI methods being slightly preferred. 
@@
 \end{figure}
-\\
+
 EF is a heterogeneous field of research, its researchers come from a wide array of backgrounds, with electrical engineers and statisticians making up the top contributors; their different educational training may explain why the split between statistical and computational intelligence methods is so marked.
@@
 \end{figure}
-\\
+
 In order to refer to the most relevant source in the field, EF outlets have been ranked by popularity and plotted in Figure \ref{fig:src_title}.
@@
 \end{figure}
-\\
+
 \section{Electricity forecasting literature review}
@@
 The opposite conclusion are drawn for density forecasting.
-\\
 Lately, the idea of combining forecasts has gained popularity in the forecasting community \cite{forecasting_big}.
@@
 It is also worth noting that older and simpler methods are still valuable (in combination with other models or on their own), these being less subject to overfitting than complex models.
-\\
+
 %GEFCom2012
@@
 GEFCom2012 had two tracks; the former about hierarchical load forecasting, the latter about wind power forecasting, see \cite{hong2014global} for a comprehensive review.
-\\
 %GEFCom 2014
@@
 Ziel et al.\ \cite{ziel2016lasso} estimates an AR model through the LASSO \cite{tibshirani1996regression} instead of the standard OLS.
-\\
 %GEFCom 2017
@@
 The GEFCom competition has also inspired the organisation of other competitions such as the RWEnpower competition in the UK, the RTE competition in France, the Tokyo electric power company competition in Japan and the BigDEAL forecasting competitions.
-\\
-\\
+
 A couple of considerations can be drawn from the above literature review.
 The EF field is characterised by heterogeneity in its forecasting techniques; methods come from statisitics, mathematics, econometrics, electrical engineering and the artificial intelligence communities.
-\\
 Every paper uses different datasets.
@@
 An additional hurdle is that some datasets are not freely accessible.
-\\
 Thus, a good understanding of the state of the art methods in EF is required to carry out a rigorous comparison between methods.
@@
 Throughout this thesis, we will address the performance of kernel methods in the context of EF.
-\\
+
 %- Kernel history evolution.
@@
 Nevertheless kernel methods received very little attention in the specific setting of EF literature.
-\\
+
 The kernel theory needed for this thesis work is covered in Section \ref{kernel_theory}. 
 For an introduction to kernel methods, we refer to \cite{learning_with_kernels, hofmann2006review, shawe2004kernel}.
-\\
+
 Kanagawa et Fukumizu introduces to the concept of kernel mean embedding \cite{pmlr}.

### 2024-06-04 thesis/input_files/metrics.tex
cleaned all problems and warnings

@@
   \end{figure}
-  \\
 It is worth noting, that the CRPS integral can be rewritten in terms of expectations.
@@
 That is because $Y_1$ and $Y_2$ both following distribution $G$ implies that $\mathbb{E}(|Y_1-Y_2|)$ corresponds to the difference of two equal constant numbers.
-\\
+
 Additionally, since $Y_1$ is just a constant, we have $Y_1=y$.
-\\
+
 Putting everything together we have obtained an alternative way of computing the CRPS score.
@@
 \end{equation}
-\\
+
 \section{Probability integral transform}
@@
   \end{figure}
-\\
+
 %RELIABILITY PLOT

### 2024-06-04 thesis/input_files/paper1.tex
cleaned all problems and warnings

@@
 We take then the inner product in the high dimensional space $\mathcal{H}$ mapped by $\phi(x)$, i.e.\
-\\
+
 \begin{align*}
@@
 \end{align*}
-\\
+
 where $\phi(x)$ is referred to as feature map while $k$ is the kernel function.
-\\
 Therefore, we can kernelize any algorithm involving a dot product by substituting $\langle x, y \rangle_{\mathcal{X}}$ with $\langle \phi(x), \phi(y) \rangle_{\mathcal{H}}$
-\\
+
 One would expect constructing the feature maps explicitly and then evaluating their inner product in $\mathcal{H}$ to be computationally expensive, and indeed it is.
@@
 This is because of the kernel trick.
-\\
 To illustrate the idea behind the kernel trick consider the following example.
@@
 For instance:
-\\
+
 $$
@@
 $$
-\\
-\\
+
 While $k([2,4], [8,9])=(16+36)^2=2704$
@@
 Let the feature space be the set of functions
-\\
+
 $$ 
@@
 \end{aligned}
-$$.
+$$
 This space is clearly closed under addition of functions and scalar multiplication. Therefore, it is a vector space.
 Letting $f(x)=\sum\limits_{i=1}^n \alpha_i k_i(x_i,x)$ and $g(x)=\sum\limits_{i=1}^m \beta_i k_i(y_i,x)$,  we can introduce an inner product on $\mathcal{F}$ as
-\\
+
 $$
@@
 $$
-\\
+
 Notice, $\langle f, f \rangle_{\mathcal{F}}= \alpha^\intercal K \alpha \geq 0 \ \forall f \in \mathcal{F}$, where $K$ is the kernel matrix.
 Moreover, $\langle f, g \rangle_{\mathcal{F}}$ is real valued, symmetric and bilinear and, hence, an inner product.
-\\
+
 Additionally, letting $g(x)=k(\cdot, x)$, we have that 
@@
 This is known as the reproducing property.
-\\
+
 Finally, we extend $\mathcal{F}$ to a complete vector space.
 Consider a fixed input $x$ and a Cauchy sequence $\{f_n\}_{n=1}^{\infty}$, we then have by means of the Cauchy-Schwarz inequality that
-\\
+
 \begin{equation}
@@
 \end{equation}
-\\
+
 We have then, that $f_n(x)$ is a Cauchy sequence of real numbers.
@@
 Having constructed the feature space $\mathcal{H}$, the mapping $\phi(x)$ is defined as
-\\
+
 \begin{equation}
@@
 \end{equation}
-\\
+
 Notice, that by the reproducing property, $f$ can be represented as the inner product with itself in the feature space, that is 
@@
     Consider a finite set of points $x_1, \cdots, x_n$, a vector $z \in \mathbb{R}^n$ and let $K_1$ and $K_2$ be the kernel matrices associated to $k_1$ and $k_2$ evaluated on these points.
-    \\
+
     We have:
-    \\
+
     $$  
@@
        Polynomial &  $k(x,y)=(x^\intercal y+c)^d$ &   c, d\\
-       Gaussian RBF &  $k(x,y)=\exp\left(-\frac{\|x-y\|_{2}^2}{2\l^2}\right)$ &   $l$\\
+       Gaussian RBF &  $k(x,y)=\exp\left(-\frac{\|x-y\|_{2}^2}{2 l^2}\right)$ &   $l$\\
        Exponential RBF/Laplacian& $k(x,y)=\exp\left(-\frac{\|x-y\|_{2}}{l}\right)$ &   $l$\\
@@
 Notice, in the Matern family, $\Gamma(\cdot)$ is the gamma function while $Y(\cdot)$ is the Bessel function of the second kind.
-\\
 When $\nu=p+\frac{1}{2}$, $p \in \mathbb{N}$, the Matern kernel can be rewritten as a product of an exponential and a polynomial of degree $p$.
-\\
-$$k(x,y)=\exp \left(-{\frac {{\sqrt {2p+1}}\|x-y|\|_{2}}{l }}\right){\frac {p!}{(2p)!}}\sum\limits_{i=0}^{p}{\frac {(p+i)!}{i!(p-i)!}}\left({\frac {2{\sqrt {2p+1}}\|x-y|\|_{2}}{l }}\right)^{p-i}$$
-\\
+
+\begin{equation*}
+    \begin{split}
+    k(x,y)&=\exp \left(-{\frac {{\sqrt {2p+1}}\|x-y|\|_{2}}{l }}\right){\frac {p!}{(2p)!}}\\
+    & \ \cdot \sum\limits_{i=0}^{p}{\frac {(p+i)!}{i!(p-i)!}}\left({\frac {2{\sqrt {2p+1}}\|x-y|\|_{2}}{l }}\right)^{p-i}
+\end{split}
+\end{equation*}
+
 For example:

### 2024-06-04 thesis/input_files/probabilistic.tex
cleaned all problems and warnings

@@
 Quantile forest allows us to estimate conditional quantiles in a non parametric fashion.
-\\
 In order to understand this algorithm, it is first necessary to cover the theory of decision trees and random forests.
@@
   \end{figure}
-\\
+
 \begin{figure}
@@
 \end{figure}
-\\
+
 Suppose now to partition the feature space into M regions $R_1,\dots,R_M$, then the model reads as follows.
@@
 \end{equation}
-\\
+
 It follows that the function minimising the sum of squares is the one with $\hat{c}_m=mean(y_i|x_i \in R_m)$.
@@
 Random forest consists in reducing the correlation between the trees by randomly selecting $m$ of the $p$ features as candidates for splitting; $m$ tipically takes a value in the order of $\sqrt{p}$ or even 1 with default value $m=\left\lfloor \frac{p}{3} \right\rfloor$, while a good minimum node size is around five.
-\\
+
 Letting $T(x;\Theta_b)$ be the $b^{th}$ bagged tree where $\Theta_b$ denotes the randomness characterising its splits, cutpoints and terminal node values, we have that the random forest regressor is given by
@@
 Therefore, the idea is that by maximising the dual function subject to the constraint $\lambda \succeq 0$, we can obtain an approximate or perfect solution to the primal problem.
-\\
 To explain why we may or may not be able to attain the best solution to the primal problem by maximising its dual, we have to introduce the concept of duality.
-\\
 We use $d^*$ to denote the optimal value of the Lagrange dual problem; we can think of it as the best lower bound on $p^*$. 
-\\
 The Inequality \ref{weak_duality} is called weak duality.
@@
 We say that strong duality holds, when the optimal duality gap is zero. In other words, the Lagrange dual bound is tight.
-\\
+
 Constraint qualifications are conditions under which strong duality holds. One of the most popular is Slater's condition.
@@
 We can now check that our optimisation problem possesses strong duality by checking Slater's condition.
-\\
+
 In our case we do not have any equality constraint, so we do not have to worry about the $h_i(x)=0$ term in Equation \ref{slater_condition}.
@@
 Thus, let us suppose that $0 \leq \xi \leq \mu$ and $0 \leq \xi^* \leq \mu$.
-\\
+
 Next, let us consider the other two inequalities and make the following ansatz.
-\\
+
 \begin{equation}
@@
 Hence, we conclude that our problem satisfies Slater's condition. Therefore, the solution of the dual and primal problems are equivalent.
-\\
+
 We end this section with the derivation of the dual problem, that is, the convex problem, which we will solve in order to get the quantiles prediction of our quantile kernel algorithm.
-\\
+
 First, take the dual function from Equation \ref{eq:kqr_min5}.
@@
 Looking at the last two equations of the system \ref{eq:lagrange_derivatives}, this implies the following two constraints $\lambda \in [0, Cq\mathbb{1}]$ and $\lambda^*\in [0, C(1-q)\mathbb{1}]$.
-\\
+
 Substitute the conditions for an optimum into \ref{eq:kqr_min5}, we obtain the dual formulation.
@@
 In order to get acquainted with the inner workings of the presented methods, this section covers an application explaining practical details and comparing results.
-\\
+
 The dataset used is the Melbourne daily maximum temperatures \citeW{hdrcde}.
@@
 The observed bimodality is that a hot day is likely to be followed by either an equally hot day or one much cooler.
-\begin{figure}[!h]
+\begin{figure}[!htp]
     \includegraphics[width=\textwidth]{images/melbourne_temperature.png}
@@
 Hyperparameters have been tuned through cross validation, see Appendix \ref{appendix:cross_validation}.
-\begin{figure}[!h]
+\begin{figure}[!htp]
     \begin{subfigure}[b]{0.5\linewidth}
       \centering
-      \includegraphics[width=1.1\textwidth]{images/melbourne_linear_quantile_regression.png} 
+      \includegraphics[width=1.0\textwidth]{images/melbourne_linear_quantile_regression.png} 
       \caption{Linear quantile regression} 
@@
       \centering
-      \includegraphics[width=1.1\textwidth]{images/melbourne_gradient_boosting_quantile_regression.png} 
+      \includegraphics[width=1.0\textwidth]{images/melbourne_gradient_boosting_quantile_regression.png} 
       \caption{Gradient boosting quantile regression} 
@@
       \centering
-      \includegraphics[width=1.1\textwidth]{images/melbourne_quantile_forest.png} 
+      \includegraphics[width=1.0\textwidth]{images/melbourne_quantile_forest.png} 
       \caption{Quantile forest} 
@@
       \centering
-      \includegraphics[width=1.1\textwidth]{images/melbourne_gaussian_rbf_kernel_quantile_regression.png}
+      \includegraphics[width=1.0\textwidth]{images/melbourne_gaussian_rbf_kernel_quantile_regression.png}
       \caption{Kernel quantile regression, Gaussian RBF} 
@@
 % tables
-\begin{table}[!h]
+\begin{table}[!htp]
 \caption{Pinball loss for Melbourne data}
@@
 
-\begin{table}[!h]
+\begin{table}[!htp]
     \caption{Pinball loss quantile-wise for Melbourne data}
@@
 As already pointed out, the quantile regression with $q=0.5$ corresponds to the standard regression problem, hence we can compare the proposed methods also in terms of the mean absolute error.
-\begin{table}[!h]
+\begin{table}[!htp]
 \caption{Mean absolute error for Melbourne data}
@@
 
-We conclude this chapter with reporting the same kind of Tables \ref{tab:kernel pinball comparison}, \ref{tab:kernel pinball comparison quantile-wise}, \ref{tab:kernel mae comparison} and Figure \ref{fig:kernel quantile regressors comparison} comparing various kernel functions, see Section \ref{kernel families}.
-
+We conclude this chapter with reporting the same kind of Tables \ref{tab:kernel pinball comparison}, \ref{tab:kernel pinball comparison quantile-wise}, \ref{tab:kernel mae comparison} and Figures \ref{fig:kernel quantile regressors comparison1},\ref{fig:kernel quantile regressors comparison2} comparing various kernel functions, see Section \ref{kernel families}.
 From these tables, we can conclude that the Matern and the Chi squared kernels are the top performers on the Melbourne weather data.
@@
 % tables
-\begin{table}[!h]
+\begin{table}[!htp]
     \caption{Kernels comparison pinball loss for Melbourne data}
@@
 
-\begin{table}[!h]
+\begin{table}[!htp]
     \caption{Kernels comparison pinball loss quantile-wise for Melbourne data}
@@
 
-\begin{table}[!h]
+\begin{table}[!htp]
     \caption{Kernels comparison mean absolute error for Melbourne data}
@@
 
-\begin{figure}[!h]
-    \begin{subfigure}[b]{0.5\linewidth}
+\begin{figure}[!htp] 
+    \begin{minipage}[b]{0.5\linewidth}
       \centering
-        \includegraphics[width=1.1\textwidth]{images/melbourne_a_laplacian_kernel_quantile_regression.png} 
+      \includegraphics[width=1.0\textwidth]{images/melbourne_a_laplacian_kernel_quantile_regression.png} 
         \caption{Absolute Laplacian} 
-        \label{} 
       \vspace{4ex}
-    \end{subfigure}%% 
-    \begin{subfigure}[b]{0.5\linewidth}
+    \end{minipage}%%
+    \begin{minipage}[b]{0.5\linewidth}
       \centering
-        \includegraphics[width=1.1\textwidth]{images/melbourne_matern_0.5_kernel_quantile_regression.png} 
+      \includegraphics[width=1.0\textwidth]{images/melbourne_matern_0.5_kernel_quantile_regression.png} 
         \caption{Matern 0.5/Laplacian}  
-        \label{} 
       \vspace{4ex}
-    \end{subfigure} 
-    \begin{subfigure}[b]{0.5\linewidth}
+    \end{minipage} 
+    \begin{minipage}[b]{0.5\linewidth}
       \centering
-        \includegraphics[width=1.1\textwidth]{images/melbourne_matern_1.5_kernel_quantile_regression.png} 
+      \includegraphics[width=1.0\textwidth]{images/melbourne_matern_1.5_kernel_quantile_regression.png} 
         \caption{Matern 1.5} 
-        \label{} 
       \vspace{4ex}
-    \end{subfigure}%%
-    \begin{subfigure}[b]{0.5\linewidth}
+    \end{minipage}%% 
+    \begin{minipage}[b]{0.5\linewidth}
       \centering
-        \includegraphics[width=1.1\textwidth]{images/melbourne_matern_2.5_kernel_quantile_regression.png}
+      \includegraphics[width=1.0\textwidth]{images/melbourne_matern_2.5_kernel_quantile_regression.png}
       \caption{Matern 2.5} 
-        \label{} 
       \vspace{4ex}
-    \end{subfigure} 
-    \begin{subfigure}[b]{0.5\linewidth}
+    \end{minipage} 
+    \begin{minipage}[b]{0.5\linewidth}
         \centering
-        \includegraphics[width=1.1\textwidth]{images/melbourne_gaussian_rbf_kernel_quantile_regression.png}
+        \includegraphics[width=1.0\textwidth]{images/melbourne_gaussian_rbf_kernel_quantile_regression.png}
         \caption{Matern $\infty$/Gaussian RBF} 
-        \label{} 
         \vspace{4ex}
-    \end{subfigure} 
-    \begin{subfigure}[b]{0.5\linewidth}
+      \end{minipage}%% 
+      \begin{minipage}[b]{0.5\linewidth}
         \centering
-        \includegraphics[width=1.1\textwidth]{images/melbourne_chi_squared_kernel_quantile_regression.png}
+        \includegraphics[width=1.0\textwidth]{images/melbourne_chi_squared_kernel_quantile_regression.png}
         \caption{Chi squared} 
-        \label{} 
         \vspace{4ex}
-    \end{subfigure} 
-    \end{figure}
-\begin{figure}[!h]\ContinuedFloat
-    \begin{subfigure}[b]{0.5\linewidth}
-        \centering
-        \includegraphics[width=1.1\textwidth]{images/melbourne_linear_kernel_quantile_regression.png}
-        \caption{Linear} 
-        \label{} 
-        \vspace{4ex}
-    \end{subfigure} 
-    \begin{subfigure}[b]{0.5\linewidth}
-        \centering
-        \includegraphics[width=1.1\textwidth]{images/melbourne_periodic_kernel_quantile_regression.png}
-        \caption{Periodic} 
-        \label{} 
-        \vspace{4ex}
-    \end{subfigure}
-    \begin{subfigure}[b]{0.5\linewidth}
-        \centering
-        \includegraphics[width=1.1\textwidth]{images/melbourne_polynomial_kernel_quantile_regression.png}
-        \caption{Polynomial} 
-        \label{} 
-        \vspace{4ex}
-    \end{subfigure} 
-    \begin{subfigure}[b]{0.5\linewidth}
-        \centering
-        \includegraphics[width=1.1\textwidth]{images/melbourne_sigmoid_kernel_quantile_regression.png}
-        \caption{Sigmoid} 
-        \label{} 
-        \vspace{4ex}
-    \end{subfigure}
-    \begin{subfigure}[b]{0.5\linewidth}
-        \centering
-        \includegraphics[width=1.1\textwidth]{images/melbourne_cosine_kernel_quantile_regression.png}
-        \caption{Cosine} 
-        \label{} 
-    \end{subfigure} 
+      \end{minipage} 
       \caption{Kernel quantile regressors comparison for Melbourne data}
-    \label{fig:kernel quantile regressors comparison} 
+      \label{fig:kernel quantile regressors comparison1} 
   \end{figure}
 
+  \begin{figure}[ht] 
+    \label{ fig8} 
+    \begin{minipage}[b]{0.5\linewidth}
+      \centering
+      \includegraphics[width=1.0\textwidth]{images/melbourne_linear_kernel_quantile_regression.png}
+        \caption{Linear}
+
+      \vspace{4ex}
+    \end{minipage}%%
+    \begin{minipage}[b]{0.5\linewidth}
+      \centering
+      \includegraphics[width=1.0\textwidth]{images/melbourne_periodic_kernel_quantile_regression.png}
+      \caption{Periodic} 
+      \vspace{4ex}
+    \end{minipage} 
+    \begin{minipage}[b]{0.5\linewidth}
+      \centering
+      \includegraphics[width=1.0\textwidth]{images/melbourne_polynomial_kernel_quantile_regression.png}
+        \caption{Polynomial}
+
+      \vspace{4ex}
+    \end{minipage}%% 
+    \begin{minipage}[b]{0.5\linewidth}
+      \centering
+      \includegraphics[width=1.0\textwidth]{images/melbourne_sigmoid_kernel_quantile_regression.png}
+        \caption{Sigmoid} 
+      \vspace{4ex}
+    \end{minipage} 
+    \begin{minipage}[b]{0.5\linewidth}
+        \centering
+        \includegraphics[width=1.0\textwidth]{images/melbourne_cosine_kernel_quantile_regression.png}
+        \caption{Cosine}  
+        \vspace{4ex}
+      \end{minipage}%% 
+      \caption{Kernel quantile regressors comparison for Melbourne data}
+      \label{fig:kernel quantile regressors comparison2}
+  \end{figure}
 % \section{Kernel density estimation}

### 2024-06-04 thesis/input_files/problemdescription.tex
cleaned all problems and warnings

@@
 Such methods are crucial in the process of taking informed decisions and for strategic planning.
-\\
 The basic idea of forecasting is that we can extract knowledge from the past in order to make educated guesses about the future.
@@
 In this thesis, our focus lies on applying forecasting techniques to the energy sector. 
-\\
 Our decision to focus on the energy market is mainly motivated by the rapid changes it has undergone. %/experienced 
@@
 Finally, the electricity market has a set of features that make it unique: electricity cannot be stored in an efficient way and supply and demand have to be matched instantly.
-\\
+
 \section{Motivation}
@@
 This enables them to better model consumer preferences.
-\\
 Transmission system operators' (TSO) main goal is to match supply and demand, generally TSO do so by increasing or decreasing the generation.
@@
 Other possible applications include: control of storage, demand side response, anomaly detection, network design and planning, simulating inputs and handling missing data.
-\\
+
 \section{Point versus probabilistic forecast}

### 2024-06-04 thesis/input_files/qr.tex
cleaned all problems and warnings

@@
 Alternatively, in order  to ease the posing of the quantile regression problem, we can formulate quantiles as the solutions to the following optimisation problems.
-\\
 For any $0<q<1$ consider the pinball loss function from Section \ref{pinball}, $\rho_q(u)=u(q-\mathbb{I}_{\{u<0\}})$. 
@@
  we can estimate quantiles by minimising the expectation of $\rho_q(y-g(x,\beta))$ with respect to the parameter $\beta$.
-\\
+
 Note, that in the special case $q=\frac{1}{2}$,  quantile regression corresponds exactly to standard regression with an absolute value loss function.
-\\
+
 It follows that the conditional linear quantile function $\hat{Q}_q=x_i\beta(q)$, can be estimated by solving
@@
 Nevertheless, we can easily solve it by employing linear programming and convex optimisation \cite{boyd2004convex}.
-\\
 Furthermore, we can extend this framework to non linear quantile regression by choosing a non linear model in place of $x\beta$ in the above Equation \ref{eq:linear_quantile_regression_minimisation_problem}.

### 2024-06-04 thesis/input_files/literaturereview.tex
cleaned all problems and warnings

@@
 Thus, understanding the complexity, strengths and weaknesses of each method is crucial for fitting the right model to the right setting.
-Within this research community, the need for more homogeneity in the choice of the error valuation metrics emerged (Section \ref{metrics}), data quality and in the way of comparing model performances \cite{EPF_review}.
+Within this research community, the need for more homogeneity in the choice of the error valuation metrics emerged, see Section \ref{metrics}, along with the need of higher data quality and uniformity in the way of comparing model performances \cite{EPF_review}.
 As a solution, \cite{lago} proposes a checklist to aid evaluating the meaningfulness of new research.

### 2024-06-05 thesis/input_files/analysis.tex
checked links urls

@@
 The kernel functions considered are: Gaussian RBF, Absolute Laplacian, Matern 0.5, Matern 1.5, Matern 2.5, Linear, Periodic, Polynomial, Sigmoid, and Cosine. 
-Details concerning the hyperparameter selection for the different kernels can be found in \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.metrics.pairwise}{sklearn.metrics.pairwise} and \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.gaussian_process}{sklearn.gaussian\_process}.
+Details concerning the hyperparameter selection for the different kernels can be found in \href{https://scikit-learn.org/stable/api/sklearn.metrics.html}{sklearn.metrics.pairwise} and \href{https://scikit-learn.org/stable/api/sklearn.gaussian_process.html}{sklearn.gaussian\_process.kernels}.
 The pool of data for this study has been created by combining \href{https://zenodo.org/records/7907883}{SECURES-Met} data (predictors) \cite{Formayer2023} and the load data from \href{https://www.energy-charts.info/index.html?l=en&c=DE}{energy charts}.

### 2024-06-05 thesis/input_files/implementation.tex
checked links urls

@@
 This estimator is compatible with in built scikit-learn methods like gridsearch, crossvalidation and scoring rules.
-Moreover, this code is compatible with sklearn in built kernel functions \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.metrics.pairwise}{sklearn.metrics.pairwise} and \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.gaussian_process.kernels}{sklearn.gaussian\_process.kernels}.
+Moreover, this code is compatible with sklearn in built kernel functions \href{https://scikit-learn.org/stable/api/sklearn.metrics.html}{sklearn.metrics.pairwise} and \href{https://scikit-learn.org/stable/api/sklearn.gaussian_process.html}{sklearn.gaussian\_process.kernels}.
 Functions supported by our kernel quantile regression include: Gaussian RBF, Absolute Laplacian, Matern 3/2, Matern 5/2, Linear, Cosine, Sigmoid, Periodic, Polynomial and custom composition of kernels.

### 2024-06-05 thesis/appendix.tex
cleaned all problems and warnings
final thesis

@@
 The rationale is that $K$-fold will randomise the order of the data, thus it might happen to use future data to predict the past; we want to avoid such behaviour in a time series setting.
-When carrying out out any kind of cross validation, we must keep consistency in the way we evaluate our predictors during model selection and in the way we perform evaluation of the test data.
+When carrying out any kind of cross validation, we must keep consistency in the way we evaluate our predictors during model selection and in the way we perform evaluation of the test data.
 Hence, for time series crossvalidation we have the timeseries split procedure, see Figure \ref{fig:crossvalidationtimeseries} for a visualisation.
@@
 \section{Source code}\label{src_code}
-The whole code for the project is hosted on
+The whole code for the project is hosted on its
 \href{https://github.com/luca-pernigo/ThesisKernelMethods}{github repo}.
@@
     \item thesis/: folder containing tex files for thesis pdf.
-    \item beamer/: folder containing tex files for beamer pptx %pptx thesis defence.
+    \item beamer/: folder containing tex files for beamer pptx. %pptx thesis defence.
     \item experiments: folder containing train, test script for each of the experiments carried out in the thesis alongside with the binary files of the pickled models.

### 2024-06-05 thesis/input_files/literaturereview.tex
final thesis

@@
 Hyndman et al.\ \cite{hyndman1996estimating} explains kernel estimator of conditional density, analyses its asymptotic behaviour and covers an application for the daily temperatures of Melbourne.
-Kaur et al.\ \cite{kaur2022energy} carries out a comparative study of techniques spanning ideas from statistic and artificial intelligence.
+Kaur et al.\ \cite{kaur2022energy} carries out a comparative study of techniques spanning ideas from statistics and artificial intelligence.
 Van der Meer et al.\ \cite{van2018review} provides another thorough analysis of the probabilistic forecasting realm by covering recent advances and identifying research gaps.
@@
 It was in the 1990s that the theory of kernel methods got popular, particularly in the field of machine learning.
-Kernels have been used in various different tasks such as support vector machines (SVM) \cite{vapnik1} \cite{vapnik2}, Gaussian process classifiers \cite{williams}, spline methods \cite{wahba}, neural networks \cite{poggio} and principal component analysis \cite{pca_scholkopf}.
+Kernels have been used in various different tasks such as support vector machines (SVM) \cite{vapnik1, vapnik2}, Gaussian process classifiers \cite{williams}, spline methods \cite{wahba}, neural networks \cite{poggio} and principal component analysis \cite{pca_scholkopf}.
 Nevertheless kernel methods received very little attention in the specific setting of EF literature.
@@
 Its weakness is that it can only take a finite number of features and this limits its expressive power.
-Smola et al.\ overcome this by translating the algorithm into the framework of kernels.
+Smola et al.\ overcomes this by translating the algorithm into the framework of kernels.
 

### 2024-06-05 thesis/input_files/metrics.tex
final thesis

@@
 Consider the time series with actual values given by $y=(y_{n+1}, y_{n+2},\dots, y_{n+h})$
-and its $h$ step ahead point forecast $\hat{y}=(\hat{y}_{n+1}, \hat{y}_{n+2},\dots, \hat{y}_{n+h})$ the mean absolute error (MAE) is defined as
+and its $h$ step ahead point forecast $\hat{y}=(\hat{y}_{n+1}, \hat{y}_{n+2},\dots, \hat{y}_{n+h})$
 \begin{definition}
@@
 \begin{cases}
-(q-1)(\hat{y}_{t,q}-y_{t}) & y_t > \hat{y}_{t,q} \\
-q(\hat{y}_{t,q}-y_t) & y_t \leq \hat{y}_{t,q}
+q(y_{t}-\hat{y}_{t,q}) & y_t > \hat{y}_{t,q} \\
+-(q-1)(y_{t}-\hat{y}_{t,q}) & y_t \leq \hat{y}_{t,q}
 \end{cases}
@@
 This makes its evaluation easier, since we know that the sample mean converges to the expectation for sufficiently big sample sizes by the law of large numbers.
-This was first pointed out by \cite{proper_scores}, where the authors take advantage of Lemma 2.2 of \cite{new_multi_test2} or equivalently identity (17) of \cite{new_multi_tes1}.
+This was first pointed out by \cite{proper_scores}, where the authors take advantage of \cite[Lemma 2.2]{new_multi_test2} or equivalently of \cite[Identity (17)]{new_multi_tes1}.
 \\

### 2024-06-05 thesis/input_files/paper1.tex
cleaned all problems and warnings
final thesis

@@
 
-While $k([2,4], [8,9])=(16+36)^2=2704$
+while $$k([2,4], [8,9])=(16+36)^2=2704$$
+
 We can see that the latter calculation is much quicker and compact.
@@
 \begin{definition}
-    A sequence $\{v_n\}_{n=1}^{\infty}$ of elements of a normed vector space $\mathcal{V}$ is a Cauchy sequence if for every $\epsilon>0$, there exist $N=N(\epsilon) \in \mathbb{N}$ such that $\|v_n-v_m\|_{\mathcal{V}}<\epsilon \ \ \forall m,n\geq N$
+    A sequence $\{v_n\}_{n=1}^{\infty}$ of elements of a normed vector space $\mathcal{V}$ is a Cauchy sequence if for every $\epsilon>0$, there exists $N=N(\epsilon) \in \mathbb{N}$ such that $\|v_n-v_m\|_{\mathcal{V}}<\epsilon \ \ \forall m,n\geq N$
 \end{definition}
@@
 Next, the matrix given by $k_1 \cdot k_2$ corresponds to the Schur product of $K_1$ and $K_2$, denote it $H$.
-Notice, $H$ is a principal submatrix of $K$, hence for any $z \in \mathbb{R}^n$ there exist a $t \in \mathbb{R}^{n^2}$ such that $z^\intercal H z= t^\intercal K t$.
+Notice, $H$ is a principal submatrix of $K$, hence for any $z \in \mathbb{R}^n$ there exists a $t \in \mathbb{R}^{n^2}$ such that $z^\intercal H z= t^\intercal K t$.
 Using the fact that K is positive semidefinite we have that for any $z$, it holds
@@
        Absolute Laplacian& $k(x,y)=\exp\left(-\frac{\|x-y\|_{1}}{l}\right)$ &   $l$\\
-       Hyperbolic/Sigmoid Kernel &  $k(x,y2)=\tanh(\gamma x^\intercal y+r)$ &  $\gamma$, r \\
+       Hyperbolic/Sigmoid &  $k(x,y)=\tanh(\gamma x^\intercal y+r)$ &  $\gamma$, r \\
        Periodic &  $k(x,y)=\exp\left(\frac{-2 sin^2\left(\frac{\pi}{p}\|x-y\|_{2} \right)}{l^2}\right) $ &   $p, l$\\
-       Chi-squared kernel &  $k(x,y)=\exp\left(-\gamma \sum\limits_i \frac{(x_i - y_i)^2}{x_i+y_i}\right) $ &   $\gamma$\\
+       Chi-squared &  $k(x,y)=\exp\left(-\gamma \sum\limits_i \frac{(x_i - y_i)^2}{x_i+y_i}\right) $ &   $\gamma$\\
         Cosine  &  $k(x,y)=\frac{x^\intercal y}{\|x\|_{2}\|y\|_{2}}$ &   \\
@@
 Kernel methods show good modularity because they work the same for any kernel and any data type.
-This concept is illustrated in Figure \ref{fig:workflow_kernels}, where an abstract kernel methods workflow is visualized.
+This concept is illustrated in Figure \ref{fig:workflow_kernels}, where an abstract kernel methods workflow is visualised.
 First we build the kernel matrix by using a kernel function to process the input data.

### 2024-06-05 thesis/input_files/point.tex
cleaned all problems and warnings
final thesis

@@
 Writing SARIMA in compact notation we have
-\\
+
 \begin{equation}
@@
 \end{figure}
-Smola et al.\ \cite{smola2004tutorial} point out that considering the dual formulation makes our optimisation problem easier to solve. Doing so, we have the equivalent optimisation problem
+Smola et al.\ \cite{smola2004tutorial} points out that considering the dual formulation makes our optimisation problem easier to solve. Doing so, we have the equivalent optimisation problem
 \begin{equation}
@@
 In this way, the current state of a cell depends on the current value $X_t$, on the previous cell activation and on the previous cell state $C_{t-1}$, see Figure \ref{fig:lstm_cell} for a visual representation.
-\begin{figure}[!h]
+\begin{figure}[!ht]
     \includegraphics[width=\textwidth]{images/lstm_cell.png}
@@
 \subsection{Kernel ridge regression}
-In the setting of kernel ridge regression, we aim at estimating $f$ such that $y=f(x)+\epsilon$. Given an RKHS $\mathcal{H}$, we can estimate $f$ by solving the optmisation problem
+In the setting of kernel ridge regression, we aim at estimating $f$ such that $y=f(x)+\epsilon$. Given a RKHS $\mathcal{H}$, we can estimate $f$ by solving the optmisation problem
 \begin{equation}\label{eq:kernel ridge}
@@
 \end{figure}
-
 \begin{figure}

### 2024-06-05 thesis/input_files/probabilistic.tex
cleaned all problems and warnings
final thesis

@@
 
-Suppose now to partition the feature space into M regions $R_1,\dots,R_M$, then the model reads as follows.
+Suppose now to partition the feature space into $M$ regions $R_1,\dots,R_M$, then the model reads as follows.
 \begin{equation}
@@
 \end{equation}
-The next step is deriving its dual formulation, since it is easier and more efficient to solve. This because the dual problem has the useful property of being always convex.
+The next step is deriving its dual formulation, since it is easier and more efficient to solve. This because the dual problem is always convex.
 
@@
 As pointed out previously, the lower bound resulting from the dual formulation is non trivial only when the Lagrange multipliers $\lambda$ are $\succeq 0$.
-Looking at the last two equations of the system \ref{eq:lagrange_derivatives}, this implies the following two constraints $\lambda \in [0, Cq\mathbb{1}]$ and $\lambda^*\in [0, C(1-q)\mathbb{1}]$.
+Looking at the last two equations of the System \ref{eq:lagrange_derivatives}, this implies the following two constraints $\lambda \in [0, Cq\mathbb{1}]$ and $\lambda^*\in [0, C(1-q)\mathbb{1}]$.
 
@@
 
-We conclude this chapter with reporting the same kind of Tables \ref{tab:kernel pinball comparison}, \ref{tab:kernel pinball comparison quantile-wise}, \ref{tab:kernel mae comparison} and Figures \ref{fig:kernel quantile regressors comparison1},\ref{fig:kernel quantile regressors comparison2} comparing various kernel functions, see Section \ref{kernel families}.
+We conclude this chapter with reporting the same kind of Tables \ref{tab:kernel pinball comparison}, \ref{tab:kernel pinball comparison quantile-wise}, \ref{tab:kernel mae comparison} and Figures \ref{fig:kernel quantile regressors comparison1}, \ref{fig:kernel quantile regressors comparison2} comparing various kernel functions, see Section \ref{kernel families}.
 From these tables, we can conclude that the Matern and the Chi squared kernels are the top performers on the Melbourne weather data.

### 2024-06-07 thesis/input_files/paper1.tex
fixed typo

@@
 This space is clearly closed under addition of functions and scalar multiplication. Therefore, it is a vector space.
-Letting $f(x)=\sum\limits_{i=1}^n \alpha_i k_i(x_i,x)$ and $g(x)=\sum\limits_{i=1}^m \beta_i k_i(y_i,x)$,  we can introduce an inner product on $\mathcal{F}$ as
+Letting $f(x)=\sum\limits_{i=1}^n \alpha_i k(x_i,x)$ and $g(x)=\sum\limits_{i=1}^m \beta_i k(y_i,x)$,  we can introduce an inner product on $\mathcal{F}$ as
 
@@
 \begin{aligned}
-    \langle f,g \rangle_{\mathcal{F}}:=\sum\limits_{i=1}^n\sum\limits_{j=1}^m a_i \beta_j k_i(x_i,y_j)=\sum\limits_{i=1}^n a_i g(x_i)= \sum\limits_{j=1}^m \beta_j f(y_j)
+    \langle f,g \rangle_{\mathcal{F}}:=\sum\limits_{i=1}^n\sum\limits_{j=1}^m a_i \beta_j k(x_i,y_j)=\sum\limits_{i=1}^n a_i g(x_i)= \sum\limits_{j=1}^m \beta_j f(y_j)
 \end{aligned}
